### 2024-10-04 study2.qmd
initial commit

---
title: "Study 2 Analysis"
author:
- name: Thomas E. Gorman
  url: https://tegorman13.github.io/
  affiliations: 
  - name:  Communication and Cognition Lab, Purdue University, USA
    affiliation-url: https://web.ics.purdue.edu/~treimer/
lightbox: true
---

```{r}
library(pacman)
pacman::p_load(dplyr,purrr,tidyr,haven,tibble,ggplot2,gt,ggh4x,lme4)

s2 <- haven::read_sav("data/Frequency & Probability Study 2 - 3-24-19.sav")
s2 <- s2 |> mutate(id = as.numeric(format(EndDate, "%H%M%S"))) |> relocate(id)

s2 <- s2 |> 
  mutate(
    state1 = case_when(
      COLORADO == 1 ~ "Colorado",
      MASSACHUSETTS == 1 ~ "Massachusetts",
      CALIFORNIA == 1 ~ "California",
      TEXAS == 1 ~ "Texas",
      TRUE ~ NA_character_
    ),
    state2 = case_when(
      COLORADO == 2 ~ "Colorado",
      MASSACHUSETTS == 2 ~ "Massachusetts",
      CALIFORNIA == 2 ~ "California",
      TEXAS == 2 ~ "Texas",
      TRUE ~ NA_character_
    )
  ) |> 
  mutate(refClass = case_when(
  REFERENCECLASS == 1 ~ "USD",
  REFERENCECLASS == 2 ~ "kWh",
  REFERENCECLASS == 3 ~ "Percentage"
)) |> 
  relocate(refClass,state1,state2, .after = id)

s2_b1 <- s2 |> select(id,state1,refClass,TENPERCENT,ROUNDED, ACFAMILY1:TOTALAPTWO1) |> 
  mutate( 
    block = 1,
    state=state1,
  pct_goal=case_when (TENPERCENT == 1 ~ "10%", TENPERCENT == 2 ~ "15%"), 
  rounded = case_when(ROUNDED == 1 ~ "Rounded", ROUNDED == 2 ~ "Not Rounded")
  ) |> relocate(id,block,refClass,pct_goal,rounded,state)  |> 
  select(-TENPERCENT,-ROUNDED, -state1)

s2_b2 <- s2 |> select(id,state2,refClass,TENPERCENT,ROUNDED, ACFAMILY2:TOTALAPTWO2) |> 
  mutate( 
    block = 2,
    state=state2,
  pct_goal=case_when (TENPERCENT == 1 ~ "15%", TENPERCENT == 2 ~ "10%"), 
  rounded = case_when(ROUNDED == 1 ~ "Not Rounded", ROUNDED == 2 ~ "Rounded")
  ) |> relocate(id,block,refClass,pct_goal,rounded,state) |> 
  select(-TENPERCENT,-ROUNDED, -state2)

# pivot data into longer format, so that each appliance type is on its own row, so for each appliance we have the family value, state, value, plan 1 value, and plan 2 value
# so for the first row for id 113029, we would have ACFAMILY1 ACSTATE1 ACAPONE1 ACAPTWO1, then the next row would be HEATFAMILY1 HEATSTATE1 HEATAPONE1 HEATAPTWO1, etc.
# so each of the bill sources, AC, HEAT, WATER, REFRIG, OTHER, will now be on its own row, with the family, state, value, plan 1 value, and plan 2 value

s2_b1_long <- s2_b1 |>
  pivot_longer(
    cols = c(ACFAMILY1:TOTALAPTWO1),
    names_to = c("appliance", "measure"),
    names_pattern = "(AC|HEAT|WATER|REFRIG|OTHER|TOTAL)(FAMILY1|STATE1|APONE1|APTWO1)",
    values_to = "value"
  ) |>
  pivot_wider(
    names_from = measure,
    values_from = value
  ) |>
  rename(
    family = FAMILY1,
    state_avg = STATE1,
    plan1 = APONE1,
    plan2 = APTWO1
  ) |>
  relocate(appliance, .after = refClass)

s2_b2_long <- s2_b2 |>
  pivot_longer(
    cols = c(ACFAMILY2:TOTALAPTWO2),
    names_to = c("appliance", "measure"),
    names_pattern = "(AC|HEAT|WATER|REFRIG|OTHER|TOTAL)(FAMILY2|STATE2|APONE2|APTWO2)",
    values_to = "value"
  ) |>
  pivot_wider(
    names_from = measure,
    values_from = value
  ) |>
  rename(
    family = FAMILY2,
    state_avg = STATE2,
    plan1 = APONE2,
    plan2 = APTWO2
  ) |>
  relocate(appliance, .after = refClass)

s2_long <- bind_rows(s2_b1_long, s2_b2_long) |> 
  arrange(id, block)

# pivot so that plan1 and plan2 are on their own rows
s2_long <- s2_long |>
  group_by(id,block) |> 
  pivot_longer(
    cols = c(plan1, plan2),
    names_to = "plan",
    values_to = "value"
  ) |> 
  mutate(change = value-family, state_dif = value-state_avg,plan=as_factor(plan))

#write.csv(s2_long,"data/study_2_tg.csv")

# compute total kWh reduction, and total % reduction 
sum1 <- s2_long |> 
  filter(appliance !="TOTAL") |> 
  group_by(id,refClass,state,pct_goal,rounded,block,plan) |> 
  summarise(total_kWh = sum(value),orig_kWh=sum(family), pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), state_dif=mean(state_dif))

```

```{r}
#| fig.width: 11
#| fig.height: 9

outliers <- sum1 |>  ungroup() |> group_by(state) |> mutate(change_mean=mean(pct_change,na.rm=TRUE),change_sd=sd(pct_change,na.rm=TRUE),
                                     z_score=(pct_change-change_mean)/change_sd,
                                     is_outlier=abs(z_score)>2.5)

outlier_id <- outliers |> filter(is_outlier) |> pull(id) |> unique()

outlier_summary <- outliers %>%
  group_by(pct_goal, rounded, refClass,state) %>%
  summarise(
    num_outliers = sum(is_outlier, na.rm = TRUE),
    total = n(),
    outlier_percentage = (num_outliers / total) * 100
  )

outlier_plot <- ggplot(outliers, aes(x = state, y = z_score, color = is_outlier)) +
  geom_jitter(alpha = 0.5) +
  facet_wrap(~ pct_goal) +
  labs(
    title = "Z-Scores of Change in Energy Usage by Appliance",
    x = "Appliance",
    y = "Z-Score",
    color = "Outlier"
  ) +
  theme_minimal()

outlier_plot

sum1 |> ggplot(aes(x=pct_change)) +
  geom_density(aes(fill=refClass), alpha=0.5) +
  ggdist::stat_halfeye() +
  geom_rug(alpha=0.1) +
  stat_boxplot(width=0.1, alpha=0.3) +
  facet_wrap(~state, scales="free_y") +
  theme_minimal() +
  labs(title="Distribution of Percent Change by State and Reference Class",
       x="Percent Change",
       y="Density") +
  scale_fill_brewer(palette="Set2")

# plot densities in a better way to account for the positive skew

# Assess skewness and normality of the data
library(e1071)
# Calculate skewness and kurtosis for the `pct_change` variable
skewness_value <- skewness(sum1$pct_change, na.rm = TRUE)
kurtosis_value <- kurtosis(sum1$pct_change, na.rm = TRUE)

cat("Skewness:", skewness_value, "\n")
cat("Kurtosis:", kurtosis_value, "\n")

# Shapiro-Wilk test for normality (sample size limitation)
shapiro_test <- shapiro.test(sum1$pct_change[1:5000])  # Using first 5000 observations if large dataset
print(shapiro_test)

sum1 |> 
  filter(!id %in% outlier_id) |>
  ggplot(aes(x=pct_change)) +
  geom_density(aes(fill=refClass), alpha=0.5) +
  ggdist::stat_halfeye() +
  geom_rug(alpha=0.1) +
  stat_boxplot(width=0.1, alpha=0.3) +
  facet_wrap(~state, scales="free_y") +
  theme_minimal() +
  labs(title="Distribution of Percent Change by State and Reference Class",
       x="Percent Change",
       y="Density") +
  scale_fill_brewer(palette="Set2")

sum1 <- sum1 |> 
  filter(!id %in% outlier_id)

```

```{r}
# plot distribution of reductions by applicance
s2_long |> ggplot(aes(x = change)) + geom_histogram(bins = 30) + facet_wrap(~appliance, scales = "free")

DT::datatable(sum1)

DT::datatable(sum1 |> group_by(id,block) |> summarise(diff(total_kWh),diff(orig_kWh)))

```

```{r}

sum1 |> ggplot(aes(x = pct_goal, y = pct_change, fill = refClass)) + 
  geom_boxplot(position = "dodge") +
  facet_wrap(~state)

 sum1 |> ggplot(aes(x = pct_goal, y = pct_change, fill = refClass)) + 
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") 
```

```{r}
#| fig.width: 12
#| fig.height: 10

  # add geom_vline with x intercept based on pct_goal of the data (which is currently a string of form 10%)
sum1 |> 
  filter(block==1) |>
  ungroup() |> 
  mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100) |>
  mutate(id=reorder(id,pct_change)) |> 
  ggplot(aes(y=id,x=pct_change)) + 
  geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
  geom_vline(aes(xintercept=goal_pct),linetype="dashed",alpha=.5) +
  ggh4x::facet_nested_wrap(refClass~pct_goal,axes="all",scales="free",ncol=2)

sum1 |> 
  ungroup() |> 
  mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100) |>
  mutate(id=reorder(id,pct_change)) |> 
  ggplot(aes(y=id,x=pct_change)) + 
  geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
  geom_vline(aes(xintercept=goal_pct),linetype="dashed",alpha=.5) +
  ggh4x::facet_nested_wrap(refClass~pct_goal,axes="all",scales="free",ncol=2)

# sum1 |> 
#   ungroup() |> 
#   mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100) |>
#   mutate(id=reorder(id,pct_change)) |> 
#   ggplot(aes(y=id,x=pct_change,col=refClass)) + 
#   geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
#   geom_vline(aes(xintercept=goal_pct),linetype="dashed",alpha=.5) +
#   ggh4x::facet_nested_wrap(block~pct_goal,axes="all",scales="free",ncol=2) 
# 
# 
# sum1 |> 
#   ungroup() |> 
#     mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100) |>
#   ggplot(aes(x=state,y=pct_change,col=state)) + 
#   geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
#     geom_hline(aes(yintercept=goal_pct),linetype="dashed",alpha=.5) +
#   ggdist::stat_halfeye()+
#   ggh4x::facet_nested_wrap(block~pct_goal,axes="all",ncol=2)

```

```{r}
#| fig.height: 11
#| 
sum1 |> 
  ungroup() |> 
  mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100,
         block=factor(block)) |>
  group_by(refClass,block,pct_goal) |>
  summarise(mean_pct_change = mean(pct_change),se_pct_change = sd(pct_change)/sqrt(n())) |>
  ggplot(aes(x=mean_pct_change,y=refClass,col=block)) + 
  geom_pointrange(aes(xmin=mean_pct_change-se_pct_change,xmax=mean_pct_change+se_pct_change),position = position_dodge(width=0.5)) +
  ggforce::facet_col(~pct_goal)

sum1 |> 
  ungroup() |> 
  mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100,
         block=factor(block)) |>
  group_by(refClass,,pct_goal,state) |>
  summarise(mean_pct_change = mean(pct_change),se_pct_change = sd(pct_change)/sqrt(n())) |>
  ggplot(aes(x=mean_pct_change,y=refClass,col=pct_goal)) + 
  geom_pointrange(aes(xmin=mean_pct_change-se_pct_change,xmax=mean_pct_change+se_pct_change),position = position_dodge(width=0.5)) +
  ggforce::facet_col(~state)

sum1 |> 
  ungroup() |> 
  mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100,
         block=factor(block)) |>
  group_by(refClass,,pct_goal,state,goal_pct) |>
  summarise(mean_pct_change = mean(pct_change),se_pct_change = sd(pct_change)/sqrt(n())) |>
  ggplot(aes(x=mean_pct_change,y=refClass,col=state)) + 
  geom_vline(aes(xintercept=goal_pct),linetype="dashed",alpha=.5) +
  geom_pointrange(aes(xmin=mean_pct_change-se_pct_change,xmax=mean_pct_change+se_pct_change),position = position_dodge(width=0.5)) +
  ggforce::facet_col(~pct_goal)

  s2_long |> 
    filter(appliance != "TOTAL") |>
  group_by(refClass,appliance,state) |>
  summarise(mean_kWh_change = mean(change),se_pct_change = sd(change)/sqrt(n())) |>
  ggplot(aes(x=mean_kWh_change,y=refClass,col=appliance)) + 
  geom_pointrange(aes(xmin=mean_kWh_change-se_pct_change,xmax=mean_kWh_change+se_pct_change),position = position_dodge(width=0.5)) +
  ggforce::facet_col(~state)

s2_long |> 
  filter(appliance != "TOTAL") |>
  ggplot(aes(x=refClass,y=change,fill=state))+
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
  scale_y_reverse() +
  facet_wrap(~appliance)

s2_long |> 
  filter(appliance != "TOTAL") |>
  ggplot(aes(x=refClass,y=state_dif,fill=state))+
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
  scale_y_reverse() +
  facet_wrap(~appliance)

s2_long |> 
  filter(appliance != "TOTAL") |>
  ggplot(aes(x=state_dif,y=value,col=state))+
  geom_point() +
  facet_wrap(~appliance)

sum1 |> 
  ggplot(aes(x=state_dif,y=pct_change,col=state)) +
  geom_point()

```

```{r}

# mixed effects model testing effect of refClass, state, and pct_goal on pct_change

m1 <- lmer(pct_change ~ refClass  + (1+plan|id) + (1|state), data = sum1)
summary(m1)

```

```{r}
#| eval: false

# compute total kWh reduction, and total % reduction 
s2_long |> group_by(id,refClass,state,pct_goal,rounded,block,plan) |> 
  filter(appliance !="TOTAL") |> 
  summarise(total_kWh = sum(value),orig_kWh=sum(family), total_pct = sum(value)/sum(family)) |>
  ggplot(aes(x = pct_goal, y = total_pct, fill = refClass)) + 
  geom_boxplot(position = "dodge") +
  facet_wrap(~state)

s2_long |> group_by(id,refClass,state,pct_goal,rounded) |> 
  filter(appliance !="TOTAL") |> 
  summarise(total_kWh = sum(reduction), total_pct = sum(reduction)/sum(family)) |>
  ggplot(aes(x = pct_goal, y = total_pct, fill = refClass)) + 
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge")

s2_long |> group_by(id,refClass,state,pct_goal,rounded) |> 
  filter(appliance !="TOTAL") |> 
  summarise(total_kWh = sum(reduction), orig_kWh=sum(family), total_pct = round(sum(reduction)/sum(family))) |>
  ggplot(aes(x = state1, y = total_pct, fill = refClass)) + 
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
  facet_wrap(~pct_goal)

s2_long |> group_by(id,refClass,state,pct_goal,rounded,plan) |> 
  filter(appliance !="TOTAL") |> 
  summarise(total_kWh = sum(reduction), total_pct = sum(reduction)/sum(family)) |>
  ggplot(aes(x = refClass, y = total_pct, fill = pct_goal)) + 
  stat_summary(fun = mean, geom = "bar", position="dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
  facet_wrap(plan~rounded)

# s2_long <- s2 |>
#   pivot_longer(
#     cols = c(ACFAMILY1:TOTALAPTWO1),
#     names_to = c(".value", "appliance"),
#     names_pattern = "(.*)(FAMILY1|STATE1|APONE1|APTWO1)$",
#     values_to = "value"
#   )

# s2 |> pivot_longer(cols = contains("FAMILY"), names_to = "Family", values_to = "kWh") |> relocate(Family,kWh) |> 
#   pivot_longer(cols = contains("STATE1"), names_to = "State", values_to = "kWh") |> relocate(State)

# 
# s2 |> 
#   pivot_longer(cols=ACFAMILY1:TOTALAPTWO1, names_to = "appliance", values_to = "family") |> relocate(appliance,family)

```

### 2024-10-04 analysis1.R

@@ -1,0 +1,63 @@
+
+
+
+pacman::p_load(dplyr,purrr,tidyr,tibble,ggplot2,gt,patchwork,ggh4x,knitr,lme4)
+data <- read.csv("processed_energy_study_data.csv")
+
+colnames(data)
+head(data,1)
+glimpse(data)
+
+
+
+data %>%
+  summarise(
+    avg_age = mean(2021 - birth_year, na.rm = TRUE),
+    gender_distribution = table(gender)
+  )
+
+avg_age <- data %>%
+  summarise(avg_age = mean(birth_year, na.rm = TRUE))
+
+final_data %>%
+  count(gender)
+
+
+
+# Reliability analysis for 'els' scale
+library(psych)
+els_items <- final_data %>%
+  select(starts_with("els")) %>%
+  mutate_all(as.numeric)
+alpha_results <- alpha(els_items)
+print(alpha_results)
+
+
+
+
+
+# Visualization of scenario responses
+data %>%
+  ggplot(aes(x = scenario, y = `01`)) +
+  geom_boxplot() +
+  labs(title = "Distribution of Response '01' Across Scenarios")
+
+
+
+model <- lmer(X01 ~ scenario + (1 | responseid), data = data)
+
+
+
+
+scenario_means <- data %>%
+  group_by(scenario) %>%
+  summarise(mean_X01 = mean(X01, na.rm = TRUE))
+
+scenario_means %>%
+  ggplot(aes(x = scenario, y = mean_X01)) +
+  geom_bar(stat = "identity") +
+  labs(title = "Mean Response 'X01' Across Scenarios",
+       x = "Scenario",
+       y = "Mean Response") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))

### 2024-10-04 read_data.R

@@ -1,0 +1,180 @@
+
+# read sav file "CLEAN NSF Frequency and Probability - 3-9-18.sav"
+
+#library(foreign)
+
+# data <- read.spss("CLEAN NSF Frequency and Probability - 3-9-18.sav", to.data.frame = TRUE)
+# colnames(data)
+# head(data,1)
+
+# read sav file "Frequency & Probability Study 1 - 3-24-19.sav"
+#data2 <- read.spss("Frequency & Probability Study 1 - 3-24-19.sav", to.data.frame = TRUE)
+# read sav file "Frequency & Probability - 9-25-18 cleaned.sav"
+#data3 <- read.spss("Frequency & Probability - 9-25-18 cleaned.sav", to.data.frame = TRUE)
+
+
+
+#------
+
+# list.files('data')
+# [1] "aggr.sav"  - aggregates for 12 conditions (RefClass *ROUNDED * TENPERCENT)                                                                        
+# [2] "CLEAN NSF Frequency and Probability - 3-9-18.sav"                                  
+# [3] "Format_1.sav"                                                                      
+# [4] "Format_3.sav"                                                                      
+# [5] "Format_new1.sav"  - 252x2353;  accuracy - no group variables?                                                              
+# [6] "Format_revised1.sav"                                                               
+# [8] "Frequency & Probability - 9-25-18 cleaned.sav"                                     
+# [9] "Frequency & Probability - 9-25-18_2.sav"                                           
+# [10] "Frequency & Probability 2 10-25-18 140p 207count.sav"                              
+# [11] "Frequency & Probability 2 3-22-19.sav"    - 204x145;  refclass; rounding; tenpercent;                                         
+# [12] "Frequency & Probability old.sav"                                                   
+# [13] "Frequency & Probability Study 1 - 3-24-19.sav" - 1840x252; ReferenceClass and States; responseid; sept 24-25th 2018                                    
+# [14] "Frequency & Probability Study 2 - 3-24-19.sav"    - 208x146; refclass; rounding; tenpercent; states;  Oct 21-22 2018                           
[…]

### 2024-10-04 study2.qmd
update plots

@@
 lightbox: true
+code-fold: true
+code-tools: true
+execute:
+  #echo: false
+  warning: false
+format: 
+  html:
+    page-layout: full
 ---
@@
 
+## Inspect response distribution for outliers
+
 ```{r}
-#| fig.width: 11
-#| fig.height: 9
 
@@
   labs(
-    title = "Z-Scores of Change in Energy Usage by Appliance",
-    x = "Appliance",
+    title = "Z-Scores of Change in Energy Usage by State - outliers colored",
+    x = "State",
     y = "Z-Score",
@@
   theme_minimal()
-
-outlier_plot
-
-sum1 |> ggplot(aes(x=pct_change)) +
-  geom_density(aes(fill=refClass), alpha=0.5) +
-  ggdist::stat_halfeye() +
-  geom_rug(alpha=0.1) +
-  stat_boxplot(width=0.1, alpha=0.3) +
-  facet_wrap(~state, scales="free_y") +
-  theme_minimal() +
-  labs(title="Distribution of Percent Change by State and Reference Class",
-       x="Percent Change",
-       y="Density") +
-  scale_fill_brewer(palette="Set2")
-
-# plot densities in a better way to account for the positive skew
 
@@
 print(shapiro_test)
+```
+
+::: {.column-page-inset-right}
+```{r}
+#| layout-ncol: 1
+#| fig.width: 12
+#| fig.height: 11
+
+#outlier_plot
+
+sum1 |> ggplot(aes(x=pct_change)) +
+  geom_density(aes(fill=state), alpha=0.5) +
+  ggdist::stat_halfeye() +
+  geom_rug(alpha=0.1) +
+  stat_boxplot(width=0.2, alpha=0.3) +
+  facet_wrap(~refClass) +
+  theme_minimal() +
+  labs(title="Distribution of Percent Change by State and Reference Class - Before outlier exclusion",
+       x="Percent Change",
+       y="Density") +
+  scale_fill_brewer(palette="Set2")
+
+# plot densities in a better way to account for the positive skew
+
+sum1 <- sum1 |> 
+  filter(!id %in% outlier_id)
 
@@
   ggplot(aes(x=pct_change)) +
-  geom_density(aes(fill=refClass), alpha=0.5) +
+  geom_density(aes(fill=state), alpha=0.5) +
   ggdist::stat_halfeye() +
   geom_rug(alpha=0.1) +
-  stat_boxplot(width=0.1, alpha=0.3) +
-  facet_wrap(~state, scales="free_y") +
+  stat_boxplot(width=0.2, alpha=0.3) +
+  facet_wrap(~refClass) +
   theme_minimal() +
-  labs(title="Distribution of Percent Change by State and Reference Class",
+  labs(title="After outlier exclusion",
        x="Percent Change",
@@
 
-sum1 <- sum1 |> 
-  filter(!id %in% outlier_id)
-
 ```
 
-```{r}
-# plot distribution of reductions by applicance
-s2_long |> ggplot(aes(x = change)) + geom_histogram(bins = 30) + facet_wrap(~appliance, scales = "free")
-
-DT::datatable(sum1)
-
-DT::datatable(sum1 |> group_by(id,block) |> summarise(diff(total_kWh),diff(orig_kWh)))
-
-```
-
-```{r}
-
-sum1 |> ggplot(aes(x = pct_goal, y = pct_change, fill = refClass)) + 
-  geom_boxplot(position = "dodge") +
-  facet_wrap(~state)
-
- sum1 |> ggplot(aes(x = pct_goal, y = pct_change, fill = refClass)) + 
-  stat_summary(fun = mean, geom = "bar", position="dodge") +
-  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") 
-```
+## Individual Participant responses
 
@@
 
-sum1 |> 
-  ungroup() |> 
-  mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100) |>
-  mutate(id=reorder(id,pct_change)) |> 
-  ggplot(aes(y=id,x=pct_change)) + 
-  geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
-  geom_vline(aes(xintercept=goal_pct),linetype="dashed",alpha=.5) +
-  ggh4x::facet_nested_wrap(refClass~pct_goal,axes="all",scales="free",ncol=2)
+```
 
-# sum1 |> 
-#   ungroup() |> 
-#   mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100) |>
-#   mutate(id=reorder(id,pct_change)) |> 
-#   ggplot(aes(y=id,x=pct_change,col=refClass)) + 
-#   geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
-#   geom_vline(aes(xintercept=goal_pct),linetype="dashed",alpha=.5) +
-#   ggh4x::facet_nested_wrap(block~pct_goal,axes="all",scales="free",ncol=2) 
-# 
-# 
-# sum1 |> 
-#   ungroup() |> 
-#     mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100) |>
-#   ggplot(aes(x=state,y=pct_change,col=state)) + 
-#   geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
-#     geom_hline(aes(yintercept=goal_pct),linetype="dashed",alpha=.5) +
-#   ggdist::stat_halfeye()+
-#   ggh4x::facet_nested_wrap(block~pct_goal,axes="all",ncol=2)
+:::
+
+## Interactive data table
+
+```{r}
+# plot distribution of reductions by applicance
+# s2_long |> ggplot(aes(x = change)) + geom_histogram(bins = 30) + facet_wrap(~appliance, scales = "free")
+DT::datatable(sum1)
+#DT::datatable(sum1 |> group_by(id,block) |> summarise(diff(total_kWh),diff(orig_kWh)))
+```
+
+## Main Effects and Interactions
+
+```{r}
+#| layout-ncol: 2
+
+sum1 |> ggplot(aes(x = refClass, y = pct_change, fill = refClass)) + 
+  stat_summary(fun = mean, geom = "bar", position="dodge") +
+  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge")
+
+sum1 |> ggplot(aes(x = pct_goal, y = pct_change, fill = pct_goal)) + 
+  stat_summary(fun = mean, geom = "bar", position="dodge") +
+  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge")
+
+sum1 |> ggplot(aes(x = pct_goal, y = pct_change, fill = refClass)) + 
+  stat_summary(fun = mean, geom = "bar", position="dodge") +
+  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge")
+
+ sum1 |> ggplot(aes(x = pct_goal, y = pct_change, fill = refClass)) + 
+  stat_summary(fun = mean, geom = "bar", position="dodge") +
+  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge")
+
+sum1 |> ggplot(aes(x = pct_goal, y = pct_change, fill = refClass)) + 
+  stat_summary(fun = mean, geom = "bar", position="dodge") +
+  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") + facet_wrap(~state)
+
+sum1 |> ggplot(aes(x = pct_goal, y = pct_change, fill = rounded)) + 
+  stat_summary(fun = mean, geom = "bar", position="dodge") +
+  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") + facet_wrap(~rounded)
 
@@
 
+```
+
+```{r}
+
+# mixed effects model testing effect of refClass, state, and pct_goal on pct_change
+# marginal effects plots
+library(ggeffects)
+
+m1 <- lmer(pct_change ~ refClass  + (1+plan|id) + (1|state), data = sum1)
+summary(m1)
+ggpredict(m1, c("refClass")) |> plot()
+
+m1 <- lmer(pct_change ~ pct_goal  + (1|id) + (1|state), data = sum1)
+summary(m1)
+ggpredict(m1, c("pct_goal")) |> plot()
+
+m1 <- lmer(pct_change ~ pct_goal*rounded  + (1|id) + (1|state), data = sum1)
+summary(m1)
+ggpredict(m1, c("pct_goal","rounded")) |> plot()
+
+m1 <- lmer(pct_change ~ refClass*pct_goal*rounded  + (1|id) + (1|state), data = sum1)
+ggpredict(m1, c("refClass","pct_goal","rounded")) |> plot()
+
+m1 <- lmer(pct_change ~ state*pct_goal*rounded  + (1|id), data = sum1)
+summary(m1)
+ggpredict(m1, c("state","pct_goal","rounded")) |> plot()
+
+# m1 <- lmer(pct_change ~ refClass*pct_goal*rounded*state  + (1|id), data = sum1)
+# ggpredict(m1, c("refClass","pct_goal","rounded","state")) |> plot()
+
+```
+
+```{r}
+#| eval: false
+
 s2_long |> 
@@
   geom_point()
-
-```
-
-```{r}
-
-# mixed effects model testing effect of refClass, state, and pct_goal on pct_change
-
-m1 <- lmer(pct_change ~ refClass  + (1+plan|id) + (1|state), data = sum1)
-summary(m1)
-
-```
-
-```{r}
-#| eval: false
 
@@
 
+# sum1 |> 
+#   ungroup() |> 
+#   mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100) |>
+#   mutate(id=reorder(id,pct_change)) |> 
+#   ggplot(aes(y=id,x=pct_change)) + 
+#   geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
+#   geom_vline(aes(xintercept=goal_pct),linetype="dashed",alpha=.5) +
+#   ggh4x::facet_nested_wrap(refClass~pct_goal,axes="all",scales="free",ncol=2)
+
+# sum1 |> 
+#   ungroup() |> 
+#   mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100) |>
+#   mutate(id=reorder(id,pct_change)) |> 
+#   ggplot(aes(y=id,x=pct_change,col=refClass)) + 
+#   geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
+#   geom_vline(aes(xintercept=goal_pct),linetype="dashed",alpha=.5) +
+#   ggh4x::facet_nested_wrap(block~pct_goal,axes="all",scales="free",ncol=2) 
+# 
+# 
+# sum1 |> 
+#   ungroup() |> 
+#     mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100) |>
+#   ggplot(aes(x=state,y=pct_change,col=state)) + 
+#   geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
+#     geom_hline(aes(yintercept=goal_pct),linetype="dashed",alpha=.5) +
+#   ggdist::stat_halfeye()+
+#   ggh4x::facet_nested_wrap(block~pct_goal,axes="all",ncol=2)
+
 # s2_long <- s2 |>

### 2024-10-04 study2.qmd
updates

@@
 lightbox: true
+toc: false
 code-fold: true
@@
   html: 
-    page-layout: full
+    grid:
+      sidebar-width: 220px
+      body-width: 1200px
+      margin-width: 170px
+      gutter-width: 1.0rem
 ---
@@
 
-::: {.column-page-inset-right}
+::: column-page-right
+
+## Before and after excluding outliers
+
 ```{r}
 #| layout-ncol: 1
-#| fig.width: 12
-#| fig.height: 11
+#| fig.width: 10
+#| fig.height: 8
 
@@
   geom_rug(alpha=0.1) +
-  stat_boxplot(width=0.2, alpha=0.3) +
+  stat_boxplot(width=0.5, alpha=0.6) +
   facet_wrap(~refClass) +
@@
 
+- Facets indicate the Reference class (%, kWh, USD) and the goal % reduction (10%, 15%)
+- dashed lines also indicate goal percentage - dots on the line indicate participants who met the goal
+
 ```{r}
-#| fig.width: 12
-#| fig.height: 10
+#| fig.width: 11
+#| fig.height: 9
 
@@
 
-:::
-
 ## Interactive data table
@@
 # s2_long |> ggplot(aes(x = change)) + geom_histogram(bins = 30) + facet_wrap(~appliance, scales = "free")
-DT::datatable(sum1)
+sum1 |> mutate(refClass=factor(refClass),pct_goal=factor(pct_goal),rounded=factor(rounded),state=factor(state)) |>
+DT::datatable(extensions = 'Buttons',
+  options = list(
+    dom = 'Blfrtip',
+    buttons = list('copy', 'csv', 'excel', 'pdf', 'print'),
+    pageLength = 7,
+    autoWidth = TRUE
+  ),
+  filter = 'top') # Add filtering at the top of each column)
 #DT::datatable(sum1 |> group_by(id,block) |> summarise(diff(total_kWh),diff(orig_kWh)))
 ```
+:::
 
 ## Main Effects and Interactions
+
+**Click on plots to enlarge**
 
@@
   stat_summary(fun = mean, geom = "bar", position="dodge") +
-  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge")
+  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
+  labs(
+    title = "Effect of Reference Class",
+    x = "Reference Class",
+    y = "% Change",
+    fill = "Reference Class"
+  ) +
+  theme_minimal()
 
@@
   stat_summary(fun = mean, geom = "bar", position="dodge") +
-  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge")
+  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
+  labs(
+    title = "Effect of Goal % Reduction",
+    x = "Goal % Reduction",
+    y = "% Change",
+    fill = "Goal % Reduction"
+  ) +
+  theme_minimal()
+
+sum1 |> ggplot(aes(x = rounded, y = pct_change, fill = rounded)) + 
+  stat_summary(fun = mean, geom = "bar", position="dodge") +
+  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
+  labs(
+    title = "Effect of Rounding",
+    x = "Rounding",
+    y = "% Change",
+    fill = "Rounding"
+  ) +
+  theme_minimal()
 
@@
   stat_summary(fun = mean, geom = "bar", position="dodge") +
-  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge")
+  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
+  labs(
+    title = "Effect of Goal % Reduction and Reference Class",
+    x = "Goal % Reduction",
+    y = "% Change",
+    fill= "Reference Class"
+  ) +
+  theme_minimal()
 
@@
   stat_summary(fun = mean, geom = "bar", position="dodge") +
-  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge")
-
-sum1 |> ggplot(aes(x = pct_goal, y = pct_change, fill = refClass)) + 
-  stat_summary(fun = mean, geom = "bar", position="dodge") +
-  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") + facet_wrap(~state)
+  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") + 
+  facet_wrap(~state) +
+  labs(
+    title = "Effect of Goal % Reduction and Reference Class - Separated by State",
+    x = "Goal % Reduction",
+    y = "% Change",
+    fill= "Reference Class"
+  ) +
+  theme_minimal()
 
@@
   stat_summary(fun = mean, geom = "bar", position="dodge") +
-  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") + facet_wrap(~rounded)
+  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") + 
+  facet_wrap(~rounded) +
+  labs(
+    title = "Effect of Goal % Reduction and Rounding",
+    x = "Goal % Reduction",
+    y = "% Change",
+    fill= "Rounding"
+  ) +
+  theme_minimal()
 
@@
 #| 
-sum1 |> 
-  ungroup() |> 
-  mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100,
-         block=factor(block)) |>
-  group_by(refClass,block,pct_goal) |>
-  summarise(mean_pct_change = mean(pct_change),se_pct_change = sd(pct_change)/sqrt(n())) |>
-  ggplot(aes(x=mean_pct_change,y=refClass,col=block)) + 
-  geom_pointrange(aes(xmin=mean_pct_change-se_pct_change,xmax=mean_pct_change+se_pct_change),position = position_dodge(width=0.5)) +
-  ggforce::facet_col(~pct_goal)
-
-sum1 |> 
-  ungroup() |> 
-  mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100,
-         block=factor(block)) |>
-  group_by(refClass,,pct_goal,state) |>
-  summarise(mean_pct_change = mean(pct_change),se_pct_change = sd(pct_change)/sqrt(n())) |>
-  ggplot(aes(x=mean_pct_change,y=refClass,col=pct_goal)) + 
-  geom_pointrange(aes(xmin=mean_pct_change-se_pct_change,xmax=mean_pct_change+se_pct_change),position = position_dodge(width=0.5)) +
-  ggforce::facet_col(~state)
-
-sum1 |> 
-  ungroup() |> 
-  mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100,
-         block=factor(block)) |>
-  group_by(refClass,,pct_goal,state,goal_pct) |>
-  summarise(mean_pct_change = mean(pct_change),se_pct_change = sd(pct_change)/sqrt(n())) |>
-  ggplot(aes(x=mean_pct_change,y=refClass,col=state)) + 
-  geom_vline(aes(xintercept=goal_pct),linetype="dashed",alpha=.5) +
-  geom_pointrange(aes(xmin=mean_pct_change-se_pct_change,xmax=mean_pct_change+se_pct_change),position = position_dodge(width=0.5)) +
-  ggforce::facet_col(~pct_goal)
+# sum1 |> 
+#   ungroup() |> 
+#   mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100,
+#          block=factor(block)) |>
+#   group_by(refClass,block,pct_goal) |>
+#   summarise(mean_pct_change = mean(pct_change),se_pct_change = sd(pct_change)/sqrt(n())) |>
+#   ggplot(aes(x=mean_pct_change,y=refClass,col=block)) + 
+#   geom_pointrange(aes(xmin=mean_pct_change-se_pct_change,xmax=mean_pct_change+se_pct_change),position = position_dodge(width=0.5)) +
+#   ggforce::facet_col(~pct_goal)
+# 
+# 
+# sum1 |> 
+#   ungroup() |> 
+#   mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100,
+#          block=factor(block)) |>
+#   group_by(refClass,,pct_goal,state) |>
+#   summarise(mean_pct_change = mean(pct_change),se_pct_change = sd(pct_change)/sqrt(n())) |>
+#   ggplot(aes(x=mean_pct_change,y=refClass,col=pct_goal)) + 
+#   geom_pointrange(aes(xmin=mean_pct_change-se_pct_change,xmax=mean_pct_change+se_pct_change),position = position_dodge(width=0.5)) +
+#   ggforce::facet_col(~state)
+# 
+# 
+# sum1 |> 
+#   ungroup() |> 
+#   mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100,
+#          block=factor(block)) |>
+#   group_by(refClass,,pct_goal,state,goal_pct) |>
+#   summarise(mean_pct_change = mean(pct_change),se_pct_change = sd(pct_change)/sqrt(n())) |>
+#   ggplot(aes(x=mean_pct_change,y=refClass,col=state)) + 
+#   geom_vline(aes(xintercept=goal_pct),linetype="dashed",alpha=.5) +
+#   geom_pointrange(aes(xmin=mean_pct_change-se_pct_change,xmax=mean_pct_change+se_pct_change),position = position_dodge(width=0.5)) +
+#   ggforce::facet_col(~pct_goal)
 

### 2024-10-04 study2.qmd
updates

@@
 library(pacman)
-pacman::p_load(dplyr,purrr,tidyr,haven,tibble,ggplot2,gt,ggh4x,lme4)
+pacman::p_load(dplyr,purrr,tidyr,here, haven,tibble,ggplot2,gt,ggh4x,lme4)
+
+options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
+walk(c("fun_plot"), ~ source(here::here(paste0("Scripts/", .x, ".R"))))
+theme_set(theme_nice())
 
@@
 #| layout-ncol: 1
-#| fig.width: 10
-#| fig.height: 8
+#| fig.width: 9
+#| fig.height: 6
 
@@
   facet_wrap(~refClass) +
-  theme_minimal() +
   labs(title="Distribution of Percent Change by State and Reference Class - Before outlier exclusion",
@@
   facet_wrap(~refClass) +
-  theme_minimal() +
   labs(title="After outlier exclusion",
@@
 ```{r}
-#| fig.width: 11
+#| fig.width: 10
 #| fig.height: 9
@@
   geom_vline(aes(xintercept=goal_pct),linetype="dashed",alpha=.5) +
-  ggh4x::facet_nested_wrap(refClass~pct_goal,axes="all",scales="free",ncol=2)
+  ggh4x::facet_nested_wrap(refClass~pct_goal,axes="all",scales="free",ncol=2)  + 
+  labs(y="Participant Id", x="Percent Change", title="Individual Performance") +
+  theme(axis.text.y = element_text(family = "Manrope Light", face = "plain", size = rel(0.7)),)
 
@@
     fill = "Reference Class"
-  ) +
-  theme_minimal()
+  )
 
@@
     fill = "Goal % Reduction"
-  ) +
-  theme_minimal()
+  )
 
@@
     fill = "Rounding"
-  ) +
-  theme_minimal()
+  )
 
@@
     fill= "Reference Class"
-  ) +
-  theme_minimal()
+  )
 
@@
     fill= "Rounding"
-  ) +
-  theme_minimal()
+  )
 

### 2024-10-04 Scripts/fun_plot.R

@@ -1,0 +1,99 @@
+
+
+
+theme_blank <- theme(
+  panel.grid = element_blank(),
+  axis.text = element_blank(),
+  axis.title = element_blank(),
+  axis.ticks = element_blank()
+)
+
+
+stat_bar <- list(stat_summary(fun=mean, geom="bar", position=position_dodge()),
+                 stat_summary(fun.data=mean_se, geom="errorbar", color="black", position=position_dodge(), linewidth=.5))
+
+ 
+round_tibble <- function(tbl, rn) {
+  tbl %>% 
+    mutate(across(where(is.numeric), ~round(., rn)))
+}
+
+
+
+
+theme_nice_dist <- function() {
+  theme_nice() +
+    theme(
+      panel.grid = element_blank(),
+      panel.spacing.x = unit(10, units = "pt"),
+      axis.ticks.x = element_line(linewidth = 0.25),
+      axis.text.y = element_blank()
+    )
+}
+
+
+# theme_set(theme_nice())
+
+theme_nice <- function() {
+  theme_minimal(base_family = "Manrope") +
+    theme(
+      panel.grid.minor = element_blank(),
+      plot.title = element_text(family = "Manrope Extrabold", face = "plain", size = rel(1.35)),
+      plot.subtitle = element_text(family = "Manrope Medium", face = "plain", size = rel(1.2)),
+      axis.title = element_text(family = "Manrope SemiBold", face = "plain", size = rel(1)),
+      axis.title.x = element_text(hjust = .5),
+      axis.title.y = element_text(hjust = .5),
+      axis.text = element_text(family = "Manrope Light", face = "plain", size = rel(0.8)),
+      strip.text = element_text(
+        family = "Manrope", face = "bold",
+        size = rel(.95), hjust = 0
+      ),
+      legend.position = "top",
+      legend.text = element_text(family = "Manrope Light", face = "plain", size = rel(0.95)),
+      strip.background = element_rect(fill = "grey90", color = NA)
+    )
+}
+
+
+theme_nice_pdf <- function() {
+ theme_minimal() +
+    theme(
+      panel.grid.minor = element_blank(),
+      plot.title = element_text( face = "plain", size = rel(1.35)), #family = "Manrope Extrabold",
+      plot.subtitle = element_text(face = "plain", size = rel(1.2)), # family = "Manrope Medium", 
+      axis.title = element_text(face = "plain", size = rel(1)), #family = "Manrope SemiBold", 
+      axis.title.x = element_text(hjust = .5),
[…]

### 2024-10-04 Scripts/fun_plot.R
update

@@ -1,4 +1,4 @@
-
+pacman::p_load(ggplot2,wesanderson,glue)
 
 
 theme_blank <- theme(
@@ -97,3 +97,49 @@ theme_nice_b <- function() {
 
 
 
+geom_flat_violin <- function(mapping = NULL, data = NULL, stat = "ydensity",
+                             position = "dodge", trim = TRUE, scale = "area",
+                             show.legend = NA, inherit.aes = TRUE, ...) {
+  layer(
+    data = data,
+    mapping = mapping,
+    stat = stat,
+    geom = GeomFlatViolin,
+    position = position,
+    show.legend = show.legend,
+    inherit.aes = inherit.aes,
+    params = list(
+      trim = trim,
+      scale = scale,
+      ...
+    )
+  )
+}
+
+
+
+
+col_themes <- tibble::lst(darjeeling = c(wes_palette("Darjeeling1"),wes_palette("Darjeeling2"), ggokabeito::palette_okabe_ito(),
+  wes_palette("AsteroidCity1"), wes_palette("AsteroidCity2")), 
+                          wes2 = wes_palette("AsteroidCity1"), 
+                          okabeito = c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999", "#000000"))
+
+# okabeito <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7", "#999999", "#000000")
+# wes2 <- wes_palette("AsteroidCity1")
+
+
+#### COLOR THEME ##### 
+# darjeeling <- c(wes_palette("Darjeeling1"),wes_palette("Darjeeling2"))
+#cat(darjeeling)
+
+
+scale_colour_discrete <- function(...) {
+  scale_colour_manual(..., values = col_themes$darjeeling)
+}
+
+scale_fill_discrete <- function(...) {
+  scale_fill_manual(..., values = col_themes$darjeeling)
+}
+
+options(ggplot2.continuous.colour=col_themes$darjeeling)
+options(ggplot2.continuous.fill = col_themes$darjeeling)

### 2024-10-09 study2.qmd
update
add calculator and education plots & tables; add % of sbjs matching target plots

@@
 library(pacman)
-pacman::p_load(dplyr,purrr,tidyr,here, haven,tibble,ggplot2,gt,ggh4x,lme4)
+pacman::p_load(dplyr,purrr,tidyr,here, haven,tibble,ggplot2,gt,ggh4x,lme4,flextable)
 
@@
 s2 <- s2 |> mutate(id = as.numeric(format(EndDate, "%H%M%S"))) |> relocate(id)
-
 s2 <- s2 |> 
@@
 )) |> 
-  relocate(refClass,state1,state2, .after = id)
+  rename(calc=MATH01) |>  # replace NA's in calc with 0's
+  mutate(calc=factor(ifelse(is.na(calc),0,calc))) |> 
+  relocate(refClass,state1,state2,calc, .after = id)
 
-s2_b1 <- s2 |> select(id,state1,refClass,TENPERCENT,ROUNDED, ACFAMILY1:TOTALAPTWO1) |> 
+# recode DEMO4 variable based on this mapping: 
+# 1 = "Some schooling, but no diploma or degree" 
+# 2 = "Highschool diploma or GED" 
+# 3 = "Some college" 
+# 4 = "College degree" 
+# 5 = "Some graduate school" 
+# 6 = "Graduate degree" 
+# 7 = "Choose not to answer"
+
+s2 <- s2 |> rename(edu=DEM04) |>
+  mutate(edu = factor(case_when(
+    edu == '1' ~ "Some schooling, but no diploma or degree",
+    edu == 2 ~ "Highschool diploma or GED",
+    edu == 3 ~ "Some college",
+    edu == 4 ~ "College degree",
+    edu == 5 ~ "Some graduate school",
+    edu == 6 ~ "Graduate degree",
+    edu == 7 ~ "Choose not to answer",
+    TRUE ~ NA_character_
+  ))) |> relocate(edu, .after = id)
+
+s2_b1 <- s2 |> select(id,state1,refClass,TENPERCENT,ROUNDED, calc, edu,ACFAMILY1:TOTALAPTWO1) |> 
   mutate( 
@@
   pct_goal=case_when (TENPERCENT == 1 ~ "10%", TENPERCENT == 2 ~ "15%"), 
+  pct=case_when (TENPERCENT == 1 ~ .10, TENPERCENT == 2 ~ .15),
   rounded = case_when(ROUNDED == 1 ~ "Rounded", ROUNDED == 2 ~ "Not Rounded")
@@
 
-s2_b2 <- s2 |> select(id,state2,refClass,TENPERCENT,ROUNDED, ACFAMILY2:TOTALAPTWO2) |> 
+s2_b2 <- s2 |> select(id,state2,refClass,TENPERCENT,ROUNDED, calc,edu,ACFAMILY2:TOTALAPTWO2) |> 
   mutate( 
@@
   pct_goal=case_when (TENPERCENT == 1 ~ "15%", TENPERCENT == 2 ~ "10%"), 
+  pct=case_when (TENPERCENT == 1 ~ .10, TENPERCENT == 2 ~ .15),
   rounded = case_when(ROUNDED == 1 ~ "Not Rounded", ROUNDED == 2 ~ "Rounded")
@@
   filter(appliance !="TOTAL") |> 
-  group_by(id,refClass,state,pct_goal,rounded,block,plan) |> 
+  group_by(id,refClass,state,pct_goal,pct,rounded,block,plan,calc,edu) |> 
   summarise(total_kWh = sum(value),orig_kWh=sum(family), pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), state_dif=mean(state_dif))
@@
                                      z_score=(pct_change-change_mean)/change_sd,
-                                     is_outlier=abs(z_score)>2.5)
+                                     is_outlier=abs(z_score)>2.75)
 
@@
 
+## Looking at % of subjects who hit the target reduction
+
+```{r}
+
+# Step 1: Calculate whether participants matched the target pct_goal
+sum1 <- s2_long |> 
+  filter(appliance != "TOTAL") |> 
+  group_by(id, refClass, state, pct_goal, pct, rounded, block, plan, calc, edu) |> 
+  summarise(
+    total_kWh = sum(value),
+    orig_kWh = sum(family),
+    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
+    state_dif = mean(state_dif),
+    .groups = "drop"
+  ) |>
+  mutate(matched_goal = (pct_change == pct))
+
+# Step 2: Summarize the percentage of participants meeting the goal for each condition
+summary_pct <- sum1 |>
+  group_by(refClass, rounded, pct_goal) |>
+  summarise(
+    matched_count = sum(matched_goal),
+    total_count = n(),
+    pct_matched = matched_count / total_count * 100,
+    .groups = "drop"
+  )
+
+# Step 3: Create plots for the percentage of subjects in each condition who matched the target pct
+ggplot(summary_pct, aes(x = refClass, y = pct_matched, fill = rounded)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  facet_wrap(~ pct_goal) +
+  labs(
+    title = "Percentage of Participants Matching Target Pct by Condition",
+    x = "Reference Class",
+    y = "Percentage Matched (%)",
+    fill = "Rounding"
+  ) +
+  theme_minimal()
+
+summary_pct |>
+  gt() |>
+  tab_header(
+    title = md("**Percentage of Participants Matching Target Pct by Condition**"),
+    subtitle = "Grouped by Reference Class, Rounding, and Pct Goal"
+  ) |>
+  fmt_number(
+    columns = c(pct_matched),
+    decimals = 2
+  ) |>
+  cols_label(
+    refClass = "Reference Class",
+    rounded = "Rounding",
+    pct_goal = "Pct Goal",
+    matched_count = "Matched Count",
+    total_count = "Total Count",
+    pct_matched = "Percentage Matched (%)"
+  ) |>
+  cols_align(
+    align = "center",
+    columns = everything()
+  ) |>
+  tab_spanner(
+    label = "Participants",
+    columns = c(matched_count, total_count, pct_matched)
+  )
+
+```
+
+```{r}
+#| layout-ncol: 2
+#| fig.height: 10
+#| fig.width: 12
+
+# Step 1: Calculate whether participants matched the target pct_goal within different tolerances
+sum1a <- s2_long |> 
+  filter(appliance != "TOTAL") |> 
+  group_by(id, refClass, state, pct_goal, pct, rounded, block, plan, calc, edu) |> 
+  summarise(
+    total_kWh = sum(value),
+    orig_kWh = sum(family),
+    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
+    state_dif = mean(state_dif),
+    .groups = "drop"
+  ) |>
+  mutate(
+    target_pct = as.numeric(sub("%", "", pct_goal)) / 100,
+    match_level = case_when(
+      abs(pct_change - target_pct) <= 0.001 ~ "Exact Match",
+      abs(pct_change - target_pct) <= 0.05 ~ "Within 5%",
+      abs(pct_change - target_pct) <= 0.10 ~ "Within 10%",
+      abs(pct_change - target_pct) <= 0.20 ~ "Within 20%",
+      TRUE ~ "Over 20%"
+    ),
+    match_level = factor(match_level, levels = c("Exact Match", "Within 5%", "Within 10%", "Within 20%", "Over 20%"))
+  )
+
+# Step 2: Summarize the percentage of participants in each match level for each condition
+summary_pct <- sum1a |>
+  group_by(refClass, rounded, pct_goal, match_level) |>
+  summarise(count = n(), .groups = "drop") |>
+  group_by(refClass, rounded, pct_goal) |>
+  mutate(
+    total_count = sum(count),
+    percentage = count / total_count * 100
+  ) |>
+  ungroup()
+
+# Step 3: Create plot
+ggplot(summary_pct, aes(x = interaction(refClass, rounded), y = percentage, fill = match_level)) +
+  geom_bar(stat = "identity", position = "stack") +
+  facet_wrap(~ pct_goal, scales = "free_x") +
+  scale_fill_brewer(palette = "Set2") +
+  labs(
+    title = "Percentage of Participants by Match Level",
+    subtitle = "Grouped by Reference Class, Rounding, and Pct Goal",
+    x = "Reference Class and Rounding",
+    y = "Percentage (%)",
+    fill = "Match Level"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+ggplot(summary_pct, aes(x = refClass, y = percentage, fill = match_level)) +
+  geom_bar(stat = "identity",position=position_dodge()) +
+  facet_wrap(rounded ~ pct_goal, scales = "free_x") +
+  scale_fill_brewer(palette = "Set2") +
+  labs(
+    title = "Percentage of Participants by Match Level",
+    subtitle = "Grouped by Reference Class, Rounding, and Pct Goal",
+    x = "Reference Class and Rounding",
+    y = "Percentage (%)",
+    fill = "Match Level"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+```
+## Cumulative plot
+
+**Click on plots to enlarge**
+
+```{r}
+#| layout-ncol: 2
+#| fig.height: 11
+#| fig.width: 12
+
+# Step 1: Calculate match levels with cumulative groups
+sum1c <- s2_long |>
+  filter(appliance != "TOTAL") |>
+  group_by(id, refClass, state, pct_goal, pct, rounded, block, plan, calc, edu) |>
+  summarise(
+    total_kWh = sum(value),
+    orig_kWh = sum(family),
+    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
+    state_dif = mean(state_dif),
+    .groups = "drop"
+  ) |>
+  mutate(
+    target_pct = as.numeric(sub("%", "", pct_goal)) / 100,
+    match_level = case_when(
+      abs(pct_change - target_pct) <= 0.001 ~ "Exact Match",
+      abs(pct_change - target_pct) <= 0.05 ~ "Within 5%",
+      abs(pct_change - target_pct) <= 0.10 ~ "Within 10%",
+      abs(pct_change - target_pct) <= 0.20 ~ "Within 20%",
+      TRUE ~ "Over 20%"
+    ),
+    match_level = factor(
+      match_level, 
+      levels = c("Exact Match", "Within 5%", "Within 10%", "Within 20%", "Over 20%")
+    )
+  )
+
+# Step 2: Create cumulative groups for each match level
+summary_pct <- sum1c |>
+  group_by(refClass, rounded, pct_goal, match_level) |>
+  summarise(count = n(), .groups = "drop") |>
+  group_by(refClass, rounded, pct_goal) |>
+  mutate(
+    total_count = sum(count),
+    cumulative_count = cumsum(count),
+    cumulative_percentage = cumulative_count / total_count * 100
+  ) |>
+  ungroup()
+
+# Step 3: Create a plot for cumulative percentages
+ggplot(summary_pct, aes(x = refClass, y = cumulative_percentage, fill = match_level)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  facet_wrap(pct_goal~rounded, scales = "free_x") +
+  scale_fill_brewer(palette = "Set2") +
+  labs(
+    title = "Cumulative Percentage of Participants by Match Level",
+    subtitle = "Grouped by Reference Class, Rounding, and Pct Goal",
+    x = "Reference Class",
+    y = "Cumulative Percentage (%)",
+    fill = "Match Level"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+ggplot(summary_pct, aes(x = refClass, y = cumulative_percentage, fill = match_level)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  facet_wrap(~rounded, scales = "free_x") +
+  scale_fill_brewer(palette = "Set2") +
+  labs(
+    title = "Cumulative Percentage of Participants by Match Level",
+    subtitle = "Grouped by Reference Class, Rounding, and Pct Goal",
+    x = "Reference Class",
+    y = "Cumulative Percentage (%)",
+    fill = "Match Level"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+ggplot(summary_pct, aes(x = refClass, y = cumulative_percentage, fill = match_level)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  scale_fill_brewer(palette = "Set2") +
+  labs(
+    title = "Cumulative Percentage of Participants by Match Level",
+    subtitle = "Grouped by Reference Class, Rounding, and Pct Goal",
+    x = "Reference Class",
+    y = "Cumulative Percentage (%)",
+    fill = "Match Level"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+ggplot(summary_pct, aes(x = rounded, y = cumulative_percentage, fill = match_level)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  scale_fill_brewer(palette = "Set2") +
+  labs(
+    title = "Cumulative Percentage of Participants by Match Level",
+    subtitle = "Grouped by Reference Class, Rounding, and Pct Goal",
+    x = "Rounding",
+    y = "Cumulative Percentage (%)",
+    fill = "Match Level"
+  ) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+```
+
 ## Main Effects and Interactions
@@
     fill= "Rounding"
+  )
+
+```
+
+## Calculator and education
+
+```{r}
+
+# table with only education level counts, just use kable
+s2 |> 
+  group_by(edu) |> 
+  summarise(n = n(), .groups = "drop") |>
+  knitr::kable(caption = "Count of Items by Education Level", format = "html")
+
+# table with only calculator counts
+
+s2 |> 
+  mutate(calc = if_else(calc == 0, "No Calculator", "Used Calculator")) |> 
+  group_by(calc) |> 
+  summarise(n = n(), .groups = "drop") |>
+  knitr::kable(caption = "Count of Items by Calculator Use", format = "html")
+
+# s2 |> 
+#   mutate(rounded = case_when(ROUNDED == 1 ~ "Not Rounded", ROUNDED == 2 ~ "Rounded")) |> 
+#   group_by(calc, refClass, rounded) |> 
+#   summarise(n = n(), .groups = "drop") |>
+#   group_by(calc, refClass) |>
+#   mutate(count_stats = as.character(n)) |>
+#   ungroup() |>
+#   tabulator(
+#     rows = c("calc", "refClass"),
+#     columns = "rounded",
+#     `Count` = as_paragraph(count_stats)
+#   ) |>
+#   as_flextable() |> 
+#   align(align = "left", part = "all") |>
+#   autofit() |>
+#   set_caption(
+#     caption = "Count of Items by Calculation Type, Reference Class, and Rounded Value",
+#     style = "Table Caption"
+#   )
+
+s2 |> 
+  mutate(
+    rounded = if_else(ROUNDED == 1, "Not Rounded", "Rounded"),
+    calc = if_else(calc == 0, "No Calculator", "Used Calculator")
+  ) |> 
+  count(calc, refClass, rounded) |>
+  pivot_wider(
+    names_from = c(calc, rounded),
+    values_from = n,
+    values_fill = 0
+  ) |>
+  gt() |>
+  tab_header(
+    title = md("**Count of Items by Calculator Use, Reference Class, and Rounding**")
+  ) |>
+  cols_align(align = "left") |>
+  tab_spanner_delim(delim = "_") |>
+  cols_label(
+    `No Calculator_Not Rounded` = "Not Rounded",
+    `No Calculator_Rounded` = "Rounded",
+    `Used Calculator_Not Rounded` = "Not Rounded",
+    `Used Calculator_Rounded` = "Rounded"
+  )
+
+# table with education level counts
+
+s2 |>
+  group_by(edu, refClass) |>
+  summarise(n = n(), .groups = "drop") |>
+  pivot_wider(names_from = refClass, values_from = n, values_fill = 0) |>
+  gt(rowname_col = "edu") |>
+  tab_header(title = md("**Count of Items by Education Level and Reference Class**")) |>
+  cols_align(align = "left", columns = everything())
+
+# table with calculator and education level counts
+s2 |>
+  mutate(calc = if_else(calc == 0, "No Calculator", "Used Calculator")) |> 
+  group_by(calc, edu) |>
+  summarise(n = n(), .groups = "drop") |>
+  pivot_wider(names_from = edu, values_from = n, values_fill = 0) |>
+  gt(rowname_col = "calc") |>
+  tab_header(title = md("**Count of Items by Calculator Use and Education Level**")) |>
+  cols_align(align = "left", columns = everything())
+
+sum1 |> ggplot(aes(x = calc, y = pct_change, fill = calc)) + 
+  stat_summary(fun = mean, geom = "bar", position="dodge") +
+  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
+  labs(
+    title = "Effect of Calculator Use",
+    x = "Calculator Use",
+    y = "% Change",
+    fill = "Calculator Use"
+  )
+
+## education
+
+sum1 |> ggplot(aes(x = edu, y = pct_change, fill = edu)) + 
+  stat_summary(fun = mean, geom = "bar", position="dodge") +
+  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
+  labs(
+    title = "Effect of Education",
+    x = "Education",
+    y = "% Change",
+    fill = "Education"
   )

### 2024-10-18 study1.qmd
study 1 analysis scripts

---
title: "Study 1 Analysis"
author:
- name: Thomas E. Gorman
  url: https://tegorman13.github.io/
  affiliations: 
  - name:  Communication and Cognition Lab, Purdue University, USA
    affiliation-url: https://web.ics.purdue.edu/~treimer/
lightbox: true
toc: false
code-fold: true
code-tools: true
execute:
  #echo: false
  warning: false
format:
  html: 
    grid:
      sidebar-width: 220px
      body-width: 1200px
      margin-width: 170px
      gutter-width: 1.0rem
---

```{r}
library(pacman)
pacman::p_load(dplyr,purrr,tidyr,here,haven,tibble,ggplot2,gt,ggh4x,lme4,flextable)

options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("fun_plot"), ~ source(here::here(paste0("Scripts/", .x, ".R"))))
theme_set(theme_nice())

```

```{r}
#| eval: FALSE

# s1 <- haven::read_sav(here::here("data/data_qualtrics/Frequency & Probability - 9-23-18_October 6, 2024_20.39.sav")) # length(unique(s1$ResponseId)) = 588
# qd <- unique(s1$ResponseId)

s1 <- haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav"))

#gd <- unique(s1$ResponseId)
# gd %in% qd
# qd %in% gd

s1 <- s1 |> 
  mutate(
    state1 = case_when(
      Colorado == 1 ~ "Colorado",
      Massachusetts == 1 ~ "Massachusetts",
      California == 1 ~ "California",
      Texas == 1 ~ "Texas",
      TRUE ~ NA_character_
    ),
    state2 = case_when(
      Colorado == 2 ~ "Colorado",
      Massachusetts == 2 ~ "Massachusetts",
      California == 2 ~ "California",
      Texas == 2 ~ "Texas",
      TRUE ~ NA_character_
    )) |> 
  mutate(refClass = case_when(
  ReferenceClass == 1 ~ "USD",
  ReferenceClass == 2 ~ "kWh",
  ReferenceClass == 3 ~ "Percentage"
)) |> 
  rename(calc=MATH01) |>  # replace NA's in calc with 0's
  rename(id=ResponseId) |>
  mutate(calc=factor(ifelse(is.na(calc),0,calc))) |> 
  relocate(refClass,state1,state2,calc, .after = id)

s1 <- s1 |> rename(edu=DEM04) |>
  mutate(edu = factor(case_when(
    edu == 1 ~ "Some schooling, but no diploma or degree",
    edu == 2 ~ "Highschool diploma or GED",
    edu == 3 ~ "Some college",
    edu == 4 ~ "College degree",
    edu == 5 ~ "Some graduate school",
    edu == 6 ~ "Graduate degree",
    edu == 7 ~ "Choose not to answer",
    TRUE ~ NA_character_
  ))) |> relocate(edu, .after = id)

# colnames(s1)
# colnames(s1)[1000:1500]
# colnames(s1)[1500:1843]

# select all columns that start with "A"
# s12 <- s1 |> select(id,state1,state2,refClass,calc,edu,starts_with("A")) |> 
#   select(-starts_with("AIND"))

# 
# t <- s12 |> filter(state1=="Texas") |> 
#   # select only columns that contain "TEX" in the name
#   select(id,state1,state2,refClass,calc,edu,contains("TEX")) 
# 
# t2 <- s12 |> filter(state2=="Texas") |> 
#   select(id,state1,state2,refClass,calc,edu,contains("TEX"))
#

trim_na_and_empty_columns <- function(df) {
  df %>%
    select(where(~!(all(is.na(.)) | (is.character(.) && all(trimws(.) == ""))))) |> 
    select(-starts_with("AIND"))
}

# Create a list of trimmed datasets for each unique ID
dAll <- s1 %>%
  group_by(id) %>%
  group_split() %>%
  map(trim_na_and_empty_columns) %>%
  set_names(unique(s1$id))

# modalities <- c(A = "kWh", D = "USD", P = "Percentage")
# states <- c(CAL = "California", MASS = "Massachusetts", COL = "Colorado", TEX = "Texas")

parse_action_plan <- function(df) {

  action_plan_cols <- df %>% 
  select(-matches("01$|02$|03$|04$|05$|06$|33$|34$|35$|36$|37$")) %>%
    select(
      matches("^[ADP]\\d+(CAL|MASS|COL|TEX)\\d{2,3}$"), #,  # Select action plan columns
      -matches("32$")                                     # Exclude columns ending with '32'
    ) %>%
    names()

  df_long <- df %>%
  select(-matches("01$|02$|03$|04$|05$|06$")) %>%
    pivot_longer(
      cols = all_of(action_plan_cols),
      names_to = c("modality", "state", "item_num"),
      names_pattern = "^([ADP])\\d+(CAL|MASS|COL|TEX)(\\d+)$",
      values_to = "value"
    ) |> select(id,state1,state2,refClass,calc,edu,modality,state,item_num,value)

  df_long <- df_long |> 
  mutate(
    appliance = case_when(
      item_num %in% c("07","08","09",7:10) ~ "Cooling",
      item_num %in% 11:14 ~ "Heating",
      item_num %in% 15:18 ~ "Water Heating",
      item_num %in% 19:22 ~ "Refrigerator",
      item_num %in% 23:26 ~ "Other Appliance",
      item_num %in% 27:30 ~ "Total kWh",
      item_num == 31 ~ "Preferred Plan",
      item_num %in% 33:37 ~ "Reduction Choices",
      TRUE ~ NA_character_
    )
  ) |> 
  # item_type - family, state, plan1 or plan2
  mutate(item_type=case_when(
    item_num %in% c("07","11","15","19","23","27") ~ "family",
    item_num %in% c("08","12","16","20","24","28") ~ "state_avg",
    item_num %in% c("09","13","17","21","25","29") ~ "plan1",
    item_num %in% c("10","14","18","22","26","30") ~ "plan2",
    TRUE ~ "other"
  ))

# pivot family and state_avg into their own separate columns, plan1 and plan2 into a combined column called plan
df_long <- df_long |> 
  filter(item_type !="other") |> 
  select(-item_num) |> 
  pivot_wider(names_from = c("item_type"), values_from = value) |> 
  pivot_longer(cols = c("plan1","plan2"), names_to = "plan", values_to = "value")

}

test <- dAll |> # map each item in dAll into parse function
  map(parse_action_plan)

s1 <- bind_rows(test)

s1 <- s1 |> group_by(id) |> 
  mutate(trial=row_number(), block=ifelse(trial>=12,2,1), plan=as_factor(plan)) |> 
  select(-trial)

saveRDS(s1, here::here("data/s1_processed.rds"))

```

```{r}
s1 <- readRDS(here::here("data/s1_processed.rds"))

s1 <- s1 |> mutate(change = value-family, state_dif = value-state_avg)

sum1 <- s1 |> 
  filter(appliance !="Total kWh") |> 
  group_by(id,refClass,state,block,plan,calc,edu) |> 
  summarise(total_kWh = sum(value),orig_kWh=sum(family), pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), state_dif=mean(state_dif)) |> 
  mutate(matched_goal = (pct_change == .15))

```

```{r}

sum1 |> ggplot(aes(x=pct_change)) +
  geom_density(aes(fill=state), alpha=0.5) +
  ggdist::stat_halfeye() +
  geom_rug(alpha=0.1) +
  stat_boxplot(width=0.5, alpha=0.6) +
  #facet_wrap(~refClass) +
  labs(title="Distribution of Percent Change by State and Reference Class - Before outlier exclusion",
       x="Percent Change",
       y="Density") +
  scale_fill_brewer(palette="Set2")

# descriptive table to do with plot
sum1 |> 
  group_by(refClass,state) |> 
  summarise(mean_pct_change=median(pct_change), sd_pct_change=sd(pct_change), n=n()) |> 
  gt() |> 
  tab_header(title="Descriptive Statistics of Percent Change by State and Reference Class")

```

```{r}
#| fig.width: 10
#| fig.height: 9

  # add geom_vline with x intercept based on pct_goal of the data (which is currently a string of form 10%)
sum1 |> 
  filter(block==1) |>
  ungroup() |> 
  mutate(id=reorder(id,pct_change)) |> 
  ggplot(aes(y=id,x=pct_change)) + 
  geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
  geom_vline(aes(xintercept=.15),linetype="dashed",alpha=.5) +
  ggh4x::facet_nested_wrap(~refClass,axes="all",scales="free",ncol=2)  + 
  labs(y="Participant Id", x="Percent Change", title="Individual Performance") +
  theme(axis.text.y = element_text(family = "Manrope Light", face = "plain", size = rel(0.7)))
```

## Looking at % of subjects who hit the target reduction

```{r}

# Step 2: Summarize the percentage of participants meeting the goal for each condition
summary_pct <- sum1 |>
  group_by(refClass) |>
  summarise(
    matched_count = sum(matched_goal),
    total_count = n(),
    pct_matched = matched_count / total_count * 100,
    .groups = "drop"
  )

# Step 3: Create plots for the percentage of subjects in each condition who matched the target pct
ggplot(summary_pct, aes(x = refClass, y = pct_matched, fill = refClass)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(
    title = "Percentage of Participants Matching Target Pct by Condition",
    x = "Reference Class",
    y = "Percentage Matched (%)",
    fill = "Rounding"
  ) +
  theme_minimal()

summary_pct |>
  gt() |>
  tab_header(
    title = md("**Percentage of Participants Matching Target Pct by Condition**"),
    subtitle = "Grouped by Reference Class, Rounding, and Pct Goal"
  ) |>
  fmt_number(
    columns = c(pct_matched),
    decimals = 2
  ) |>
  cols_label(
    refClass = "Reference Class",
    rounded = "Rounding",
    pct_goal = "Pct Goal",
    matched_count = "Matched Count",
    total_count = "Total Count",
    pct_matched = "Percentage Matched (%)"
  ) |>
  cols_align(
    align = "center",
    columns = everything()
  ) |>
  tab_spanner(
    label = "Participants",
    columns = c(matched_count, total_count, pct_matched)
  )

```

### 2024-10-18 study1.qmd
render initial study 1 analysis

@@
 
+# for id R_1kFwgoKAmNTwCs3 - when state== TEX, appliance == Heating, check values of family column
+s1 |> filter(id=="R_1kFwgoKAmNTwCs3" & state=="TEX" & appliance=="Heating") |> select(family) 
+# for id R_1kFwgoKAmNTwCs3 - when state== TEX, appliance == Heating, set values of family column to 6118
+s1 <- s1 |> mutate(family=ifelse(id=="R_1kFwgoKAmNTwCs3" & state=="TEX" & appliance=="Heating",6118,family))
+
 saveRDS(s1, here::here("data/s1_processed.rds"))
@@
 
+k <- sum1 |> group_by(id) |>  summarize(total_kWh=sum(total_kWh),orig_kWh=sum(orig_kWh),pct_change=round((orig_kWh-total_kWh)/orig_kWh,3)) |> 
+  mutate(matched_goal = (pct_change == .15))
+
+outliers <- sum1 |>  ungroup() |> group_by(state) |> mutate(change_mean=mean(pct_change,na.rm=TRUE),change_sd=sd(pct_change,na.rm=TRUE),
+                                     z_score=(pct_change-change_mean)/change_sd,
+                                     is_outlier=abs(z_score)>2.75)
+outlier_id <- outliers |> filter(is_outlier) |> pull(id) |> unique()
 ```
@@
   tab_header(title="Descriptive Statistics of Percent Change by State and Reference Class")
+
+sum1 <- sum1 |> filter(!id %in% outlier_id)
 
@@
     refClass = "Reference Class",
-    rounded = "Rounding",
-    pct_goal = "Pct Goal",
     matched_count = "Matched Count",

### 2024-10-18 study2.qmd
render initial study 1 analysis

@@
   labs(y="Participant Id", x="Percent Change", title="Individual Performance") +
-  theme(axis.text.y = element_text(family = "Manrope Light", face = "plain", size = rel(0.7)),)
-
+  theme(axis.text.y = element_text(family = "Manrope Light", face = "plain", size = rel(0.7)))
 ```

### 2024-10-23 study1.qmd
updates

@@
 library(pacman)
-pacman::p_load(dplyr,purrr,tidyr,here,haven,tibble,ggplot2,gt,ggh4x,lme4,flextable)
+pacman::p_load(dplyr,purrr,tidyr,here,haven,tibble,brms,bayestestR,emmeans,tidybayes,modelsummary,
+               ggplot2,gt,ggh4x,lme4,flextable,kableExtra)
 
@@
 theme_set(theme_nice())
-
 ```
@@
 s1 <- s1 |> group_by(id) |> 
-  mutate(trial=row_number(), block=ifelse(trial>=12,2,1), plan=as_factor(plan)) |> 
+  mutate(trial=row_number(), block=ifelse(trial>=12,2,1), plan=as_factor(plan),pct_goal=.15) |> 
   select(-trial)
@@
 
-s1 <- s1 |> mutate(change = value-family, state_dif = value-state_avg)
+s1 <- s1 |> mutate(change = value-family, state_dif = value-state_avg,calc = if_else(calc == 0, "No Calculator", "Used Calculator"))
 
-sum1 <- s1 |> 
+s1_agg <- s1 |> 
   filter(appliance !="Total kWh") |> 
-  group_by(id,refClass,state,block,plan,calc,edu) |> 
+  group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
   summarise(total_kWh = sum(value),orig_kWh=sum(family), pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), state_dif=mean(state_dif)) |> 
-  mutate(matched_goal = (pct_change == .15))
+  mutate(matched_goal = (pct_change == pct_goal))
 
-k <- sum1 |> group_by(id) |>  summarize(total_kWh=sum(total_kWh),orig_kWh=sum(orig_kWh),pct_change=round((orig_kWh-total_kWh)/orig_kWh,3)) |> 
-  mutate(matched_goal = (pct_change == .15))
+k <- s1_agg |> group_by(id,pct_goal) |>  summarize(total_kWh=sum(total_kWh),orig_kWh=sum(orig_kWh),pct_change=round((orig_kWh-total_kWh)/orig_kWh,3)) |> 
+  mutate(matched_goal = (pct_change == pct_goal))
 
-outliers <- sum1 |>  ungroup() |> group_by(state) |> mutate(change_mean=mean(pct_change,na.rm=TRUE),change_sd=sd(pct_change,na.rm=TRUE),
+s1_agg4 <- s1_agg |> group_by(id,refClass) |> 
+  summarise(mg=sum(matched_goal),n=n(), pct=mg/n)
+
+s1_agg2 <- s1_agg |> group_by(id,refClass,state) |> 
+  summarise(mg=sum(matched_goal),n=n(), pct=mg/n)
+
+outliers <- s1_agg |>  ungroup() |> group_by(state) |> mutate(change_mean=mean(pct_change,na.rm=TRUE),change_sd=sd(pct_change,na.rm=TRUE),
                                      z_score=(pct_change-change_mean)/change_sd,
-                                     is_outlier=abs(z_score)>2.75)
+                                     is_outlier=abs(z_score)>3.0)
 outlier_id <- outliers |> filter(is_outlier) |> pull(id) |> unique()
+
+```
+
+```{r}
+#| panel: tabset
+#| output: asis
+
+t1 <- datasummary_crosstab(data=s1_agg |> group_by(id) |> slice(1),
+                           calc~refClass,
+                           statistic = 1 ~ N + Percent("row"),
+                           output='markdown')
+t2 <- datasummary(data=s1_agg4 ,(` `= pct)*refClass ~ Mean+Median+Mode+sd+Histogram,output='markdown',title="% of trials where goal was matched")
+t3 <- datasummary(data=s1_agg, pct_change*refClass~Mean+Median+Mode, output='markdown')
+list_of_tables <- list(t1,t2,t3)
+names(list_of_tables) <- c("Calculator by Ref Class tallies","% of trials with matched goal","% change in plan")
+
+for (itable in names(list_of_tables)) {
+  cat(sprintf("\n## %s\n\n", itable))
+  print(list_of_tables[[itable]])
+}
+
 ```
@@
 
-sum1 |> ggplot(aes(x=pct_change)) +
+plot_hist_condit <- function(fcondit,dat,pvar) {
+  full_range <- dat |> pull({{pvar}}) |> range(na.rm=TRUE)
+  dat |>
+    filter(refClass == fcondit) |> 
+    ggplot(aes(.data[[pvar]])) +
+    geom_bar(fill = 'dodgerblue4',width=1) +
+    theme_minimal() #+ coord_cartesian(xlim = full_range)
+}
+
+s1_agg4 |> group_by(refClass) |> 
+  summarise(Mean=mean(pct), Median=median(pct),Mode=modeest::mlv(pct,method='mfv'), sd=sd(pct), n=n()) |> 
+  mutate(Distribution=refClass) |> rename("Reference Class"=refClass) |>  
+  gt() |> 
+  tab_header(title="Descriptive Statistics of % of trials with matched goal") |> 
+  text_transform(
+    locations=cells_body(columns='Distribution'),
+    fn=function(column) {
+      map(column,plot_hist_condit,dat=s1_agg4,pvar='mg') |> 
+        ggplot_image(height=px(50),aspect_ratio=2)
+    }
+  )
+
+```
+
+```{r}
+#| eval: FALSE
+
+m_zib <- brm(
+  bf(pct ~ refClass + (1|id),
+     zi ~ refClass + (1|id)),
+  family = zero_inflated_beta(),
+  data = s1_agg4
+)
+
+m_bb <- brm(
+  pct ~ refClass + (1|id),
+  family = beta_binomial(),
+  data = s1_agg4
+)
+
+m_bb <- brm(
+    successes | trials(4) ~ refClass,
+    family = beta_binomial(),
+    data = s1_agg4 |> 
+        mutate(successes = pct * 4)  # convert proportion back to count
+)
+# Regression Coefficients:
+#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# Intercept             -0.52      0.01    -0.55    -0.49 1.00     4003     3155
+# refClassPercentage    -0.73      0.02    -0.77    -0.69 1.00     3603     2648
+# refClassUSD           -1.74      0.02    -1.79    -1.69 1.00     3478     3007
+# 
+# Further Distributional Parameters:
+#     Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# phi     0.33      0.00     0.33     0.34 1.00     3255     2429
+
+# > m_bb |> emmeans(~refClass, type="response")
+#  refClass   prob lower.HPD
+#  upper.HPD
+#  kWh        0.37      0.37      0.38
+#  Percentage 0.22      0.22      0.23
+#  USD        0.09      0.09      0.10
+
+m_bb |> emmeans(~refClass)
+m_bb |> emmeans(~refClass, type="response")
+
+m_bb_re <- brm(
+    successes | trials(4) ~ refClass + (1|id),
+    family = beta_binomial(),
+    cores=4,
+    data = s1_agg4 |> 
+        mutate(successes = pct * 4)
+)
+
+m_bb2 <- brm(
+    mg | trials(4) ~ 0+refClass,
+    family = beta_binomial(link = "identity", link_phi = "identity"),
+    chains=2,
+    cores=2,
+    data = s1_agg4
+)
+
+m_bb3 <- brm(
+    mg | trials(2) ~ 0+refClass,
+    family = beta_binomial(link = "identity", link_phi = "identity"),
+    chains=2,
+    cores=2,
+    data = s1_agg2
+)
+# Regression Coefficients:
+#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# refClasskWh            0.38      0.00     0.38     0.38 1.00     1900     1385
+# refClassPercentage     0.22      0.00     0.22     0.22 1.00     1868     1549
+# refClassUSD            0.09      0.00     0.09     0.10 1.00     2608     1328
+# 
+# Further Distributional Parameters:
+#     Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# phi     0.17      0.00     0.17     0.18 1.00      813      940
+
+m_bb4 <- brm(
+    mg | trials(2) ~ refClass + (1|id),
+    family = beta_binomial(),
+    chains=2,
+    cores=2,
+    data = s1_agg2
+)
+
+m_t1 <- brm(matched_goal ~ refClass + (1|id), family = bernoulli(), data = s1_agg)
+ m_t1 |> emmeans(~refClass, type="response")
+
+# 
+# m_zib <- brm(
+#   bf(pct ~ refClass + (1|id),
+#      zi ~ refClass + (1|id)),
+#   family = zero_inflated_beta(),
+#   iter=2000,
+#   cores=4,
+#   data = s1_agg4 |> mutate( pct = ifelse(pct == 1, 0.999, ifelse(pct == 0, 0.001, pct)))
+# )
+# Multilevel Hyperparameters:
+# ~id (Number of levels: 242) 
+#                  Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# sd(Intercept)        0.09      0.07     0.00     0.26 1.00     2164     1682
+# sd(zi_Intercept)     0.77      0.63     0.03     2.32 1.00     2846     1826
+# 
+# Regression Coefficients:
+#                       Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# Intercept                -0.37      0.16    -0.67    -0.06 1.00     5376     2635
+# zi_Intercept             -8.17      3.46   -17.15    -3.80 1.00     2673     1375
+# refClassPercentage       -0.43      0.22    -0.87     0.00 1.00     5054     2984
+# refClassUSD              -0.69      0.21    -1.10    -0.28 1.00     4759     2772
+# zi_refClassPercentage    -0.04      4.78   -10.32    10.04 1.00     2037     1627
+# zi_refClassUSD            0.47      4.50    -8.53    10.19 1.00     2181     1665
+
+ s1_agg |> group_by(id,refClass) |> 
+  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
+   group_by(refClass) |> 
+  summarise(mean_pct=mean(pct), sd_pct=sd(pct), n=n())
+
+ post <- posterior_samples(m_t1)
+nsims <- 1000
+re_sims <- rnorm(nsims, 0, mean(post$sd_id__Intercept))
+```
+
+```{r}
+#| eval: false
+s1_agg |> ggplot(aes(x=pct_change)) +
   geom_density(aes(fill=state), alpha=0.5) +
@@
 # descriptive table to do with plot
-sum1 |> 
+s1_agg |> 
   group_by(refClass,state) |> 
@@
 
-sum1 <- sum1 |> filter(!id %in% outlier_id)
+s1_agg <- s1_agg |> filter(!id %in% outlier_id)
 
@@
   # add geom_vline with x intercept based on pct_goal of the data (which is currently a string of form 10%)
-sum1 |> 
+s1_agg |> 
   filter(block==1) |>
@@
 # Step 2: Summarize the percentage of participants meeting the goal for each condition
-summary_pct <- sum1 |>
+summary_pct <- s1_agg |>
   group_by(refClass) |>
@@
   theme_minimal()
+
+s1_agg |> group_by(id,refClass) |> 
+  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
+  ggplot(aes(x=pct)) +
+  geom_density(aes(fill=refClass), alpha=0.5)
+
+# histogram
+s1_agg |> group_by(id,refClass) |> 
+  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
+  ggplot(aes(x=pct)) +
+  geom_bar(aes(fill=refClass)) +
+  ggdist::stat_halfeye() +
+  geom_rug(alpha=0.1) +
+  stat_boxplot(width=0.5, alpha=0.6) +
+  facet_wrap(~refClass) +
+  labs(title="Distributions of % of trials that matched goal",
+       x="Percent of trials that matched goal",
+       y="Number of Participants",
+       fill="Reference Class") +
+  scale_fill_brewer(palette="Set2") + 
+  scale_y_continuous(breaks=seq(0,80,5)) + 
+  scale_x_continuous(breaks=seq(0,1,.25),labels=scales::percent_format(accuracy=1))
+
+# s1_agg |> group_by(id,refClass) |> 
+#   summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
+#   group_by(refClass) |>
+#   summarise(mean_pct=mean(pct), sd_pct=sd(pct), n=n())
+
+s1_agg |> group_by(id,refClass) |> 
+  summarise(mg=sum(matched_goal), n=n(), pct=mg/n) |> 
+  group_by(refClass,mg) |>
+  summarise(n=n()) |>
+  setNames(c("refClass", "# of trials with matched goal", "n")) |>
+  knitr::kable(booktabs=T,col.names = c("Reference Class", 
+                            "# of correct trials ",
+                            "n"),
+               format = "html") |> 
+  kableExtra::kable_styling(full_width = F) |> column_spec(1:3, width = "2cm")
 
@@
 ```
+
+```{r}
+#| eval: FALSE
+
+contingency_table <- s1_agg %>%
+    group_by(refClass) %>%
+    summarise(
+      matched = sum(matched_goal),
+      unmatched = n() - sum(matched_goal)
+    ) %>%
+    select(matched, unmatched) %>%
+    as.matrix()
+
+# Perform the chi-squared test
+chisq_test <- chisq.test(contingency_table)
+
+# View the results
+chisq_test
+
+s1_aggb <- s1_agg %>%
+  mutate(matched_goal = as.numeric(matched_goal))
+
+m1_s1 <- brms::brm(
+  matched_goal ~ refClass,        # Logistic regression with reference class as predictor
+  data = s1_agg,                    # The dataset
+  family = bernoulli(),           # Binary outcome model
+ # prior = set_prior("normal(0, 1)", class = "b"),  # Prior for coefficients
+  iter = 2000,                    # Number of iterations
+  chains = 4,                     # Number of MCMC chains
+  cores = 4,                      # Number of cores for parallel computation
+  seed = 123                      # Set seed for reproducibility
+)
+# Regression Coefficients: - prior
+#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# Intercept             -0.53      0.12    -0.76    -0.30 1.00     3182     2941
+# refClassPercentage    -0.76      0.18    -1.12    -0.41 1.00     2565     2685
+# refClassUSD           -1.64      0.21    -2.06    -1.24 1.00     2576     2941
+# Regression Coefficients: - no prior
+#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# Intercept             -0.49      0.12    -0.72    -0.27 1.00     3259     2975
+# refClassPercentage    -0.81      0.19    -1.18    -0.45 1.00     2364     2585
+# refClassUSD           -1.73      0.21    -2.15    -1.32 1.00     2643     2942
+
+m2_s1 <- brms::brm(
+  matched_goal ~ refClass + (1|id),       
+  data = s1_agg,                    
+  family = bernoulli(),           
+  #prior = set_prior("normal(0, 1)", class = "b"),  
+  iter = 2000,                    
+  chains = 4,                     
+  cores = 4,                      
+  seed = 123                      
+)
+# Regression Coefficients:
+#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# Intercept             -3.97      0.95    -6.07    -2.38 1.00     1001     1715
+# refClassPercentage    -0.54      0.75    -1.97     0.96 1.00     1523     2223
+# refClassUSD           -2.26      0.76    -3.74    -0.77 1.00     1729     2581
+# Regression Coefficients: - no prior
+#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# Intercept             -2.04      0.97    -4.11    -0.35 1.01      616     1384
+# refClassPercentage    -3.16      1.35    -5.97    -0.65 1.00      542     1234
+# refClassUSD           -6.84      1.68   -10.42    -3.87 1.01      581     1096
+
+m3_s1 <- brms::brm(
+  matched_goal ~ refClass + (1|id) + (1|state), 
+  data = s1_agg,                    
+  family = bernoulli(),           
+  prior = set_prior("normal(0, 1)", class = "b"),  
+  iter = 2000,                    #
+  chains = 4,                     
+  cores = 4,                      
+  seed = 123                      
+)
+# Regression Coefficients:
+#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# Intercept             -4.08      1.05    -6.39    -2.20 1.00      651     1280
+# refClassPercentage    -0.52      0.78    -2.05     1.02 1.01      769     1583
+# refClassUSD           -2.25      0.78    -3.78    -0.72 1.00     1036     2002
+
+m4_s1 <- brms::brm(
+  matched_goal ~ refClass + (1+plan|id) + (1|state), 
+  data = s1_agg,                    
+  family = bernoulli(),           
+  iter = 4000,                    #
+  chains = 4,                     
+  cores = 4,                      
+  seed = 123                      
+)
+# Regression Coefficients:
+#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# Intercept             -4.24      1.12    -6.60    -2.18 1.00     1280      962
+# refClassPercentage    -0.49      0.79    -1.99     1.12 1.00     2347     4012
+# refClassUSD           -2.22      0.78    -3.74    -0.71 1.00     2395     4132
+
+m5_s1 <- brms::brm(
+  matched_goal ~ refClass + (1+plan|id) + (1|state), 
+  data = s1_agg,                    
+  family = bernoulli(),           
+  iter = 4000,                    #
+  chains = 4,                     
+  cores = 4,                      
+  seed = 123                      
+)
+# Regression Coefficients:
+#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# Intercept             -2.20      1.15    -4.56     0.02 1.00     1819     2875
+# refClassPercentage    -3.34      1.53    -6.59    -0.64 1.00     1139     2377
+# refClassUSD           -7.37      1.90   -11.64    -4.13 1.00     1706     3083
+
+```
+```{r}
+#| eval: FALSE
+
+stanplot(m5_s1)
+
+mcmc_plot(m5_s1, pars = c("b_Intercept","refClassPercentage", "refClassUSD"), regex = FALSE)
+mcmc_plot(m5_s1, regex_pars=c('b_*'), regex = FALSE)
+
+m5_s1 |> emmeans(~refClass)
+#  refClass   emmean lower.HPD upper.HPD
+#  kWh          -2.2      -4.6       0.0
+#  Percentage   -5.4      -8.5      -2.9
+#  USD          -9.4     -13.4      -6.3
+# 
+# Point estimate displayed: median 
+# Results are given on the logit (not the response) scale. 
+# HPD interval probability: 0.95
+
+m5_s1 |> emmeans(~refClass) |> gather_emmeans_draws() |> 
+  ggplot(aes(x=refClass,y=.value,fill=refClass)) +
+  stat_dist_pointinterval() + stat_halfeye(alpha=.2) +
+  stat_lineribbon(alpha = .25, size = 1, .width = c(.95))
+
+# transform the log-odds to odds
+
+```

### 2024-10-23 study2.qmd
updates

--- 2024-10-09 study2.qmd
@@
 ```{r}
+
 library(pacman)
@@
 s2 <- s2 |> mutate(id = as.numeric(format(EndDate, "%H%M%S"))) |> relocate(id) 
+s2 <- s2 |> filter(!(id==203309)) # NA for some responses
 s2 <- s2 |> 
@@
 
-# recode DEMO4 variable based on this mapping: 
-# 1 = "Some schooling, but no diploma or degree" 
-# 2 = "Highschool diploma or GED" 
-# 3 = "Some college" 
-# 4 = "College degree" 
-# 5 = "Some graduate school" 
-# 6 = "Graduate degree" 
-# 7 = "Choose not to answer"
-
 s2 <- s2 |> rename(edu=DEM04) |>
@@
   ))) |> relocate(edu, .after = id) 
+```
+
+```{r}
+#| eval: false
 
@@
   pct_goal=case_when (TENPERCENT == 1 ~ "15%", TENPERCENT == 2 ~ "10%"), 
-  pct=case_when (TENPERCENT == 1 ~ .10, TENPERCENT == 2 ~ .15),
+  pct=case_when (TENPERCENT == 1 ~ .15, TENPERCENT == 2 ~ .10),
   rounded = case_when(ROUNDED == 1 ~ "Not Rounded", ROUNDED == 2 ~ "Rounded")
@@
   select(-TENPERCENT,-ROUNDED, -state2)
-
-# pivot data into longer format, so that each appliance type is on its own row, so for each appliance we have the family value, state, value, plan 1 value, and plan 2 value
-# so for the first row for id 113029, we would have ACFAMILY1 ACSTATE1 ACAPONE1 ACAPTWO1, then the next row would be HEATFAMILY1 HEATSTATE1 HEATAPONE1 HEATAPTWO1, etc.
-# so each of the bill sources, AC, HEAT, WATER, REFRIG, OTHER, will now be on its own row, with the family, state, value, plan 1 value, and plan 2 value
 
@@
 #write.csv(s2_long,"data/study_2_tg.csv")
+saveRDS(s2_long, here::here("data/s2_processed.rds"))
+
+```
+
+```{r}
+
+s2_long <- readRDS(here::here("data/s2_processed.rds"))
 
 # compute total kWh reduction, and total % reduction 
-sum1 <- s2_long |> 
+s2_agg <- s2_long |> 
   filter(appliance !="TOTAL") |> 
@@
 
+# Step 1: Calculate whether participants matched the target pct_goal
+s2_agg <- s2_long |> 
+  filter(appliance != "TOTAL") |> 
+  group_by(id, refClass, state, pct_goal, pct, rounded, block, plan, calc, edu) |> 
+  summarise(
+    total_kWh = sum(value),
+    orig_kWh = sum(family),
+    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
+    state_dif = mean(state_dif),
+    .groups = "drop"
+  ) |>
+  mutate(matched_goal = (pct_change == pct),
+  matched_goal2 = abs(pct_change-pct)<.03)
+
+```
+
+```{r}
+s2_agg |> group_by(id,refClass) |> 
+  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
+  group_by(refClass) |>
+  summarise(mean_pct=mean(pct), sd_pct=sd(pct), n=n())
+
+s2_agg |> group_by(id,refClass) |> 
+  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
+  group_by(refClass,mg) |>
+  summarise(n=n())
+
+s2_agg |> group_by(id,refClass) |> 
+  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
+  ggplot(aes(x=mg)) +
+  geom_histogram(aes(fill=refClass), alpha=0.5, bins=20) +
+  ggdist::stat_halfeye() +
+  geom_rug(alpha=0.1) +
+  stat_boxplot(width=0.5, alpha=0.6) +
+  facet_wrap(~refClass) +
+  labs(title="Distribution of Percent Change by State and Reference Class - Before outlier exclusion",
+       x="Percent Change",
+       y="Density") +
+  scale_fill_brewer(palette="Set2") + 
+  scale_y_continuous(breaks=seq(0,80,5))
+
+s2_agg |> group_by(id,refClass,pct_goal) |> 
+  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
+  ggplot(aes(x=mg)) +
+  geom_histogram(aes(fill=refClass), alpha=0.5, bins=20) +
+  ggdist::stat_halfeye() +
+  geom_rug(alpha=0.1) +
+  stat_boxplot(width=0.5, alpha=0.6) +
+  facet_wrap(pct_goal~refClass) +
+  labs(title="Distribution of Percent Change by State and Reference Class - Before outlier exclusion",
+       x="Percent Change",
+       y="Density") +
+  scale_fill_brewer(palette="Set2") + 
+  scale_y_continuous(breaks=seq(0,80,5))
 ```
@@
 
-outliers <- sum1 |>  ungroup() |> group_by(state) |> mutate(change_mean=mean(pct_change,na.rm=TRUE),change_sd=sd(pct_change,na.rm=TRUE),
+outliers <- s2_agg |>  ungroup() |> group_by(state) |> mutate(change_mean=mean(pct_change,na.rm=TRUE),change_sd=sd(pct_change,na.rm=TRUE),
                                      z_score=(pct_change-change_mean)/change_sd,
@@
 # Calculate skewness and kurtosis for the `pct_change` variable
-skewness_value <- skewness(sum1$pct_change, na.rm = TRUE)
-kurtosis_value <- kurtosis(sum1$pct_change, na.rm = TRUE)
+skewness_value <- skewness(s2_agg$pct_change, na.rm = TRUE)
+kurtosis_value <- kurtosis(s2_agg$pct_change, na.rm = TRUE)
 
@@
 # Shapiro-Wilk test for normality (sample size limitation)
-shapiro_test <- shapiro.test(sum1$pct_change[1:5000])  # Using first 5000 observations if large dataset
+shapiro_test <- shapiro.test(s2_agg$pct_change[1:5000])  # Using first 5000 observations if large dataset
 print(shapiro_test)
@@
 
-sum1 |> ggplot(aes(x=pct_change)) +
+s2_agg |> ggplot(aes(x=pct_change)) +
   geom_density(aes(fill=state), alpha=0.5) +
@@
 
-sum1 <- sum1 |> 
+s2_agg <- s2_agg |> 
   filter(!id %in% outlier_id)
 
-sum1 |> 
+s2_agg |> 
   filter(!id %in% outlier_id) |>
@@
   # add geom_vline with x intercept based on pct_goal of the data (which is currently a string of form 10%)
-sum1 |> 
+s2_agg |> 
   filter(block==1) |>
@@
 # s2_long |> ggplot(aes(x = change)) + geom_histogram(bins = 30) + facet_wrap(~appliance, scales = "free")
-sum1 |> mutate(refClass=factor(refClass),pct_goal=factor(pct_goal),rounded=factor(rounded),state=factor(state)) |>
+s2_agg |> mutate(refClass=factor(refClass),pct_goal=factor(pct_goal),rounded=factor(rounded),state=factor(state)) |>
 DT::datatable(extensions = 'Buttons',
@@
 
-# Step 1: Calculate whether participants matched the target pct_goal
-sum1 <- s2_long |> 
-  filter(appliance != "TOTAL") |> 
-  group_by(id, refClass, state, pct_goal, pct, rounded, block, plan, calc, edu) |> 
-  summarise(
-    total_kWh = sum(value),
-    orig_kWh = sum(family),
-    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
-    state_dif = mean(state_dif),
-    .groups = "drop"
-  ) |>
-  mutate(matched_goal = (pct_change == pct))
+# 203309
 
 # Step 2: Summarize the percentage of participants meeting the goal for each condition
-summary_pct <- sum1 |>
+summary_pct <- s2_agg |>
   group_by(refClass, rounded, pct_goal) |>
@@
 
-sum1 |> ggplot(aes(x = refClass, y = pct_change, fill = refClass)) + 
+s2_agg |> ggplot(aes(x = refClass, y = pct_change, fill = refClass)) + 
   stat_summary(fun = mean, geom = "bar", position="dodge") +
@@
 
-sum1 |> ggplot(aes(x = pct_goal, y = pct_change, fill = pct_goal)) + 
+s2_agg |> ggplot(aes(x = pct_goal, y = pct_change, fill = pct_goal)) + 
   stat_summary(fun = mean, geom = "bar", position="dodge") +
@@
 
-sum1 |> ggplot(aes(x = rounded, y = pct_change, fill = rounded)) + 
+s2_agg |> ggplot(aes(x = rounded, y = pct_change, fill = rounded)) + 
   stat_summary(fun = mean, geom = "bar", position="dodge") +
@@
 
-sum1 |> ggplot(aes(x = pct_goal, y = pct_change, fill = refClass)) + 
+s2_agg |> ggplot(aes(x = pct_goal, y = pct_change, fill = refClass)) + 
   stat_summary(fun = mean, geom = "bar", position="dodge") +
@@
 
-sum1 |> ggplot(aes(x = pct_goal, y = pct_change, fill = rounded)) + 
+s2_agg |> ggplot(aes(x = pct_goal, y = pct_change, fill = rounded)) + 
   stat_summary(fun = mean, geom = "bar", position="dodge") +
@@
 
-sum1 |> ggplot(aes(x = calc, y = pct_change, fill = calc)) + 
+s2_agg |> ggplot(aes(x = calc, y = pct_change, fill = calc)) + 
   stat_summary(fun = mean, geom = "bar", position="dodge") +
@@
 
-sum1 |> ggplot(aes(x = edu, y = pct_change, fill = edu)) + 
+s2_agg |> ggplot(aes(x = edu, y = pct_change, fill = edu)) + 
   stat_summary(fun = mean, geom = "bar", position="dodge") +
@@
 ```{r}
+#| eval: FALSE
+
+contingency_table <- sum1 %>%
+    group_by(refClass) %>%
+    summarise(
+      matched = sum(matched_goal),
+      unmatched = n() - sum(matched_goal)
+    ) %>%
+    select(matched, unmatched) %>%
+    as.matrix()
+
+# Perform the chi-squared test
+chisq_test <- chisq.test(contingency_table)
+
+# View the results
+chisq_test
+
+sum1b <- sum1 %>%
+  mutate(matched_goal = as.numeric(matched_goal))
+
+library(brms)
+bayesian_model <- brms::brm(
+  matched_goal ~ refClass,        # Logistic regression with reference class as predictor
+  data = sum1,                    # The dataset
+  family = bernoulli(),           # Binary outcome model
+  prior = set_prior("normal(0, 1)", class = "b"),  # Prior for coefficients
+  iter = 2000,                    # Number of iterations
+  chains = 4,                     # Number of MCMC chains
+  cores = 4,                      # Number of cores for parallel computation
+  seed = 123                      # Set seed for reproducibility
+)
+# Regression Coefficients:
+#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# Intercept             -1.27      0.14    -1.55    -1.00 1.00     3696     3393
+# refClassPercentage    -0.56      0.22    -0.98    -0.14 1.00     3308     3065
+# refClassUSD           -0.95      0.25    -1.45    -0.46 1.00     2895     2701
+
+bm2 <- brms::brm(
+  matched_goal ~ refClass + (1|id),        # Logistic regression with reference class as predictor
+  data = sum1,                    # The dataset
+  family = bernoulli(),           # Binary outcome model
+  prior = set_prior("normal(0, 1)", class = "b"),  # Prior for coefficients
+  iter = 2000,                    # Number of iterations
+  chains = 4,                     # Number of MCMC chains
+  cores = 4,                      # Number of cores for parallel computation
+  seed = 123                      # Set seed for reproducibility
+)
+# Regression Coefficients:
+#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# Intercept             -1.92      0.31    -2.56    -1.37 1.00     2039     2669
+# refClassPercentage    -0.73      0.36    -1.43    -0.02 1.00     2586     3142
+# refClassUSD           -1.15      0.41    -1.95    -0.37 1.00     3079     3071
+
+bm3 <- brms::brm(
+  matched_goal ~ refClass*rounded + (1|id),        # Logistic regression with reference class as predictor
+  data = sum1,                    # The dataset
+  family = bernoulli(),           # Binary outcome model
+  prior = set_prior("normal(0, 1)", class = "b"),  # Prior for coefficients
+  iter = 2000,                    # Number of iterations
+  chains = 4,                     # Number of MCMC chains
+  cores = 4,                      # Number of cores for parallel computation
+  seed = 123                      # Set seed for reproducibility
+)
+# Regression Coefficients:
+#                                   Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# Intercept                            -1.80      0.34    -2.49    -1.17 1.00     1896     2384
+# refClassPercentage                   -1.48      0.45    -2.37    -0.61 1.00     3108     3324
+# refClassUSD                          -1.54      0.47    -2.53    -0.66 1.00     3309     3226
+# roundedRounded                       -0.59      0.30    -1.17     0.02 1.00     4310     2857
+# refClassPercentage:roundedRounded     1.71      0.45     0.83     2.59 1.00     4076     3070
+# refClassUSD:roundedRounded            1.00      0.50     0.02     1.96 1.00     4399     3144
+
+bm4 <- brms::brm(
+  matched_goal ~ refClass*rounded + (1|id) +(1|state),        # Logistic regression with reference class as predictor
+  data = sum1,                    # The dataset
+  family = bernoulli(),           # Binary outcome model
+  prior = set_prior("normal(0, 1)", class = "b"),  # Prior for coefficients
+  iter = 2000,                    # Number of iterations
+  chains = 4,                     # Number of MCMC chains
+  cores = 4,                      # Number of cores for parallel computation
+  seed = 123                      # Set seed for reproducibility
+)
+# Regression Coefficients:
+#                                   Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# Intercept                            -1.77      0.44    -2.62    -0.81 1.01      526      187
+# refClassPercentage                   -1.49      0.44    -2.34    -0.62 1.00     1352     1519
+# refClassUSD                          -1.53      0.48    -2.49    -0.62 1.00     1836     2504
+# roundedRounded                       -0.58      0.29    -1.16    -0.03 1.00     2933     2938
+# refClassPercentage:roundedRounded     1.72      0.45     0.85     2.59 1.00     1936     2501
+# refClassUSD:roundedRounded            1.00      0.48     0.05     1.91 1.00     2460     2472
+
+```
+
+```{r}
+#| eval: false
 
@@
 
-m1 <- lmer(pct_change ~ refClass  + (1+plan|id) + (1|state), data = sum1)
+m1 <- lmer(pct_change ~ refClass  + (1+plan|id) + (1|state), data = s2_agg)
 summary(m1)
@@
 
-m1 <- lmer(pct_change ~ pct_goal  + (1|id) + (1|state), data = sum1)
+m1 <- lmer(pct_change ~ pct_goal  + (1|id) + (1|state), data = s2_agg)
 summary(m1)
@@
 
-m1 <- lmer(pct_change ~ pct_goal*rounded  + (1|id) + (1|state), data = sum1)
+m1 <- lmer(pct_change ~ pct_goal*rounded  + (1|id) + (1|state), data = s2_agg)
 summary(m1)
@@
 
-m1 <- lmer(pct_change ~ refClass*pct_goal*rounded  + (1|id) + (1|state), data = sum1)
+m1 <- lmer(pct_change ~ refClass*pct_goal*rounded  + (1|id) + (1|state), data = s2_agg)
 ggpredict(m1, c("refClass","pct_goal","rounded")) |> plot()
 
-m1 <- lmer(pct_change ~ state*pct_goal*rounded  + (1|id), data = sum1)
+m1 <- lmer(pct_change ~ state*pct_goal*rounded  + (1|id), data = s2_agg)
 summary(m1)

### 2024-10-23 Scripts/fit_models.R

@@ -1,0 +1,85 @@
+
+library(pacman)
+pacman::p_load(dplyr,purrr,tidyr,here,tibble,brms,bayestestR, ggplot2,gt,ggh4x,lme4,flextable)
+
+options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
+options(brms.backend="cmdstanr",mc.cores=4)
+
+
+
+
+
+# prior = c(prior(normal(30,200),lb=0,class= Intercept))
+# modelName <- "combo_testVxBand_RF_5K.rds"
+# combo_vx <- brm(vx ~ condit * fb * bandOrder * bandInt + (1 + bandInt|id),
+#                 data=testAll,file=paste0(here::here("data/model_cache",modelName)),
+#                 iter=5000,chains=4,silent=0, prior=prior, 
+#                 control=list(adapt_delta=0.94, max_treedepth=11))
+# bayestestR::describe_posterior(combo_vx)
+
+#  posterior_summary(bmtd,variable=r_bandInt_params)
+
+
+# get_coef_details <- function(model, term_name) {
+#   term_details <- broom.mixed::tidy(model) %>%
+#     filter(term == term_name) %>%
+#     select(estimate, conf.low, conf.high) %>%
+#     mutate(across(where(is.numeric), ~round(.x, 2)))
+#   
+#   pd_results <- bayestestR::p_direction(model, effects = "fixed", parameters = term_name)
+#   # Directly extract the pd value for the specified term
+#   pd_value <- ifelse(!is.null(pd_results$pd), pd_results$pd[1], NA_real_)
+#   
+#   term_details$pd <- paste0(round(pd_value * 100, 2),"%")  # Ensure it's a percentage and rounded
+#   return(term_details)
+# }
+
+# bayestestR::p_direction(e1_vxBMM)[, "conditVaried"]
+
+
+# condEffects <- function(m){
+#   m |> ggplot(aes(x = bandInt, y = .value, color = condit, fill = condit)) + 
+#     stat_dist_pointinterval() + stat_halfeye(alpha=.2) +
+#     stat_lineribbon(alpha = .25, size = 1, .width = c(.95)) +
+#     theme(axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) +
+#     ylab("Predicted X Velocity") + xlab("Band")
+# }
+
+
+
+# GetModelStats <- function(model, type="brms") {
[…]

### 2024-10-24 results.qmd
update with results page

---
title: "Results"
author:
- name: Thomas E. Gorman
  url: https://tegorman13.github.io/
  affiliations: 
  - name:  Communication and Cognition Lab, Purdue University, USA
    affiliation-url: https://web.ics.purdue.edu/~treimer/
lightbox: true
toc: false
code-fold: true
code-tools: true
execute:
  #echo: false
  warning: false
format:
  html: 
    grid:
      sidebar-width: 220px
      body-width: 1200px
      margin-width: 170px
      gutter-width: 1.0rem
  hugo-md:
    include: true
    html-math-method: mathjax
    output-file: results_hugo.md
  gfm:
    echo: true
    output-file: results_gfm.md
---

```{r}

library(pacman)
pacman::p_load(dplyr,purrr,tidyr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,
               ggplot2,gt,knitr,kableExtra,ggh4x,lme4,flextable)

options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
options(brms.backend="cmdstanr",mc.cores=4)

walk(c("fun_plot"), ~ source(here::here(paste0("Scripts/", .x, ".R"))))
theme_set(theme_nice())

s1 <- readRDS(here::here("data/s1_processed.rds"))

s1_agg <- s1 |> 
  filter(appliance !="Total kWh") |> 
  group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
  summarise(total_kWh = sum(value),orig_kWh=sum(family), 
            pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), state_dif=mean(state_dif)) |> 
  mutate(matched_goal = (pct_change == pct_goal))

s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n)

s2_long <- readRDS(here::here("data/s2_processed.rds"))

s2_agg <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id, refClass, state, pct_goal, pct, rounded, block, plan, calc, edu) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(matched_goal = (pct_change == pct),
         matched_goal2 = abs(pct_change-pct)<.03)

s2_agg4 <- s2_agg |> group_by(id,refClass,calc) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n)

s2_agg2 <- s2_agg |> group_by(id,refClass,rounded,pct_goal,pct,calc) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n)

```

## Experiment 1

```{r}
##| label: tbl-s1-bb
##| tbl-cap: "E1. beta-binomial regression results

s1_bb1 <- brm(
  mg | trials(4) ~ refClass,
  family = beta_binomial(),
  cores=4,
  iter=5000,
  data = s1_agg4,
  file=paste0(here::here("data/model_cache",'s1_bb1.rds'))
)

mted1 <- as.data.frame(describe_posterior(s1_bb1, centrality = "Mean"))[, c(1,2,4,5,6)]
colnames(mted1) <- c("Term", "Estimate"," 95% CrI Lower", " 95% CrI Upper", "pd")

mted1 |>
  mutate(across(c("Estimate", " 95% CrI Lower", " 95% CrI Upper"), 
                \(x) plogis(x))) |>
  mutate(across(where(is.numeric), \(x) round(x, 3))) |>
  tibble::remove_rownames() |>
  mutate(Term = stringr::str_remove(Term, "b_")) |>
  kable(booktabs = TRUE)

```

```{r}
#| label: fig-s1-mg1
#| fig-cap: Study 1
#| fig-width: 11
#| fig-height: 8

plot(conditional_effects(s1_bb1, effects="refClass"),points=FALSE)

posterior_predict(s1_bb1,ndraws=200) |> array_branch(margin=1) |> 
  map_dfr(
    function(yrep_iter){
      s1_agg4 %>% ungroup() |>  mutate(pred=yrep_iter)
    },
    .id='iter'
  ) |> mutate(iter=as.numeric(iter)) |> 
  ggplot(aes(pred, group=iter)) +
  # Scale density to match the count scale
  geom_line(stat='density', aes(y=..count..), 
            alpha=0.7, color='blue') +
  geom_bar(data=s1_agg4, aes(mg, fill=refClass),
           inherit.aes=FALSE) +
  facet_wrap(~refClass)
```

```{r}
##| label: tbl-s2-bb
##| tbl-cap: "Study 2. beta-binomial regression results
##| 
##|

s2_bb2_r_g <- brm(
  mg | trials(2) ~ refClass*rounded*pct_goal,
  family = beta_binomial(),
  data = s2_agg2,
  cores=4,
  iter=5000,
  file=paste0(here::here("data/model_cache",'s2_bb2_r_g.rds'))
)

mted2 <- as.data.frame(describe_posterior(s2_bb2_r_g, centrality = "Mean"))[, c(1,2,4,5,6)]
colnames(mted2) <- c("Term", "Estimate"," 95% CrI Lower", " 95% CrI Upper", "pd")

mted2 |>
  mutate(across(c("Estimate", " 95% CrI Lower", " 95% CrI Upper"), 
                \(x) plogis(x))) |>
  mutate(across(where(is.numeric), \(x) round(x, 3))) |>
  tibble::remove_rownames() |>
  mutate(Term = stringr::str_remove(Term, "b_")) |>
  kable(booktabs = TRUE)
```

```{r}
#| label: fig-s2-mg1
#| fig-cap: Study 2
#| fig-width: 11
#| fig-height: 9

# plot(conditional_effects(s2_bb2_r_g, 
#                          effects = "rounded:pct_goal", 
#                          conditions=make_conditions(s2_bb2_r_g,"refClass" )),
#      points=FALSE,plot=TRUE)

patchwork::wrap_plots(plot(conditional_effects(s2_bb2_r_g),points=FALSE,plot=FALSE))

```

```{r}
#| label: fig-s2-mg2
#| fig-cap: Study 2
#| fig-width: 11
#| fig-height: 10

condEffects <- function(m,xvar){
  m |> ggplot(aes(x = {{xvar}}, y = .value, color = refClass, fill = refClass)) + 
  stat_dist_pointinterval() + 
  stat_halfeye(alpha=.1, height=.5) +
  theme(legend.title=element_blank(),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) 
}

pe3td <- s2_agg2 |>  ggplot(aes(x =refClass, y = mg,fill=pct_goal)) +
    stat_summary(geom = "bar", position=position_dodge(), fun = mean) +
    stat_summary(geom = "errorbar", position=position_dodge(.9), fun.data = mean_se, width = .4, alpha = .7) + 
    facet_wrap(~rounded,ncol=1) +
  theme(legend.title=element_blank(),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5)) +
  labs(x="Band", y="Deviation From Target")

pe3ce <- s2_bb2_r_g |> emmeans( ~refClass *rounded*pct_goal, type="response") |>
  gather_emmeans_draws() |>
 condEffects(pct_goal) + labs(y="Absolute Deviation From Band", x="Band Type") + 
 facet_wrap(~refClass,ncol=1)

s2_bb2_r_g |> emmeans(pairwise ~ refClass * rounded, type="response") |> 
  pluck("contrasts") |>
  gather_emmeans_draws() |> 
  group_by(contrast,.draw) |> summarise(value=mean(.value), n=n()) |> 
   ggplot(aes(x=value,y=contrast,fill=contrast)) +
  stat_halfeye() + 
  geom_vline(xintercept=0, alpha=.5) +
  labs(x="Model Error Difference",y="Contrast") + theme(legend.position="none")

```

### 2024-10-24 study1.qmd
update with results page

@@
       gutter-width: 1.0rem
+  hugo-md:
+    include: true
+    html-math-method: mathjax
+    output-file: s1_hugo.md
+  gfm:
+    echo: true
+    output-file: s1_gfm.md
 ---
@@
   rename(calc=MATH01) |>  # replace NA's in calc with 0's
-  rename(id=ResponseId) |>
+  rename(id2=ResponseId) |>
+  mutate(id=row_number()) |> relocate(id) |> 
   mutate(calc=factor(ifelse(is.na(calc),0,calc))) |> 
@@
       values_to = "value"
-    ) |> select(id,state1,state2,refClass,calc,edu,modality,state,item_num,value)
+    ) |> select(id,state1,state2,refClass,calc,edu,modality,state,item_num,value,id2)
 
@@
 # for id R_1kFwgoKAmNTwCs3 - when state== TEX, appliance == Heating, check values of family column
-s1 |> filter(id=="R_1kFwgoKAmNTwCs3" & state=="TEX" & appliance=="Heating") |> select(family) 
+s1 |> filter(id2=="R_1kFwgoKAmNTwCs3" & state=="TEX" & appliance=="Heating") |> select(family) 
 # for id R_1kFwgoKAmNTwCs3 - when state== TEX, appliance == Heating, set values of family column to 6118
-s1 <- s1 |> mutate(family=ifelse(id=="R_1kFwgoKAmNTwCs3" & state=="TEX" & appliance=="Heating",6118,family))
+s1 <- s1 |> mutate(family=ifelse(id2=="R_1kFwgoKAmNTwCs3" & state=="TEX" & appliance=="Heating",6118,family))
+
+s1 <- s1 |> 
+  mutate(change = value-family, state_dif = value-state_avg,calc = if_else(calc == 0, "No Calculator", "Used Calculator")) |> 
+  mutate(value=as.numeric(value))
 
@@
 s1 <- readRDS(here::here("data/s1_processed.rds"))
-
-s1 <- s1 |> mutate(change = value-family, state_dif = value-state_avg,calc = if_else(calc == 0, "No Calculator", "Used Calculator"))
 

### 2024-10-24 study2.qmd
update with results page

@@
   warning: false
+cache: true
 format:
@@
       gutter-width: 1.0rem
+  hugo-md:
+    include: true
+    html-math-method: mathjax
+    output-file: s2_hugo.md
+  gfm:
+    echo: true
+    output-file: s2_gfm.md
 ---
@@
 
+#202009
+
 s2 <- haven::read_sav("data/Frequency & Probability Study 2 - 3-24-19.sav")
-s2 <- s2 |> mutate(id = as.numeric(format(EndDate, "%H%M%S"))) |> relocate(id) 
-s2 <- s2 |> filter(!(id==203309)) # NA for some responses
+s2 <- s2 |> #mutate(id = as.numeric(format(EndDate, "%H%M%S"))) |> relocate(id) |> 
+  mutate(id=paste0("s2_",row_number())) |> relocate(id)
+s2 <- s2 |> filter(!(id=="s2_78")) # NA for some responses
 s2 <- s2 |> 
@@
   mutate(calc=factor(ifelse(is.na(calc),0,calc))) |> 
+  mutate(calc = if_else(calc == 0, "No Calculator", "Calculator")) |> 
   relocate(refClass,state1,state2,calc, .after = id)
@@
   ggplot(aes(x=mg)) +
-  geom_histogram(aes(fill=refClass), alpha=0.5, bins=20) +
+  geom_bar(aes(fill=refClass)) +
   ggdist::stat_halfeye() +
@@
   scale_y_continuous(breaks=seq(0,80,5))
+
 ```
@@
                                      z_score=(pct_change-change_mean)/change_sd,
-                                     is_outlier=abs(z_score)>2.75)
+                                     is_outlier=abs(z_score)>3.0)
 

### 2024-10-24 Scripts/fit_models.R

@@ -1,9 +1,248 @@
 
 library(pacman)
-pacman::p_load(dplyr,purrr,tidyr,here,tibble,brms,bayestestR, ggplot2,gt,ggh4x,lme4,flextable)
+pacman::p_load(dplyr,purrr,tidyr,here,tibble,brms,bayestestR,emmeans,tidybayes,modelsummary,
+               ggplot2,gt,ggh4x,lme4,flextable)
 
 options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
 options(brms.backend="cmdstanr",mc.cores=4)
+
+
+
+
+
+
+
+s1 <- readRDS(here::here("data/s1_processed.rds"))
+
+s1_agg <- s1 |> 
+  filter(appliance !="Total kWh") |> 
+  group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
+  summarise(total_kWh = sum(value),orig_kWh=sum(family), 
+            pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), state_dif=mean(state_dif)) |> 
+  mutate(matched_goal = (pct_change == pct_goal))
+
+s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
+  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) 
+
+
+
+s1_bb1 <- brm(
+  mg | trials(4) ~ refClass,
+  family = beta_binomial(),
+  cores=4,
+  iter=5000,
+  data = s1_agg4,
+  file=paste0(here::here("data/model_cache",'s1_bb1.rds'))
+)
+# Regression Coefficients:
+#   Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# Intercept             -0.58      0.01    -0.61    -0.56 1.00     9576     7793
+# refClassPercentage    -0.70      0.02    -0.74    -0.66 1.00     8559     7667
+# refClassUSD           -1.74      0.02    -1.78    -1.69 1.00     8391     7522
+# 
+# Further Distributional Parameters:
+#   Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# phi     0.33      0.00     0.32     0.34 1.00     9328     7470
+
+
+s1_bb1_c <- brm(
+  mg | trials(4) ~ refClass*calc,
+  family = beta_binomial(),
+  data = s1_agg4,
[…]

### 2024-10-24 study1.qmd
improve figures

@@
   labs(title="Distributions of % of trials that matched goal",
-       x="Percent of trials that matched goal",
+       x="% of trials (out of 4) that matched goal",
        y="Number of Participants",
@@
   scale_fill_brewer(palette="Set2") + 
-  scale_y_continuous(breaks=seq(0,80,5)) + 
-  scale_x_continuous(breaks=seq(0,1,.25),labels=scales::percent_format(accuracy=1))
+  scale_y_continuous(breaks=seq(0,80,5)) + scale_x_continuous(breaks=seq(0,1,.25),labels=scales::percent_format(accuracy=1))
 
@@
   group_by(refClass,mg) |>
-  summarise(n=n()) |>
-  setNames(c("refClass", "# of trials with matched goal", "n")) |>
-  knitr::kable(booktabs=T,col.names = c("Reference Class", 
-                            "# of correct trials ",
-                            "n"),
-               format = "html") |> 
-  kableExtra::kable_styling(full_width = F) |> column_spec(1:3, width = "2cm")
+  summarise(n=n()) |>  rename("# matched goal"=mg,"# participants"=n) |> gt() |>  tab_header(
+    title = md("**Number of Trials participants match goal by Reference class**"),
+    subtitle = "Each participant had 4 trials where they attempt to create a plan to match goal"
+  )
+
+  # setNames(c("refClass", "# of trials with matched goal", "n")) |>
+  # knitr::kable(booktabs=T,col.names = c("Reference Class", 
+  #                           "# of correct trials ",
+  #                           "n"),
+  #              format = "html") |> 
+  # kableExtra::kable_styling(full_width = F) |> column_spec(1:3, width = "2cm")
 
@@
   tab_header(
-    title = md("**Percentage of Participants Matching Target Pct by Condition**"),
+    title = md("**Percentage of Trials Matching Target Pct by Condition**"),
     subtitle = "Grouped by Reference Class, Rounding, and Pct Goal"

### 2024-10-24 study2.qmd
improve figures

@@
   group_by(refClass) |>
-  summarise(mean_pct=mean(pct), sd_pct=sd(pct), n=n())
+  summarise(mean_pct=mean(pct), sd_pct=sd(pct), n=n()) |> gt()
 
@@
   group_by(refClass,mg) |>
-  summarise(n=n())
+  summarise(n=n()) |> 
+  rename("# matched goal"=mg,"# participants"=n) |> gt() |>  tab_header(
+    title = md("**Number of Trials participants match goal by Reference class**"),
+    subtitle = "Each participant had 4 trials where they attempt to create a plan to match goal"
+  )
 
@@
   summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
-  ggplot(aes(x=mg)) +
+  ggplot(aes(x=pct)) +
   geom_bar(aes(fill=refClass)) +
@@
   facet_wrap(~refClass) +
-  labs(title="Distribution of Percent Change by State and Reference Class - Before outlier exclusion",
-       x="Percent Change",
-       y="Density") +
+  labs(title="Distributions of % of trials that matched goal",
+       x="% of trials (out of 4) that matched goal",
+       y="Number of Participants") +
   scale_fill_brewer(palette="Set2") + 
-  scale_y_continuous(breaks=seq(0,80,5))
+  scale_y_continuous(breaks=seq(0,80,5))  + 
+  scale_x_continuous(breaks=seq(0,1,.25),labels=scales::percent_format(accuracy=1))
 
@@
   facet_wrap(pct_goal~refClass) +
-  labs(title="Distribution of Percent Change by State and Reference Class - Before outlier exclusion",
-       x="Percent Change",
-       y="Density") +
+  labs(title="Distributions of % of trials that matched goal",
+       x="% of trials (out of 2) that matched goal",
+       y="Number of Participants") +
   scale_fill_brewer(palette="Set2") + 
-  scale_y_continuous(breaks=seq(0,80,5))
+  scale_y_continuous(breaks=seq(0,80,5))  + 
+  scale_x_continuous(breaks=seq(0,1,.25),labels=scales::percent_format(accuracy=1))
 

### 2024-10-26 study1.qmd
add new design + stimuli page, and re-render all sites. s1 also has some new data checks for the state freq oddities.

@@
 s1 <- s1 |> 
-  mutate(change = value-family, state_dif = value-state_avg,calc = if_else(calc == 0, "No Calculator", "Used Calculator")) |> 
+  mutate(change = value-family, 
+         state_p_dif = value-state_avg,
+         state_f_dif= family-state_avg,
+         less_avg = value<state_avg,
+         calc = if_else(calc == 0, "No Calculator", "Used Calculator")) |> 
   mutate(value=as.numeric(value))
@@
   group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
-  summarise(total_kWh = sum(value),orig_kWh=sum(family), pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), state_dif=mean(state_dif)) |> 
+  summarise(total_kWh = sum(value),orig_kWh=sum(family), 
+            pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
+            n_change = sum(value!=family),
+            state_p_dif=mean(state_p_dif),
+            state_f_dif=mean(state_f_dif),
+            n_less_avg = sum(less_avg)) |> 
   mutate(matched_goal = (pct_change == pct_goal))
@@
 outlier_id <- outliers |> filter(is_outlier) |> pull(id) |> unique()
+
+# plot frequency dist of values of n_change
+s1_agg |> ggplot(aes(x=n_change)) +
+  geom_bar(fill = 'dodgerblue4',width=1) +
+  theme_minimal()
+
+# examine relationship between n_change and whether goal was matched
+s1_agg |> ggplot(aes(x=n_change,fill=matched_goal)) +
+  geom_bar(position="dodge") +
+  theme_minimal() +
+  labs(title="Relationship between # of changes and whether goal was matched",
+       x="# of changes",
+       y="Count") +
+  scale_fill_brewer(palette="Set2")
+
+# alternative way to examine the relationship, not a plot
+s1_agg |> group_by(n_change,matched_goal) |> summarise(n=n()) |> 
+  gt() |> tab_header(title="Relationship between # of changes and whether goal was matched")
+
+s1 |> 
+  filter(appliance !="Total kWh") |> 
+  group_by(state, appliance) |> 
+  summarise(family=mean(family),state_avg=mean(state_avg),dif=family-state_avg)
+
+s1 |> ungroup() |> mutate(diff=family-state_avg) |> group_by(state,appliance) |>  summarize(min(diff))
+
+s1 |> ungroup() |> mutate(diff=family-state_avg) |> group_by(state) |> filter(diff<0) |> select(id,state1,state2,state,appliance,family,state_avg,plan,value,diff)
+
+# examine the relationship between n_less_avg (a value btw 0 and 5) and whether goal was matched
+s1_agg |> ggplot(aes(x=n_less_avg,fill=matched_goal)) +
+  geom_bar(position="dodge") +
+  facet_wrap(~state) +
+  theme_minimal() +
+  labs(title="Relationship between # of appliances changed to below state average and whether goal was matched",
+       x="# of appliances changed to below state average",
+       y="Count") +
+  scale_x_continuous(breaks=seq(1,5,1))
 
@@
 ```
+
+## Data checks
+
+```{r}
+
+library(tidyverse)
+
+# for every combination of state, appliance, and refClass, calculate the number of distinct values of family, and of state_avg
+s1 |> group_by(state,appliance,refClass) |> 
+  summarise(n_distinct_family=n_distinct(family),
+            n_distinct_state_avg=n_distinct(state_avg), 
+            # paste a string of each distinct value, separated by a comma
+            distinct_family=paste(unique(family),collapse=","),
+            distinct_state_avg=paste(unique(state_avg),collapse=",")) |>
+  filter(n_distinct_family>1 | n_distinct_state_avg>1)
+
+s1 |> group_by(state,appliance,refClass) |> 
+  mutate(n_distinct_family=n_distinct(family),
+            n_distinct_state_avg=n_distinct(state_avg), 
+            # paste a string of each distinct value, separated by a comma
+            distinct_family=paste(unique(family),collapse=","),
+            distinct_state_avg=paste(unique(state_avg),collapse=",")) |>
+  filter(n_distinct_family>1 | n_distinct_state_avg>1) |> 
+  group_by(state,appliance,refClass,family,state_avg) 
+  # the count of each unique value of family, and each unique value of state avg
+
+s1 |> 
+    group_by(state, appliance, refClass) |> 
+    mutate(n_distinct_family = n_distinct(family),
+           n_distinct_state_avg = n_distinct(state_avg), 
+           distinct_family = paste(unique(family), collapse = ","),
+           distinct_state_avg = paste(unique(state_avg), collapse = ",")) |>
+    filter(n_distinct_family > 1 | n_distinct_state_avg > 1) |> 
+    group_by(state, appliance, refClass, family, state_avg) |> 
+    mutate(state_avg_count = n_distinct(state_avg),
+           family_count = n()) |> 
+    ungroup() |> 
+    group_by(state, appliance, refClass, state_avg) |> 
+    mutate(state_avg_count = n()) |> 
+    distinct(state, appliance, refClass, family, state_avg, family_count, state_avg_count) |> 
+  arrange(state,appliance,refClass)
+
+haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav"))  |> filter(ResponseId=="R_2QMsYuHS385pTj1") -> k
+# find if any columns equal 5.26
+
+haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav"))  |> select_if(~any(. ==11441.0)) |> 
+  # filter out any NAs from any column that is left (we don't know the column names)
+  filter_all(any_vars(!is.na(.))) |> 
+  # retain only cases of 11441.0
+  filter_all(any_vars(. == 11441.0))
+
+haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav"))  |> # return the names of any columns that contain 11441.0
+  select_if(~any(. ==11441.0)) |> 
+  names() |> 
+  str_c(collapse=",") |> 
+  cat() # P3TEX15
+
+# do the same procedure for each of these values, include a line break after the output from each number
+low_freq <- c(5.26, 4534, 340,11667,5.3,400, 29046,5600,5987,34265.0,5600.0	)
+
+# for(val in low_freq) {
+#   haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav")) |>
+#     select_if(~any(. == val)) |>
+#     names() |>
+#     str_c(collapse = ",") |>
+#     cat()
+#   cat("\n")  # Add line break after each output
+# }
+
+low_freq |>
+  walk(function(val) {
+    cat(val, "\n")  # Print the current value
+    haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav")) |>
+      select_if(~any(. == val)) |>
+      names() |>
+      str_c(collapse = ",") |>
+      cat()
+    cat("\n\n")  # Add two line breaks before next value
+  })
+
+k %>% select_if(~any(is.numeric(.)) & . <10)
+haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav")) |> select(P3TEX15) |> unique()
+
+haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav")) |> 
+  #select(ResponseId, P3TEX15) |> 
+  filter(!is.na(P3TEX15))
+
+# A tibble: 2 × 2
+#   P3TEX15 `n()`
+#     <dbl> <int>
+# 1    5.26    10
+# 2   NA      242
+
+```

### 2024-10-31 misc/outlier.qmd
outlier checks and some model updates

---
title: "Outliers and discrepancies"
author:
- name: Thomas E. Gorman
  url: https://tegorman13.github.io/
  affiliations: 
  - name:  Communication and Cognition Lab, Purdue University, USA
    affiliation-url: https://web.ics.purdue.edu/~treimer/
lightbox: true
toc: false
code-fold: true
code-tools: true
execute:
  #echo: false
  warning: false
format:
  html: 
    grid:
      sidebar-width: 220px
      body-width: 1200px
      margin-width: 170px
      gutter-width: 1.0rem
  # hugo-md:
  #   include: true
  #   html-math-method: mathjax
  #   output-file: results_hugo.md
  # gfm:
  #   echo: true
  #   output-file: results_gfm.md
---

```{r}
library(pacman)
pacman::p_load(dplyr,purrr,tidyr,here, haven,tibble,ggplot2,gt,ggh4x,lme4,flextable,ggdist)

s1 <- readRDS(here::here("data/s1_processed.rds"))

s2_long <- readRDS(here::here("data/s2_processed.rds"))

s1_agg <- s1 |> 
  filter(appliance !="Total kWh") |> 
  group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
  summarise(total_kWh = sum(value),orig_kWh=sum(family), 
            pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
            n_change = sum(value!=family),
            state_p_dif=mean(state_p_dif),
            state_f_dif=mean(state_f_dif),
            n_less_avg = sum(less_avg),
            duration=first(Duration__in_seconds_)) |> 
  mutate(matched_goal = (pct_change == pct_goal), 
             error = pct_change - pct_goal,
            abs_error = abs(error))

# s1_agg |> filter(matched_goal==FALSE, abs_error<.02) |> pull(abs_error)

s1_agg4 <- s1_agg |> group_by(id,refClass) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n)

s1_agg2 <- s1_agg |> group_by(id,refClass,state) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n)

outliers <- s1_agg |>  ungroup() |> group_by(state) |> mutate(change_mean=mean(pct_change,na.rm=TRUE),change_sd=sd(pct_change,na.rm=TRUE),
                                     z_score=(pct_change-change_mean)/change_sd,
                                     is_outlier=abs(z_score)>3.0)
outlier_id <- outliers |> filter(is_outlier) |> pull(id) |> unique()

```

```{r}
#| eval: false

# quick summary of s1$Duration_in_seconds variable
summary(s1$Duration__in_seconds_)
  #  Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
  #   176     846    1184    1659    1688   47914

s1 |> group_by(refClass) |> 
  summarize(mean(Duration__in_seconds_),sd(Duration__in_seconds_),median(Duration__in_seconds_))

s1 |> ggplot(aes(x=Duration__in_seconds_)) + geom_histogram() + facet_wrap(~refClass)

s1 |>  mutate(id=reorder(id,Duration__in_seconds_)) |> 
  ggplot(aes(y=id,x=Duration__in_seconds_)) + 
  geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
  ggh4x::facet_nested_wrap(~refClass,axes="all",scales="free",ncol=2)

s1_agg |> filter(duration<10000) |>
  ggplot(aes(x=duration,y=abs_error)) + geom_point() +
  facet_wrap(~refClass)

s1_agg |> ggplot(aes(x=duration)) + geom_density() + facet_wrap(~refClass) + stat_halfeye()

s2_long |> filter(id=="s2_10") |> head(20)

s2_long |> filter(id %in% unique(s2_long$id)[1:30]) |> 
  filter(appliance!="TOTAL",state=="California") |> 
  ggplot(aes(x=appliance,y=value)) + geom_point(aes(shape=plan)) +
  geom_point(aes(y=state_avg),color="red") +
  geom_point(aes(y=family),color="blue") +
  facet_wrap(~id) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

s2_long |> 
  filter(appliance!="TOTAL",state=="California") |> 
   ggplot(aes(x=appliance,y=family)) + geom_point() + facet_wrap(~rounded)

s2_long |> 
  filter(appliance!="TOTAL",state=="California") |> 
  group_by(appliance,family,rounded) |> 
  # counts for each distinct value
  summarise(n=n())

s2_long |> 
  filter(appliance!="TOTAL",state=="California",plan=="plan1") |> 
  group_by(appliance,family,rounded) |> 
  summarise(n=n()) |>
  ggplot(aes(x=appliance,y=family)) + geom_point(aes(size=n,alpha=.2)) + 
  geom_text(aes(label=n),vjust=1,size=8) +
  facet_wrap(~rounded)

s2_long |> 
  filter(appliance!="TOTAL",plan=="plan1") |> 
  group_by(appliance,family,rounded,state) |> 
  summarise(n=n()) |>
  ggplot(aes(x=appliance,y=family)) + geom_point(aes(size=n,alpha=.2)) + 
  geom_text(aes(label=n),vjust=1,size=7) +
  facet_wrap(state~rounded)

# bar plot version of the above
s2_long |> 
  filter(appliance!="TOTAL",state=="California") |> 
  group_by(appliance,family,rounded) |> 
  summarise(n=n()) |>
  ggplot(aes(x=appliance,y=family,fill=rounded)) + geom_col() + facet_wrap(~rounded)

s2_long %>%
  filter(appliance != "TOTAL", plan == "plan1") %>%
  group_by(appliance, rounded, state) %>%
  mutate(
    distinct_families = n_distinct(family),
    color_flag = distinct_families > 1  # Flag as TRUE if more than 1 distinct value
  ) %>%
  ungroup() %>%
  # Continue grouping for plotting
  group_by(appliance, family, rounded, state,color_flag) %>%
  summarise(n = n(), .groups = "drop") %>%
  ggplot(aes(x = appliance, y = family,color = color_flag,fill=color_flag)) +
 geom_point(aes(alpha = .1, color = color_flag), position = position_jitter(width = 0.15)) +
  geom_text(aes(label = n), size = 4, position = position_jitter(width = 0.45)) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "blue"), guide = "none") +
  facet_wrap(state ~ rounded,scales="free_y") +
  labs(
    x = "Appliance",
    y = "Family Value",
    size = "Frequency (n)"
  )

s2_long %>%
  filter(appliance != "TOTAL", plan == "plan1") %>%
  group_by(appliance, rounded, state) %>%
  mutate(
    distinct_families = n_distinct(state_avg),
    color_flag = distinct_families > 1  # Flag as TRUE if more than 1 distinct value
  ) %>%
  ungroup() %>%
  # Continue grouping for plotting
  group_by(appliance, state_avg, rounded, state,color_flag) %>%
  summarise(n = n(), .groups = "drop") %>%
  ggplot(aes(x = appliance, y = state_avg,color = color_flag,fill=color_flag)) +
 geom_point(aes(alpha = .1, color = color_flag), position = position_jitter(width = 0.15)) +
  geom_text(aes(label = n), size = 4, position = position_jitter(width = 0.15)) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "blue"), guide = "none") +
  facet_wrap(state ~ rounded,scales="free_y") +
  labs(
    x = "Appliance",
    y = "Family Value",
    size = "Frequency (n)"
  )

s1 %>%
  filter(appliance != "Total kWh", plan == "plan1") %>%
  group_by(appliance, state,refClass) %>%
  mutate(
    distinct_families = n_distinct(family),
    color_flag = distinct_families > 1  # Flag as TRUE if more than 1 distinct value
  ) %>%
  ungroup() %>%
  # Continue grouping for plotting
  group_by(appliance, family, state,color_flag,refClass) %>%
  summarise(n = n(), .groups = "drop") %>%
  filter(color_flag) |> 
  ggplot(aes(x = appliance, y = family,color = color_flag,fill=color_flag)) +
 #geom_point(aes(alpha = .1, color = color_flag), position = position_jitter(width = 0.15)) +
  geom_text(aes(label = n), size = 4, position = position_jitter(width = 0.75,height=.1)) +
  scale_color_manual(values = c("FALSE" = "black", "TRUE" = "blue"), guide = "none") +
  facet_wrap(refClass~state,scales="free_y") +
  labs(
    x = "Appliance",
    y = "Family Value",
    size = "Frequency (n)"
  )

```

## Task data checks

```{r}

library(tidyverse)
library(haven)

# for every combination of state, appliance, and refClass, calculate the number of distinct values of family, and of state_avg
s1 |> group_by(state,appliance,refClass) |> 
  mutate(state_avg=as.numeric(state_avg)) |> 
  summarise(n_distinct_family=n_distinct(family),
            n_distinct_state_avg=n_distinct(state_avg), 
            # paste a string of each distinct value, separated by a comma
            distinct_family=paste(unique(family),collapse=","),
            distinct_state_avg=paste(unique(state_avg),collapse=",")) |>
  filter(n_distinct_family>1 | n_distinct_state_avg>1)

s2_long |> group_by(state,appliance,refClass,rounded) |> 
  mutate(state_avg=as.numeric(state_avg)) |> 
  summarise(n_distinct_family=n_distinct(family),
            n_distinct_state_avg=n_distinct(state_avg), 
            # paste a string of each distinct value, separated by a comma
            distinct_family=paste(unique(family),collapse=","),
            distinct_state_avg=paste(unique(state_avg),collapse=",")) |>
  filter(n_distinct_family>1 | n_distinct_state_avg>1)

s1 |> group_by(state,appliance,refClass) |> 
  mutate(n_distinct_family=n_distinct(family),
            n_distinct_state_avg=n_distinct(state_avg), 
            # paste a string of each distinct value, separated by a comma
            distinct_family=paste(unique(family),collapse=","),
            distinct_state_avg=paste(unique(state_avg),collapse=",")) |>
  group_by(state,appliance,refClass,family,state_avg) |>
  select(-calc,-edu,-modality,-id2,-Duration__in_seconds_) |> 
    filter(n_distinct_family>1 | n_distinct_state_avg>1) |> head(20)

s1 |> 
    group_by(state, appliance, refClass) |> 
    mutate(n_distinct_family = n_distinct(family),
           n_distinct_state_avg = n_distinct(state_avg), 
           distinct_family = paste(unique(family), collapse = ","),
           distinct_state_avg = paste(unique(state_avg), collapse = ",")) |>
    filter(n_distinct_family > 1 | n_distinct_state_avg > 1) |> 
    group_by(state, appliance, refClass, family, state_avg) |> 
    mutate(state_avg_count = n_distinct(state_avg),
           family_count = n()) |> 
    ungroup() |> 
    group_by(state, appliance, refClass, state_avg) |> 
    mutate(state_avg_count = n()) |> 
    distinct(state, appliance, refClass, family, state_avg, family_count, state_avg_count) |> 
  arrange(state,appliance,refClass) |> head(15)

problematic_ids <- c("R_1kFwgoKAmNTwCs3") # Add other IDs as needed

raw_subset <-  haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav")) %>%
  filter(ResponseId %in% problematic_ids) %>%
  select(ResponseId, starts_with("P1"), starts_with("P2"), 
         starts_with("P3"), starts_with("P4"))

print(raw_subset)

s2_long |> filter(state=="Texas",appliance=="HEAT", rounded=="Not Rounded",refClass=="kWh") |> arrange(-family)
s2_long |> filter(id=="s2_10") |> head(20) # no effort

```

r$> s1 |> filter(family==34265) |> select(id,state1,state2,refClass,id2,state,appliance,family,state_avg,plan,value,block,pct_goal)
# A tibble: 2 × 13
# Groups:   id [1]
     id state1 state2        refClass id2               state appliance     family state_avg plan  value block pct_goal
  <int> <chr>  <chr>         <chr>    <chr>             <chr> <chr>          <dbl> <dbl+lbl> <fct> <dbl> <dbl>    <dbl>
1    16 Texas  Massachusetts USD      R_3gNKA8IQr0uOSh3 TEX   Water Heating  34265 4396      plan1  5486     1     0.15
2    16 Texas  Massachusetts USD      R_3gNKA8IQr0uOSh3 TEX   Water Heating  34265 4396      plan2  6414     1     0.15

s2_long |> group_by(state,appliance,refClass,rounded) |> 
  mutate(state_avg=as.numeric(state_avg)) |> 
  summarise(n_distinct_family=n_distinct(family),
            n_distinct_state_avg=n_distinct(state_avg), 
            # paste a string of each distinct value, separated by a comma
            distinct_family=paste(unique(family),collapse=","),
            distinct_state_avg=paste(unique(state_avg),collapse=",")) |>
  filter(n_distinct_family>1 | n_distinct_state_avg>1) 
 # A tibble: 8 × 8
# Groups:   state, appliance, refClass [8]
  state         appliance refClass   rounded     n_distinct_family n_distinct_state_avg distinct_family distinct_state_avg
  <chr>         <chr>     <chr>      <chr>                   <int>                <int> <chr>           <chr>             
1 California    HEAT      kWh        Not Rounded                 2                    1 3000,6157       5597              
2 Colorado      HEAT      USD        Rounded                     1                    2 18000           16411,NA          
3 Colorado      WATER     Percentage Not Rounded                 1                    2 11667           1142,5832         
4 Massachusetts AC        Percentage Rounded                     2                    1 0,1             322               
5 Massachusetts AC        USD        Rounded                     2                    1 0,1             322               
6 Texas         HEAT      Percentage Not Rounded                 2                    1 6118,5          5099              
7 Texas         HEAT      kWh        Not Rounded                 2                    1 8539,6118       5099              
8 Texas         WATER     Percentage Not Rounded                 2                    1 5.257,5257      4396

r$> s2_long |> filter(family==3000,state=="California",appliance=="HEAT")
# A tibble: 2 × 16
# Groups:   id, block [1]
  id    block refClass appliance pct_goal rounded     state      calc          edu            pct family state_avg plan  value change state_dif
  <chr> <dbl> <chr>    <chr>     <chr>    <chr>       <chr>      <chr>         <fct>        <dbl>  <dbl>     <dbl> <fct> <dbl>  <dbl>     <dbl>
1 s2_11     1 kWh      HEAT      10%      Not Rounded California No Calculator Some college   0.1   3000      5597 plan1  4000   1000     -1597
2 s2_11     1 kWh      HEAT      10%      Not Rounded California No Calculator Some college   0.1   3000      5597 plan2  5000   2000      -597

s2_long |> filter(family==1)
  id    block refClass   appliance pct_goal rounded state         calc          edu              pct family state_avg plan  value change state_dif
  <chr> <dbl> <chr>      <chr>     <chr>    <chr>   <chr>         <chr>         <fct>          <dbl>  <dbl>     <dbl> <fct> <dbl>  <dbl>     <dbl>
1 s2_8      1 Percentage AC        15%      Rounded Massachusetts Calculator    College degree  0.15      1       322 plan1     1      0      -321
2 s2_8      1 Percentage AC        15%      Rounded Massachusetts Calculator    College degree  0.15      1       322 plan2     1      0      -321
3 s2_9      2 USD        AC        10%      Rounded Massachusetts No Calculator College degree  0.1       1       322 plan1     1      0      -321
4 s2_9      2 USD        AC        10%      Rounded Massachusetts No Calculator College degree  0.1       1       322 plan2     1      0      -321

## Wells Family 15 R and Smith Family 10 E Percentage - 5.257 for Texas water heating

 s2_long |> filter(family==5.257)
# A tibble: 8 × 16
# Groups:   id, block [4]
  id    block refClass   appliance pct_goal rounded     state calc          edu               pct family state_avg plan  value change state_dif
  <chr> <dbl> <chr>      <chr>     <chr>    <chr>       <chr> <chr>         <fct>           <dbl>  <dbl>     <dbl> <fct> <dbl>  <dbl>     <dbl>
1 s2_1      2 Percentage WATER     10%      Not Rounded Texas Calculator    Some college      0.1   5.26      4396 plan1  5000  4995.       604
2 s2_1      2 Percentage WATER     10%      Not Rounded Texas Calculator    Some college      0.1   5.26      4396 plan2  5000  4995.       604
3 s2_2      2 Percentage WATER     10%      Not Rounded Texas No Calculator Graduate degree   0.1   5.26      4396 plan1  4396  4391.         0
4 s2_2      2 Percentage WATER     10%      Not Rounded Texas No Calculator Graduate degree   0.1   5.26      4396 plan2  5257  5252.       861
5 s2_3      2 Percentage WATER     10%      Not Rounded Texas Calculator    Graduate degree   0.1   5.26      4396 plan1  3527  3522.      -869
6 s2_3      2 Percentage WATER     10%      Not Rounded Texas Calculator    Graduate degree   0.1   5.26      4396 plan2  3027  3022.     -1369
7 s2_4      2 Percentage WATER     10%      Not Rounded Texas Calculator    Graduate degree   0.1   5.26      4396 plan1  4800  4795.       404
8 s2_4      2 Percentage WATER     10%      Not Rounded Texas Calculator    Graduate degree   0.1   5.26      4396 plan2  5000  4995.       604

```{r}
library(dplyr)
library(flextable)
library(officer)

correct_values <- tribble(
  ~state, ~appliance, ~family, ~state_avg,
  # Texas values
  "Texas", "Cooling", 6573, 4249,
  "Texas", "Heating", 6118, 5099,
  "Texas", "Water Heating", 5257, 4396,
  "Texas", "Refrigerator", 2639, 1318,
  "Texas", "Other Appliance", 8459, 7883,
  "Texas", "Total kWh", 29046, 22945,

  # California values
  "California", "Cooling", 2581, 1289,
  "California", "Heating", 6157, 5597,
  "California", "Water Heating", 5061, 4601,
  "California", "Refrigerator", 1266, 1055,
  "California", "Other Appliance", 7608, 6916,
  "California", "Total kWh", 22673, 19458,

  # Colorado values
  "Colorado", "Cooling", 697, 498,
  "Colorado", "Heating", 18052, 16411,
  "Colorado", "Water Heating", 11667, 5832,
  "Colorado", "Refrigerator", 1370, 1142,
  "Colorado", "Other Appliance", 7882, 6652,
  "Colorado", "Total kWh", 39768, 30535,

  # Massachusetts values
  "Massachusetts", "Cooling", 419, 322,
  "Massachusetts", "Heating", 26751, 19108,
  "Massachusetts", "Water Heating", 10543, 5070,
  "Massachusetts", "Refrigerator", 1230, 1025,
  "Massachusetts", "Other Appliance", 7350, 6682,
  "Massachusetts", "Total kWh", 46293, 32207
)

# First, create a lookup table of correct values
correct_lookup <- correct_values %>%
  select(state, appliance, family) %>%
  mutate(state = case_when(
    state == "Texas" ~ "TEX",
    state == "California" ~ "CAL",
    state == "Colorado" ~ "COL",
    state == "Massachusetts" ~ "MASS",
    TRUE ~ state
  ))

# Modify the main table
s1 %>%
  filter(appliance != "Total kWh", plan == "plan1") %>%
  group_by(appliance, state, refClass) %>%
  mutate(
    distinct_families = n_distinct(family),
    color_flag = distinct_families > 1
  ) %>%
  ungroup() %>%
  filter(color_flag) %>%
  group_by(appliance, family, state, color_flag, refClass) %>%
  summarise(n = n(), .groups = "drop") %>%
  select(-color_flag) %>%
  # Join with correct values to create is_correct flag
  left_join(correct_lookup, 
            by = c("state", "appliance"),
            suffix = c("", "_correct")) %>%
  mutate(is_correct = abs(as.numeric(family) - family_correct) < 0.1) %>%
  select(-family_correct) %>%
  # Reorganize and format
  select(state, refClass, appliance, family, n, is_correct) %>%  # Swapped order here
  arrange(state, appliance, refClass, family) %>%
  flextable() %>%
  # Merge cells
  merge_v(j = "state") %>%
  merge_v(j = c("appliance", "refClass")) %>%
  # Conditional formatting for both family and count
  color(i = ~ is_correct == TRUE, j = c("family", "n"), color = "darkgreen") %>%
  color(i = ~ is_correct == FALSE, j = c("family", "n"), color = "darkred") %>%
  bg(i = ~ is_correct == TRUE, j = c("family", "n"), bg = "#e6ffe6") %>%
  bg(i = ~ is_correct == FALSE, j = c("family", "n"), bg = "#ffe6e6") %>%
  # Style improvements
  theme_vanilla() %>%
  align(j = c("family", "n"), align = "right") %>%
  border_outer() %>%
  border_inner() %>%
  fontsize(size = 11) %>%
  padding(padding = 4) %>%
  # Add section borders
  hline(i = ~ state != lag(state, default = first(state)), 
        border = officer::fp_border(color = "gray40", width = 2)) %>%
  colformat_double(j = "family", digits = 1) %>%
  set_header_labels(
    state = "State",
    appliance = "Appliance",
    family = "Family",
    refClass = "Reference Class",
    n = "Count",
    is_correct = "Correct Value"
  ) %>%
  autofit()

library(dplyr)
library(flextable)
library(officer)

s1 %>%
  filter(appliance != "Total kWh", plan == "plan1") %>%
  group_by(appliance, state, refClass) %>%
  mutate(distinct_state_avg = n_distinct(state_avg)) %>%
  ungroup() %>%
  filter(distinct_state_avg > 1) %>%
  group_by(appliance, state_avg, state, refClass) %>%
  summarise(n = n(), .groups = "drop") %>%
  # Identify outliers (rows where n is 1)
  mutate(is_outlier = n == 1) %>%
  # Reorganize and format
  select(state, refClass, appliance, state_avg, n) %>%
  arrange(state, appliance, refClass, state_avg) %>%
  flextable() %>%
  # Merge cells
  merge_v(j = "state") %>%
  merge_v(j = c("appliance", "refClass")) %>%
  # Conditional formatting for outliers
  color(i = ~ n == 1, j = c("state_avg", "n"), color = "darkred") %>%
  color(i = ~ n > 1, j = c("state_avg", "n"), color = "darkgreen") %>%
  bg(i = ~ n == 1, j = c("state_avg", "n"), bg = "#ffe6e6") %>%
  bg(i = ~ n > 1, j = c("state_avg", "n"), bg = "#e6ffe6") %>%
  # Style improvements
  theme_vanilla() %>%
  align(j = c("state_avg", "n"), align = "right") %>%
  border_outer() %>%
  border_inner() %>%
  fontsize(size = 11) %>%
  padding(padding = 4) %>%
  # Add section borders
  hline(i = ~ state != lag(state, default = first(state)), 
        border = officer::fp_border(color = "gray40", width = 2)) %>%
  colformat_double(j = "state_avg", digits = 0) %>%
  set_header_labels(
    state = "State",
    appliance = "Appliance",
    state_avg = "State Average",
    refClass = "Reference Class",
    n = "Count"
  ) %>%
  autofit()
```

### Alternate discrep frequency tables

```{r}
library(dplyr)
library(gt)

table_data <- s1 %>%
  filter(appliance != "Total kWh", plan == "plan1") %>%
  group_by(appliance, state, refClass) %>%
  mutate(distinct_families = n_distinct(family)) %>%
  ungroup() %>%
  filter(distinct_families > 1) %>%
  group_by(appliance, family, state, refClass) %>%
  summarise(n = n(), .groups = "drop") %>%
  arrange(state, refClass, appliance)

# Prepare the data with separate columns for Family and Frequency
table_data_compact <- table_data %>%
  arrange(state, appliance, refClass, desc(n)) %>%  # Sort by descending frequency
  select(state, refClass, appliance, family, n) %>%
  group_by(state, refClass, appliance) %>%
  summarise(
    Family = paste(family[order(-n)], collapse = ", "),  # Order family values by descending frequency
    Frequency = paste(n[order(-n)], collapse = ", "),    # Order frequency values in descending order
    .groups = "drop"
  )

# Create the GT table with the new format
table_data_compact %>%
  gt(rowname_col = "state") %>%
  tab_header(
    title = "Compact Frequency Table by State, RefClass, and Appliance"
  ) %>%
  # Label columns for clarity
  cols_label(
    refClass = "Reference Class",
    appliance = "Appliance",
    Family = "Family last year",
    Frequency = "Frequency (n)"
  ) %>%
  # Style the table headers
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(columns = everything())
  ) %>%
  # Center the row group headers for each state
  tab_style(
    style = list(cell_text(weight = "bold", align = "center")),
    locations = cells_row_groups(groups = everything())
  ) %>%
  # Style the appliance rows with lighter borders for each appliance within the state
  tab_options(
    row_group.border.bottom.color = "grey",
    row_group.border.bottom.width = px(1),
    table.font.size = px(13),
    data_row.padding = px(3),  # Compact padding
    row.striping.include_stub = TRUE,
    row.striping.background_color = "lightgrey",
    table.width = pct(100)
  )

#######

library(dplyr)
library(gt)

# Prepare the data, focusing on `state_avg` instead of `family`
table_data_state_avg <- s1 %>%
  filter(appliance != "Total kWh", plan == "plan1") %>%
  group_by(appliance, state, refClass) %>%
  mutate(distinct_state_avg = n_distinct(state_avg)) %>%
  ungroup() %>%
  filter(distinct_state_avg > 1) %>%
  group_by(appliance, state_avg, state, refClass) %>%
  summarise(n = n(), .groups = "drop") %>%
  arrange(state, refClass, appliance, desc(n)) %>%  # Sort by descending frequency
  select(state, refClass, appliance, state_avg, n) %>%
  group_by(state, refClass, appliance) %>%
  summarise(
    State_Avg = paste(state_avg[order(-n)], collapse = ", "),  # Order state_avg values by descending frequency
    Frequency = paste(n[order(-n)], collapse = ", "),          # Order frequency values in descending order
    .groups = "drop"
  )

# Create the GT table with the new format
table_data_state_avg %>%
  gt(rowname_col = "state") %>%
  tab_header(
    title = "Compact Frequency Table by RefClass, Appliance, and State Average"
  ) %>%
  # Label columns for clarity
  cols_label(
    refClass = "Reference Class",
    appliance = "Appliance",
    State_Avg = "State Avg",
    Frequency = "Frequency (n)"
  ) %>%
  # Style the table headers
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(columns = everything())
  ) %>%
  # Center the row group headers for each state
  tab_style(
    style = list(cell_text(weight = "bold", align = "center")),
    locations = cells_row_groups(groups = everything())
  ) %>%
  # Style the appliance rows with lighter borders for each appliance within the state
  tab_options(
    row_group.border.bottom.color = "grey",
    row_group.border.bottom.width = px(1),
    table.font.size = px(13),
    data_row.padding = px(3),  # Compact padding
    row.striping.include_stub = TRUE,
    row.striping.background_color = "lightgrey",
    table.width = pct(100)
  )
```

```{r}
#| eval: false

library(dplyr)
library(purrr)
library(tidyr)
library(haven)
library(here)

# Define a mapping between state codes and full state names
state_code_map <- c(
  "CAL" = "California",
  "TEX" = "Texas",
  "COL" = "Colorado",
  "MASS" = "Massachusetts"
)

# Revised parse_action_plan function
parse_action_plan <- function(df) {

  # Extract participant-specific states
  participant_states <- c(unique(df$state1), unique(df$state2))

  # Select action plan columns based on state codes
  action_plan_cols <- df %>% 
    select(matches("^[ADP]\\d+(CAL|MASS|COL|TEX)\\d{2,3}$")) %>%
    select(-matches("32$")) %>%  # Exclude columns ending with '32'
    names()

  # Pivot longer to reshape data
  df_long <- df %>%
    pivot_longer(
      cols = all_of(action_plan_cols),
      names_to = c("modality", "state_code", "item_num"),
      names_pattern = "^([ADP])\\d+(CAL|MASS|COL|TEX)(\\d+)$",
      values_to = "value"
    ) %>%
    mutate(
      # Map state codes to full state names
      state_full = state_code_map[state_code],
      # Ensure that the state_full is one of the participant's assigned states
      state_assigned = if_else(state_full %in% participant_states, state_full, NA_character_)
    ) %>%
    filter(!is.na(state_assigned)) %>%  # Keep only matching states
    select(id, state1, state2, refClass, calc, edu, modality, state_assigned, item_num, value, id2, Duration__in_seconds_)

  # Assign appliance based on item_num
  df_long <- df_long %>%
    mutate(
      appliance = case_when(
        item_num %in% c("07","08","09","10") ~ "Cooling",
        item_num %in% c("11","12","13","14") ~ "Heating",
        item_num %in% c("15","16","17","18") ~ "Water Heating",
        item_num %in% c("19","20","21","22") ~ "Refrigerator",
        item_num %in% c("23","24","25","26") ~ "Other Appliance",
        item_num %in% c("27","28","29","30") ~ "Total kWh",
        item_num == "31" ~ "Preferred Plan",
        item_num %in% c("33","34","35","36","37") ~ "Reduction Choices",
        TRUE ~ NA_character_
      ),
      # Assign item_type based on item_num
      item_type = case_when(
        item_num %in% c("07","11","15","19","23","27") ~ "family",
        item_num %in% c("08","12","16","20","24","28") ~ "state_avg",
        item_num %in% c("09","13","17","21","25","29") ~ "plan1",
        item_num %in% c("10","14","18","22","26","30") ~ "plan2",
        TRUE ~ "other"
      )
    )

  # Continue processing only relevant item_types
  df_long <- df_long %>%
    filter(item_type != "other") %>%
    select(-item_num) %>%
    pivot_wider(names_from = "item_type", values_from = "value") %>%
    pivot_longer(cols = c("plan1", "plan2"), names_to = "plan", values_to = "value")

  return(df_long)
}

# Read and preprocess the raw data
s1 <- haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav"))

s1 <- s1 %>% 
  mutate(
    state1 = case_when(
      Colorado == 1 ~ "Colorado",
      Massachusetts == 1 ~ "Massachusetts",
      California == 1 ~ "California",
      Texas == 1 ~ "Texas",
      TRUE ~ NA_character_
    ),
    state2 = case_when(
      Colorado == 2 ~ "Colorado",
      Massachusetts == 2 ~ "Massachusetts",
      California == 2 ~ "California",
      Texas == 2 ~ "Texas",
      TRUE ~ NA_character_
    ),
    refClass = case_when(
      ReferenceClass == 1 ~ "USD",
      ReferenceClass == 2 ~ "kWh",
      ReferenceClass == 3 ~ "Percentage",
      TRUE ~ NA_character_
    ),
    calc = factor(if_else(is.na(MATH01), 0, MATH01)),
    id = row_number()
  ) %>%
  rename(id2 = ResponseId, edu = DEM04) %>%
  mutate(
    edu = factor(case_when(
      edu == 1 ~ "Some schooling, but no diploma or degree",
      edu == 2 ~ "Highschool diploma or GED",
      edu == 3 ~ "Some college",
      edu == 4 ~ "College degree",
      edu == 5 ~ "Some graduate school",
      edu == 6 ~ "Graduate degree",
      edu == 7 ~ "Choose not to answer",
      TRUE ~ NA_character_
    ))
  ) %>%
  relocate(id, edu, refClass, state1, state2, calc, .after = id2)

# Function to trim NA and empty columns
trim_na_and_empty_columns <- function(df) {
  df %>%
    select(where(~ !(all(is.na(.)) | (is.character(.) && all(trimws(.) == ""))))) %>% 
    select(-starts_with("AIND"))
}

# Create a list of trimmed datasets for each unique ID
dAll <- s1 %>%
  group_by(id) %>%
  group_split() %>%
  map(trim_na_and_empty_columns) %>%
  set_names(unique(s1$id))

# Apply the revised parse_action_plan function to all participants
test <- dAll %>%
  map(parse_action_plan)

# Combine all parsed data
s1_processed <- bind_rows(test)

# Handle specific data corrections (e.g., participant "R_1kFwgoKAmNTwCs3")
s1_processed <- s1_processed %>%
  mutate(
    family = if_else(id2 == "R_1kFwgoKAmNTwCs3" & state_assigned == "Texas" & appliance == "Heating", 6118, family)
  )

# Final mutations and calculations
s1_processed <- s1_processed %>%
  mutate(
    value = as.numeric(value),
    change = value - family,
    state_p_dif = value - as.numeric(state_avg),
    state_f_dif = family - as.numeric(state_avg),
    less_avg = value < as.numeric(state_avg),
    calc = if_else(calc == 0, "No Calculator", "Used Calculator")
  )

s1_processed |> group_by(state_assigned,appliance) |> 
  mutate(state_avg=as.numeric(state_avg)) |> 
  summarise(n_distinct_family=n_distinct(family),
            n_distinct_state_avg=n_distinct(state_avg), 
            # paste a string of each distinct value, separated by a comma
            distinct_family=paste(unique(family),collapse=","),
            distinct_state_avg=paste(unique(state_avg),collapse=",")) |>
  filter(n_distinct_family>1 | n_distinct_state_avg>1)

# Convert family and state_avg to numeric
s1_processed <- s1_processed %>%
  mutate(
    family_value = as.numeric(family),
    state_avg_value = as.numeric(state_avg)
  )

# Identify discrepancies and include participant IDs
duplicate_checks <- s1_processed %>%
  group_by(state_assigned, appliance, refClass) %>%
  summarise(
    n_unique_family = n_distinct(family_value),
    n_unique_state_avg = n_distinct(state_avg_value),
    families = paste(unique(family_value), collapse = ", "),
    state_avgs = paste(unique(state_avg_value), collapse = ", "),
    participant_ids = paste(unique(id[duplicated(family_value) | duplicated(state_avg_value)]), collapse = ", "),
    .groups = 'drop'
  ) %>%
  filter(n_unique_family > 1 | n_unique_state_avg > 1)

if(nrow(duplicate_checks) > 0){
  cat("Found inconsistencies in family or state_avg values:\n")
  print(duplicate_checks)
} else {
  cat("No inconsistencies found in family or state_avg values.\n")
}
```

```{r}
#| label: s1_check
#| 
s1_processed <- s1 |> mutate(state_assigned=state)

s1_check <- s1_processed %>% 
  mutate(
    family_value = as.numeric(family),
    state_avg_value = as.numeric(state_avg)
  ) %>% 
  select(id, id2, state_assigned, appliance, family_value, state_avg_value, refClass,state)

# Rename columns in correct_values
correct_values <- correct_values %>%
  rename(
    family_correct = family,
    state_avg_correct = state_avg
  )

# Join participant data with correct values
comparison <- left_join(s1_check, correct_values, by = c("state", "appliance"))

# Compute differences between participant values and correct values
comparison <- comparison %>%
  mutate(
    family_diff = family_value - family_correct,
    state_avg_diff = state_avg_value - state_avg_correct
  )

# Identify discrepancies
discrepancies <- comparison %>%
  filter(family_diff != 0 | state_avg_diff != 0)

  if (nrow(discrepancies) > 0) {
  cat("Participants with discrepancies in family or state_avg values:\n")
  print(
    discrepancies %>%
      select(
        id,
        id2,
        state,
        appliance,
        refClass,
        family_value,
        family_correct,
        family_diff,
        state_avg_value,
        state_avg_correct,
        state_avg_diff
      )
  )
} else {
  cat("No discrepancies found between participant values and correct values.\n")
}

# Function to inspect a participant's data
inspect_participant <- function(participant_id) {
  participant_data <- comparison %>%
    filter(id == participant_id) %>%
    select(
      id,
      id2,
      state,
      appliance,
      refClass,
        family_value,
        family_correct,
        family_diff,
        state_avg_value,
        state_avg_correct,
        state_avg_diff
    )

 print(participant_data, n=24)
}

# Example usage:
inspect_participant(1)   # Replace with the participant ID you want to investigate
inspect_participant(11)

```

```{r}

correct_values <- tribble(
  ~state, ~appliance, ~family, ~state_avg,
  # Texas values
  "Texas", "Cooling", 6573, 4249,
  "Texas", "Heating", 6118, 5099,
  "Texas", "Water Heating", 5257, 4396,
  "Texas", "Refrigerator", 2639, 1318,
  "Texas", "Other Appliance", 8459, 7883,
  "Texas", "Total kWh", 29046, 22945,

  # California values
  "California", "Cooling", 2581, 1289,
  "California", "Heating", 6157, 5597,
  "California", "Water Heating", 5061, 4601,
  "California", "Refrigerator", 1266, 1055,
  "California", "Other Appliance", 7608, 6916,
  "California", "Total kWh", 22673, 19458,

  # Colorado values
  "Colorado", "Cooling", 697, 498,
  "Colorado", "Heating", 18052, 16411,
  "Colorado", "Water Heating", 11667, 5832,
  "Colorado", "Refrigerator", 1370, 1142,
  "Colorado", "Other Appliance", 7882, 6652,
  "Colorado", "Total kWh", 39768, 30535,

  # Massachusetts values
  "Massachusetts", "Cooling", 419, 322,
  "Massachusetts", "Heating", 26751, 19108,
  "Massachusetts", "Water Heating", 10543, 5070,
  "Massachusetts", "Refrigerator", 1230, 1025,
  "Massachusetts", "Other Appliance", 7350, 6682,
  "Massachusetts", "Total kWh", 46293, 32207
)
```

```{r}
#| eval: false
#| 
# Function to inspect raw data for a specific participant and state
inspect_raw_values <- function(raw_data, response_id, state_code) {
  # Get participant's raw data
  participant_data <- raw_data %>%
    filter(ResponseId == response_id)

  # Get relevant columns for the state
  relevant_cols <- names(participant_data)[str_detect(names(participant_data), 
                                                    paste0("^[ADP]\\d+", state_code))]

  # Create a mapping for item numbers to appliance types
  appliance_map <- tribble(
    ~item_nums, ~appliance,
    "07|08", "Cooling",
    "11|12", "Heating",
    "15|16", "Water Heating",
    "19|20", "Refrigerator",
    "23|24", "Other Appliance",
    "27|28", "Total kWh"
  )

  # Extract and organize values
  results <- map_dfr(appliance_map$item_nums, function(nums) {
    cols <- relevant_cols[str_detect(relevant_cols, paste0("(", nums, ")$"))]
    values <- participant_data %>% select(all_of(cols)) %>% unlist()

    tibble(
      column = names(values),
      value = as.numeric(values),
      appliance = appliance_map$appliance[appliance_map$item_nums == nums]
    )
  })

  return(results)
}

# Let's check participant 1's Texas Water Heating issue
raw_data <- haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav"))

# Check Participant 1 (R_2QMsYuHS385pTj1)
print("Participant 1 - Texas Water Heating Values:")
p1_tex <- inspect_raw_values(raw_data, "R_2QMsYuHS385pTj1", "TEX")
print(p1_tex %>% filter(str_detect(appliance, "Water")),n=24)

# Check Participant 11 (R_263zmNZDB4Pqn9h)
print("\nParticipant 11 - Colorado Heating Values:")
p11_col <- inspect_raw_values(raw_data, "R_263zmNZDB4Pqn9h", "COL")
print(p11_col %>% filter(str_detect(appliance, "Heating")))

# Function to compare raw vs processed values
compare_values <- function(raw_data, processed_data, response_id, state, appliance_type) {
  # Get raw values
  raw_vals <- inspect_raw_values(raw_data, response_id, 
                                case_when(
                                  state == "Texas" ~ "TEX",
                                  state == "Colorado" ~ "COL",
                                  state == "California" ~ "CAL",
                                  state == "Massachusetts" ~ "MASS"
                                )) %>%
    filter(appliance == appliance_type)

  # Get processed values
  proc_vals <- processed_data %>%
    filter(id2 == response_id, 
           state == !!state,
           appliance == appliance_type) %>%
    select(family, state_avg_value)

  # Compare
  tibble(
    source = "Raw Data",
    family = raw_vals$value[str_detect(raw_vals$column, "07$|11$|15$|19$|23$|27$")],
    state_avg = raw_vals$value[str_detect(raw_vals$column, "08$|12$|16$|20$|24$|28$")]
  ) %>%
    bind_rows(
      tibble(
        source = "Processed Data",
        family = first(proc_vals$family),
        state_avg = first(proc_vals$state_avg_value)
      )
    )
}

# Example comparisons
print("\nComparison for Participant 1 - Texas Water Heating:")
compare_values(raw_data, s1_processed, "R_2QMsYuHS385pTj1", "Texas", "Water Heating")
unique(raw_data$P3TEX15)
s1_processed |> filter(state2=="Texas",appliance=="Water Heating",state_assigned=="Texas",plan=="plan1")

print("\nComparison for Participant 11 - Colorado Heating:")
compare_values(raw_data, s1_processed, "R_263zmNZDB4Pqn9h", "Colorado", "Heating")

# Check if the discrepancy is consistent across different blocks/presentations
check_all_presentations <- function(raw_data, response_id, state_code, appliance_pattern) {
  participant_data <- raw_data %>%
    filter(ResponseId == response_id)

  all_cols <- names(participant_data)[str_detect(names(participant_data), 
                                               paste0("^[ADP]\\d+", state_code)) &
                                      str_detect(names(participant_data), appliance_pattern)]

  participant_data %>%
    select(all_of(all_cols)) %>%
    gather(key = "column", value = "value") %>%
    arrange(column)
}

# Example: Check all presentations of water heating for participant 1
print("\nAll presentations of water heating for Participant 1:")
check_all_presentations(raw_data, "R_2QMsYuHS385pTj1", "TEX", "15|16")

unique(raw_data$P3TEX15)
```

## Just refClass Percentage and state 2 Texas have the issues with family water heating = 5.26

r$> s1_processed |> filter(state2=="Texas",appliance=="Water Heating",state_assigned=="Texas",plan=="plan1") |> head(15)
# A tibble: 15 × 17
      id state1        state2 refClass   calc  edu             modality state_assigned id2               Duration__in_seconds_ appliance     family    state_avg plan  value     family_value state_avg_value
   <int> <chr>         <chr>  <chr>      <fct> <fct>           <chr>    <chr>          <chr>                             <dbl> <chr>         <dbl+lbl> <dbl+lbl> <chr> <dbl+lbl>        <dbl>           <dbl>
 1     1 Colorado      Texas  Percentage 1     Graduate degree P        Texas          R_2QMsYuHS385pTj1                  1008 Water Heating    5.26   4396      plan1  700              5.26            4396
 2     2 Colorado      Texas  Percentage 1     College degree  P        Texas          R_2tzIyZavpl1RZBY                  1126 Water Heating    5.26   4396      plan1 4396              5.26            4396
 3     3 Colorado      Texas  Percentage 1     College degree  P        Texas          R_3Gm1Y4QzSlyrEWo                   870 Water Heating    5.26   4396      plan1 4468              5.26            4396
 4     4 Colorado      Texas  Percentage 1     College degree  P        Texas          R_Djh2q728ibyNWjD                  2763 Water Heating    5.26   4396      plan1 1000              5.26            4396
 5     5 Colorado      Texas  Percentage 0     Graduate degree P        Texas          R_3fAXNbIZ7kFieSu                   920 Water Heating    5.26   4396      plan1 3756              5.26            4396
 6     6 Colorado      Texas  Percentage 0     Graduate degree P        Texas          R_1JECzl10KgckJXg                   878 Water Heating    5.26   4396      plan1 5200              5.26            4396
 7     7 Colorado      Texas  Percentage 1     Graduate degree P        Texas          R_C2IQAZim25sij4d                  2680 Water Heating    5.26   4396      plan1 4257              5.26            4396
 8     8 Colorado      Texas  Percentage 1     Some college    P        Texas          R_5bbi3tpMOhuGBIR                  1628 Water Heating    5.26   4396      plan1 5100              5.26            4396
 9     9 Colorado      Texas  Percentage 1     Graduate degree P        Texas          R_2BsVSAOpehXW9Cr                   724 Water Heating    5.26   4396      plan1 5217              5.26            4396
10    10 Colorado      Texas  Percentage 1     Some college    P        Texas          R_325F8wRln52w1QN                   913 Water Heating    5.26   4396      plan1 4468              5.26            4396
11    12 Colorado      Texas  kWh        1     Graduate degree A        Texas          R_abEpO6g4xKOlwwF                   644 Water Heating 5257      4396      plan1 8459           5257               4396
12    14 Colorado      Texas  USD        0     College degree  D        Texas          R_2VsaxwvgHXasQx3                   625 Water Heating 5257       400      plan1 4000           5257                400
13    24 Massachusetts Texas  USD        1     College degree  D        Texas          R_AEvC2EEL0GltodP                  1047 Water Heating 5257      4396      plan1 4000           5257               4396
14    36 Colorado      Texas  USD        0     Graduate degree D        Texas          R_1nO7uu36tUOSbfw                   390 Water Heating 5257      4396      plan1 4500           5257               4396
15    38 Massachusetts Texas  kWh        1     College degree  A        Texas          R_6f13I34krdFacjT                   973 Water Heating 5257      4396      plan1 4500           5257               4396

### 2024-10-31 results.qmd
outlier checks and some model updates

@@
   summarise(total_kWh = sum(value),orig_kWh=sum(family), 
-            pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), state_dif=mean(state_dif)) |> 
-  mutate(matched_goal = (pct_change == pct_goal))
+            pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
+            n_change = sum(value!=family),
+            state_p_dif=mean(state_p_dif),
+            state_f_dif=mean(state_f_dif),
+            n_less_avg = sum(less_avg)) |> 
+  mutate(matched_goal = (pct_change == pct_goal), 
+             error = pct_change - pct_goal,
+            abs_error = abs(error))
 
@@
 
-mted1 <- as.data.frame(describe_posterior(s1_bb1, centrality = "Mean"))[, c(1,2,4,5,6)]
-colnames(mted1) <- c("Term", "Estimate"," 95% CrI Lower", " 95% CrI Upper", "pd")
+mted1 <- as.data.frame(describe_posterior(s1_bb1, centrality = "Median"))[, c(1,2,4,5,6)]
+#colnames(mted1) <- c("Term", "Estimate"," 95% CrI Lower", " 95% CrI Upper", "pd")
+colnames(mted1) <- c("Term", "Estimate", "95% CrI Lower", "95% CrI Upper", "pd")
 
-mted1 |>
-  mutate(across(c("Estimate", " 95% CrI Lower", " 95% CrI Upper"), 
-                \(x) plogis(x))) |>
+intercept <- mted1$Estimate[mted1$Term == "b_Intercept"]
+
+# mted1 |>
+#   mutate(
+#     Estimate_prob = plogis(Estimate + if_else(Term == "b_Intercept", 0, intercept)),
+#     Lower_prob = plogis(`95% CrI Lower` + if_else(Term == "b_Intercept", 0, intercept)),
+#     Upper_prob = plogis(`95% CrI Upper` + if_else(Term == "b_Intercept", 0, intercept))
+#   )
+
+t1 <- mted1 |>
+  mutate(Term = stringr::str_remove(Term, "b_")) |>
+  mutate(across(c("Estimate", "95% CrI Lower", "95% CrI Upper"), 
+                \(x) if_else(Term == "Intercept", plogis(x), plogis(x + intercept)))) |>
   mutate(across(where(is.numeric), \(x) round(x, 3))) |>
+  mutate(Term = if_else(Term == "Intercept", "Intercept (kWh)", Term)) |>
   tibble::remove_rownames() |>
-  mutate(Term = stringr::str_remove(Term, "b_")) |>
   kable(booktabs = TRUE)
 
+# plogis(-.58+ -.70) = .22
+
+# parameters::parameters(s1_bb1, effect = "fixed")
+
 ```
+
+::: {#tbl-s1-bb}
+|Term               | Estimate| 95% CrI Lower| 95% CrI Upper| pd|
+|:------------------|--------:|-------------:|-------------:|--:|
+|Intercept (kWh)    |     0.36|          0.35|          0.36|  1|
+|refClassPercentage |     0.22|          0.21|          0.22|  1|
+|refClassUSD        |     0.09|          0.09|          0.09|  1|
+
+: **Experiment 1 main effect**. Beta binomial regression results for the main effect of Reference Class.
+
+:::
 
@@
   facet_wrap(~refClass)
+
+s1_bb1 %>%
+  gather_draws(b_Intercept, b_refClassPercentage, b_refClassUSD) %>%
+   mutate(Term = stringr::str_remove(.variable, "b_")) |>
+   mutate(prob=plogis(.value)) |>
+  mutate(Term = if_else(Term == "Intercept", "kWh", Term)) |>
+  ggplot(aes(x = prob, fill = Term)) +
+  geom_density(alpha = 0.6) +
+    stat_halfeye() +
+  theme_minimal() +
+  labs(title = "Posterior Distributions of Reference Class Estimates",
+       x = "Coefficient Value",
+       fill = "Reference Class")
+
 ```
@@
 )
+mted2 <- as.data.frame(describe_posterior(s2_bb2_r_g, centrality = "Mean"))[, c(1,2,4,5,6)]
+colnames(mted2) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")
 
-mted2 <- as.data.frame(describe_posterior(s2_bb2_r_g, centrality = "Mean"))[, c(1,2,4,5,6)]
-colnames(mted2) <- c("Term", "Estimate"," 95% CrI Lower", " 95% CrI Upper", "pd")
+# mted2 |>
+#   mutate(across(c("Estimate", " 95% CrI Lower", " 95% CrI Upper"), 
+#                 \(x) plogis(x))) |>
+#   mutate(across(where(is.numeric), \(x) round(x, 3))) |>
+#   tibble::remove_rownames() |>
+#   mutate(Term = stringr::str_remove(Term, "b_")) |>
+#   kable(booktabs = TRUE)
 
  mted2 |>
-  mutate(across(c("Estimate", " 95% CrI Lower", " 95% CrI Upper"), 
-                \(x) plogis(x))) |>
+  mutate(Term = stringr::str_remove(Term, "b_")) |>
+  mutate(across(c("Estimate", "95% CrI Lower", "95% CrI Upper"), 
+                \(x) if_else(Term == "Intercept", plogis(x), plogis(x + intercept)))) |>
   mutate(across(where(is.numeric), \(x) round(x, 3))) |>
+  mutate(Term = if_else(Term == "Intercept", "Intercept (kWh)", Term)) |>
   tibble::remove_rownames() |>
-  mutate(Term = stringr::str_remove(Term, "b_")) |>
   kable(booktabs = TRUE)
+
 ```
@@
 
+s2_bb2_r_g |> emmeans( ~refClass, type="response") |>
+  gather_emmeans_draws() |>
+ condEffects(refClass) + labs(y="Absolute Deviation From Band", x="Band Type")
+
 s2_bb2_r_g |> emmeans(pairwise ~ refClass * rounded, type="response") |> 

### 2024-10-31 study1.qmd
outlier checks and some model updates

@@
       values_to = "value"
-    ) |> select(id,state1,state2,refClass,calc,edu,modality,state,item_num,value,id2)
+    ) |> select(id,state1,state2,refClass,calc,edu,modality,state,item_num,value,id2,Duration__in_seconds_)
 
@@
             state_f_dif=mean(state_f_dif),
-            n_less_avg = sum(less_avg)) |> 
-  mutate(matched_goal = (pct_change == pct_goal))
+            n_less_avg = sum(less_avg),
+            duration=first(Duration__in_seconds_)) |> 
+  mutate(matched_goal = (pct_change == pct_goal), 
+             error = pct_change - pct_goal,
+            abs_error = abs(error),
+            close_match = abs_error <= 0.02)
 
-k <- s1_agg |> group_by(id,pct_goal) |>  summarize(total_kWh=sum(total_kWh),orig_kWh=sum(orig_kWh),pct_change=round((orig_kWh-total_kWh)/orig_kWh,3)) |> 
-  mutate(matched_goal = (pct_change == pct_goal))
+# s1_agg |> filter(matched_goal==FALSE, abs_error<.02) |> pull(abs_error)
 
@@
 
+ggplot(s1_agg |> filter(abs_error<1), aes(x = refClass, y = abs_error, fill = refClass)) +
+    geom_boxplot() +
+    labs(
+        title = "Absolute Error by Reference Class",
+        x = "Reference Class",
+        y = "Absolute Error"
+    ) +
+    theme_minimal()
+
+ggplot(s1_agg |> filter(abs_error<1), aes(x = abs_error, fill = refClass)) +
+    geom_histogram(binwidth = 0.05, alpha = 0.6, position = "identity") +
+    labs(
+        title = "Distribution of Absolute Error by Reference Class",
+        x = "Absolute Error",
+        y = "Frequency"
+    ) +
+    theme_minimal() + facet_wrap(~refClass)
 ```
@@
     matched_count = sum(matched_goal),
+    matched_count2 = sum(close_match),
     total_count = n(),
@@
     title = "Percentage of Participants Matching Target Pct by Condition",
+    x = "Reference Class",
+    y = "Percentage Matched (%)",
+    fill = "Rounding"
+  ) +
+  theme_minimal()
+
+ggplot(summary_pct, aes(x = refClass, y = matched_count2, fill = refClass)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  labs(
+    title = "Percentage of Participants Matching Target Pct by Condition - 2% leeway",
     x = "Reference Class",

### 2024-10-31 Scripts/fit_models.R

@@ -18,8 +18,18 @@ s1_agg <- s1 |>
   filter(appliance !="Total kWh") |> 
   group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
   summarise(total_kWh = sum(value),orig_kWh=sum(family), 
-            pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), state_dif=mean(state_dif)) |> 
-  mutate(matched_goal = (pct_change == pct_goal))
+            pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
+            n_change = sum(value!=family),
+            state_p_dif=mean(state_p_dif),
+            state_f_dif=mean(state_f_dif),
+            n_less_avg = sum(less_avg)) |> 
+  mutate(matched_goal = (pct_change == pct_goal), 
+             error = pct_change - pct_goal,
+            abs_error = abs(error),
+            close_match = abs_error <= 0.02)
+
+
+
 
 s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
   summarise(mg=sum(matched_goal),n=n(), pct=mg/n) 
@@ -43,6 +53,72 @@ s1_bb1 <- brm(
 # Further Distributional Parameters:
 #   Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
 # phi     0.33      0.00     0.32     0.34 1.00     9328     7470
+
+
+pp_check(s1_bb1, type = "stat_grouped", stat = "mean", group = "refClass")
+
+
+
+
+s1_bb1 %>%
+  gather_draws(b_Intercept, b_refClassPercentage, b_refClassUSD) %>%
+  ggplot(aes(x = .value, fill = .variable)) +
+  geom_density(alpha = 0.6) +
+  theme_minimal() +
+  labs(title = "Posterior Distributions of Fixed Effects for s1_bb1",
+       x = "Coefficient Value",
+       fill = "Reference Class")
+
+# Visualize fixed effects with credible intervals
+s1_bb1 %>%
+  gather_draws(b_Intercept, b_refClassPercentage, b_refClassUSD) %>%
+  ggplot(aes(x = .value, y = .variable)) +
+  stat_halfeye() +
+  theme_minimal() +
+  labs(title = "Posterior Estimates of Log-Odds by Reference Class (Model s1_bb1)",
+       x = "Log-Odds Estimate",
+       y = "Reference Class")
+
+s1_bb1 |> emmeans(~refClass, type="response")
[…]

### 2024-10-31 misc/outlier.qmd
fix

@@
 
+## Inspects counts for each distinct value of prior year family use, and state average
+
 ```{r}
@@
   flextable() %>%
+  add_header_lines(values = "Family prior year use per utility counts (green=most frequent)") %>%
+  bold(part = "header", i = 1) %>%
   # Merge cells
@@
 
+library(dplyr)
+library(flextable)
+library(officer)
+
 s1 %>%
@@
   flextable() %>%
+  # Add title
+  add_header_lines(values = "State Avg. per utility counts (green=most frequent)") %>%
+  bold(part = "header", i = 1) %>%
   # Merge cells

### 2024-10-31 misc/indiv.qmd
indiv patterns page

---
title: "Response Patterns"
author:
- name: Thomas E. Gorman
  url: https://tegorman13.github.io/
  affiliations: 
  - name:  Communication and Cognition Lab, Purdue University, USA
    affiliation-url: https://web.ics.purdue.edu/~treimer/
lightbox: true
toc: false
code-fold: true
code-tools: true
execute:
  #echo: false
  warning: false
format:
  html: 
    grid:
      sidebar-width: 220px
      body-width: 1200px
      margin-width: 170px
      gutter-width: 1.0rem
  # hugo-md:
  #   include: true
  #   html-math-method: mathjax
  #   output-file: results_hugo.md
  # gfm:
  #   echo: true
  #   output-file: results_gfm.md
---

```{r}
library(pacman)
pacman::p_load(dplyr,purrr,tidyr,here, haven,tibble,ggplot2,gt,ggh4x,lme4,flextable,ggdist)

s1 <- readRDS(here::here("data/s1_processed.rds"))

s2_long <- readRDS(here::here("data/s2_processed.rds"))

s1_agg <- s1 |> 
  filter(appliance !="Total kWh") |> 
  group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
  summarise(total_kWh = sum(value),orig_kWh=sum(family), 
            pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
            n_change = sum(value!=family),
            state_p_dif=mean(state_p_dif),
            state_f_dif=mean(state_f_dif),
            n_less_avg = sum(less_avg),
            duration=first(Duration__in_seconds_)) |> 
  mutate(matched_goal = (pct_change == pct_goal), 
             error = pct_change - pct_goal,
            abs_error = abs(error))
```

## Study 1 - comparison against state average
```{r}

s2_long |> filter(id %in% unique(s2_long$id)[1:30]) |> 
  filter(appliance!="TOTAL",state=="California") |> 
  ggplot(aes(x=appliance,y=value)) + geom_point(aes(shape=plan)) +
  geom_point(aes(y=state_avg),color="red") +
  geom_point(aes(y=family),color="blue") +
  facet_wrap(~id) + theme(axis.text.x = element_text(angle = 45, hjust = 1))

s2_long |> filter(id %in% unique(s2_long$id)[1:30]) |> 
  filter(appliance!="TOTAL",state=="California") |> 
  ggplot(aes(x=appliance,y=value)) + geom_point(aes(shape=plan)) +
  geom_point(aes(y=state_avg),color="red") +
  geom_point(aes(y=family),color="blue") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

s1 |> 
filter(appliance !="Total kWh") |> 
group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
mutate(total_kWh = sum(value),orig_kWh=sum(family), 
        pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
        n_change = sum(value!=family),
        state_p_dif=mean(state_p_dif),
        state_f_dif=mean(state_f_dif),
        n_less_avg = sum(less_avg),
        duration=first(Duration__in_seconds_)) |> 
mutate(matched_goal = (pct_change == pct_goal), 
            error = pct_change - pct_goal,
        abs_error = abs(error)) |> 
          filter(appliance!="TOTAL",state=="CAL",value<20000) |> 
  ggplot(aes(x=appliance,y=value)) + geom_point(aes(shape=matched_goal)) +
  geom_point(aes(y=state_avg),color="red") +
  geom_point(aes(y=family),color="blue") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

s1 |> 
filter(appliance !="Total kWh") |> 
group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
mutate(total_kWh = sum(value),orig_kWh=sum(family), 
        pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
        n_change = sum(value!=family),
        state_p_dif=mean(state_p_dif),
        state_f_dif=mean(state_f_dif),
        n_less_avg = sum(less_avg),
        duration=first(Duration__in_seconds_)) |> 
mutate(matched_goal = (pct_change == pct_goal), 
            error = pct_change - pct_goal,
        abs_error = abs(error)) |> 
          filter(appliance!="TOTAL",state=="CAL",value<20000) |> 
  ggplot(aes(x=appliance,y=value)) + geom_point(aes(shape=matched_goal)) +
  geom_point(aes(y=state_avg),color="red") +
  geom_point(aes(y=family),color="blue") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(~matched_goal)

s1 |> 
filter(appliance !="Total kWh") |> 
group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
mutate(total_kWh = sum(value),orig_kWh=sum(family), 
        pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
        n_change = sum(value!=family),
        state_p_dif=mean(state_p_dif),
        state_f_dif=mean(state_f_dif),
        n_less_avg = sum(less_avg),
        duration=first(Duration__in_seconds_)) |> 
mutate(matched_goal = (pct_change == pct_goal), 
            error = pct_change - pct_goal,
        abs_error = abs(error)) |> 
          filter(appliance!="TOTAL",value<20000) |> 
  ggplot(aes(x=appliance,y=value)) + geom_point(aes(shape=matched_goal)) +
  geom_point(aes(y=state_avg),color="red") +
  geom_point(aes(y=family),color="blue") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(state~matched_goal,scales="free_y")
```

## Study 2 - comparison against state average

```{r}

s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id, refClass, appliance, state, pct_goal, pct, rounded, block, plan, calc, edu) |> 
  mutate(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(matched_goal = (pct_change == pct),
  matched_goal2 = abs(pct_change-pct)<.03) |> 
    ggplot(aes(x=appliance,y=value)) + geom_point(aes(shape=matched_goal)) +
    geom_point(aes(y=state_avg),color="red") +
    geom_point(aes(y=family),color="blue") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    facet_wrap(state~matched_goal,scales="free_y")

s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id, refClass, appliance, state, pct_goal, pct, rounded, block, plan, calc, edu) |> 
  mutate(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(matched_goal = (pct_change == pct),
  matched_goal2 = abs(pct_change-pct)<.03) |> 
    ggplot(aes(x=appliance,y=value)) + geom_point(aes(shape=matched_goal)) +
    geom_point(aes(y=state_avg),color="red") +
    geom_point(aes(y=family),color="blue") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    facet_wrap(refClass~state,scales="free_y") +
    labs(title="Study 2 - plans vs. state averages and prior year values",subtitle="Blue points are prior year values, red points are state averages, and black points are current year values")

library(ggplot2)
library(dplyr)
s2_long %>%
  filter(appliance != "TOTAL") %>%
  group_by(id, refClass, appliance, state, pct_goal, pct, rounded, block, plan, calc, edu) %>%
  mutate(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) %>%
  mutate(
    matched_goal = (pct_change == pct),
    matched_goal2 = abs(pct_change - pct) < 0.03
  ) %>%
  ggplot(aes(x = appliance)) +
  # Points for current plan, with conditional color and shape for matched goal
  geom_point(aes(
    y = value,
    color = ifelse(matched_goal2, "Matched Goal", "Current Plan"),
    shape = ifelse(matched_goal2, "Matched Goal", "Current Plan"),
    size = ifelse(matched_goal2, 5, 2),  # Larger size for matched goal
    alpha = ifelse(matched_goal2, 1, 0.6),  # Less transparency for matched goal
    stroke = ifelse(matched_goal2, 1.5, 0)  # Outline for matched goal points
  )) +
  # Points for state average (red) with color mapped for legend
  geom_point(aes(y = state_avg, color = "State Average"), shape = 17, size = 3) +
  # Points for family prior year (blue) with color mapped for legend
  geom_point(aes(y = family, color = "Prior Year (Family)"), shape = 15, size = 3) +
  # Custom colors for each legend entry
  scale_color_manual(
    values = c(
      "Matched Goal" = "#00FF00",  # Bright green for matched goal
      "Current Plan" = "black",
      "State Average" = "red",
      "Prior Year (Family)" = "blue"
    ),
    name = "Legend"
  ) +
  scale_shape_manual(
    values = c(
      "Matched Goal" = 8,  # Star shape for matched goal
      "Current Plan" = 16  # Default shape for current plan
    )
  ) +
  scale_size_continuous(range = c(2, 1)) +  # Use continuous scale for size
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  facet_wrap(refClass ~ state, scales = "free_y") +
  labs(
    title = "Study 2 - Plans vs. State Averages and Prior Year Values",
    subtitle = "Blue points are prior year values, red points are state averages, black points are current plans, bright green star points indicate plans that match goal"
  )

```

```{r}

s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id, refClass, appliance, state, pct_goal, pct, rounded, block, plan, calc, edu) |> 
  mutate(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(
    matched_goal = (pct_change == pct),
    matched_goal2 = abs(pct_change-pct) < .03
  ) |> 
  ggplot(aes(x = appliance)) + 
    # Non-matching plans (plotted first, in background)
    geom_point(data = . %>% filter(!matched_goal2),
               aes(y = value, color = "Plan (Not Matching)", shape = "Plan"), 
               size = 2, alpha = 0.4) +
    # Matching plans
    geom_point(data = . %>% filter(matched_goal2),
               aes(y = value, color = "Plan (Matching Goal)", shape = "Plan"), 
               size = 2.5) +
    # Prior year values
    geom_point(aes(y = family, color = "Prior Year", shape = "Prior Year"), 
               size = 3.5) +
    # State averages with star shape, plotted last and largest
    geom_point(aes(y = state_avg, color = "State Average", shape = "State Average"), 
               size = 3, stroke = 2) +
    scale_color_manual(
      name = "Values",
      values = c(
        "State Average" = "#FF0000",     # Bright red
        "Prior Year" = "#4169E1",        # Royal blue
        "Plan (Matching Goal)" = "#228B22",
        "Plan (Not Matching)" = "#CD5C5C"
      )
    ) +
    scale_shape_manual(
      name = "Type",
      values = c(
        "State Average" = 8,    # Star
        "Prior Year" = 17,      # Triangle
        "Plan" = 16            # Circle
      )
    ) +
    theme_minimal() +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      panel.grid.minor = element_blank(),
      strip.background = element_rect(fill = "lightgray", color = NA),
      strip.text = element_text(face = "bold"),
      legend.position = "bottom",
      legend.title = element_text(face = "bold")
    ) +
    facet_wrap(refClass ~ state, scales = "free_y") +
    labs(
      title = "Study 2 - Plans vs. State Averages and Prior Year Values",
      x = "Appliance",
      y = "Value"
    )
```

### 2024-11-04 misc/discrep.qmd
s1 discrepancy checks; index page reset

---
title: "Discrep_Patterns"
author:
- name: Thomas E. Gorman
  url: https://tegorman13.github.io/
  affiliations: 
  - name:  Communication and Cognition Lab, Purdue University, USA
    affiliation-url: https://web.ics.purdue.edu/~treimer/
lightbox: true
toc: false
code-fold: show
code-tools: true
execute:
  warning: false
---

## Study 1

```{r}
library(pacman)
pacman::p_load(dplyr,purrr,tidyr,here,tibble, haven,
               ggplot2,gt,knitr,kableExtra,ggh4x,lme4,flextable)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("fun_plot"), ~ source(here::here(paste0("Scripts/", .x, ".R"))))
theme_set(theme_nice())

s1 <- readRDS(here::here("data/s1_processed.rds"))
s2_long <- readRDS(here::here("data/s2_processed.rds"))

```

```{r}

s1 %>%
  filter(appliance != "Total kWh", plan == "plan1") %>%
  group_by(appliance, state, refClass) %>%
  mutate(distinct_state_avg = n_distinct(state_avg)) %>%
  ungroup() %>%
  filter(distinct_state_avg > 1) %>%
  group_by(appliance, state_avg, state, refClass) |> arrange(id)

s1 |> group_by(state,appliance,refClass) |> 
  mutate(state_avg=as.numeric(state_avg)) |> 
  mutate(n_distinct_family=n_distinct(family),
            n_distinct_state_avg=n_distinct(state_avg), 
            # paste a string of each distinct value, separated by a comma
            distinct_family=paste(unique(family),collapse=","),
            distinct_state_avg=paste(unique(state_avg),collapse=",")) |>
  filter(n_distinct_family>1 | n_distinct_state_avg>1)

s1 %>%
  filter(appliance != "Total kWh", plan == "plan1") %>%
  group_by(appliance, state, refClass) %>%
  mutate(
    distinct_families = n_distinct(family),
    distinct_state_avgs = n_distinct(state_avg),
  ) %>%
  ungroup() %>%
  filter(distinct_families > 1 | distinct_state_avgs>1) %>%
  group_by(appliance, family, state, refClass) %>%
  summarise(n = n(), .groups = "drop") |> gt()

```

```{r}

state_code_map <- c(
  "CAL" = "California",
  "TEX" = "Texas",
  "COL" = "Colorado",
  "MASS" = "Massachusetts"
)

state_to_code <- state_code_map
code_to_state <- setNames(names(state_code_map), state_code_map)
lookup_code <- function(state) {
  if (state %in% names(state_to_code)) {
    return(state_to_code[[state]])
  } else if (state %in% names(code_to_state)) {
    return(code_to_state[[state]])
  } else {
    return(NA)  # Return NA if not found in either mapping
  }
}

correct_values <- tribble(
  ~state, ~appliance, ~family, ~state_avg,
  # Texas values
  "Texas", "Cooling", 6573, 4249,
  "Texas", "Heating", 6118, 5099,
  "Texas", "Water Heating", 5257, 4396,
  "Texas", "Refrigerator", 2639, 1318,
  "Texas", "Other Appliance", 8459, 7883,
  "Texas", "Total kWh", 29046, 22945,

  # California values
  "California", "Cooling", 2581, 1289,
  "California", "Heating", 6157, 5597,
  "California", "Water Heating", 5061, 4601,
  "California", "Refrigerator", 1266, 1055,
  "California", "Other Appliance", 7608, 6916,
  "California", "Total kWh", 22673, 19458,

  # Colorado values
  "Colorado", "Cooling", 697, 498,
  "Colorado", "Heating", 18052, 16411,
  "Colorado", "Water Heating", 11667, 5832,
  "Colorado", "Refrigerator", 1370, 1142,
  "Colorado", "Other Appliance", 7982, 6652,
  "Colorado", "Total kWh", 39768, 30535,

  # Massachusetts values
  "Massachusetts", "Cooling", 419, 322,
  "Massachusetts", "Heating", 26751, 19108,
  "Massachusetts", "Water Heating", 10543, 5070,
  "Massachusetts", "Refrigerator", 1230, 1025,
  "Massachusetts", "Other Appliance", 7350, 6682,
  "Massachusetts", "Total kWh", 46293, 32207
) |> 
  rowwise() |>
  mutate(state_abrv=lookup_code(state))

 s1 |> 
       filter(#appliance != "Total kWh", 
       plan == "plan1") %>%
        group_by(state, appliance, refClass) |> 
        mutate(n_distinct_family = n_distinct(family),
               n_distinct_state_avg = n_distinct(state_avg), 
               distinct_family = paste(unique(family), collapse = ","),
               distinct_state_avg = paste(unique(state_avg), collapse = ",")) |>
        filter(n_distinct_family > 1 | n_distinct_state_avg > 1) |> 
        group_by(state, appliance, refClass, family, state_avg) |> 
        mutate(state_avg_count = n_distinct(state_avg),
               family_count = n()) |> 
        ungroup() |> 
        group_by(state, appliance, refClass, state_avg) |> 
        mutate(state_avg_count = n()) |> 
        distinct(state, appliance, refClass, family, state_avg, family_count, state_avg_count) |> 
      arrange(state,appliance,refClass,-family_count) |> 
      rowwise() |> 
      mutate(fam_in = family %in% correct_values$family,
             state_in = state_avg %in% correct_values$state_avg) |>
       kable() # print(n=25)

# Function to check values
check_values <- function(data, reference) {
  # Join the datasets
  data %>%
    left_join(reference |> rename(family_true=family,state_true=state_avg), by = c("state" = "state_abrv", "appliance" = "appliance")) %>%
    mutate(
      # Convert to numeric to ensure proper comparison
      family_numeric = as.numeric(family),
      state_avg_numeric = as.numeric(state_avg),
      # Flag discrepancies (using a small tolerance for floating point differences)
      family_discrepant =family_numeric != family_true,
      state_avg_discrepant = state_avg_numeric != state_true
    )
}

discrepant_counts <- s1 %>%
  # Remove total kWh as specified
  #filter(appliance != "Total kWh") %>%
  # Use only one plan to avoid duplicates
  filter(plan == "plan1") %>%
  # Check against correct values
  check_values(correct_values) %>%
  select(id,id2,starts_with("sta"),starts_with("fam"),appliance,refClass) |> 
  # Filter to show only discrepant cases
  filter(family_discrepant | state_avg_discrepant) %>%
  # Get unique subject IDs and their discrepancies
  group_by(refClass,state) %>%
  summarise(
    discrepant_states = paste(unique(state), collapse = ", "),
    discrepant_appliances = paste(unique(appliance), collapse = ", "),
    n_family_discrepancies = sum(family_discrepant),
    n_state_avg_discrepancies = sum(state_avg_discrepant),
    .groups = 'drop'
  )

# Identify discrepant subjects
discrepant_subjects <- s1 %>%
  # Remove total kWh as specified
  #filter(appliance != "Total kWh") %>%
  # Use only one plan to avoid duplicates
  filter(plan == "plan1") %>%
  # Check against correct values
  check_values(correct_values) %>%
  select(id,id2,refClass,starts_with("sta"),starts_with("fam"),appliance) |> 
  # Filter to show only discrepant cases
  filter(family_discrepant | state_avg_discrepant) %>%
  # Get unique subject IDs and their discrepancies
  group_by(id, id2,refClass) %>%
  summarise(
    discrepant_states = paste(unique(state), collapse = ", "),
    discrepant_appliances = paste(unique(appliance), collapse = ", "),
    n_family_discrepancies = sum(family_discrepant),
    n_state_avg_discrepancies = sum(state_avg_discrepant),
    .groups = 'drop'
  )

discrepant_counts |> gt()
discrepant_subjects |> gt()

s1_discrep_ids <- discrepant_subjects |> pull(id) |> unique()

s1_discrep_ids

# for every combination of state, appliance, and refClass, calculate the number of distinct values of family, and of state_avg
s1 |> group_by(state,appliance,refClass) |> 
  filter(appliance != "Total kWh") %>%
  filter(!(id %in% s1_discrep_ids)) |>
  mutate(state_avg=as.numeric(state_avg)) |> 
  summarise(n_distinct_family=n_distinct(family),
            n_distinct_state_avg=n_distinct(state_avg), 
            # paste a string of each distinct value, separated by a comma
            distinct_family=paste(unique(family),collapse=","),
            distinct_state_avg=paste(unique(state_avg),collapse=",")) |>
  filter(n_distinct_family>1 | n_distinct_state_avg>1)

# <!-- issue in slides where colorado cooling is shown as 489 in a few cases instead of 498 - all of the data has 498. 498 corresponds with total value as well  -->
s1 |> filter(state=="COL",appliance=="Cooling") |> pull(state_avg) |> unique() #498
s2_long |> filter(state=="Colorado",appliance=="AC") |> pull(state_avg) |> unique() #498

s1 |> select(id,id2,starts_with("sta"),starts_with("fam"),appliance,refClass,plan,value) |> check_values(correct_values) %>% 
filter(plan=="plan1") |> 
select(-state_f_dif,-state_p_dif,-plan,-state.y,-state_avg) |> ungroup() |> as_tibble() |> 
relocate(id,state, appliance,value,family_numeric,family_true,state_avg_numeric,state_true, family_discrepant,state_avg_discrepant) |>
# mutate to turn applilance, state and id columns into factors
mutate(across(c(appliance,state,id),factor)) |>
#reactable::reactable()
  DT::datatable(extensions = 'Buttons',
    options = list(
       dom = 'Blfrtip',
       buttons = list('copy', 'csv', 'excel', 'pdf', 'print'),
      pageLength = 25,
      autoWidth = TRUE
    ),
    filter = 'top')

```

```{r}

haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav")) |> select(P3TEX15) |> unique()

s1 |> filter(id==18,plan=="plan1")

haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav")) |>
  filter(ResponseId == "R_vB56XspT1LqhrkB") |>
  # Select columns with non-NA values and not empty strings, handling character and numeric columns separately
  select(
    where(~ is.numeric(.) && !all(is.na(.))),
    where(~ is.character(.) && !all(is.na(.)) && !all(. == ""))
  ) |>
  # Remove columns that start with "AIN"
  select(-starts_with("AIN"))

haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav"))  |> select_if(~any(. ==5124)) |> 
  # filter out any NAs from any column that is left (we don't know the column names)
  filter_all(any_vars(!is.na(.)))

```

```{r}

s2_long %>%
  filter(appliance != "TOTAL", plan == "plan1") %>%
  group_by(appliance, rounded, state) %>%
  mutate(
    distinct_families = n_distinct(family),
    color_flag = distinct_families > 1  # Flag as TRUE if more than 1 distinct value
  ) %>%
  ungroup() %>%
  # Continue grouping for plotting
  group_by(appliance, family, rounded, state,color_flag) %>%
  summarise(n = n(), .groups = "drop")

```

### 2024-11-04 misc/indiv.qmd
update
s1 discrepancy checks; index page reset

@@
         abs_error = abs(error)) |> 
-          filter(appliance!="TOTAL",value<20000) |> 
-  ggplot(aes(x=appliance,y=value)) + geom_point(aes(shape=matched_goal)) +
-  geom_point(aes(y=state_avg),color="red") +
-  geom_point(aes(y=family),color="blue") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  facet_wrap(state~matched_goal,scales="free_y")
+          filter(appliance!="TOTAL",state=="CAL",value<20000) |> 
+  ggplot(aes(x=appliance,y=value)) + 
+  geom_jitter(aes(color=matched_goal)) +
+geom_point(aes(y=state_avg),color="red",size=10) +
+  geom_point(aes(y=family),color="blue",size=5)
+
+s1 |> 
+filter(appliance !="Total kWh") |> 
+group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
+mutate(total_kWh = sum(value),orig_kWh=sum(family), 
+        pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
+        n_change = sum(value!=family),
+        state_p_dif=mean(state_p_dif),
+        state_f_dif=mean(state_f_dif),
+        n_less_avg = sum(less_avg),
+        duration=first(Duration__in_seconds_)) |> 
+mutate(matched_goal = (pct_change == pct_goal), 
+            error = pct_change - pct_goal,
+        abs_error = abs(error)) |> 
+          filter(appliance!="TOTAL",value<20000,family<25000) |> 
+  ggplot(aes(x=appliance,y=value)) + 
+  geom_jitter(aes(color=matched_goal)) +
+ # gghalves::geom_half_violin(aes(fill=matched_goal)) +
+  geom_boxplot(aes(fill=matched_goal),alpha=.5) +
+geom_point(aes(y=state_avg),color="red",size=7,shape=8) +
+  geom_point(aes(y=family),color="blue",size=7,shape=17) +
+  facet_wrap(~state)
+
+s1 |> 
+filter(appliance !="Total kWh") |> 
+group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
+mutate(total_kWh = sum(value),orig_kWh=sum(family), 
+        pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
+        n_change = sum(value!=family),
+        state_p_dif=mean(state_p_dif),
+        state_f_dif=mean(state_f_dif),
+        n_less_avg = sum(less_avg),
+        duration=first(Duration__in_seconds_)) |> 
+mutate(matched_goal = (pct_change == pct_goal), 
+            error = pct_change - pct_goal,
+        abs_error = abs(error)) |> 
+          filter(appliance!="TOTAL",value<20000,family<25000) |> 
+  ggplot(aes(x=appliance,y=state_p_dif)) + 
+  geom_jitter(aes(color=matched_goal)) +
+  geom_point(aes(y=state_avg),color="red",size=10) +
+  geom_point(aes(y=family),color="blue",size=5)
+
+s1 |> 
+filter(appliance !="Total kWh") |> 
+group_by(id,refClass,state,block,plan,calc,edu,pct_goal,appliance) |> 
+summarize(total_kWh = sum(value),orig_kWh=sum(family), 
+            adjust=orig_kWh-total_kWh,
+        pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
+        n_change = sum(value!=family),
+        state_p_dif=mean(state_p_dif),
+        state_f_dif=mean(state_f_dif),
+        n_less_avg = sum(less_avg),
+        duration=first(Duration__in_seconds_)) |> 
+mutate(matched_goal = (pct_change == pct_goal), 
+            error = pct_change - pct_goal,
+        abs_error = abs(error)) |> 
+          filter(appliance!="TOTAL") |> 
+  ggplot(aes(x=appliance,y=adjust)) + 
+  geom_jitter(aes(color=matched_goal)) +
+  facet_wrap(~state)
+
 ```

### 2024-11-04 misc/outlier.qmd
indiv patterns page
s1 discrepancy checks; index page reset

@@
 2    16 Texas  Massachusetts USD      R_3gNKA8IQr0uOSh3 TEX   Water Heating  34265 4396      plan2  6414     1     0.15
+
+ s1 |> filter(state_avg==29046)
+# A tibble: 2 × 21
+# Groups:   id [1]
+     id state1   state2 refClass calc  edu   modality state id2   Duration__in_seconds_ appliance family state_avg plan  value block
+  <int> <chr>    <chr>  <chr>    <chr> <fct> <chr>    <chr> <chr>                 <dbl> <chr>      <dbl> <dbl+lbl> <fct> <dbl> <dbl>
+1    12 Colorado Texas  kWh      Used… Grad… A        TEX   R_ab…                   644 Total kWh  29046 29046     plan1 29046     2
+2    12 Colorado Texas  kWh      Used… Grad… A        TEX   R_ab…                   644 Total kWh  29046 29046     plan2 29046     2
+ℹ 5 more variables: pct_goal <dbl>, change <dbl>, state_p_dif <dbl>, state_f_dif <dbl>, less_avg <lgl>
 
@@
 
+state_code_map <- c(
+  "CAL" = "California",
+  "TEX" = "Texas",
+  "COL" = "Colorado",
+  "MASS" = "Massachusetts"
+)
+
 # Join participant data with correct values

### 2024-11-04 study1.qmd
indiv patterns page
s1 discrepancy checks; index page reset

@@
            distinct_state_avg = paste(unique(state_avg), collapse = ",")) |>
-    filter(n_distinct_family > 1 | n_distinct_state_avg > 1) |> 
+    filter(n_distinct_family > 1 | n_distinct_state_avg > 1,plan=="plan1") |> 
     group_by(state, appliance, refClass, family, state_avg) |> 
@@
 # find if any columns equal 5.26
+
+haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav")) |>
+  filter(ResponseId == "R_vB56XspT1LqhrkB") |>
+  # Select columns with non-NA values and not empty strings, handling character and numeric columns separately
+  select(
+    where(~ is.numeric(.) && !all(is.na(.))),
+    where(~ is.character(.) && !all(is.na(.)) && !all(. == ""))
+  ) |>
+  # Remove columns that start with "AIN"
+  select(-starts_with("AIN"))
 

### 2024-11-04 study2.qmd
outlier checks and some model updates
indiv patterns page
s1 discrepancy checks; index page reset

@@
 library(pacman)
-pacman::p_load(dplyr,purrr,tidyr,here, haven,tibble,ggplot2,gt,ggh4x,lme4,flextable)
+pacman::p_load(dplyr,purrr,tidyr,here, haven,tibble,ggplot2,gt,ggh4x,lme4,flextable,ggh4x)
 
@@
 
-s2 |> 
-  mutate(
-    rounded = if_else(ROUNDED == 1, "Not Rounded", "Rounded"),
-    calc = if_else(calc == 0, "No Calculator", "Used Calculator")
-  ) |> 
-  count(calc, refClass, rounded) |>
-  pivot_wider(
-    names_from = c(calc, rounded),
-    values_from = n,
-    values_fill = 0
-  ) |>
-  gt() |>
-  tab_header(
-    title = md("**Count of Items by Calculator Use, Reference Class, and Rounding**")
-  ) |>
-  cols_align(align = "left") |>
-  tab_spanner_delim(delim = "_") |>
-  cols_label(
-    `No Calculator_Not Rounded` = "Not Rounded",
-    `No Calculator_Rounded` = "Rounded",
-    `Used Calculator_Not Rounded` = "Not Rounded",
-    `Used Calculator_Rounded` = "Rounded"
-  )
+# s2 |> 
+#   mutate(
+#     rounded = if_else(ROUNDED == 1, "Not Rounded", "Rounded"),
+#     calc = if_else(calc == 0, "No Calculator", "Used Calculator")
+#   ) |> 
+#   count(calc, refClass, rounded) |>
+#   pivot_wider(
+#     names_from = c(calc, rounded),
+#     values_from = n,
+#     values_fill = 0
+#   ) |>
+#   gt() |>
+#   tab_header(
+#     title = md("**Count of Items by Calculator Use, Reference Class, and Rounding**")
+#   ) |>
+#   cols_align(align = "left") |>
+#   tab_spanner_delim(delim = "_") |>
+#   cols_label(
+#     `No Calculator_Not Rounded` = "Not Rounded",
+#     `No Calculator_Rounded` = "Rounded",
+#     `Used Calculator_Not Rounded` = "Not Rounded",
+#     `Used Calculator_Rounded` = "Rounded"
+#   )
 

### 2024-11-04 misc/discrep.qmd
discrep update

@@
 
+correct_values_s2 <- tribble(
+  ~state, ~appliance, ~family_exact, ~family_rounded, ~state_avg,
+  # Texas values (Smith Family)
+  "Texas", "Cooling", 6573, 7000, 4249,
+  "Texas", "Heating", 6118, 6000, 5099,
+  "Texas", "Water Heating", 5257, 5000, 4396,
+  "Texas", "Refrigerator", 2639, 3000, 1318,
+  "Texas", "Other Appliance", 8459, 8000, 7883,
+  "Texas", "Total kWh", 29046, 29000, 22945,
+
+  # California values (Adams Family)
+  "California", "Cooling", 2581, 3000, 1289,
+  "California", "Heating", 6157, 6000, 5597,
+  "California", "Water Heating", 5061, 5000, 4601,
+  "California", "Refrigerator", 1266, 1000, 1055,
+  "California", "Other Appliance", 7608, 8000, 6916,
+  "California", "Total kWh", 22673, 23000, 19458,
+
+  # Colorado values (Wells Family)
+  "Colorado", "Cooling", 697, 1000, 498,
+  "Colorado", "Heating", 18052, 18000, 16411,
+  "Colorado", "Water Heating", 11667, 12000, 5832,
+  "Colorado", "Refrigerator", 1370, 1000, 1142,
+  "Colorado", "Other Appliance", 7982, 8000, 6652,
+  "Colorado", "Total kWh", 39768, 40000, 30535,
+
+  # Massachusetts values (Davis Family)
+  "Massachusetts", "Cooling", 419, 0, 322,
+  "Massachusetts", "Heating", 26751, 27000, 19108,
+  "Massachusetts", "Water Heating", 10543, 11000, 5070,
+  "Massachusetts", "Refrigerator", 1230, 1000, 1025,
+  "Massachusetts", "Other Appliance", 7350, 7000, 6682,
+  "Massachusetts", "Total kWh", 46293, 46000, 32207
+) %>%
+  rowwise() %>%
+  mutate(
+    state_abrv = case_when(
+      state == "Texas" ~ "TEX",
+      state == "California" ~ "CAL",
+      state == "Colorado" ~ "COL",
+      state == "Massachusetts" ~ "MASS"
+    ),
+    family_name = case_when(
+      state == "Texas" ~ "Smith Family",
+      state == "California" ~ "Adams Family",
+      state == "Colorado" ~ "Wells Family",
+      state == "Massachusetts" ~ "Davis Family"
+    ))
+
 s2_long %>%
-  filter(appliance != "TOTAL", plan == "plan1") %>%
-  group_by(appliance, rounded, state) %>%
+  filter( plan == "plan1") %>%
+  group_by(appliance, rounded, state,pct_goal) %>%
   mutate(
     distinct_families = n_distinct(family),
-    color_flag = distinct_families > 1  # Flag as TRUE if more than 1 distinct value
+    distinct_state_avg = n_distinct(state_avg),
   ) %>% 
-  ungroup() %>%
-  # Continue grouping for plotting
-  group_by(appliance, family, rounded, state,color_flag) %>%
-  summarise(n = n(), .groups = "drop")
+  filter(distinct_families > 1 | distinct_state_avg > 1)  |> 
+   group_by(appliance, rounded, state,pct_goal, family) |>
+   mutate(family_count = n()) |>
+   group_by(appliance, rounded, state,pct_goal, state_avg) |>
+    mutate(state_avg_count = n()) |>
+  distinct(rounded,state,refClass,appliance,family,state_avg,state_avg_count,family_count) |> 
+     rowwise() |> 
+  mutate(
+    fam_in = if (rounded == "Rounded") {
+      family %in% correct_values_s2$family_rounded
+    } else {
+      family %in% correct_values_s2$family_exact
+    },
+    state_in = state_avg %in% correct_values_s2$state_avg
+  ) |> arrange(rounded,refClass,state,appliance,family_count) |> 
+  relocate(rounded,pct_goal,refClass,state,appliance,family,family_count,state_avg,state_avg_count,fam_in,state_in) |> 
+  kable()
+
+  # ungroup() %>%
+  # group_by(appliance, family, rounded, state) %>%
+  # summarise(n = n(), .groups = "drop")
 

### 2024-11-04 misc/discrep.qmd
compute study 2 discrepancy counts

@@
 s1_discrep_ids
+#saveRDS(s1_discrep_ids,here::here("data/s1_discrep_ids.rds"))
 
@@
 
+## Study 2
+
+### Counts per unique value of family prior year, and state average
+
 ```{r}
@@
     ),
+    app_abrv = case_when(
+      appliance == "Cooling" ~ "AC",
+      appliance == "Heating" ~ "HEAT",
+      appliance == "Water Heating" ~ "WATER",
+      appliance == "Refrigerator" ~ "REFRIG",
+      appliance == "Other Appliance" ~ "OTHER",
+      appliance == "Total kWh" ~ "TOTAL"
+    ),
     family_name = case_when(
@@
 ```
+
+### Discrepant Subjects
+
+```{r}
+
+s2_long %>%
+  filter(plan == "plan1",
+  id !="s2_208") %>% # s2_208 is just missing a value for a total column - but has all their app values
+  group_by(appliance, rounded, state, pct_goal) %>%
+  mutate(
+    distinct_families = n_distinct(family),
+    distinct_state_avg = n_distinct(state_avg),
+  ) %>% 
+  filter(distinct_families > 1 | distinct_state_avg > 1) %>%
+  group_by(appliance, rounded, state, pct_goal, family) %>%
+  mutate(family_count = n()) %>%
+  group_by(appliance, rounded, state, pct_goal, state_avg) %>%
+  mutate(state_avg_count = n()) %>%
+  distinct(rounded, state, refClass, appliance, family, state_avg, 
+          state_avg_count, family_count) %>%
+  mutate(
+    # Check family values
+    fam_in = map2_lgl(rounded, family, ~ {
+      if (.x == "Rounded") {
+        .y %in% correct_values_s2$family_rounded
+      } else {
+        .y %in% correct_values_s2$family_exact
+      }
+    }),
+    # Check state averages
+    state_in = pmap_lgl(list(state, appliance, state_avg), function(s, a, sa) {
+      correct_row <- correct_values_s2 %>%
+        filter(state == s, app_abrv == a)
+        #print(nrow(correct_row))  
+      if(nrow(correct_row) == 0) FALSE else sa == correct_row$state_avg
+    })
+  ) %>%
+  arrange(rounded, refClass, state, appliance, family_count) %>%
+  relocate(rounded, pct_goal, refClass, state, appliance, family, 
+          family_count, state_avg, state_avg_count, fam_in, state_in) |> kable()
+
+discrepant_subjects_s2 <- s2_long |> 
+  filter(plan == "plan1") |>
+  group_by(id, rounded, refClass) |> 
+  mutate(
+    # Compute state_in based on exact match for state, appliance abbreviation, and state_avg
+    state_in = pmap_lgl(list(state, appliance, state_avg), function(s, a, sa) {
+      correct_row <- correct_values_s2 %>%
+        filter(state == s, app_abrv == a)
+      if (nrow(correct_row) == 0) FALSE else sa == correct_row$state_avg
+    }),
+    # Compute fam_in based on `rounded` value and appropriate family column in `correct_values_s2`
+    fam_in = pmap_lgl(list(state, appliance, family, rounded), function(s, a, fam, rnd) {
+      correct_row <- correct_values_s2 %>%
+        filter(state == s, app_abrv == a)
+      if (nrow(correct_row) == 0) {
+        FALSE
+      } else if (rnd == "Rounded") {
+        fam == correct_row$family_rounded
+      } else {
+        fam == correct_row$family_exact
+      }
+    })
+  )
+
+discrepant_subjects_s2 |> filter(state_in==FALSE | fam_in==FALSE) |>  arrange(state, appliance,refClass,pct_goal,rounded) |> select(-calc,-edu, -plan,-state_dif) |> kable()
+
+s2_discrep_ids <- discrepant_subjects_s2 |> filter(state_in==FALSE | fam_in==FALSE) |> pull(id) |> unique()
+s2_discrep_ids
+# saveRDS(s2_discrep_ids,here::here("data/s2_discrep_ids.rds"))
+
+# mutate(
+#   fam_in = if_else(rounded == "Rounded",
+#                   family %in% correct_values_s2$family_rounded,
+#                   family %in% correct_values_s2$family_exact),
+# # Use `map_lgl` to evaluate each row individually
+#   state_in = map_lgl(
+#     seq_along(appliance),
+#     ~ any(correct_values_s2$appliance[correct_values_s2$state_avg == state_avg[.x]] == appliance[.x])
+#   ) )
+
+```
+
+```{r}
+
+# 3 s2_185     2 USD      HEAT      10%      Rounded Colorado Calculator    Some graduate school   0.1  18000     16411 plan1 17000  -1000       589
+# 4 s2_196     2 USD      HEAT      10%      Rounded Colorado Calculator    College degree         0.1  18000        NA plan1 17000  -1000        NA
+# 5 s2_197     2 USD      HEAT      10%      Rounded Colorado Calculator    Some college           0.1  18000     16411 plan1 17000  -1000       589
+
+#  s2 |>
+#   mutate(id=paste0("s2_",row_number())) |> relocate(id) |> filter(id=="s2_196") |> 
+#    select(-starts_with("AIN")) |> select(starts_with("HEAT"))
+
+# s2 |> filter(id=="s2_196") |> select(starts_with("HEAT"))
+
+s2_long |> filter(state=="Colorado",state_avg==1142,appliance=="WATER")
+
+haven::read_sav(here::here("data/Frequency & Probability Study 2 - 3-24-19.sav")) |>
+  mutate(id=paste0("s2_",row_number())) |> 
+  relocate(id) |> filter(id=="s2_196") |> 
+   select(-starts_with("AIN")) |> select(starts_with("HEAT"))
+
+#s2 |> filter(id=="s2_208") |> select(starts_with("T"))
+s2_long |> filter_all(any_vars(is.na(.)))
+
+```

### 2024-11-04 study2.qmd
compute study 2 discrepancy counts

@@
   mutate(id=paste0("s2_",row_number())) |> relocate(id)
-s2 <- s2 |> filter(!(id=="s2_78")) # NA for some responses
+s2 <- s2 |> filter(!(id=="s2_78")) |> # NA for some responses
+  filter(!(id=="s2_196")) # NA for state_avg for COL HEAT - (HEATSTATE2)
+  # filter(!(id=="s2_208")) # NA for response for California not rounded Total (TOTALAPONE1 and TOTALAPTWO1) - leave them in since all the acutal app values are okay
 s2 <- s2 |> 
@@
   mutate(change = value-family, state_dif = value-state_avg,plan=as_factor(plan))
+
+# s2 |> filter(id=="s2_208") |> select(starts_with("T"))
+# s2_long |> filter_all(any_vars(is.na(.)))
 

### 2024-11-04 misc/discrep.qmd
update with before and after discrepancy exclusion tables and figures

@@
 pacman::p_load(dplyr,purrr,tidyr,here,tibble, haven,
-               ggplot2,gt,knitr,kableExtra,ggh4x,lme4,flextable)
+               ggplot2,gt,knitr,kableExtra,ggh4x,ggdist,gghalves,patchwork,lme4,flextable)
 options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
@@
 ```
+
+## Data after exclusions
+
+### Study 1
+
+```{r}
+
+# number of unique ids per condition
+s1 |> group_by(refClass) |> 
+filter(!(id %in% s1_discrep_ids)) |>
+rename("Reference Class" = refClass) |>
+summarise(n=n_distinct(id)) |> gt() |> tab_header(title="Study 1: Number of unique IDs per condition")
+
+s1 |> group_by(refClass,state) |> 
+  filter(!(id %in% s1_discrep_ids)) |>
+  rename("Reference Class" = refClass) |>
+  summarise(n=n_distinct(id)) |> 
+  rowwise() |> 
+  mutate(state=lookup_code(state)) |> 
+  pivot_wider(names_from=state,values_from=n) |>
+  gt() |> tab_header(title="Study 1: Number of unique IDs per condition by state")
+
+s1 |> group_by(refClass,calc) |> 
+  filter(!(id %in% s1_discrep_ids)) |>
+  rename("Reference Class" = refClass) |>
+  summarise(n=n_distinct(id)) |> 
+  pivot_wider(names_from=calc,values_from=n) |>
+  gt() |> tab_header(title="Study 1: Number of unique IDs per condition by calculation")
+
+s2_long |> group_by(refClass) |>
+filter(!(id %in% s2_discrep_ids)) |>
+rename("Reference Class" = refClass) |>
+summarise(n=n_distinct(id)) |> gt() |> tab_header(title="Study 2: Number of unique IDs per condition")
+
+s2_long |> group_by(refClass,calc) |>
+filter(!(id %in% s2_discrep_ids)) |>
+rename("Reference Class" = refClass) |>
+mutate(calc=case_when(calc=="Calculator" ~ "Used Calculator",
+                      calc=="No Calculator" ~ "No Calculator")) |>
+summarise(n=n_distinct(id)) |>
+pivot_wider(names_from=calc,values_from=n) |>
+gt() |> tab_header(title="Study 2: Number of unique IDs per condition by calculation")
+
+# s2_long |> group_by(id) |> summarise(nrc=n_distinct(refClass),npct=n_distinct(pct_goal),nround=n_distinct(rounded))
+
+```
+
+### Performance before and after exclusions - Study 1
+
+```{r}
+#| fig-width: 12
+#| fig-height: 8
+
+s1_agg <- s1 |> 
+  filter(appliance !="Total kWh") |> 
+  group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
+  summarise(total_kWh = sum(value),orig_kWh=sum(family), 
+            pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
+            n_change = sum(value!=family),
+            state_p_dif=mean(state_p_dif),
+            state_f_dif=mean(state_f_dif),
+            n_less_avg = sum(less_avg),
+            duration=first(Duration__in_seconds_)) |> 
+  mutate(matched_goal = (pct_change == pct_goal), 
+             error = pct_change - pct_goal,
+            abs_error = abs(error),
+            close_match = abs_error <= 0.02)
+
+s1_agg4 <- s1_agg |> group_by(id,refClass) |> 
+  summarise(mg=sum(matched_goal),n=n(), pct=mg/n)
+
+# s1_agg4 |> group_by(refClass) |> 
+#   summarise(mean=mean(pct),sd=sd(pct),n=n()) |> 
+#   gt() |> tab_header(title="Study 1: Performance before exclusions")
+
+# s1_agg4 |> 
+#   filter(!(id %in% s1_discrep_ids)) |> 
+#   group_by(refClass) |> 
+#   summarise(mean=mean(pct),sd=sd(pct),n=n()) |> 
+#   gt() |> tab_header(title="Study 1: Performance after exclusions")
+
+bind_rows(
+  s1_agg4 |> 
+    group_by(refClass) |> 
+    summarise(
+      mean = mean(pct),
+      sd = sd(pct),
+      n = n(),
+      dataset = "Before exclusions"
+    ),
+  s1_agg4 |> 
+    filter(!(id %in% s1_discrep_ids)) |> 
+    group_by(refClass) |> 
+    summarise(
+      mean = mean(pct),
+      sd = sd(pct),
+      n = n(),
+      dataset = "After exclusions"
+    )
+) |>
+  pivot_wider(
+    names_from = dataset,
+    values_from = c(mean, sd, n),
+    names_glue = "{dataset}_{.value}"
+  ) |>
+  gt() |>
+  tab_header(
+    title = "Study 1: Performance Before and After Exclusions"
+  ) |>
+  fmt_number(
+    columns = contains("mean"),
+    decimals = 2
+  ) |>
+  fmt_number(
+    columns = contains("sd"),
+    decimals = 2
+  ) |>
+  cols_label(
+    refClass = "Reference Class",
+    `Before exclusions_mean` = "Mean",
+    `Before exclusions_sd` = "SD",
+    `Before exclusions_n` = "N",
+    `After exclusions_mean` = "Mean",
+    `After exclusions_sd` = "SD",
+    `After exclusions_n` = "N"
+  ) |>
+  tab_spanner(
+    label = "Before Exclusions",
+    columns = starts_with("Before")
+  ) |>
+  tab_spanner(
+    label = "After Exclusions",
+    columns = starts_with("After")
+  )
+
+sp1b <- s1_agg4 |> 
+  ggplot(aes(x=refClass,y=pct,fill=refClass)) +
+  stat_summary(fun=mean, geom="bar") +
+  stat_summary(fun.data=mean_cl_normal, geom="errorbar", width=0.2) +
+  labs(title="Mean % meeting goal - Before Exclusions",
+       x="Reference Class",
+       y="% of trials meeting goal",
+       fill="Reference Class")
+
+sp1a <- s1_agg4 |> 
+  filter(!(id %in% s1_discrep_ids)) |>
+  ggplot(aes(x=refClass,y=pct,fill=refClass)) +
+  stat_summary(fun=mean, geom="bar") +
+  stat_summary(fun.data=mean_cl_normal, geom="errorbar", width=0.2) +
+    labs(title="Mean % meeting goal - After Exclusions",
+       x="Reference Class",
+       y="% of trials meeting goal",
+       fill="Reference Class")
+
+sp2b <- s1_agg |> group_by(id,refClass) |> 
+  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
+  ggplot(aes(x=pct)) +
+  geom_bar(aes(fill=refClass)) +
+  ggdist::stat_halfeye() +
+  geom_rug(alpha=0.1) +
+  stat_boxplot(width=0.5, alpha=0.6) +
+  facet_wrap(~refClass) +
+  labs(title="Distributions of % of trials that matched goal - Before Exclusions",
+       x="% of trials (out of 4) that matched goal",
+       y="Number of Participants",
+       fill="Reference Class") +
+  scale_y_continuous(limits=c(0,85),breaks=seq(0,85,5)) + scale_x_continuous(breaks=seq(0,1,.25),labels=scales::percent_format(accuracy=1))
+
+sp2a <- s1_agg |> 
+  filter(!(id %in% s1_discrep_ids)) |> 
+  group_by(id,refClass) |> 
+  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
+  ggplot(aes(x=pct)) +
+  geom_bar(aes(fill=refClass)) +
+  ggdist::stat_halfeye() +
+  geom_rug(alpha=0.1) +
+  stat_boxplot(width=0.5, alpha=0.6) +
+  facet_wrap(~refClass) +
+  labs(title="Distributions of % of trials that matched goal - After Exclusions",
+       x="% of trials (out of 4) that matched goal",
+       y="Number of Participants",
+       fill="Reference Class") +
+  scale_y_continuous(limits=c(0,85),breaks=seq(0,85,5)) + scale_x_continuous(breaks=seq(0,1,.25),labels=scales::percent_format(accuracy=1))
+
+sp1b + sp1a
+
+sp2b + sp2a
+
+```
+
+### After Exclusions - Study 2
+
+```{r}
+#| fig-width: 12
+#| fig-height: 8
+
+s2_agg1 <- s2_long |> 
+  filter(appliance != "TOTAL") |> 
+  group_by(id, refClass,pct,plan) |> 
+  summarise(
+    total_kWh = sum(value),
+    orig_kWh = sum(family),
+    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
+    state_dif = mean(state_dif),
+    .groups = "drop"
+  ) |>
+  mutate(matched_goal = (pct_change == pct),
+  matched_goal2 = abs(pct_change-pct)<.03)
+
+s2_agg4 <- s2_agg1|> group_by(id,refClass) |> 
+  summarise(mg=sum(matched_goal),n=n(), pct=mg/n)
+
+bind_rows(
+  s2_agg4 |> 
+    group_by(refClass) |> 
+    summarise(
+      mean = mean(pct),
+      sd = sd(pct),
+      n = n(),
+      dataset = "Before exclusions"
+    ),
+  s2_agg4 |> 
+    filter(!(id %in% s2_discrep_ids)) |> 
+    group_by(refClass) |> 
+    summarise(
+      mean = mean(pct),
+      sd = sd(pct),
+      n = n(),
+      dataset = "After exclusions"
+    )) |> 
+    pivot_wider(names_from = dataset,
+    values_from = c(mean, sd, n),
+    names_glue = "{dataset}_{.value}") |> gt() |>
+    tab_header(title = "Study 2: Performance Before and After Exclusions") |>
+    cols_label(
+    refClass = "Reference Class",
+    `Before exclusions_mean` = "Mean",
+    `Before exclusions_sd` = "SD",
+    `Before exclusions_n` = "N",
+    `After exclusions_mean` = "Mean",
+    `After exclusions_sd` = "SD",
+    `After exclusions_n` = "N"
+  ) |>
+  tab_spanner(
+    label = "Before Exclusions",
+    columns = starts_with("Before")
+  ) |>
+  tab_spanner(
+    label = "After Exclusions",
+    columns = starts_with("After"))
+
+s2p1b <- s2_agg4 |> 
+  ggplot(aes(x=refClass,y=pct,fill=refClass)) +
+  stat_summary(fun=mean, geom="bar") +
+  stat_summary(fun.data=mean_cl_normal, geom="errorbar", width=0.2) +
+  labs(title="Mean % meeting goal - Before Exclusions",
+       x="Reference Class",
+       y="% of trials meeting goal",
+       fill="Reference Class")
+
+s2p1a <- s2_agg4 |> 
+  filter(!(id %in% s2_discrep_ids)) |>
+  ggplot(aes(x=refClass,y=pct,fill=refClass)) +
+  stat_summary(fun=mean, geom="bar") +
+  stat_summary(fun.data=mean_cl_normal, geom="errorbar", width=0.2) +
+  labs(title="Mean % meeting goal - After Exclusions",
+       x="Reference Class",
+       y="% of trials meeting goal",
+       fill="Reference Class")
+
+s2p2b <- s2_agg1 |> group_by(id,refClass) |>
+  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
+  ggplot(aes(x=pct)) +
+  geom_bar(aes(fill=refClass)) +
+  ggdist::stat_halfeye() +
+  geom_rug(alpha=0.1) +
+  stat_boxplot(width=0.5, alpha=0.6) +
+  facet_wrap(~refClass) +
+  labs(title="Distributions of % of trials that matched goal - Before Exclusions",
+       x="% of trials (out of 4) that matched goal",
+       y="Number of Participants",
+       fill="Reference Class") + 
+       # make title text smaller
+        theme(plot.title = element_text(size=10)) +
+  scale_y_continuous(limits=c(0,85),breaks=seq(0,85,5)) + scale_x_continuous(breaks=seq(0,1,.25),labels=scales::percent_format(accuracy=1))
+
+s2p2a <- s2_agg1 |> 
+  filter(!(id %in% s2_discrep_ids)) |> 
+  group_by(id,refClass) |> 
+  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
+  ggplot(aes(x=pct)) +
+  geom_bar(aes(fill=refClass)) +
+  ggdist::stat_halfeye() +
+  geom_rug(alpha=0.1) +
+  stat_boxplot(width=0.5, alpha=0.6) +
+  facet_wrap(~refClass) +
+  labs(title="Distributions of % of trials that matched goal - After Exclusions",
+       x="% of trials (out of 4) that matched goal",
+       y="Number of Participants",
+       fill="Reference Class") +
+  scale_y_continuous(limits=c(0,85),breaks=seq(0,85,5)) + scale_x_continuous(breaks=seq(0,1,.25),labels=scales::percent_format(accuracy=1))
+
+s2p1b + s2p1a
+
+s2p2b + s2p2a
+
+```

### 2024-11-04 misc/indiv.qmd
update with before and after discrepancy exclusion tables and figures

@@
 ## Study 1 - comparison against state average
+
 ```{r}
-
 s2_long |> filter(id %in% unique(s2_long$id)[1:30]) |> 
@@
   facet_wrap(~id) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
+```{r}
 

### 2024-11-04 misc/discrep.qmd
add proportion tables and SE

@@
 
-### Study 1
+### Study 1 - Subject counts
 
@@
 
+prop.table(xtabs(~refClass,s1)) |> kable()
+prop.table(xtabs(~refClass+calc,s1)) |> kable()
+prop.table(xtabs(~refClass+state,s1)) |> kable()
+
+# s2_long |> group_by(id) |> summarise(nrc=n_distinct(refClass),npct=n_distinct(pct_goal),nround=n_distinct(rounded))
+
+```
+
+### Study 2 - Subject counts 
+```{r}
+
 s2_long |> group_by(refClass) |>
@@
 
-# s2_long |> group_by(id) |> summarise(nrc=n_distinct(refClass),npct=n_distinct(pct_goal),nround=n_distinct(rounded))
+prop.table(xtabs(~refClass,s2_long)) |> kable()
+prop.table(xtabs(~refClass+state,s2_long))
+prop.table(xtabs(~refClass+calc,
+s2_long |> mutate(calc=case_when(calc=="Calculator" ~ "Used Calculator",
+                      calc=="No Calculator" ~ "No Calculator")))) |> kable()
+```
 
-```
+##  Performance before and after exclusions
 
@@
       n = n(),
+      se=sd(pct)/sqrt(n),
       dataset = "Before exclusions"
@@
       n = n(),
+      se=sd(pct)/sqrt(n),
       dataset = "After exclusions"
@@
     names_from = dataset,
-    values_from = c(mean, sd, n),
+    values_from = c(mean, sd, se,n),
     names_glue = "{dataset}_{.value}"
@@
     `Before exclusions_sd` = "SD",
+    `Before exclusions_se` = "SE",
     `Before exclusions_n` = "N",
@@
     `After exclusions_sd` = "SD",
+    `After exclusions_se` = "SE",
     `After exclusions_n` = "N"
@@
     pivot_wider(names_from = dataset,
-    values_from = c(mean, sd, n),
+    values_from = c(mean, sd,se, n),
     names_glue = "{dataset}_{.value}") |> gt() |>
@@
 
+s1_agg4 |> 
+    group_by(refClass) |> 
+    summarise(
+      mean = mean(pct),
+      sd = sd(pct),
+      n = n(),
+      se=sd(pct)/sqrt(n),
+      dataset = "Before exclusions"
+    )
+
+s1_agg4 |> 
+    group_by(refClass) |> 
+    summarise(
+      mean = mean(pct),
+      sd = sd(pct),
+      n = n(),
+      se=sd/sqrt(n),
+      dataset = "Before exclusions"
+    )
+
+s1_agg4 |> 
+    group_by(refClass) |> 
+    summarise(
+      mean = mean(pct),
+      sd = sd(pct),
+      n = n(),
+      se = sd/sqrt(n),  # Using sd calculated above
+      dataset = "After exclusions"
+    )
+
 ```

### 2024-11-05 misc/discrep.qmd
update response pattern plots

@@
 
-s1_agg4 |> 
-    group_by(refClass) |> 
-    summarise(
-      mean = mean(pct),
-      sd = sd(pct),
-      n = n(),
-      se=sd(pct)/sqrt(n),
-      dataset = "Before exclusions"
-    )
-
-s1_agg4 |> 
-    group_by(refClass) |> 
-    summarise(
-      mean = mean(pct),
-      sd = sd(pct),
-      n = n(),
-      se=sd/sqrt(n),
-      dataset = "Before exclusions"
-    )
-
-s1_agg4 |> 
-    group_by(refClass) |> 
-    summarise(
-      mean = mean(pct),
-      sd = sd(pct),
-      n = n(),
-      se = sd/sqrt(n),  # Using sd calculated above
-      dataset = "After exclusions"
-    )
-
 ```

### 2024-11-05 misc/indiv.qmd
update response pattern plots

@@
 
-s1 <- readRDS(here::here("data/s1_processed.rds"))
+#https://arelbundock.com/posts/quarto_figures/index.html
+knitr::opts_chunk$set(
+    out.width = "75%", # enough room to breath
+    fig.width = 6,     # reasonable size
+    fig.asp = 0.618,   # golden ratio
+    fig.align = "center" # mostly what I want
+)
 
-s2_long <- readRDS(here::here("data/s2_processed.rds"))
+out2fig = function(out.width, out.width.default = 0.7, fig.width.default = 6) {
+    fig.width.default * out.width / out.width.default 
+}
+
+s1 <- readRDS(here::here("data/s1_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds"))))
+s2_long <- readRDS(here::here("data/s2_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds"))))
 
@@
 ```{r}
+#| fig-asp: .8
+
+#| 
 s2_long |> filter(id %in% unique(s2_long$id)[1:30]) |> 
@@
 ```{r}
+#| fig.height: 13
+#| fig.width: 12
 
@@
                aes(y = value, color = "Plan (Not Matching)", shape = "Plan"), 
-               size = 2, alpha = 0.4) +
+               size = 1, alpha = 0.4) +
     # Matching plans
@@
                aes(y = value, color = "Plan (Matching Goal)", shape = "Plan"), 
-               size = 2.5) +
+               size = 1.5) +
     # Prior year values
     geom_point(aes(y = family, color = "Prior Year", shape = "Prior Year"), 
-               size = 3.5) +
+               size = 2.5) +
     # State averages with star shape, plotted last and largest
     geom_point(aes(y = state_avg, color = "State Average", shape = "State Average"), 
-               size = 3, stroke = 2) +
+               size = 2, stroke = 2) +
     scale_color_manual(

### 2024-11-06 misc/_inspect_data.qmd
reorganize a bit

```{r}
library(pacman)
pacman::p_load(dplyr,purrr,tidyr,haven,tibble,ggplot2,gt,ggh4x,lme4)

d <- read_spss("Frequency & Probability Study 1 - 3-24-19.sav")
d |> glimpse()
colnames(d)[1:500]
colnames(d[1000:1840])
head(d,1)

```

| **ORDER VARIABLE FORMAT** |  |  |
| Colorado/Wells Family | COLORADO | 1 = first<br>2 = second<br>Missing = not received |
| Massachusetts/Davis Family | MASSACHUSETTS | (same as COLORADO) |
| California/Adams Family | CALIFORNIA | (same as COLORADO) |
| Texas/Smith Family | TEXAS | (same as COLORADO) |
| Reference class received | REFERENCECLASS | 1=USD<br>2=kWh<br>3=Percentage |
| Was participant's first scenario rounded? | ROUNDING1 | 1=yes<br>0=no |
| Was participant's second scenario rounded? | ROUNDING2 | (same as ROUNDING1) |
| Did the participant receive the rounded scenario first or second? | ROUNDED | 1 = first<br>2 = second |
| Was participant's first scenario a 10% reduction goal? | TENPERCENT1 | 1=yes<br>0=no |
| Was participant's second scenario a 10% reduction goal? | TENPERCENT2 | (same as TENPERCENT1) |
| Did the participant receive the 10% reduction goal first or second? | TENPERCENT | 1 = first<br>2 = second |

```{r}

# for each ResponseId, how many unique ReferenceClass values are there?

d |> 
  group_by(ResponseId) |> 
  summarise(n_distinct(ReferenceClass)) |> head()

d |> 
  group_by(ResponseId, ReferenceClass) |> 
  summarise(n())

pmat <- d |> 
  group_by(ResponseId, ReferenceClass, Texas, California,Colorado,Massachusetts) |> 
  summarise(n=n())

# head(pmat)
# # A tibble: 6 × 7
# # Groups:   ResponseId, ReferenceClass, Texas, California, Colorado [6]
#   ResponseId        ReferenceClass Texas California Colorado Massachusetts     n
#   <chr>                      <dbl> <dbl>      <dbl>    <dbl>         <dbl> <int>
# 1 R_081G93DbeOgMjjX              1    NA          1        2            NA     1
# 2 R_08qP6LR1m2VzuQp              1    NA          2        1            NA     1
# 3 R_0keShcI9atm6rux              3     2         NA       NA             1     1
# 4 R_10DdSXDfNIuNmFY              3    NA          1       NA             2     1
# 5 R_11c5U65IWfuEeCI              3    NA          1        2            NA     1
# 6 R_12AnbKDp26Lk7NZ              1    NA          2        1            NA     1

# replace NA in pmat with 0
pmat[is.na(pmat)] <- 0
head(pmat)

# head(pmat)
# # A tibble: 6 × 7
# # Groups:   ResponseId, ReferenceClass, Texas, California, Colorado [6]
#   ResponseId        ReferenceClass Texas California Colorado Massachusetts     n
#   <chr>                      <dbl> <dbl>      <dbl>    <dbl>         <dbl> <int>
# 1 R_081G93DbeOgMjjX              1     0          1        2             0     1
# 2 R_08qP6LR1m2VzuQp              1     0          2        1             0     1
# 3 R_0keShcI9atm6rux              3     2          0        0             1     1
# 4 R_10DdSXDfNIuNmFY              3     0          1        0             2     1
# 5 R_11c5U65IWfuEeCI              3     0          1        2             0     1
# 6 R_12AnbKDp26Lk7NZ              1     0          2        1             0     1

# create new column, condit,  A string with ReferenceClass, _, first state name,_, second state name
#   ResponseId        ReferenceClass Texas California Colorado Massachusetts     n condit
#   <chr>                      <dbl> <dbl>      <dbl>    <dbl>         <dbl> <int> <chr> 
# 1 R_081G93DbeOgMjjX              1     0          1        2             0     1 10120 
# e.g., for the first subject, condit = 1_California_Colorado
# The state names with values of 0 are not included in the condit string

pmat <- pmat |> 
  mutate(condit = paste0(ReferenceClass, "_", 
                         ifelse(Texas == 1, "Texas", 
                                ifelse(California == 1, "California", 
                                       ifelse(Colorado == 1, "Colorado", 
                                              ifelse(Massachusetts == 1, "Massachusetts", "NA")
                                       )
                                )
                         ), "_", 
                         ifelse(Texas == 2, "Texas", 
                                ifelse(California == 2, "California", 
                                       ifelse(Colorado == 2, "Colorado", 
                                              ifelse(Massachusetts == 2, "Massachusetts", "NA")
                                       )
                                )
                         )
  )
  )

# count number of unique responseID for each condit

pmat |> 
  group_by(condit) |> 
  summarise(n_distinct(ResponseId)) |> head()

# # A tibble: 6 × 2
#   condit                     `n_distinct(ResponseId)`
#   <chr>                                         <int>
# 1 1_California_Colorado                            10
# 2 1_California_Massachusetts                        9
# 3 1_Colorado_California                            12
# 4 1_Colorado_Texas                                 14
# 5 1_Massachusetts_California                       11
# 6 1_Massachusetts_Texas                            11

```

### 2024-11-06 analysis1.R

@@ -1,63 +1,0 @@
-
-
-
-pacman::p_load(dplyr,purrr,tidyr,tibble,ggplot2,gt,patchwork,ggh4x,knitr,lme4)
-data <- read.csv("processed_energy_study_data.csv")
-
-colnames(data)
-head(data,1)
-glimpse(data)
-
-
-
-data %>%
-  summarise(
-    avg_age = mean(2021 - birth_year, na.rm = TRUE),
-    gender_distribution = table(gender)
-  )
-
-avg_age <- data %>%
-  summarise(avg_age = mean(birth_year, na.rm = TRUE))
-
-final_data %>%
-  count(gender)
-
-
-
-# Reliability analysis for 'els' scale
-library(psych)
-els_items <- final_data %>%
-  select(starts_with("els")) %>%
-  mutate_all(as.numeric)
-alpha_results <- alpha(els_items)
-print(alpha_results)
-
-
-
-
-
-# Visualization of scenario responses
-data %>%
-  ggplot(aes(x = scenario, y = `01`)) +
-  geom_boxplot() +
-  labs(title = "Distribution of Response '01' Across Scenarios")
-
-
-
-model <- lmer(X01 ~ scenario + (1 | responseid), data = data)
-
-
-
-
-scenario_means <- data %>%
-  group_by(scenario) %>%
-  summarise(mean_X01 = mean(X01, na.rm = TRUE))
-
-scenario_means %>%
-  ggplot(aes(x = scenario, y = mean_X01)) +
-  geom_bar(stat = "identity") +
-  labs(title = "Mean Response 'X01' Across Scenarios",
-       x = "Scenario",
-       y = "Mean Response") +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))

### 2024-11-07 misc/ordinal.qmd
finished mpa abstract

---
title: "Ordinal Models"
author:
- name: Thomas E. Gorman
  url: https://tegorman13.github.io/
  affiliations: 
  - name:  Communication and Cognition Lab, Purdue University, USA
    affiliation-url: https://web.ics.purdue.edu/~treimer/
#lightbox: true
toc: false
code-fold: true
code-tools: true
execute:
  #echo: false
  warning: false
format:
  html: 
    grid:
      sidebar-width: 220px
      body-width: 1200px
      margin-width: 170px
      gutter-width: 1.0rem
  hugo-md:
    include: true
    html-math-method: mathjax
    output-file: ordinal_hugo.md
  gfm:
    echo: true
    output-file: ordinal.md
---

## Ordinal Regression alternative

```{r}

pacman::p_load(dplyr,purrr,tidyr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,
               ggplot2,gt,knitr,kableExtra,ggh4x,lme4,flextable,pander)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("fun_plot"), ~ source(here::here(paste0("Scripts/", .x, ".R"))))
theme_set(theme_nice())
options(brms.backend="cmdstanr",mc.cores=4)

s1 <- readRDS(here::here("data/s1_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds"))))
s2_long <- readRDS(here::here("data/s2_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds"))))

```

## Proportions in aggregate data

```{r}

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                close_match = abs_error <= 0.03) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        )

ggplot(data = s1_agg, aes(x = accuracy_level, fill = refClass)) +
  geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Accuracy Levels by Reference Class",
       x = "Accuracy Level",
       y = "Percentage of Participants",
       fill = "Reference Class") +
  theme_minimal()

observed_props <- s1_agg |>
  group_by(refClass, accuracy_level) |>
  summarise(n = n()) |>
  group_by(refClass) |>
  mutate(prop = n/sum(n)) |>
  pivot_wider(
    names_from = accuracy_level,
    values_from = c(n, prop)
  )

  # Calculate proportions
prop_acc_s1 <- s1_agg %>%
    group_by(refClass, accuracy_level) %>%
    summarise(count = n()) %>%
    group_by(refClass) %>%
    mutate(Probability = count / sum(count)) %>%
    ungroup()

# Plot
ggplot(prop_acc_s1, aes(x = accuracy_level, y = Probability, fill = refClass)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    labs(title = "Distribution of Accuracy Levels by Reference Class",
        x = "Accuracy Level",
        y = "Probability",
        fill = "Reference Class") +
    theme_minimal() +
    scale_fill_brewer(palette = "Set1")

```

## Fit Ordinal model accuracy_level ~ refClass + (1|id) + (1|state)
```{r}

ordinal_model_s1 <- brm(
  accuracy_level ~ refClass + (1|id) + (1|state),
  data = s1_agg,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 1.5), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s1_ordinal3.rds')) # Cache for efficiency
)

summary(ordinal_model_s1)

#ordinal_model_s1 |> emmeans(~refClass, type="response")

pp_check(ordinal_model_s1)

describe_posterior(ordinal_model_s1, centrality = "Mode") |> 
  filter(stringr::str_detect(Parameter, "b_")) |> 
  mutate(Parameter = stringr::str_remove(Parameter, "b_")) |> 
  mutate(across(where(is.numeric), round, 3)) |> 
  kable(booktabs = TRUE)

library(tidybayes)

# Get predicted probabilities
pred_summary <- ordinal_model_s1 |>
    epred_draws(newdata = data.frame(refClass = c("kWh", "Percentage", "USD")),
                ndraws = 1000, re_formula = NA) |>
    group_by(refClass, .category) |>
    summarise(
        mean_prob = mean(.epred),
        lower_ci = quantile(.epred, 0.025),
        upper_ci = quantile(.epred, 0.975)
    )
pred_summary |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy level by reference class")

ggplot(pred_summary, 
        aes(x = .category, y = mean_prob, fill = refClass )) +
    geom_col(position = position_dodge(width = 0.8)) +
    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), 
                    position = position_dodge(width = 0.8)) +
    labs(x = "Reference Class", 
        y = "Predicted Probability",
        fill = "Accuracy Level",
        title = "Predicted Probabilities of Accuracy Levels by Reference Class") +
    scale_fill_brewer(palette = "RdYlBu") +
    theme_minimal()

ggplot(pred_summary, 
        aes(x = refClass, y = mean_prob, fill = .category )) +
    geom_col(position = position_dodge(width = 0.8)) +
    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), 
                    position = position_dodge(width = 0.8)) +
    labs(x = "Reference Class", 
        y = "Predicted Probability",
        fill = "Accuracy Level",
        title = "Predicted Probabilities of Accuracy Levels by Reference Class") +
    scale_fill_brewer(palette = "RdYlBu") +
    theme_minimal()

# Calculate cumulative probabilities
cum_probs <- pred_summary |>
  group_by(refClass) |>
  arrange(refClass, .category) |>
  mutate(
    cum_prob = cumsum(mean_prob),
    cum_lower = cumsum(lower_ci),
    cum_upper = cumsum(upper_ci)
  )

ggplot(cum_probs, 
    aes(x = refClass, y = cum_prob, color = .category)) +
geom_point(size = 3,position=position_dodge()) +
geom_errorbar(aes(ymin = cum_lower, ymax = cum_upper), 
            width = 0.2,position=position_dodge()) +
labs(x = "Reference Class",
    y = "Cumulative Probability",
    color = "Accuracy Level",
    title = "Cumulative Probabilities of Accuracy Levels") +
theme_minimal() +
scale_color_brewer(palette = "Blues")

summary_table <- pred_summary |>
  pivot_wider(
    names_from = .category,
    values_from = c(mean_prob, lower_ci, upper_ci)
  ) |>
  kable(digits = 3)

print(summary_table)

# Convert log-odds to odds ratios
posterior_samples <- as.data.frame(ordinal_model_s1)
odds_ratios <- data.frame(
  Percentage_vs_kWh = exp(posterior_samples$b_refClassPercentage),
  USD_vs_kWh = exp(posterior_samples$b_refClassUSD)
)

# Calculate summary statistics
odds_ratio_summary <- data.frame(
  comparison = c("Percentage vs kWh", "USD vs kWh"),
  odds_ratio = c(mean(odds_ratios$Percentage_vs_kWh),
                 mean(odds_ratios$USD_vs_kWh)),
  ci_lower = c(quantile(odds_ratios$Percentage_vs_kWh, 0.025),
               quantile(odds_ratios$USD_vs_kWh, 0.025)),
  ci_upper = c(quantile(odds_ratios$Percentage_vs_kWh, 0.975),
               quantile(odds_ratios$USD_vs_kWh, 0.975))
)

# Plot predicted probabilities
ggplot(pred_summary, aes(x = refClass, y = mean_prob, fill = .category)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci),
                position = position_dodge(width = 0.9), width = 0.2) +
  labs(y = "Predicted Probability", x = "Reference Class", fill = "Accuracy Level") +
  ggtitle("Study 1: Predicted Probabilities of Accuracy Level by Reference Class") +
  theme_minimal()

```

```{r}

bayesplot::color_scheme_set(wes_palettes[1])

pp_check(ordinal_model_s1, type = "bars", ndraws = 1000, fatten = 2) +
    scale_x_continuous("y", breaks = 1:7) +
    scale_y_continuous(NULL, breaks = NULL, expand = expansion(mult = c(0, 0.05))) +
    ggtitle("Data with posterior predictions",
            subtitle = "N = 100")

pp_check(ordinal_model_s1, type = "bars_grouped", group="refClass", fatten = 2) +
    scale_x_continuous("y", breaks = 1:7) +
    scale_y_continuous(NULL, breaks = NULL, expand = expansion(mult = c(0, 0.05))) +
    ggtitle("Data with posterior predictions",
            subtitle = expression(list(italic(N[A])==44, italic(N[B])==44))) +
    theme(legend.background = element_blank(),
            legend.position = c(.9, .75))

f <- fitted(ordinal_model_s1,newdata_pred <- expand.grid(
    refClass = c("kWh", "Percentage", "USD"),
    id = unique(s1_agg$id),
    state = unique(s1_agg$state)
    ))

f |> str()

f <-
  rbind(
    f[, , 1],
    f[, , 2],
    f[, , 3]
  ) |> data.frame() |> bind_cols(expand_grid(new_data, Y=1:3))

```

## Comparing Marginal vs Conditional Predictions

```{r}

# Create newdata with all necessary variables
newdata_pred <- expand.grid(
  refClass = c("kWh", "Percentage", "USD"),
  id = unique(s1_agg$id),
  state = unique(s1_agg$state)
)

# Population-level predictions
pred_probs_marginal <- ordinal_model_s1 |>
  epred_draws(
    newdata = data.frame(refClass = c("kWh", "Percentage", "USD")),
    ndraws = 1000, 
    re_formula = NA
  )

# Conditional predictions (including random effects)
pred_probs_conditional <- ordinal_model_s1 |>
  epred_draws(
    newdata = newdata_pred,
    ndraws = 1000, 
    re_formula = NULL
  )

# Compare the two approaches
summary_comparison <- bind_rows(
    pred_probs_marginal |> 
        group_by(refClass, .category) |>
        summarise(
        mean_prob = mean(.epred),
        sd_prob = sd(.epred),
        lower_ci = quantile(.epred, 0.025),
        upper_ci = quantile(.epred, 0.975),
        .groups = 'drop'
        ) |>
        mutate(type = "Marginal"),

    pred_probs_conditional |> 
        group_by(refClass, .category) |>
        summarise(
        mean_prob = mean(.epred),
        sd_prob = sd(.epred),
        lower_ci = quantile(.epred, 0.025),
        upper_ci = quantile(.epred, 0.975),
        .groups = 'drop'
        ) |>
        mutate(type = "Conditional")
)

# View the comparison
summary_comparison |>
  arrange(type, refClass, .category) |>
  kable(digits = 3)

# Visualize the comparison
ggplot(summary_comparison, 
        aes(x = .category, y = mean_prob, fill = refClass)) +
    geom_col(position = position_dodge()) +
    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), 
            position = position_dodge(width = 0.8)) +
    facet_wrap(~type) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = "Comparing Marginal vs Conditional Predictions",
        x = "Accuracy Level",
        y = "Predicted Probability")
```

## Individual Differences in Predictions

```{r}
# Extract random effects for subjects with corrected ID handling
ranef_draws <- ordinal_model_s1 |>
  spread_draws(r_id[id,term]) |>
  mutate(id = as.integer(as.character(id))) # Convert factor to integer

# Get subject-specific adjustments
subject_adjustments <- ranef_draws |>
  group_by(id) |>
  summarise(
    mean_adj = mean(r_id),
    lower = quantile(r_id, 0.025),
    upper = quantile(r_id, 0.975)
  )

# Get reference class for each subject
subject_refclass <- s1_agg |>
  group_by(id) |>
  summarise(refClass = first(refClass))

# Combine adjustments with reference class
subject_plot_data <- subject_adjustments |>
  left_join(subject_refclass, by = "id")

# Create caterpillar plot
ggplot(subject_plot_data, 
       aes(x = reorder(factor(id), mean_adj), 
           y = mean_adj, 
           color = refClass)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower, ymax = upper), alpha = 0.5) +
  coord_flip() +
  facet_wrap(~refClass, scales = "free_y") +
  theme_minimal() +
  labs(title = "Subject-Level Random Effects by Reference Class",
       y = "Subject-Specific Adjustment",
       x = "Subject ID") +
  theme(legend.position = "none")

# Density plot of random effects by condition
ggplot(subject_plot_data, 
       aes(x = mean_adj, fill = refClass)) +
  geom_density(alpha = 0.5) +
  theme_minimal() +
  labs(title = "Distribution of Subject-Level Random Effects by Reference Class",
       x = "Subject-Specific Adjustment",
       y = "Density")

# Summary statistics by condition
random_effects_summary <- subject_plot_data |>
  group_by(refClass) |>
  summarise(
    n = n(),
    mean_adjustment = mean(mean_adj),
    sd_adjustment = sd(mean_adj),
    median_adjustment = median(mean_adj),
    q25 = quantile(mean_adj, 0.25),
    q75 = quantile(mean_adj, 0.75)
  )

print(random_effects_summary)

pop_preds <- ordinal_model_s1 |>
    epred_draws(
        newdata = data.frame(refClass = c("kWh", "Percentage", "USD")),
        ndraws = 1000, 
        re_formula = NA
    ) |>
    group_by(refClass, .category) |>
    summarise(
        pop_mean = mean(.epred),
        pop_lower = quantile(.epred, 0.025),
        pop_upper = quantile(.epred, 0.975)
    )

# Get subject-specific predictions
subject_preds <- ordinal_model_s1 |>
    epred_draws(
        newdata = s1_agg,
        ndraws = 100  # fewer draws for computational efficiency
    ) |>
    group_by(id, refClass, .category) |>
    summarise(
        subj_mean = mean(.epred)
    )

# Plot 1: Population effects with subject-specific points
ggplot() +
    # Add subject-specific points
    geom_jitter(data = subject_preds,
                aes(x = .category, y = subj_mean, color = refClass),
                alpha = 0.2, width = 0.2, height = 0) +
    # Add population-level effects
    geom_point(data = pop_preds,
                aes(x = .category, y = pop_mean, fill = refClass),
                size = 3, shape = 21, color = "black") +
    geom_errorbar(data = pop_preds,
                aes(x = .category, ymin = pop_lower, ymax = pop_upper),
                width = 0.2) +
    facet_wrap(~refClass) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = "Population-Level Effects with Subject-Specific Predictions",
        x = "Accuracy Level",
        y = "Predicted Probability")

# Plot 2: Violin plot with population effects
ggplot() +
    # Add violin plot of subject-specific effects
    geom_violin(data = subject_preds,
                aes(x = .category, y = subj_mean, fill = refClass),
                alpha = 0.3) +
    # Add population-level points
    geom_point(data = pop_preds,
                aes(x = .category, y = pop_mean),
                size = 3, color = "black") +
    geom_errorbar(data = pop_preds,
                aes(x = .category, ymin = pop_lower, ymax = pop_upper),
                width = 0.2) +
    facet_wrap(~refClass) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(title = "Distribution of Subject-Specific Effects with Population Estimates",
        x = "Accuracy Level",
        y = "Predicted Probability")

# Plot 3: Subject deviations from population mean
subject_deviations <- subject_preds |>
    left_join(pop_preds, by = c("refClass", ".category")) |>
    mutate(deviation = subj_mean - pop_mean)

ggplot(subject_deviations,
        aes(x = deviation, fill = refClass)) +
    geom_density(alpha = 0.5) +
    geom_vline(xintercept = 0, linetype = "dashed") +
    facet_grid(refClass ~ .category) +
    theme_minimal() +
    labs(title = "Subject Deviations from Population Mean",
        x = "Deviation from Population Mean",
        y = "Density")

# Plot 4: Ridgeline plot of subject effects
library(ggridges)
ggplot(subject_preds,
        aes(x = subj_mean, y = .category, fill = refClass)) +
    geom_density_ridges(alpha = 0.5) +
    geom_vline(data = pop_preds,
                aes(xintercept = pop_mean),
                linetype = "dashed") +
    facet_wrap(~refClass) +
    theme_minimal() +
    labs(title = "Distribution of Subject-Specific Effects by Accuracy Level",
        x = "Predicted Probability",
        y = "Accuracy Level")
```

```{r}

library(tidybayes)
library(stringr)

predicted_probs_s1 <- ordinal_model_s1 %>%
    # gather draws that contain refClass ("b_refClassPercentage"    "b_refClassUSD"  )
    gather_draws(b_refClassPercentage, b_refClassUSD) %>%
    mutate(refClass = case_when(
    str_detect(.variable, "Percentage") ~ "Percentage",
    str_detect(.variable, "USD") ~ "USD",
    TRUE ~ "kWh"
  )) %>%
    group_by(refClass) %>%
    summarise(effect = mean(.value)) %>%
    mutate(
        group = refClass,
        prob_exact = plogis(effect + (-1.85)),
        prob_0_2 = plogis(effect + 0.22) - plogis(effect + (-1.85)),
        prob_2_15 = plogis(effect + 2.86) - plogis(effect + 0.22),
        prob_over15 = 1 - plogis(effect + 2.86)
    ) %>%
    pivot_longer(
        cols = starts_with("prob"),
        names_to = "Accuracy_Level",
        values_to = "Probability"
    ) %>%
    mutate(
        Accuracy_Level = recode(Accuracy_Level,
                                "prob_exact" = "Exact match",
                                "prob_0_2" = "0.01-2% error",
                                "prob_2_15" = "2.01-15% error",
                                "prob_over15" = "Over 15% error")
    )

ggplot(predicted_probs_s1, aes(x = Accuracy_Level, y = Probability, fill = refClass)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Predicted Probabilities of Accuracy Levels by Reference Class",
       x = "Accuracy Level",
       y = "Probability",
       fill = "Reference Class") +
  theme_minimal()

# Calculate Odds Ratios
odds_ratios_s1 <- posterior_samples(ordinal_model_s1) %>%
  mutate(
    OR_Percentage = exp(`b_refClassPercentage`),
    OR_USD = exp(`b_refClassUSD`)
  ) %>%
  select(OR_Percentage, OR_USD) %>%
  summarise_all(list(mean = mean, lower = ~ quantile(., 0.025), upper = ~ quantile(., 0.975))) %>%
  pivot_longer(cols = everything(),
               names_to = c("Reference Class", ".value"),
               names_pattern = "OR_(.*)_(.*)") %>%
  mutate(
    Reference_Class = case_when(
      str_detect(name, "Percentage") ~ "Percentage",
      str_detect(name, "USD") ~ "USD",
      TRUE ~ "kWh"
    )
  )

odds_ratios_s1 <- posterior_summary(ordinal_model_s1, prob = 0.95) %>%
  as.data.frame() %>%
  rownames_to_column(var = "Term") %>%
  filter(str_detect(Term, "refClass")) %>%
  mutate(
    OR = exp(Estimate),
    lower = exp(`Q2.5`),
    upper = exp(`Q97.5`)
  ) %>%
  select(Term, OR, lower, upper)

# Display Odds Ratios
odds_ratios_s1 %>%
  rename(`Reference Class` = Term,
         `Odds Ratio` = OR,
         `95% CrI Lower` = lower,
         `95% CrI Upper` = upper) %>%
  mutate(`Reference Class` = str_replace(`Reference Class`, "refClass", "")) %>%
  kable(booktabs = TRUE, caption = "Study 1: Odds Ratios for Matching Goals by Reference Class") %>%
  kable_styling(full_width = FALSE)

library(caret)

# Create confusion matrix
observed <- s1_agg$accuracy_level
predicted <- predict(ordinal_model_s1, newdata = s1_agg, type = "response")
predicted_class <- apply(predicted, 1, function(x) which.max(x))

confusion_matrix <- confusionMatrix(
  factor(predicted_class, levels = 1:4, labels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error")),
  factor(observed, levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"))
)

# Display Confusion Matrix
confusion_matrix$table %>%
  as.data.frame.matrix() %>%
  rownames_to_column(var = "Predicted") %>%
  kable(booktabs = TRUE, caption = "Study 1: Confusion Matrix for Matched Goals") %>%
  kable_styling(full_width = FALSE)

```

```{r}

ordinal_model_s1_prb <- brm(
  accuracy_level ~ refClass + (1|id) + (1|state),
  data = s1_agg,
  family = cumulative("probit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 2), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s1_ordinal_probit1.rds')) # Cache for efficiency
)

summary(ordinal_model_s1_prb)

pp_check(ordinal_model_s1_prb, type = "bars_grouped", group="refClass", fatten = 2) +
    scale_x_continuous("y", breaks = 1:7) +
    scale_y_continuous(NULL, breaks = NULL, expand = expansion(mult = c(0, 0.05))) +
    ggtitle("Data with posterior predictions",
            subtitle = expression(list(italic(N[A])==44, italic(N[B])==44))) +
    theme(legend.background = element_blank(),
            legend.position = c(.9, .75))

loo_logit <- loo(ordinal_model_s1)
loo_probit <- loo(ordinal_model_s1_prb)
loo_compare(loo_logit, loo_probit)

waic(ordinal_model_s1)
waic(ordinal_model_s1_prb)

# compare
loo_compare(loo(ordinal_model_s1), loo(ordinal_model_s1_prb))
compare_models(ordinal_model_s1, ordinal_model_s1_prb)

# Obtain predicted probabilities
predicted_probs_s1 <- ordinal_model_s1_prb %>%
    gather_draws(b_refClassPercentage, b_refClassUSD) %>%
    mutate(refClass = case_when(
        str_detect(.variable, "Percentage") ~ "Percentage",
        str_detect(.variable, "USD") ~ "USD",
        TRUE ~ "kWh"
    )) %>%
    group_by(refClass) %>%
    summarise(effect = mean(.value)) %>%
    mutate(
        group = refClass,
        prob_exact = plogis(effect + (-1.85)),
        prob_0_2 = plogis(effect + 0.22) - plogis(effect + (-1.85)),
        prob_2_15 = plogis(effect + 2.86) - plogis(effect + 0.22),
        prob_over15 = 1 - plogis(effect + 2.86)
    ) %>%
    pivot_longer(
        cols = starts_with("prob"),
        names_to = "Accuracy_Level",
        values_to = "Probability"
    ) %>%
    mutate(
        Accuracy_Level = recode(Accuracy_Level,
                                "prob_exact" = "Exact match",
                                "prob_0_2" = "0.01-2% error",
                                "prob_2_15" = "2.01-15% error",
                                "prob_over15" = "Over 15% error")
    )

    # Plot predicted probabilities
    ggplot(predicted_probs_s1, aes(x = Accuracy_Level, y = Probability, fill = refClass)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(title = "Predicted Probabilities of Accuracy Levels by Reference Class",
        x = "Accuracy Level",
        y = "Probability",
        fill = "Reference Class") +
    theme_minimal()

# Predicted probabilities from logit model
pred_probs_logit <- posterior_epred(ordinal_model_s1, newdata = pred_data, re_formula = NA)

# Predicted probabilities from probit model
pred_probs_probit <- posterior_epred(ordinal_model_s1_prb, newdata = pred_data, re_formula = NA)

# Compare predicted probabilities
mean_pred_logit <- apply(pred_probs_logit, 2, mean)
mean_pred_probit <- apply(pred_probs_probit, 2, mean)

comparison <- data.frame(
  Reference_Class = pred_data$refClass,
  Mean_Prob_Logit = mean_pred_logit,
  Mean_Prob_Probit = mean_pred_probit
)

print(comparison)

```

# Study 2

```{r}

s2_agg1 <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id, state,refClass,pct,pct_goal,plan,rounded) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(matched_goal = (pct_change == pct),
                error = pct_change - pct,
                abs_error = abs(error)) |> 
      ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        )

ggplot(data = s2_agg1, aes(x = accuracy_level, fill = refClass)) +
  geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Accuracy Levels by Reference Class",
       x = "Accuracy Level",
       y = "Percentage of Participants",
       fill = "Reference Class") +
  theme_minimal()

observed_props2 <- s2_agg1 |>
  group_by(refClass, accuracy_level) |>
  summarise(n = n()) |>
  group_by(refClass) |>
  mutate(prop = n/sum(n)) |>
  pivot_wider(
    names_from = accuracy_level,
    values_from = c(n, prop)
  )

  # Calculate proportions
prop_acc_s2 <- s2_agg1 %>%
    group_by(refClass, accuracy_level) %>%
    summarise(count = n()) %>%
    group_by(refClass) %>%
    mutate(Probability = count / sum(count)) %>%
    ungroup()

# Plot
ggplot(prop_acc_s2, aes(x = accuracy_level, y = Probability, fill = refClass)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    labs(title = "Distribution of Accuracy Levels by Reference Class",
        x = "Accuracy Level",
        y = "Probability",
        fill = "Reference Class") +
    theme_minimal() +
    scale_fill_brewer(palette = "Set1")

ggplot(data = s2_agg1, aes(x = accuracy_level, fill = refClass)) +
  geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Accuracy Levels by Reference Class",
       x = "Accuracy Level",
       y = "Percentage of Participants",
       fill = "Reference Class") +
  theme_minimal() + facet_wrap(~rounded+pct_goal)

prop4_acc_s2 <- s2_agg1 %>%
    group_by(refClass, accuracy_level,rounded,pct_goal) %>%
    summarise(count = n()) %>%
    group_by(refClass) %>%
    mutate(Probability = count / sum(count)) %>%
    ungroup()

# Plot

s2_agg1 %>%
    group_by( accuracy_level,rounded) %>%
    summarise(count = n()) %>%
    group_by(rounded) %>%
    mutate(Probability = count / sum(count)) %>%
    ungroup()

ggplot(prop4_acc_s2, aes(x = accuracy_level, y = Probability, fill = refClass)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    labs(title = "Distribution of Accuracy Levels by Reference Class",
        x = "Accuracy Level",
        y = "Probability",
        fill = "Reference Class") +
    theme_minimal() +
    scale_fill_brewer(palette = "Set1") + facet_wrap(~rounded+pct_goal)

ggplot(prop4_acc_s2, aes(x = accuracy_level, y = Probability, fill = refClass)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    labs(title = "Distribution of Accuracy Levels by Reference Class",
        x = "Accuracy Level",
        y = "Probability",
        fill = "Reference Class") +
    theme_minimal() +
    scale_fill_brewer(palette = "Set1")

```

# no interaction model used in MPA
```{r}

s2_agg1 <- s2_agg1 %>%
  mutate(
    refClass = factor(refClass, levels = c("kWh", "Percentage", "USD")),
    rounded = factor(rounded, levels = c("Not Rounded", "Rounded")),
    pct_goal = factor(pct_goal, levels = c("10%", "15%"))
  )

ordinal_model_s2_probit <- brm(
  accuracy_level ~ refClass +rounded+pct_goal + (1|id)+ (1|state),
  data = s2_agg1,
  family = cumulative("probit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 2), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_op.rds')) # Cache for efficiency
)

summary(ordinal_model_s2_probit)
# Regression Coefficients:
#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept[1]          -1.12      0.38    -1.88    -0.37 1.01      686     1070
# Intercept[2]          -0.28      0.38    -1.04     0.47 1.01      693     1085
# Intercept[3]           1.81      0.38     1.04     2.56 1.01      702     1248
# refClassPercentage     0.58      0.45    -0.32     1.49 1.01      446      741
# refClassUSD            1.19      0.47     0.29     2.11 1.01      583     1087
# roundedRounded        -0.37      0.10    -0.57    -0.17 1.00     5613     2928
# pct_goal15%           -0.26      0.10    -0.46    -0.06 1.00     5136     2842

pred_summary_s2 <- ordinal_model_s2_probit %>%
  epred_draws(newdata = s2_agg1, re_formula = NA) %>%
  group_by(refClass, rounded, pct_goal, .category) %>%
  summarise(
    mean_prob = mean(.epred),
    lower_ci = quantile(.epred, 0.025),
    upper_ci = quantile(.epred, 0.975),
    .groups = "drop"
  )

pred_summary_s2 |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy level by reference class")

# Extract posterior samples
posterior_samples_s2 <- as.data.frame(ordinal_model_s2_probit)

odds_ratios_s2 <- posterior_samples_s2 %>%
    transmute(
        refClass_Percentage_vs_kWh = exp(b_refClassPercentage),
        refClass_USD_vs_kWh = exp(b_refClassUSD),
        rounded_Yes_vs_No = exp(b_roundedRounded),
        pct_goal_15_vs_10 = exp(`b_pct_goal15%`)  # Note the backticks here
    )

# Calculate summary statistics
odds_ratio_summary_s2 <- data.frame(
    comparison = c("Percentage vs kWh", "USD vs kWh", "Rounded vs Not", "15% Goal vs 10% Goal"),
    odds_ratio = c(
        mean(odds_ratios_s2$refClass_Percentage_vs_kWh),
        mean(odds_ratios_s2$refClass_USD_vs_kWh),
        mean(odds_ratios_s2$rounded_Yes_vs_No),
        mean(odds_ratios_s2$pct_goal_15_vs_10)
    ),
    ci_lower = c(
        quantile(odds_ratios_s2$refClass_Percentage_vs_kWh, 0.025),
        quantile(odds_ratios_s2$refClass_USD_vs_kWh, 0.025),
        quantile(odds_ratios_s2$rounded_Yes_vs_No, 0.025),
        quantile(odds_ratios_s2$pct_goal_15_vs_10, 0.025)
    ),
    ci_upper = c(
        quantile(odds_ratios_s2$refClass_Percentage_vs_kWh, 0.975),
        quantile(odds_ratios_s2$refClass_USD_vs_kWh, 0.975),
        quantile(odds_ratios_s2$rounded_Yes_vs_No, 0.975),
        quantile(odds_ratios_s2$pct_goal_15_vs_10, 0.975)
    )
)

odds_ratio_summary_s2
#             comparison odds_ratio ci_lower ci_upper
# 1    Percentage vs kWh       1.99     0.73     4.43
# 2           USD vs kWh       3.69     1.33     8.29
# 3       Rounded vs Not       0.70     0.57     0.84
# 4 15% Goal vs 10% Goal       0.78     0.63     0.94

ggplot(pred_summary, 
        aes(x = .category, y = mean_prob, fill = refClass )) +
    geom_col(position = position_dodge(width = 0.8)) +
    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), 
                    position = position_dodge(width = 0.8)) +
    labs(x = "Reference Class", 
        y = "Predicted Probability",
        fill = "Accuracy Level",
        title = "Predicted Probabilities of Accuracy Levels by Reference Class") +
    scale_fill_brewer(palette = "RdYlBu") +
    theme_minimal()

```

```{r}

# Fit the full model with interactions
ordinal_model_s2 <- brm(
  accuracy_level ~ refClass * rounded * pct_goal + (1|id) + (1|state),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99),
  prior = c(
    prior(normal(0, 2), class = "Intercept"),
    prior(normal(0, 2), class = "b")
  ),
  file = "s2_ordinal_full6.rds"
)

pred_summary_s2 <- ordinal_model_s2 %>%
  epred_draws(newdata = s2_agg1, re_formula = NA) %>%
  group_by(refClass, rounded, pct_goal, .category) %>%
  summarise(
    mean_prob = mean(.epred),
    lower_ci = quantile(.epred, 0.025),
    upper_ci = quantile(.epred, 0.975),
    .groups = "drop"
  )

# Extract posterior samples
posterior_samples_s2 <- as.data.frame(ordinal_model_s2)

odds_ratios_s2 <- posterior_samples_s2 %>%
    transmute(
        refClass_Percentage_vs_kWh = exp(b_refClassPercentage),
        refClass_USD_vs_kWh = exp(b_refClassUSD),
        rounded_Yes_vs_No = exp(b_roundedRounded),
        pct_goal_15_vs_10 = exp(`b_pct_goal15%`)  # Note the backticks here
    )
# Calculate summary statistics
odds_ratio_summary_s2 <- data.frame(
    comparison = c("Percentage vs kWh", "USD vs kWh", "Rounded vs Not", "15% Goal vs 10% Goal"),
    odds_ratio = c(
        mean(odds_ratios_s2$refClass_Percentage_vs_kWh),
        mean(odds_ratios_s2$refClass_USD_vs_kWh),
        mean(odds_ratios_s2$rounded_Yes_vs_No),
        mean(odds_ratios_s2$pct_goal_15_vs_10)
    ),
    ci_lower = c(
        quantile(odds_ratios_s2$refClass_Percentage_vs_kWh, 0.025),
        quantile(odds_ratios_s2$refClass_USD_vs_kWh, 0.025),
        quantile(odds_ratios_s2$rounded_Yes_vs_No, 0.025),
        quantile(odds_ratios_s2$pct_goal_15_vs_10, 0.025)
    ),
    ci_upper = c(
        quantile(odds_ratios_s2$refClass_Percentage_vs_kWh, 0.975),
        quantile(odds_ratios_s2$refClass_USD_vs_kWh, 0.975),
        quantile(odds_ratios_s2$rounded_Yes_vs_No, 0.975),
        quantile(odds_ratios_s2$pct_goal_15_vs_10, 0.975)
    )
)

odds_ratio_summary_s2
```

            comparison odds_ratio ci_lower ci_upper
1    Percentage vs kWh       3.31     0.43     11.3
2           USD vs kWh       6.64     0.84     25.2
3       Rounded vs Not       0.41     0.09      1.1
4 15% Goal vs 10% Goal       0.51     0.12      1.5

```{r}

s2_agg1 <- s2_agg1 %>%
  mutate(
    refClass = factor(refClass, levels = c("kWh", "Percentage", "USD")),
    rounded = factor(rounded, levels = c("Not Rounded", "Rounded")),
    pct_goal = factor(pct_goal, levels = c("10%", "15%"))
  )

# Fit the full model with interactions
ordinal_model_s2 <- brm(
  accuracy_level ~ refClass * rounded * pct_goal + (1|id) + (1|state),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99),
  prior = c(
    prior(normal(0, 2), class = "Intercept"),
    prior(normal(0, 2), class = "b")
  ),
  file = "s2_ordinal_full6.rds"
)

pred_summary_s2 <- ordinal_model_s2 %>%
  epred_draws(newdata = s2_agg1, re_formula = NA) %>%
  group_by(refClass, rounded, pct_goal, .category) %>%
  summarise(
    mean_prob = mean(.epred),
    lower_ci = quantile(.epred, 0.025),
    upper_ci = quantile(.epred, 0.975),
    .groups = "drop"
  )

# Extract posterior samples
posterior_samples_s2 <- as.data.frame(ordinal_model_s2)

# Compute odds ratios for the predictors
odds_ratios_s2 <- posterior_samples_s2 %>%
  transmute(
    refClass_Percentage_vs_kWh = exp(b_refClassPercentage),
    refClass_USD_vs_kWh = exp(b_refClassUSD),
    rounded_Yes_vs_No = exp(b_roundedRounded),
    pct_goal_15_vs_10 = exp(`b_pct_goal15%`)  # Note the backticks here
  )

# Summarize odds ratios
odds_ratio_summary_s2 <- odds_ratios_s2 %>%
  summarise(
    Percentage_vs_kWh = mean(refClass_Percentage_vs_kWh),
    USD_vs_kWh = mean(refClass_USD_vs_kWh),
    Rounded_vs_Not = mean(rounded_Yes_vs_No),
    Goal15_vs_Goal10 = mean(pct_goal_15_vs_10)
  )

odds_ratio_summary_s2

posterior_samples_s2 <- as.data.frame(ordinal_model_s2)

odds_ratio_summary_s2 <- data.frame(
  comparison = c("Percentage vs kWh", "USD vs kWh", "Rounded vs Not", "15% Goal vs 10% Goal"),
  odds_ratio = c(
    mean(odds_ratios_s2$Percentage_vs_kWh),
    mean(odds_ratios_s2$USD_vs_kWh),
    mean(odds_ratios_s2$Rounded_vs_Not),
    mean(odds_ratios_s2$Goal15_vs_10)
  ),
  ci_lower = c(
    quantile(odds_ratios_s2$Percentage_vs_kWh, 0.025),
    quantile(odds_ratios_s2$USD_vs_kWh, 0.025),
    quantile(odds_ratios_s2$Rounded_vs_Not, 0.025),
    quantile(odds_ratios_s2$Goal15_vs_10, 0.025)
  ),
  ci_upper = c(
    quantile(odds_ratios_s2$Percentage_vs_kWh, 0.975),
    quantile(odds_ratios_s2$USD_vs_kWh, 0.975),
    quantile(odds_ratios_s2$Rounded_vs_Not, 0.975),
    quantile(odds_ratios_s2$Goal15_vs_10, 0.975)
  )
)

```

### 2024-11-07 overview.qmd
finished mpa abstract

---
title: "Effect of Reference Class on Planning to Conserve"
author:
- name: Thomas E. Gorman
  url: https://tegorman13.github.io/
  affiliations: 
  - name:  Communication and Cognition Lab, Purdue University, USA
    affiliation-url: https://web.ics.purdue.edu/~treimer/
#lightbox: true
toc: false
code-fold: true
code-tools: true
execute:
  #echo: false
  warning: false
format:
  html: 
    grid:
      sidebar-width: 220px
      body-width: 1200px
      margin-width: 170px
      gutter-width: 1.0rem
  hugo-md:
    include: true
    html-math-method: mathjax
    output-file: overview_hugo.md
  gfm:
    echo: true
    output-file: overview_gfm.md
---

```{r}

pacman::p_load(dplyr,purrr,tidyr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,
               ggplot2,gt,knitr,kableExtra,ggh4x,lme4,flextable,pander)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("fun_plot"), ~ source(here::here(paste0("Scripts/", .x, ".R"))))
theme_set(theme_nice())

s1 <- readRDS(here::here("data/s1_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds"))))
s2_long <- readRDS(here::here("data/s2_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds"))))

state_code_map <- c(
  "CAL" = "California",
  "TEX" = "Texas",
  "COL" = "Colorado",
  "MASS" = "Massachusetts"
)

state_to_code <- state_code_map
code_to_state <- setNames(names(state_code_map), state_code_map)
lookup_code <- function(state) {
  if (state %in% names(state_to_code)) {
    return(state_to_code[[state]])
  } else if (state %in% names(code_to_state)) {
    return(code_to_state[[state]])
  } else {
    return(NA)  # Return NA if not found in either mapping
  }
}

```

```{r}

s1 |> group_by(refClass) |> 
    rename("Reference Class" = refClass) |>
    summarise(n=n_distinct(id)) |> 
    pander::pandoc.table(caption="Study 1: Number of unique IDs per condition")

s1 |> group_by(refClass,state) |> 
    rename("Reference Class" = refClass) |>
    summarise(n=n_distinct(id)) |> 
    rowwise() |> 
    mutate(state=lookup_code(state)) |> 
    pivot_wider(names_from=state,values_from=n) |>
    pander::pandoc.table(caption="Study 1: Number of unique IDs per condition by state")

s1 |> group_by(refClass,calc) |> 
    rename("Reference Class" = refClass) |>
    summarise(n=n_distinct(id)) |> 
    pivot_wider(names_from=calc,values_from=n) |>
    pander::pandoc.table(caption="Study 1: Number of unique IDs per condition by calculator use")

prop.table(xtabs(~refClass,s1)) |> kable()
prop.table(xtabs(~refClass+calc,s1)) |> kable()
prop.table(xtabs(~refClass+state,s1)) |> kable()

```

### Study 2 - Subject counts 
```{r}

s2_long |> group_by(refClass) |>
    rename("Reference Class" = refClass) |>
    summarise(n=n_distinct(id)) |> 
    pander::pandoc.table(caption="Study 2: Number of unique IDs per condition")

s2_long |> group_by(refClass,calc) |>
    rename("Reference Class" = refClass) |>
    mutate(calc=case_when(calc=="Calculator" ~ "Used Calculator",
                        calc=="No Calculator" ~ "No Calculator")) |>
    summarise(n=n_distinct(id)) |>
    pivot_wider(names_from=calc,values_from=n) |>
    pander::pandoc.table(caption="Study 2: Number of unique IDs per condition by calculator use")

prop.table(xtabs(~refClass,s2_long)) |> kable()
prop.table(xtabs(~refClass+state,s2_long))
prop.table(xtabs(~refClass+calc,
s2_long |> mutate(calc=case_when(calc=="Calculator" ~ "Used Calculator",
                      calc=="No Calculator" ~ "No Calculator")))) |> kable()
```

## Performance

### Study 1

```{r}
#| fig-width: 12
#| fig-height: 8

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                close_match = abs_error <= 0.03)

s1_agg4 <- s1_agg |> group_by(id,refClass) |> 
    summarise(mg=sum(matched_goal),n=n(), pct=mg/n)

# overall pct of subjects who matched their goal
s1_agg4 |> ungroup() |>
    summarise(
    mean = mean(pct),
    sd = sd(pct),
    n = n(),
    se=sd(pct)/sqrt(n)
) |> pander::pandoc.table(caption="Study 1: Proportion of participants who matched their goal overall")

# proportion matching vs close match

s1_agg |> pivot_longer(c(matched_goal,close_match),names_to="match_type",values_to="match") |> 
    group_by(id,match_type) |>
    summarise(mg=sum(match),n=n(), pct=mg/n) |>
    group_by(match_type) |>
    summarise(
    mean = mean(pct),
    sd = sd(pct),
    n = n(),
    se=sd(pct)/sqrt(n)
) |> pander::pandoc.table(caption="Study 1: Proportion of participants who matched their goal or were within 3% of their goal")

# refClass
s1_agg4 |> 
group_by(refClass) |> 
summarise(
    mean = mean(pct),
    sd = sd(pct),
    n = n(),
    se=sd(pct)/sqrt(n)
) |> pander::pandoc.table(caption="Study 1: Proportion of participants who matched their goal by condition")

```

### Study 2

```{r}

s2_agg1 <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id, state,refClass,pct,pct_goal,plan,rounded) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(matched_goal = (pct_change == pct),
  close_match = abs(pct_change-pct)<.03)

s2_agg4 <- s2_agg1|> group_by(id,refClass) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n)

# overall pct of subjects who matched their goal
s2_agg4 |> ungroup() |>
    summarise(
    mean = mean(pct),
    sd = sd(pct),
    n = n(),
    se=sd(pct)/sqrt(n)
) |> pander::pandoc.table(caption="Study 2: Proportion of participants who matched their goal overall")

# proportion matching vs close match

s2_agg1 |> pivot_longer(c(matched_goal,close_match),names_to="match_type",values_to="match") |> 
    group_by(id,match_type) |>
    summarise(mg=sum(match),n=n(), pct=mg/n) |>
    group_by(match_type) |>
    summarise(
    mean = mean(pct),
    sd = sd(pct),
    n = n(),
    se=sd(pct)/sqrt(n)
) |> pander::pandoc.table(caption="Study 2: Proportion of participants who matched their goal or were within 3% of their goal")

# refClass only 
s2_agg4 |> 
    group_by(refClass) |> 
    summarise(
    mean = mean(pct),
    sd = sd(pct),
    n = n(),
    se=sd(pct)/sqrt(n)
) |> pander::pandoc.table(caption="Study 2: Proportion of participants who matched their goal by condition")

# rounded only

s2_agg1 |> group_by(id,rounded) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
    group_by(rounded) |>
    summarise(
    mean = mean(pct),
    sd = sd(pct),
    n = n(),
    se=sd(pct)/sqrt(n)
) |> pander::pandoc.table(caption="Study 2: Proportion of participants who matched their goal by rounding condition")

# pct_goal only

s2_agg1 |> group_by(id,pct_goal) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
    group_by(pct_goal) |>
    summarise(
    mean = mean(pct),
    sd = sd(pct),
    n = n(),
    se=sd(pct)/sqrt(n)
) |> pander::pandoc.table(caption="Study 2: Proportion of participants who matched their goal by goal condition")

# refClass x rounded

s2_agg1 |> group_by(id, refClass,rounded) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
    group_by(refClass,rounded) |>
    summarise(
    mean = mean(pct),
    sd = sd(pct),
    n = n(),
    se=sd(pct)/sqrt(n)
) |> pander::pandoc.table(caption="Study 2: Proportion of participants who matched their goal by condition and rounding")

# refClass x pct_goal

s2_agg1 |> group_by(id, refClass,pct_goal) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
    group_by(refClass,pct_goal) |>
    summarise(
    mean = mean(pct),
    sd = sd(pct),
    n = n(),
    se=sd(pct)/sqrt(n)
) |> pander::pandoc.table(caption="Study 2: Proportion of participants who matched their goal by condition and goal")

# pct_goal x rounded

s2_agg1 |> group_by(id, pct_goal,rounded) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
    group_by(pct_goal,rounded) |>
    summarise(
    mean = mean(pct),
    sd = sd(pct),
    n = n(),
    se=sd(pct)/sqrt(n)
) |> pander::pandoc.table(caption="Study 2: Proportion of participants who matched their goal by goal and rounding")

# pct_goal x rounded x refClass

s2_agg1 |> group_by(id, pct_goal,rounded,refClass) |> 
  summarise(mg=sum(matched_goal),n=n(), pct=mg/n) |> 
    group_by(pct_goal,rounded,refClass) |>
    summarise(
    mean = mean(pct),
    sd = sd(pct),
    n = n(),
    se=sd(pct)/sqrt(n)
) |> pander::pandoc.table(caption="Study 2: Proportion of participants who matched their goal by goal, rounding, and condition")

```

## Analysis

### Study 1

```{r}

# binomial regression on trial level data

s1_br <- brm(
    matched_goal ~ refClass + (1|id),
    data = s1_agg,
    family = bernoulli(),
    cores=4,
    iter=2000,
    control = list(adapt_delta = 0.99),
    file=paste0(here::here("data/model_cache",'s1_br.rds')),
    )

summary(s1_br)

mr_s1 <- as.data.frame(describe_posterior(s1_br, centrality = "Mean"))[, c(1,2,4,5,6)]
colnames(mr_s1) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")

intercept <- mr_s1$Estimate[mr_s1$Term == "b_Intercept"]

 mr_s1 |>
    mutate(Term = stringr::str_remove(Term, "b_")) |>
    mutate(across(c("Estimate", "95% CrI Lower", "95% CrI Upper"), 
                    \(x) if_else(Term == "Intercept", plogis(x), plogis(x + intercept)))) |>
    mutate(across(where(is.numeric), \(x) round(x, 3))) |>
    mutate(Term = if_else(Term == "Intercept", "Intercept (kWh)", Term)) |>
    tibble::remove_rownames() |>
    kable(booktabs = TRUE)

s1_br2 <- brm(
    close_match ~ refClass + (1|id) + (1|state),
    data = s1_agg,
    family = bernoulli(),
    cores=4,
    iter=2000,
    control = list(adapt_delta = 0.99),
    file=paste0(here::here("data/model_cache",'s1_br2.rds')),
    )
summary(s1_br2)

s1_br2 |> emmeans(~refClass, type="response")

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                close_match = abs_error <= 0.03) |>
    ungroup() |> # Add ungroup here
    mutate(
        accuracy_level = factor(
            case_when(
                abs_error <= 0.01 ~ "1% or less",
                abs_error <= 0.05 ~ "5% or less",
                abs_error <= 0.10 ~ "10% or less",
                TRUE ~ "Greater than 10%"  # Capture all remaining cases
            ), 
            levels = c("1% or less", "5% or less", "10% or less", "Greater than 10%"),
            ordered = TRUE
        )
    )

```

## Ordinal Regression alternative

```{r}

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                close_match = abs_error <= 0.03) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.10 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        )
    # mutate(
    #     accuracy_level = factor(
    #         case_when(
    #             abs_error <= 0.01 ~ "1% or less",
    #             abs_error <= 0.05 ~ "5% or less",
    #             abs_error <= 0.10 ~ "10% or less",
    #             TRUE ~ "Greater than 10%"  # Capture all remaining cases
    #         ), 
    #         levels = c("1% or less", "5% or less", "10% or less", "Greater than 10%"),
    #         ordered = TRUE
    #     )
    # )

# plot accuracy level by refClass

s1_agg |> 
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.10 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        ) |>
    ggplot(aes(x=refClass,fill=accuracy_level)) +
    geom_bar(position="fill") +
    labs(title="Study 1: Accuracy Level by Reference Class",
        x="Reference Class",
        y="Proportion",
        fill="Accuracy Level") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

ordinal_model_s1 <- brm(
  accuracy_level ~ refClass + (1|id) + (1|state),
  data = s1_agg,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 1.5), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s1_ordinal3.rds')) # Cache for efficiency
)

pred_summary <- ordinal_model_s1 |>
    epred_draws(newdata = data.frame(refClass = c("kWh", "Percentage", "USD")),
                ndraws = 1000, re_formula = NA) |>
    group_by(refClass, .category) |>
    summarise(
        mean_prob = mean(.epred),
        lower_ci = quantile(.epred, 0.025),
        upper_ci = quantile(.epred, 0.975)
    )
pred_summary |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy level by reference class")

ggplot(pred_summary, 
       aes(x = refClass, y = mean_prob, fill = .category)) +
  geom_col(position = "stack") +
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), 
                position = position_stack()) +
  labs(x = "Reference Class", 
       y = "Predicted Probability",
       fill = "Accuracy Level",
       title = "Predicted Probabilities of Accuracy Levels by Reference Class") +
  theme_minimal() +
  scale_fill_brewer(palette = "Blues")

# Calculate cumulative probabilities
cum_probs <- pred_summary |>
  group_by(refClass) |>
  arrange(refClass, .category) |>
  mutate(
    cum_prob = cumsum(mean_prob),
    cum_lower = cumsum(lower_ci),
    cum_upper = cumsum(upper_ci)
  )

ggplot(cum_probs, 
    aes(x = refClass, y = cum_prob, color = .category)) +
geom_point(size = 3,position=position_dodge()) +
geom_errorbar(aes(ymin = cum_lower, ymax = cum_upper), 
            width = 0.2,position=position_dodge()) +
labs(x = "Reference Class",
    y = "Cumulative Probability",
    color = "Accuracy Level",
    title = "Cumulative Probabilities of Accuracy Levels") +
theme_minimal() +
scale_color_brewer(palette = "Blues")

summary_table <- pred_summary |>
  pivot_wider(
    names_from = .category,
    values_from = c(mean_prob, lower_ci, upper_ci)
  ) |>
  kable(digits = 3)

print(summary_table)

# Convert log-odds to odds ratios
posterior_samples <- as.data.frame(ordinal_model_s1)
odds_ratios <- data.frame(
  Percentage_vs_kWh = exp(posterior_samples$b_refClassPercentage),
  USD_vs_kWh = exp(posterior_samples$b_refClassUSD)
)

# Calculate summary statistics
odds_ratio_summary <- data.frame(
  comparison = c("Percentage vs kWh", "USD vs kWh"),
  odds_ratio = c(mean(odds_ratios$Percentage_vs_kWh),
                 mean(odds_ratios$USD_vs_kWh)),
  ci_lower = c(quantile(odds_ratios$Percentage_vs_kWh, 0.025),
               quantile(odds_ratios$USD_vs_kWh, 0.025)),
  ci_upper = c(quantile(odds_ratios$Percentage_vs_kWh, 0.975),
               quantile(odds_ratios$USD_vs_kWh, 0.975))
)

# Plot predicted probabilities
ggplot(pred_summary, aes(x = refClass, y = mean_prob, fill = .category)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci),
                position = position_dodge(width = 0.9), width = 0.2) +
  labs(y = "Predicted Probability", x = "Reference Class", fill = "Accuracy Level") +
  ggtitle("Study 1: Predicted Probabilities of Accuracy Level by Reference Class") +
  theme_minimal()

```

#### Frequentist alternative

```{r}
s1_glm <- glm(matched_goal ~ refClass, data = s1_agg, family = binomial())
summary(s1_glm)

# Odds Ratios and Confidence Intervals
exp(cbind(Odds_Ratio = coef(s1_glm), confint(s1_glm)))

# Chi-Square Test for Proportions in Study 1
s1_table <- table(s1_agg$refClass, s1_agg$matched_goal)
chisq.test(s1_table)

```

### Study 2

```{r}

# binomial regression on trial level data

s2_br_3 <- brm(
    matched_goal ~ refClass+rounded+pct_goal + (1|id) + (1|state),
    data = s2_agg1,
    family = bernoulli(),
    cores=4,
    iter=2000,
    control = list(adapt_delta = 0.99),
    file=paste0(here::here("data/model_cache",'s2_br_3.rds')),
    )

summary(s2_br_3)

mr_s2_3 <- as.data.frame(describe_posterior(s2_br_3, centrality = "Mean"))[, c(1,2,4,5,6)]
colnames(mr_s2_3) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")

 mr_s2_3 |>
    mutate(Term = stringr::str_remove(Term, "b_")) |>
    mutate(across(where(is.numeric), \(x) round(x, 3))) |>
    mutate(Term = if_else(Term == "Intercept", "Intercept (kWh)", Term)) |>
    tibble::remove_rownames() |>
    kable(booktabs = TRUE)

s2_br2_3 <- brm(
    close_match ~ refClass+rounded+pct_goal + (1|id) + (1|state),
    data = s2_agg1,
    family = bernoulli(),
    cores=4,
    iter=2000,
    control = list(adapt_delta = 0.99),
    file=paste0(here::here("data/model_cache",'s2_br2_3.rds')),
    )

summary(s2_br2_3)

#s2_br2_3 |> emmeans(~refClass, type="response")

mr_s2_3_2 <- as.data.frame(describe_posterior(s2_br2_3, centrality = "Mean"))[, c(1,2,4,5,6)]
colnames(mr_s2_3_2) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")

mr_s2_3_2 |>
    mutate(Term = stringr::str_remove(Term, "b_")) |>
    mutate(across(where(is.numeric), \(x) round(x, 3))) |>
    mutate(Term = if_else(Term == "Intercept", "Intercept (kWh)", Term)) |>
    tibble::remove_rownames() |>
    kable(booktabs = TRUE)

```

```{r}

s2_glm <- glm(matched_goal ~ refClass + rounded + pct_goal, data = s2_agg1, family = binomial())
summary(s2_glm)

# Odds Ratios and Confidence Intervals
exp(cbind(Odds_Ratio = coef(s2_glm), confint(s2_glm)))

# Chi-Square Test for Proportions in Study 2
s2_table <- table(s2_agg1$refClass, s2_agg1$matched_goal)
chisq.test(s2_table)

```

### 2024-11-07 Scripts/fit_models.R

@@ -919,17 +919,17 @@ s1_zib_close_c <- brm(
    control = list(adapt_delta = .94, max_treedepth = 13),
   file=paste0(here::here("data/model_cache",'s1_zib_close_c.rds'))
 )
-summary(s1_zib_close_c)
-Regression Coefficients:
-                   Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
-Intercept             -0.88      0.93    -2.74     0.95 1.01      439     1421
-refClassPercentage    -1.90      0.78    -3.49    -0.43 1.01      783     1322
-refClassUSD           -3.78      0.80    -5.47    -2.30 1.00      908     1529
-calcUsedCalculator     2.59      0.76     1.16     4.10 1.01     1070     1957
-
-Further Distributional Parameters:
-   Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
-zi     0.13      0.07     0.01     0.27 1.02      216      774
+# summary(s1_zib_close_c)
+# Regression Coefficients:
+#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# Intercept             -0.88      0.93    -2.74     0.95 1.01      439     1421
+# refClassPercentage    -1.90      0.78    -3.49    -0.43 1.01      783     1322
+# refClassUSD           -3.78      0.80    -5.47    -2.30 1.00      908     1529
+# calcUsedCalculator     2.59      0.76     1.16     4.10 1.01     1070     1957
+# 
+# Further Distributional Parameters:
+#    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# zi     0.13      0.07     0.01     0.27 1.02      216      774
 
 s1_zib_close_c |> emmeans(~refClass, type="response")
 #  refClass   prob lower.HPD upper.HPD

### 2024-11-20 manuscript/manuscript.qmd
add manuscript, and restructure website

---
title: "Planning to Save Energy: How Information Format Affects Accuracy"
aliases:
  - /manuscript.html
author:
    - name: Thomas E. Gorman
      url: https://tegorman13.github.io/
      orcid: "0000-0001-5366-5442"
      affiliations: 
        - name:  Purdue University, USA
          city: West Lafayette
          region: IN
          country: USA
          postal-code: 47907
          affiliation-url: https://web.ics.purdue.edu/~treimer/
      attributes: 
          corresponding: true
    - name: Torsten Reimer
      affiliation: Brian Lamb School of Communication, Purdue University
      affiliation-url: https://cla.purdue.edu/communication/
      url: https://web.ics.purdue.edu/~treimer/
      orcid: 0000-0002-7419-0076
    - name: Juan Pablo Loaiza Ramirez
      affiliation: Brian Lamb School of Communication, Purdue University
      affiliation-url: https://cla.purdue.edu/communication/
      url: https://web.ics.purdue.edu/~treimer/
      orcid: 0000-0001-9663-0522
format:
    # html: 
    #     grid:
    #         sidebar-width: 220px
    #         body-width: 1200px
    #         margin-width: 170px
    #         gutter-width: 1.0rem
    wordcount-html: 
        #output-file: index.html
        wordcount-banner: true
        theme: 
            - zephyr
            - ../assets/custom.scss
        toc: true
        shift-heading-level-by: 1
        title-block-banner: "#000D51"
        include-in-header:
            - text: |
                <style type="text/css">
                h2.hide {
                    display: none;
                }
                </style>
    # hugo-md:
    #     include: true
    #     html-math-method: mathjax
    #     output-file: manuscript_hugo.md
    # gfm:
    #     echo: true
    #     output-file: manuscript_gfm.md
    hikmah-manuscript-pdf:
        output-file: rep_study.pdf
        mainfont: "Linux Libertine O"
        mainfontoptions:
        - "Numbers=Proportional"
        - "Numbers=OldStyle"
        mathfont: "Libertinus Math"
        knitr:
            opts_chunk:
                dev: "cairo_pdf"
    #hikmah-manuscript-docx: default
format-links:   
    - html
    - gfm
    - format: hikmah-manuscript-pdf
      text: PDF version
      icon: file-pdf
date: 2024-11-19
abstract: |
    This study examines how presenting energy information in different formats (kWh, %, USD) affects planning accuracy. Across two experiments, the kWh format generally led to better accuracy, while the USD format consistently led to the worst performance. These findings highlight the importance of effective information presentation to promote energy conservation.
---

```{r}
#| label: setup
#| include: false

knitr::opts_chunk$set(
  echo = FALSE, include = FALSE, warning = FALSE, message = FALSE,
  fig.width = 6, fig.height = (6 * 0.618), fig.retina = 3,
  out.width = "100%", fig.align = "center"
)
```

# Introduction

# Experiment 1

## Methods

![Example trial in the energy planning task. Participants are shown the prior year electricity use of a household, and are tasked with creating a plan for the next year that will meet the energy reduction goal. Study 1 manipulates the format of the reduction goal to be either a percentage (15% given as goal reduction), kilowatt hours (5965 kWh given), or USD ($656)](/assets/images/Smith_10R_Wells_15E.png){#fig-task}

## Results

# Experiment 2

## Methods

The questions in study 2 are quite similar to those in study 1, but we also included a rounding manipulation (rounded vs. not rounded), and a manipulation of the goal (10% reduction vs. 15% rediction)

Note that reference class remains a between-subjects variable, while percent goal, rounding, and state are within-subjects variables.

In study 2, the new design is a 4 state temperature (2 warm vs. 2 cold states) X 2 task goal (10% vs. 15%) X 2 last year’s usage for the family and the state average (exact vs. rounded numbers) within X 3 task reference class (USD vs. Percentage vs. kWh) between. A note on the percentage goals is that they are rounded to the full dollar. For instance, for the Davis family, 10% of their \$6,943.95 cost of residential energy usage is \$694.395; this is rounded to \$694. For the rounded condition, participants are also told that “The numbers for their \[the family’s\] usage have been rounded for convenience.” In terms of randomization, blocks were combined so that there was a block for each condition and order combination. As an example, one block is the smith family with a 10% goal and rounded numbers for the family’s usage first, and the Wells family second, with a 15% goal and exact numbers for the family’s usage. Another block has these but in reverse, where the participant receives the Wells family first and the Smith family second. In total, there are 96 conditions, representing each possible order (2 orders x 4 states x 2 possible significant digit values, and 3 reference classes).

## Results

# Discusion

### 2024-11-20 misc/indiv.qmd
add manuscript, and restructure website

@@
 title: "Response Patterns"
-author:
-- name: Thomas E. Gorman
-  url: https://tegorman13.github.io/
-  affiliations: 
-  - name:  Communication and Cognition Lab, Purdue University, USA
-    affiliation-url: https://web.ics.purdue.edu/~treimer/
-lightbox: true
-toc: false
-code-fold: true
-code-tools: true
 execute:

### 2024-11-20 misc/ordinal.qmd
add manuscript, and restructure website

@@
 title: "Ordinal Models"
-author:
-- name: Thomas E. Gorman
-  url: https://tegorman13.github.io/
-  affiliations: 
-  - name:  Communication and Cognition Lab, Purdue University, USA
-    affiliation-url: https://web.ics.purdue.edu/~treimer/
-#lightbox: true
-toc: false
-code-fold: true
-code-tools: true
 execute:
@@
 
-pacman::p_load(dplyr,purrr,tidyr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,
+pacman::p_load(dplyr,purrr,tidyr,here,tibble,stringr,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,
                ggplot2,gt,knitr,kableExtra,ggh4x,lme4,flextable,pander)
 options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
-walk(c("fun_plot"), ~ source(here::here(paste0("Scripts/", .x, ".R"))))
+walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
 theme_set(theme_nice())
@@
 s2_long <- readRDS(here::here("data/s2_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds"))))
-
 ```
@@
 
-library(tidybayes)
-
 # Get predicted probabilities
@@
     )
-pred_summary |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy level by reference class")
-
-ggplot(pred_summary, 
-        aes(x = .category, y = mean_prob, fill = refClass )) +
-    geom_col(position = position_dodge(width = 0.8)) +
-    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), 
-                    position = position_dodge(width = 0.8)) +
-    labs(x = "Reference Class", 
-        y = "Predicted Probability",
-        fill = "Accuracy Level",
-        title = "Predicted Probabilities of Accuracy Levels by Reference Class") +
-    scale_fill_brewer(palette = "RdYlBu") +
-    theme_minimal()
-
-ggplot(pred_summary, 
-        aes(x = refClass, y = mean_prob, fill = .category )) +
-    geom_col(position = position_dodge(width = 0.8)) +
-    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), 
-                    position = position_dodge(width = 0.8)) +
-    labs(x = "Reference Class", 
-        y = "Predicted Probability",
-        fill = "Accuracy Level",
-        title = "Predicted Probabilities of Accuracy Levels by Reference Class") +
-    scale_fill_brewer(palette = "RdYlBu") +
-    theme_minimal()
-
-# Calculate cumulative probabilities
-cum_probs <- pred_summary |>
-  group_by(refClass) |>
-  arrange(refClass, .category) |>
-  mutate(
-    cum_prob = cumsum(mean_prob),
-    cum_lower = cumsum(lower_ci),
-    cum_upper = cumsum(upper_ci)
-  )
-
-ggplot(cum_probs, 
-    aes(x = refClass, y = cum_prob, color = .category)) +
-geom_point(size = 3,position=position_dodge()) +
-geom_errorbar(aes(ymin = cum_lower, ymax = cum_upper), 
-            width = 0.2,position=position_dodge()) +
-labs(x = "Reference Class",
-    y = "Cumulative Probability",
-    color = "Accuracy Level",
-    title = "Cumulative Probabilities of Accuracy Levels") +
-theme_minimal() +
-scale_color_brewer(palette = "Blues")
-
-summary_table <- pred_summary |>
-  pivot_wider(
-    names_from = .category,
-    values_from = c(mean_prob, lower_ci, upper_ci)
-  ) |>
-  kable(digits = 3)
-
-print(summary_table)
+pred_summary |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy")
 
@@
 
+odds_ratio_summary |> kable()
+# |comparison        | odds_ratio| ci_lower| ci_upper|
+# |:-----------------|----------:|--------:|--------:|
+# |Percentage vs kWh |        2.3|     0.62|      6.2|
+# |USD vs kWh        |       14.1|     4.08|     38.0|
+
+odds_ratio_summary |> pander::pandoc.table(caption="Study 1: Odds ratios of accuracy")
+
 # Plot predicted probabilities
@@
 
-```{r}
-
-bayesplot::color_scheme_set(wes_palettes[1])
+|comparison        | odds_ratio| ci_lower| ci_upper|
+|:-----------------|----------:|--------:|--------:|
+|Percentage vs kWh |        2.3|     0.62|      6.2|
+|USD vs kWh        |       14.1|     4.08|     38.0|
+
+## S1  posterior predictive checks
+
+```{r}
+
+bayesplot::color_scheme_set(wes_palettes[1]$BottleRocket1[1:6])
 
@@
 
-f |> str()
-
-f <-
-  rbind(
-    f[, , 1],
-    f[, , 2],
-    f[, , 3]
-  ) |> data.frame() |> bind_cols(expand_grid(new_data, Y=1:3))
+# Create new_data with one id and state
+new_data <- expand.grid(
+  refClass = c("kWh", "Percentage", "USD"),
+  id = unique(s1_agg$id)[1],      # Take first id
+  state = unique(s1_agg$state)[1] # Take first state
+)
+
+f_df <- as.data.frame.table(f) |>
+  rename(Probability = Freq) |>
+  # Extract Y levels from the Var3 column
+  mutate(Y = str_extract(Var3, "(?<=P\\(Y = ).*(?=\\))")) |>
+  # Join with new_data using crossing instead of bind_cols
+  crossing(new_data) |>
+  # Select and arrange columns as needed
+  select(refClass, id, state, Y, Probability)
 
@@
 
-```{r}
-
-library(tidybayes)
-library(stringr)
+## Alt s1 visual of predicted probs
+
+```{r}
 
@@
 
-# Calculate Odds Ratios
-odds_ratios_s1 <- posterior_samples(ordinal_model_s1) %>%
-  mutate(
-    OR_Percentage = exp(`b_refClassPercentage`),
-    OR_USD = exp(`b_refClassUSD`)
-  ) %>%
-  select(OR_Percentage, OR_USD) %>%
-  summarise_all(list(mean = mean, lower = ~ quantile(., 0.025), upper = ~ quantile(., 0.975))) %>%
-  pivot_longer(cols = everything(),
-               names_to = c("Reference Class", ".value"),
-               names_pattern = "OR_(.*)_(.*)") %>%
-  mutate(
-    Reference_Class = case_when(
-      str_detect(name, "Percentage") ~ "Percentage",
-      str_detect(name, "USD") ~ "USD",
-      TRUE ~ "kWh"
-    )
-  )
-
-odds_ratios_s1 <- posterior_summary(ordinal_model_s1, prob = 0.95) %>%
-  as.data.frame() %>%
-  rownames_to_column(var = "Term") %>%
-  filter(str_detect(Term, "refClass")) %>%
-  mutate(
-    OR = exp(Estimate),
-    lower = exp(`Q2.5`),
-    upper = exp(`Q97.5`)
-  ) %>%
-  select(Term, OR, lower, upper)
-
-# Display Odds Ratios
-odds_ratios_s1 %>%
-  rename(`Reference Class` = Term,
-         `Odds Ratio` = OR,
-         `95% CrI Lower` = lower,
-         `95% CrI Upper` = upper) %>%
-  mutate(`Reference Class` = str_replace(`Reference Class`, "refClass", "")) %>%
-  kable(booktabs = TRUE, caption = "Study 1: Odds Ratios for Matching Goals by Reference Class") %>%
-  kable_styling(full_width = FALSE)
-
-library(caret)
-
-# Create confusion matrix
-observed <- s1_agg$accuracy_level
-predicted <- predict(ordinal_model_s1, newdata = s1_agg, type = "response")
-predicted_class <- apply(predicted, 1, function(x) which.max(x))
-
-confusion_matrix <- confusionMatrix(
-  factor(predicted_class, levels = 1:4, labels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error")),
-  factor(observed, levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"))
-)
-
-# Display Confusion Matrix
-confusion_matrix$table %>%
-  as.data.frame.matrix() %>%
-  rownames_to_column(var = "Predicted") %>%
-  kable(booktabs = TRUE, caption = "Study 1: Confusion Matrix for Matched Goals") %>%
-  kable_styling(full_width = FALSE)
-
-```
+```
+
+## s1 probit
 
@@
 # compare
-loo_compare(loo(ordinal_model_s1), loo(ordinal_model_s1_prb))
-compare_models(ordinal_model_s1, ordinal_model_s1_prb)
+rstanarm::loo_compare(loo_logit, loo_probit)
 
@@
 
+pred_data <- data.frame(
+  refClass = c("kWh", "Percentage", "USD"),
+  id = unique(s1_agg$id)[1],      # Take first id
+  state = unique(s1_agg$state)[1] # Take first state
+)
+
 # Predicted probabilities from logit model
@@
 
-odds_ratio_summary_s2
 #             comparison odds_ratio ci_lower ci_upper
@@
 # 4 15% Goal vs 10% Goal       0.78     0.63     0.94
+
+odds_ratio_summary_s2 |> pander::pandoc.table(caption="Study 2: Odds ratios of accuracy")
 
@@
 
-```{r}
+## Interaction model
+
+```{r}
+#| eval: false
 
@@
   ),
-  file = "s2_ordinal_full6.rds"
+  file = paste0(here::here("data/model_cache","s2_ordinal_full6.rds"))
 )
@@
   )
+
+# |refClass   |rounded     |pct_goal |.category      | mean_prob| lower_ci| upper_ci|
+# |:----------|:-----------|:--------|:--------------|---------:|--------:|--------:|
+# |kWh        |Not Rounded |10%      |Exact match    |      0.09|     0.02|     0.25|
+# |kWh        |Not Rounded |10%      |0.01-2% error  |      0.20|     0.06|     0.36|
+# |kWh        |Not Rounded |10%      |2.01-15% error |      0.64|     0.38|     0.77|
+# |kWh        |Not Rounded |10%      |Over 15% error |      0.07|     0.01|     0.19|
+# |kWh        |Not Rounded |15%      |Exact match    |      0.19|     0.05|     0.44|
+# |kWh        |Not Rounded |15%      |0.01-2% error  |      0.29|     0.14|     0.41|
+# |kWh        |Not Rounded |15%      |2.01-15% error |      0.49|     0.22|     0.73|
+# |kWh        |Not Rounded |15%      |Over 15% error |      0.03|     0.01|     0.09|
+# |kWh        |Rounded     |10%      |Exact match    |      0.22|     0.06|     0.49|
+# |kWh        |Rounded     |10%      |0.01-2% error  |      0.31|     0.16|     0.41|
+# |kWh        |Rounded     |10%      |2.01-15% error |      0.45|     0.18|     0.71|
+# |kWh        |Rounded     |10%      |Over 15% error |      0.02|     0.00|     0.07|
+# |kWh        |Rounded     |15%      |Exact match    |      0.28|     0.07|     0.61|
+# |kWh        |Rounded     |15%      |0.01-2% error  |      0.32|     0.18|     0.42|
+# |kWh        |Rounded     |15%      |2.01-15% error |      0.38|     0.11|     0.68|
+# |kWh        |Rounded     |15%      |Over 15% error |      0.02|     0.00|     0.06|
+# |Percentage |Not Rounded |10%      |Exact match    |      0.04|     0.01|     0.13|
+# |Percentage |Not Rounded |10%      |0.01-2% error  |      0.12|     0.03|     0.28|
+# |Percentage |Not Rounded |10%      |2.01-15% error |      0.70|     0.53|     0.78|
+# |Percentage |Not Rounded |10%      |Over 15% error |      0.14|     0.03|     0.38|
+# |Percentage |Not Rounded |15%      |Exact match    |      0.10|     0.02|     0.29|
+# |Percentage |Not Rounded |15%      |0.01-2% error  |      0.21|     0.06|     0.37|
+# |Percentage |Not Rounded |15%      |2.01-15% error |      0.63|     0.34|     0.77|
+# |Percentage |Not Rounded |15%      |Over 15% error |      0.07|     0.01|     0.20|
+# |Percentage |Rounded     |10%      |Exact match    |      0.15|     0.03|     0.40|
+# |Percentage |Rounded     |10%      |0.01-2% error  |      0.26|     0.09|     0.39|
+# |Percentage |Rounded     |10%      |2.01-15% error |      0.55|     0.24|     0.75|
+# |Percentage |Rounded     |10%      |Over 15% error |      0.04|     0.01|     0.13|
+# |Percentage |Rounded     |15%      |Exact match    |      0.13|     0.03|     0.35|
+# |Percentage |Rounded     |15%      |0.01-2% error  |      0.25|     0.09|     0.39|
+# |Percentage |Rounded     |15%      |2.01-15% error |      0.57|     0.28|     0.76|
+# |Percentage |Rounded     |15%      |Over 15% error |      0.05|     0.01|     0.14|
+# |USD        |Not Rounded |10%      |Exact match    |      0.02|     0.00|     0.07|
+# |USD        |Not Rounded |10%      |0.01-2% error  |      0.07|     0.01|     0.19|
+# |USD        |Not Rounded |10%      |2.01-15% error |      0.67|     0.42|     0.78|
+# |USD        |Not Rounded |10%      |Over 15% error |      0.24|     0.06|     0.56|
+# |USD        |Not Rounded |15%      |Exact match    |      0.06|     0.01|     0.18|
+# |USD        |Not Rounded |15%      |0.01-2% error  |      0.14|     0.03|     0.32|
+# |USD        |Not Rounded |15%      |2.01-15% error |      0.68|     0.47|     0.78|
+# |USD        |Not Rounded |15%      |Over 15% error |      0.13|     0.02|     0.37|
+# |USD        |Rounded     |10%      |Exact match    |      0.05|     0.01|     0.18|
+# |USD        |Rounded     |10%      |0.01-2% error  |      0.13|     0.03|     0.32|
+# |USD        |Rounded     |10%      |2.01-15% error |      0.68|     0.47|     0.78|
+# |USD        |Rounded     |10%      |Over 15% error |      0.13|     0.02|     0.37|
+# |USD        |Rounded     |15%      |Exact match    |      0.04|     0.01|     0.14|
+# |USD        |Rounded     |15%      |0.01-2% error  |      0.11|     0.02|     0.28|
+# |USD        |Rounded     |15%      |2.01-15% error |      0.69|     0.50|     0.78|
+# |USD        |Rounded     |15%      |Over 15% error |      0.15|     0.03|     0.41|
 
@@
 
-odds_ratio_summary_s2
-```
-
-            comparison odds_ratio ci_lower ci_upper
-1    Percentage vs kWh       3.31     0.43     11.3
-2           USD vs kWh       6.64     0.84     25.2
-3       Rounded vs Not       0.41     0.09      1.1
-4 15% Goal vs 10% Goal       0.51     0.12      1.5
-
-```{r}
+#  odds_ratio_summary_s2
+#             comparison odds_ratio ci_lower ci_upper
+# 1    Percentage vs kWh       3.31     0.43     11.3
+# 2           USD vs kWh       6.64     0.84     25.2
+# 3       Rounded vs Not       0.41     0.09      1.1
+# 4 15% Goal vs 10% Goal       0.51     0.12      1.5
+
+odds_ratio_summary_s2 |> kable()
+
+```
+
+|comparison           | odds_ratio| ci_lower| ci_upper|
+|:--------------------|----------:|--------:|--------:|
+|Percentage vs kWh    |       3.31|     0.43|     11.2|
+|USD vs kWh           |       6.64|     0.84|     25.2|
+|Rounded vs Not       |       0.41|     0.09|      1.1|
+|15% Goal vs 10% Goal |       0.51|     0.12|      1.5|
+
+```{r}
+#| eval: false
 
@@
 odds_ratio_summary_s2
-
-posterior_samples_s2 <- as.data.frame(ordinal_model_s2)
-
-odds_ratio_summary_s2 <- data.frame(
-  comparison = c("Percentage vs kWh", "USD vs kWh", "Rounded vs Not", "15% Goal vs 10% Goal"),
-  odds_ratio = c(
-    mean(odds_ratios_s2$Percentage_vs_kWh),
-    mean(odds_ratios_s2$USD_vs_kWh),
-    mean(odds_ratios_s2$Rounded_vs_Not),
-    mean(odds_ratios_s2$Goal15_vs_10)
-  ),
-  ci_lower = c(
-    quantile(odds_ratios_s2$Percentage_vs_kWh, 0.025),
-    quantile(odds_ratios_s2$USD_vs_kWh, 0.025),
-    quantile(odds_ratios_s2$Rounded_vs_Not, 0.025),
-    quantile(odds_ratios_s2$Goal15_vs_10, 0.025)
-  ),
-  ci_upper = c(
-    quantile(odds_ratios_s2$Percentage_vs_kWh, 0.975),
-    quantile(odds_ratios_s2$USD_vs_kWh, 0.975),
-    quantile(odds_ratios_s2$Rounded_vs_Not, 0.975),
-    quantile(odds_ratios_s2$Goal15_vs_10, 0.975)
-  )
-)
-
-```
+#   Percentage_vs_kWh USD_vs_kWh Rounded_vs_Not Goal15_vs_Goal10
+# 1               3.3        6.6           0.41             0.51
+
+```

### 2024-11-20 misc/outlier.qmd
add manuscript, and restructure website

@@
 title: "Outliers and discrepancies"
-author:
-- name: Thomas E. Gorman
-  url: https://tegorman13.github.io/
-  affiliations: 
-  - name:  Communication and Cognition Lab, Purdue University, USA
-    affiliation-url: https://web.ics.purdue.edu/~treimer/
-lightbox: true
-toc: false
-code-fold: true
-code-tools: true
-execute:
-  #echo: false
-  warning: false
 format:

### 2024-11-20 mpa/mpa_analysis.qmd
add manuscript, and restructure website

--- 2024-11-20 misc/ordinal.qmd
@@
 ---
-title: "Ordinal Models"
+title: "MPA abstract analysis"
 execute:
-  #echo: false
     warning: false
@@
         html-math-method: mathjax
-    output-file: ordinal_hugo.md
+        output-file: mpa_hugo.md
     gfm:
         echo: true
-    output-file: ordinal.md
+        output-file: mpa_gfm.md
 ---
 
-## Ordinal Regression alternative
-
 ```{r}
-
-pacman::p_load(dplyr,purrr,tidyr,here,tibble,stringr,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,
+pacman::p_load(dplyr,purrr,tidyr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,
                ggplot2,gt,knitr,kableExtra,ggh4x,lme4,flextable,pander)
@@
 s2_long <- readRDS(here::here("data/s2_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds"))))
-```
-
-## Proportions in aggregate data
-
-```{r}
-
-s1_agg <- s1 |> 
-    filter(appliance !="Total kWh") |> 
-    group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
-    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
-                pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
-                n_change = sum(value!=family),
-                state_p_dif=mean(state_p_dif),
-                state_f_dif=mean(state_f_dif),
-                n_less_avg = sum(less_avg),
-                duration=first(Duration__in_seconds_)) |> 
-    mutate(matched_goal = (pct_change == pct_goal), 
-                error = pct_change - pct_goal,
-                abs_error = abs(error),
-                close_match = abs_error <= 0.03) |>
-    ungroup() |> # Add ungroup here
-        mutate(
-            accuracy_level = factor(
-                case_when(
-                    abs_error == 0.00 ~ "Exact match",
-                    abs_error <= 0.02 ~ "0.01-2% error",
-                    abs_error <= 0.15 ~ "2.01-15% error",
-                    TRUE ~ "Over 15% error"  # Capture all remaining cases
-                ), 
-                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
-                ordered = TRUE
-            )
-        )
-
-ggplot(data = s1_agg, aes(x = accuracy_level, fill = refClass)) +
-  geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
-  scale_y_continuous(labels = scales::percent) +
-  labs(title = "Accuracy Levels by Reference Class",
-       x = "Accuracy Level",
-       y = "Percentage of Participants",
-       fill = "Reference Class") +
-  theme_minimal()
-
-observed_props <- s1_agg |>
-  group_by(refClass, accuracy_level) |>
-  summarise(n = n()) |>
-  group_by(refClass) |>
-  mutate(prop = n/sum(n)) |>
-  pivot_wider(
-    names_from = accuracy_level,
-    values_from = c(n, prop)
-  )
-
-  # Calculate proportions
-prop_acc_s1 <- s1_agg %>%
-    group_by(refClass, accuracy_level) %>%
-    summarise(count = n()) %>%
-    group_by(refClass) %>%
-    mutate(Probability = count / sum(count)) %>%
-    ungroup()
-
-# Plot
-ggplot(prop_acc_s1, aes(x = accuracy_level, y = Probability, fill = refClass)) +
-    geom_bar(stat = "identity", position = position_dodge()) +
-    labs(title = "Distribution of Accuracy Levels by Reference Class",
-        x = "Accuracy Level",
-        y = "Probability",
-        fill = "Reference Class") +
-    theme_minimal() +
-    scale_fill_brewer(palette = "Set1")
-
 ```
@@
     iter = 2000,
-  control = list(adapt_delta = 0.99), # Recommended for ordinal models
-  prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
-            prior(normal(0, 1.5), class = "b")), # Priors for predictors
-  file = paste0(here::here("data/model_cache",'s1_ordinal3.rds')) # Cache for efficiency
+    control = list(adapt_delta = 0.99), 
+    prior = c(prior(normal(0, 2), class = "Intercept"), 
+                prior(normal(0, 1.5), class = "b")), 
+    file = paste0(here::here("data/model_cache",'s1_ordinal3.rds')) 
 )
@@
 summary(ordinal_model_s1)
-
-#ordinal_model_s1 |> emmeans(~refClass, type="response")
 
@@
 |USD vs kWh        |       14.1|     4.08|     38.0|
-
-## S1  posterior predictive checks
-
-```{r}
-
-bayesplot::color_scheme_set(wes_palettes[1]$BottleRocket1[1:6])
-
-pp_check(ordinal_model_s1, type = "bars", ndraws = 1000, fatten = 2) +
-    scale_x_continuous("y", breaks = 1:7) +
-    scale_y_continuous(NULL, breaks = NULL, expand = expansion(mult = c(0, 0.05))) +
-    ggtitle("Data with posterior predictions",
-            subtitle = "N = 100")
-
-pp_check(ordinal_model_s1, type = "bars_grouped", group="refClass", fatten = 2) +
-    scale_x_continuous("y", breaks = 1:7) +
-    scale_y_continuous(NULL, breaks = NULL, expand = expansion(mult = c(0, 0.05))) +
-    ggtitle("Data with posterior predictions",
-            subtitle = expression(list(italic(N[A])==44, italic(N[B])==44))) +
-    theme(legend.background = element_blank(),
-            legend.position = c(.9, .75))
-
-f <- fitted(ordinal_model_s1,newdata_pred <- expand.grid(
-    refClass = c("kWh", "Percentage", "USD"),
-    id = unique(s1_agg$id),
-    state = unique(s1_agg$state)
-    ))
-
-# Create new_data with one id and state
-new_data <- expand.grid(
-  refClass = c("kWh", "Percentage", "USD"),
-  id = unique(s1_agg$id)[1],      # Take first id
-  state = unique(s1_agg$state)[1] # Take first state
-)
-
-f_df <- as.data.frame.table(f) |>
-  rename(Probability = Freq) |>
-  # Extract Y levels from the Var3 column
-  mutate(Y = str_extract(Var3, "(?<=P\\(Y = ).*(?=\\))")) |>
-  # Join with new_data using crossing instead of bind_cols
-  crossing(new_data) |>
-  # Select and arrange columns as needed
-  select(refClass, id, state, Y, Probability)
-
-```
-
-## Comparing Marginal vs Conditional Predictions
-
-```{r}
-
-# Create newdata with all necessary variables
-newdata_pred <- expand.grid(
-  refClass = c("kWh", "Percentage", "USD"),
-  id = unique(s1_agg$id),
-  state = unique(s1_agg$state)
-)
-
-# Population-level predictions
-pred_probs_marginal <- ordinal_model_s1 |>
-  epred_draws(
-    newdata = data.frame(refClass = c("kWh", "Percentage", "USD")),
-    ndraws = 1000, 
-    re_formula = NA
-  )
-
-# Conditional predictions (including random effects)
-pred_probs_conditional <- ordinal_model_s1 |>
-  epred_draws(
-    newdata = newdata_pred,
-    ndraws = 1000, 
-    re_formula = NULL
-  )
-
-# Compare the two approaches
-summary_comparison <- bind_rows(
-    pred_probs_marginal |> 
-        group_by(refClass, .category) |>
-        summarise(
-        mean_prob = mean(.epred),
-        sd_prob = sd(.epred),
-        lower_ci = quantile(.epred, 0.025),
-        upper_ci = quantile(.epred, 0.975),
-        .groups = 'drop'
-        ) |>
-        mutate(type = "Marginal"),
-
-    pred_probs_conditional |> 
-        group_by(refClass, .category) |>
-        summarise(
-        mean_prob = mean(.epred),
-        sd_prob = sd(.epred),
-        lower_ci = quantile(.epred, 0.025),
-        upper_ci = quantile(.epred, 0.975),
-        .groups = 'drop'
-        ) |>
-        mutate(type = "Conditional")
-)
-
-# View the comparison
-summary_comparison |>
-  arrange(type, refClass, .category) |>
-  kable(digits = 3)
-
-# Visualize the comparison
-ggplot(summary_comparison, 
-        aes(x = .category, y = mean_prob, fill = refClass)) +
-    geom_col(position = position_dodge()) +
-    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), 
-            position = position_dodge(width = 0.8)) +
-    facet_wrap(~type) +
-    theme_minimal() +
-    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-    labs(title = "Comparing Marginal vs Conditional Predictions",
-        x = "Accuracy Level",
-        y = "Predicted Probability")
-```
-
-## Individual Differences in Predictions
-
-```{r}
-# Extract random effects for subjects with corrected ID handling
-ranef_draws <- ordinal_model_s1 |>
-  spread_draws(r_id[id,term]) |>
-  mutate(id = as.integer(as.character(id))) # Convert factor to integer
-
-# Get subject-specific adjustments
-subject_adjustments <- ranef_draws |>
-  group_by(id) |>
-  summarise(
-    mean_adj = mean(r_id),
-    lower = quantile(r_id, 0.025),
-    upper = quantile(r_id, 0.975)
-  )
-
-# Get reference class for each subject
-subject_refclass <- s1_agg |>
-  group_by(id) |>
-  summarise(refClass = first(refClass))
-
-# Combine adjustments with reference class
-subject_plot_data <- subject_adjustments |>
-  left_join(subject_refclass, by = "id")
-
-# Create caterpillar plot
-ggplot(subject_plot_data, 
-       aes(x = reorder(factor(id), mean_adj), 
-           y = mean_adj, 
-           color = refClass)) +
-  geom_point() +
-  geom_errorbar(aes(ymin = lower, ymax = upper), alpha = 0.5) +
-  coord_flip() +
-  facet_wrap(~refClass, scales = "free_y") +
-  theme_minimal() +
-  labs(title = "Subject-Level Random Effects by Reference Class",
-       y = "Subject-Specific Adjustment",
-       x = "Subject ID") +
-  theme(legend.position = "none")
-
-# Density plot of random effects by condition
-ggplot(subject_plot_data, 
-       aes(x = mean_adj, fill = refClass)) +
-  geom_density(alpha = 0.5) +
-  theme_minimal() +
-  labs(title = "Distribution of Subject-Level Random Effects by Reference Class",
-       x = "Subject-Specific Adjustment",
-       y = "Density")
-
-# Summary statistics by condition
-random_effects_summary <- subject_plot_data |>
-  group_by(refClass) |>
-  summarise(
-    n = n(),
-    mean_adjustment = mean(mean_adj),
-    sd_adjustment = sd(mean_adj),
-    median_adjustment = median(mean_adj),
-    q25 = quantile(mean_adj, 0.25),
-    q75 = quantile(mean_adj, 0.75)
-  )
-
-print(random_effects_summary)
-
-pop_preds <- ordinal_model_s1 |>
-    epred_draws(
-        newdata = data.frame(refClass = c("kWh", "Percentage", "USD")),
-        ndraws = 1000, 
-        re_formula = NA
-    ) |>
-    group_by(refClass, .category) |>
-    summarise(
-        pop_mean = mean(.epred),
-        pop_lower = quantile(.epred, 0.025),
-        pop_upper = quantile(.epred, 0.975)
-    )
-
-# Get subject-specific predictions
-subject_preds <- ordinal_model_s1 |>
-    epred_draws(
-        newdata = s1_agg,
-        ndraws = 100  # fewer draws for computational efficiency
-    ) |>
-    group_by(id, refClass, .category) |>
-    summarise(
-        subj_mean = mean(.epred)
-    )
-
-# Plot 1: Population effects with subject-specific points
-ggplot() +
-    # Add subject-specific points
-    geom_jitter(data = subject_preds,
-                aes(x = .category, y = subj_mean, color = refClass),
-                alpha = 0.2, width = 0.2, height = 0) +
-    # Add population-level effects
-    geom_point(data = pop_preds,
-                aes(x = .category, y = pop_mean, fill = refClass),
-                size = 3, shape = 21, color = "black") +
-    geom_errorbar(data = pop_preds,
-                aes(x = .category, ymin = pop_lower, ymax = pop_upper),
-                width = 0.2) +
-    facet_wrap(~refClass) +
-    theme_minimal() +
-    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-    labs(title = "Population-Level Effects with Subject-Specific Predictions",
-        x = "Accuracy Level",
-        y = "Predicted Probability")
-
-# Plot 2: Violin plot with population effects
-ggplot() +
-    # Add violin plot of subject-specific effects
-    geom_violin(data = subject_preds,
-                aes(x = .category, y = subj_mean, fill = refClass),
-                alpha = 0.3) +
-    # Add population-level points
-    geom_point(data = pop_preds,
-                aes(x = .category, y = pop_mean),
-                size = 3, color = "black") +
-    geom_errorbar(data = pop_preds,
-                aes(x = .category, ymin = pop_lower, ymax = pop_upper),
-                width = 0.2) +
-    facet_wrap(~refClass) +
-    theme_minimal() +
-    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-    labs(title = "Distribution of Subject-Specific Effects with Population Estimates",
-        x = "Accuracy Level",
-        y = "Predicted Probability")
-
-# Plot 3: Subject deviations from population mean
-subject_deviations <- subject_preds |>
-    left_join(pop_preds, by = c("refClass", ".category")) |>
-    mutate(deviation = subj_mean - pop_mean)
-
-ggplot(subject_deviations,
-        aes(x = deviation, fill = refClass)) +
-    geom_density(alpha = 0.5) +
-    geom_vline(xintercept = 0, linetype = "dashed") +
-    facet_grid(refClass ~ .category) +
-    theme_minimal() +
-    labs(title = "Subject Deviations from Population Mean",
-        x = "Deviation from Population Mean",
-        y = "Density")
-
-# Plot 4: Ridgeline plot of subject effects
-library(ggridges)
-ggplot(subject_preds,
-        aes(x = subj_mean, y = .category, fill = refClass)) +
-    geom_density_ridges(alpha = 0.5) +
-    geom_vline(data = pop_preds,
-                aes(xintercept = pop_mean),
-                linetype = "dashed") +
-    facet_wrap(~refClass) +
-    theme_minimal() +
-    labs(title = "Distribution of Subject-Specific Effects by Accuracy Level",
-        x = "Predicted Probability",
-        y = "Accuracy Level")
-```
-
-## Alt s1 visual of predicted probs
-
-```{r}
-
-predicted_probs_s1 <- ordinal_model_s1 %>%
-    # gather draws that contain refClass ("b_refClassPercentage"    "b_refClassUSD"  )
-    gather_draws(b_refClassPercentage, b_refClassUSD) %>%
-    mutate(refClass = case_when(
-    str_detect(.variable, "Percentage") ~ "Percentage",
-    str_detect(.variable, "USD") ~ "USD",
-    TRUE ~ "kWh"
-  )) %>%
-    group_by(refClass) %>%
-    summarise(effect = mean(.value)) %>%
-    mutate(
-        group = refClass,
-        prob_exact = plogis(effect + (-1.85)),
-        prob_0_2 = plogis(effect + 0.22) - plogis(effect + (-1.85)),
-        prob_2_15 = plogis(effect + 2.86) - plogis(effect + 0.22),
-        prob_over15 = 1 - plogis(effect + 2.86)
-    ) %>%
-    pivot_longer(
-        cols = starts_with("prob"),
-        names_to = "Accuracy_Level",
-        values_to = "Probability"
-    ) %>%
-    mutate(
-        Accuracy_Level = recode(Accuracy_Level,
-                                "prob_exact" = "Exact match",
-                                "prob_0_2" = "0.01-2% error",
-                                "prob_2_15" = "2.01-15% error",
-                                "prob_over15" = "Over 15% error")
-    )
-
-ggplot(predicted_probs_s1, aes(x = Accuracy_Level, y = Probability, fill = refClass)) +
-  geom_bar(stat = "identity", position = "dodge") +
-  labs(title = "Predicted Probabilities of Accuracy Levels by Reference Class",
-       x = "Accuracy Level",
-       y = "Probability",
-       fill = "Reference Class") +
-  theme_minimal()
-
-```
-
-## s1 probit
-
-```{r}
-
-ordinal_model_s1_prb <- brm(
-  accuracy_level ~ refClass + (1|id) + (1|state),
-  data = s1_agg,
-  family = cumulative("probit"),
-  cores = 4,
-  iter = 2000,
-  control = list(adapt_delta = 0.99), # Recommended for ordinal models
-  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
-            prior(normal(0, 2), class = "b")), # Priors for predictors
-  file = paste0(here::here("data/model_cache",'s1_ordinal_probit1.rds')) # Cache for efficiency
-)
-
-summary(ordinal_model_s1_prb)
-
-pp_check(ordinal_model_s1_prb, type = "bars_grouped", group="refClass", fatten = 2) +
-    scale_x_continuous("y", breaks = 1:7) +
-    scale_y_continuous(NULL, breaks = NULL, expand = expansion(mult = c(0, 0.05))) +
-    ggtitle("Data with posterior predictions",
-            subtitle = expression(list(italic(N[A])==44, italic(N[B])==44))) +
-    theme(legend.background = element_blank(),
-            legend.position = c(.9, .75))
-
-loo_logit <- loo(ordinal_model_s1)
-loo_probit <- loo(ordinal_model_s1_prb)
-loo_compare(loo_logit, loo_probit)
-
-waic(ordinal_model_s1)
-waic(ordinal_model_s1_prb)
-
-# compare
-rstanarm::loo_compare(loo_logit, loo_probit)
-
-# Obtain predicted probabilities
-predicted_probs_s1 <- ordinal_model_s1_prb %>%
-    gather_draws(b_refClassPercentage, b_refClassUSD) %>%
-    mutate(refClass = case_when(
-        str_detect(.variable, "Percentage") ~ "Percentage",
-        str_detect(.variable, "USD") ~ "USD",
-        TRUE ~ "kWh"
-    )) %>%
-    group_by(refClass) %>%
-    summarise(effect = mean(.value)) %>%
-    mutate(
-        group = refClass,
-        prob_exact = plogis(effect + (-1.85)),
-        prob_0_2 = plogis(effect + 0.22) - plogis(effect + (-1.85)),
-        prob_2_15 = plogis(effect + 2.86) - plogis(effect + 0.22),
-        prob_over15 = 1 - plogis(effect + 2.86)
-    ) %>%
-    pivot_longer(
-        cols = starts_with("prob"),
-        names_to = "Accuracy_Level",
-        values_to = "Probability"
-    ) %>%
-    mutate(
-        Accuracy_Level = recode(Accuracy_Level,
-                                "prob_exact" = "Exact match",
-                                "prob_0_2" = "0.01-2% error",
-                                "prob_2_15" = "2.01-15% error",
-                                "prob_over15" = "Over 15% error")
-    )
-
-    # Plot predicted probabilities
-    ggplot(predicted_probs_s1, aes(x = Accuracy_Level, y = Probability, fill = refClass)) +
-    geom_bar(stat = "identity", position = "dodge") +
-    labs(title = "Predicted Probabilities of Accuracy Levels by Reference Class",
-        x = "Accuracy Level",
-        y = "Probability",
-        fill = "Reference Class") +
-    theme_minimal()
-
-pred_data <- data.frame(
-  refClass = c("kWh", "Percentage", "USD"),
-  id = unique(s1_agg$id)[1],      # Take first id
-  state = unique(s1_agg$state)[1] # Take first state
-)
-
-# Predicted probabilities from logit model
-pred_probs_logit <- posterior_epred(ordinal_model_s1, newdata = pred_data, re_formula = NA)
-
-# Predicted probabilities from probit model
-pred_probs_probit <- posterior_epred(ordinal_model_s1_prb, newdata = pred_data, re_formula = NA)
-
-# Compare predicted probabilities
-mean_pred_logit <- apply(pred_probs_logit, 2, mean)
-mean_pred_probit <- apply(pred_probs_probit, 2, mean)
-
-comparison <- data.frame(
-  Reference_Class = pred_data$refClass,
-  Mean_Prob_Logit = mean_pred_logit,
-  Mean_Prob_Probit = mean_pred_probit
-)
-
-print(comparison)
-
-```
 
@@
 
+odds_ratio_summary_s2 |> kable()
+
 odds_ratio_summary_s2 |> pander::pandoc.table(caption="Study 2: Odds ratios of accuracy")
@@
 
-## Interaction model
-
-```{r}
-#| eval: false
-
-# Fit the full model with interactions
-ordinal_model_s2 <- brm(
-  accuracy_level ~ refClass * rounded * pct_goal + (1|id) + (1|state),
-  data = s2_agg1,
-  family = cumulative("logit"),
-  cores = 4,
-  iter = 2000,
-  control = list(adapt_delta = 0.99),
-  prior = c(
-    prior(normal(0, 2), class = "Intercept"),
-    prior(normal(0, 2), class = "b")
-  ),
-  file = paste0(here::here("data/model_cache","s2_ordinal_full6.rds"))
-)
-
-pred_summary_s2 <- ordinal_model_s2 %>%
-  epred_draws(newdata = s2_agg1, re_formula = NA) %>%
-  group_by(refClass, rounded, pct_goal, .category) %>%
-  summarise(
-    mean_prob = mean(.epred),
-    lower_ci = quantile(.epred, 0.025),
-    upper_ci = quantile(.epred, 0.975),
-    .groups = "drop"
-  )
-
-# |refClass   |rounded     |pct_goal |.category      | mean_prob| lower_ci| upper_ci|
-# |:----------|:-----------|:--------|:--------------|---------:|--------:|--------:|
-# |kWh        |Not Rounded |10%      |Exact match    |      0.09|     0.02|     0.25|
-# |kWh        |Not Rounded |10%      |0.01-2% error  |      0.20|     0.06|     0.36|
-# |kWh        |Not Rounded |10%      |2.01-15% error |      0.64|     0.38|     0.77|
-# |kWh        |Not Rounded |10%      |Over 15% error |      0.07|     0.01|     0.19|
-# |kWh        |Not Rounded |15%      |Exact match    |      0.19|     0.05|     0.44|
-# |kWh        |Not Rounded |15%      |0.01-2% error  |      0.29|     0.14|     0.41|
-# |kWh        |Not Rounded |15%      |2.01-15% error |      0.49|     0.22|     0.73|
-# |kWh        |Not Rounded |15%      |Over 15% error |      0.03|     0.01|     0.09|
-# |kWh        |Rounded     |10%      |Exact match    |      0.22|     0.06|     0.49|
-# |kWh        |Rounded     |10%      |0.01-2% error  |      0.31|     0.16|     0.41|
-# |kWh        |Rounded     |10%      |2.01-15% error |      0.45|     0.18|     0.71|
-# |kWh        |Rounded     |10%      |Over 15% error |      0.02|     0.00|     0.07|
-# |kWh        |Rounded     |15%      |Exact match    |      0.28|     0.07|     0.61|
-# |kWh        |Rounded     |15%      |0.01-2% error  |      0.32|     0.18|     0.42|
-# |kWh        |Rounded     |15%      |2.01-15% error |      0.38|     0.11|     0.68|
-# |kWh        |Rounded     |15%      |Over 15% error |      0.02|     0.00|     0.06|
-# |Percentage |Not Rounded |10%      |Exact match    |      0.04|     0.01|     0.13|
-# |Percentage |Not Rounded |10%      |0.01-2% error  |      0.12|     0.03|     0.28|
-# |Percentage |Not Rounded |10%      |2.01-15% error |      0.70|     0.53|     0.78|
-# |Percentage |Not Rounded |10%      |Over 15% error |      0.14|     0.03|     0.38|
-# |Percentage |Not Rounded |15%      |Exact match    |      0.10|     0.02|     0.29|
-# |Percentage |Not Rounded |15%      |0.01-2% error  |      0.21|     0.06|     0.37|
-# |Percentage |Not Rounded |15%      |2.01-15% error |      0.63|     0.34|     0.77|
-# |Percentage |Not Rounded |15%      |Over 15% error |      0.07|     0.01|     0.20|
-# |Percentage |Rounded     |10%      |Exact match    |      0.15|     0.03|     0.40|
-# |Percentage |Rounded     |10%      |0.01-2% error  |      0.26|     0.09|     0.39|
-# |Percentage |Rounded     |10%      |2.01-15% error |      0.55|     0.24|     0.75|
-# |Percentage |Rounded     |10%      |Over 15% error |      0.04|     0.01|     0.13|
-# |Percentage |Rounded     |15%      |Exact match    |      0.13|     0.03|     0.35|
-# |Percentage |Rounded     |15%      |0.01-2% error  |      0.25|     0.09|     0.39|
-# |Percentage |Rounded     |15%      |2.01-15% error |      0.57|     0.28|     0.76|
-# |Percentage |Rounded     |15%      |Over 15% error |      0.05|     0.01|     0.14|
-# |USD        |Not Rounded |10%      |Exact match    |      0.02|     0.00|     0.07|
-# |USD        |Not Rounded |10%      |0.01-2% error  |      0.07|     0.01|     0.19|
-# |USD        |Not Rounded |10%      |2.01-15% error |      0.67|     0.42|     0.78|
-# |USD        |Not Rounded |10%      |Over 15% error |      0.24|     0.06|     0.56|
-# |USD        |Not Rounded |15%      |Exact match    |      0.06|     0.01|     0.18|
-# |USD        |Not Rounded |15%      |0.01-2% error  |      0.14|     0.03|     0.32|
-# |USD        |Not Rounded |15%      |2.01-15% error |      0.68|     0.47|     0.78|
-# |USD        |Not Rounded |15%      |Over 15% error |      0.13|     0.02|     0.37|
-# |USD        |Rounded     |10%      |Exact match    |      0.05|     0.01|     0.18|
-# |USD        |Rounded     |10%      |0.01-2% error  |      0.13|     0.03|     0.32|
-# |USD        |Rounded     |10%      |2.01-15% error |      0.68|     0.47|     0.78|
-# |USD        |Rounded     |10%      |Over 15% error |      0.13|     0.02|     0.37|
-# |USD        |Rounded     |15%      |Exact match    |      0.04|     0.01|     0.14|
-# |USD        |Rounded     |15%      |0.01-2% error  |      0.11|     0.02|     0.28|
-# |USD        |Rounded     |15%      |2.01-15% error |      0.69|     0.50|     0.78|
-# |USD        |Rounded     |15%      |Over 15% error |      0.15|     0.03|     0.41|
-
-# Extract posterior samples
-posterior_samples_s2 <- as.data.frame(ordinal_model_s2)
-
-odds_ratios_s2 <- posterior_samples_s2 %>%
-    transmute(
-        refClass_Percentage_vs_kWh = exp(b_refClassPercentage),
-        refClass_USD_vs_kWh = exp(b_refClassUSD),
-        rounded_Yes_vs_No = exp(b_roundedRounded),
-        pct_goal_15_vs_10 = exp(`b_pct_goal15%`)  # Note the backticks here
-    )
-# Calculate summary statistics
-odds_ratio_summary_s2 <- data.frame(
-    comparison = c("Percentage vs kWh", "USD vs kWh", "Rounded vs Not", "15% Goal vs 10% Goal"),
-    odds_ratio = c(
-        mean(odds_ratios_s2$refClass_Percentage_vs_kWh),
-        mean(odds_ratios_s2$refClass_USD_vs_kWh),
-        mean(odds_ratios_s2$rounded_Yes_vs_No),
-        mean(odds_ratios_s2$pct_goal_15_vs_10)
-    ),
-    ci_lower = c(
-        quantile(odds_ratios_s2$refClass_Percentage_vs_kWh, 0.025),
-        quantile(odds_ratios_s2$refClass_USD_vs_kWh, 0.025),
-        quantile(odds_ratios_s2$rounded_Yes_vs_No, 0.025),
-        quantile(odds_ratios_s2$pct_goal_15_vs_10, 0.025)
-    ),
-    ci_upper = c(
-        quantile(odds_ratios_s2$refClass_Percentage_vs_kWh, 0.975),
-        quantile(odds_ratios_s2$refClass_USD_vs_kWh, 0.975),
-        quantile(odds_ratios_s2$rounded_Yes_vs_No, 0.975),
-        quantile(odds_ratios_s2$pct_goal_15_vs_10, 0.975)
-    )
-)
-
-#  odds_ratio_summary_s2
-#             comparison odds_ratio ci_lower ci_upper
-# 1    Percentage vs kWh       3.31     0.43     11.3
-# 2           USD vs kWh       6.64     0.84     25.2
-# 3       Rounded vs Not       0.41     0.09      1.1
-# 4 15% Goal vs 10% Goal       0.51     0.12      1.5
-
-odds_ratio_summary_s2 |> kable()
-
-```
-
 |comparison           | odds_ratio| ci_lower| ci_upper|
 |:--------------------|----------:|--------:|--------:|
-|Percentage vs kWh    |       3.31|     0.43|     11.2|
-|USD vs kWh           |       6.64|     0.84|     25.2|
-|Rounded vs Not       |       0.41|     0.09|      1.1|
-|15% Goal vs 10% Goal |       0.51|     0.12|      1.5|
-
-```{r}
-#| eval: false
-
-s2_agg1 <- s2_agg1 %>%
-  mutate(
-    refClass = factor(refClass, levels = c("kWh", "Percentage", "USD")),
-    rounded = factor(rounded, levels = c("Not Rounded", "Rounded")),
-    pct_goal = factor(pct_goal, levels = c("10%", "15%"))
-  )
-
-# Fit the full model with interactions
-ordinal_model_s2 <- brm(
-  accuracy_level ~ refClass * rounded * pct_goal + (1|id) + (1|state),
-  data = s2_agg1,
-  family = cumulative("logit"),
-  cores = 4,
-  iter = 2000,
-  control = list(adapt_delta = 0.99),
-  prior = c(
-    prior(normal(0, 2), class = "Intercept"),
-    prior(normal(0, 2), class = "b")
-  ),
-  file = paste0(here::here("data/model_cache","s2_ordinal_full6.rds"))
-)
-
-pred_summary_s2 <- ordinal_model_s2 %>%
-  epred_draws(newdata = s2_agg1, re_formula = NA) %>%
-  group_by(refClass, rounded, pct_goal, .category) %>%
-  summarise(
-    mean_prob = mean(.epred),
-    lower_ci = quantile(.epred, 0.025),
-    upper_ci = quantile(.epred, 0.975),
-    .groups = "drop"
-  )
-
-# Extract posterior samples
-posterior_samples_s2 <- as.data.frame(ordinal_model_s2)
-
-# Compute odds ratios for the predictors
-odds_ratios_s2 <- posterior_samples_s2 %>%
-  transmute(
-    refClass_Percentage_vs_kWh = exp(b_refClassPercentage),
-    refClass_USD_vs_kWh = exp(b_refClassUSD),
-    rounded_Yes_vs_No = exp(b_roundedRounded),
-    pct_goal_15_vs_10 = exp(`b_pct_goal15%`)  # Note the backticks here
-  )
-
-# Summarize odds ratios
-odds_ratio_summary_s2 <- odds_ratios_s2 %>%
-  summarise(
-    Percentage_vs_kWh = mean(refClass_Percentage_vs_kWh),
-    USD_vs_kWh = mean(refClass_USD_vs_kWh),
-    Rounded_vs_Not = mean(rounded_Yes_vs_No),
-    Goal15_vs_Goal10 = mean(pct_goal_15_vs_10)
-  )
-
-odds_ratio_summary_s2
-#   Percentage_vs_kWh USD_vs_kWh Rounded_vs_Not Goal15_vs_Goal10
-# 1               3.3        6.6           0.41             0.51
-
-```
+|Percentage vs kWh    |       1.99|     0.73|     4.43|
+|USD vs kWh           |       3.69|     1.33|     8.29|
+|Rounded vs Not       |       0.70|     0.57|     0.84|
+|15% Goal vs 10% Goal |       0.78|     0.63|     0.94|

### 2024-11-20 nb/overview.qmd
add manuscript, and restructure website

--- 2024-11-07 overview.qmd
@@
 options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
-walk(c("fun_plot"), ~ source(here::here(paste0("Scripts/", .x, ".R"))))
+walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
 theme_set(theme_nice())

### 2024-11-20 nb/study1.qmd
add manuscript, and restructure website

--- 2024-11-04 study1.qmd
@@
 title: "Study 1 Analysis"
-author:
-- name: Thomas E. Gorman
-  url: https://tegorman13.github.io/
-  affiliations: 
-  - name:  Communication and Cognition Lab, Purdue University, USA
-    affiliation-url: https://web.ics.purdue.edu/~treimer/
-lightbox: true
 toc: false
-code-fold: true
-code-tools: true
 execute:
@@
 
-saveRDS(s1, here::here("data/s1_processed.rds"))
+#saveRDS(s1, here::here("data/s1_processed.rds"))
 
@@
 
+```
+
+```{r}
+
 # examine relationship between n_change and whether goal was matched
@@
   scale_x_continuous(breaks=seq(1,5,1)) 
-
 ```

### 2024-11-20 nb/study2.qmd
add manuscript, and restructure website

--- 2024-11-04 study2.qmd
@@
 title: "Study 2 Analysis"
-author:
-- name: Thomas E. Gorman
-  url: https://tegorman13.github.io/
-  affiliations: 
-  - name:  Communication and Cognition Lab, Purdue University, USA
-    affiliation-url: https://web.ics.purdue.edu/~treimer/
 lightbox: true
 toc: false
-code-fold: true
-code-tools: true
+date: today
 execute:
@@
   html:
+
     grid:
@@
 
-s2 <- haven::read_sav("data/Frequency & Probability Study 2 - 3-24-19.sav")
+s2 <- haven::read_sav(here::here("data/Frequency & Probability Study 2 - 3-24-19.sav"))
 s2 <- s2 |> #mutate(id = as.numeric(format(EndDate, "%H%M%S"))) |> relocate(id) |> 
   mutate(id=paste0("s2_",row_number())) |> relocate(id)
-s2 <- s2 |> filter(!(id=="s2_78")) # NA for some responses
+s2 <- s2 |> filter(!(id=="s2_78")) |> # NA for some responses
+  filter(!(id=="s2_196")) # NA for state_avg for COL HEAT - (HEATSTATE2)
+  # filter(!(id=="s2_208")) # NA for response for California not rounded Total (TOTALAPONE1 and TOTALAPTWO1) - leave them in since all the acutal app values are okay
 s2 <- s2 |> 
@@
 
+# s2 |> filter(id=="s2_208") |> select(starts_with("T"))
+# s2_long |> filter_all(any_vars(is.na(.)))
+
 #write.csv(s2_long,"data/study_2_tg.csv")
-saveRDS(s2_long, here::here("data/s2_processed.rds"))
+#saveRDS(s2_long, here::here("data/s2_processed.rds"))
 

### 2024-11-20 paper.qmd
add manuscript, and restructure website

---
title: Representation Project
subtitle: "Planning to Save Energy: How Information Format Affects Accuracy"
toc: false
sidebar: false
repo-actions: false
---

```{=html}
<p class="text-center">
  <a class="btn btn-success btn-sm mb-1" href="/manuscript/output/manuscript.html" target="manuscript">
  <i class="fa-solid fa-file-lines"></i> &ensp;HTML
  </a>

  <a class="btn btn-success btn-sm mb-1" href="/manuscript/output/manuscript.html" target="_blank">
  <i class="fa-solid fa-arrow-up-right-from-square"></i>&ensp;HTML <small>(new window)</small>
  </a>

  <a class="btn btn-danger btn-sm mb-1" href="/manuscript/output/rep_study.pdf" target="manuscript">
  <i class="fa-solid fa-file-pdf"></i>&ensp;PDF
  </a>

  <a class="btn btn-danger btn-sm mb-1" target="_blank" href="/manuscript/output/rep_study.pdf">
  <i class="fa-solid fa-file-arrow-down"></i>&ensp;PDF <small>(download)</small>
  </a>

</p>

<iframe src="/manuscript/output/manuscript.html" frameborder="0" style="overflow:hidden;height:88%;width:100%;border:2px solid #ccc;" height="88%" width="100%" name="manuscript"></iframe>
```

### 2024-11-20 Scripts/test.R

@@ -1,0 +1,59 @@
+# Load necessary libraries
+pacman::p_load(dplyr, purrr, tidyr, here, tibble, brms, rstan, bayestestR, emmeans, tidybayes, modelsummary, ggplot2, gt, knitr, kableExtra, ggh4x, lme4, flextable, pander)
+
+# Load the ggplot2 package
+library(ggplot2)
+
+# Create a sample dataset
+set.seed(123)
+data <- data.frame(
+  value = c(rnorm(1000, mean = 0, sd = 1), rnorm(1000, mean = 1, sd = 1)),
+  group = rep(c("A", "B"), each = 1000)
+)
+
+# Define thresholds
+thresholds <- c(-1, 0, 1, 2)
+
+# Create the plot
+ggplot(data, aes(x = value, fill = group)) +
+  geom_density(alpha = 0.5) +
+  scale_fill_manual(values = c("orange", "brown")) +
+  geom_vline(xintercept = thresholds, linetype = "dotted") +
+  labs(title = "Underlying latent scale for Y, given X",
+       fill = "X") +
+  theme_minimal() +
+  xlab(NULL) +
+  ylab(NULL) +
+  xlim(-5, 6)
+
+
+
+
+
+
+
+# Load the ggplot2 package
+library(ggplot2)
+
+# Create a sample dataset
+set.seed(123)
+data <- data.frame(
+  value = c(rnorm(1000, mean = 0, sd = 1), rnorm(1000, mean = 1, sd = 1)),
+  group = rep(c("A", "B"), each = 1000)
+)
+
+# Define thresholds
+thresholds <- c(-1, 0, 1, 2)
+
+# Create the plot
+ggplot(data, aes(x = value, fill = group)) +
+  geom_density(alpha = 0.5) +
+  scale_fill_manual(values = c("orange", "brown")) +
+  geom_vline(xintercept = thresholds, linetype = "dotted") +
+  labs(title = "Underlying latent scale for Y, given X",
+       fill = "X") +
+  theme_minimal() +
+  xlab(NULL) +
+  ylab(NULL) +
+  xlim(-5, 6) +
+  annotate("text", x = thresholds, y = 0.4, label = c(expression(theta[1]), expression(theta[2]), expression(theta[3]), expression(theta[4])), angle = 0, vjust = -0.5,hjust=.2)

### 2024-11-20 misc/indiv.qmd
minor tweak

@@
   #   output-file: results_hugo.md
-  # gfm:
-  #   echo: true
-  #   output-file: results_gfm.md
+  gfm:
+    echo: true
+    output-file: indiv_gfm.md
 ---

### 2024-11-20 misc/outlier.qmd
minor tweak

@@
   #   output-file: results_hugo.md
-  # gfm:
-  #   echo: true
-  #   output-file: results_gfm.md
+  gfm:
+    echo: true
+    output-file: outliers_gfm.md
 ---

### 2024-11-20 manuscript/manuscript.qmd
minor tweak
update with some results

@@
         shift-heading-level-by: 1
+        code-fold: true
+        code-summary: "Show the code"
         title-block-banner: "#000D51"
@@
     #     output-file: manuscript_hugo.md
-    # gfm:
-    #     echo: true
-    #     output-file: manuscript_gfm.md
+    gfm:
+        echo: true
+        output-file: manuscript_gfm.md
     hikmah-manuscript-pdf:
@@
                 dev: "cairo_pdf"
-    #hikmah-manuscript-docx: default
-format-links:   
-    - html
-    - gfm
-    - format: hikmah-manuscript-pdf
-      text: PDF version
-      icon: file-pdf
+    hikmah-manuscript-docx: default
+# format-links:   
+#     - html
+#     - gfm
+#     - format: hikmah-manuscript-pdf
+#       text: PDF version
+#       icon: file-pdf
 date: 2024-11-19
@@
 
-```{r}
-#| label: setup
-#| include: false
+# Introduction
 
-knitr::opts_chunk$set(
-  echo = FALSE, include = FALSE, warning = FALSE, message = FALSE,
-  fig.width = 6, fig.height = (6 * 0.618), fig.retina = 3,
-  out.width = "100%", fig.align = "center"
-)
-```
+### Literature Review
 
-# Introduction
+Energy poverty continues to be a pervasive issue in the United States @memmottSociodemographicDisparitiesEnergy2021 .
+This challenge partly arises from difficulties in converting information across numerical formats, impeding the development of precise energy reduction plans @reimerNumericCommunicationRisk2015.
+Prior research by @canfieldPerceptionsElectricityuseCommunications2017 demonstrated that presenting energy information in tabular formats enhances comprehension relative to graphs.
+
+### Hypotheses
+
+Building on these findings and informed by prior work showing that frequencies (like absolute units in kWh) are easier to comprehend and facilitate more precise decision-making compared to percentages, our study also utilizes a tabular format, but manipulates whether participants must consider energy information presented as absolute units (kWh), percentages (%), or monetary costs (USD).
+We hypothesize that presenting information in absolute units (kWh) will lead to more accurate household energy conservation planning.
 
@@
 
+See @fig-task for an example of a planning trial as it was seen by participants.
+
 ## Methods
+
+### Participants
+
+We implemented our task and surveys on Qualtrics, and recruited participants through Amazon Mechanical Turk.
+In Experiment 1, 252 participants were intially recruited, but data from 17 participants were corrupted due to experimenter error, leaving a final sample of 235 participants.
+Most participants (76%) reported using a calculator to complete the task.
+
+### Materials and Design
+
+The study employed a mixed design with reference class (kWh, percentage, USD) as a between-subjects factor and state/family scenario as a within-subjects factor.
+Each participant completed energy reduction planning tasks for two different states, with state order counterbalanced across participants.
+The family scenarios featured four households in different climate regions: Texas (Smith family) and California (Adams family) representing warm climates, and Colorado (Wells family) and Massachusetts (Davis family) representing cold climates.
+We obtain average utility use from each state by CITE SOURCE FOR STATE AVGS?
+
+### Procedure
+
+Participants received energy usage data for two hypothetical families and were tasked with creating action plans to meet specified reduction goals by allocating usage across five appliance categories:  heating, cooling, water heating, refrigerator, and an other appliances (e.g., TV, lighting).
+
+For each family scenario, the particpipants were shown a table containing the families utility usage from the prior year, alongside the state averages for each appliance category (both prior year usage and stage averages are always shown in kWh).
+For each scenario, participants were asked to create two possible action plans to achieve the target reduction in total household energy usage (see @fig-task).
+Depending on their reference class condition, the target reduction amount presented either in kilowatt-hours (kWh), as percentages of total household usage, or in U.S.
+dollars.
+In all conditions, the target reduction was equivalent to a 15% reduction in total household kWh.
 
@@
 Participants are shown the prior year electricity use of a household, and are tasked with creating a plan for the next year that will meet the energy reduction goal.
-Study 1 manipulates the format of the reduction goal to be either a percentage (15% given as goal reduction), kilowatt hours (5965 kWh given), or USD ($656)](/assets/images/Smith_10R_Wells_15E.png){#fig-task}
+Study 1 manipulates the format of the reduction goal to be either a percentage (15% given as goal reduction), kilowatt hours (5965 kWh given), or USD (\$656)](/assets/images/Smith_10R_Wells_15E.png){#fig-task}
+
+Additional data collected included:
+
+-   **Energy Literacy Quiz**: An 8-item questionnaire assessing participants' knowledge of energy consumption and conversion.
+-   **Calculator Usage Tracking**: Questions determined whether participants used a calculator, paper/pen, or other methods to complete the tasks.
+-   **Demographic Survey**: Collected information on gender, age, income, education, employment status, and state of residence.
+-   **Environmental Attitudes Survey**: Assessed participants' pro-environmental attitudes and perceived importance of energy conservation.
 
 ## Results
+
+```{r}
+
+pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,ggplot2,gt,knitr,kableExtra,ggh4x,patchwork,lme4,flextable,pander)
+
+options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
+walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
+theme_set(theme_nice())
+
+s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
+  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
+  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))
+
+s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
+  filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
+  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))
+
+```
+
+```{r}
+#| label: tbl-s1-agg
+#| caption: "Study 1: Summary of planning accuracy by reference class.
+The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."
+
+s1_agg <- s1 |> 
+    filter(appliance !="Total kWh") |> 
+    group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
+    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
+                pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
+                n_change = sum(value!=family),
+                state_p_dif=mean(state_p_dif),
+                state_f_dif=mean(state_f_dif),
+                n_less_avg = sum(less_avg),
+                duration=first(Duration__in_seconds_)) |> 
+    mutate(matched_goal = (pct_change == pct_goal), 
+                error = pct_change - pct_goal,
+                abs_error = abs(error),
+                close_match = abs_error <= 0.03) |>
+    ungroup() |> # Add ungroup here
+        mutate(
+            accuracy_level = factor(
+                case_when(
+                    abs_error == 0.00 ~ "Exact match",
+                    abs_error <= 0.02 ~ "0.01-2% error",
+                    abs_error <= 0.15 ~ "2.01-15% error",
+                    TRUE ~ "Over 15% error"  # Capture all remaining cases
+                ), 
+                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+                ordered = TRUE
+            )
+        ) |> relocate(accuracy_level, .after= "pct_change")
+
+s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
+    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),mean_abs_error=mean(abs_error),n_accuracy=first(n_accuracy)) |> 
+    mutate(accuracy_level = factor(
+            case_when(
+                mean_abs_error < 0.02 ~ "Exact match",
+                mean_abs_error <= 0.02 ~ "0.01-2% error",
+                mean_abs_error <= 0.15 ~ "2.01-15% error",
+                TRUE ~ "Over 15% error"  # Capture all remaining cases
+            ), 
+            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+            ordered = TRUE
+        ))
+
+# proportion matching vs close match
+
+# proportion matching vs close match
+
+# overall pct of subjects who matched their goal
+s1_agg4 |> group_by(refClass) |>
+    summarise(
+    'Avg. % meeting goal' = mean(pct),
+    'Avg. Deviation From Goal' = median(mean_pct_change),
+    # sd = sd(pct),
+    # n = n(),
+    #se=sd(pct)/sqrt(n)
+) |>   mutate(across(where(is.numeric), \(x) round(x, 3))) %>% 
+  kable(escape=FALSE,booktabs=TRUE,align=c("l"))
+
+ #pander::pandoc.table(caption="Study 1: Proportion of participants who matched their goal overall")
+
+```
+
+For our primary analyses of participants' ability to create accurate energy-saving plans, we employed an accuracy level binning approach by categorizing responses into four distinct levels: Exact match, 0.01–2% error, 2.01–15% error, and Over 15% error.
+
+```{r}
+#| label: tbl-s1-prop
+#| caption: "Study 1: The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages).
+The combined group column reflects the percentage of participants in each accuracy level when aggregating across across all reference class conditions."
+
+# compute percentage of subjects per accuracy level per group
+observed_props_s1 <- s1_agg |>
+  group_by(refClass, accuracy_level) |>
+  summarise(n = n()) |>
+  group_by(refClass) |>
+  mutate(prop = n/sum(n)) |>
+  mutate(n_prop=paste0(n," (",round(prop*100,1),"%)" ), pct_grp=paste0(round(prop*100,1), "%")) |> ungroup()
+
+observed_props_s1 |> 
+  mutate(n_total=sum(n)/4) |> 
+  group_by(accuracy_level) |>
+  mutate(ns=sum(n)/4) |> 
+  mutate(Total = paste0(round(ns/n_total*100,1), "%")) |>
+  select('Reference Class'=refClass, 'Accuracy Level'=accuracy_level, '% in Group'=pct_grp, "Combined Groups %" =Total) |>
+  pivot_wider(
+    names_from = 'Reference Class',
+    values_from = c('% in Group')
+  ) |> relocate("Combined Groups %" , .after=last_col()) |> 
+  kable(escape=FALSE,booktabs=TRUE,align=c("l"))
+
+```
+
+```{r}
+#| label: fig-s1-plot
+#| caption: "Study 1: Proportion of participants in each accuracy level, colored by reference class. A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
+
+prop_acc_s1 <- s1_agg %>%
+    group_by(refClass, accuracy_level) %>%
+    summarise(count = n()) %>%
+    group_by(accuracy_level) |>
+    mutate(Probability = count / sum(count)) %>%
+    ungroup()
+
+ggplot(prop_acc_s1, aes(x = accuracy_level, y = Probability, fill = refClass)) +
+    geom_bar(stat = "identity", position = position_dodge()) +
+  scale_y_continuous(labels = scales::percent) +
+    labs(title = "S1. % of Participants within each Accuracy Bin",
+        x = "Accuracy Level",
+        y = "Percentage of Participants",
+        fill = "Reference Class") +
+    theme_minimal() 
+```
+
+We analyzed planning accuracy using Bayesian ordinal regression.
+The dependent variable, plan error, was computed by binning the goal deviation into four ordered levels: exact match (0% error), minor deviations (0.01-2% error), moderate deviations (2.01-15% error), and major deviations (>15% error).
+For each comparison, we provide posterior odds ratios (OR) and their 95% CIs.
+
+```{r}
+#| label: tbl-s1-ord
+#| caption: "Study 1: Odds ratios of for group comparisons. "
+
+ordinal_model_s1 <- brm(
+    accuracy_level ~ refClass + (1|id) + (1|state),
+    data = s1_agg,
+    family = cumulative("logit"),
+    cores = 4,
+    iter = 2000,
+    control = list(adapt_delta = 0.99), 
+    prior = c(prior(normal(0, 2), class = "Intercept"), 
+                prior(normal(0, 1.5), class = "b")), 
+    file = paste0(here::here("data/model_cache",'s1_ordinal3.rds')) 
+)
+
+#summary(ordinal_model_s1)
+
+# describe_posterior(ordinal_model_s1, centrality = "Mode") |> 
+#     filter(stringr::str_detect(Parameter, "b_")) |> 
+#     mutate(Parameter = stringr::str_remove(Parameter, "b_")) |> 
+#     mutate(across(where(is.numeric), round, 3)) |> 
+#     kable(booktabs = TRUE)
+
+# Get predicted probabilities
+pred_summary <- ordinal_model_s1 |>
+    epred_draws(newdata = data.frame(refClass = c("kWh", "Percentage", "USD")),
+                ndraws = 1000, re_formula = NA) |>
+    group_by(refClass, .category) |>
+    summarise(
+        mean_prob = mean(.epred),
+        lower_ci = quantile(.epred, 0.025),
+        upper_ci = quantile(.epred, 0.975)
+    )
+#pred_summary |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy")
+
+# Convert log-odds to odds ratios
+posterior_samples <- as.data.frame(ordinal_model_s1)
+odds_ratios <- data.frame(
+  Percentage_vs_kWh = exp(posterior_samples$b_refClassPercentage),
+  USD_vs_kWh = exp(posterior_samples$b_refClassUSD)
+)
+
+# Calculate summary statistics
+odds_ratio_summary <- data.frame(
+    comparison = c("Percentage vs kWh", "USD vs kWh"),
+    odds_ratio = c(mean(odds_ratios$Percentage_vs_kWh),
+                    mean(odds_ratios$USD_vs_kWh)),
+    ci_lower = c(quantile(odds_ratios$Percentage_vs_kWh, 0.025),
+                quantile(odds_ratios$USD_vs_kWh, 0.025)),
+    ci_upper = c(quantile(odds_ratios$Percentage_vs_kWh, 0.975),
+                quantile(odds_ratios$USD_vs_kWh, 0.975))
+)
+
+odds_ratio_summary |> kable(escape=FALSE,booktabs=TRUE,align=c("l")) 
+# |comparison        | odds_ratio| ci_lower| ci_upper|
+# |:-----------------|----------:|--------:|--------:|
+# |Percentage vs kWh |        2.3|     0.62|      6.2|
+# |USD vs kWh        |       14.1|     4.08|     38.0|
+
+#odds_ratio_summary |> pander::pandoc.table(caption="Study 1: Odds ratios of accuracy")
+
+```
+
+```{r}
+#| fig.width: 11
+#| fig.height: 7
+
+bayesplot::color_scheme_set(wes_palettes[1]$BottleRocket1[1:6])
+
+# Plot predicted probabilities
+ggplot(pred_summary, aes(x = refClass, y = mean_prob, fill = .category)) +
+    geom_bar(stat = "identity", position = position_dodge()) +
+    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci),
+                    position = position_dodge(width = 0.9), width = 0.2) +
+    labs(y = "Predicted Probability", x = "Reference Class", fill = "Accuracy Level") +
+    ggtitle("Study 1: Predicted Probabilities of Accuracy Level by Reference Class") +
+    theme_minimal()
+
+#pp_check(ordinal_model_s1) + ggtitle("Study 1: Posterior Predictive Checks for Ordinal Regression Model") + theme_minimal()
+
+pp_check(ordinal_model_s1, type = "bars_grouped", group="refClass", fatten = 2) +
+    # scale_x_continuous("y", breaks = 1:7) +
+    # scale_y_continuous(NULL, breaks = NULL, expand = expansion(mult = c(0, 0.05))) +
+    ggtitle("Data with posterior predictions",
+            subtitle = "S1. red bars are empirical data, black points are model predictions and credible intervals") +
+    labs(x = "Accuracy Level: 1= exact match,   +1= 0.01-2% error, +2= 2.01-15% error, +3= Over 15% error") +
+    theme_minimal()
+    theme(legend.background = element_blank() #,legend.position = c(.9, .75)
+    )
+
+```
 
@@
 
-The questions in study 2 are quite similar to those in study 1, but we also included a rounding manipulation (rounded vs. not rounded), and a manipulation of the goal (10% reduction vs. 15% rediction)
+The experimental procedures in study 2 are quite similar to those in study 1, but we also included a rounding manipulation (rounded vs.
+not rounded), and a manipulation of the goal (10% reduction vs.
+15% rediction).
+We recruited 206 participants from Amazon Mechanical Turk, but data from from 10 participants were corrupted due to experimenter error, leaving a final sample of 196 participants.
 
 Note that reference class remains a between-subjects variable, while percent goal, rounding, and state are within-subjects variables.
-
 In study 2, the new design is a 4 state temperature (2 warm vs.
@@
 kWh) between.
-A note on the percentage goals is that they are rounded to the full dollar.
-For instance, for the Davis family, 10% of their \$6,943.95 cost of residential energy usage is \$694.395; this is rounded to \$694.
-For the rounded condition, participants are also told that “The numbers for their \[the family’s\] usage have been rounded for convenience.” In terms of randomization, blocks were combined so that there was a block for each condition and order combination.
-As an example, one block is the smith family with a 10% goal and rounded numbers for the family’s usage first, and the Wells family second, with a 15% goal and exact numbers for the family’s usage.
-Another block has these but in reverse, where the participant receives the Wells family first and the Smith family second.
-In total, there are 96 conditions, representing each possible order (2 orders x 4 states x 2 possible significant digit values, and 3 reference classes).
 
@@
 
+```{r}
+#| label: tbl-s2-agg
+#| caption: "Study 2: Summary of planning accuracy by reference class.
+The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."
+
+s2_agg1 <- s2_long |> 
+  filter(appliance != "TOTAL") |> 
+  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
+  summarise(
+    total_kWh = sum(value),
+    orig_kWh = sum(family),
+    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
+    state_dif = mean(state_dif),
+    .groups = "drop"
+  ) |>
+  mutate(matched_goal = (pct_change == pct),
+                error = pct_change - pct,
+                abs_error = abs(error)) |> 
+      ungroup() |> # Add ungroup here
+        mutate(
+            accuracy_level = factor(
+                case_when(
+                    abs_error == 0.00 ~ "Exact match",
+                    abs_error <= 0.02 ~ "0.01-2% error",
+                    abs_error <= 0.15 ~ "2.01-15% error",
+                    TRUE ~ "Over 15% error"  # Capture all remaining cases
+                ), 
+                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+                ordered = TRUE
+            )
+        )
+
+s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc) |> 
+    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),mean_abs_error=mean(abs_error),n_accuracy=first(n_accuracy)) |> 
+    mutate(accuracy_level = factor(
+            case_when(
+                mean_abs_error < 0.02 ~ "Exact match",
+                mean_abs_error <= 0.02 ~ "0.01-2% error",
+                mean_abs_error <= 0.15 ~ "2.01-15% error",
+                TRUE ~ "Over 15% error"  # Capture all remaining cases
+            ), 
+            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+            ordered = TRUE
+        ))
+
+# overall pct of subjects who matched their goal
+s2_agg4 |> group_by(refClass) |>
+    summarise(
+    'Avg. % meeting goal' = mean(pct),
+    'Avg. Deviation From Goal' = median(mean_pct_change),
+    # sd = sd(pct),
+    # n = n(),
+    #se=sd(pct)/sqrt(n)
+) |>   mutate(across(where(is.numeric), \(x) round(x, 3))) %>% 
+  kable(escape=FALSE,booktabs=TRUE,align=c("l"))
+
+```
+
+```{r}
+#| label: tbl-s2-prop
+#| caption: "Study 2: The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages).
+The combined group column reflects the percentage of participants in each accuracy level when aggregating across across all reference class conditions."
+
+# compute percentage of subjects per accuracy level per group
+observed_props_s2 <- s2_agg1 |>
+  group_by(refClass, accuracy_level) |>
+  summarise(n = n()) |>
+  group_by(refClass) |>
+  mutate(prop = n/sum(n)) |>
+  mutate(n_prop=paste0(n," (",round(prop*100,1),"%)" ), pct_grp=paste0(round(prop*100,1), "%")) |> ungroup()
+
+observed_props_s2 |> 
+  mutate(n_total=sum(n)/4) |> 
+  group_by(accuracy_level) |>
+  mutate(ns=sum(n)/4) |> 
+  mutate(Total = paste0(round(ns/n_total*100,1), "%")) |>
+  select('Reference Class'=refClass, 'Accuracy Level'=accuracy_level, '% in Group'=pct_grp, "Combined Groups %" =Total) |>
+  pivot_wider(
+    names_from = 'Reference Class',
+    values_from = c('% in Group')
+  ) |> relocate("Combined Groups %" , .after=last_col()) |> 
+  kable(escape=FALSE,booktabs=TRUE,align=c("l"))
+
+```
+
+```{r}
+#| label: fig-s2-plot
+#| caption: "Study 2: Proportion of participants in each accuracy level, colored by reference class. A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
+
+# % of entire sample 
+prop_combo_s2 <- s2_agg1 %>%
+    group_by(refClass, accuracy_level) %>%
+    summarise(count = n()) %>%
+    group_by(refClass) %>%
+    mutate(Probability = count / sum(count)) %>%
+    ungroup()
+
+ggplot(prop_combo_s2, aes(x = accuracy_level, y = Probability, fill = refClass)) +
+    geom_bar(stat = "identity", position = position_dodge()) +
+  scale_y_continuous(labels = scales::percent) +
+    labs(title = "S1. % of Participants within each Accuracy Bin",
+        x = "Accuracy Level",
+        y = "Percentage of Participants",
+        fill = "Reference Class") +
+    theme_minimal() 
+```
+
+```{r}
+#| label: fig-s2-plot2
+#| caption: "Study 2: Proportion of participants in each accuracy level, colored by reference class, and seprated in facets based on the levels of reduction goal, and rounding.
+A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
+#| fig.width: 11
+#| fig.height: 9
+
+# % of entire sample 
+prop_combo_s2_full <- s2_agg1 %>%
+    group_by(refClass, accuracy_level, pct_goal,rounded) %>%
+    summarise(count = n()) %>%
+    group_by(refClass) %>%
+    mutate(Probability = count / sum(count)) %>%
+    ungroup()
+
+ggplot(prop_combo_s2_full, aes(x = accuracy_level, y = Probability, fill = refClass)) +
+    geom_bar(stat = "identity", position = position_dodge()) +
+  scale_y_continuous(labels = scales::percent) +
+    labs(title = "S1. % of Participants within each Accuracy Bin",
+        x = "Accuracy Level",
+        y = "Percentage of Participants",
+        fill = "Reference Class") +
+    facet_wrap(~pct_goal*rounded) +
+    theme_minimal() 
+```
+
+### Individual Differences
+
+```{r}
+#| label: fig-s2-indv
+#| caption: "Study 2: Individual performance in the energy planning task, colored by reference class.
+The dashed line represents the target reduction goal.
+Participants are shown along the y axis, those who fall above or below the dashed line have not met the target goal.
+The x-axis represents the percent change in energy usage from the prior year."
+#| fig.width: 11
+#| fig.height: 9
+
+s2_agg1 |> group_by(id,refClass,calc,pct_goal,pct_change) |> 
+    filter(plan=="plan1",rounded=="Rounded") |> 
+    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),mean_abs_error=mean(abs_error),n_accuracy=first(n_accuracy)) |>
+  ungroup() |> 
+  mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100) |>
+  filter(mean_abs_error <= 0.50) |>
+  mutate(id=reorder(id,pct_change)) |> 
+  ggplot(aes(y=id,x=mean_pct_change,col=refClass)) + 
+  geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
+  geom_vline(aes(xintercept=goal_pct),linetype="dashed",alpha=.5) +
+  ggh4x::facet_nested_wrap(~pct_goal,axes="all",scales="free",ncol=2)  + 
+  labs(y="Participant Id", x="Percent Change", title="Individual Performance") +
+  theme(axis.text.y = element_text(family = "Manrope Light", face = "plain", size = rel(0.7))) + 
+    scale_x_continuous(breaks = seq(0, 0.5, by = 0.05),
+                    labels = scales::percent_format(accuracy = 1))
+
+```
+
+```{r}
+#| label: fig-s2-indv2
+#| caption: "Study 2: Respones patterns for a subset of individiual participants.
+Black points are participant responses, red points are the state average, and blue points are the family average.
+The x-axis represents the appliance category, and the y-axis represents the energy usage in kWh."
+#| fig.width: 11
+#| fig.height: 10
+
+s2_long |> filter(id %in% unique(s2_long$id)[1:30]) |> 
+  filter(appliance!="TOTAL",state=="California") |> 
+  ggplot(aes(x=appliance,y=value)) + geom_point(aes(shape=plan)) +
+  geom_point(aes(y=state_avg),color="red") +
+  geom_point(aes(y=family),color="blue") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+#| 
+s2_long |> filter(id %in% unique(s2_long$id)[1:30]) |> 
+  filter(appliance!="TOTAL",state=="California") |> 
+  ggplot(aes(x=appliance,y=value)) + geom_point(aes(shape=plan)) +
+  geom_point(aes(y=state_avg),color="red") +
+  geom_point(aes(y=family),color="blue") +
+  facet_wrap(~id) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
 # Discusion
+
+# References

### 2024-11-20 nb/overview.qmd
update with some results

@@
 ---
-title: "Effect of Reference Class on Planning to Conserve"
-author:
-- name: Thomas E. Gorman
-  url: https://tegorman13.github.io/
-  affiliations: 
-  - name:  Communication and Cognition Lab, Purdue University, USA
-    affiliation-url: https://web.ics.purdue.edu/~treimer/
-#lightbox: true
-toc: false
-code-fold: true
-code-tools: true
+title: "Results Overview"
 execute:
@@
 
-pacman::p_load(dplyr,purrr,tidyr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,
-               ggplot2,gt,knitr,kableExtra,ggh4x,lme4,flextable,pander)
+pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,ggplot2,gt,knitr,kableExtra,ggh4x,patchwork,lme4,flextable,pander)
+
 options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
@@
 
-s1 <- readRDS(here::here("data/s1_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds"))))
-s2_long <- readRDS(here::here("data/s2_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds"))))
+s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
+  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
+  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))
+
+s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
+  filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
+  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))
 
@@
 
+### Study 1 Subject Counts
+
 ```{r}
+#| label: tbl-s1-counts
+# total subjects
+n_s1 <- s1 |> ungroup() |> summarise(n=n_distinct(id))
+n_s1_removed <-  readRDS(here::here("data/s1_discrep_ids.rds")) |> length()
+
+print(paste("Study 1: Total subjects:", n_s1$n, "Removed subjects:", n_s1_removed))
 
@@
 ### Study 2 - Subject counts
+
 ```{r}
+
+n_s2 <- s2_long |> ungroup() |> summarise(n=n_distinct(id))
+n_s2_removed <-  readRDS(here::here("data/s2_discrep_ids.rds")) |> length()
+
+print(paste("Study 2: Total subjects:", n_s2$n, "Removed subjects:", n_s2_removed))
 
@@
 
-### Study 2
+### Study 2 Performance
 
@@
 
+## Group comparison via proportions
+
+### Study 1
+
+```{r}
+#| label: tbl-s1-agg
+
+s1_agg <- s1 |> 
+    filter(appliance !="Total kWh") |> 
+    group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
+    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
+                pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
+                n_change = sum(value!=family),
+                state_p_dif=mean(state_p_dif),
+                state_f_dif=mean(state_f_dif),
+                n_less_avg = sum(less_avg),
+                duration=first(Duration__in_seconds_)) |> 
+    mutate(matched_goal = (pct_change == pct_goal), 
+                error = pct_change - pct_goal,
+                abs_error = abs(error),
+                close_match = abs_error <= 0.03) |>
+    ungroup() |> # Add ungroup here
+        mutate(
+            accuracy_level = factor(
+                case_when(
+                    abs_error == 0.00 ~ "Exact match",
+                    abs_error <= 0.02 ~ "0.01-2% error",
+                    abs_error <= 0.15 ~ "2.01-15% error",
+                    TRUE ~ "Over 15% error"  # Capture all remaining cases
+                ), 
+                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+                ordered = TRUE
+            )
+        ) |> relocate(accuracy_level, .after= "pct_change")
+
+s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
+    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),mean_abs_error=mean(abs_error),n_accuracy=first(n_accuracy)) |> 
+    mutate(accuracy_level = factor(
+            case_when(
+                mean_abs_error < 0.02 ~ "Exact match",
+                mean_abs_error <= 0.02 ~ "0.01-2% error",
+                mean_abs_error <= 0.15 ~ "2.01-15% error",
+                TRUE ~ "Over 15% error"  # Capture all remaining cases
+            ), 
+            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+            ordered = TRUE
+        ))
+
+# s1_agg |> arrange(id,accuracy_level) |> 
+#   select(id,refClass,state,block,plan,total_kWh,orig_kWh,pct_change,accuracy_level,n_change,matched_goal,error,abs_error) |> print(n=20)
+
+observed_props <- s1_agg |>
+  group_by(refClass, accuracy_level) |>
+  summarise(n = n()) |>
+  group_by(refClass) |>
+  mutate(prop = n/sum(n)) |>
+  pivot_wider(
+    names_from = accuracy_level,
+    values_from = c(n, prop)
+  )
+
+prop_acc_s1 <- s1_agg %>%
+    group_by(refClass, accuracy_level) %>%
+    summarise(count = n()) %>%
+    group_by(refClass) %>%
+    mutate(Probability = count / sum(count)) %>%
+    ungroup()
+
+# compute percentage of subjects per accuracy level per group
+observed_props_s1 <- s1_agg |>
+  group_by(refClass, accuracy_level) |>
+  summarise(n = n()) |>
+  group_by(refClass) |>
+  mutate(prop = n/sum(n)) |>
+  mutate(n_prop=paste0(n," (",round(prop*100,1),"%)" ), pct_grp=paste0(round(prop*100,1), "%")) |> ungroup()
+
+observed_props_s1 |> 
+  mutate(n_total=sum(n)/4) |> 
+  group_by(accuracy_level) |>
+  mutate(ns=sum(n)/4) |> 
+  mutate(Total = paste0(round(ns/n_total*100,1), "%")) |>
+  select('Reference Class'=refClass, 'Accuracy Level'=accuracy_level, '% in Group'=pct_grp, "Combined Groups %" =Total) |>
+  pivot_wider(
+    names_from = 'Reference Class',
+    values_from = c('% in Group')
+  ) |> relocate("Combined Groups %" , .after=last_col()) |> 
+  pander::pandoc.table(caption="Study 1: Proportion of participants per accuracy level by condition")
+
+# observed_props_s1 |> 
+#   select('Reference'=refClass, 'Accuracy Level'=accuracy_level, 'n (% group)'=n_prop) |>
+#   pivot_wider(
+#     names_from = 'Accuracy Level',
+#     values_from = c('n (% group)')
+#   )
+
+# % of accuracy_level
+prop_acc_s1 <- s1_agg %>%
+    group_by(refClass, accuracy_level) %>%
+    summarise(count = n()) %>%
+    group_by(accuracy_level) |>
+    mutate(Probability = count / sum(count)) %>%
+    ungroup()
+
+# % of entire sample 
+prop_acc_s1 <- s1_agg %>%
+    group_by(refClass, accuracy_level) %>%
+    summarise(count = n()) %>%
+    ungroup() |> 
+    mutate(Probability = count / sum(count)) %>%
+    ungroup()
+
+```
+
+::: {.content-visible when-format="html"}
+
+```{r}
+#| label: tbl-s1-interactive
+
+s1_agg4 |> 
+    DT::datatable(extensions = 'Buttons',
+    options = list(
+       dom = 'Blfrtip',
+       buttons = list('copy', 'csv', 'excel', 'pdf', 'print'),
+      pageLength = 25,
+      autoWidth = TRUE
+    ),
+    filter = 'top', caption="Exp 1. Participant global averages")
+
+s1_agg |> 
+  select(id,refClass,calc,state,plan,pct_change,accuracy_level,n_change,matched_goal,abs_error,total_kWh,orig_kWh) |> 
+    group_by(id,refClass,calc) |> 
+    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+    mutate(mean_abs_error=mean(abs_error), .after=abs_error) |> 
+    mutate(across(where(is.numeric), \(x) round(x, 3))) %>%
+    DT::datatable(extensions = 'Buttons',
+    options = list(
+       dom = 'Blfrtip',
+       buttons = list('copy', 'csv', 'excel', 'pdf', 'print'),
+      pageLength = 25,
+      autoWidth = TRUE
+    ),
+    filter = 'top', caption="Exp 1. Participant trial level performance")
+
+s1_agg |> filter(id==124) |> 
+  select(id,refClass,state,block,plan,total_kWh,orig_kWh,pct_change,accuracy_level,n_change,matched_goal,abs_error) |> 
+  group_by(id) |> mutate(mean_abs_error=mean(abs_error), .after=abs_error) |> 
+  pander::pandoc.table(caption="Study 1: Example of participant performance (sbj 124")
+
+```
+
+:::
+
+### Study 2
+
+```{r}
+#| label: tbl-s2-agg
+#| fig-width: 12
+
+s2_agg1 <- s2_long |> 
+  filter(appliance != "TOTAL") |> 
+  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
+  summarise(
+    total_kWh = sum(value),
+    orig_kWh = sum(family),
+    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
+    state_dif = mean(state_dif),
+    .groups = "drop"
+  ) |>
+  mutate(matched_goal = (pct_change == pct),
+                error = pct_change - pct,
+                abs_error = abs(error)) |> 
+      ungroup() |> # Add ungroup here
+        mutate(
+            accuracy_level = factor(
+                case_when(
+                    abs_error == 0.00 ~ "Exact match",
+                    abs_error <= 0.02 ~ "0.01-2% error",
+                    abs_error <= 0.15 ~ "2.01-15% error",
+                    TRUE ~ "Over 15% error"  # Capture all remaining cases
+                ), 
+                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+                ordered = TRUE
+            )
+        )
+
+s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc) |> 
+    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),mean_abs_error=mean(abs_error),n_accuracy=first(n_accuracy)) |> 
+    mutate(accuracy_level = factor(
+            case_when(
+                mean_abs_error < 0.02 ~ "Exact match",
+                mean_abs_error <= 0.02 ~ "0.01-2% error",
+                mean_abs_error <= 0.15 ~ "2.01-15% error",
+                TRUE ~ "Over 15% error"  # Capture all remaining cases
+            ), 
+            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+            ordered = TRUE
+        ))
+
+# s1_agg |> arrange(id,accuracy_level) |> 
+#   select(id,refClass,state,block,plan,total_kWh,orig_kWh,pct_change,accuracy_level,n_change,matched_goal,error,abs_error) |> print(n=20)
+
+# compute percentage of subjects per accuracy level per group
+observed_props <- s2_agg1 |>
+  group_by(refClass, accuracy_level) |>
+  summarise(n = n()) |>
+  group_by(refClass) |>
+  mutate(prop = n/sum(n)) |>
+  mutate(n_prop=paste0(n," (",round(prop*100,1),"%)" ))
+
+# confirm that the proportions are with respect to the total number of subjects in each refClass
+# 74/.268 = 276; 69*4=276;  38/.138=275;  .435 + .0797 + .21 + .275 = 1
+
+# compute percentage of subjects per accuracy level per group
+observed_props_s2 <- s2_agg1 |>
+  group_by(refClass, accuracy_level) |>
+  summarise(n = n()) |>
+  group_by(refClass) |>
+  mutate(prop = n/sum(n)) |>
+  mutate(n_prop=paste0(n," (",round(prop*100,1),"%)" ), pct_grp=paste0(round(prop*100,1), "%")) |> ungroup()
+
+observed_props_s2 |> 
+  mutate(n_total=sum(n)/4) |> 
+  group_by(accuracy_level) |>
+  mutate(ns=sum(n)/4) |> 
+  mutate(Total = paste0(round(ns/n_total*100,1), "%")) |>
+  select('Reference Class'=refClass, 'Accuracy Level'=accuracy_level, '% in Group'=pct_grp, "Combined Groups %" =Total) |>
+  pivot_wider(
+    names_from = 'Reference Class',
+    values_from = c('% in Group')
+  ) |> relocate("Combined Groups %" , .after=last_col()) |> 
+  pander::pandoc.table(caption="Study 1: Proportion of participants per accuracy level by condition")
+
+observed_props_s2 |>  select('Reference'=refClass, 'Accuracy Level'=accuracy_level, 'n (% group)'=n_prop) |>
+  pivot_wider(
+    names_from = 'Accuracy Level',
+    values_from = c('n (% group)')
+  )
+
+# % of accuracy_level
+prop_acc_s2 <- s2_agg1 %>%
+    group_by(refClass, accuracy_level) %>%
+    summarise(count = n()) %>%
+    group_by(accuracy_level) |>
+    mutate(Probability = count / sum(count)) %>%
+    ungroup()
+
+# % of entire sample 
+prop_combo_s2 <- s2_agg1 %>%
+    group_by(refClass, accuracy_level) %>%
+    summarise(count = n()) %>%
+    group_by(refClass) %>%
+    mutate(Probability = count / sum(count)) %>%
+    ungroup()
+
+```
+
+## Study 2 proportion in accuracy plots
+
+```{r}
+
+p1 <- ggplot(prop_combo_s2, aes(x = accuracy_level, y = Probability, fill = refClass)) +
+    geom_bar(stat = "identity", position = position_dodge()) +
+  scale_y_continuous(labels = scales::percent) +
+    labs(title = "Within-Group %'s per Accuracy bin",
+        x = "Accuracy Level",
+        y = "Percentage of Participants",
+        fill = "Reference Class") +
+    theme_minimal()
+
+ p2 <- ggplot(data = s2_agg1, aes(x = accuracy_level, fill = refClass)) +
+  geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
+  scale_y_continuous(labels = scales::percent) +
+  labs(title = "Combined Participant %'s per bin",
+       x = "Accuracy Level",
+       y = "Percentage of Participants",
+       fill = "Reference Class") +
+  theme_minimal()
+
+ p3 <- ggplot(prop_acc_s2, aes(x = accuracy_level, y = Probability, fill = refClass)) +
+    geom_bar(stat = "identity", position = position_dodge()) +
+  scale_y_continuous(labels = scales::percent) +
+    labs(title = "Aggregate by Accuracy Bin",
+        x = "Accuracy Level",
+        y = "Percentage of Participants",
+        fill = "Reference Class") +
+    theme_minimal()
+
+ # combined plots, and unify axes and legends
+{ p1 + p2 +p3 + plot_layout(guides = 'collect') & theme(legend.position = 'right') } + plot_annotation(title = "Study 2: Accuracy Level Proportions")
+
+```
+
 ## Analysis
@@
 ```{r}
+#| eval: false
+#| include: false
+
+pacman::p_load(brms,rstan,bayestestR,emmeans,tidybayes,modelsummary)
 
 # binomial regression on trial level data
-
 s1_br <- brm(
@@
 ```{r}
+#| eval: false
+#| include: false
 
@@
 ```{r}
+#| eval: false
+#| include: false
+
 s1_glm <- glm(matched_goal ~ refClass, data = s1_agg, family = binomial())
@@
 ```{r}
-
+#| eval: false
+#| include: false
 # binomial regression on trial level data
@@
 ```{r}
-
+#| eval: false
+#| include: false
+#| 
 s2_glm <- glm(matched_goal ~ refClass + rounded + pct_goal, data = s2_agg1, family = binomial())

### 2024-11-20 paper.qmd
update with some results

@@
 
+  <a class="btn btn-warning btn-sm mb-1" target="_blank" href="/Manuscript/output/manuscript.docx">
+  <iconify-icon inline="" icon="fa6-solid:file-arrow-down"></iconify-icon> &ensp;docx <small>(download)</small>
+  </a>
+
 </p>
 
-<iframe src="/manuscript/output/manuscript.html" frameborder="0" style="overflow:hidden;height:88%;width:100%;border:2px solid #ccc;" height="88%" width="100%" name="manuscript"></iframe>
+<iframe src="/manuscript/output/manuscript.html" frameborder="0" style="overflow:hidden;height:75%;width:95%;border:2px solid #ccc;" height="75%" width="95%" name="manuscript"></iframe>
 ```

### 2024-11-20 Scripts/setup.R

@@ -1,0 +1,12 @@
+
+
+
+
+#| label: setup
+#| include: false
+
+knitr::opts_chunk$set(
+    echo = FALSE, include = FALSE, warning = FALSE, message = FALSE,
+    fig.width = 6, fig.height = (6 * 0.618), fig.retina = 3,
+    out.width = "100%", fig.align = "center"
+)

### 2024-12-04 manuscript/manuscript.qmd
introduction outline. Fixed plot bug.

@@
     affiliations: 
-        - name:  Purdue University, USA
+      - ref: cclab
+      - ref: clara
+    attributes: 
+      corresponding: true
+  - name: Torsten Reimer
+    url: https://web.ics.purdue.edu/~treimer/
+    orcid: 0000-0002-7419-0076
+    affiliations:
+      - ref: cclab
+      - ref: clara
+  - name: Juan Pablo Loaiza Ramirez
+    url: https://www.cla.purdue.edu/directory/profiles/loaiza-ramirez,juan-pablo.html
+    orcid: 0000-0001-9663-0522
+    affiliations:
+      - ref: cclab
+      - ref: clara
+  - name: Hayden Barber
+    url: https://www.sdstate.edu/directory/hayden-barber
+    orcid: 0000-0002-3465-8615
+    affiliations:
+      - ref: sdsu
+affiliations:
+  - id: cclab
+    name: Communication and Cognition Lab, Purdue University
     city: West Lafayette
@@
     postal-code: 47907
-          affiliation-url: https://web.ics.purdue.edu/~treimer/
-      attributes: 
-          corresponding: true
-    - name: Torsten Reimer
-      affiliation: Brian Lamb School of Communication, Purdue University
-      affiliation-url: https://cla.purdue.edu/communication/
     url: https://web.ics.purdue.edu/~treimer/
-      orcid: 0000-0002-7419-0076
-    - name: Juan Pablo Loaiza Ramirez
-      affiliation: Brian Lamb School of Communication, Purdue University
-      affiliation-url: https://cla.purdue.edu/communication/
-      url: https://web.ics.purdue.edu/~treimer/
-      orcid: 0000-0001-9663-0522
+  - id: clara
+    name: College of Liberal Arts Research Academy
+    department: Purdue University
+    city: West Lafayette
+    region: IN
+    country: USA
+    postal-code: 47907
+    url: https://cla.purdue.edu/about/college-initiatives/research-academy/
+  - id: sdsu
+    name: School of Communication & Journalism, South Dakota State University
+    city: Brookings
+    region: SD
+    country: USA
+    url: https://www.sdstate.edu/directory/hayden-barber
 format:
@@
                 </style>
-    # hugo-md:
-    #     include: true
-    #     html-math-method: mathjax
-    #     output-file: manuscript_hugo.md
+    hugo-md:
+        include: true
+        echo: false
+        html-math-method: mathjax
+        output-file: manuscript_hugo.md
     gfm:
@@
 #       icon: file-pdf
-date: 2024-11-19
+date: 2024-12-03
 abstract: |
-    This study examines how presenting energy information in different formats (kWh, %, USD) affects planning accuracy.
-Across two experiments, the kWh format generally led to better accuracy, while the USD format consistently led to the worst performance.
-These findings highlight the importance of effective information presentation to promote energy conservation.
+    This study aims to examine how the format of energy information impacts individuals’ ability to develop precise energy reduction plans.
+By manipulating the reference class (kWh, %, USD) and assessing planning accuracy, we seek to determine which format facilitates better comprehension and decision-making.Across two experiments, the kWh format generally led to better accuracy, while the USD format consistently led to the worst performance.
+These findings highlight the importance of effective information presentation to promote energy conservation, and may contribute to the development of more effective energy communication strategies that can enhance conservation efforts.
 ---
@@
 
+- highlight the significant contribution of residential energy consumption to carbon emissions and the potential for substantial reductions.
+Make point about urgency of climate change?
+(IPCC (2014) and EIA (2012) ).
+
+- Electricity bills are a primary source of energy-use information for consumers and offer a promising avenue for enhancing communication about energy consumption (Fischer, 2008)
+
 ### Literature Review
+
+- Canfield et al.
+(2017) found that tables were more effective than graphs for conveying specific electricity usage data, likely because tables facilitate straightforward point reading.
+However, they also noted that individuals with lower energy literacy had reduced comprehension across all formats.
+- Canfield et al.'s (2017) findings on preferences for historical use information and the impact of neighbor comparisons
+- The concept of cognitive fit posits that performance improves when the information presentation format aligns with the task requirements (Vessey, 1991)
+- alignment can reduce cognitive load and enhance accuracy in planning (Shah & Freedman, 2009)
+- Reimer et al. (2015) provide  context on how numerical formats affect risk perception, the reference class problem, and the benefits of natural frequencies.
 
@@
 Prior research by @canfieldPerceptionsElectricityuseCommunications2017 demonstrated that presenting energy information in tabular formats enhances comprehension relative to graphs.
+
+The way numerical information is presented can significantly affect how individuals process and use that information (Reimer et al., 2015).
+The reference class problem highlights that numbers without clear reference points can lead to misinterpretation, as the meaning of a statistic depends on the category or class it refers to (Gigerenzer & Edwards, 2003; Reimer et al., 2015).
+Presenting energy information in absolute units (e.g., kWh) provides a clear reference class, potentially enhancing comprehension.
+
+Furthermore, research suggests that natural frequencies and absolute numbers are generally easier for individuals to understand compared to percentages or probabilities (Gigerenzer & Hoffrage, 1995; Hoffrage et al., 2000).
+In the context of energy conservation, using absolute units may facilitate more accurate planning and decision-making by aligning with intuitive cognitive processing.
+
+Despite existing studies on energy-use communication and format effects, limited research has explored how different numerical representations influence consumers’ ability to create accurate energy conservation plans.
+Specifically, there is a gap in understanding how presenting energy information in absolute units versus percentages or monetary terms affects the precision of planning appliance-specific reductions.
+Addressing this gap is crucial for developing effective interventions that promote energy conservation behaviors.
 
@@
     labs(x = "Accuracy Level: 1= exact match,   +1= 0.01-2% error, +2= 2.01-15% error, +3= Over 15% error") +
-    theme_minimal()
+    theme_minimal() +
     theme(legend.background = element_blank() #,legend.position = c(.9, .75)
@@
   labs(y="Participant Id", x="Percent Change", title="Individual Performance") +
-  theme(axis.text.y = element_text(family = "Manrope Light", face = "plain", size = rel(0.7))) + 
+  theme(axis.text.y=element_text(face = "plain", size = rel(0.7))) + 
     scale_x_continuous(breaks = seq(0, 0.5, by = 0.05),

### 2024-12-04 Scripts/test.R

@@ -27,33 +27,8 @@ ggplot(data, aes(x = value, fill = group)) +
   xlim(-5, 6)
 
 
+# test 
 
 
 
-
-
-# Load the ggplot2 package
-library(ggplot2)
-
-# Create a sample dataset
-set.seed(123)
-data <- data.frame(
-  value = c(rnorm(1000, mean = 0, sd = 1), rnorm(1000, mean = 1, sd = 1)),
-  group = rep(c("A", "B"), each = 1000)
-)
-
-# Define thresholds
-thresholds <- c(-1, 0, 1, 2)
-
-# Create the plot
-ggplot(data, aes(x = value, fill = group)) +
-  geom_density(alpha = 0.5) +
-  scale_fill_manual(values = c("orange", "brown")) +
-  geom_vline(xintercept = thresholds, linetype = "dotted") +
-  labs(title = "Underlying latent scale for Y, given X",
-       fill = "X") +
-  theme_minimal() +
-  xlab(NULL) +
-  ylab(NULL) +
-  xlim(-5, 6) +
-  annotate("text", x = thresholds, y = 0.4, label = c(expression(theta[1]), expression(theta[2]), expression(theta[3]), expression(theta[4])), angle = 0, vjust = -0.5,hjust=.2)
+#simulate from a guassian

### 2024-12-07 docs/misc/discrep.qmd
update results section. And some formating improvements.

--- 2024-11-05 misc/discrep.qmd
@@
 title: "Discrep_Patterns"
-author:
-- name: Thomas E. Gorman
-  url: https://tegorman13.github.io/
-  affiliations: 
-  - name:  Communication and Cognition Lab, Purdue University, USA
-    affiliation-url: https://web.ics.purdue.edu/~treimer/
-lightbox: true
 toc: false
-code-fold: show
-code-tools: true
 execute:

### 2024-12-07 docs/misc/odds_ratio.qmd
update results section. And some formating improvements.

```{r}

pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,ggplot2,gt,knitr,kableExtra,ggh4x,patchwork,lme4,flextable,pander)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)

walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
theme_set(theme_nice())

s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                close_match = abs_error <= 0.03) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

```

```{r}
#| label: tbl-s1-ord
#| tbl-cap: "Study 1: Odds ratios of for group comparisons. "

ordinal_model_s1 <- brm(
    accuracy_level ~ refClass + (1|id) + (1|state),
    data = s1_agg,
    family = cumulative("logit"),
    cores = 4,
    iter = 5000,
    save_pars = save_pars(all = TRUE),
    control = list(adapt_delta = 0.99), 
    prior = c(prior(normal(0, 3), class = "Intercept"), 
                prior(normal(0, 3), class = "b")), 
    file = paste0(here::here("data/model_cache",'s1_ordinal3.rds')) 
)
# 
# ordinal_model_s1_null <- brm(
#     accuracy_level ~  (1|id) + (1|state),
#     data = s1_agg,
#     family = cumulative("logit"),
#     cores = 4,
#     iter = 5000,
#     save_pars = save_pars(all = TRUE),
#     control = list(adapt_delta = 0.99), 
#     prior = c(prior(normal(0, 3), class = "Intercept")),
#     file = paste0(here::here("data/model_cache",'s1_null.rds')) 
# )

#colnames(mted1) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")

as.data.frame(describe_posterior(ordinal_model_s1, centrality = "Mean"))[, c(1,2,4,5,6)] |> 
  setNames(c("Parameter", "Estimate", "CI_Lower", "CI_Upper", "pd")) |> 
  mutate(Parameter = stringr::str_remove(Parameter, "b_"))

# ordinal_model_s1 |> emmeans(~refClass) |> contrast(method="pairwise") |>  
#   gather_emmeans_draws() %>%
#   mean_hdi(.width = .95)

# describe_posterior(ordinal_model_s1, centrality = "Mode") |> 
#     filter(stringr::str_detect(Parameter, "b_")) |> 
#     mutate(Parameter = stringr::str_remove(Parameter, "b_")) |> 
#     mutate(across(where(is.numeric), round, 3)) |> 
#     kable(booktabs = TRUE)

# Get predicted probabilities
pred_summary <- ordinal_model_s1 |>
    epred_draws(newdata = data.frame(refClass = c("kWh", "Percentage", "USD")),
                ndraws = 1000, re_formula = NA) |>
    group_by(refClass, Category=.category) |>
    summarise(
        mean_prob = mean(.epred),
        lower_ci = quantile(.epred, 0.025),
        upper_ci = quantile(.epred, 0.975)
    )
#pred_summary |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy")

# Convert log-odds to odds ratios
posterior_samples <- as.data.frame(ordinal_model_s1)
odds_ratios <- data.frame(
  Percentage_vs_kWh = exp(posterior_samples$b_refClassPercentage),
  USD_vs_kWh = exp(posterior_samples$b_refClassUSD)
)

# Calculate summary statistics
odds_ratio_summary <- data.frame(
    comparison = c("Percentage vs kWh", "USD vs kWh"),
    odds_ratio = c(mean(odds_ratios$Percentage_vs_kWh),
                    mean(odds_ratios$USD_vs_kWh)),
    ci_lower = c(quantile(odds_ratios$Percentage_vs_kWh, 0.025),
                quantile(odds_ratios$USD_vs_kWh, 0.025)),
    ci_upper = c(quantile(odds_ratios$Percentage_vs_kWh, 0.975),
                quantile(odds_ratios$USD_vs_kWh, 0.975))
)

odds_ratio_summary |> kable(escape=FALSE,booktabs=TRUE,align=c("l")) 
# |comparison        | odds_ratio| ci_lower| ci_upper|
# |:-----------------|----------:|--------:|--------:|
# |Percentage vs kWh |        2.3|     0.62|      6.2|
# |USD vs kWh        |       14.1|     4.08|     38.0|

#odds_ratio_summary |> pander::pandoc.table(caption="Study 1: Odds ratios of accuracy")

# odds ratios of fixed effects
as.data.frame(fixef(ordinal_model_s1)[,-2])|> as.data.frame() %>%
    rownames_to_column(var = "Parameter") %>%
    mutate(across(where(is.numeric), exp)) |>
    filter(!stringr::str_detect(Parameter, "Intercept"))

describe_posterior(ordinal_model_s1, centrality = "Median")[, c(1,2,4,5)] %>%
  as_tibble() |> # remove rows with Intercept
  mutate(across(where(is.numeric), exp)) |>
  filter(!stringr::str_detect(Parameter, "Intercept")) |>
  mutate(Parameter = stringr::str_remove(Parameter, "b_"))

or_table <- data.frame(exp(fixef(ordinal_model_s1)))
or_table |>
  rownames_to_column(var = "Term") |>
  as_flextable() |>
  set_header_labels(values = list(
    OR = "Odds Ratio",
    `2.5 %` = "Lower 95% CI",
    `97.5 %` = "Upper 95% CI"
  )) |>
  flextable::add_footer_lines(values = "Note: Odds ratios derived from the ordinal logistic regression model.") |> 
  align(align = 'center', part = 'all') |>
  fontsize(size = 10, part = "all")

posterior_draws <- as_draws_df(ordinal_model_s1) |> select(starts_with("b_"))

# Compute posterior summaries and exponentiate to get odds ratios
or_summary <- posterior_draws %>%
  summarise(across(starts_with("b_"),
                   list(mean = ~mean(.x),
                        sd = ~sd(.x),
                        lower95 = ~quantile(.x, 0.025),
                        upper95 = ~quantile(.x, 0.975)))) %>%
  tidyr::pivot_longer(cols = everything(),
                      names_to = c("parameter", ".value"),
                      names_pattern = "(b_.*)_(.*)") %>%
  mutate(odds_ratio = exp(mean),
         lower95_or = exp(lower95),
         upper95_or = exp(upper95))

or_table <- or_summary %>%
  select(parameter, odds_ratio, lower95_or, upper95_or) %>%
  mutate(parameter = gsub("b_", "", parameter)) %>%
  rename("Predictor" = parameter,
         "OR" = odds_ratio,
         "2.5% CI" = lower95_or,
         "97.5% CI" = upper95_or) %>%
  kbl(digits = 2, booktabs = TRUE, align = "lccc",
      caption = "Odds Ratios for Ordinal Model") %>%
  kableExtra::kable_styling(full_width = FALSE)

print(or_table)

```

```{r}

summary(ordinal_model_s1)

pp_check(ordinal_model_s1, type = "bars_grouped", group="refClass", fatten = 2) +
  scale_x_continuous("Response Category", breaks = 1:4, 
            labels = c("Exact", "0.01-2%", "2.01-15%", ">15%")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  ggtitle("Posterior Predictive Check - ordinal_model_s1") +
  theme_minimal() +
  theme(
    legend.background = element_blank(),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )

pp_check(ordinal_model_s1, type = "dens_overlay_grouped", group="refClass", fatten = 2,ndraws=100) +   scale_x_continuous("Response Category", breaks = 1:4, 
            labels = c("Exact", "0.01-2%", "2.01-15%", ">15%")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  ggtitle("Posterior Predictive Check - ordinal_model_s1") +
  theme_minimal() +
  theme(
    legend.background = element_blank(),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

```{r}

multinomial_model_s1 <- brm(
  accuracy_level ~ refClass + (1|id) + (1|state),
  data = s1_agg,
  family = categorical(), # Note: categorical() for multinomial
  cores = 4,
  iter = 1000,
  save_pars = save_pars(all = TRUE),
  control = list(adapt_delta = 0.99, max_treedepth=12),
#   prior = c(prior(normal(0, 3), class = "Intercept"), 
#             prior(normal(0, 3), class = "b")),
  file = paste0(here::here("data/model_cache",'s1_multinomial.rds')) 
)

summary(multinomial_model_s1)
pp_check(multinomial_model_s1, type = "bars_grouped", group="refClass", fatten = 2,ndraws=100) + labs(title = "Study 1: Posterior predictive check for the multinomial model") + theme_minimal()

```

```{r}
#| eval: false

loo_ordinal <- LOO(ordinal_model_s1, moment_match=TRUE)
loo_multinomial <- LOO(multinomial_model_s1, moment_match=TRUE, reloo=TRUE)
loo_compare(loo_ordinal, loo_multinomial)

loo_ordinal <- LOO(ordinal_model_s1)
loo_multinomial <- LOO(multinomial_model_s1)
loo_compare(loo_ordinal, loo_multinomial)

```

# S2

```{r}
#| label: tbl-s2-agg
#| tbl-cap: "Study 2: Summary of planning accuracy by reference class. The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."

s2_agg1 <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(matched_goal = (pct_change == pct),
                error = pct_change - pct,
                abs_error = abs(error)) |> 
      ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        )

s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),mean_abs_error=mean(abs_error),n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

# overall pct of subjects who matched their goal
s2_agg4 |> group_by(refClass) |>
    summarise(
    'Avg. % meeting goal' = mean(pct),
    'Avg. Deviation From Goal' = median(mean_pct_change),
    # sd = sd(pct),
    # n = n(),
    #se=sd(pct)/sqrt(n)
) |>   mutate(across(where(is.numeric), \(x) round(x, 3))) %>% 
  kable(escape=FALSE,booktabs=TRUE,align=c("l"))

```

```{r}

ordinal_model_s2_logit <- brm(
  accuracy_level ~ refClass +rounded+pct_goal + (1|id)+ (1|state),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 2), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_add.rds')) # Cache for efficiency
)
summary(ordinal_model_s2_logit)

pred_summary_s2 <- ordinal_model_s2_logit %>%
  epred_draws(newdata = s2_agg1, re_formula = NA,ndraws=200) %>%
  group_by("Reference Class"=refClass, rounded, "% Goal"=pct_goal, Category=.category) %>%
  summarise(
    mean_prob = mean(.epred),
    lower_ci = quantile(.epred, 0.025),
    upper_ci = quantile(.epred, 0.975),
    .groups = "drop"
  )

# Create a new data frame with the predictor values
new_data <- expand.grid(
  refClass = c("kWh", "Percentage", "USD"),
  id = unique(s2_agg1$id)[1],      # Take first id
  state = unique(s2_agg1$state)[1], # Take first state,
  rounded = unique(s2_agg1$rounded)[1], # Take first rounded
  pct_goal = unique(s2_agg1$pct_goal)[1] # Take first pct_goal
)

# Generate predicted probabilities
predicted_probabilities <- predict(ordinal_model_s2_logit, newdata = new_data, 
                                  summary = FALSE, ndraws = 1000)

# Calculate the predicted probability of "Exact match"
predicted_probability_exact_match <- mean(predicted_probabilities <= 1)

print(predicted_probability_exact_match)

pred_summary_s2 |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy level by reference class",split.table=Inf)

# Extract posterior samples
posterior_samples_s2 <- as.data.frame(ordinal_model_s2_logit)

odds_ratios_s2 <- posterior_samples_s2 %>%
    transmute(
        refClass_Percentage_vs_kWh = exp(b_refClassPercentage),
        refClass_USD_vs_kWh = exp(b_refClassUSD),
        rounded_Yes_vs_No = exp(b_roundedRounded),
        pct_goal_15_vs_10 = exp(`b_pct_goal15%`)  # Note the backticks here
    )

# Calculate summary statistics
odds_ratio_summary_s2 <- data.frame(
    comparison = c("Percentage vs kWh", "USD vs kWh", "Rounded vs Not", "15% Goal vs 10% Goal"),
    odds_ratio = c(
        mean(odds_ratios_s2$refClass_Percentage_vs_kWh),
        mean(odds_ratios_s2$refClass_USD_vs_kWh),
        mean(odds_ratios_s2$rounded_Yes_vs_No),
        mean(odds_ratios_s2$pct_goal_15_vs_10)
    ),
    ci_lower = c(
        quantile(odds_ratios_s2$refClass_Percentage_vs_kWh, 0.025),
        quantile(odds_ratios_s2$refClass_USD_vs_kWh, 0.025),
        quantile(odds_ratios_s2$rounded_Yes_vs_No, 0.025),
        quantile(odds_ratios_s2$pct_goal_15_vs_10, 0.025)
    ),
    ci_upper = c(
        quantile(odds_ratios_s2$refClass_Percentage_vs_kWh, 0.975),
        quantile(odds_ratios_s2$refClass_USD_vs_kWh, 0.975),
        quantile(odds_ratios_s2$rounded_Yes_vs_No, 0.975),
        quantile(odds_ratios_s2$pct_goal_15_vs_10, 0.975)
    )
)

#             comparison odds_ratio ci_lower ci_upper
# 1    Percentage vs kWh       1.99     0.73     4.43
# 2           USD vs kWh       3.69     1.33     8.29
# 3       Rounded vs Not       0.70     0.57     0.84
# 4 15% Goal vs 10% Goal       0.78     0.63     0.94

odds_ratio_summary_s2 |> kable()

odds_ratio_summary_s2 |> pander::pandoc.table(caption="Study 2: Odds ratios of accuracy")

```

### 2024-12-07 docs/misc/ordinal.qmd
update results section. And some formating improvements.

--- 2024-11-20 misc/ordinal.qmd
@@
     theme_minimal()
+
+marginal_effects(ordinal_model_s2_probit)
+
+```
+
+## S2 logit version
+```{r}
+
+ordinal_model_s2_logit <- brm(
+  accuracy_level ~ refClass +rounded+pct_goal + (1|id)+ (1|state),
+  data = s2_agg1,
+  family = cumulative("logit"),
+  cores = 4,
+  iter = 2000,
+  control = list(adapt_delta = 0.99), # Recommended for ordinal models
+  prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
+            prior(normal(0, 2), class = "b")), # Priors for predictors
+  file = paste0(here::here("data/model_cache",'s2_logit_add.rds')) # Cache for efficiency
+)
+summary(ordinal_model_s2_logit)
+# Regression Coefficients:
+#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# Intercept[1]          -2.13      0.64    -3.39    -0.86 1.00      639     1328
+# Intercept[2]          -0.62      0.63    -1.89     0.63 1.00      658     1366
+# Intercept[3]           3.15      0.65     1.88     4.42 1.00      717     1638
+# refClassPercentage     0.83      0.75    -0.64     2.33 1.00      452     1253
+# refClassUSD            1.87      0.79     0.31     3.35 1.00      508     1260
+# roundedRounded        -0.66      0.18    -1.01    -0.31 1.00     4276     3095
+# pct_goal15%           -0.44      0.18    -0.79    -0.10 1.00     6263     3109
+
+pred_summary_s2 <- ordinal_model_s2_logit %>%
+  epred_draws(newdata = s2_agg1, re_formula = NA) %>%
+  group_by(refClass, rounded, pct_goal, .category) %>%
+  summarise(
+    mean_prob = mean(.epred),
+    lower_ci = quantile(.epred, 0.025),
+    upper_ci = quantile(.epred, 0.975),
+    .groups = "drop"
+  )
+
+pred_summary_s2 |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy level by reference class")
+
+```
+
+### logit vs. probit
+
+```{r}
+
+loo_logit <- loo(ordinal_model_s2_logit)
+loo_probit <- loo(ordinal_model_s2_probit)
+loo_compare(loo_logit, loo_probit)
+
+waic(ordinal_model_s2_logit)
+waic(ordinal_model_s2_probit)
+
+# compare
+rstanarm::loo_compare(loo_logit, loo_probit)
+
+pp_check(ordinal_model_s2_logit, type = "bars_grouped", group="refClass", fatten = 2) +
+  scale_x_continuous("Response Category", breaks = 1:4, 
+            labels = c("Exact", "0.01-2%", "2.01-15%", ">15%")) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
+  ggtitle("Posterior Predictive Check by Reference Class - logit") +
+  theme_minimal() +
+  theme(
+    legend.background = element_blank(),
+    legend.position = "bottom",
+    panel.grid.minor = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 1))
+
+pp_check(ordinal_model_s2_probit, type = "bars_grouped", group="refClass", fatten = 2) +
+  scale_x_continuous("Response Category", breaks = 1:4, 
+            labels = c("Exact", "0.01-2%", "2.01-15%", ">15%")) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
+  ggtitle("Posterior Predictive Check by Reference Class - probit") +
+  theme_minimal() +
+  theme(
+    legend.background = element_blank(),
+    legend.position = "bottom",
+    panel.grid.minor = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 1))
 

### 2024-12-07 manuscript/manuscript.qmd
update results section. And some formating improvements.

@@
   - /manuscript.html
-author:
-  - name: Thomas E. Gorman
-    url: https://tegorman13.github.io/
-    orcid: "0000-0001-5366-5442"
-    affiliations: 
-      - ref: cclab
-      - ref: clara
-    attributes: 
-      corresponding: true
-  - name: Torsten Reimer
-    url: https://web.ics.purdue.edu/~treimer/
-    orcid: 0000-0002-7419-0076
-    affiliations:
-      - ref: cclab
-      - ref: clara
-  - name: Juan Pablo Loaiza Ramirez
-    url: https://www.cla.purdue.edu/directory/profiles/loaiza-ramirez,juan-pablo.html
-    orcid: 0000-0001-9663-0522
-    affiliations:
-      - ref: cclab
-      - ref: clara
-  - name: Hayden Barber
-    url: https://www.sdstate.edu/directory/hayden-barber
-    orcid: 0000-0002-3465-8615
-    affiliations:
-      - ref: sdsu
-affiliations:
-  - id: cclab
-    name: Communication and Cognition Lab, Purdue University
-    city: West Lafayette
-    region: IN
-    country: USA
-    postal-code: 47907
-    url: https://web.ics.purdue.edu/~treimer/
-  - id: clara
-    name: College of Liberal Arts Research Academy
-    department: Purdue University
-    city: West Lafayette
-    region: IN
-    country: USA
-    postal-code: 47907
-    url: https://cla.purdue.edu/about/college-initiatives/research-academy/
-  - id: sdsu
-    name: School of Communication & Journalism, South Dakota State University
-    city: Brookings
-    region: SD
-    country: USA
-    url: https://www.sdstate.edu/directory/hayden-barber
 format:
-    # html: 
-    #     grid:
-    #         sidebar-width: 220px
-    #         body-width: 1200px
-    #         margin-width: 170px
-    #         gutter-width: 1.0rem
     wordcount-html: 
@@
         shift-heading-level-by: 1
-        code-fold: true
-        code-summary: "Show the code"
         title-block-banner: "#000D51"
@@
                 </style>
+    html: 
+      echo: true
+      code-fold: show
+      output-file: manuscript-code.html
     hugo-md:
-        include: true
+        include: false
         echo: false
@@
     hikmah-manuscript-docx: default
-# format-links:   
-#     - html
-#     - gfm
-#     - format: hikmah-manuscript-pdf
-#       text: PDF version
-#       icon: file-pdf
-date: 2024-12-03
+format-links:   
+    - text: HTML
+      href: manuscript.html
+      icon: file-minus
+    - text: HTML (with code)
+      href: manuscript-code.html
+      icon: file-plus
+    - format: hikmah-manuscript-pdf
+      text: PDF version
+      icon: file-pdf
+    - hikmah-manuscript-docx
+    - gfm
+date: 2024-12-07
 abstract: |
@@
 
-- Electricity bills are a primary source of energy-use information for consumers and offer a promising avenue for enhancing communication about energy consumption (Fischer, 2008)
+- Electricity bills are a primary source of energy-use information for consumers and offer a promising avenue for enhancing communication about energy consumption [@fischerFeedbackHouseholdElectricity2008]
 
@@
 
-- Canfield et al.
-(2017) found that tables were more effective than graphs for conveying specific electricity usage data, likely because tables facilitate straightforward point reading.
+- @canfieldPerceptionsElectricityuseCommunications2017 found that tables were more effective than graphs for conveying specific electricity usage data, likely because tables facilitate straightforward point reading.
 However, they also noted that individuals with lower energy literacy had reduced comprehension across all formats.
-- Canfield et al.'s (2017) findings on preferences for historical use information and the impact of neighbor comparisons
-- The concept of cognitive fit posits that performance improves when the information presentation format aligns with the task requirements (Vessey, 1991)
-- alignment can reduce cognitive load and enhance accuracy in planning (Shah & Freedman, 2009)
+- @canfieldPerceptionsElectricityuseCommunications2017's findings on preferences for historical use information and the impact of neighbor comparisons
+- The concept of cognitive fit posits that performance improves when the information presentation format aligns with the task requirements [@vesseyCognitiveFitTheoryBased1991]
+- alignment can reduce cognitive load and enhance accuracy in planning [@shahBarLineGraph2011]
 - Reimer et al. (2015) provide  context on how numerical formats affect risk perception, the reference class problem, and the benefits of natural frequencies.
 
-Energy poverty continues to be a pervasive issue in the United States @memmottSociodemographicDisparitiesEnergy2021 .
+Energy poverty continues to be a pervasive issue in the United States @memmottSociodemographicDisparitiesEnergy2021.
 This challenge partly arises from difficulties in converting information across numerical formats, impeding the development of precise energy reduction plans @reimerNumericCommunicationRisk2015.
@@
 
-The way numerical information is presented can significantly affect how individuals process and use that information (Reimer et al., 2015).
-The reference class problem highlights that numbers without clear reference points can lead to misinterpretation, as the meaning of a statistic depends on the category or class it refers to (Gigerenzer & Edwards, 2003; Reimer et al., 2015).
+The way numerical information is presented can significantly affect how individuals process and use that information [@reimerNumericCommunicationRisk2015] .
+The reference class problem highlights that numbers without clear reference points can lead to misinterpretation, as the meaning of a statistic depends on the category or class it refers to [@gigerenzerSimpleToolsUnderstanding2003; @reimerNumericCommunicationRisk2015].
 Presenting energy information in absolute units (e.g., kWh) provides a clear reference class, potentially enhancing comprehension.
 
-Furthermore, research suggests that natural frequencies and absolute numbers are generally easier for individuals to understand compared to percentages or probabilities (Gigerenzer & Hoffrage, 1995; Hoffrage et al., 2000).
+Furthermore, research suggests that natural frequencies and absolute numbers are generally easier for individuals to understand compared to percentages or probabilities [@gigerenzerHowImproveBayesian1995 @hoffrageCommunicatingStatisticalInformation2000].
 In the context of energy conservation, using absolute units may facilitate more accurate planning and decision-making by aligning with intuitive cognitive processing.
@@
 ```{r}
+pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,ggplot2,gt,knitr,kableExtra,ggh4x,patchwork,lme4,flextable,pander)
+options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
 
-pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,ggplot2,gt,knitr,kableExtra,ggh4x,patchwork,lme4,flextable,pander)
-
-options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
 walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
@@
 #| label: tbl-s1-agg
-#| caption: "Study 1: Summary of planning accuracy by reference class.
+#| tbl-cap: "Study 1: Summary of planning accuracy by reference class.
 The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."
@@
 For our primary analyses of participants' ability to create accurate energy-saving plans, we employed an accuracy level binning approach by categorizing responses into four distinct levels: Exact match, 0.01–2% error, 2.01–15% error, and Over 15% error.
+The current analysis employs a cumulative ordinal regression model, implemented via a Bayesian hierarchical framework [@burknerBrmsPackageBayesian2017] .
+This approach allows the estimation of threshold parameters and regression coefficients that characterize how changes in predictor variables (such as the reference class: kWh, percentage, or USD) relate to probabilities of being in each accuracy category.
+All analyses were carried out in R [@rcoreteamLanguageEnvironmentStatistical2020] and the tidyverse package [@wickhamWelcomeTidyverse2019].
 
@@
 #| label: tbl-s1-prop
-#| caption: "Study 1: The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages).
+#| tbl-cap: "Study 1: The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages).
 The combined group column reflects the percentage of participants in each accuracy level when aggregating across across all reference class conditions."
@@
 
+see @fig-s1-plot
+
 ```{r}
 #| label: fig-s1-plot
-#| caption: "Study 1: Proportion of participants in each accuracy level, colored by reference class. A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
+#| fig-cap: "Study 1: Proportion of participants in each accuracy level, colored by reference class. A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
 
@@
 #| label: tbl-s1-ord
-#| caption: "Study 1: Odds ratios of for group comparisons. "
+##| tbl-cap: "Study 1: Ordinal Regression Model Results."
 
@@
     cores = 4,
-    iter = 2000,
+    iter = 5000,
+    save_pars = save_pars(all = TRUE),
     control = list(adapt_delta = 0.99), 
-    prior = c(prior(normal(0, 2), class = "Intercept"), 
-                prior(normal(0, 1.5), class = "b")), 
+    prior = c(prior(normal(0, 3), class = "Intercept"), 
+                prior(normal(0, 3), class = "b")), 
     file = paste0(here::here("data/model_cache",'s1_ordinal3.rds')) 
@@
 
-#summary(ordinal_model_s1)
+#colnames(mted1) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")
 
-# describe_posterior(ordinal_model_s1, centrality = "Mode") |> 
-#     filter(stringr::str_detect(Parameter, "b_")) |> 
-#     mutate(Parameter = stringr::str_remove(Parameter, "b_")) |> 
-#     mutate(across(where(is.numeric), round, 3)) |> 
-#     kable(booktabs = TRUE)
+# as.data.frame(describe_posterior(ordinal_model_s1, centrality = "Mean"))[, c(1,2,4,5,6)] |> 
+#   setNames(c("Parameter", "Estimate", "CI_Lower", "CI_Upper", "pd")) |> 
+#   mutate(Parameter = stringr::str_remove(Parameter, "b_")) |> kable(escape=FALSE,booktabs=TRUE,align=c("l"), row.names = FALSE)
 
@@
                 ndraws = 1000, re_formula = NA) |>
-    group_by(refClass, .category) |>
+    group_by(refClass, Category=.category) |>
     summarise(
@@
 
-# Convert log-odds to odds ratios
-posterior_samples <- as.data.frame(ordinal_model_s1)
-odds_ratios <- data.frame(
-  Percentage_vs_kWh = exp(posterior_samples$b_refClassPercentage),
-  USD_vs_kWh = exp(posterior_samples$b_refClassUSD)
-)
+#plot(conditional_effects(ordinal_model_s1),points=TRUE)
 
-# Calculate summary statistics
-odds_ratio_summary <- data.frame(
-    comparison = c("Percentage vs kWh", "USD vs kWh"),
-    odds_ratio = c(mean(odds_ratios$Percentage_vs_kWh),
-                    mean(odds_ratios$USD_vs_kWh)),
-    ci_lower = c(quantile(odds_ratios$Percentage_vs_kWh, 0.025),
-                quantile(odds_ratios$USD_vs_kWh, 0.025)),
-    ci_upper = c(quantile(odds_ratios$Percentage_vs_kWh, 0.975),
-                quantile(odds_ratios$USD_vs_kWh, 0.975))
-)
+# or_table |>
+#       rownames_to_column(var = "Term") |>
+#       filter(!stringr::str_detect(Term, "Intercept")) |> kable(escape=FALSE,booktabs=TRUE,align=c("l"))
 
-odds_ratio_summary |> kable(escape=FALSE,booktabs=TRUE,align=c("l")) 
-# |comparison        | odds_ratio| ci_lower| ci_upper|
-# |:-----------------|----------:|--------:|--------:|
-# |Percentage vs kWh |        2.3|     0.62|      6.2|
-# |USD vs kWh        |       14.1|     4.08|     38.0|
-
-#odds_ratio_summary |> pander::pandoc.table(caption="Study 1: Odds ratios of accuracy")
-
+# or_table <- data.frame(exp(fixef(ordinal_model_s1)))
+# or_table |>
+#   rownames_to_column(var = "Term") |>
+#   filter(!stringr::str_detect(Term, "Intercept")) |>
+#   as_flextable() |>
+#   set_header_labels(values = list(
+#     OR = "Odds Ratio",
+#     `2.5 %` = "Lower 95% CI",
+#     `97.5 %` = "Upper 95% CI"
+#   )) |>
+#   flextable::add_footer_lines(values = "Note: Odds ratios derived from the ordinal logistic regression model.") |> 
+#   align(align = 'center', part = 'all') |>
+#   fontsize(size = 10, part = "all")
 ```
 
+::: {#tbl-s1-or}
+|Parameter          |Estimate |CI_Lower |CI_Upper |pd   |
+|:------------------|:--------|:--------|:--------|:----|
+|Intercept[1]       |-1.94    |-3.19    |-0.74    |1.00 |
+|Intercept[2]       |0.20     |-1.02    |1.42     |0.63 |
+|Intercept[3]       |4.36     |3.11     |5.65     |1.00 |
+|refClassPercentage |0.85     |-0.54    |2.23     |0.88 |
+|refClassUSD        |2.72     |1.41     |4.06     |1.00 |
+
+**Experiment 1**: Ordinal Regression results.
+
+:::
+
+::: {#tbl-s1-epred}
+|Term               |Estimate |Est.Error |Q2.5 |Q97.5 |
+|:------------------|:--------|:---------|:----|:-----|
+|refClassPercentage |2.3      |2         |0.58 |9.3   |
+|refClassUSD        |15.2     |2         |4.11 |58.1  |
+
+**Experiment 1**: Odds ratios for group comparisons.
+
+:::
+
+The ordinal model is parameterized with thresholds (intercepts), and positive coefficients can indicate that it is more difficult to achieve higher accuracy categories in the USD condition. 
+The model output suggests that, compared to the kWh condition, the USD condition shows a positive coefficient (Estimate = 2.72, 95% CI: 1.41 to 4.06) for the ordinal outcome.
+At least to me, this positive coefficient appears to indicate that, relative to the kWh reference class, participants in the USD condition are more likely to fall into higher numerical categories of the dependent variable coding.
+However, because the dependent variable is ordered from best (Exact match) to worst (Over 15% error), care is needed in interpretation.
+The Percentage condition coefficient (Estimate = 0.85, 95% CI: -0.54 to 2.23) is more uncertain, with its credible interval overlapping zero.
+Posterior predictive checks (Figure 3) showed that the ordinal model provided a reasonable fit to the observed data (see @fig-s1-plot2).
+
 ```{r}
+#| label: fig-s1-plot2
+#| fig-cap: "Study 1: Proportion of participants in each accuracy level, colored by reference class, and seprated in facets based on the levels of reduction goal.
+A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
 #| fig.width: 11
@@
 
-bayesplot::color_scheme_set(wes_palettes[1]$BottleRocket1[1:6])
+#bayesplot::color_scheme_set(wes_palettes[1]$BottleRocket1[1:6])
 
 # Plot predicted probabilities
-ggplot(pred_summary, aes(x = refClass, y = mean_prob, fill = .category)) +
-    geom_bar(stat = "identity", position = position_dodge()) +
-    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci),
-                    position = position_dodge(width = 0.9), width = 0.2) +
-    labs(y = "Predicted Probability", x = "Reference Class", fill = "Accuracy Level") +
-    ggtitle("Study 1: Predicted Probabilities of Accuracy Level by Reference Class") +
-    theme_minimal()
+# ggplot(pred_summary, aes(x = refClass, y = mean_prob, fill = Category)) +
+#     geom_bar(stat = "identity", position = position_dodge()) +
+#     geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci),
+#                     position = position_dodge(width = 0.9), width = 0.2) +
+#     labs(y = "Predicted Probability", x = "Reference Class", fill = "Accuracy Level") +
+#     ggtitle("Study 1: Predicted Probabilities of Accuracy Level by Reference Class") +
+#     theme_minimal()
 
@@
 pp_check(ordinal_model_s1, type = "bars_grouped", group="refClass", fatten = 2) +
-    # scale_x_continuous("y", breaks = 1:7) +
-    # scale_y_continuous(NULL, breaks = NULL, expand = expansion(mult = c(0, 0.05))) +
-    ggtitle("Data with posterior predictions",
-            subtitle = "S1. red bars are empirical data, black points are model predictions and credible intervals") +
-    labs(x = "Accuracy Level: 1= exact match,   +1= 0.01-2% error, +2= 2.01-15% error, +3= Over 15% error") +
+  scale_x_continuous("Response Category", breaks = 1:4, 
+            labels = c("Exact", "0.01-2%", "2.01-15%", ">15%")) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
+  ggtitle("Posterior Predictive Check by Reference Class") +
   theme_minimal() +
-    theme(legend.background = element_blank() #,legend.position = c(.9, .75)
+  theme(
+    legend.background = element_blank(),
+    legend.position = "bottom",
+    panel.grid.minor = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 1)
   )
-
 ```
@@
 #| label: tbl-s2-agg
-#| caption: "Study 2: Summary of planning accuracy by reference class.
+#| tbl-cap: "Study 2: Summary of planning accuracy by reference class.
 The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."
@@
                 error = pct_change - pct,
-                abs_error = abs(error)) |> 
+                abs_error = abs(error),
+                log_abs_error=log(abs(error)+.001)) |> 
       ungroup() |> # Add ungroup here
@@
     mutate(n_accuracy = n_distinct(accuracy_level)) |> 
-    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),mean_abs_error=mean(abs_error),n_accuracy=first(n_accuracy)) |> 
+    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),
+    mean_abs_error=mean(abs_error),mean_log_abs_error=mean(log_abs_error),
+    n_accuracy=first(n_accuracy)) |> 
     mutate(accuracy_level = factor(
@@
     'Avg. % meeting goal' = mean(pct),
-    'Avg. Deviation From Goal' = median(mean_pct_change),
+    'Avg. Abs. Deviation From Goal' = median(mean_pct_change),
+    'Log Deviation' = (median(mean_log_abs_error)),
     # sd = sd(pct),
@@
 
+# s2_agg1 |> group_by(id,refClass,calc) |> 
+#     mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+#     summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),
+#     mean_abs_error=mean(abs_error),mean_log_abs_error=mean(log_abs_error),
+#     n_accuracy=first(n_accuracy))
+
+# s2_agg4 %>% ggplot(aes(x=refClass,y=mean_log_abs_error,fill=refClass)) + 
+#   stat_summary(fun=mean, geom="bar") +
+#   stat_summary(fun.data=mean_cl_normal, geom="errorbar", width=0.2) +
+#   geom_jitter(alpha = 0.2, width = 0.2, height = 0)
+
+# s2_agg4 |> 
+#   ggplot(aes(x = mean_log_abs_error,fill = refClass)) +
+#   geom_density(alpha = 0.7)
+
+# library(ggridges)
+# s2_agg1 |>  ggplot(aes(y = refClass, x=log_abs_error,fill=refClass,col=refClass)) +
+#     geom_density_ridges(alpha = 0.5) +
+#     geom_vline(aes(xintercept=(log_abs_error)))
+
 ```
@@
 #| label: tbl-s2-prop
-#| caption: "Study 2: The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages).
+#| tbl-cap: "Study 2: The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages).
 The combined group column reflects the percentage of participants in each accuracy level when aggregating across across all reference class conditions."
@@
 #| label: fig-s2-plot
-#| caption: "Study 2: Proportion of participants in each accuracy level, colored by reference class. A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
+#| tbl-cap: "Study 2: Proportion of participants in each accuracy level, colored by reference class. A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
 
@@
 #| label: fig-s2-plot2
-#| caption: "Study 2: Proportion of participants in each accuracy level, colored by reference class, and seprated in facets based on the levels of reduction goal, and rounding.
+#| fig-cap: "Study 2: Proportion of participants in each accuracy level, colored by reference class, and seprated in facets based on the levels of reduction goal, and rounding.
 A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
@@
 
+```{r}
+#| label: tbl-s2-ord
+##| tbl-cap: "Study 2: Ordinal Regression Model Results."
+
+ordinal_model_s2_logit <- brm(
+  accuracy_level ~ refClass +rounded+pct_goal + (1|id)+ (1|state),
+  data = s2_agg1,
+  family = cumulative("logit"),
+  cores = 4,
+  iter = 2000,
+  control = list(adapt_delta = 0.99), # Recommended for ordinal models
+  prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
+            prior(normal(0, 2), class = "b")), # Priors for predictors
+  file = paste0(here::here("data/model_cache",'s2_logit_add.rds')) # Cache for efficiency
+)
+#summary(ordinal_model_s2_logit)
+
+# as.data.frame(describe_posterior(ordinal_model_s2_logit, centrality = "Mean"))[, c(1,2,4,5,6)] |> 
+#   setNames(c("Parameter", "Estimate", "CI_Lower", "CI_Upper", "pd")) |> 
+#   mutate(Parameter = stringr::str_remove(Parameter, "b_")) |> 
+#   kable(escape = FALSE, booktabs = TRUE, align = c("l"), row.names = FALSE)
+
+pred_summary_s2 <- ordinal_model_s2_logit %>%
+  epred_draws(newdata = s2_agg1, re_formula = NA,ndraws=200) %>%
+ # group_by("Reference Class"=refClass, rounded, "% Goal"=pct_goal, Category=.category) %>%
+  group_by("Reference Class"=refClass, Category=.category) %>%
+  summarise(
+    mean_prob = mean(.epred),
+    lower_ci = quantile(.epred, 0.025),
+    upper_ci = quantile(.epred, 0.975),
+    .groups = "drop"
+  )
+
+#pred_summary_s2 |> kable(escape=FALSE,booktabs=TRUE,align=c("l"), row.names = FALSE)
+
+# Create a new data frame with the predictor values
+# new_data <- expand.grid(
+#   refClass = c("kWh", "Percentage", "USD"),
+#   id = unique(s2_agg1$id)[1],      # Take first id
+#   state = unique(s2_agg1$state)[1], # Take first state,
+#   rounded = unique(s2_agg1$rounded)[1], # Take first rounded
+#   pct_goal = unique(s2_agg1$pct_goal)[1] # Take first pct_goal
+# )
+
+# or_table <- data.frame(exp(fixef(ordinal_model_s2_logit)))
+# or_table |>
+#   rownames_to_column(var = "Term") |>
+#   filter(!stringr::str_detect(Term, "Intercept")) |>
+#   as_flextable() |>
+#   set_header_labels(values = list(
+#     OR = "Odds Ratio",
+#     `2.5 %` = "Lower 95% CI",
+#     `97.5 %` = "Upper 95% CI"
+#   )) |>
+#   flextable::add_footer_lines(values = "Note: Odds ratios derived from the ordinal logistic regression model.") |> 
+#   align(align = 'center', part = 'all') |>
+#   fontsize(size = 10, part = "all")
+
+```
+
+::: {#tbl-s2-or}
+|Parameter          |Estimate |CI_Lower |CI_Upper |pd   |
+|:------------------|:--------|:--------|:--------|:----|
+|Intercept[1]       |-2.13    |-3.39    |-0.86    |1.00 |
+|Intercept[2]       |-0.62    |-1.89    |0.63     |0.84 |
+|Intercept[3]       |3.15     |1.88     |4.42     |1.00 |
+|refClassPercentage |0.83     |-0.64    |2.33     |0.87 |
+|refClassUSD        |1.87     |0.31     |3.35     |0.99 |
+|roundedRounded     |-0.66    |-1.01    |-0.31    |1.00 |
+|pct_goal15%        |-0.44    |-0.79    |-0.10    |0.99 |
+
+**Experiment 2.** Ordinal Regression Model Results.
+
+:::
+
+::: {#tbl-s2-epred}
+|comparison           | odds_ratio| ci_lower| ci_upper|
+|:--------------------|----------:|--------:|--------:|
+|Percentage vs kWh    |       3.02|     0.53|    10.31|
+|USD vs kWh           |       8.80|     1.37|    28.42|
+|Rounded vs Not       |       0.53|     0.36|     0.73|
+|15% Goal vs 10% Goal |       0.66|     0.45|     0.91|
+
+**Experiment 2.** Odds ratios for group comparisons.
+
+:::
+
+```{r}
+#| label: fig-s2-plot3
+
+# ggplot(pred_summary_s2, 
+#         aes(x = .category, y = mean_prob, fill = refClass )) +
+#     geom_col(position = position_dodge(width = 0.8)) +
+#     geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), 
+#                     position = position_dodge(width = 0.8)) +
+#     labs(x = "Reference Class", 
+#         y = "Predicted Probability",
+#         fill = "Accuracy Level",
+#         title = "Predicted Probabilities of Accuracy Levels by Reference Class") +
+#     scale_fill_brewer(palette = "RdYlBu") +
+#     theme_minimal()
+
+pp_check(ordinal_model_s2_logit, type = "bars_grouped", group="refClass", fatten = 2) +
+  scale_x_continuous("Response Category", breaks = 1:4, 
+            labels = c("Exact", "0.01-2%", "2.01-15%", ">15%")) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
+  ggtitle("Posterior Predictive Check by Reference Class") +
+  theme_minimal() +
+  theme(
+    legend.background = element_blank(),
+    legend.position = "bottom",
+    panel.grid.minor = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 1))
+
+```
+
 ### Individual Differences
+
+see @fig-s2-indv
 
@@
 #| label: fig-s2-indv
-#| caption: "Study 2: Individual performance in the energy planning task, colored by reference class.
+#| fig-cap: "Study 2: Individual performance in the energy planning task, colored by reference class.
 The dashed line represents the target reduction goal.
@@
 #| label: fig-s2-indv2
-#| caption: "Study 2: Respones patterns for a subset of individiual participants.
+#| fig-cap: "Study 2: Respones patterns for a subset of individiual participants.
 Black points are participant responses, red points are the state average, and blue points are the family average.
@@
 
-#| 
 s2_long |> filter(id %in% unique(s2_long$id)[1:30]) |> 
@@
 
+Karjalainen 2011 - people prefer information about price [@karjalainenConsumerPreferencesFeedback2011]
+
 # References
+::: {#refs}
+:::

### 2024-12-07 manuscript/test.qmd
update results section. And some formating improvements.

--- 2024-12-04 manuscript/manuscript.qmd
@@
 title: "Planning to Save Energy: How Information Format Affects Accuracy"
-aliases:
-  - /manuscript.html
 author:
@@
         html-math-method: mathjax
-        output-file: manuscript_hugo.md
+        output-file: test_hugo.md
     gfm:
         echo: true
-        output-file: manuscript_gfm.md
+        output-file: test_gfm.md
     hikmah-manuscript-pdf:
-        output-file: rep_study.pdf
+        output-file: test.pdf
         mainfont: "Linux Libertine O"
@@
 
-For each family scenario, the particpipants were shown a table containing the families utility usage from the prior year, alongside the state averages for each appliance category (both prior year usage and stage averages are always shown in kWh).
-For each scenario, participants were asked to create two possible action plans to achieve the target reduction in total household energy usage (see @fig-task).
-Depending on their reference class condition, the target reduction amount presented either in kilowatt-hours (kWh), as percentages of total household usage, or in U.S.
-dollars.
-In all conditions, the target reduction was equivalent to a 15% reduction in total household kWh.
-
-![Example trial in the energy planning task.
-Participants are shown the prior year electricity use of a household, and are tasked with creating a plan for the next year that will meet the energy reduction goal.
-Study 1 manipulates the format of the reduction goal to be either a percentage (15% given as goal reduction), kilowatt hours (5965 kWh given), or USD (\$656)](/assets/images/Smith_10R_Wells_15E.png){#fig-task}
-
 Additional data collected included:
@@
 -   **Environmental Attitudes Survey**: Assessed participants' pro-environmental attitudes and perceived importance of energy conservation.
-
-## Results
-
-```{r}
-
-pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,ggplot2,gt,knitr,kableExtra,ggh4x,patchwork,lme4,flextable,pander)
-
-options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
-walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
-theme_set(theme_nice())
-
-s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
-  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
-  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))
-
-s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
-  filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
-  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))
-
-```
-
-```{r}
-#| label: tbl-s1-agg
-#| caption: "Study 1: Summary of planning accuracy by reference class.
-The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."
-
-s1_agg <- s1 |> 
-    filter(appliance !="Total kWh") |> 
-    group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
-    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
-                pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
-                n_change = sum(value!=family),
-                state_p_dif=mean(state_p_dif),
-                state_f_dif=mean(state_f_dif),
-                n_less_avg = sum(less_avg),
-                duration=first(Duration__in_seconds_)) |> 
-    mutate(matched_goal = (pct_change == pct_goal), 
-                error = pct_change - pct_goal,
-                abs_error = abs(error),
-                close_match = abs_error <= 0.03) |>
-    ungroup() |> # Add ungroup here
-        mutate(
-            accuracy_level = factor(
-                case_when(
-                    abs_error == 0.00 ~ "Exact match",
-                    abs_error <= 0.02 ~ "0.01-2% error",
-                    abs_error <= 0.15 ~ "2.01-15% error",
-                    TRUE ~ "Over 15% error"  # Capture all remaining cases
-                ), 
-                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
-                ordered = TRUE
-            )
-        ) |> relocate(accuracy_level, .after= "pct_change")
-
-s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
-    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
-    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),mean_abs_error=mean(abs_error),n_accuracy=first(n_accuracy)) |> 
-    mutate(accuracy_level = factor(
-            case_when(
-                mean_abs_error < 0.02 ~ "Exact match",
-                mean_abs_error <= 0.02 ~ "0.01-2% error",
-                mean_abs_error <= 0.15 ~ "2.01-15% error",
-                TRUE ~ "Over 15% error"  # Capture all remaining cases
-            ), 
-            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
-            ordered = TRUE
-        ))
-
-# proportion matching vs close match
-
-# proportion matching vs close match
-
-# overall pct of subjects who matched their goal
-s1_agg4 |> group_by(refClass) |>
-    summarise(
-    'Avg. % meeting goal' = mean(pct),
-    'Avg. Deviation From Goal' = median(mean_pct_change),
-    # sd = sd(pct),
-    # n = n(),
-    #se=sd(pct)/sqrt(n)
-) |>   mutate(across(where(is.numeric), \(x) round(x, 3))) %>% 
-  kable(escape=FALSE,booktabs=TRUE,align=c("l"))
-
- #pander::pandoc.table(caption="Study 1: Proportion of participants who matched their goal overall")
-
-```
-
-For our primary analyses of participants' ability to create accurate energy-saving plans, we employed an accuracy level binning approach by categorizing responses into four distinct levels: Exact match, 0.01–2% error, 2.01–15% error, and Over 15% error.
-
-```{r}
-#| label: tbl-s1-prop
-#| caption: "Study 1: The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages).
-The combined group column reflects the percentage of participants in each accuracy level when aggregating across across all reference class conditions."
-
-# compute percentage of subjects per accuracy level per group
-observed_props_s1 <- s1_agg |>
-  group_by(refClass, accuracy_level) |>
-  summarise(n = n()) |>
-  group_by(refClass) |>
-  mutate(prop = n/sum(n)) |>
-  mutate(n_prop=paste0(n," (",round(prop*100,1),"%)" ), pct_grp=paste0(round(prop*100,1), "%")) |> ungroup()
-
-observed_props_s1 |> 
-  mutate(n_total=sum(n)/4) |> 
-  group_by(accuracy_level) |>
-  mutate(ns=sum(n)/4) |> 
-  mutate(Total = paste0(round(ns/n_total*100,1), "%")) |>
-  select('Reference Class'=refClass, 'Accuracy Level'=accuracy_level, '% in Group'=pct_grp, "Combined Groups %" =Total) |>
-  pivot_wider(
-    names_from = 'Reference Class',
-    values_from = c('% in Group')
-  ) |> relocate("Combined Groups %" , .after=last_col()) |> 
-  kable(escape=FALSE,booktabs=TRUE,align=c("l"))
-
-```
-
-```{r}
-#| label: fig-s1-plot
-#| caption: "Study 1: Proportion of participants in each accuracy level, colored by reference class. A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
-
-prop_acc_s1 <- s1_agg %>%
-    group_by(refClass, accuracy_level) %>%
-    summarise(count = n()) %>%
-    group_by(accuracy_level) |>
-    mutate(Probability = count / sum(count)) %>%
-    ungroup()
-
-ggplot(prop_acc_s1, aes(x = accuracy_level, y = Probability, fill = refClass)) +
-    geom_bar(stat = "identity", position = position_dodge()) +
-  scale_y_continuous(labels = scales::percent) +
-    labs(title = "S1. % of Participants within each Accuracy Bin",
-        x = "Accuracy Level",
-        y = "Percentage of Participants",
-        fill = "Reference Class") +
-    theme_minimal() 
-```
-
-We analyzed planning accuracy using Bayesian ordinal regression.
-The dependent variable, plan error, was computed by binning the goal deviation into four ordered levels: exact match (0% error), minor deviations (0.01-2% error), moderate deviations (2.01-15% error), and major deviations (>15% error).
-For each comparison, we provide posterior odds ratios (OR) and their 95% CIs.
-
-```{r}
-#| label: tbl-s1-ord
-#| caption: "Study 1: Odds ratios of for group comparisons. "
-
-ordinal_model_s1 <- brm(
-    accuracy_level ~ refClass + (1|id) + (1|state),
-    data = s1_agg,
-    family = cumulative("logit"),
-    cores = 4,
-    iter = 2000,
-    control = list(adapt_delta = 0.99), 
-    prior = c(prior(normal(0, 2), class = "Intercept"), 
-                prior(normal(0, 1.5), class = "b")), 
-    file = paste0(here::here("data/model_cache",'s1_ordinal3.rds')) 
-)
-
-#summary(ordinal_model_s1)
-
-# describe_posterior(ordinal_model_s1, centrality = "Mode") |> 
-#     filter(stringr::str_detect(Parameter, "b_")) |> 
-#     mutate(Parameter = stringr::str_remove(Parameter, "b_")) |> 
-#     mutate(across(where(is.numeric), round, 3)) |> 
-#     kable(booktabs = TRUE)
-
-# Get predicted probabilities
-pred_summary <- ordinal_model_s1 |>
-    epred_draws(newdata = data.frame(refClass = c("kWh", "Percentage", "USD")),
-                ndraws = 1000, re_formula = NA) |>
-    group_by(refClass, .category) |>
-    summarise(
-        mean_prob = mean(.epred),
-        lower_ci = quantile(.epred, 0.025),
-        upper_ci = quantile(.epred, 0.975)
-    )
-#pred_summary |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy")
-
-# Convert log-odds to odds ratios
-posterior_samples <- as.data.frame(ordinal_model_s1)
-odds_ratios <- data.frame(
-  Percentage_vs_kWh = exp(posterior_samples$b_refClassPercentage),
-  USD_vs_kWh = exp(posterior_samples$b_refClassUSD)
-)
-
-# Calculate summary statistics
-odds_ratio_summary <- data.frame(
-    comparison = c("Percentage vs kWh", "USD vs kWh"),
-    odds_ratio = c(mean(odds_ratios$Percentage_vs_kWh),
-                    mean(odds_ratios$USD_vs_kWh)),
-    ci_lower = c(quantile(odds_ratios$Percentage_vs_kWh, 0.025),
-                quantile(odds_ratios$USD_vs_kWh, 0.025)),
-    ci_upper = c(quantile(odds_ratios$Percentage_vs_kWh, 0.975),
-                quantile(odds_ratios$USD_vs_kWh, 0.975))
-)
-
-odds_ratio_summary |> kable(escape=FALSE,booktabs=TRUE,align=c("l")) 
-# |comparison        | odds_ratio| ci_lower| ci_upper|
-# |:-----------------|----------:|--------:|--------:|
-# |Percentage vs kWh |        2.3|     0.62|      6.2|
-# |USD vs kWh        |       14.1|     4.08|     38.0|
-
-#odds_ratio_summary |> pander::pandoc.table(caption="Study 1: Odds ratios of accuracy")
-
-```
-
-```{r}
-#| fig.width: 11
-#| fig.height: 7
-
-bayesplot::color_scheme_set(wes_palettes[1]$BottleRocket1[1:6])
-
-# Plot predicted probabilities
-ggplot(pred_summary, aes(x = refClass, y = mean_prob, fill = .category)) +
-    geom_bar(stat = "identity", position = position_dodge()) +
-    geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci),
-                    position = position_dodge(width = 0.9), width = 0.2) +
-    labs(y = "Predicted Probability", x = "Reference Class", fill = "Accuracy Level") +
-    ggtitle("Study 1: Predicted Probabilities of Accuracy Level by Reference Class") +
-    theme_minimal()
-
-#pp_check(ordinal_model_s1) + ggtitle("Study 1: Posterior Predictive Checks for Ordinal Regression Model") + theme_minimal()
-
-pp_check(ordinal_model_s1, type = "bars_grouped", group="refClass", fatten = 2) +
-    # scale_x_continuous("y", breaks = 1:7) +
-    # scale_y_continuous(NULL, breaks = NULL, expand = expansion(mult = c(0, 0.05))) +
-    ggtitle("Data with posterior predictions",
-            subtitle = "S1. red bars are empirical data, black points are model predictions and credible intervals") +
-    labs(x = "Accuracy Level: 1= exact match,   +1= 0.01-2% error, +2= 2.01-15% error, +3= Over 15% error") +
-    theme_minimal() +
-    theme(legend.background = element_blank() #,legend.position = c(.9, .75)
-    )
-
-```
-
-# Experiment 2
-
-## Methods
-
-The experimental procedures in study 2 are quite similar to those in study 1, but we also included a rounding manipulation (rounded vs.
-not rounded), and a manipulation of the goal (10% reduction vs.
-15% rediction).
-We recruited 206 participants from Amazon Mechanical Turk, but data from from 10 participants were corrupted due to experimenter error, leaving a final sample of 196 participants.
-
-Note that reference class remains a between-subjects variable, while percent goal, rounding, and state are within-subjects variables.
-In study 2, the new design is a 4 state temperature (2 warm vs.
-2 cold states) X 2 task goal (10% vs.
-15%) X 2 last year’s usage for the family and the state average (exact vs.
-rounded numbers) within X 3 task reference class (USD vs.
-Percentage vs.
-kWh) between.
-
-## Results
-
-```{r}
-#| label: tbl-s2-agg
-#| caption: "Study 2: Summary of planning accuracy by reference class.
-The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."
-
-s2_agg1 <- s2_long |> 
-  filter(appliance != "TOTAL") |> 
-  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
-  summarise(
-    total_kWh = sum(value),
-    orig_kWh = sum(family),
-    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
-    state_dif = mean(state_dif),
-    .groups = "drop"
-  ) |>
-  mutate(matched_goal = (pct_change == pct),
-                error = pct_change - pct,
-                abs_error = abs(error)) |> 
-      ungroup() |> # Add ungroup here
-        mutate(
-            accuracy_level = factor(
-                case_when(
-                    abs_error == 0.00 ~ "Exact match",
-                    abs_error <= 0.02 ~ "0.01-2% error",
-                    abs_error <= 0.15 ~ "2.01-15% error",
-                    TRUE ~ "Over 15% error"  # Capture all remaining cases
-                ), 
-                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
-                ordered = TRUE
-            )
-        )
-
-s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc) |> 
-    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
-    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),mean_abs_error=mean(abs_error),n_accuracy=first(n_accuracy)) |> 
-    mutate(accuracy_level = factor(
-            case_when(
-                mean_abs_error < 0.02 ~ "Exact match",
-                mean_abs_error <= 0.02 ~ "0.01-2% error",
-                mean_abs_error <= 0.15 ~ "2.01-15% error",
-                TRUE ~ "Over 15% error"  # Capture all remaining cases
-            ), 
-            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
-            ordered = TRUE
-        ))
-
-# overall pct of subjects who matched their goal
-s2_agg4 |> group_by(refClass) |>
-    summarise(
-    'Avg. % meeting goal' = mean(pct),
-    'Avg. Deviation From Goal' = median(mean_pct_change),
-    # sd = sd(pct),
-    # n = n(),
-    #se=sd(pct)/sqrt(n)
-) |>   mutate(across(where(is.numeric), \(x) round(x, 3))) %>% 
-  kable(escape=FALSE,booktabs=TRUE,align=c("l"))
-
-```
-
-```{r}
-#| label: tbl-s2-prop
-#| caption: "Study 2: The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages).
-The combined group column reflects the percentage of participants in each accuracy level when aggregating across across all reference class conditions."
-
-# compute percentage of subjects per accuracy level per group
-observed_props_s2 <- s2_agg1 |>
-  group_by(refClass, accuracy_level) |>
-  summarise(n = n()) |>
-  group_by(refClass) |>
-  mutate(prop = n/sum(n)) |>
-  mutate(n_prop=paste0(n," (",round(prop*100,1),"%)" ), pct_grp=paste0(round(prop*100,1), "%")) |> ungroup()
-
-observed_props_s2 |> 
-  mutate(n_total=sum(n)/4) |> 
-  group_by(accuracy_level) |>
-  mutate(ns=sum(n)/4) |> 
-  mutate(Total = paste0(round(ns/n_total*100,1), "%")) |>
-  select('Reference Class'=refClass, 'Accuracy Level'=accuracy_level, '% in Group'=pct_grp, "Combined Groups %" =Total) |>
-  pivot_wider(
-    names_from = 'Reference Class',
-    values_from = c('% in Group')
-  ) |> relocate("Combined Groups %" , .after=last_col()) |> 
-  kable(escape=FALSE,booktabs=TRUE,align=c("l"))
-
-```
-
-```{r}
-#| label: fig-s2-plot
-#| caption: "Study 2: Proportion of participants in each accuracy level, colored by reference class. A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
-
-# % of entire sample 
-prop_combo_s2 <- s2_agg1 %>%
-    group_by(refClass, accuracy_level) %>%
-    summarise(count = n()) %>%
-    group_by(refClass) %>%
-    mutate(Probability = count / sum(count)) %>%
-    ungroup()
-
-ggplot(prop_combo_s2, aes(x = accuracy_level, y = Probability, fill = refClass)) +
-    geom_bar(stat = "identity", position = position_dodge()) +
-  scale_y_continuous(labels = scales::percent) +
-    labs(title = "S1. % of Participants within each Accuracy Bin",
-        x = "Accuracy Level",
-        y = "Percentage of Participants",
-        fill = "Reference Class") +
-    theme_minimal() 
-```
-
-```{r}
-#| label: fig-s2-plot2
-#| caption: "Study 2: Proportion of participants in each accuracy level, colored by reference class, and seprated in facets based on the levels of reduction goal, and rounding.
-A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
-#| fig.width: 11
-#| fig.height: 9
-
-# % of entire sample 
-prop_combo_s2_full <- s2_agg1 %>%
-    group_by(refClass, accuracy_level, pct_goal,rounded) %>%
-    summarise(count = n()) %>%
-    group_by(refClass) %>%
-    mutate(Probability = count / sum(count)) %>%
-    ungroup()
-
-ggplot(prop_combo_s2_full, aes(x = accuracy_level, y = Probability, fill = refClass)) +
-    geom_bar(stat = "identity", position = position_dodge()) +
-  scale_y_continuous(labels = scales::percent) +
-    labs(title = "S1. % of Participants within each Accuracy Bin",
-        x = "Accuracy Level",
-        y = "Percentage of Participants",
-        fill = "Reference Class") +
-    facet_wrap(~pct_goal*rounded) +
-    theme_minimal() 
-```
-
-### Individual Differences
-
-```{r}
-#| label: fig-s2-indv
-#| caption: "Study 2: Individual performance in the energy planning task, colored by reference class.
-The dashed line represents the target reduction goal.
-Participants are shown along the y axis, those who fall above or below the dashed line have not met the target goal.
-The x-axis represents the percent change in energy usage from the prior year."
-#| fig.width: 11
-#| fig.height: 9
-
-s2_agg1 |> group_by(id,refClass,calc,pct_goal,pct_change) |> 
-    filter(plan=="plan1",rounded=="Rounded") |> 
-    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
-    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),mean_abs_error=mean(abs_error),n_accuracy=first(n_accuracy)) |>
-  ungroup() |> 
-  mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100) |>
-  filter(mean_abs_error <= 0.50) |>
-  mutate(id=reorder(id,pct_change)) |> 
-  ggplot(aes(y=id,x=mean_pct_change,col=refClass)) + 
-  geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
-  geom_vline(aes(xintercept=goal_pct),linetype="dashed",alpha=.5) +
-  ggh4x::facet_nested_wrap(~pct_goal,axes="all",scales="free",ncol=2)  + 
-  labs(y="Participant Id", x="Percent Change", title="Individual Performance") +
-  theme(axis.text.y=element_text(face = "plain", size = rel(0.7))) + 
-    scale_x_continuous(breaks = seq(0, 0.5, by = 0.05),
-                    labels = scales::percent_format(accuracy = 1))
-
-```
-
-```{r}
-#| label: fig-s2-indv2
-#| caption: "Study 2: Respones patterns for a subset of individiual participants.
-Black points are participant responses, red points are the state average, and blue points are the family average.
-The x-axis represents the appliance category, and the y-axis represents the energy usage in kWh."
-#| fig.width: 11
-#| fig.height: 10
-
-s2_long |> filter(id %in% unique(s2_long$id)[1:30]) |> 
-  filter(appliance!="TOTAL",state=="California") |> 
-  ggplot(aes(x=appliance,y=value)) + geom_point(aes(shape=plan)) +
-  geom_point(aes(y=state_avg),color="red") +
-  geom_point(aes(y=family),color="blue") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-#| 
-s2_long |> filter(id %in% unique(s2_long$id)[1:30]) |> 
-  filter(appliance!="TOTAL",state=="California") |> 
-  ggplot(aes(x=appliance,y=value)) + geom_point(aes(shape=plan)) +
-  geom_point(aes(y=state_avg),color="red") +
-  geom_point(aes(y=family),color="blue") +
-  facet_wrap(~id) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
-```
-
-# Discusion
-
-# References

### 2024-12-07 misc/log_error.qmd
update results section. And some formating improvements.

```{r}

pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,ggplot2,gt,knitr,kableExtra,ggh4x,patchwork,lme4,flextable,pander)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)

walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
theme_set(theme_nice())

s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                close_match = abs_error <= 0.03) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

```

```{r}

pacman::p_load(lmerTest) #provides p-values in lmer summary
# Log transformation of absolute error (add a small constant to avoid log(0))
s1_agg$log_abs_error <- log(s1_agg$abs_error + 0.001)

s1_agg$log_abs_error <- log(s1_agg$abs_error + 0.001)
# Linear mixed-effects model with log-transformed error
error_model <- lmer(log_abs_error ~ refClass + (1|id) + (1|state), data = s1_agg)
summary(error_model)

error_model_table <- broom.mixed::tidy(error_model) %>%
mutate(Estimate = exp(estimate)) %>%
rename('Coefficient' = term,
        'Estimate (Exp)' = Estimate) %>%
select(effect, Coefficient, `Estimate (Exp)`, std.error, statistic)

error_model_table %>%
kable(escape = FALSE, booktabs = TRUE, align = c("lccc"), caption = "Effect of Reference Class on Log-Transformed Error Magnitude (Mixed Model)") %>%
kable_styling(full_width = FALSE)

# diagnostic plots to assess model assumptions
plot(error_model) # residual plot
qqnorm(resid(error_model))
qqline(resid(error_model))

#rename coefficients and add stars to indicate significance
coef_map2 <- c("refClassPercentage" = "Reference Class: Percentage",
              "refClassUSD" = "Reference Class: USD",
              "(Intercept)" = "Intercept")
error_model_table <- modelsummary(error_model,
                                  stars=c('*' = .1, '**' = .05, '***' = .01),
                                  coef_rename = coef_map2,
                                  gof_omit = 'AIC|BIC|Log.Lik.|RMSE|Std.Dev.|VC|sigma|R2 Within',
                                  title = 'Linear Mixed-Effects Model on Log-Transformed Error')

error_model_table
```

Table: Effect of Reference Class on Log-Transformed Error Magnitude (Mixed Model)

|effect   |    Coefficient     | Estimate (Exp) | std.error |statistic |
|:--------|:------------------:|:--------------:|:---------:|:---------|
|fixed    |    (Intercept)     |      0.01      |   0.22    |-19.1     |
|fixed    | refClassPercentage |      1.64      |   0.32    |1.6       |
|fixed    |    refClassUSD     |      4.12      |   0.29    |4.8       |
|ran_pars |  sd__(Intercept)   |      6.26      |    NA     |NA        |
|ran_pars |  sd__(Intercept)   |      1.12      |    NA     |NA        |
|ran_pars |  sd__Observation   |      2.73      |    NA     |NA        |

Fixed effects:
                   Estimate Std. Error      df t value             Pr(>|t|)    
(Intercept)          -4.244      0.222 162.883  -19.11 < 0.0000000000000002 ***
refClassPercentage    0.496      0.316 232.000    1.57                 0.12    
refClassUSD           1.416      0.295 231.832    4.80            0.0000028 ***

```{r}

error_model2 <- glmer(
  log_abs_error ~ refClass + (1|id) + (1|state),
  data = s1_agg,
  family = gaussian(link = "identity")
)
summary(error_model2)

# Publication-worthy table
error_model_table <- broom.mixed::tidy(error_model2) %>%
  mutate(Estimate = exp(estimate)) %>%
  rename('Coefficient' = term,
         'Estimate (Exp)' = Estimate) %>%
  select(effect,Coefficient, `Estimate (Exp)`,std.error,statistic)

error_model_table %>%
  kable(escape = FALSE, booktabs = TRUE, align = c("lccc"), caption = "Effect of Reference Class on Log-Transformed Error Magnitude (Mixed Model)")

```

Table: Effect of Reference Class on Log-Transformed Error Magnitude (Mixed Model)

|effect   |    Coefficient     | Estimate (Exp) | std.error |statistic |
|:--------|:------------------:|:--------------:|:---------:|:---------|
|fixed    |    (Intercept)     |      0.01      |   0.22    |-19.1     |
|fixed    | refClassPercentage |      1.64      |   0.32    |1.6       |
|fixed    |    refClassUSD     |      4.12      |   0.29    |4.8       |
|ran_pars |  sd__(Intercept)   |      6.26      |    NA     |NA        |
|ran_pars |  sd__(Intercept)   |      1.12      |    NA     |NA        |
|ran_pars |  sd__Observation   |      2.73      |    NA     |NA        |

## Bayesian version

```{r}

log_error_model_s1 <- brm(
  log_abs_error ~ refClass + (1|id) + (1|state),
  data = s1_agg,
  family = gaussian(link = "identity"),
  prior = c(prior(normal(0, 3), class = "b"),
            prior(cauchy(0, 1), class = "sd")),
  iter = 2000, warmup = 1000, chains = 4, cores = 4, 
  file = paste0(here::here("data/model_cache",'log_error_model_s1.rds')) 
)

summary(log_error_model_s1)

# exp and make table
log_error_model_table <- as.data.frame(describe_posterior(log_error_model_s1, centrality = "Mean"))[, c(1,2,4,5,6)] |> 
  setNames(c("Parameter", "Estimate", "CI_Lower", "CI_Upper", "pd")) |> 
  mutate(Parameter = stringr::str_remove(Parameter, "b_")) |> 
    mutate(across(where(is.numeric), exp)) |>
  kable(booktabs = TRUE)

# plot density of raw data
s1_agg |> 
  ggplot(aes(x = abs_error)) +
  geom_density(fill = "skyblue", alpha = 0.7) +
  facet_wrap(~ refClass, scales = "free") +ggtitle("Error density by reference class")

# plot density of log-transformed data
s1_agg |> 
  ggplot(aes(x = log_abs_error)) +
  geom_density(fill = "skyblue", alpha = 0.7) +
  facet_wrap(~ refClass, scales = "free") + ggtitle("Log-transformed error density by reference class")

# Regression Coefficients:
#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept             -4.23      0.25    -4.72    -3.73 1.01      630      722
# refClassPercentage     0.51      0.31    -0.08     1.12 1.00      583     1050
# refClassUSD            1.42      0.30     0.86     2.01 1.01      502      795

pp_check(log_error_model_s1, type = "dens_overlay_grouped", group="refClass", fatten = 2,ndraws=100) +
  ggtitle("Posterior Predictive Check - log_error_model_s1") +
  theme_minimal() +
  theme(
    legend.background = element_blank(),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1))

```

|Parameter          | Estimate| CI_Lower| CI_Upper|  pd|
|:------------------|--------:|--------:|--------:|---:|
|Intercept          |     0.01|     0.01|     0.02| 2.7|
|refClassPercentage |     1.66|     0.92|     3.08| 2.6|
|refClassUSD        |     4.15|     2.36|     7.49| 2.7|

```{r}

model_list <- list("Ordinal Model (Accuracy)" = ordinal_model_s1,
                   "Gamma Model (Error Magnitude)" = log_error_model_s1)

ms_table <- modelsummary(model_list,
                         exponentiate = c(TRUE, FALSE), # first model OR, second model no transform
                         statistic = "{estimate} [{conf.low}, {conf.high}]",
                         output = "kableExtra",
                         booktabs = TRUE,
                         align = "lcc",
                         title = "Model Comparison: Accuracy (Ordinal) and Error Magnitude (Gamma)",
                         stars = TRUE) %>%
  kable_styling(full_width = FALSE)

print(ms_table)

```

## quantile regression

```{r}

library(quantreg)

# Fit quantile regression for median and upper quantiles
error_qr <- rq(abs_error ~ refClass, data = s1_agg, tau = c(0.5, 0.9))

# Summarize results
summary(error_qr)

# Publication-worthy table

```

Call: rq(formula = abs_error ~ refClass, tau = c(0.5, 0.9), data = s1_agg)

tau: [1] 0.9

Coefficients:
                   coefficients lower bd upper bd
(Intercept)         0.3720       0.2818   0.5205 
refClassPercentage -0.1180      -0.2234  -0.0031 
refClassUSD        -0.0140      -0.1405   0.1016

```{r}

#--------------------------------------
# 2. Model the effect of refClass on error magnitude
#--------------------------------------
# Considering abs_error as a continuous variable:
# Because this variable might be zero or very small, check its range. If strictly positive, Gamma can be used.
# If zero values exist, consider adding a small constant or using a zero-inflated model.
# Here, we assume abs_error > 0.

# Fit a Gamma regression model with a log link:
error_model_gamma <- brm(
  abs_error ~ refClass + (1 | id) + (1 | state),
  data = s1_agg,
  family = Gamma(link = "log"),
  cores = 4, iter = 5000, control = list(adapt_delta = 0.95),
  prior = c(
    prior(normal(0, 1), class = "b"),
    prior(normal(0, 3), class = "Intercept")
  )
)
```

```{r}

 mix <- mixture(gaussian, gaussian)

    # Priors for the mixture model (adjust as needed)
    prior_mix <- c(
      prior(normal(0, 4), class = "b", dpar = "mu1"),
      prior(normal(0, 4), class = "b", dpar = "mu2"),
      prior(cauchy(0, .5), class = "sd", dpar = "mu1"),
      prior(cauchy(0, .5), class = "sd", dpar = "mu2"),
      prior(dirichlet(2, 2), class = "theta") # Prior for mixing proportions
    )

# Fit the mixture model
mixture_model_s1 <- brm(
  log_abs_error ~ refClass + (1|id),
  data = s1_agg,
  family = mix,
  prior = prior_mix,
  iter = 2000, chains = 2, cores = 2,
  file = paste0(here::here("data/model_cache",'mixture_model_s1.rds'))
)

summary(mixture_model_s1)
pp_check(mixture_model_s1, type = "dens_overlay_grouped", group="refClass", ndraws=100) + ggtitle("mixture_model_s1") + theme_minimal()

pp_check(mixture_model_s1, type = "dens_overlay_grouped", group="refClass", ndraws=100) 
```

### 2024-12-07 Scripts/fun_plot.R

@@ -35,7 +35,8 @@ theme_nice_dist <- function() {
 # theme_set(theme_nice())
 
 theme_nice <- function() {
-  theme_minimal(base_family = "Manrope") +
+  #theme_minimal(base_family = "Manrope") +
+  theme_minimal() +
     theme(
       panel.grid.minor = element_blank(),
       plot.title = element_text(family = "Manrope Extrabold", face = "plain", size = rel(1.35)),

### 2024-12-08 docs/misc/log_error.qmd
big update to results section. Added a few paragraphs to Intro.

--- 2024-12-07 misc/log_error.qmd
@@
     filter(appliance !="Total kWh") |> 
-    group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
+    group_by(id,refClass,state,block,plan,edu,pct_goal,calc) |> 
     summarise(total_kWh = sum(value),orig_kWh=sum(family), 
-                pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
+                pct_change = abs(round((orig_kWh-total_kWh)/orig_kWh,3)), 
                 n_change = sum(value!=family),
@@
                 abs_error = abs(error),
-                close_match = abs_error <= 0.03) |>
+                log_abs_error=log(abs(error)+.007), 
+                close_match = abs_error <= 0.02) |>
     ungroup() |> # Add ungroup here
@@
         ) |> relocate(accuracy_level, .after= "pct_change")
+
+s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
+    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+    summarise(
+    mg=sum(matched_goal),
+    mgc=sum(close_match),
+    n=n(), 
+    pct=mg/n,
+    pct_close=mgc/n,
+    mean_pct_change=mean(pct_change),
+    mean_abs_error=mean(abs_error),
+    mean_log_abs_error=mean(log_abs_error),
+    n_accuracy=first(n_accuracy)) |> 
+    mutate(accuracy_level = factor(
+            case_when(
+                mean_abs_error < 0.02 ~ "Exact match",
+                mean_abs_error <= 0.02 ~ "0.01-2% error",
+                mean_abs_error <= 0.15 ~ "2.01-15% error",
+                TRUE ~ "Over 15% error"  # Capture all remaining cases
+            ), 
+            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+            ordered = TRUE
+        ))
+
+s2_agg1 <- s2_long |> 
+  filter(appliance != "TOTAL") |> 
+  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
+  summarise(
+    total_kWh = sum(value),
+    orig_kWh = sum(family),
+    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
+    state_dif = mean(state_dif),
+    .groups = "drop"
+  ) |>
+  mutate(
+    matched_goal = (pct_change == pct),
+    close_match = abs(pct_change - pct) <= 0.02,
+                error = pct_change - pct,
+                abs_error = abs(error),
+                log_abs_error=log(abs(error)+.007)) |> 
+      ungroup() |> 
+        mutate(
+            accuracy_level = factor(
+                case_when(
+                    abs_error == 0.00 ~ "Exact match",
+                    abs_error <= 0.02 ~ "0.01-2% error",
+                    abs_error <= 0.15 ~ "2.01-15% error",
+                    TRUE ~ "Over 15% error"  # Capture all remaining cases
+                ), 
+                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+                ordered = TRUE
+            )
+        )
+
+s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc) |> 
+    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+    summarise(
+    mg=sum(matched_goal),
+    mgc=sum(close_match),
+    n=n(), 
+    pct=mg/n,
+    pct_close=mgc/n,
+    mean_pct_change=mean(pct_change),
+    mean_abs_error=mean(abs_error),
+    mean_log_abs_error=mean(log_abs_error),
+    n_accuracy=first(n_accuracy)) |> 
+    mutate(accuracy_level = factor(
+            case_when(
+                mean_abs_error < 0.02 ~ "Exact match",
+                mean_abs_error <= 0.02 ~ "0.01-2% error",
+                mean_abs_error <= 0.15 ~ "2.01-15% error",
+                TRUE ~ "Over 15% error"  # Capture all remaining cases
+            ), 
+            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+            ordered = TRUE
+        ))
 
@@
 ```
+
+# S2
+
+```{r}
+
+log_error_model_s2_c1 <- brm(
+  log_abs_error ~ refClass*calc + (1|id) + (1|state),
+  data = s2_agg1,
+  family = gaussian(link = "identity"),
+  prior = c(prior(normal(0, 3), class = "b"),
+            prior(cauchy(0, 1), class = "sd")),
+  iter = 2000, warmup = 1000, chains = 4, cores = 4, 
+  file = paste0(here::here("data/model_cache",'log_error_model_s2_c1.rds')) 
+)
+
+summary(log_error_model_s2_c1)
+conditional_effects(log_error_model_s2_c1)
+
+log_error_model_s2_c1rf <- brm(
+  log_abs_error ~ refClass + (1|id) + (1|state) + (1|calc),
+  data = s2_agg1,
+  family = gaussian(link = "identity"),
+  prior = c(prior(normal(0, 3), class = "b"),
+            prior(cauchy(0, 1), class = "sd")),
+  iter = 2000, warmup = 1000, chains = 4, cores = 4, 
+  file = paste0(here::here("data/model_cache",'log_error_model_s2_c1rf.rds')) 
+)
+
+summary(log_error_model_s2_c1rf)
+
+log_error_model_s2_cAdd_int3 <- brm(
+  log_abs_error ~ (refClass*rounded*pct_goal)+calc + (1|id) + (1|state),
+  data = s2_agg1,
+  family = gaussian(link = "identity"),
+  prior = c(prior(normal(0, 3), class = "b"),
+            prior(cauchy(0, 1), class = "sd")),
+  iter = 3000, warmup = 1000, chains = 4, cores = 4, 
+  file = paste0(here::here("data/model_cache",'log_error_model_s2_cAdd_int3.rds')) 
+)
+
+summary(log_error_model_s2_cAdd_int3)
+conditional_effects(log_error_model_s2_cAdd_int3)
+
+log_error_model_s2_c1rf_int3 <- brm(
+  log_abs_error ~ refClass*rounded*pct_goal + (1|id) + (1|state) + (1|calc),
+  data = s2_agg1,
+  family = gaussian(link = "identity"),
+  prior = c(prior(normal(0, 3), class = "b"),
+            prior(cauchy(0, 1), class = "sd")),
+  iter = 3000, warmup = 1000, chains = 4, cores = 4, 
+  file = paste0(here::here("data/model_cache",'log_error_model_s2_c1rf_int3.rds')) 
+)
+
+summary(log_error_model_s2_c1rf_int3)
+
+conditional_effects(log_error_model_s2_c1rf_int3)
+
+log_error_model_s2_c1rf_add3 <- brm(
+  log_abs_error ~ refClass+rounded+pct_goal + (1|id) + (1|state) + (1|calc),
+  data = s2_agg1,
+  family = gaussian(link = "identity"),
+  prior = c(prior(normal(0, 3), class = "b"),
+            prior(cauchy(0, 1), class = "sd")),
+  iter = 3000, warmup = 1000, chains = 4, cores = 4, 
+  file = paste0(here::here("data/model_cache",'log_error_model_s2_c1rf_add3.rds')) 
+)
+
+summary(log_error_model_s2_c1rf_add3)
+conditional_effects(log_error_model_s2_c1rf_add3)
+
+```
+
+```{r}
+
+```

### 2024-12-08 manuscript/appendix.qmd
big update to results section. Added a few paragraphs to Intro.

```{r}
#| cache: false
pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,
    ggplot2,gt,knitr,kableExtra,ggh4x,patchwork, ggridges,ggstance,lme4,flextable,pander)

options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)

walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))

theme_set(theme_nice())

s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,edu,pct_goal,calc) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = abs(round((orig_kWh-total_kWh)/orig_kWh,3)), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007), 
                close_match = abs_error <= 0.02) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

s2_agg1 <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(matched_goal = (pct_change == pct),
                error = pct_change - pct,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007)) |> 
      ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        )

s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

```

## Study 1

### Amount of change

```{r}

# plot mean_pct_change distribution
s1_agg4 |> ggplot(aes(x = refClass, y = mean_pct_change, fill = refClass)) +
    geom_boxplot() +
    labs(title = "S1. Mean % Change by Reference Class",
        x = "Reference Class",
        y = "Mean % Change",
        fill = "Reference Class") +
    theme_minimal()

# plot mean_pct_change density
s1_agg4 |> ggplot(aes(x = mean_pct_change, fill = refClass)) +
    geom_density(alpha = 0.5) +
    labs(title = "S1. Mean % Change Density by Reference Class",
        x = "Mean % Change",
        fill = "Reference Class") +
    theme_minimal()

```

## Study2

### Goal Manipulation

```{r}

ggplot(s2_agg1, aes(y = pct_goal, x = log_abs_error, color = calc)) +
  geom_density_ridges(aes(fill = calc), alpha = 0.2, scale = 0.5,
                      jittered_points = TRUE, point_alpha = 0.7,point_size=.4,
                      position = position_raincloud(width = 0.05, height = 0.1,
                                                    ygap = 0.05)) +
  geom_boxploth(width = 0.05, alpha = 0.2, outlier.shape = NA, show.legend = FALSE) +
  #scale_y_discrete(expand = expansion(mult = c(0.2, 0.4))) +
 # guides(fill = "none", color = guide_legend(reverse = TRUE)) +
  guides(fill = "none") +
  labs(x = "Log Absolute Deviation", y = "Reference Class", color = "") +
  theme(legend.position = "top")

ggplot(prop_combo_s2, aes(x = accuracy_level, y = Probability, fill = pct_goal)) +
    geom_bar(stat = "identity", position = position_dodge()) +
  scale_y_continuous(labels = scales::percent) +
    labs(title = "S1. % of Participants within each Accuracy Bin",
        x = "Accuracy Level",
        y = "Percentage of Participants",
        fill = "Reference Class") +
    theme_minimal()

ggplot(data = s2_agg1, aes(x = accuracy_level, fill = pct_goal)) +
  geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Accuracy Levels by Reference Class",
       x = "Accuracy Level",
       y = "Percentage of Participants",
       fill = "Reference Class") +
  theme_minimal()

ggplot(data = s2_agg1, aes(x = accuracy_level, fill = pct_goal)) +
geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
scale_y_continuous(labels = scales::percent) +
facet_wrap(~refClass) +
labs(title = "Accuracy Levels by Reference Class",
      x = "Accuracy Level",
      y = "Percentage of Participants",
      fill = "Reference Class") +
theme_minimal()

```

### Rounding

```{r}

ggplot(s2_agg1, aes(y = refClass, x = log_abs_error, color = rounded)) +
  geom_density_ridges(aes(fill = rounded), alpha = 0.2, scale = 0.5,
                      jittered_points = TRUE, point_alpha = 0.7,point_size=.4,
                      position = position_raincloud(width = 0.05, height = 0.1,
                                                    ygap = 0.05)) +
  geom_boxploth(width = 0.05, alpha = 0.2, outlier.shape = NA, show.legend = FALSE) +
  #scale_y_discrete(expand = expansion(mult = c(0.2, 0.4))) +
 # guides(fill = "none", color = guide_legend(reverse = TRUE)) +
  guides(fill = "none") +
  labs(x = "Log Absolute Deviation", y = "Reference Class", color = "") +
  theme(legend.position = "top")

ggplot(data = s2_agg1, aes(x = accuracy_level, fill = rounded)) +
geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
scale_y_continuous(labels = scales::percent) +
facet_wrap(~refClass) +
labs(title = "Accuracy Levels by Reference Class",
      x = "Accuracy Level",
      y = "Percentage of Participants",
      fill = "Reference Class") +
theme_minimal()

```

### Tables

### Distance plots
```{r}
#| label: tbl-s2-agg
#| tbl-cap: "Study 2: Summary of planning accuracy by reference class. The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."

s2_agg1 |> group_by(id,refClass,calc) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy))

s2_agg4 %>% ggplot(aes(x=refClass,y=mean_log_abs_error,fill=refClass)) + 
  stat_summary(fun=mean, geom="bar") +
  stat_summary(fun.data=mean_cl_normal, geom="errorbar", width=0.2) +
  geom_jitter(alpha = 0.2, width = 0.2, height = 0)

s2_agg4 |> 
  ggplot(aes(x = mean_log_abs_error,fill = refClass)) +
  geom_density(alpha = 0.7)

```

### 2024-12-08 manuscript/manuscript.qmd
big update to results section. Added a few paragraphs to Intro.

@@
         output-file: manuscript_hugo.md
+        suppress-bibliography: true
     gfm:
@@
         output-file: manuscript_gfm.md
+        suppress-bibliography: true
     hikmah-manuscript-pdf:
@@
                 dev: "cairo_pdf"
-    hikmah-manuscript-docx: default
+    #hikmah-manuscript-docx: default
+    docx: default
 format-links:   
@@
       icon: file-pdf
-    - hikmah-manuscript-docx
+    - docx
     - gfm
@@
 
-- @canfieldPerceptionsElectricityuseCommunications2017 found that tables were more effective than graphs for conveying specific electricity usage data, likely because tables facilitate straightforward point reading.
-However, they also noted that individuals with lower energy literacy had reduced comprehension across all formats.
-- @canfieldPerceptionsElectricityuseCommunications2017's findings on preferences for historical use information and the impact of neighbor comparisons
-- The concept of cognitive fit posits that performance improves when the information presentation format aligns with the task requirements [@vesseyCognitiveFitTheoryBased1991]
-- alignment can reduce cognitive load and enhance accuracy in planning [@shahBarLineGraph2011]
-- Reimer et al. (2015) provide  context on how numerical formats affect risk perception, the reference class problem, and the benefits of natural frequencies.
-
 Energy poverty continues to be a pervasive issue in the United States @memmottSociodemographicDisparitiesEnergy2021.
@@
 Presenting energy information in absolute units (e.g., kWh) provides a clear reference class, potentially enhancing comprehension.
+
+The concept of cognitive fit posits that performance improves when the information presentation format aligns with the task requirements [@vesseyCognitiveFitTheoryBased1991], and that such an alignment can reduce cognitive load and enhance accuracy in planning [@shahBarLineGraph2011].
+For instance, tables are generally more effective than graphs for conveying specific electricity usage data because they facilitate straightforward point reading [@canfieldPerceptionsElectricityuseCommunications2017].
+However, the effectiveness of the format varies with the type of information and individual differences, such as energy literacy, which significantly impacts comprehension and conservation intent.
+Moreover, the unit in which numerical information is presented influences how decision-makers evaluate and choose between options, with default units increasing value sensitivity [@herberzKiloWhatDefault2020].
+In the context of energy, presenting information in terms of multiple translations can increase preference for options aligned with activated objectives, such as pro-environmental values [@ungemachTranslatedAttributesChoice2018].
+Furthermore, mental accounting mechanisms, where individuals create mental budgets linking specific consumption acts to specific payments, significantly impact energy decisions and behaviors [@hahnelMentalAccountingMechanisms2020].
+
+Evidence from research on energy consumption feedback, normative comparisons, and eco-feedback platforms suggests that comprehensible and contextually meaningful data presentations can improve users’ ability to plan reductions, especially when these formats are integrated into daily routines [@canfieldPerceptionsElectricityuseCommunications2017; @fischerFeedbackHouseholdElectricity2008; @kimMySmartEEcofeedbackGaming2022; @schwartzWhatPeopleConsumption2015].
+Furthermore, temporal and monetary frames have been shown to alter decision quality, with monthly costs or absolute consumption levels often encouraging more energy-efficient intentions than abstract annual or percentage-based metrics [@gillConsumerMatchingCosts2022; @larrickMPGIllusion2008].
+In this context, tailoring reference classes to align with intuitive cognitive processes can help bridge the gap between aggregate reduction goals and targeted, appliance-specific conservation strategies.
 
@@
 We implemented our task and surveys on Qualtrics, and recruited participants through Amazon Mechanical Turk.
-In Experiment 1, 252 participants were intially recruited, but data from 17 participants were corrupted due to experimenter error, leaving a final sample of 235 participants.
+In Experiment 1, 252 participants were initially recruited, but data from 17 participants were corrupted due to experimenter error, leaving a final sample of 235 participants.
 Most participants (76%) reported using a calculator to complete the task.
@@
 
-Participants received energy usage data for two hypothetical families and were tasked with creating action plans to meet specified reduction goals by allocating usage across five appliance categories:  heating, cooling, water heating, refrigerator, and an other appliances (e.g., TV, lighting).
+Participants received energy usage data for two hypothetical families and were tasked with creating action plans to meet specified reduction goals by allocating usage across five appliance categories:  heating, cooling, water heating, refrigerator, and and other appliances (e.g., TV, lighting).
 
-For each family scenario, the particpipants were shown a table containing the families utility usage from the prior year, alongside the state averages for each appliance category (both prior year usage and stage averages are always shown in kWh).
+For each family scenario, the participants were shown a table containing the families utility usage from the prior year, alongside the state averages for each appliance category (both prior year usage and stage averages are always shown in kWh).
 For each scenario, participants were asked to create two possible action plans to achieve the target reduction in total household energy usage (see @fig-task).
@@
 
+### Data Analysis
+
+All preprocessing and analyses were carried out in R [@rcoreteamLanguageEnvironmentStatistical2020] and the tidyverse package [@wickhamWelcomeTidyverse2019].
+Mixed Bayesian regressions were fit using the brms package [@burknerBrmsPackageBayesian2017], with participants and family scenario (states) set as random effects.
+
 ```{r}
-pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,ggplot2,gt,knitr,kableExtra,ggh4x,patchwork,lme4,flextable,pander)
+#| cache: false
+pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,
+    ggplot2,gt,knitr,kableExtra,ggh4x,patchwork, ggridges,ggstance,lme4,flextable,pander)
+
 options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
@@
 walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
+
 theme_set(theme_nice())
@@
   filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
+  filter(!(id %in% readRDS(here::here("data/s1_grp_outlier_ids.rds")))) |>
   mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))
@@
   filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
+  filter(!(id %in% readRDS(here::here("data/s2_grp_outlier_ids.rds"))) ) |>
   mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))
@@
 ```{r}
-#| label: tbl-s1-agg
-#| tbl-cap: "Study 1: Summary of planning accuracy by reference class.
-The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."
 
@@
     mutate(n_accuracy = n_distinct(accuracy_level)) |> 
-    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),mean_abs_error=mean(abs_error),n_accuracy=first(n_accuracy)) |> 
+    summarise(
+    mg=sum(matched_goal),
+    mgc=sum(close_match),
+    n=n(), 
+    pct=mg/n,
+    pct_close=mgc/n,
+    mean_pct_change=mean(pct_change),
+    mean_abs_error=mean(abs_error),
+    mean_log_abs_error=mean(log_abs_error),
+    n_accuracy=first(n_accuracy)) |> 
     mutate(accuracy_level = factor(
@@
 
-# proportion matching vs close match
+```
 
-# proportion matching vs close match
+```{r}
+#| eval: false
+##| label: tbl-s1-agg
+##| tbl-cap: "Study 1: Summary of planning accuracy by reference class.
+The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."
 
 # overall pct of subjects who matched their goal
-s1_agg4 |> group_by(refClass) |>
+s1_agg4 |> group_by('Reference Class' = refClass) |>
     summarise(
-    'Avg. % meeting goal' = mean(pct),
-    'Avg. Deviation From Goal' = median(mean_pct_change),
+    'Avg. % Change' = mean(mean_pct_change),
+    '% meeting goal (exact)' = mean(pct),
+    '% meeting goal (close match)' = mean(pct_close),
+    'Abs. Deviation' = median(mean_abs_error),
+    'Log Abs. Deviation' = (median(mean_log_abs_error)),
     # sd = sd(pct),
@@
 
-For our primary analyses of participants' ability to create accurate energy-saving plans, we employed an accuracy level binning approach by categorizing responses into four distinct levels: Exact match, 0.01–2% error, 2.01–15% error, and Over 15% error.
-The current analysis employs a cumulative ordinal regression model, implemented via a Bayesian hierarchical framework [@burknerBrmsPackageBayesian2017] .
-This approach allows the estimation of threshold parameters and regression coefficients that characterize how changes in predictor variables (such as the reference class: kWh, percentage, or USD) relate to probabilities of being in each accuracy category.
-All analyses were carried out in R [@rcoreteamLanguageEnvironmentStatistical2020] and the tidyverse package [@wickhamWelcomeTidyverse2019].
+::: {#tbl-s1-agg}
+
+|Reference Class |Avg. % Change |% meeting goal (exact) |% meeting goal (close match) |Abs. Deviation |Log Abs. Deviation |
+|:---------------|:-------------|:----------------------|:----------------------------|:--------------|:------------------|
+|kWh             |0.22          |0.38                   |0.54                         |0.03           |-3.7               |
+|Percentage      |0.21          |0.22                   |0.40                         |0.06           |-3.1               |
+|USD             |0.23          |0.10                   |0.22                         |0.10           |-2.4               |
+
+Study 1: Summary of planning accuracy by reference class.
+The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal
+
+:::
+
+@tbl-s1-agg that participants in the kWh condition met the target goal 38% of the time, compared to 22% for the Percentage condition and 10% for the USD condition.
+Moreover, the kWh reference class exhibited smaller deviations from the target reduction, suggesting that participants performed more accurately when the goal was framed in kWh rather than percentages or USD.
+
+As shown in @tbl-s1-agg, participants in the kWh condition exactly met the target reduction goal 38% of the time, significantly outperforming those in the Percentage (22%) and USD (10%) conditions.
+Furthermore, the kWh reference class exhibited notably smaller mean absolute deviations (0.03) compared to Percentage (0.06) and USD (0.10), suggesting that presenting the reduction goal in absolute units facilitated more precise allocations.
+
+```{r}
+#| label: fig-s1-log-dist
+#| fig-cap: "Study 1: Distribution of log absolute error by reference class."
+#| fig-width: 11
+#| fig-height: 8
+
+s1_ld <- ggplot(s1_agg, aes(y = refClass, x = log_abs_error, color = refClass)) +
+  geom_density_ridges(aes(fill = refClass), alpha = 0.2, scale = 0.5,
+                      jittered_points = TRUE, point_alpha = 0.7,point_size=.4,
+                      position = position_raincloud(width = 0.05, height = 0.1,
+                                                    ygap = 0.05)) +
+  geom_boxploth(width = 0.05, alpha = 0.2, outlier.shape = NA, show.legend = FALSE) +
+  #scale_y_discrete(expand = expansion(mult = c(0.2, 0.4))) +
+ # guides(fill = "none", color = guide_legend(reverse = TRUE)) +
+  guides(fill = "none", color = "none") +
+  labs(x = "Log Absolute Deviation", y = "Reference Class", color = "Reference Class") +
+  theme(legend.position = "top")
+
+s1_ldc <- ggplot(s1_agg, aes(y = refClass, x = log_abs_error, color = calc)) +
+  geom_density_ridges(aes(fill = calc), alpha = 0.2, scale = 0.5,
+                      jittered_points = TRUE, point_alpha = 0.7, point_size = .4,
+                      position = position_raincloud(width = 0.05, height = 0.1,
+                                                    ygap = 0.05)) +
+  geom_boxploth(width = 0.05, alpha = 0.2, outlier.shape = NA, show.legend = FALSE) +
+  scale_color_brewer(palette = "Set1") +
+  scale_fill_brewer(palette = "Set1") +
+  guides(fill = "none") +
+  labs(x = "Log Absolute Deviation", y = "Reference Class", color = "") +
+  theme(legend.position = "top")
+
+s1_ld | s1_ldc
+
+```
 
@@
 
-see @fig-s1-plot
-
 ```{r}
@@
 
-We analyzed planning accuracy using Bayesian ordinal regression.
-The dependent variable, plan error, was computed by binning the goal deviation into four ordered levels: exact match (0% error), minor deviations (0.01-2% error), moderate deviations (2.01-15% error), and major deviations (>15% error).
+We next categorized responses into four accuracy levels (exact match [0% error], minor deviations [0.01–2%], moderate deviations [2.01–15%], and major deviations [>15%]) for our primary statistical modeling.
+Using Bayesian ordinal regression, we modeled the ordered accuracy outcome as a function of the reference class condition, while controlling for random variation across participants and family scenarios:
+
+$$
+\text{Accuracy Level} \sim \text{Reference Class} + \text{Calculator} + (1|\text{id}) + (1|\text{Family Scenario})
+$$
+
+This approach allowed us to estimate thresholds (intercepts) and regression coefficients that capture how different reference classes affect the likelihood of achieving higher accuracy categories.
 For each comparison, we provide posterior odds ratios (OR) and their 95% CIs.
+This approach allows the estimation of threshold parameters and regression coefficients that characterize how changes in predictor variables (such as the reference class: kWh, percentage, or USD) relate to probabilities of being in each accuracy category.
 
 ```{r}
-#| label: tbl-s1-ord
+#| eval: true
+##| label: tbl-s1-ord
 ##| tbl-cap: "Study 1: Ordinal Regression Model Results."
@@
 ordinal_model_s1 <- brm(
-    accuracy_level ~ refClass + (1|id) + (1|state),
-    data = s1_agg,
+    accuracy_level ~ refClass +calc + (1|id) + (1|state),
+    data = s1_agg_filtered,
     family = cumulative("logit"),
     cores = 4,
-    iter = 5000,
-    save_pars = save_pars(all = TRUE),
-    control = list(adapt_delta = 0.99), 
+    iter = 2000,
+    control = list(adapt_delta = 0.97), 
     prior = c(prior(normal(0, 3), class = "Intercept"), 
                 prior(normal(0, 3), class = "b")), 
-    file = paste0(here::here("data/model_cache",'s1_ordinal3.rds')) 
+    file = paste0(here::here("data/model_cache",'s1_ordinal_grpclean_calc_2.5.rds')) 
 )
-
-#colnames(mted1) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")
 
@@
 # Get predicted probabilities
-pred_summary <- ordinal_model_s1 |>
-    epred_draws(newdata = data.frame(refClass = c("kWh", "Percentage", "USD")),
-                ndraws = 1000, re_formula = NA) |>
-    group_by(refClass, Category=.category) |>
-    summarise(
-        mean_prob = mean(.epred),
-        lower_ci = quantile(.epred, 0.025),
-        upper_ci = quantile(.epred, 0.975)
-    )
+# pred_summary <- ordinal_model_s1 |>
+#     epred_draws(newdata = data.frame(refClass = c("kWh", "Percentage", "USD")),
+#                 ndraws = 1000, re_formula = NA) |>
+#     group_by(refClass, Category=.category) |>
+#     summarise(
+#         mean_prob = mean(.epred),
+#         lower_ci = quantile(.epred, 0.025),
+#         upper_ci = quantile(.epred, 0.975)
+#     )
 #pred_summary |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy")
 
-#plot(conditional_effects(ordinal_model_s1),points=TRUE)
+# odds ratios of fixed effects
+# as.data.frame(fixef(ordinal_model_s1)[,-2])|> as.data.frame() %>%
+#     rownames_to_column(var = "Parameter") %>%
+#     mutate(across(where(is.numeric), exp)) |>
+#     filter(!stringr::str_detect(Parameter, "Intercept")) |> 
+#     filter(!stringr::str_detect(Parameter, "calc")) |> 
+#     # rename columns to |comparison           | odds_ratio| ci_lower| ci_upper|
+#     rename(Comparison = Parameter, odds_ratio = Estimate, ci_lower = Q2.5, ci_upper = Q97.5) |>
+#     # rename levels to Percentage vs kWh; USD vs kWh; 
+#     mutate(Comparison = case_when(
+#         str_detect(Comparison, "refClassPercentage") ~ "Percentage vs kWh",
+#         str_detect(Comparison, "refClassUSD") ~ "USD vs kWh",
+#         TRUE ~ Comparison
+#     )) |> kable(escape=FALSE,booktabs=TRUE,align=c("l"))
 
-# or_table |>
-#       rownames_to_column(var = "Term") |>
-#       filter(!stringr::str_detect(Term, "Intercept")) |> kable(escape=FALSE,booktabs=TRUE,align=c("l"))
-
-# or_table <- data.frame(exp(fixef(ordinal_model_s1)))
-# or_table |>
-#   rownames_to_column(var = "Term") |>
-#   filter(!stringr::str_detect(Term, "Intercept")) |>
-#   as_flextable() |>
-#   set_header_labels(values = list(
-#     OR = "Odds Ratio",
-#     `2.5 %` = "Lower 95% CI",
-#     `97.5 %` = "Upper 95% CI"
-#   )) |>
-#   flextable::add_footer_lines(values = "Note: Odds ratios derived from the ordinal logistic regression model.") |> 
-#   align(align = 'center', part = 'all') |>
-#   fontsize(size = 10, part = "all")
 ```
 
-::: {#tbl-s1-or}
+::: {#tbl-s1-reg}
+
 |Parameter          |Estimate |CI_Lower |CI_Upper |pd   |
 |:------------------|:--------|:--------|:--------|:----|
-|Intercept[1]       |-1.94    |-3.19    |-0.74    |1.00 |
-|Intercept[2]       |0.20     |-1.02    |1.42     |0.63 |
-|Intercept[3]       |4.36     |3.11     |5.65     |1.00 |
-|refClassPercentage |0.85     |-0.54    |2.23     |0.88 |
-|refClassUSD        |2.72     |1.41     |4.06     |1.00 |
+|Intercept[1]       |-3.8     |-5.45    |-2.28    |1.00 |
+|Intercept[2]       |-1.7     |-3.29    |-0.15    |0.98 |
+|Intercept[3]       |2.8      |1.27     |4.40     |1.00 |
+|refClassPercentage |1.3      |0.01     |2.66     |0.98 |
+|refClassUSD        |2.8      |1.52     |4.04     |1.00 |
+|calcUsedCalculator |-2.8     |-4.09    |-1.56    |1.00 |
 
 **Experiment 1**: Ordinal Regression results.
+Ordinal regression results.
+Positive coefficients for the reference class predictors indicate that those conditions are associated with higher error categories relative to the kWh baseline.
 
@@
 
-::: {#tbl-s1-epred}
-|Term               |Estimate |Est.Error |Q2.5 |Q97.5 |
-|:------------------|:--------|:---------|:----|:-----|
-|refClassPercentage |2.3      |2         |0.58 |9.3   |
-|refClassUSD        |15.2     |2         |4.11 |58.1  |
+::: {#tbl-s1-ord}
+|Comparison        |odds_ratio |ci_lower |ci_upper |
+|:-----------------|:----------|:--------|:--------|
+|Percentage vs kWh |3.7        |1.0      |14       |
+|USD vs kWh        |15.7       |4.6      |57       |
 
-**Experiment 1**: Odds ratios for group comparisons.
+**Experiment 1**: Odds ratios for group comparisons. Odds ratios greater than 1 indicate increased odds of falling into a worse accuracy category compared to the kWh condition.
 
@@
 
-The ordinal model is parameterized with thresholds (intercepts), and positive coefficients can indicate that it is more difficult to achieve higher accuracy categories in the USD condition. 
-The model output suggests that, compared to the kWh condition, the USD condition shows a positive coefficient (Estimate = 2.72, 95% CI: 1.41 to 4.06) for the ordinal outcome.
-At least to me, this positive coefficient appears to indicate that, relative to the kWh reference class, participants in the USD condition are more likely to fall into higher numerical categories of the dependent variable coding.
-However, because the dependent variable is ordered from best (Exact match) to worst (Over 15% error), care is needed in interpretation.
-The Percentage condition coefficient (Estimate = 0.85, 95% CI: -0.54 to 2.23) is more uncertain, with its credible interval overlapping zero.
-Posterior predictive checks (Figure 3) showed that the ordinal model provided a reasonable fit to the observed data (see @fig-s1-plot2).
+As shown in @tbl-s1-reg, the reference class coefficients are positive for both the Percentage (Estimate = 1.3, 95% CI: 0.01 to 2.66, pd = 0.98) and USD (Estimate = 2.8, 95% CI: 1.52 to 4.04, pd = 1.00) conditions, relative to the kWh baseline.
+This indicates that, compared to the kWh condition, participants in both the Percentage and USD conditions were more likely to produce plans that fell into higher error categories.
+Moreover, the odds ratios (see Table 3) suggest that the USD condition led to a notably higher likelihood of large errors compared to the kWh baseline (OR = 15.7), while the Percentage condition also demonstrated increased odds (OR = 3.7) but was somewhat less detrimental to accuracy than USD.
+These results align with our descriptive findings and further clarify that framing the target reductions in absolute kWh units may facilitate significantly more accurate planning.
+Posterior predictive checks showed that the ordinal model provided a reasonable fit to the observed data (see @fig-s1-ppd).
 
 ```{r}
-#| label: fig-s1-plot2
+#| label: fig-s1-ppd
 #| fig-cap: "Study 1: Proportion of participants in each accuracy level, colored by reference class, and seprated in facets based on the levels of reduction goal.
 A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
-#| fig.width: 11
-#| fig.height: 7
-
-#bayesplot::color_scheme_set(wes_palettes[1]$BottleRocket1[1:6])
-
-# Plot predicted probabilities
-# ggplot(pred_summary, aes(x = refClass, y = mean_prob, fill = Category)) +
-#     geom_bar(stat = "identity", position = position_dodge()) +
-#     geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci),
-#                     position = position_dodge(width = 0.9), width = 0.2) +
-#     labs(y = "Predicted Probability", x = "Reference Class", fill = "Accuracy Level") +
-#     ggtitle("Study 1: Predicted Probabilities of Accuracy Level by Reference Class") +
-#     theme_minimal()
-
-#pp_check(ordinal_model_s1) + ggtitle("Study 1: Posterior Predictive Checks for Ordinal Regression Model") + theme_minimal()
 
@@
 ```
+
+## Experiment 1: Discussion
+
+Experiment 1 examined how different numerical representations of energy reduction goals influenced participants’ planning accuracy.
+In line with our hypothesis that absolute units would yield better accuracy, the kWh condition supported significantly more precise energy reduction plans than did either the Percentage or USD conditions.
+Although the Percentage format was detrimental to accuracy relative to kWh, it was the USD condition that consistently produced the poorest outcomes, suggesting that monetary terms, while intuitive in everyday contexts, may not serve as effective reference classes for planning appliance-specific reductions in energy use.
+
+Experiment 2 will extend these findings by examining whether additional variables, such as the difficulty of the reduction goal or the rounding of numerical values, further interact with reference class conditions, thereby providing a more comprehensive understanding of how to optimize energy information presentation for improved planning accuracy.
 
@@
 ```{r}
-#| label: tbl-s2-agg
-#| tbl-cap: "Study 2: Summary of planning accuracy by reference class.
-The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."
-
 s2_agg1 <- s2_long |> 
@@
   ) |>
-  mutate(matched_goal = (pct_change == pct),
+  mutate(
+    matched_goal = (pct_change == pct),
+    close_match = abs(pct_change - pct) <= 0.02,
                 error = pct_change - pct,
                 abs_error = abs(error),
-                log_abs_error=log(abs(error)+.001)) |> 
-      ungroup() |> # Add ungroup here
+                log_abs_error=log(abs(error)+.007)) |> 
+      ungroup() |> 
         mutate(
@@
     mutate(n_accuracy = n_distinct(accuracy_level)) |> 
-    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),
-    mean_abs_error=mean(abs_error),mean_log_abs_error=mean(log_abs_error),
+    summarise(
+    mg=sum(matched_goal),
+    mgc=sum(close_match),
+    n=n(), 
+    pct=mg/n,
+    pct_close=mgc/n,
+    mean_pct_change=mean(pct_change),
+    mean_abs_error=mean(abs_error),
+    mean_log_abs_error=mean(log_abs_error),
     n_accuracy=first(n_accuracy)) |> 
@@
         ))
+```
+
+```{r}
+#| eval: false
+##| label: tbl-s2-agg
+##| tbl-cap: "Study 2: Summary of planning accuracy by reference class.
+The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."
 
 # overall pct of subjects who matched their goal
-s2_agg4 |> group_by(refClass) |>
+s2_agg4 |> group_by('Reference Class' = refClass) |>
     summarise(
-    'Avg. % meeting goal' = mean(pct),
-    'Avg. Abs. Deviation From Goal' = median(mean_pct_change),
-    'Log Deviation' = (median(mean_log_abs_error)),
+    #'Avg. % Change' = mean(mean_pct_change),
+    '% meeting goal (exact)' = mean(pct),
+    '% meeting goal (close match)' = mean(pct_close),
+    'Abs. Deviation' = median(mean_abs_error),
+    'Log Abs. Deviation' = (median(mean_log_abs_error)),
     # sd = sd(pct),
@@
 
-# s2_agg1 |> group_by(id,refClass,calc) |> 
-#     mutate(n_accuracy = n_distinct(accuracy_level)) |> 
-#     summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),
-#     mean_abs_error=mean(abs_error),mean_log_abs_error=mean(log_abs_error),
-#     n_accuracy=first(n_accuracy))
+```
 
-# s2_agg4 %>% ggplot(aes(x=refClass,y=mean_log_abs_error,fill=refClass)) + 
-#   stat_summary(fun=mean, geom="bar") +
-#   stat_summary(fun.data=mean_cl_normal, geom="errorbar", width=0.2) +
-#   geom_jitter(alpha = 0.2, width = 0.2, height = 0)
+::: {#tbl-s2-agg}
 
-# s2_agg4 |> 
-#   ggplot(aes(x = mean_log_abs_error,fill = refClass)) +
-#   geom_density(alpha = 0.7)
+|Reference Class |% meeting goal (exact) |% meeting goal (close match) |Abs. Deviation |Log Abs. Deviation |
+|:---------------|:----------------------|:----------------------------|:--------------|:------------------|
+|kWh             |0.44                   |0.52                         |0.02           |-3.9               |
+|Percentage      |0.28                   |0.42                         |0.06           |-3.2               |
+|USD             |0.20                   |0.29                         |0.10           |-2.4               |
 
-# library(ggridges)
-# s2_agg1 |>  ggplot(aes(y = refClass, x=log_abs_error,fill=refClass,col=refClass)) +
-#     geom_density_ridges(alpha = 0.5) +
-#     geom_vline(aes(xintercept=(log_abs_error)))
+Study 2: Summary of planning accuracy by reference class.
+The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal
 
+:::
+
+```{r}
+#| label: fig-s2-log-dist
+#| fig-cap: "Study 1: Distribution of log absolute error by reference class."
+#| fig-width: 11
+#| fig-height: 8
+
+s2_ld <- ggplot(s2_agg1, aes(y = refClass, x = log_abs_error, color = refClass)) +
+  geom_density_ridges(aes(fill = refClass), alpha = 0.2, scale = 0.5,
+                      jittered_points = TRUE, point_alpha = 0.7,point_size=.4,
+                      position = position_raincloud(width = 0.05, height = 0.1,
+                                                    ygap = 0.05)) +
+  geom_boxploth(width = 0.05, alpha = 0.2, outlier.shape = NA, show.legend = FALSE) +
+  #scale_y_discrete(expand = expansion(mult = c(0.2, 0.4))) +
+ # guides(fill = "none", color = guide_legend(reverse = TRUE)) +
+  guides(fill = "none", color = "none") +
+  labs(x = "Log Absolute Deviation", y = "Reference Class", color = "Reference Class") +
+  theme(legend.position = "top")
+
+s2_ldc <- ggplot(s2_agg1, aes(y = refClass, x = log_abs_error, color = calc)) +
+  geom_density_ridges(aes(fill = calc), alpha = 0.2, scale = 0.5,
+                      jittered_points = TRUE, point_alpha = 0.7,point_size=.4,
+                      position = position_raincloud(width = 0.05, height = 0.1,
+                                                    ygap = 0.05)) +
+  geom_boxploth(width = 0.05, alpha = 0.2, outlier.shape = NA, show.legend = FALSE) +
+  scale_color_brewer(palette = "Set1") +
+  scale_fill_brewer(palette = "Set1") +
+  guides(fill = "none") +
+  labs(x = "Log Absolute Deviation", y = "Reference Class", color = "") +
+  theme(legend.position = "top")
+
+s2_ld + s2_ldc
 ```
@@
 prop_combo_s2 <- s2_agg1 %>%
-    group_by(refClass, accuracy_level) %>%
+    group_by(refClass, accuracy_level,pct_goal) %>%
     summarise(count = n()) %>%
@@
     theme_minimal()
+
 ```
+
+@tbl-s2-prop shows that, once again, participants in the kWh condition achieved closer alignment with the target goals (44% exact matches), followed by Percentage (27%) and USD (18%).
+These percentages are consistent with the patterns observed in Study 1, reinforcing the conclusion that providing goals in kWh supports better accuracy.
 
@@
 
-# Create a new data frame with the predictor values
-# new_data <- expand.grid(
-#   refClass = c("kWh", "Percentage", "USD"),
-#   id = unique(s2_agg1$id)[1],      # Take first id
-#   state = unique(s2_agg1$state)[1], # Take first state,
-#   rounded = unique(s2_agg1$rounded)[1], # Take first rounded
-#   pct_goal = unique(s2_agg1$pct_goal)[1] # Take first pct_goal
-# )
-
-# or_table <- data.frame(exp(fixef(ordinal_model_s2_logit)))
-# or_table |>
-#   rownames_to_column(var = "Term") |>
-#   filter(!stringr::str_detect(Term, "Intercept")) |>
-#   as_flextable() |>
-#   set_header_labels(values = list(
-#     OR = "Odds Ratio",
-#     `2.5 %` = "Lower 95% CI",
-#     `97.5 %` = "Upper 95% CI"
-#   )) |>
-#   flextable::add_footer_lines(values = "Note: Odds ratios derived from the ordinal logistic regression model.") |> 
-#   align(align = 'center', part = 'all') |>
-#   fontsize(size = 10, part = "all")
+as.data.frame(fixef(ordinal_model_s2_logit)[,-2])|> as.data.frame() %>%
+    rownames_to_column(var = "Parameter") %>%
+    mutate(across(where(is.numeric), exp)) |>
+    filter(!stringr::str_detect(Parameter, "Intercept")) |> 
+    # rename columns to |comparison           | odds_ratio| ci_lower| ci_upper|
+    rename(comparison = Parameter, odds_ratio = Estimate, ci_lower = Q2.5, ci_upper = Q97.5) |>
+    # rename levels to Percentage vs kWh; USD vs kWh; Rounded vs Not; 15% Goal vs 10% Goal
+    mutate(comparison = case_when(
+        str_detect(comparison, "refClassPercentage") ~ "Percentage vs kWh",
+        str_detect(comparison, "refClassUSD") ~ "USD vs kWh",
+        str_detect(comparison, "roundedRounded") ~ "Rounded vs Not",
+        str_detect(comparison, "pct_goal15%") ~ "15% Goal vs 10% Goal",
+        TRUE ~ comparison
+    )) |> kable(escape=FALSE,booktabs=TRUE,align=c("l"))
 
@@
 
-::: {#tbl-s2-or}
+::: {#tbl-s2-reg}
 |Parameter          |Estimate |CI_Lower |CI_Upper |pd   |
@@
 
-::: {#tbl-s2-epred}
+::: {#tbl-s2-ord}
 |comparison           | odds_ratio| ci_lower| ci_upper|
@@
 
+We again employed Bayesian ordinal logistic regression to model the
+probability of participants falling into each accuracy category as a
+function of reference class, rounding, and goal level (@tbl-s2-ord and
+@tbl-s2-reg). Results indicated that the kWh condition served as a baseline
+for higher accuracy. Compared to kWh, the USD reference class increased
+the odds of falling into lower-accuracy bins (Odds Ratio = 8.80, 95% CI:
+1.37 to 28.42). The Percentage condition showed a similar trend, though
+the credible intervals were more uncertain. Notably, the “Rounded”
+condition showed an advantage: rounded usage information reduced the
+likelihood of errors (OR = 0.53, 95% CI: 0.36 to 0.73). Moreover, when
+the goal was more challenging (15% vs. 10%), accuracy generally declined
+(OR = 0.66, 95% CI: 0.45 to 0.91). Thus, while rounding facilitated more
+accurate responses, the more difficult goal reduced overall accuracy.
+Crucially, the kWh condition’s advantage persisted across these
+additional manipulations, reinforcing the conclusion from Experiment 1
+that absolute units support better accuracy in energy reduction
+planning.
+
 ```{r}
-#| label: fig-s2-plot3
-
-# ggplot(pred_summary_s2, 
-#         aes(x = .category, y = mean_prob, fill = refClass )) +
-#     geom_col(position = position_dodge(width = 0.8)) +
-#     geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), 
-#                     position = position_dodge(width = 0.8)) +
-#     labs(x = "Reference Class", 
-#         y = "Predicted Probability",
-#         fill = "Accuracy Level",
-#         title = "Predicted Probabilities of Accuracy Levels by Reference Class") +
-#     scale_fill_brewer(palette = "RdYlBu") +
-#     theme_minimal()
+#| label: fig-s2-ppd
 
@@
 
-see @fig-s2-indv
-
 ```{r}
@@
 The x-axis represents the percent change in energy usage from the prior year."
-#| fig.width: 11
-#| fig.height: 9
+#| fig-width: 11
+#| fig-height: 9
 
@@
 
-# Discusion
+# General Discusion
 
 Karjalainen 2011 - people prefer information about price [@karjalainenConsumerPreferencesFeedback2011]
+
+{{< pagebreak >}}
 

### 2024-12-08 misc/indiv.qmd
big update to results section. Added a few paragraphs to Intro.

@@
 
+```{r}
+
+s1_agg2 <- s1 |> 
+    filter(appliance !="Total kWh") |> 
+    group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
+    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
+                pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
+                n_change = sum(value!=family),
+                state_p_dif=mean(state_p_dif),
+                state_f_dif=mean(state_f_dif),
+                n_less_avg = sum(less_avg),
+                duration=first(Duration__in_seconds_)) |> 
+    mutate(matched_goal = (pct_change == pct_goal), 
+                error = pct_change - pct_goal,
+                abs_error = abs(error),
+                log_abs_error=log(abs(error)+.007), 
+                close_match = abs_error <= 0.03) |>
+    ungroup() |> # Add ungroup here
+        mutate(
+            accuracy_level = factor(
+                case_when(
+                    abs_error == 0.00 ~ "Exact match",
+                    abs_error <= 0.02 ~ "0.01-2% error",
+                    abs_error <= 0.15 ~ "2.01-15% error",
+                    TRUE ~ "Over 15% error"  # Capture all remaining cases
+                ), 
+                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+                ordered = TRUE
+            )
+        ) |> relocate(accuracy_level, .after= "pct_change")
+
+s1_agg4 <- s1_agg2 |> group_by(id,refClass,calc) |> 
+    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),
+    mean_abs_error=mean(abs_error),
+    mean_log_abs_error=mean(log_abs_error),
+    n_accuracy=first(n_accuracy)) |> 
+    mutate(accuracy_level = factor(
+            case_when(
+                mean_abs_error < 0.02 ~ "Exact match",
+                mean_abs_error <= 0.02 ~ "0.01-2% error",
+                mean_abs_error <= 0.15 ~ "2.01-15% error",
+                TRUE ~ "Over 15% error"  # Capture all remaining cases
+            ), 
+            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+            ordered = TRUE
+        ))
+
+s1_agg2 |> 
+  ggplot(aes(x =log_abs_error,fill=refClass)) +
+  geom_density(alpha=.5) +
+  geom_boxplot(aes(x =log_abs_error,fill=refClass),y=-.1, width=.1, alpha = 0.5)
+
+library(ggridges)
+library(ggstance)
+ggplot(s1_agg2, aes(y = refClass, x = log_abs_error, color = refClass)) +
+  geom_density_ridges(aes(fill = refClass), alpha = 0.2, scale = 0.5,
+                      jittered_points = TRUE, point_alpha = 0.7,point_size=.4,
+                      position = position_raincloud(width = 0.05, height = 0.3,
+                                                    ygap = 0.05)) +
+  geom_boxploth(width = 0.05, alpha = 0.2, outlier.shape = NA, show.legend = FALSE) +
+  #scale_y_discrete(expand = expansion(mult = c(0.2, 0.4))) +
+ # guides(fill = "none", color = guide_legend(reverse = TRUE)) +
+  guides(fill = "none", color = "none") +
+  labs(x = "Log Absolute Deviation", y = "Reference Class", color = "Reference Class") +
+  theme(legend.position = "top")
+
+s1_agg2 |> 
+  ggplot(aes(x = log_abs_error,fill=refClass,col=refClass)) +
+  ggdist::stat_halfeye(alpha=.7) +
+    geom_point(
+    aes(y = -.05),
+    size = 4, alpha = 0.75,
+    shape = 21,
+    position = position_jitter(
+      height = 0.01,
+      seed = 2343
+    )
+  ) +
+  facet_wrap(~ refClass,ncol=1) + ggtitle("Log-transformed error density by reference class")
+
+s1_agg4 |> 
+  ggplot(aes(x = mean_log_abs_error,fill=refClass,col=refClass)) +
+  ggdist::stat_halfeye(alpha=.7) +
+    geom_point(
+    aes(y = -.1),
+    size = 2, alpha = 0.75,
+    shape = 21,
+    position = position_jitter(
+      height = 0.01,
+      seed = 2343
+    )
+  ) +
+  facet_wrap(~ refClass,ncol=1) + ggtitle("Log-transformed error density by reference class")
+
+```
+
 ## Study 2 - comparison against state average

### 2024-12-08 misc/odds_ratio.qmd
big update to results section. Added a few paragraphs to Intro.

--- 2024-12-07 docs/misc/odds_ratio.qmd
@@
         ) |> relocate(accuracy_level, .after= "pct_change")
+
+s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
+    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+    summarise(
+    mg=sum(matched_goal),
+    mgc=sum(close_match),
+    n=n(), 
+    pct=mg/n,
+    pct_close=mgc/n,
+    mean_pct_change=mean(pct_change),
+    mean_abs_error=mean(abs_error),
+    mean_log_abs_error=mean(log_abs_error),
+    n_accuracy=first(n_accuracy)) |> 
+    mutate(accuracy_level = factor(
+            case_when(
+                mean_abs_error < 0.02 ~ "Exact match",
+                mean_abs_error <= 0.02 ~ "0.01-2% error",
+                mean_abs_error <= 0.15 ~ "2.01-15% error",
+                TRUE ~ "Over 15% error"  # Capture all remaining cases
+            ), 
+            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+            ordered = TRUE
+        ))
 
@@
 
+## S1 exact match
+
+```{r}
+
+# Assuming s1_agg is your dataframe from Experiment 1, and it's already loaded
+
+# Fit a Bayesian logistic regression model
+exact_match_model <- brm(
+  matched_goal ~ refClass + (1|id) + (1|state),
+  data = s1_agg,
+  family = bernoulli(link = "logit"),
+  cores = 4,
+  iter = 2000,
+  control = list(adapt_delta = 0.97),
+  prior = c(
+    prior(normal(0, 2), class = "Intercept"),
+    prior(normal(0, 2), class = "b")
+  ),
+  file = paste0(here::here("data/model_cache", 's1_exact_match_model.rds'))
+)
+summary(exact_match_model)
+
+exact_match_model_c_add <- brm(
+  matched_goal ~ refClass + calc+ (1|id) + (1|state),
+  data = s1_agg,
+  family = bernoulli(link = "logit"),
+  cores = 4,
+  iter = 2000,
+  control = list(adapt_delta = 0.97),
+  prior = c(
+    prior(normal(0, 2), class = "Intercept"),
+    prior(normal(0, 2), class = "b")
+  ),
+  file = paste0(here::here("data/model_cache", 's1_exact_match_model_c_add.rds'))
+)
+summary(exact_match_model_c_add)
+
+exact_match_model_c_int <- brm(
+  matched_goal ~ refClass * calc+ (1|id) + (1|state),
+  data = s1_agg,
+  family = bernoulli(link = "logit"),
+  cores = 4,
+  iter = 2000,
+  control = list(adapt_delta = 0.97),
+  prior = c(
+    prior(normal(0, 2), class = "Intercept"),
+    prior(normal(0, 2), class = "b")
+  ),
+  file = paste0(here::here("data/model_cache", 's1_exact_match_model_c_int.rds'))
+)
+summary(exact_match_model_c_int )
+conditional_effects(exact_match_model_c_int)
+
+```
+
+## outliers
+
+```{r}
+
+identify_outliers_group <- function(x) {
+  Q1 <- quantile(x, 0.25, na.rm = TRUE)
+  Q3 <- quantile(x, 0.75, na.rm = TRUE)
+  IQR_value <- Q3 - Q1
+  lower_bound <- Q1 - 3.3 * IQR_value
+  upper_bound <- Q3 + 3.3 * IQR_value
+  return(x < lower_bound | x > upper_bound)
+}
+
+# 2. Apply the Outlier Detection to abs_error and log_abs_error within each group
+s1_agg <- s1_agg %>%
+  group_by(refClass, calc) %>%
+  mutate(
+    outlier_abs_error = identify_outliers_group(abs_error),
+    outlier_log_abs_error = identify_outliers_group(log_abs_error)
+  ) %>%
+  ungroup()
+
+#s1_agg |> relocate(outlier_abs_error,.after=abs_error) |> sort(desc(outlier_abs_error)) |> head(5)
+s1_agg |> filter(outlier_abs_error) |> select(id,refClass,calc,abs_error,log_abs_error) |> # keep unique id and refClass pairs
+  distinct(id,refClass)
+
+# 3. Summarize the Number and Proportion of Outliers by Group
+outlier_summary_by_group <- s1_agg %>%
+  group_by(refClass, calc) %>%
+  summarise(
+    total_observations = n(),
+    outliers_abs_error = sum(outlier_abs_error),
+    percent_outliers_abs_error = round((outliers_abs_error / total_observations) * 100, 2),
+    outliers_log_abs_error = sum(outlier_log_abs_error),
+    percent_outliers_log_abs_error = round((outliers_log_abs_error / total_observations) * 100, 2)
+  ) %>%
+  ungroup()
+
+# Display the outlier summary by group
+print("Outlier Summary by Reference Class and Calculator Usage:")
+print(outlier_summary_by_group)
+
+s1_agg_no_outliers <- s1_agg %>%
+  filter(!outlier_abs_error & !outlier_log_abs_error)
+
+# Re-run the exact match comparison without outliers
+
+# Create a contingency table of reference class vs. exact match without outliers
+exact_match_table_no_outliers <- s1_agg_no_outliers %>%
+  mutate(exact_match = ifelse(matched_goal, "Exact Match", "No Exact Match")) %>%
+  count(refClass, exact_match) %>%
+  pivot_wider(names_from = exact_match, values_from = n, values_fill = 0)
+
+# Perform Chi-Square Test of Independence without outliers
+chi_square_test_no_outliers <- chisq.test(exact_match_table_no_outliers[, -1])
+
+# Display results
+print("Chi-Square Test Results (Without Outliers):")
+print(chi_square_test_no_outliers)
+
+# Optionally, re-run logistic regression without outliers
+logistic_model_no_outliers <- glm(matched_goal ~ refClass, data = s1_agg_no_outliers, family = binomial)
+summary(logistic_model_no_outliers)
+
+# Calculate Odds Ratios
+odds_ratios_no_outliers <- exp(coef(logistic_model_no_outliers))
+print("Odds Ratios (Without Outliers):")
+print(odds_ratios_no_outliers)
+
+# Calculate 95% Confidence Intervals for Odds Ratios without outliers
+conf_intervals_no_outliers <- exp(confint(logistic_model_no_outliers))
+print("95% Confidence Intervals for Odds Ratios (Without Outliers):")
+print(conf_intervals_no_outliers)
+
+s1_clean <- s1_agg %>%
+  # Remove trials that are outliers in abs_error or log_abs_error
+  #filter(!(outlier_abs_error | outlier_log_abs_error))
+  # remove any id that has any outlier_abs_error
+  filter(!(id %in% (s1_agg |> filter(outlier_abs_error) |> select(id) |> distinct() |> pull(id))))
+
+length(unique(s1_agg$id)) 
+length(unique(s1_clean$id))
+
+# 2. Summary of Cleaned Data
+cleaned_summary <- s1_clean %>%
+  group_by(refClass, calc) %>%
+  summarise(
+    total_observations = n(),
+    exact_matches = sum(matched_goal),
+    percent_exact = round((exact_matches / total_observations) * 100, 2),
+    mean_abs_error = mean(abs_error, na.rm = TRUE),
+    mean_log_abs_error = mean(log_abs_error, na.rm = TRUE)
+  ) %>%
+  ungroup()
+
+# Display the cleaned data summary
+print("Cleaned Data Summary by Reference Class and Calculator Usage:")
+print(cleaned_summary)
+
+ordinal_model_s1_clean_calc <- brm(
+    accuracy_level ~ refClass +calc + (1|id) + (1|state),
+    data = s1_clean,
+    family = cumulative("logit"),
+    cores = 4,
+    iter = 2000,
+    save_pars = save_pars(all = TRUE),
+    control = list(adapt_delta = 0.97), 
+    prior = c(prior(normal(0, 3), class = "Intercept"), 
+                prior(normal(0, 3), class = "b")), 
+    file = paste0(here::here("data/model_cache",'s1_ordinal_clean_calc_3.3.rds')) 
+)
+
+summary(ordinal_model_s1_clean_calc)
+
+ordinal_model_s1_clean_calc_int <- brm(
+    accuracy_level ~ refClass * calc + (1|id) + (1|state),
+    data = s1_clean,
+    family = cumulative("logit"),
+    cores = 4,
+    iter = 2000,
+    save_pars = save_pars(all = TRUE),
+    control = list(adapt_delta = 0.97), 
+    prior = c(prior(normal(0, 3), class = "Intercept"), 
+                prior(normal(0, 3), class = "b")), 
+    file = paste0(here::here("data/model_cache",'s1_ordinal_clean_calc_int.rds')) 
+)
+
+summary(ordinal_model_s1_clean_calc_int)
+conditional_effects(ordinal_model_s1_clean_calc_int)
+
+# logistic regression on exact match with clean data (frequentist)
+logistic_model_clean <- glm(matched_goal ~ refClass, data = s1_clean, family = binomial)
+summary(logistic_model_clean)
+
+# logistic regression on exact match with clean data (frequentist) - mixed effects
+logistic_model_clean_mixed <- glmer(matched_goal ~ refClass + (1|id) + (1|state), data = s1_clean, family = binomial)
+summary(logistic_model_clean_mixed)
+
+# logistic regression on exact match with clean data (frequentist) - mixed effects
+logistic_model_clean_mixed <- glmer(matched_goal ~ refClass+calc + (1|id) , data = s1_clean, family = binomial)
+summary(logistic_model_clean_mixed)
+
+# logistic regression on exact match with clean data (bayesian)
+logistic_model_clean_bayes <- brm(
+  matched_goal ~ refClass + (1|id) + (1|state),
+  data = s1_clean,
+  family = bernoulli(link = "logit"),
+  cores = 4,
+  iter = 2000,
+  control = list(adapt_delta = 0.97),
+  prior = c(
+    prior(normal(0, 2), class = "Intercept"),
+    prior(normal(0, 2), class = "b")
+  ),
+  file = paste0(here::here("data/model_cache", 's1_exact_match_model_clean.rds'))
+)
+summary(logistic_model_clean_bayes)
+conditional_effects(logistic_model_clean_bayes)
+
+```
+
+## outlier from avg.
+
+```{r}
+
+group_stats <- s1_agg |>
+  group_by(refClass, calc) |>
+  summarise(
+    group_mean_abs_error = mean(abs_error, na.rm = TRUE),
+    group_sd_abs_error = sd(abs_error, na.rm = TRUE),
+    .groups = "drop"
+  )
+
+# Calculate participant-level average abs_error
+participant_stats <- s1_agg |>
+  group_by(id, refClass, calc) |>
+  summarise(
+    avg_abs_error = mean(abs_error, na.rm = TRUE),
+    .groups = "drop"
+  )
+
+# Join group stats with participant stats
+participant_stats <- participant_stats |>
+  left_join(group_stats, by = c("refClass", "calc"))
+
+# Identify participants to exclude based on the 2.5 SD threshold
+participants_to_exclude <- participant_stats |>
+  mutate(
+    exclusion_threshold = group_mean_abs_error + 2.5 * group_sd_abs_error,
+    exclude = avg_abs_error > exclusion_threshold
+  ) |>
+  filter(exclude)
+
+ordinal_model_s1_clean_calc <- brm(
+    accuracy_level ~ refClass +calc + (1|id) + (1|state),
+    data = s1_clean,
+    family = cumulative("logit"),
+    cores = 4,
+    iter = 2000,
+    save_pars = save_pars(all = TRUE),
+    control = list(adapt_delta = 0.97), 
+    prior = c(prior(normal(0, 3), class = "Intercept"), 
+                prior(normal(0, 3), class = "b")), 
+    file = paste0(here::here("data/model_cache",'s1_ordinal_clean_calc_3.3.rds')) 
+)
+
+s1_agg_filtered <- s1_agg |>
+  filter(!(id %in% participants_to_exclude$id))
+
+length(unique(s1_agg$id)) 
+length(unique(s1_agg_filtered$id))
+
+ordinal_model_s1_clean_calc <- brm(
+    accuracy_level ~ refClass +calc + (1|id) + (1|state),
+    data = s1_agg_filtered,
+    family = cumulative("logit"),
+    cores = 4,
+    iter = 2000,
+    control = list(adapt_delta = 0.97), 
+    prior = c(prior(normal(0, 3), class = "Intercept"), 
+                prior(normal(0, 3), class = "b")), 
+    file = paste0(here::here("data/model_cache",'s1_ordinal_grpclean_calc_2.5.rds')) 
+)
+
+summary(ordinal_model_s1_clean_calc)
+conditional_effects(ordinal_model_s1_clean_calc)
+
+mixed_model_abs_error <- lmer(log_abs_error ~ refClass + calc+ (1|id) + (1|state), data = s1_agg_filtered)
+summary(mixed_model_abs_error)
+# with p value
+library(lmerTest)
+mixed_model_abs_error <- lmer(log_abs_error ~ refClass + calc+ (1|id) + (1|state), data = s1_agg_filtered)
+summary(mixed_model_abs_error)
+
+```
+
+### clearn log error
+```{r}
+
+s1_clean |> 
+  ggplot(aes(x = abs_error)) +
+  geom_density(fill = "skyblue", alpha = 0.7) +
+  facet_wrap(~ refClass, scales = "free") +ggtitle("Error density by reference class")
+
+# plot density of log-transformed data
+s1_clean |> 
+  ggplot(aes(x = log_abs_error)) +
+  geom_density(fill = "skyblue", alpha = 0.7) +
+  facet_wrap(~ refClass, scales = "free") + ggtitle("Log-transformed error density by reference class")
+
+# mixed effects regression on abs_error (frequentist)
+mixed_model_abs_error <- lmer(abs_error ~ refClass + (1|id) + (1|state), data = s1_clean)
+summary(mixed_model_abs_error)
+
+mixed_model_abs_error <- lmer(abs_error ~ refClass*calc + (1|id) + (1|state), data = s1_clean)
+summary(mixed_model_abs_error)
+
+```
+
 # S2
@@
 
+# Create a new data frame with the predictor values
+# new_data <- expand.grid(
+#   refClass = c("kWh", "Percentage", "USD"),
+#   id = unique(s2_agg1$id)[1],      # Take first id
+#   state = unique(s2_agg1$state)[1], # Take first state,
+#   rounded = unique(s2_agg1$rounded)[1], # Take first rounded
+#   pct_goal = unique(s2_agg1$pct_goal)[1] # Take first pct_goal
+# )
+
+# or_table <- data.frame(exp(fixef(ordinal_model_s2_logit)))
+# or_table |>
+#   rownames_to_column(var = "Term") |>
+#   filter(!stringr::str_detect(Term, "Intercept")) |>
+#   as_flextable() |>
+#   set_header_labels(values = list(
+#     OR = "Odds Ratio",
+#     `2.5 %` = "Lower 95% CI",
+#     `97.5 %` = "Upper 95% CI"
+#   )) |>
+#   flextable::add_footer_lines(values = "Note: Odds ratios derived from the ordinal logistic regression model.") |> 
+#   align(align = 'center', part = 'all') |>
+#   fontsize(size = 10, part = "all")
+
 ```
+
+# Interaction s2
+
+```{r}
+
+ordinal_model_s2_logit_int2 <- brm(
+  accuracy_level ~ refClass +(rounded*pct_goal) + (1|id)+ (1|state),
+  data = s2_agg1,
+  family = cumulative("logit"),
+  cores = 4,
+  iter = 2000,
+  control = list(adapt_delta = 0.99), # Recommended for ordinal models
+  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
+            prior(normal(0, 3), class = "b")), # Priors for predictors
+  file = paste0(here::here("data/model_cache",'s2_logit_int2.rds')) # Cache for efficiency
+)
+summary(ordinal_model_s2_logit_int2)
+
+conditional_effects(ordinal_model_s2_logit_int2)
+
+ordinal_model_s2_logit_int3 <- brm(
+  accuracy_level ~ refClass * rounded*pct_goal + (1|id)+ (1|state),
+  data = s2_agg1,
+  family = cumulative("logit"),
+  cores = 4,
+  iter = 3000,
+  control = list(adapt_delta = 0.99), # Recommended for ordinal models
+  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
+            prior(normal(0, 3), class = "b")), # Priors for predictors
+  file = paste0(here::here("data/model_cache",'s2_logit_int3.rds')) # Cache for efficiency
+)
+summary(ordinal_model_s2_logit_int3)
+
+conditional_effects(ordinal_model_s2_logit_int3)
+
+ordinal_model_s2_logit_c1rf_int3<- brm(
+  accuracy_level ~ refClass * rounded*pct_goal + (1|id)+ (1|state) + (1|calc),
+  data = s2_agg1,
+  family = cumulative("logit"),
+  cores = 4,
+  iter = 3000,
+  control = list(adapt_delta = 0.99), # Recommended for ordinal models
+  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
+            prior(normal(0, 3), class = "b")), # Priors for predictors
+  file = paste0(here::here("data/model_cache",'s2_logit_c1rf_int3.rds')) # Cache for efficiency
+)
+summary(ordinal_model_s2_logit_c1rf_int3)
+
+conditional_effects(ordinal_model_s2_logit_c1rf_int3)
+
+ordinal_model_s2_logit_c1add_int3<- brm(
+  accuracy_level ~ (refClass * rounded*pct_goal) + calc + (1|id)+ (1|state),
+  data = s2_agg1,
+  family = cumulative("logit"),
+  cores = 4,
+  iter = 3000,
+  control = list(adapt_delta = 0.97), # Recommended for ordinal models
+  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
+            prior(normal(0, 3), class = "b")), # Priors for predictors
+  file = paste0(here::here("data/model_cache",'s2_logit_c1add_int3.rds')) # Cache for efficiency
+)
+summary(ordinal_model_s2_logit_c1add_int3)
+
+conditional_effects(ordinal_model_s2_logit_c1add_int3)
+
+```

### 2024-12-08 misc/outlier.qmd
big update to results section. Added a few paragraphs to Intro.

@@
 ---
+
+```{r}
+
+pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,ggplot2,gt,knitr,kableExtra,ggh4x,patchwork,lme4,flextable,pander)
+options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
+
+walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
+theme_set(theme_nice())
+
+s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
+  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
+  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))
+
+s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
+  filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
+  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))
+
+s1_agg <- s1 |> 
+    filter(appliance !="Total kWh") |> 
+    group_by(id,refClass,state,block,plan,edu,pct_goal,calc) |> 
+    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
+                pct_change = abs(round((orig_kWh-total_kWh)/orig_kWh,3)), 
+                n_change = sum(value!=family),
+                state_p_dif=mean(state_p_dif),
+                state_f_dif=mean(state_f_dif),
+                n_less_avg = sum(less_avg),
+                duration=first(Duration__in_seconds_)) |> 
+    mutate(matched_goal = (pct_change == pct_goal), 
+                error = pct_change - pct_goal,
+                abs_error = abs(error),
+                log_abs_error=log(abs(error)+.007), 
+                close_match = abs_error <= 0.02) |>
+    ungroup() |> # Add ungroup here
+        mutate(
+            accuracy_level = factor(
+                case_when(
+                    abs_error == 0.00 ~ "Exact match",
+                    abs_error <= 0.02 ~ "0.01-2% error",
+                    abs_error <= 0.15 ~ "2.01-15% error",
+                    TRUE ~ "Over 15% error"  # Capture all remaining cases
+                ), 
+                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+                ordered = TRUE
+            )
+        ) |> relocate(accuracy_level, .after= "pct_change")
+
+s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
+    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+    summarise(
+    mg=sum(matched_goal),
+    mgc=sum(close_match),
+    n=n(), 
+    pct=mg/n,
+    pct_close=mgc/n,
+    mean_pct_change=mean(pct_change),
+    mean_abs_error=mean(abs_error),
+    mean_log_abs_error=mean(log_abs_error),
+    n_accuracy=first(n_accuracy)) |> 
+    mutate(accuracy_level = factor(
+            case_when(
+                mean_abs_error < 0.02 ~ "Exact match",
+                mean_abs_error <= 0.02 ~ "0.01-2% error",
+                mean_abs_error <= 0.15 ~ "2.01-15% error",
+                TRUE ~ "Over 15% error"  # Capture all remaining cases
+            ), 
+            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+            ordered = TRUE
+        ))
+
+s2_agg1 <- s2_long |> 
+  filter(appliance != "TOTAL") |> 
+  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
+  summarise(
+    total_kWh = sum(value),
+    orig_kWh = sum(family),
+    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
+    state_dif = mean(state_dif),
+    .groups = "drop"
+  ) |>
+  mutate(
+    matched_goal = (pct_change == pct),
+    close_match = abs(pct_change - pct) <= 0.02,
+                error = pct_change - pct,
+                abs_error = abs(error),
+                log_abs_error=log(abs(error)+.007)) |> 
+      ungroup() |> 
+        mutate(
+            accuracy_level = factor(
+                case_when(
+                    abs_error == 0.00 ~ "Exact match",
+                    abs_error <= 0.02 ~ "0.01-2% error",
+                    abs_error <= 0.15 ~ "2.01-15% error",
+                    TRUE ~ "Over 15% error"  # Capture all remaining cases
+                ), 
+                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+                ordered = TRUE
+            )
+        )
+
+s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc) |> 
+    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+    summarise(
+    mg=sum(matched_goal),
+    mgc=sum(close_match),
+    n=n(), 
+    pct=mg/n,
+    pct_close=mgc/n,
+    mean_pct_change=mean(pct_change),
+    mean_abs_error=mean(abs_error),
+    mean_log_abs_error=mean(log_abs_error),
+    n_accuracy=first(n_accuracy)) |> 
+    mutate(accuracy_level = factor(
+            case_when(
+                mean_abs_error < 0.02 ~ "Exact match",
+                mean_abs_error <= 0.02 ~ "0.01-2% error",
+                mean_abs_error <= 0.15 ~ "2.01-15% error",
+                TRUE ~ "Over 15% error"  # Capture all remaining cases
+            ), 
+            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+            ordered = TRUE
+        ))
+
+```
+
+## outlier from avg.
+
+### S1
+
+```{r}
+
+group_stats <- s1_agg |>
+  group_by(refClass, calc) |>
+  summarise(
+    group_mean_abs_error = mean(abs_error, na.rm = TRUE),
+    group_sd_abs_error = sd(abs_error, na.rm = TRUE),
+    .groups = "drop"
+  )
+
+# Calculate participant-level average abs_error
+participant_stats <- s1_agg |>
+  group_by(id, refClass, calc) |>
+  summarise(
+    avg_abs_error = mean(abs_error, na.rm = TRUE),
+    .groups = "drop"
+  )
+
+# Join group stats with participant stats
+participant_stats <- participant_stats |>
+  left_join(group_stats, by = c("refClass", "calc"))
+
+# Identify participants to exclude based on the 2.5 SD threshold
+participants_to_exclude <- participant_stats |>
+  mutate(
+    exclusion_threshold = group_mean_abs_error + 2.5 * group_sd_abs_error,
+    exclude = avg_abs_error > exclusion_threshold
+  ) |>
+  filter(exclude)
+
+#saveRDS(participants_to_exclude$id,here::here("data/s1_grp_outlier_ids.rds"))
+
+length(unique(s1_agg$id)) 
+length(unique(s1_agg_filtered$id))
+
+s1_agg_filtered <- s1_agg |>
+  filter(!(id %in% participants_to_exclude$id))
+
+summary(ordinal_model_s1_clean_calc)
+conditional_effects(ordinal_model_s1_clean_calc)
+
+mixed_model_abs_error <- lmer(log_abs_error ~ refClass + calc+ (1|id) + (1|state), data = s1_agg_filtered)
+summary(mixed_model_abs_error)
+# with p value
+library(lmerTest)
+mixed_model_abs_error <- lmer(log_abs_error ~ refClass + calc+ (1|id) + (1|state), data = s1_agg_filtered)
+summary(mixed_model_abs_error)
+
+```
+
+### S2
+
+```{r}
+
+group_stats <- s2_agg1 |>
+  group_by(refClass, calc) |>
+  summarise(
+    group_mean_abs_error = mean(abs_error, na.rm = TRUE),
+    group_sd_abs_error = sd(abs_error, na.rm = TRUE),
+    .groups = "drop"
+  )
+
+# Calculate participant-level average abs_error
+participant_stats <- s2_agg1 |>
+  group_by(id, refClass, calc) |>
+  summarise(
+    avg_abs_error = mean(abs_error, na.rm = TRUE),
+    .groups = "drop"
+  )
+
+# Join group stats with participant stats
+participant_stats <- participant_stats |>
+  left_join(group_stats, by = c("refClass", "calc"))
+
+# Identify participants to exclude based on the 2.5 SD threshold
+participants_to_exclude <- participant_stats |>
+  mutate(
+    exclusion_threshold = group_mean_abs_error + 2.5 * group_sd_abs_error,
+    exclude = avg_abs_error > exclusion_threshold
+  ) |>
+  filter(exclude)
+
+s2_agg_filtered <- s2_agg1 |>
+  filter(!(id %in% participants_to_exclude$id))
+
+#saveRDS(participants_to_exclude$id,here::here("data/s2_grp_outlier_ids.rds"))
+
+mixed_model_abs_error <- lmer(log_abs_error ~ refClass + calc+ (1|id) + (1|state), data = s2_agg_filtered)
+summary(mixed_model_abs_error)
+# with p value
+library(lmerTest)
+mixed_model_abs_error <- lmer(log_abs_error ~ refClass + calc+ (1|id) + (1|state), data = s2_agg_filtered)
+summary(mixed_model_abs_error)
+
+library(lmerTest)
+mixed_model_abs_error <- lmer(log_abs_error ~ refClass*rounded*pct_goal + calc+ (1|id) + (1|state), data = s2_agg_filtered)
+summary(mixed_model_abs_error)
+
+ordinal_model_s2_logit <- brm(
+  accuracy_level ~ (refClass*rounded*pct_goal)+calc + (1|id)+ (1|state),
+  data = s2_agg_filtered,
+  family = cumulative("logit"),
+  cores = 4,
+  iter = 2000,
+  control = list(adapt_delta = 0.97), # Recommended for ordinal models
+  prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
+            prior(normal(0, 2), class = "b")), # Priors for predictors
+  file = paste0(here::here("data/model_cache",'s2_logit_grpclean_calc_int.rds')) # Cache for efficiency
+)
+summary(ordinal_model_s2_logit)
+
+conditional_effects(ordinal_model_s2_logit)
+
+```
+
+# old
 

### 2024-12-09 docs/misc/outlier.qmd
minor link fix

--- 2024-12-08 misc/outlier.qmd
@@
 
+prob_exact <- posterior_preds[, , "Exact match"]
+prob_minor <- posterior_preds[, , "0.01-2% error"]
+prob_moderate <- posterior_preds[, , "2.01-15% error"]
+prob_major <- posterior_preds[, , "Over 15% error"]
+
+# Compute mean probabilities across all posterior samples
+s2_agg_filtered <- s2_agg_filtered %>%
+  mutate(
+    prob_exact = colMeans(prob_exact),
+    prob_minor = colMeans(prob_minor),
+    prob_moderate = colMeans(prob_moderate),
+    prob_major = colMeans(prob_major)
+  )
+
+emm_refClass_manual <- s2_agg_filtered %>%
+  group_by(refClass) %>%
+  summarise(
+    mean_prob_exact = mean(prob_exact),
+    mean_prob_minor = mean(prob_minor),
+    mean_prob_moderate = mean(prob_moderate),
+    mean_prob_major = mean(prob_major)
+  )
+
+print(emm_refClass_manual)
+
+ggplot(emm_refClass_manual, aes(x = refClass)) +
+  geom_bar(aes(y = mean_prob_exact), stat = "identity", fill = "steelblue") +
+  geom_errorbar(aes(ymin = mean_prob_exact - 0.05, ymax = mean_prob_exact + 0.05), width = 0.2) +
+  labs(title = "Estimated Marginal Means for Reference Class",
+       x = "Reference Class",
+       y = "Probability of Exact Match") +
+  theme_minimal()
 ```

### 2024-12-09 paper.qmd
update results section. And some formating improvements.
minor link fix

@@
 
+  <a class="btn btn-success btn-sm mb-1" href="/manuscript/output/manuscript-code.html" target="_blank">
+   <i class="fa-solid fa-arrow-up-right-from-square"></i>&ensp;HTML <small>(with code)</small>
+  </a>
+
   <a class="btn btn-danger btn-sm mb-1" href="/manuscript/output/rep_study.pdf" target="manuscript">
@@
 
-  <a class="btn btn-warning btn-sm mb-1" target="_blank" href="/Manuscript/output/manuscript.docx">
+  <a class="btn btn-warning btn-sm mb-1" target="_blank" href="/Manuscript/output/Manuscript.docx">
   <iconify-icon inline="" icon="fa6-solid:file-arrow-down"></iconify-icon> &ensp;docx <small>(download)</small>

### 2024-12-09 paper.qmd
fix docx link

@@
 
-  <a class="btn btn-warning btn-sm mb-1" target="_blank" href="/Manuscript/output/Manuscript.docx">
+  <a class="btn btn-warning btn-sm mb-1" target="_blank" href="/manuscript/output/manuscript.docx">
   <iconify-icon inline="" icon="fa6-solid:file-arrow-down"></iconify-icon> &ensp;docx <small>(download)</small>

### 2024-12-09 paper.qmd
fix word link

@@
   <a class="btn btn-warning btn-sm mb-1" target="_blank" href="/manuscript/output/manuscript.docx">
-  <iconify-icon inline="" icon="fa6-solid:file-arrow-down"></iconify-icon> &ensp;docx <small>(download)</small>
+  <i class="fa-solid fa-file-word"></i>&ensp;docx <small>(download)</small>
   </a>

### 2024-12-11 manuscript/appendix.qmd
Analysis on energy literacy

@@
+## Energy Literacy Questions
+
+| **ENERGY LITERACY SURVEY** |  |  |
+| Dewaters and Powers 2011. Assesses participant<br>knowledge of energy names, use, and sources |  |  |
+| The amount of ELECTRICAL ENERGY<br>(ELECTRICITY) we use is measured in units called ...
+| ELS01 | 1 = "Kilowatt (kW)"<br>2 = "Kilowatt-hours (kWh)"<br>3 = "British Thermal Units (BTU)"<br>4 = "Volts (V)"<br>5 = "Horsepower (HP)" |
+
+| Question | Variable | Values |
+|--------------------------|--------------------------|---------------------|
+| The amount of ENERGY consumed by an electrical appliance is equal to the power rating of the appliance (watts or kilowatts) ...
+| ELS02 | 1 = "Multiplied by the cost of electricity"<br>2 = "Added to the cost of electricity"<br>3 = "Multiplied by the time it's used"<br>4 = "Divided by the time it's used"<br>5 = "Added to the time it's used" |
+| When you turn on an incandescent light bulb, which of the following energy conversion takes place?
+| ELS03 | 1 = "Electrical energy to radiant energy (light)"<br>2 = "Chemical energy to radiant energy (light)"<br>3 = "Electrical energy to radiant energy (light) and thermal energy (heat)"<br>4 = "Chemical energy to radiant energy (light) and thermal energy (heat)"<br>5 = "Electrical energy to radiant energy (light) and mechanical energy" |
+| The best reason to buy an ENERGY STAR® appliance is ...
+| ELS04 | 1 = "ENERGY STAR appliances are usually bigger"<br>2 = "ENERGY STAR appliances cost more"<br>3 = "ENERGY STAR appliances use less energy"<br>4 = "ENERGY STAR appliances are more modern looking"<br>5 = "ENERGY STAR appliances cost less" |
+| Which uses the MOST ENERGY in the average American home in one year?
+| ELS05 | 1 = "Refrigerating food and beverages"<br>2 = "Washing and drying clothing"<br>3 = "Heating and cooling rooms"<br>4 = "Heating and cooling water"<br>5 = "Lighting the home" |
+| Which of the following items uses the MOST ELECTRICITY in the average home in one year? | ELS06 | 1 = "Lights"<br>2 = "Refrigerator"<br>3 = "Telephone"<br>4 = "Television"<br>5 = "Computer" |
+| Which of the following sources provides most of the ELECTRICITY in the United States?
+| ELS07 | 1 = "Nuclear power"<br>2 = "Burning petroleum"<br>3 = "Burning coal"<br>4 = "Solar energy"<br>5 = "Water (hydro) power" |
+| Some people think that if we run out of fossil fuels we can just switch over to electric cars.
+What is wrong with this idea?
+| ELS08 | 1 = "Most electricity is currently produced from fossil fuels (coal, oil, natural gas)"<br>2 = "Switching to electric cars will make unemployment rates go up"<br>3 = "It has been proven that it is impossible to build electric cars in great quantities"<br>4 = "You can't use electricity to operate a car"<br>5 = "There is nothing wrong with this idea" |
+
 ```{r}

### 2024-12-11 manuscript/manuscript.qmd
Analysis on energy literacy

@@
 
--   **Energy Literacy Quiz**: An 8-item questionnaire assessing participants' knowledge of energy consumption and conversion.
+-   **Energy Literacy Quiz**: An 8-item questionnaire assessing participants' knowledge of energy consumption and conversion [ @dewatersEnergyLiteracySecondary2011].
 -   **Calculator Usage Tracking**: Questions determined whether participants used a calculator, paper/pen, or other methods to complete the tasks.
@@
 
+```{r}
+#| label: fig-s1-els
+
+s1_els_log_error <- brm(
+    log_abs_error ~ els + (1|id) + (1|state),
+    data = s1_agg,
+    family = gaussian(),
+    cores = 4,
+    iter = 2000,
+    control = list(adapt_delta = 0.97), 
+    prior = c(prior(normal(0, 3), class = "Intercept"), 
+                prior(normal(0, 3), class = "b")), 
+    file = paste0(here::here("data/model_cache",'s1_els_log_error.rds')) 
+)
+
+# summary(s1_els_log_error)
+# conditional_effects(s1_els_log_error)
+
+# Create the conditional effects plot
+conditional_effects_plot <- conditional_effects(s1_els_log_error)
+
+# Extract the data for plotting
+plot_data <- conditional_effects_plot[[1]]
+
+# Create the plot
+ggplot(plot_data, aes(x = els, y = estimate__)) +
+  geom_line(color = "blue", size = 1) +
+  geom_ribbon(aes(ymin = lower__, ymax = upper__), alpha = 0.2) +
+  labs(
+    x = "Energy Literacy Score",
+    y = "Log Absolute Error",
+    title = "Conditional Effect of Energy Literacy on Log Absolute Error"
+  ) +
+  theme_minimal()
+```
+
+To further investigate individual factors that may influence planning accuracy, we examined the relationship between participants' energy literacy scores and their performance on the task.
+Energy literacy was assessed using an 8-item questionnaire adapted from [@dewatersEnergyLiteracySecondary2011], which covers topics such as energy units, appliance energy consumption, and sources of electricity.
+A Bayesian linear regression model was fit with log-transformed absolute error as the outcome variable and energy literacy score as the predictor, controlling for random effects of participant and state: log_abs_error ~ els + (1|id) + (1|state).
+Results indicated a significant negative relationship between energy literacy and log absolute error (Estimate = -2.35, 95% CI: -2.88 to -1.81), suggesting that participants with higher energy literacy scores tended to have smaller deviations from the target reduction goal, and thus more accurate plans overall (@fig-s1-els).
+
 ## Experiment 1: Discussion
@@
 
-We again employed Bayesian ordinal logistic regression to model the
-probability of participants falling into each accuracy category as a
-function of reference class, rounding, and goal level (@tbl-s2-ord and
-@tbl-s2-reg). Results indicated that the kWh condition served as a baseline
-for higher accuracy. Compared to kWh, the USD reference class increased
-the odds of falling into lower-accuracy bins (Odds Ratio = 8.80, 95% CI:
-1.37 to 28.42). The Percentage condition showed a similar trend, though
-the credible intervals were more uncertain. Notably, the “Rounded”
-condition showed an advantage: rounded usage information reduced the
-likelihood of errors (OR = 0.53, 95% CI: 0.36 to 0.73). Moreover, when
-the goal was more challenging (15% vs. 10%), accuracy generally declined
-(OR = 0.66, 95% CI: 0.45 to 0.91). Thus, while rounding facilitated more
-accurate responses, the more difficult goal reduced overall accuracy.
-Crucially, the kWh condition’s advantage persisted across these
-additional manipulations, reinforcing the conclusion from Experiment 1
-that absolute units support better accuracy in energy reduction
-planning.
+We again employed Bayesian ordinal logistic regression to model the probability of participants falling into each accuracy category as a function of reference class, rounding, and goal level (@tbl-s2-ord and @tbl-s2-reg).
+Results indicated that the kWh condition served as a baseline for higher accuracy.
+Compared to kWh, the USD reference class increased the odds of falling into lower-accuracy bins (Odds Ratio = 8.80, 95% CI: 1.37 to 28.42).
+The Percentage condition showed a similar trend, though the credible intervals were more uncertain.
+Notably, the “Rounded” condition showed an advantage: rounded usage information reduced the likelihood of errors (OR = 0.53, 95% CI: 0.36 to 0.73).
+Moreover, when the goal was more challenging (15% vs. 10%), accuracy generally declined (OR = 0.66, 95% CI: 0.45 to 0.91).
+Thus, while rounding facilitated more accurate responses, the more difficult goal reduced overall accuracy.
+Crucially, the kWh condition’s advantage persisted across these additional manipulations, reinforcing the conclusion from Experiment 1 that absolute units support better accuracy in energy reduction planning.
 
@@
 ```
+
+```{r}
+#| label: fig-s2-els
+
+s2_els_log_error <- brm(
+    log_abs_error ~ els + (1|id) + (1|state),
+    data = s2_agg1,
+    family = gaussian(),
+    cores = 4,
+    iter = 3000,
+    control = list(adapt_delta = 0.97), 
+    prior = c(prior(normal(0, 3), class = "Intercept"), 
+                prior(normal(0, 3), class = "b")), 
+    file = paste0(here::here("data/model_cache",'s2_els_log_error.rds')) 
+)
+#summary(s2_els_log_error)
+
+conditional_effects_plot <- conditional_effects(s2_els_log_error)
+plot_data <- conditional_effects_plot[[1]]
+
+ggplot(plot_data, aes(x = els, y = estimate__)) +
+  geom_line(color = "blue", size = 1) +
+  geom_ribbon(aes(ymin = lower__, ymax = upper__), alpha = 0.2) +
+  labs(
+    x = "Energy Literacy Score",
+    y = "Log Absolute Error",
+    title = "Conditional Effect of Energy Literacy on Log Absolute Error"
+  ) +
+  theme_minimal()
+
+```
+
+As in Experiment 1, we further investigated the role of individual differences in energy literacy in predicting planning accuracy.
+A Bayesian linear regression model, analogous to the one used in Experiment 1 (log_abs_error ~ els + (1|id) + (1|state)), revealed a significant negative relationship between energy literacy scores and log-transformed absolute error (Estimate = -3.21, 95% CI: -3.89 to -2.52).
+This finding indicates that participants with higher energy literacy tended to produce more accurate plans, exhibiting smaller deviations from the target reduction goals.
+The conditional effect plot (@fig-s2-els) visually confirms this relationship, showing a clear decreasing trend in log absolute error as energy literacy increases.
+
+## Experiment 2: Discussion
 

### 2024-12-11 misc/indiv.qmd
Analysis on energy literacy

@@
 
-s1 <- readRDS(here::here("data/s1_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds"))))
-s2_long <- readRDS(here::here("data/s2_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds"))))
+s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
+  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
+  filter(!(id %in% readRDS(here::here("data/s1_grp_outlier_ids.rds")))) |>
+  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))
+
+s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
+  filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
+  filter(!(id %in% readRDS(here::here("data/s2_grp_outlier_ids.rds"))) ) |>
+  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))
 
@@
 ```
+
+```{r}
+
+# Assuming you have a data frame 's2_long' with energy literacy scores ('els')
+# and a data frame 's2_agg1' with planning data
+
+# Calculate the deviation of each participant's plan from the state average and family usage
+s2_long <- s2_long |>
+  mutate(dev_state = abs(value - state_avg),
+         dev_family = abs(value - family))
+
+# Merge energy literacy scores
+s2_long <- merge(s2_long, s2_agg1[, c("id", "els")], by = "id")
+
+# Model the relationship between energy literacy and deviations
+model_dev_state <- lm(dev_state ~ els * appliance, data = s2_long)
+model_dev_family <- lm(dev_family ~ els * appliance, data = s2_long)
+
+summary(model_dev_state)
+summary(model_dev_family)
+
+# Visualize the relationship
+ggplot(s2_long, aes(x = els, y = dev_state, color = appliance)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~appliance) +
+  labs(title = "Deviation from State Average vs. Energy Literacy")
+
+ggplot(s2_long, aes(x = els, y = dev_family, color = appliance)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~appliance) +
+  labs(title = "Deviation from Family Usage vs. Energy Literacy")
+```
+
+```{r}
+s2_agg1 <- s2_long |> 
+  filter(appliance != "TOTAL") |> 
+  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded,els) |> 
+  summarise(
+    total_kWh = sum(value),
+    orig_kWh = sum(family),
+    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
+    state_dif = mean(state_dif),
+    .groups = "drop"
+  ) |>
+  mutate(
+    matched_goal = (pct_change == pct),
+    close_match = abs(pct_change - pct) <= 0.02,
+                error = pct_change - pct,
+                abs_error = abs(error),
+                log_abs_error=log(abs(error)+.007)) |> 
+      ungroup() |> 
+        mutate(
+            accuracy_level = factor(
+                case_when(
+                    abs_error == 0.00 ~ "Exact match",
+                    abs_error <= 0.02 ~ "0.01-2% error",
+                    abs_error <= 0.15 ~ "2.01-15% error",
+                    TRUE ~ "Over 15% error"  # Capture all remaining cases
+                ), 
+                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+                ordered = TRUE
+            )
+        )
+
+s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc,els) |> 
+    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+    summarise(
+    mg=sum(matched_goal),
+    mgc=sum(close_match),
+    n=n(), 
+    pct=mg/n,
+    pct_close=mgc/n,
+    mean_pct_change=mean(pct_change),
+    mean_abs_error=mean(abs_error),
+    mean_log_abs_error=mean(log_abs_error),
+    n_accuracy=first(n_accuracy)) |> 
+    mutate(accuracy_level = factor(
+            case_when(
+                mean_abs_error < 0.02 ~ "Exact match",
+                mean_abs_error <= 0.02 ~ "0.01-2% error",
+                mean_abs_error <= 0.15 ~ "2.01-15% error",
+                TRUE ~ "Over 15% error"  # Capture all remaining cases
+            ), 
+            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+            ordered = TRUE
+        ))
+```
+
+```{r}
+
+# Example: Use k-means clustering on participant-level mean accuracy measures
+participant_summary <- s2_agg1 %>%
+  group_by(id) %>%
+  summarize(mean_log_error = mean(log_abs_error),
+            mean_els = mean(els),
+            mean_pct_correct = mean(accuracy_level == "exact_match"))
+
+# Run clustering (e.g., 3 clusters)
+set.seed(123)
+km <- kmeans(participant_summary[, c("mean_log_error", "mean_els", "mean_pct_correct")], centers = 3)
+
+participant_summary$cluster <- factor(km$cluster)
+
+ggplot(participant_summary, aes(x = mean_els, y = mean_log_error, color = cluster)) +
+  geom_point(alpha = 0.7) +
+  theme_minimal() +
+  labs(x = "Mean Energy Literacy", y = "Mean Log Absolute Error",
+       title = "Clustering Participants by Accuracy and Literacy",
+       color = "Cluster")
+
+```
+
+```{r}
+#| eval: false
+# Assuming you have a data frame 's2_long' with energy literacy scores ('els')
+# and a data frame 's2_agg1' with planning data
+
+# Calculate the deviation of each participant's plan from the state average and family usage
+s2_long <- s2_long |>
+  mutate(dev_state = abs(value - state_avg),
+         dev_family = abs(value - family))
+
+# Merge energy literacy scores
+s2_long <- s2_long |> left_join(s2_agg1, by = c("id", "refClass", "calc", "state", "pct", "pct_goal", "plan", "rounded", "els"))
+
+# Model the relationship between energy literacy and deviations
+model_dev_state <- lm(dev_state ~ els * appliance, data = s2_long)
+model_dev_family <- lm(dev_family ~ els * appliance, data = s2_long)
+
+summary(model_dev_state)
+summary(model_dev_family)
+
+# Visualize the relationship
+ggplot(s2_long, aes(x = els, y = dev_state, color = appliance)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~appliance) +
+  labs(title = "Deviation from State Average vs. Energy Literacy")
+
+ggplot(s2_long, aes(x = els, y = dev_family, color = appliance)) +
+  geom_point() +
+  geom_smooth(method = "lm") +
+  facet_wrap(~appliance) +
+  labs(title = "Deviation from Family Usage vs. Energy Literacy")
+
+# Prepare the data for clustering (e.g., using the deviations from state average or family usage)
+cluster_data <- s2_long |>
+  group_by(id, appliance) |>
+  summarize(mean_dev_state = mean(dev_state),
+            mean_dev_family = mean(dev_family)) |>
+  pivot_wider(names_from = appliance, values_from = c(mean_dev_state, mean_dev_family)) |>
+  ungroup() |>
+  select(-id) # Remove ID for clustering
+
+# Scale the data
+cluster_data_scaled <- scale(cluster_data)
+
+# Determine the optimal number of clusters (e.g., using the elbow method)
+wss <- (nrow(cluster_data_scaled) - 1) * sum(apply(cluster_data_scaled, 2, var))
+for (i in 2:15) wss[i] <- sum(kmeans(cluster_data_scaled, centers = i)$withinss)
+plot(1:15, wss, type = "b", xlab = "Number of Clusters", ylab = "Within groups sum of squares")
+
+# Perform k-means clustering (e.g., with 3 clusters)
+kmeans_result <- kmeans(cluster_data_scaled, centers = 3)
+
+# Add cluster assignments to the original data
+s2_long$cluster <- kmeans_result$cluster[match(s2_long$id, unique(s2_long$id))]
+
+# Visualize the clusters
+ggplot(s2_long, aes(x = appliance, y = value, color = factor(cluster))) +
+  geom_point() +
+  facet_wrap(~cluster) +
+  labs(title = "Clusters of Planning Strategies")
+
+```

### 2024-12-11 nb/s1_survey.qmd
Analysis on energy literacy

```{r}
library(pacman)
pacman::p_load(dplyr,purrr,tidyr,here,haven,tibble,
               ggplot2,gt,ggh4x,lme4,flextable,kableExtra,brms,emmeans,tidybayes)

options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
theme_set(theme_nice())
```

```{r}

s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
  filter(!(id %in% readRDS(here::here("data/s1_grp_outlier_ids.rds")))) |>
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s1_agg <- s1 |> 
    filter(appliance !="Total kWh") |> 
    group_by(id,refClass,state,block,plan,edu,pct_goal,calc,els) |> 
    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
                pct_change = abs(round((orig_kWh-total_kWh)/orig_kWh,3)), 
                n_change = sum(value!=family),
                state_p_dif=mean(state_p_dif),
                state_f_dif=mean(state_f_dif),
                n_less_avg = sum(less_avg),
                duration=first(Duration__in_seconds_)) |> 
    mutate(matched_goal = (pct_change == pct_goal), 
                error = pct_change - pct_goal,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007), 
                close_match = abs_error <= 0.02) |>
    ungroup() |> # Add ungroup here
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        ) |> relocate(accuracy_level, .after= "pct_change")

s1_agg4 <- s1_agg |> group_by(id,refClass,calc,els) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

```

```{r}

# Scatter plots with regression lines
ggplot(s1_agg4, aes(x = els, y = mean_abs_error)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +  # Add a linear regression line with standard error
  labs(title = "ELS Score vs. Mean Absolute Error",
       x = "ELS Score",
       y = "Mean Absolute Error") +
  theme_bw()

ggplot(s1_agg4, aes(x = els, y = mean_log_abs_error)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "ELS Score vs. Mean Log Absolute Error",
       x = "ELS Score",
       y = "Mean Log Absolute Error") +
  theme_bw()

ggplot(s1_agg4, aes(x = accuracy_level, y = els, fill = accuracy_level)) +
  geom_boxplot() +
  labs(title = "ELS Score by Accuracy Level",
       x = "Accuracy Level",
       y = "ELS Score") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(s1_agg4, aes(x = calc, y = els, fill = calc)) +
          geom_boxplot() +
          labs(title = "ELS Score by Calculator Use",
               x = "Calculator Use",
               y = "ELS Score") +
          theme_bw()

unique(s1_agg4$accuracy_level)

s1_agg4 |> group_by(accuracy_level) |> summarize(n=n())
s1_agg |> group_by(accuracy_level) |> summarize(n=n())

# Calculate correlations
cor.test(s1_agg4$els, s1_agg4$mean_abs_error)
cor.test(s1_agg4$els, s1_agg4$mean_log_abs_error)

# ANOVA
anova_result <- aov(els ~ accuracy_level, data = s1_agg4)
summary(anova_result)

# Tukey's HSD post-hoc test
TukeyHSD(anova_result)

# ordinal regression
polr_model <- MASS::polr(accuracy_level ~ els + refClass + calc, data = s1_agg4)
summary(polr_model)
emmeans::emmeans(polr_model, pairwise ~ els)

library(ordinal)    
library(parameters)

m_equal <- clm(accuracy_level ~ els, 
    data = s1_agg, 
    link="probit")

parameters::model_parameters(m_equal) |> 
  insight::print_html()
summary(m_equal)
# formula: accuracy_level ~ els
# data:    s1_agg

#  link   threshold nobs logLik   AIC     niter max.grad cond.H 
#  probit flexible  1008 -1213.66 2435.32 8(0)  4.05e-11 6.0e+01

# Coefficients:
#     Estimate Std. Error z value            Pr(>|z|)    
# els   -1.957      0.133   -14.7 <0.0000000000000002 ***
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1

# Threshold coefficients:
#                               Estimate Std. Error z value
# Exact match|0.01-2% error       -2.052      0.099  -20.72
# 0.01-2% error|2.01-15% error    -1.574      0.093  -16.92
# 2.01-15% error|Over 15% error   -0.350      0.083   -4.21

mean2=coef(m_equal)["els"]
Thresholds <- coef(m_equal)[1:3]

polr_model_ref_interaction <- MASS::polr(accuracy_level ~ els * refClass + calc, data = s1_agg4)
summary(polr_model_ref_interaction)
# Example for refClass = USD, calc = No Calculator
new_data <- expand.grid(els = seq(min(s1_agg4$els), max(s1_agg4$els), length.out = 100),
                        refClass = "USD",
                        calc = "No Calculator")

predicted_probs <- predict(polr_model_ref_interaction, newdata = new_data, type = "probs")

# Combine with new_data for plotting
predicted_probs_df <- cbind(new_data, predicted_probs)

# Reshape to long format for ggplot
predicted_probs_long <- tidyr::pivot_longer(predicted_probs_df, 
                                            cols = `Exact match`:`Over 15% error`,
                                            names_to = "accuracy_level",
                                            values_to = "probability")

# Plot
ggplot(predicted_probs_long, aes(x = els, y = probability, color = accuracy_level)) +
  geom_line(linewidth=1.5) +
  labs(title = "Predicted Probabilities of Accuracy Levels by ELS",
       subtitle = "Reference Class = USD, No Calculator",
       x = "ELS Score",
       y = "Probability") +
  theme_bw()

library(lme4)
mixed_model <- lmer(log_abs_error ~ els + (1 | id), data = s1_agg)
emmeans_model <- emmeans::emmeans(mixed_model, pairwise ~ els)
summary(emmeans_model)

```

```{r}

s1_els_log_error <- brm(
    log_abs_error ~ els + (1|id) + (1|state),
    data = s1_agg,
    family = gaussian(),
    cores = 4,
    iter = 2000,
    control = list(adapt_delta = 0.97), 
    prior = c(prior(normal(0, 3), class = "Intercept"), 
                prior(normal(0, 3), class = "b")), 
    file = paste0(here::here("data/model_cache",'s1_els_log_error.rds')) 
)

summary(s1_els_log_error)
# Regression Coefficients:
#           Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept    -1.65      0.23    -2.12    -1.18 1.00      921     1465
# els          -2.35      0.27    -2.88    -1.81 1.00      753      997

conditional_effects(s1_els_log_error)

# Create the conditional effects plot
conditional_effects_plot <- conditional_effects(s1_els_log_error)

# Extract the data for plotting
plot_data <- conditional_effects_plot[[1]]

# Create the plot
ggplot(plot_data, aes(x = els, y = estimate__)) +
  geom_line(color = "blue", size = 1) +
  geom_ribbon(aes(ymin = lower__, ymax = upper__), alpha = 0.2) +
  labs(
    x = "Energy Literacy Score",
    y = "Log Absolute Error",
    title = "Conditional Effect of Energy Literacy on Log Absolute Error"
  ) +
  theme_minimal()

```

# Initial exploration from raw data

## Survey Questions and Energy Literacy Survey

Table 1: Survey Questions and Energy Literacy Survey

| Question | Variable | Values |
|----------|----------|--------|
| "Did you use a calculator?" Used a calculator | MATH01 | 1 = Selected <br> Missing = Not selected |
| Used a paper/pen | MATH02 | 1 = Selected <br> Missing = Not selected |
| Used something else | MATH03 | 1 = Selected <br> Missing = Not selected |
| Used something else text entry | MATH04 | 1 = Selected <br> Missing = Not selected |
| ENERGY LITERACY SURVEY |  |  |
| Dewaters and Powers 2011. Assesses participant knowledge of energy names, use, and sources <br> Question about electrical energy units (kWh) | ELS01 | 1 = "Kilowatt (kW)" <br> 2 = "Kilowatt-hours (kWh)" <br> 3 = "British Thermal Units (BTU)" <br> 4 = "Volts (V)" <br> 5 = "Horsepower (HP)" |
| Question about energy consumed and appliance power rating | ELS02 | 1 = "Multiplied by the cost of electricity" <br> 2 = "Added to the cost of electricity" <br> 3 = "Multiplied by the time it's used" <br> 4 = "Divided by the time it's used" <br> 5 = "Added to the time it's used" |
| Question about incandescent lightbulb conversion | ELS03 | 1 = "Electrical energy to radiant energy (light)" <br> 2 = "Chemical energy to radiant energy (light)" <br> 3 = "Electrical energy to radiant energy (light) and thermal energy (heat)" <br> 4 = "Chemical energy to radiant energy (light) and thermal energy (heat)" <br> 5 = "Electrical energy to radiant energy (light) and mechanical energy" |
| Question about reason to buy energy star appliances | ELS04 | 1 = "ENERGY STAR appliances are usually bigger" <br> 2 = "ENERGY STAR appliances cost more" <br> 3 = "ENERGY STAR appliances use less energy" <br> 4 = "ENERGY STAR appliances are more modern looking" <br> 5 = "ENERGY STAR appliances cost less" |
| Question about which appliances uses the most energy | ELS05 | 1 = "Refrigerating food and beverages" <br> 2 = "Washing and drying clothing" <br> 3 = "Heating and cooling rooms" <br> 4 = "Heating and cooling water" <br> 5 = "Lighting the home" |
| Question about which appliance uses the most electricity | ELS06 | 1 = "Lights" <br> 2 = "Refrigerator" <br> 3 = "Telephone" <br> 4 = "Television" <br> 5 = "Computer" |
| Question about which source provides most electricity in the US | ELS07 | 1 = "Nuclear power" <br> 2 = "Burning petroleum" <br> 3 = "Burning coal" <br> 4 = "Solar energy" <br> 5 = "Water (hydro) power" |
| Question about the problem with electric cars | ELS08 | 1 = "Most electricity is currently produced from fossil fuels (coal, oil, natural gas)" <br> 2 = "Switching to electric cars will make unemployment rates go up" <br> 3 = "It has been proven that it is impossible to build electric cars in great quantities" <br> 4 = "You can't use electricity to operate a car" <br> 5 = "There is nothing wrong with this idea" |

### Correct answers to the questions in the DeWaters & Powers (2011) energy literacy survey:

	1.	ELS01 - Electrical energy units (kWh):
	•	Correct Answer: 2 = Kilowatt-hours (kWh)
	2.	ELS02 - Energy consumed and appliance power rating:
	•	Correct Answer: 3 = Multiplied by the time it’s used
	3.	ELS03 - Incandescent lightbulb conversion:
	•	Correct Answer: 3 = Electrical energy to radiant energy (light) and thermal energy (heat)
	4.	ELS04 - Reason to buy ENERGY STAR appliances:
	•	Correct Answer: 3 = ENERGY STAR appliances use less energy
	5.	ELS05 - Which appliances use the most energy:
	•	Correct Answer: 3 = Heating and cooling rooms
	6.	ELS06 - Which appliance uses the most electricity:
	•	Correct Answer: 2 = Refrigerator
	7.	ELS07 - Which source provides most electricity in the US:
	•	Correct Answer: 3 = Burning coal (Note: While this was accurate for the period of the study, energy sources may have shifted since then.)
	8.	ELS08 - Problem with electric cars:
	•	Correct Answer: 1 = Most electricity is currently produced from fossil fuels (coal, oil, natural gas)

```{r}
# Dewaters and Powers 2011. Assesses participants knowledge of energy names, use, and sources
# - DeWaters, J. E., & Powers, S. E. (2011). Energy literacy of secondary students in New York State (USA): A measure of knowledge, affect, and behavior. Energy Policy, 39(3), 1699–1710. https://doi.org/10.1016/j.enpol.2010.12.049

s1 <- haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav"))
s1 <- s1 |> 
  mutate(
    state1 = case_when(
      Colorado == 1 ~ "Colorado",
      Massachusetts == 1 ~ "Massachusetts",
      California == 1 ~ "California",
      Texas == 1 ~ "Texas",
      TRUE ~ NA_character_
    ),
    state2 = case_when(
      Colorado == 2 ~ "Colorado",
      Massachusetts == 2 ~ "Massachusetts",
      California == 2 ~ "California",
      Texas == 2 ~ "Texas",
      TRUE ~ NA_character_
    )) |> 
  mutate(refClass = case_when(
  ReferenceClass == 1 ~ "USD",
  ReferenceClass == 2 ~ "kWh",
  ReferenceClass == 3 ~ "Percentage"
)) |> 
  rename(calc=MATH01) |>  # replace NA's in calc with 0's
  rename(id2=ResponseId) |>
  mutate(id=row_number()) |> relocate(id) |> 
  mutate(calc=factor(ifelse(is.na(calc),0,calc))) |> 
  relocate(refClass,state1,state2,calc, .after = id)

els <- s1 |> select(id,id2,calc,refClass,state1,state2, ELS01:ELS08)

els2 <- els |> 
  pivot_longer(cols = ELS01:ELS08, names_to = "question", values_to = "response") |> 
  # Convert response to numeric to avoid label issues
  mutate(response = as.numeric(response)) |>
  mutate(question = case_when(
    question == "ELS01" ~ "Electrical energy units (kWh)",
    question == "ELS02" ~ "Energy consumed and appliance power rating",
    question == "ELS03" ~ "Incandescent lightbulb conversion",
    question == "ELS04" ~ "Reason to buy energy star appliances",
    question == "ELS05" ~ "Which appliances uses the most energy",
    question == "ELS06" ~ "Which appliance uses the most electricity",
    question == "ELS07" ~ "Which source provides most electricity in the US",
    question == "ELS08" ~ "Problem with electric cars",
    TRUE ~ NA_character_
  )) |> 
  mutate(correct = case_when(
    question == "Electrical energy units (kWh)" & response == 2 ~ "Correct",
    question == "Energy consumed and appliance power rating" & response == 3 ~ "Correct",
    question == "Incandescent lightbulb conversion" & response == 3 ~ "Correct",
    question == "Reason to buy energy star appliances" & response == 3 ~ "Correct",
    question == "Which appliances uses the most energy" & response == 3 ~ "Correct",
    question == "Which appliance uses the most electricity" & response == 2 ~ "Correct",
    question == "Which source provides most electricity in the US" & response == 3 ~ "Correct",
    question == "Problem with electric cars" & response == 1 ~ "Correct",
    TRUE ~ "Incorrect"
  ))

els_id <- els2 |> group_by(id,id2,state1,state2,refClass) |> 
    summarize(els=mean(correct=="Correct"), .groups="drop")

# plot distribution fo answers to each els question. Separate questions by facets, color by correct/incorrect, make facets go in same order as questions
els2 |> 
  ggplot(aes(x=response, fill=correct)) +
  geom_bar(position="dodge", stat="count") +
  facet_wrap(~question, scales="free_y", ncol=2) +
  scale_fill_manual(values=c("Correct"="green","Incorrect"="red")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title="Energy Literacy Survey", x="Response", y="Count") +
  theme(legend.position = "none")

ggplot(els2, aes(x = as.factor(response), fill = correct)) +
  geom_bar(position = "dodge") +
  facet_wrap(~question, ncol = 2, scales = "free_x") + # Facet by question, 2 columns, free x-axis scales
  scale_fill_manual(values = c("Correct" = "#00ba38", "Incorrect" = "#f8766d")) + # Custom colors
  labs(
    title = "Distribution of Answers to Energy Literacy Survey Questions",
    subtitle = "DeWaters & Powers (2011)",
    x = "Response Option",
    y = "Number of Respondents",
    fill = "Correctness"
  ) +
  theme_bw() + # Use a clean theme
  theme(
    strip.text = element_text(size = 10, face = "bold"), # Adjust facet label appearance
    axis.text.x = element_text(size = 8, angle = 45, hjust = 1), # Rotate x-axis labels for readability
    axis.text.y = element_text(size = 8),
    axis.title = element_text(size = 10),
    plot.title = element_text(size = 12, face = "bold"),
    plot.subtitle = element_text(size = 10),
    legend.position = "bottom", # Move legend to the bottom
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9)
  )

# score per id
els2 |> 
  group_by(id) |> 
  summarise(score=sum(correct=="Correct")/n(), .groups="drop") |> 
  ggplot(aes(x=score)) +
  geom_histogram(binwidth=.1, fill="blue", color="black") +
  labs(title="Energy Literacy Survey Score", x="Score", y="Count") +
  theme_minimal()

# score per id, color by ReferenceClass
els2 |> 
  group_by(id, refClass) |> 
  summarise(score=sum(correct=="Correct")/n(), .groups="drop") |> 
  ggplot(aes(x=score, fill=refClass)) +
  geom_histogram(binwidth=.1, color="black") +
  labs(title="Energy Literacy Survey Score", x="Score", y="Count") +
  theme_minimal()

# compare refClass averages, stat_halfeye
els2 |> 
  group_by(id,refClass) |> 
  summarise(score=sum(correct=="Correct")/n(), .groups="drop") |> 
  ggplot(aes(x=refClass, y=score, col=refClass)) +
  stat_summary(fun.data = "mean_cl_boot", geom = "pointrange") +
    labs(title="Energy Literacy Survey Score", x="Reference Class", y="Score")

els2 |>
  group_by(refClass, question) |>
  summarize(mean_correct = mean(correct == "Correct", na.rm = TRUE)) |>
  ggplot(aes(x = question, y = mean_correct, fill = refClass)) +
  geom_col(position = "dodge") +
  labs(title = "Mean Correctness by Reference Class and Question",
       x = "Question",
       y = "Proportion Correct",
       fill = "Reference Class") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

r$> els_id
# A tibble: 252 × 6
      id id2               state1   state2 refClass     els
   <int> <chr>             <chr>    <chr>  <chr>      <dbl>
 1     1 R_2QMsYuHS385pTj1 Colorado Texas  Percentage 0.125
 2     2 R_2tzIyZavpl1RZBY Colorado Texas  Percentage 0.75 
 3     3 R_3Gm1Y4QzSlyrEWo Colorado Texas  Percentage 0.75 
 4     4 R_Djh2q728ibyNWjD Colorado Texas  Percentage 0.875
 5     5 R_3fAXNbIZ7kFieSu Colorado Texas  Percentage 0.125
 6     6 R_1JECzl10KgckJXg Colorado Texas  Percentage 0.375
 7     7 R_C2IQAZim25sij4d Colorado Texas  Percentage 0.25 
 8     8 R_5bbi3tpMOhuGBIR Colorado Texas  Percentage 0.75 
 9     9 R_2BsVSAOpehXW9Cr Colorado Texas  Percentage 0.25 
10    10 R_325F8wRln52w1QN Colorado Texas  Percentage 0.625

head(s1_new |> filter(appliance=="Cooling",block==1,plan=="plan1"),n=16)
# A tibble: 16 × 22
# Groups:   id [16]
      id id2           state1 state2 refClass calc  edu   modality state Duration__in_seconds_ appliance   els family state_avg plan  value block pct_goal change state_p_dif state_f_dif less_avg
   <int> <chr>         <chr>  <chr>  <chr>    <chr> <fct> <chr>    <chr>                 <dbl> <chr>     <dbl>  <dbl> <dbl+lbl> <fct> <dbl> <dbl>    <dbl>  <dbl>       <dbl>       <dbl> <lgl>   
 1     1 R_2QMsYuHS38… Color… Texas  Percent… Used… Grad… P        COL                    1008 Cooling   0.125    697  498      plan1   600     1     0.15    -97         102         199 FALSE   
 2     2 R_2tzIyZavpl… Color… Texas  Percent… Used… Coll… P        COL                    1126 Cooling   0.75     697  498      plan1   500     1     0.15   -197           2         199 FALSE   
 3     3 R_3Gm1Y4QzSl… Color… Texas  Percent… Used… Coll… P        COL                     870 Cooling   0.75     697  498      plan1   592     1     0.15   -105          94         199 FALSE   
 4     4 R_Djh2q728ib… Color… Texas  Percent… Used… Coll… P        COL                    2763 Cooling   0.875    697  498      plan1   400     1     0.15   -297         -98         199 TRUE    
 5     5 R_3fAXNbIZ7k… Color… Texas  Percent… No C… Grad… P        COL                     920 Cooling   0.125    697  498      plan1   320     1     0.15   -377        -178         199 TRUE    
 6     6 R_1JECzl10Kg… Color… Texas  Percent… No C… Grad… P        COL                     878 Cooling   0.375    697  498      plan1   550     1     0.15   -147          52         199 FALSE   
 7     7 R_C2IQAZim25… Color… Texas  Percent… Used… Grad… P        COL                    2680 Cooling   0.25     697  498      plan1   500     1     0.15   -197           2         199 FALSE   
 8     8 R_5bbi3tpMOh… Color… Texas  Percent… Used… Some… P        COL                    1628 Cooling   0.75     697  498      plan1   500     1     0.15   -197           2         199 FALSE   
 9     9 R_2BsVSAOpeh… Color… Texas  Percent… Used… Grad… P        COL                     724 Cooling   0.25     697  498      plan1   650     1     0.15    -47         152         199 FALSE   
10    10 R_325F8wRln5… Color… Texas  Percent… Used… Some… P        COL                     913 Cooling   0.625    697  498      plan1   593     1     0.15   -104          95         199 FALSE

```{r}

library(dplyr)
library(tidyr)
library(corrplot)

# Create a wide dataframe where each column is a question, and the values are 1 (Correct) or 0 (Incorrect)
els_wide <- els2 |>
  mutate(correct_binary = ifelse(correct == "Correct", 1, 0)) |>
  # mutate number correct for each sbj, then filter out those with less than 5, 
    group_by(id) |>
    mutate(n_correct = sum(correct_binary)) |>
    filter(n_correct >= 2) |>
    ungroup() |>
  pivot_wider(id_cols = c(id), names_from = question, values_from = correct_binary) |>
  select(-id) # Remove the ID column for correlation analysis

# Calculate the correlation matrix
cor_matrix <- cor(els_wide, use = "pairwise.complete.obs") # Handle missing values

# Visualize the correlation matrix using corrplot
corrplot(cor_matrix, 
         method = "color",  # Use colored squares to represent correlations
         type = "upper",    # Show only the upper triangle
         tl.col = "black",  # Color of text labels
         tl.srt = 45,       # Rotate text labels for better readability
         addCoef.col = "black", # Add correlation coefficients to the plot
         number.cex = 0.7,   # Adjust the size of the correlation coefficients
         diag = FALSE       # Exclude the diagonal (which is always 1)
)

els2 |>
  group_by(refClass, question) |>
  summarize(mean_correct = mean(correct == "Correct", na.rm = TRUE)) |>
  ggplot(aes(x = question, y = mean_correct, fill = refClass)) +
  geom_col(position = "dodge") +
  labs(title = "Mean Correctness by Reference Class and Question",
       x = "Question",
       y = "Proportion Correct",
       fill = "Reference Class") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

els2 |>
  group_by(calc, question) |>
  summarize(mean_correct = mean(correct == "Correct", na.rm = TRUE)) |>
  ggplot(aes(x = question, y = mean_correct, fill = calc)) +
  geom_col(position = "dodge") +
  labs(title = "Mean Correctness by Calculator Use",
       x = "Question",
       y = "Proportion Correct",
       fill = "Calculator Use") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

library(factoextra)

# Perform PCA
pca_result <- prcomp(els_wide, scale. = TRUE)

# Visualize the results
fviz_pca_var(pca_result, col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE)

```

### 2024-12-11 nb/s2_survey.qmd
Analysis on energy literacy

```{r}

pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,
    ggplot2,gt,knitr,kableExtra,ggh4x,patchwork, ggridges,ggstance,lme4,flextable,pander)

options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
theme_set(theme_nice())

s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
  filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
  filter(!(id %in% readRDS(here::here("data/s2_grp_outlier_ids.rds"))) ) |>
  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))

s2_agg1 <- s2_long |> 
  filter(appliance != "TOTAL") |> 
  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded,els) |> 
  summarise(
    total_kWh = sum(value),
    orig_kWh = sum(family),
    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
    state_dif = mean(state_dif),
    .groups = "drop"
  ) |>
  mutate(
    matched_goal = (pct_change == pct),
    close_match = abs(pct_change - pct) <= 0.02,
                error = pct_change - pct,
                abs_error = abs(error),
                log_abs_error=log(abs(error)+.007)) |> 
      ungroup() |> 
        mutate(
            accuracy_level = factor(
                case_when(
                    abs_error == 0.00 ~ "Exact match",
                    abs_error <= 0.02 ~ "0.01-2% error",
                    abs_error <= 0.15 ~ "2.01-15% error",
                    TRUE ~ "Over 15% error"  # Capture all remaining cases
                ), 
                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
                ordered = TRUE
            )
        )

s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc,els) |> 
    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
    summarise(
    mg=sum(matched_goal),
    mgc=sum(close_match),
    n=n(), 
    pct=mg/n,
    pct_close=mgc/n,
    mean_pct_change=mean(pct_change),
    mean_abs_error=mean(abs_error),
    mean_log_abs_error=mean(log_abs_error),
    n_accuracy=first(n_accuracy)) |> 
    mutate(accuracy_level = factor(
            case_when(
                mean_abs_error < 0.02 ~ "Exact match",
                mean_abs_error <= 0.02 ~ "0.01-2% error",
                mean_abs_error <= 0.15 ~ "2.01-15% error",
                TRUE ~ "Over 15% error"  # Capture all remaining cases
            ), 
            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
            ordered = TRUE
        ))

```

```{r}

unique(s2_agg4$accuracy_level)

s2_agg4 |> group_by(accuracy_level) |> summarize(n=n())
s2_agg1 |> group_by(accuracy_level) |> summarize(n=n())

# Calculate correlations
cor.test(s2_agg4$els, s2_agg4$mean_abs_error)
cor.test(s2_agg4$els, s2_agg4$mean_log_abs_error)
cor.test(s2_agg1$els, s2_agg1$log_abs_error)

# ANOVA
anova_result <- aov(els ~ accuracy_level, data = s2_agg4)
summary(anova_result)

# Tukey's HSD post-hoc test
TukeyHSD(anova_result)

# ordinal regression
polr_model <- MASS::polr(accuracy_level ~ els + refClass + calc, data = s2_agg4)
summary(polr_model)
emmeans::emmeans(polr_model, pairwise ~ els)

library(ordinal)    
library(parameters)

m_equal <- clm(accuracy_level ~ els, 
    data = s2_agg1, 
    link="probit")

parameters::model_parameters(m_equal) |> 
  insight::print_html()
summary(m_equal)

# Fit a mixed-effects model with accuracy level as the outcome
library(lme4)
mixed_model <- lmer(abs_error ~ els + (1 | id), data = s2_agg1)

# Post-hoc pairwise comparisons
emmeans_model <- emmeans(mixed_model, pairwise ~ els)
summary(emmeans_model)

# Scatter plots with regression lines
ggplot(s2_agg4, aes(x = els, y = mean_abs_error)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +  # Add a linear regression line with standard error
  labs(title = "ELS Score vs. Mean Absolute Error",
       x = "ELS Score",
       y = "Mean Absolute Error") +
  theme_bw()

ggplot(s2_agg4, aes(x = els, y = mean_log_abs_error)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "ELS Score vs. Mean Log Absolute Error",
       x = "ELS Score",
       y = "Mean Log Absolute Error") +
  theme_bw()

ggplot(s2_agg4, aes(x = accuracy_level, y = els, fill = accuracy_level)) +
  geom_boxplot() +
  labs(title = "ELS Score by Accuracy Level",
       x = "Accuracy Level",
       y = "ELS Score") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(s2_agg4, aes(x = calc, y = els, fill = calc)) +
          geom_boxplot() +
          labs(title = "ELS Score by Calculator Use",
               x = "Calculator Use",
               y = "ELS Score") +
          theme_bw()

polr_model_ref_interaction <- MASS::polr(accuracy_level ~ els * refClass + calc, data = s2_agg1)
summary(polr_model_ref_interaction)

# Example for refClass = USD, calc = No Calculator
new_data <- expand.grid(els = seq(min(s2_agg1$els), max(s2_agg1$els), length.out = 100),
                        refClass = "USD",
                        calc = "No Calculator")

predicted_probs <- predict(polr_model_ref_interaction, newdata = new_data, type = "probs")

# Combine with new_data for plotting
predicted_probs_df <- cbind(new_data, predicted_probs)

# Reshape to long format for ggplot
predicted_probs_long <- tidyr::pivot_longer(predicted_probs_df, 
                                            cols = `Exact match`:`Over 15% error`,
                                            names_to = "accuracy_level",
                                            values_to = "probability")

# Plot
ggplot(predicted_probs_long, aes(x = els, y = probability, color = accuracy_level)) +
  geom_line(linewidth=1.5) +
  labs(title = "Predicted Probabilities of Accuracy Levels by ELS",
       subtitle = "Reference Class = USD, No Calculator",
       x = "ELS Score",
       y = "Probability") +
  theme_bw()

s2_agg <- s2_agg1 %>%
  mutate(els_quartile = ntile(els, 4))

# Plot distribution of log_abs_error by ELS quartile and reference class
ggplot(s2_agg, aes(x = factor(els_quartile), y = log_abs_error, fill = refClass)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, position = position_dodge(width = 0.8)) +
  labs(x = "Energy Literacy Quartile", y = "Log Absolute Error",
       title = "Distribution of Planning Accuracy by Energy Literacy Quartile and Reference Class") +
  theme_minimal()

```

```{r}

s2_els_log_error <- brm(
    log_abs_error ~ els + (1|id) + (1|state),
    data = s2_agg1,
    family = gaussian(),
    cores = 4,
    iter = 3000,
    control = list(adapt_delta = 0.97), 
    prior = c(prior(normal(0, 3), class = "Intercept"), 
                prior(normal(0, 3), class = "b")), 
    file = paste0(here::here("data/model_cache",'s2_els_log_error.rds')) 
)

summary(s2_els_log_error)
# Regression Coefficients:
#           Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept    -1.10      0.27    -1.62    -0.56 1.00      975     1675
# els          -3.21      0.35    -3.89    -2.52 1.00      837     1508

# library(performance)
#r2(s2_els_log_error)
 # Conditional R2: 0.800 (95% CI [0.782, 0.815])
 #  Marginal R2: 0.270 (95% CI [0.186, 0.350])

 #icc(s2_els_log_error)
# Intraclass Correlation Coefficient
  # Adjusted ICC: 0.734
  # Unadjusted ICC: 0.541
#check_model(s2_els_log_error)

# Create the conditional effects plot
conditional_effects_plot <- conditional_effects(s2_els_log_error)

# Extract the data for plotting
plot_data <- conditional_effects_plot[[1]]

# Create the plot
ggplot(plot_data, aes(x = els, y = estimate__)) +
  geom_line(color = "blue", size = 1) +
  geom_ribbon(aes(ymin = lower__, ymax = upper__), alpha = 0.2) +
  labs(
    x = "Energy Literacy Score",
    y = "Log Absolute Error",
    title = "Conditional Effect of Energy Literacy on Log Absolute Error"
  ) +
  theme_minimal()

s2_els_log_error_full_add <- brm(
    log_abs_error ~ els+refClass+calc+rounded+pct_goal + (1|id) + (1|state),
    data = s2_agg1,
    family = gaussian(),
    cores = 4,
    iter = 4000,
    control = list(adapt_delta = 0.97), 
    prior = c(prior(normal(0, 3), class = "Intercept"), 
                prior(normal(0, 3), class = "b")), 
    file = paste0(here::here("data/model_cache",'s2_els_log_error_full_add.rds')) 
)
summary(s2_els_log_error_full_add)

check_model(s2_els_log_error_full_add)

s2_els_acc_full_add <- brm(
    accuracy_level ~ els+refClass+calc+rounded+pct_goal + (1|id) + (1|state),
    data = s2_agg1,
    cores = 4,
    family = cumulative("logit"),
    control = list(adapt_delta = 0.97), 
    prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 2), class = "b")), # Priors for predictors
    file = paste0(here::here("data/model_cache",'s2_els_acc_error_full_add.rds')) 
)
summary(s2_els_acc_full_add)
# Regression Coefficients:
#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept[1]          -5.67      0.84    -7.31    -4.01 1.00     1039     1978
# Intercept[2]          -4.22      0.83    -5.84    -2.59 1.00     1032     2049
# Intercept[3]          -0.50      0.80    -2.04     1.08 1.00      964     1707
# els                   -6.38      0.95    -8.29    -4.59 1.00      972     1605
# refClassPercentage     0.55      0.58    -0.57     1.66 1.00      950     1652
# refClassUSD            1.07      0.62    -0.13     2.31 1.00      844     1563
# calcNoCalculator       3.18      0.67     1.90     4.55 1.01      872     1585
# roundedRounded        -0.65      0.18    -0.99    -0.31 1.00     5287     3387
# pct_goal15%           -0.43      0.17    -0.76    -0.08 1.00     6491     3227

#check_model(s2_els_acc_full_add)

s2_els_acc_full_int1 <- brm(
    accuracy_level ~ (els*refClass)+(rounded*pct_goal)+calc + (1|id) + (1|state),
    data = s2_agg1,
    cores = 4,
    family = cumulative("logit"),
    control = list(adapt_delta = 0.97), 
    prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 2), class = "b")), # Priors for predictors
    file = paste0(here::here("data/model_cache",'s2_els_acc_full_int1.rds')) 
)
summary(s2_els_acc_full_int1)
conditional_effects(s2_els_acc_full_int1)
# Intercept[1]                  -5.64      0.92    -7.48    -3.82 1.00     1119     2097
# Intercept[2]                  -4.19      0.91    -6.00    -2.36 1.00     1084     2079
# Intercept[3]                  -0.45      0.90    -2.23     1.31 1.00     1074     1995
# els                           -6.13      1.10    -8.30    -3.95 1.00     1140     2308
# refClassPercentage             0.97      0.96    -0.87     2.90 1.00     1139     1717
# refClassUSD                    1.08      1.02    -0.90     3.11 1.01     1066     1943
# roundedRounded                -0.88      0.48    -1.83     0.02 1.00      795     1583
# pct_goal15%                   -0.66      0.48    -1.60     0.24 1.01      809     1652
# calcNoCalculator               3.18      0.68     1.86     4.55 1.00      956     1621
# els:refClassPercentage        -0.66      1.32    -3.36     1.89 1.00     1270     1956
# els:refClassUSD               -0.10      1.41    -2.83     2.64 1.00     1650     2359
# roundedRounded:pct_goal15%     0.46      0.90    -1.22     2.20 1.00      737     1534

#check_model(s2_els_acc_full_int1)
```

# Original

| **ENERGY LITERACY SURVEY** |  |  |
| Dewaters and Powers 2011. Assesses participant<br>knowledge of energy names, use, and sources |  |  |
| The amount of ELECTRICAL ENERGY<br>(ELECTRICITY) we use is measured in units called ... | ELS01 | 1 = "Kilowatt (kW)"<br>2 = "Kilowatt-hours (kWh)"<br>3 = "British Thermal Units (BTU)"<br>4 = "Volts (V)"<br>5 = "Horsepower (HP)" |

| Question | Variable | Values |
|--------------------------|--------------------------|---------------------|
| The amount of ENERGY consumed by an electrical appliance is equal to the power rating of the appliance (watts or kilowatts) ... | ELS02 | 1 = "Multiplied by the cost of electricity"<br>2 = "Added to the cost of electricity"<br>3 = "Multiplied by the time it's used"<br>4 = "Divided by the time it's used"<br>5 = "Added to the time it's used" |
| When you turn on an incandescent light bulb, which of the following energy conversion takes place? | ELS03 | 1 = "Electrical energy to radiant energy (light)"<br>2 = "Chemical energy to radiant energy (light)"<br>3 = "Electrical energy to radiant energy (light) and thermal energy (heat)"<br>4 = "Chemical energy to radiant energy (light) and thermal energy (heat)"<br>5 = "Electrical energy to radiant energy (light) and mechanical energy" |
| The best reason to buy an ENERGY STAR® appliance is ... | ELS04 | 1 = "ENERGY STAR appliances are usually bigger"<br>2 = "ENERGY STAR appliances cost more"<br>3 = "ENERGY STAR appliances use less energy"<br>4 = "ENERGY STAR appliances are more modern looking"<br>5 = "ENERGY STAR appliances cost less" |
| Which uses the MOST ENERGY in the average American home in one year? | ELS05 | 1 = "Refrigerating food and beverages"<br>2 = "Washing and drying clothing"<br>3 = "Heating and cooling rooms"<br>4 = "Heating and cooling water"<br>5 = "Lighting the home" |
| Which of the following items uses the MOST ELECTRICITY in the average home in one year? | ELS06 | 1 = "Lights"<br>2 = "Refrigerator"<br>3 = "Telephone"<br>4 = "Television"<br>5 = "Computer" |
| Which of the following sources provides most of the ELECTRICITY in the United States? | ELS07 | 1 = "Nuclear power"<br>2 = "Burning petroleum"<br>3 = "Burning coal"<br>4 = "Solar energy"<br>5 = "Water (hydro) power" |
| Some people think that if we run out of fossil fuels we can just switch over to electric cars. What is wrong with this idea? | ELS08 | 1 = "Most electricity is currently produced from fossil fuels (coal, oil, natural gas)"<br>2 = "Switching to electric cars will make unemployment rates go up"<br>3 = "It has been proven that it is impossible to build electric cars in great quantities"<br>4 = "You can't use electricity to operate a car"<br>5 = "There is nothing wrong with this idea" |

```{r}

library(pacman)
pacman::p_load(dplyr,purrr,tidyr,here, haven,tibble,ggplot2,gt,ggh4x,lme4,flextable,ggh4x)

options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
theme_set(theme_nice())

#202009

s2 <- haven::read_sav(here::here("data/Frequency & Probability Study 2 - 3-24-19.sav"))
s2 <- s2 |> #mutate(id = as.numeric(format(EndDate, "%H%M%S"))) |> relocate(id) |> 
  mutate(id=paste0("s2_",row_number())) |> relocate(id)
s2 <- s2 |> filter(!(id=="s2_78")) |> # NA for some responses
  filter(!(id=="s2_196")) # NA for state_avg for COL HEAT - (HEATSTATE2)
  # filter(!(id=="s2_208")) # NA for response for California not rounded Total (TOTALAPONE1 and TOTALAPTWO1) - leave them in since all the acutal app values are okay
s2 <- s2 |> 
  mutate(
    state1 = case_when(
      COLORADO == 1 ~ "Colorado",
      MASSACHUSETTS == 1 ~ "Massachusetts",
      CALIFORNIA == 1 ~ "California",
      TEXAS == 1 ~ "Texas",
      TRUE ~ NA_character_
    ),
    state2 = case_when(
      COLORADO == 2 ~ "Colorado",
      MASSACHUSETTS == 2 ~ "Massachusetts",
      CALIFORNIA == 2 ~ "California",
      TEXAS == 2 ~ "Texas",
      TRUE ~ NA_character_
    )
  ) |> 
  mutate(refClass = case_when(
  REFERENCECLASS == 1 ~ "USD",
  REFERENCECLASS == 2 ~ "kWh",
  REFERENCECLASS == 3 ~ "Percentage"
)) |> 
  rename(calc=MATH01) |>  # replace NA's in calc with 0's
  mutate(calc=factor(ifelse(is.na(calc),0,calc))) |> 
  mutate(calc = if_else(calc == 0, "No Calculator", "Calculator")) |> 
  relocate(refClass,state1,state2,calc, .after = id)

s2 <- s2 |> rename(edu=DEM04) |>
  mutate(edu = factor(case_when(
    edu == '1' ~ "Some schooling, but no diploma or degree",
    edu == 2 ~ "Highschool diploma or GED",
    edu == 3 ~ "Some college",
    edu == 4 ~ "College degree",
    edu == 5 ~ "Some graduate school",
    edu == 6 ~ "Graduate degree",
    edu == 7 ~ "Choose not to answer",
    TRUE ~ NA_character_
  ))) |> relocate(edu, .after = id)

els <- s2 |> select(id,calc,refClass,state1,state2, ELS01:ELS08)

els2 <- els |> 
  pivot_longer(cols = ELS01:ELS08, names_to = "question", values_to = "response") |> 
  # Convert response to numeric to avoid label issues
  mutate(response = as.numeric(response)) |>
  mutate(question = case_when(
    question == "ELS01" ~ "Electrical energy units (kWh)",
    question == "ELS02" ~ "Energy consumed and appliance power rating",
    question == "ELS03" ~ "Incandescent lightbulb conversion",
    question == "ELS04" ~ "Reason to buy energy star appliances",
    question == "ELS05" ~ "Which appliances uses the most energy",
    question == "ELS06" ~ "Which appliance uses the most electricity",
    question == "ELS07" ~ "Which source provides most electricity in the US",
    question == "ELS08" ~ "Problem with electric cars",
    TRUE ~ NA_character_
  )) |> 
  mutate(correct = case_when(
    question == "Electrical energy units (kWh)" & response == 2 ~ "Correct",
    question == "Energy consumed and appliance power rating" & response == 3 ~ "Correct",
    question == "Incandescent lightbulb conversion" & response == 3 ~ "Correct",
    question == "Reason to buy energy star appliances" & response == 3 ~ "Correct",
    question == "Which appliances uses the most energy" & response == 3 ~ "Correct",
    question == "Which appliance uses the most electricity" & response == 2 ~ "Correct",
    question == "Which source provides most electricity in the US" & response == 3 ~ "Correct",
    question == "Problem with electric cars" & response == 1 ~ "Correct",
    TRUE ~ "Incorrect"
  ))

els_id <- els2 |> group_by(id,state1,state2,refClass) |> 
    summarize(els=mean(correct=="Correct"), .groups="drop")

```

```{r}

ggplot(els2, aes(x = as.factor(response), fill = correct)) +
  geom_bar(position = "dodge") +
  facet_wrap(~question, ncol = 2, scales = "free_x") + # Facet by question, 2 columns, free x-axis scales
  scale_fill_manual(values = c("Correct" = "#00ba38", "Incorrect" = "#f8766d")) + # Custom colors
  labs(
    title = "Distribution of Answers to Energy Literacy Survey Questions",
    subtitle = "DeWaters & Powers (2011)",
    x = "Response Option",
    y = "Number of Respondents",
    fill = "Correctness"
  ) +
  theme_bw() + # Use a clean theme
  theme(
    strip.text = element_text(size = 10, face = "bold"), # Adjust facet label appearance
    axis.text.x = element_text(size = 8, angle = 45, hjust = 1), # Rotate x-axis labels for readability
    axis.text.y = element_text(size = 8),
    axis.title = element_text(size = 10),
    plot.title = element_text(size = 12, face = "bold"),
    plot.subtitle = element_text(size = 10),
    legend.position = "bottom", # Move legend to the bottom
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9)
  )

# compare refClass averages, stat_halfeye
els2 |> 
  group_by(id,refClass) |> 
  summarise(score=sum(correct=="Correct")/n(), .groups="drop") |> 
  ggplot(aes(x=refClass, y=score, col=refClass)) +
  stat_summary(fun.data = "mean_cl_boot", geom = "pointrange") +
    labs(title="Energy Literacy Survey Score", x="Reference Class", y="Score")

els2 |>
  group_by(refClass, question) |>
  summarize(mean_correct = mean(correct == "Correct", na.rm = TRUE)) |>
  ggplot(aes(x = question, y = mean_correct, fill = refClass)) +
  geom_col(position = "dodge") +
  labs(title = "Mean Correctness by Reference Class and Question",
       x = "Question",
       y = "Proportion Correct",
       fill = "Reference Class") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```

### 2024-12-11 nb/study1.qmd
Analysis on energy literacy

@@
 
+# Energy Literacy Score (ELS)
+
+els <- s1 |> select(id,id2,calc,refClass,state1,state2, ELS01:ELS08)
+
+els2 <- els |> 
+  pivot_longer(cols = ELS01:ELS08, names_to = "question", values_to = "response") |> 
+  mutate(response = as.numeric(response)) |>
+  mutate(question = case_when(
+    question == "ELS01" ~ "Electrical energy units (kWh)",
+    question == "ELS02" ~ "Energy consumed and appliance power rating",
+    question == "ELS03" ~ "Incandescent lightbulb conversion",
+    question == "ELS04" ~ "Reason to buy energy star appliances",
+    question == "ELS05" ~ "Which appliances uses the most energy",
+    question == "ELS06" ~ "Which appliance uses the most electricity",
+    question == "ELS07" ~ "Which source provides most electricity in the US",
+    question == "ELS08" ~ "Problem with electric cars",
+    TRUE ~ NA_character_
+  )) |> 
+  mutate(correct = case_when(
+    question == "Electrical energy units (kWh)" & response == 2 ~ "Correct",
+    question == "Energy consumed and appliance power rating" & response == 3 ~ "Correct",
+    question == "Incandescent lightbulb conversion" & response == 3 ~ "Correct",
+    question == "Reason to buy energy star appliances" & response == 3 ~ "Correct",
+    question == "Which appliances uses the most energy" & response == 3 ~ "Correct",
+    question == "Which appliance uses the most electricity" & response == 2 ~ "Correct",
+    question == "Which source provides most electricity in the US" & response == 3 ~ "Correct",
+    question == "Problem with electric cars" & response == 1 ~ "Correct",
+    TRUE ~ "Incorrect"
+  ))
+
+els_id <- els2 |> group_by(id,id2,state1,state2,refClass) |> 
+    summarize(els=mean(correct=="Correct"), .groups="drop")
+
 s1 <- s1 |> rename(edu=DEM04) |>
@@
       values_to = "value"
-    ) |> select(id,state1,state2,refClass,calc,edu,modality,state,item_num,value,id2,Duration__in_seconds_)
+    ) |> select(id,id2,state1,state2,refClass,calc,edu,modality,state,item_num,value,Duration__in_seconds_)
 
@@
 
+#s1$change==s1_old$change
+#s1$change==s1_new$change
+#head(s1_new |> filter(appliance=="Cooling",block==1,plan=="plan1"),n=16)
+
+# join with els_id, by id,id2,state1,state2,refClass
+s1 <- s1 |> left_join(els_id, by=c("id","id2","state1","state2","refClass")) |> relocate(els,.after=appliance)
+
 #saveRDS(s1, here::here("data/s1_processed.rds"))
@@
     filter(appliance !="Total kWh") |> 
-  group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
+    group_by(id,refClass,state,block,plan,edu,pct_goal,calc,els) |> 
     summarise(total_kWh = sum(value),orig_kWh=sum(family), 
-            pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
+                pct_change = abs(round((orig_kWh-total_kWh)/orig_kWh,3)), 
                 n_change = sum(value!=family),
@@
                 abs_error = abs(error),
-            close_match = abs_error <= 0.02)
+                log_abs_error=log(abs(error)+.007), 
+                close_match = abs_error <= 0.02) |>
+    ungroup() |> # Add ungroup here
+        mutate(
+            accuracy_level = factor(
+                case_when(
+                    abs_error == 0.00 ~ "Exact match",
+                    abs_error <= 0.02 ~ "0.01-2% error",
+                    abs_error <= 0.15 ~ "2.01-15% error",
+                    TRUE ~ "Over 15% error"  # Capture all remaining cases
+                ), 
+                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+                ordered = TRUE
+            )
+        ) |> relocate(accuracy_level, .after= "pct_change")
+
+s1_agg4 <- s1_agg |> group_by(id,refClass,calc,els) |> 
+    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+    summarise(
+    mg=sum(matched_goal),
+    mgc=sum(close_match),
+    n=n(), 
+    pct=mg/n,
+    pct_close=mgc/n,
+    mean_pct_change=mean(pct_change),
+    mean_abs_error=mean(abs_error),
+    mean_log_abs_error=mean(log_abs_error),
+    n_accuracy=first(n_accuracy)) |> 
+    mutate(accuracy_level = factor(
+            case_when(
+                mean_abs_error < 0.02 ~ "Exact match",
+                mean_abs_error <= 0.02 ~ "0.01-2% error",
+                mean_abs_error <= 0.15 ~ "2.01-15% error",
+                TRUE ~ "Over 15% error"  # Capture all remaining cases
+            ), 
+            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+            ordered = TRUE
+        ))
 
@@
 
-s1_agg4 <- s1_agg |> group_by(id,refClass) |> 
-  summarise(mg=sum(matched_goal),n=n(), pct=mg/n)
+# s1_agg4 <- s1_agg |> group_by(id,refClass,els) |> 
+#   summarise(mg=sum(matched_goal),n=n(), pct=mg/n)
 

### 2024-12-11 nb/study2.qmd
Analysis on energy literacy

@@
   ))) |> relocate(edu, .after = id)
+
+els <- s2 |> select(id,calc,refClass,state1,state2, ELS01:ELS08)
+
+els2 <- els |> 
+  pivot_longer(cols = ELS01:ELS08, names_to = "question", values_to = "response") |> 
+  # Convert response to numeric to avoid label issues
+  mutate(response = as.numeric(response)) |>
+  mutate(question = case_when(
+    question == "ELS01" ~ "Electrical energy units (kWh)",
+    question == "ELS02" ~ "Energy consumed and appliance power rating",
+    question == "ELS03" ~ "Incandescent lightbulb conversion",
+    question == "ELS04" ~ "Reason to buy energy star appliances",
+    question == "ELS05" ~ "Which appliances uses the most energy",
+    question == "ELS06" ~ "Which appliance uses the most electricity",
+    question == "ELS07" ~ "Which source provides most electricity in the US",
+    question == "ELS08" ~ "Problem with electric cars",
+    TRUE ~ NA_character_
+  )) |> 
+  mutate(correct = case_when(
+    question == "Electrical energy units (kWh)" & response == 2 ~ "Correct",
+    question == "Energy consumed and appliance power rating" & response == 3 ~ "Correct",
+    question == "Incandescent lightbulb conversion" & response == 3 ~ "Correct",
+    question == "Reason to buy energy star appliances" & response == 3 ~ "Correct",
+    question == "Which appliances uses the most energy" & response == 3 ~ "Correct",
+    question == "Which appliance uses the most electricity" & response == 2 ~ "Correct",
+    question == "Which source provides most electricity in the US" & response == 3 ~ "Correct",
+    question == "Problem with electric cars" & response == 1 ~ "Correct",
+    TRUE ~ "Incorrect"
+  ))
+
+els_id <- els2 |> group_by(id,calc,refClass) |> 
+    summarize(els=mean(correct=="Correct"), .groups="drop")
+
 ```
@@
 
-#write.csv(s2_long,"data/study_2_tg.csv")
-#saveRDS(s2_long, here::here("data/s2_processed.rds"))
+# join with els
+s2_long <- s2_long |> left_join(els_id,by=c("id","refClass","calc")) |> 
+  mutate(els=ifelse(is.na(els),0,els)) |> 
+  relocate(els,.after=edu)
+
+# write.csv(s2_long,"data/study_2_tg.csv")
+# saveRDS(s2_long, here::here("data/s2_processed.rds"))
 

### 2024-12-17 manuscript/intro.qmd
some minor updates

# Planning to Save Energy: How Information Format Affects Accuracy

## Abstract

Effective communication of energy consumption information is crucial for promoting residential energy conservation. This study investigates how different numerical representations of energy reduction goals influence consumers' ability to create accurate conservation plans. Across two experiments, we examined the impact of presenting energy information in kilowatt-hours (kWh), percentages, or U.S. dollars (USD) on planning accuracy. Participants completed a simulated household planning task in which they allocated energy usage across multiple appliances, with the goal presented in either kilowatt-hours (kWh), percentages, or monetary costs. Results across both experiments showed that presenting reduction goals in absolute units (kWh) led to significantly greater accuracy compared to percentage-based or monetary formats. Furthermore, we found that higher energy literacy was associated with more accurate planning. These findings demonstrate that absolute units (kWh) are more effective for communicating energy-saving goals, and highlight the potential value of educational interventions to improve consumer energy literacy.

## Introduction

The need to promote sustainable household energy consumption has become increasingly urgent as the demand for energy continues to rise, placing both financial and environmental burdens on individuals and society (DeWaters & Powers, 2011, Memmott et al., 2021).  Interventions aimed at changing households’ energy behaviors have often focused on providing information, but research suggests that simply offering more or different content is not always sufficient (Abrahamse et al., 2005; Fischer, 2008). Instead, it may be necessary to consider the way that energy information is presented, as the chosen format could influence whether individuals can effectively use it to plan for energy reductions (Canfield et al., 2017; Herrmann et al., 2018).

Research in numerical cognition demonstrates that format significantly affects how individuals process information. Presenting numerical values without clear reference points can lead to misinterpretations (Gigerenzer & Edwards, 2003; Reimer et al., 2015). While seemingly straightforward, formats like percentages often require additional mental processing, making comparisons less intuitive. Conversely, absolute units, such as kilowatt-hours (kWh), offer a concrete reference point for energy consumption. Providing information in absolute units may better align with natural cognitive tendencies to think in concrete frequencies or units (Hoffrage et al., 2000), reducing cognitive load and increasing planning precision.

**Procedural Knowledge and Action Planning**

Another related challenge is bridging the gap between general awareness and the specific steps needed to reduce energy use. Even individuals who wish to save energy may struggle to translate high-level goals into detailed action plans, a process requiring what can be viewed as procedural knowledge: the skills to develop and implement effective strategies (Abrahamse et al., 2005; Attari & Rajagopal, 2015; von Grabe, 2020). To create meaningful conservation efforts, individuals must be able to use the provided data—presented in certain units or formats—to identify where to focus, assess trade-offs, and devise actionable steps. Format matters in this context because it can facilitate or hinder the calculations and estimations needed to allocate time, money, and effort toward efficient energy use (Canfield et al., 2017; Herberz et al., 2020; Schwartz et al., 2015).

### Hypotheses

Building on prior work on how numerical formats influence understanding and action (Canfield et al., 2017; Gigerenzer & Hoffrage, 1995; Reimer et al., 2015), the present study examines how the choice of representing energy data—specifically in kWh, percentages, or USD—affects the accuracy of people’s energy reduction plans. We focus on whether certain formats make it easier to identify targeted actions and thereby improve planning precision.  We hypothesize that presenting information in absolute units (kWh) will lead to more accurate household energy conservation planning.

## Methods

### Participants

We implemented our task and surveys on Qualtrics, and recruited participants through Amazon Mechanical Turk. In Experiment 1, 252 participants were initially recruited, but data from 17 participants were corrupted due to experimenter error, leaving a final sample of 235 participants. Most participants (76%) reported using a calculator to complete the task.

### Materials and Design

The study employed a mixed design with reference class (kWh, percentage, USD) as a between-subjects factor and state/family scenario as a within-subjects factor. Each participant completed energy reduction planning tasks for two different states, with state order counterbalanced across participants. The family scenarios featured four households in different climate regions: Texas (Smith family) and California (Adams family) representing warm climates, and Colorado (Wells family) and Massachusetts (Davis family) representing cold climates. We obtain average utility use from each state by CITE SOURCE FOR STATE AVGS?

### Procedure

Participants received energy usage data for two hypothetical families and were tasked with creating action plans to meet specified reduction goals by allocating usage across five appliance categories:  heating, cooling, water heating, refrigerator, and and other appliances (e.g., TV, lighting).

For each family scenario, the participants were shown a table containing the families utility usage from the prior year, alongside the state averages for each appliance category (both prior year usage and stage averages are always shown in kWh). For each scenario, participants were asked to create two possible action plans to achieve the target reduction in total household energy usage (see fig-task). Depending on their reference class condition, the target reduction amount presented either in kilowatt-hours (kWh), as percentages of total household usage, or in U.S. dollars. In all conditions, the target reduction was equivalent to a 15% reduction in total household kWh.

### 2024-12-17 manuscript/manuscript.qmd
some minor updates

@@
 abstract: |
-    This study aims to examine how the format of energy information impacts individuals’ ability to develop precise energy reduction plans.
-By manipulating the reference class (kWh, %, USD) and assessing planning accuracy, we seek to determine which format facilitates better comprehension and decision-making.Across two experiments, the kWh format generally led to better accuracy, while the USD format consistently led to the worst performance.
-These findings highlight the importance of effective information presentation to promote energy conservation, and may contribute to the development of more effective energy communication strategies that can enhance conservation efforts.
+    Effective communication of energy consumption information is crucial for promoting residential energy conservation.
+This study investigates how different numerical representations of energy reduction goals influence consumers' ability to create accurate conservation plans.
+Across two experiments, we examined the impact of presenting energy information in kilowatt-hours (kWh), percentages, or U.S.
+dollars (USD) on planning accuracy.
+Participants completed a simulated household planning task in which they allocated energy usage across multiple appliances, with the goal presented in either kilowatt-hours (kWh), percentages, or monetary costs.
+Results across both experiments showed that presenting reduction goals in absolute units (kWh) led to significantly greater accuracy compared to percentage-based or monetary formats.
+Furthermore, we found that higher energy literacy was associated with more accurate planning.
+These findings demonstrate that absolute units (kWh) are more effective for communicating energy-saving goals, and highlight the potential value of educational interventions to improve consumer energy literacy.
 ---
@@
 
-Energy poverty continues to be a pervasive issue in the United States @memmottSociodemographicDisparitiesEnergy2021.
+Energy poverty continues to be a pervasive issue in the United States [@memmottSociodemographicDisparitiesEnergy2021].
 This challenge partly arises from difficulties in converting information across numerical formats, impeding the development of precise energy reduction plans @reimerNumericCommunicationRisk2015.
@@
 
-The way numerical information is presented can significantly affect how individuals process and use that information [@reimerNumericCommunicationRisk2015] .
+The way numerical information is presented can significantly affect how individuals process and use that information [@reimerNumericCommunicationRisk2015].
 The reference class problem highlights that numbers without clear reference points can lead to misinterpretation, as the meaning of a statistic depends on the category or class it refers to [@gigerenzerSimpleToolsUnderstanding2003; @reimerNumericCommunicationRisk2015].
@@
 @tbl-s1-agg that participants in the kWh condition met the target goal 38% of the time, compared to 22% for the Percentage condition and 10% for the USD condition.
-Moreover, the kWh reference class exhibited smaller deviations from the target reduction, suggesting that participants performed more accurately when the goal was framed in kWh rather than percentages or USD.
+Moreover, the kWh reference class exhibited smaller deviations from the target reduction, suggesting that participants performed more accurately when the goal was framed in kWh rather than when percentages or USD.
 
@@
 
-Experiment 1 examined how different numerical representations of energy reduction goals influenced participants’ planning accuracy.
+Experiment 1 examined how different numerical representations of energy reduction goals affected participants’ planning accuracy.
 In line with our hypothesis that absolute units would yield better accuracy, the kWh condition supported significantly more precise energy reduction plans than did either the Percentage or USD conditions.
@@
 
-The experimental procedures in study 2 are quite similar to those in study 1, but we also included a rounding manipulation (rounded vs.
-not rounded), and a manipulation of the goal (10% reduction vs.
-15% rediction).
+The experimental procedures in Experiment 2 are quite similar to those in Experiment 1.
+Experiment 2 employed a 2 (task goal: 10% vs.
+15% reduction) x 2 (last year's usage: exact vs.
+rounded) within-subjects design, with a between-subjects manipulation of the reference class (USD vs.
+Percentage vs.
+kWh)..
 We recruited 206 participants from Amazon Mechanical Turk, but data from from 10 participants were corrupted due to experimenter error, leaving a final sample of 196 participants.
-
-Note that reference class remains a between-subjects variable, while percent goal, rounding, and state are within-subjects variables.
-In study 2, the new design is a 4 state temperature (2 warm vs.
-2 cold states) X 2 task goal (10% vs.
-15%) X 2 last year’s usage for the family and the state average (exact vs.
-rounded numbers) within X 3 task reference class (USD vs.
-Percentage vs.
-kWh) between.
 
@@
 A Bayesian linear regression model, analogous to the one used in Experiment 1 (log_abs_error ~ els + (1|id) + (1|state)), revealed a significant negative relationship between energy literacy scores and log-transformed absolute error (Estimate = -3.21, 95% CI: -3.89 to -2.52).
-This finding indicates that participants with higher energy literacy tended to produce more accurate plans, exhibiting smaller deviations from the target reduction goals.
+This finding indicates that participants with higher energy literacy scores tended to produce more accurate plans, exhibiting smaller deviations from the target reduction goals.
 The conditional effect plot (@fig-s2-els) visually confirms this relationship, showing a clear decreasing trend in log absolute error as energy literacy increases.
@@
 ## Experiment 2: Discussion
+
+Experiment 2 expanded on the findings of Experiment 1 by introducing additional manipulations of goal difficulty (10% vs.
+15% reduction) and numerical presentation (rounded vs.
+exact numbers).
+The results largely corroborated and extended those of Experiment 1, providing further evidence for the advantages of presenting energy reduction goals in absolute units (kWh).
 
@@
 
+Across two experiments, we consistently found that presenting energy reduction goals in absolute units (kWh) led to more accurate planning compared to percentage-based or monetary representations.
+This advantage persisted across variations in goal difficulty and numerical presentation, suggesting a robust effect of reference class on planning accuracy.
+
 Karjalainen 2011 - people prefer information about price [@karjalainenConsumerPreferencesFeedback2011]
+
+Mention @blaschBoundedlyRationalConsumers2019 - better long-term appliance selection with information presented in monetary terms.
 

### 2024-12-17 manuscript/test.qmd
some minor updates

@@
 
+This study aims to examine how the format of energy information impacts individuals’ ability to develop precise energy reduction plans.
+By manipulating the reference class (kWh, %, USD) and assessing planning accuracy, we seek to determine which format facilitates better comprehension and decision-making.Across two experiments, the kWh format generally led to better accuracy, while the USD format consistently led to the worst performance.
+These findings highlight the importance of effective information presentation to promote energy conservation, and may contribute to the development of more effective energy communication strategies that can enhance conservation efforts.
+
 # Introduction
@@
 - @canfieldPerceptionsElectricityuseCommunications2017's findings on preferences for historical use information and the impact of neighbor comparisons
-- The concept of cognitive fit posits that performance improves when the information presentation format aligns with the task requirements [@vesseyCognitiveFitTheoryBased1991]
+
 - alignment can reduce cognitive load and enhance accuracy in planning [@shahBarLineGraph2011]
@@
 Addressing this gap is crucial for developing effective interventions that promote energy conservation behaviors.
+
+@abrahamseReviewInterventionStudies2005
+
+@attariEnablingEnergyConservation2015
+
+@dewatersEnergyLiteracySecondary2011
+
+@gigerenzerSimpleToolsUnderstanding2003
+
+@larrickMPGIllusion2008
+
+@blaschBoundedlyRationalConsumers2019
+
+@fischerFeedbackHouseholdElectricity2008
+
+@schwartzWhatPeopleConsumption2015
+
+@vongrabeUsingInstanceBasedLearning2020
+
+@gigerenzerHowImproveBayesian1995
+
+@attariPublicPerceptionsEnergy2010
+
+@herrmannDoesDataVisualization2018
+
+@vandenbroekHeuristicsEnergyJudgement2019
+
+@herberzKiloWhatDefault2020
+
+@marghetisSimpleInterventionsCan2019
+
+@lagomarsinoNeedChargeRight2022
+
+@hahnelMentalAccountingMechanisms2020
+
+@delgadoOpportunitiesGreaterEnergy2018
+
+@schleyCognitiveAccessibilityJudgments2015
+
+@langeWorkEnvironmentalProtection2021
+
+@tiedeNumeracyNumericAttention2022
 

### 2024-12-17 misc/indiv.qmd
some minor updates

@@
 
+# Model the relationship between calculator use, reference class, and absolute error
+model_calc <- lm(log_abs_error ~ calc * refClass, data = s2_long)
+summary(model_calc)
+
+# Visualize the relationship
+ggplot(s2_long, aes(x = refClass, y = log(abs_error), fill = calc)) +
+  geom_boxplot() +
+  labs(title = "Planning Accuracy by Calculator Use and Reference Class")
+
 ```

### 2024-12-18 docs/misc/indiv.qmd
update - marginal effects

--- 2024-12-17 misc/indiv.qmd
@@
     filter(appliance !="Total kWh") |> 
-  group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
+    group_by(id,refClass,state,block,plan,edu,pct_goal,calc) |> 
     summarise(total_kWh = sum(value),orig_kWh=sum(family), 
-            pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
+                pct_change = abs(round((orig_kWh-total_kWh)/orig_kWh,3)), 
                 n_change = sum(value!=family),
@@
                 error = pct_change - pct_goal,
-            abs_error = abs(error))
+                abs_error = abs(error),
+                log_abs_error=log(abs(error)+.007), 
+                close_match = abs_error <= 0.02) |>
+    ungroup() |> # Add ungroup here
+        mutate(
+            accuracy_level = factor(
+                case_when(
+                    abs_error == 0.00 ~ "Exact match",
+                    abs_error <= 0.05 ~ "0.01-5% error",
+                    TRUE ~ "Over 5% error"  # Capture all remaining cases
+                ), 
+                levels = c("Exact match", "0.01-5% error", "Over 5% error"),
+                ordered = TRUE
+            )
+        ) |> relocate(accuracy_level, .after= "pct_change")
+
+s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
+    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+    summarise(
+    mg=sum(matched_goal),
+    mgc=sum(close_match),
+    n=n(), 
+    pct=mg/n,
+    pct_close=mgc/n,
+    mean_pct_change=mean(pct_change),
+    mean_abs_error=mean(abs_error),
+    mean_log_abs_error=mean(log_abs_error),
+    n_accuracy=first(n_accuracy)) |> 
+    mutate(accuracy_level = factor(
+            case_when(
+                mean_abs_error < 0.02 ~ "Exact match",
+                mean_abs_error <= 0.05 ~ "0.01-5% error",
+                TRUE ~ "Over 5% error"  # Capture all remaining cases
+            ), 
+            levels = c("Exact match", ".01-5% error", "Over 5% error"),
+            ordered = TRUE
+        ))
 ```
@@
 # Assuming you have a data frame 's2_long' with energy literacy scores ('els')
-# and a data frame 's2_agg1' with planning data
+# and a data frame 's2_agg' with planning data
 
@@
 # Merge energy literacy scores
-s2_long <- merge(s2_long, s2_agg1[, c("id", "els")], by = "id")
+s2_long <- merge(s2_long, s2_agg[, c("id", "els")], by = "id")
 
@@
 ```{r}
-s2_agg1 <- s2_long |> 
+
+s2_agg <- s2_long |> 
   filter(appliance != "TOTAL") |> 
-  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded,els) |> 
+  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
   summarise(
@@
                     abs_error == 0.00 ~ "Exact match",
-                    abs_error <= 0.02 ~ "0.01-2% error",
-                    abs_error <= 0.15 ~ "2.01-15% error",
-                    TRUE ~ "Over 15% error"  # Capture all remaining cases
+                    abs_error <= 0.05 ~ "0.01-5% error",
+                    TRUE ~ "Over 5% error"  # Capture all remaining cases
                 ), 
-                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+                levels = c("Exact match","0.01-5% error", "Over 5% error"),
                 ordered = TRUE
@@
 
-s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc,els) |> 
+s2_agg4 <- s2_agg |> group_by(id,refClass,calc) |> 
     mutate(n_accuracy = n_distinct(accuracy_level)) |> 
@@
                 mean_abs_error < 0.02 ~ "Exact match",
-                mean_abs_error <= 0.02 ~ "0.01-2% error",
-                mean_abs_error <= 0.15 ~ "2.01-15% error",
-                TRUE ~ "Over 15% error"  # Capture all remaining cases
+                mean_abs_error <= 0.05 ~ ".01-5% error",
+                TRUE ~ "Over 5% error"  # Capture all remaining cases
             ), 
-            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+            levels = c("Exact match", "01-5% error", "Over 5% error"),
             ordered = TRUE
@@
 # Example: Use k-means clustering on participant-level mean accuracy measures
-participant_summary <- s2_agg1 %>%
+participant_summary <- s2_agg %>%
   group_by(id) %>%
@@
 # Merge energy literacy scores
-s2_long <- s2_long |> left_join(s2_agg1, by = c("id", "refClass", "calc", "state", "pct", "pct_goal", "plan", "rounded", "els"))
+s2_long <- s2_long |> left_join(s2_agg, by = c("id", "refClass", "calc", "state", "pct", "pct_goal", "plan", "rounded", "els"))
 

### 2024-12-18 docs/misc/log_error.qmd
update - marginal effects

--- 2024-12-07 misc/log_error.qmd
@@
                     abs_error == 0.00 ~ "Exact match",
-                    abs_error <= 0.02 ~ "0.01-2% error",
-                    abs_error <= 0.15 ~ "2.01-15% error",
-                    TRUE ~ "Over 15% error"  # Capture all remaining cases
+                    abs_error <= 0.05 ~ "0.01-5% error",
+                    TRUE ~ "Over 5% error"  # Capture all remaining cases
                 ), 
-                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+                levels = c("Exact match", "0.01-5% error", "Over 5% error"),
                 ordered = TRUE
@@
         ) |> relocate(accuracy_level, .after= "pct_change")
+
+s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
+    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+    summarise(
+    mg=sum(matched_goal),
+    mgc=sum(close_match),
+    n=n(), 
+    pct=mg/n,
+    pct_close=mgc/n,
+    mean_pct_change=mean(pct_change),
+    mean_abs_error=mean(abs_error),
+    mean_log_abs_error=mean(log_abs_error),
+    n_accuracy=first(n_accuracy)) |> 
+    mutate(accuracy_level = factor(
+            case_when(
+                mean_abs_error < 0.02 ~ "Exact match",
+                mean_abs_error <= 0.05 ~ "0.01-5% error",
+                TRUE ~ "Over 5% error"  # Capture all remaining cases
+            ), 
+            levels = c("Exact match", ".01-5% error", "Over 5% error"),
+            ordered = TRUE
+        ))
+
+s2_agg <- s2_long |> 
+  filter(appliance != "TOTAL") |> 
+  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
+  summarise(
+    total_kWh = sum(value),
+    orig_kWh = sum(family),
+    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
+    state_dif = mean(state_dif),
+    .groups = "drop"
+  ) |>
+  mutate(
+    matched_goal = (pct_change == pct),
+    close_match = abs(pct_change - pct) <= 0.02,
+                error = pct_change - pct,
+                abs_error = abs(error),
+                log_abs_error=log(abs(error)+.007)) |> 
+      ungroup() |> 
+        mutate(
+            accuracy_level = factor(
+                case_when(
+                    abs_error == 0.00 ~ "Exact match",
+                    abs_error <= 0.05 ~ "0.01-5% error",
+                    TRUE ~ "Over 5% error"  # Capture all remaining cases
+                ), 
+                levels = c("Exact match","0.01-5% error", "Over 5% error"),
+                ordered = TRUE
+            )
+        )
+
+s2_agg4 <- s2_agg |> group_by(id,refClass,calc) |> 
+    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+    summarise(
+    mg=sum(matched_goal),
+    mgc=sum(close_match),
+    n=n(), 
+    pct=mg/n,
+    pct_close=mgc/n,
+    mean_pct_change=mean(pct_change),
+    mean_abs_error=mean(abs_error),
+    mean_log_abs_error=mean(log_abs_error),
+    n_accuracy=first(n_accuracy)) |> 
+    mutate(accuracy_level = factor(
+            case_when(
+                mean_abs_error < 0.02 ~ "Exact match",
+                mean_abs_error <= 0.05 ~ ".01-5% error",
+                TRUE ~ "Over 5% error"  # Capture all remaining cases
+            ), 
+            levels = c("Exact match", "01-5% error", "Over 5% error"),
+            ordered = TRUE
+        ))
 
@@
 ```
+
+# S2
+
+```{r}
+
+log_error_model_s2_c1 <- brm(
+  log_abs_error ~ refClass*calc + (1|id) + (1|state),
+  data = s2_agg,
+  family = gaussian(link = "identity"),
+  prior = c(prior(normal(0, 3), class = "b"),
+            prior(cauchy(0, 1), class = "sd")),
+  iter = 2000, warmup = 1000, chains = 4, cores = 4, 
+  file = paste0(here::here("data/model_cache",'log_error_model_s2_c1.rds')) 
+)
+
+summary(log_error_model_s2_c1)
+conditional_effects(log_error_model_s2_c1)
+
+log_error_model_s2_c1rf <- brm(
+  log_abs_error ~ refClass + (1|id) + (1|state) + (1|calc),
+  data = s2_agg,
+  family = gaussian(link = "identity"),
+  prior = c(prior(normal(0, 3), class = "b"),
+            prior(cauchy(0, 1), class = "sd")),
+  iter = 2000, warmup = 1000, chains = 4, cores = 4, 
+  file = paste0(here::here("data/model_cache",'log_error_model_s2_c1rf.rds')) 
+)
+
+summary(log_error_model_s2_c1rf)
+
+log_error_model_s2_cAdd_int3 <- brm(
+  log_abs_error ~ (refClass*rounded*pct_goal)+calc + (1|id) + (1|state),
+  data = s2_agg,
+  family = gaussian(link = "identity"),
+  prior = c(prior(normal(0, 3), class = "b"),
+            prior(cauchy(0, 1), class = "sd")),
+  iter = 3000, warmup = 1000, chains = 4, cores = 4, 
+  file = paste0(here::here("data/model_cache",'log_error_model_s2_cAdd_int3.rds')) 
+)
+
+summary(log_error_model_s2_cAdd_int3)
+conditional_effects(log_error_model_s2_cAdd_int3)
+
+log_error_model_s2_c1rf_int3 <- brm(
+  log_abs_error ~ refClass*rounded*pct_goal + (1|id) + (1|state) + (1|calc),
+  data = s2_agg,
+  family = gaussian(link = "identity"),
+  prior = c(prior(normal(0, 3), class = "b"),
+            prior(cauchy(0, 1), class = "sd")),
+  iter = 3000, warmup = 1000, chains = 4, cores = 4, 
+  file = paste0(here::here("data/model_cache",'log_error_model_s2_c1rf_int3.rds')) 
+)
+
+summary(log_error_model_s2_c1rf_int3)
+
+conditional_effects(log_error_model_s2_c1rf_int3)
+
+log_error_model_s2_c1rf_add3 <- brm(
+  log_abs_error ~ refClass+rounded+pct_goal + (1|id) + (1|state) + (1|calc),
+  data = s2_agg,
+  family = gaussian(link = "identity"),
+  prior = c(prior(normal(0, 3), class = "b"),
+            prior(cauchy(0, 1), class = "sd")),
+  iter = 3000, warmup = 1000, chains = 4, cores = 4, 
+  file = paste0(here::here("data/model_cache",'log_error_model_s2_c1rf_add3.rds')) 
+)
+
+summary(log_error_model_s2_c1rf_add3)
+conditional_effects(log_error_model_s2_c1rf_add3)
+
+```
+
+```{r}
+
+```

### 2024-12-18 manuscript/appendix.qmd
some minor updates
update - marginal effects

@@
                 mean_abs_error < 0.02 ~ "Exact match",
-                mean_abs_error <= 0.02 ~ "0.01-2% error",
-                mean_abs_error <= 0.15 ~ "2.01-15% error",
-                TRUE ~ "Over 15% error"  # Capture all remaining cases
+                mean_abs_error <= 0.05 ~ "0.01-5% error",
+                TRUE ~ "Over 5% error"  # Capture all remaining cases
             ), 
-            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+            levels = c("Exact match", ".01-5% error", "Over 5% error"),
             ordered = TRUE
@@
 
-s2_agg1 <- s2_long |> 
+s2_agg <- s2_long |> 
   filter(appliance != "TOTAL") |> 
@@
   ) |>
-  mutate(matched_goal = (pct_change == pct),
+  mutate(
+    matched_goal = (pct_change == pct),
+    close_match = abs(pct_change - pct) <= 0.02,
                 error = pct_change - pct,
@@
                 log_abs_error=log(abs(error)+.007)) |> 
-      ungroup() |> # Add ungroup here
+      ungroup() |> 
         mutate(
@@
 
-s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc) |> 
+s2_agg4 <- s2_agg |> group_by(id,refClass,calc) |> 
     mutate(n_accuracy = n_distinct(accuracy_level)) |> 
-    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),
-    mean_abs_error=mean(abs_error),mean_log_abs_error=mean(log_abs_error),
+    summarise(
+    mg=sum(matched_goal),
+    mgc=sum(close_match),
+    n=n(), 
+    pct=mg/n,
+    pct_close=mgc/n,
+    mean_pct_change=mean(pct_change),
+    mean_abs_error=mean(abs_error),
+    mean_log_abs_error=mean(log_abs_error),
     n_accuracy=first(n_accuracy)) |> 
@@
 ## Study2
+
+```{r}
+ordinal_model_s2_logit <- brm(
+  accuracy_level ~ refClass + calc+pct_goal+rounded + (1|id)+ (1|state),
+  data = s2_agg,
+  family = cumulative("logit"),
+  cores = 4,
+  iter = 3000,
+  control = list(adapt_delta = 0.99), # Recommended for ordinal models
+  prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
+            prior(normal(0, 2), class = "b")), # Priors for predictors
+  file = paste0(here::here("data/model_cache",'s2_acc3_add.rds')) # Cache for efficiency
+)
+```
+
+# diagnostics
+```{r}
+plot(ordinal_model_s2_logit, condition = "refClass")
+```
+
+### Tables
+
+```{r}
+#| label: tbl-s2-prop
+#| tbl-cap: "Study 2: The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages).
+The combined group column reflects the percentage of participants in each accuracy level when aggregating across across all reference class conditions."
+
+# observed_props_s2 <- s2_agg |>
+#   group_by(id,refClass, accuracy_level,calc) |>
+#   summarise(n = n()) |>
+#   group_by(refClass) |>
+#   mutate(prop = n/sum(n)) |>
+#   group_by(refClass,accuracy_level) |>
+#   summarise(n = sum(n), prop = sum(prop)) |>
+#   mutate(n_prop=paste0(n," (",round(prop*100,1),"%)" ), pct_grp=paste0(round(prop*100,1), "%")) |> ungroup()
+#
+
+# compute percentage of subjects per accuracy level per group
+observed_props_s2 <- s2_agg |>
+  group_by(refClass, accuracy_level) |>
+  summarise(n = n()) |>
+  group_by(refClass) |>
+  mutate(prop = n/sum(n)) |>
+  mutate(n_prop=paste0(n," (",round(prop*100,1),"%)" ), pct_grp=paste0(round(prop*100,1), "%")) |> ungroup()
+
+observed_props_s2 |> 
+  mutate(n_total=sum(n)/3) |> 
+  group_by(accuracy_level) |>
+  mutate(ns=sum(n)/4) |> 
+  mutate(Total = paste0(round(ns/n_total*100,1), "%")) |>
+  select('Reference Class'=refClass, 'Accuracy Level'=accuracy_level, '% in Group'=pct_grp, "Combined Groups %" =Total) |>
+  pivot_wider(
+    names_from = 'Reference Class',
+    values_from = c('% in Group')
+  ) |> relocate("Combined Groups %" , .after=last_col()) |> 
+  kable(escape=FALSE,booktabs=TRUE,align=c("l"))
+
+```
+
+```{r}
+#| label: tbl-s2-prop
+#| tbl-cap: "Study 2: The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages).
+The combined group column reflects the percentage of participants in each accuracy level when aggregating across across all reference class conditions."
+
+# compute percentage of subjects per accuracy level per group
+observed_props_s2 <- s2_agg |>
+  group_by(refClass, accuracy_level) |>
+  summarise(n = n()) |>
+  group_by(refClass) |>
+  mutate(prop = n/sum(n)) |>
+  mutate(n_prop=paste0(n," (",round(prop*100,1),"%)" ), pct_grp=paste0(round(prop*100,1), "%")) |> ungroup()
+
+observed_props_s2 |> 
+  mutate(n_total=sum(n)/3) |> 
+  group_by(accuracy_level) |>
+  mutate(ns=sum(n)/4) |> 
+  mutate(Total = paste0(round(ns/n_total*100,1), "%")) |>
+  select('Reference Class'=refClass, 'Accuracy Level'=accuracy_level, '% in Group'=pct_grp, "Combined Groups %" =Total) |>
+  pivot_wider(
+    names_from = 'Reference Class',
+    values_from = c('% in Group')
+  ) |> relocate("Combined Groups %" , .after=last_col()) |> 
+  kable(escape=FALSE,booktabs=TRUE,align=c("l"))
+
+```
+
+@tbl-s2-prop shows that, once again, participants in the kWh condition achieved closer alignment with the target goals (44% exact matches), followed by Percentage (27%) and USD (18%).
+These percentages are consistent with the patterns observed in Study 1, reinforcing the conclusion that providing goals in kWh supports better accuracy.
+
+```{r}
+#| label: fig-s1-plot
+#| fig-cap: "Study 1: Proportion of participants in each accuracy level, colored by reference class. A larger % of participants in the Exact Match bin indicates better performance."
+
+# prop_acc_s1 <- s1_agg %>%
+#     group_by(refClass, accuracy_level) %>%
+#     summarise(count = n()) %>%
+#     group_by(accuracy_level) |>
+#     mutate(Probability = count / sum(count)) %>%
+#     ungroup()
+
+# ggplot(prop_acc_s1, aes(x = accuracy_level, y = Probability, fill = refClass)) +
+#     geom_bar(stat = "identity", position = position_dodge()) +
+#   scale_y_continuous(labels = scales::percent) +
+#     labs(title = "S1. % of Participants within each Accuracy Bin",
+#         x = "Accuracy Level",
+#         y = "Percentage of Participants",
+#         fill = "Reference Class") +
+#     theme_minimal()
+
+ggplot(data = s1_agg, aes(x = refClass, fill = accuracy_level)) +
+  geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
+  scale_y_continuous(labels = scales::percent) +
+  labs(title = "Accuracy Levels by Reference Class",
+       x = "Reference Class Condition",
+       y = "Percentage of Participants",
+       fill = "Accuracy Level") +
+  theme_minimal()
+
+```
+
+```{r}
+#| label: fig-s2-plot2
+#| fig-cap: "Experiment 2.
+Proportion of participants in each accuracy level, colored by reference class, and seprated in facets based on the levels of reduction goal, and rounding.
+A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
+#| fig.width: 11
+#| fig.height: 9
+
+ggplot(data = s2_agg, aes(x = refClass, fill = accuracy_level)) +
+  geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
+  scale_y_continuous(labels = scales::percent) +
+  facet_wrap(~rounded+pct_goal) +
+  labs(title = "Accuracy Levels by Reference Class",
+       x = "Reference Class Condition",
+       y = "Percentage of Participants",
+       fill = "Accuracy Level") +
+  theme_minimal()
+
+```
+
+```{r}
+#| label: fig-s2-plot
+#| tbl-cap: "Study 2: Proportion of participants in each accuracy level, colored by reference class. A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
+
+# % of entire sample 
+prop_combo_s2 <- s2_agg %>%
+    group_by(refClass, accuracy_level,pct_goal) %>%
+    summarise(count = n()) %>%
+    group_by(refClass) %>%
+    mutate(Probability = count / sum(count)) %>%
+    ungroup()
+
+ggplot(prop_combo_s2, aes(x = accuracy_level, y = Probability, fill = refClass)) +
+    geom_bar(stat = "identity", position = position_dodge()) +
+  scale_y_continuous(labels = scales::percent) +
+    labs(title = "S1. % of Participants within each Accuracy Bin",
+        x = "Accuracy Level",
+        y = "Percentage of Participants",
+        fill = "Reference Class") +
+    theme_minimal()
+
+# ggplot(data = s2_agg, aes(x = refClass, fill = accuracy_level)) +
+#   geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
+#   scale_y_continuous(labels = scales::percent) +
+#   labs(title = "Accuracy Levels by Reference Class",
+#        x = "Reference Class Condition",
+#        y = "Percentage of Participants",
+#        fill = "Accuracy Level") +
+#   theme_minimal()
+
+```
+
+### Model Tables
+
+```{r}
+
+marginaleffects_by_pct <- avg_slopes(
+      ordinal_model_s2_logit,
+      variables = "refClass",
+      by = c("pct_goal", "accuracy_level")
+    )
+    print(marginaleffects_by_pct)
+
+ #         Group     Term         Contrast pct_goal  Estimate    2.5 %  97.5 %
+ # Exact match   refClass Percentage - kWh      10% -0.073089 -0.19345  0.0421
+ # Exact match   refClass USD - kWh             10% -0.150377 -0.26763 -0.0324
+ # 0.01-5% error refClass Percentage - kWh      10%  0.000166 -0.01468  0.0223
+ # 0.01-5% error refClass USD - kWh             10% -0.018067 -0.05721  0.0100
+ # Over 5% error refClass Percentage - kWh      10%  0.073590 -0.04623  0.1885
+ # Over 5% error refClass USD - kWh             10%  0.171389  0.03806  0.2953
+ # Exact match   refClass Percentage - kWh      15% -0.076254 -0.19838  0.0448
+ # Exact match   refClass USD - kWh             15% -0.158520 -0.28022 -0.0346
+ # 0.01-5% error refClass Percentage - kWh      15%  0.003111 -0.00833  0.0299
+ # 0.01-5% error refClass USD - kWh             15% -0.007757 -0.04633  0.0197
+ # Over 5% error refClass Percentage - kWh      15%  0.071884 -0.04499  0.1843
+ # Over 5% error refClass USD - kWh             15%  0.168805  0.03638  0.2940
+
+avg_predictions(ordinal_model_s2_logit, newdata = datagrid(id=unique,refClass=unique),by="refClass")
+
+ #         Group   refClass Estimate 2.5 % 97.5 %
+ # Exact match   kWh           0.420 0.336  0.513
+ # Exact match   Percentage    0.340 0.266  0.427
+ # Exact match   USD           0.254 0.179  0.347
+ # 0.01-5% error kWh           0.209 0.173  0.244
+ # 0.01-5% error Percentage    0.213 0.182  0.247
+ # 0.01-5% error USD           0.199 0.160  0.233
+ # Over 5% error kWh           0.368 0.291  0.457
+ # Over 5% error Percentage    0.444 0.360  0.532
+ # Over 5% error USD           0.544 0.446  0.648    
+ #
+
+ame2 <- avg_slopes(
+  ordinal_model_s2_logit, 
+  variables = "refClass"
+)
+ame2 |>  kable(escape=FALSE,booktabs=TRUE,align=c("l"))
+
+ #         Group         Contrast  Estimate   2.5 %   97.5 %
+ # Exact match   Percentage - kWh -0.072820 -0.1920  0.04245
+ # Exact match   USD - kWh        -0.150425 -0.2668 -0.03252
+ # 0.01-5% error Percentage - kWh  0.000449 -0.0127  0.02116
+ # 0.01-5% error USD - kWh        -0.015279 -0.0524  0.00957
+ # Over 5% error Percentage - kWh  0.072572 -0.0453  0.18601
+ # Over 5% error USD - kWh         0.168592  0.0373  0.29260
+```
+
+```{r}
+
+me <- conditional_effects(ordinal_model_s2_logit, effects = "refClass",categorical=TRUE)
+plot(me, points = TRUE)
+```
 
@@
 
+### Individual Differences
+
+As in Experiment 1, we further investigated the role of individual differences in energy literacy in predicting planning accuracy.
+A Bayesian linear regression model, analogous to the one used in Experiment 1 (log_abs_error ~ els + (1|id) + (1|state)), revealed a significant negative relationship between energy literacy scores and log-transformed absolute error (Estimate = -3.21, 95% CI: -3.89 to -2.52).
+This finding indicates that participants with higher energy literacy scores tended to produce more accurate plans, exhibiting smaller deviations from the target reduction goals.
+The conditional effect plot (@fig-s2-els) visually confirms this relationship, showing a clear decreasing trend in log absolute error as energy literacy increases.
+
+```{r}
+#| label: fig-s2-indv
+#| fig-cap: "Study 2: Individual performance in the energy planning task, colored by reference class.
+The dashed line represents the target reduction goal.
+Participants are shown along the y axis, those who fall above or below the dashed line have not met the target goal.
+The x-axis represents the percent change in energy usage from the prior year."
+#| fig-width: 11
+#| fig-height: 9
+
+s2_agg |> group_by(id,refClass,calc,pct_goal,pct_change) |> 
+    filter(plan=="plan1",rounded=="Rounded") |> 
+    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),mean_abs_error=mean(abs_error),n_accuracy=first(n_accuracy)) |>
+  ungroup() |> 
+  mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100) |>
+  filter(mean_abs_error <= 0.50) |>
+  mutate(id=reorder(id,pct_change)) |> 
+  ggplot(aes(y=id,x=mean_pct_change,col=refClass)) + 
+  geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
+  geom_vline(aes(xintercept=goal_pct),linetype="dashed",alpha=.5) +
+  ggh4x::facet_nested_wrap(~pct_goal,axes="all",scales="free",ncol=2)  + 
+  labs(y="Participant Id", x="Percent Change", title="Individual Performance") +
+  theme(axis.text.y=element_text(face = "plain", size = rel(0.7))) + 
+    scale_x_continuous(breaks = seq(0, 0.5, by = 0.05),
+                    labels = scales::percent_format(accuracy = 1))
+
+```
+
+```{r}
+#| label: fig-s2-indv2
+#| fig-cap: "Study 2: Respones patterns for a subset of individiual participants.
+Black points are participant responses, red points are the state average, and blue points are the family average.
+The x-axis represents the appliance category, and the y-axis represents the energy usage in kWh."
+#| fig.width: 11
+#| fig.height: 10
+
+s2_long |> filter(id %in% unique(s2_long$id)[1:30]) |> 
+  filter(appliance!="TOTAL",state=="California") |> 
+  ggplot(aes(x=appliance,y=value)) + geom_point(aes(shape=plan)) +
+  geom_point(aes(y=state_avg),color="red") +
+  geom_point(aes(y=family),color="blue") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+s2_long |> filter(id %in% unique(s2_long$id)[1:30]) |> 
+  filter(appliance!="TOTAL",state=="California") |> 
+  ggplot(aes(x=appliance,y=value)) + geom_point(aes(shape=plan)) +
+  geom_point(aes(y=state_avg),color="red") +
+  geom_point(aes(y=family),color="blue") +
+  facet_wrap(~id) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
+
 ### Tables
@@
 ```
+
+Additional data collected included:
+
+-   **Energy Literacy Quiz**: An 8-item questionnaire assessing participants' knowledge of energy consumption and conversion [ @dewatersEnergyLiteracySecondary2011].
+-   **Calculator Usage Tracking**: Questions determined whether participants used a calculator, paper/pen, or other methods to complete the tasks.
+-   **Demographic Survey**: Collected information on gender, age, income, education, employment status, and state of residence.
+-   **Environmental Attitudes Survey**: Assessed participants' pro-environmental attitudes and perceived importance of energy conservation.

### 2024-12-18 manuscript/manuscript.qmd
update - marginal effects

@@
 -   **Calculator Usage Tracking**: Questions determined whether participants used a calculator, paper/pen, or other methods to complete the tasks.
--   **Demographic Survey**: Collected information on gender, age, income, education, employment status, and state of residence.
--   **Environmental Attitudes Survey**: Assessed participants' pro-environmental attitudes and perceived importance of energy conservation.
 
@@
 #| cache: false
-pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,
-    ggplot2,gt,knitr,kableExtra,ggh4x,patchwork, ggridges,ggstance,lme4,flextable,pander)
+pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,
+    ggplot2,gt,knitr,kableExtra,ggh4x,patchwork, ggridges,ggstance,lme4,flextable,pander,marginaleffects)
 
@@
 
-```{r}
-#| label: fig-s1-plot
-#| fig-cap: "Study 1: Proportion of participants in each accuracy level, colored by reference class. A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
-
-prop_acc_s1 <- s1_agg %>%
-    group_by(refClass, accuracy_level) %>%
-    summarise(count = n()) %>%
-    group_by(accuracy_level) |>
-    mutate(Probability = count / sum(count)) %>%
-    ungroup()
-
-ggplot(prop_acc_s1, aes(x = accuracy_level, y = Probability, fill = refClass)) +
-    geom_bar(stat = "identity", position = position_dodge()) +
-  scale_y_continuous(labels = scales::percent) +
-    labs(title = "S1. % of Participants within each Accuracy Bin",
-        x = "Accuracy Level",
-        y = "Percentage of Participants",
-        fill = "Reference Class") +
-    theme_minimal() 
-```
-
-We next categorized responses into four accuracy levels (exact match [0% error], minor deviations [0.01–2%], moderate deviations [2.01–15%], and major deviations [>15%]) for our primary statistical modeling.
+We next categorized responses into three accuracy levels (exact match [0% error], minor deviations [0.01–5%], and large deviations [>5%]) for our primary statistical modeling.
 Using Bayesian ordinal regression, we modeled the ordered accuracy outcome as a function of the reference class condition, while controlling for random variation across participants and family scenarios:
@@
     accuracy_level ~ refClass +calc + (1|id) + (1|state),
-    data = s1_agg_filtered,
+    data = s1_agg,
     family = cumulative("logit"),
     cores = 4,
-    iter = 2000,
-    control = list(adapt_delta = 0.97), 
-    prior = c(prior(normal(0, 3), class = "Intercept"), 
-                prior(normal(0, 3), class = "b")), 
-    file = paste0(here::here("data/model_cache",'s1_ordinal_grpclean_calc_2.5.rds')) 
+    iter = 4000,
+    control = list(adapt_delta = 0.98), 
+    prior = c(prior(normal(0, 4), class = "Intercept"), 
+                prior(normal(0, 4), class = "b")), 
+    file = paste0(here::here("data/model_cache",'s1_acc3_add.rds')) 
 )
 
-# as.data.frame(describe_posterior(ordinal_model_s1, centrality = "Mean"))[, c(1,2,4,5,6)] |> 
-#   setNames(c("Parameter", "Estimate", "CI_Lower", "CI_Upper", "pd")) |> 
-#   mutate(Parameter = stringr::str_remove(Parameter, "b_")) |> kable(escape=FALSE,booktabs=TRUE,align=c("l"), row.names = FALSE)
+t1 <- as.data.frame(describe_posterior(ordinal_model_s1, centrality = "Mean"))[, c(1,2,4,5,6)] |> 
+  setNames(c("Parameter", "Estimate", "CI_Lower", "CI_Upper", "pd")) |> 
+  mutate(Parameter = stringr::str_remove(Parameter, "b_")) |> kable(escape=FALSE,booktabs=TRUE,align=c("l"), row.names = FALSE)
 
@@
 
-# odds ratios of fixed effects
+#odds ratios of fixed effects
 # as.data.frame(fixef(ordinal_model_s1)[,-2])|> as.data.frame() %>%
@@
 |:------------------|:--------|:--------|:--------|:----|
-|Intercept[1]       |-3.8     |-5.45    |-2.28    |1.00 |
-|Intercept[2]       |-1.7     |-3.29    |-0.15    |0.98 |
-|Intercept[3]       |2.8      |1.27     |4.40     |1.00 |
-|refClassPercentage |1.3      |0.01     |2.66     |0.98 |
-|refClassUSD        |2.8      |1.52     |4.04     |1.00 |
-|calcUsedCalculator |-2.8     |-4.09    |-1.56    |1.00 |
+|Intercept[1]       |-4.21    |-5.90    |-2.58    |1.00 |
+|Intercept[2]       |-0.89    |-2.49    |0.71     |0.87 |
+|refClassPercentage |1.44     |0.07     |2.88     |0.98 |
+|refClassUSD        |3.13     |1.81     |4.50     |1.00 |
+|calcUsedCalculator |-3.30    |-4.80    |-1.92    |1.00 |
 
@@
 |:-----------------|:----------|:--------|:--------|
-|Percentage vs kWh |3.7        |1.0      |14       |
-|USD vs kWh        |15.7       |4.6      |57       |
+|Percentage vs kWh |4.2        |1.1      |18       |
+|USD vs kWh        |22.9       |6.1      |90       |
 
@@
 pp_check(ordinal_model_s1, type = "bars_grouped", group="refClass", fatten = 2) +
-  scale_x_continuous("Response Category", breaks = 1:4, 
-            labels = c("Exact", "0.01-2%", "2.01-15%", ">15%")) +
+  scale_x_continuous("Response Category", breaks = 1:3, 
+            labels = c("Exact", "0.01-5%", ">5%")) +
   scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
@@
 #| label: fig-s1-els
-
+#| fig-cap: "Experiment 1.
+Conditional effect of energy literacy on log absolute error.
+The plot shows the relationship between energy literacy score and log absolute error, controlling for random effects of participant and state.
+Higher energy literacy scores are associated with smaller deviations from the target reduction goal, indicating more accurate planning."
 s1_els_log_error <- brm(
@@
 ```{r}
-s2_agg1 <- s2_long |> 
+
+s2_agg <- s2_long |> 
   filter(appliance != "TOTAL") |> 
@@
 
-s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc) |> 
+s2_agg4 <- s2_agg |> group_by(id,refClass,calc) |> 
     mutate(n_accuracy = n_distinct(accuracy_level)) |> 
@@
 
-Study 2: Summary of planning accuracy by reference class.
+Experiment 2: Summary of planning accuracy by reference class.
 The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal
@@
 #| label: fig-s2-log-dist
-#| fig-cap: "Study 1: Distribution of log absolute error by reference class."
+#| fig-cap: "Study 1: Distribution of log absolute error by reference class.
+Displays the distributions of log absolute deviation from the target reduction goal across the three reference class conditions and is also faceted by calculator usage.
+Lower values indicate greater accuracy."
 #| fig-width: 11
@@
 
-s2_ld <- ggplot(s2_agg1, aes(y = refClass, x = log_abs_error, color = refClass)) +
+s2_ld <- ggplot(s2_agg, aes(y = refClass, x = log_abs_error, color = refClass)) +
   geom_density_ridges(aes(fill = refClass), alpha = 0.2, scale = 0.5,
@@
 
-s2_ldc <- ggplot(s2_agg1, aes(y = refClass, x = log_abs_error, color = calc)) +
+s2_ldc <- ggplot(s2_agg, aes(y = refClass, x = log_abs_error, color = calc)) +
   geom_density_ridges(aes(fill = calc), alpha = 0.2, scale = 0.5,
@@
 ```{r}
-#| label: tbl-s2-prop
-#| tbl-cap: "Study 2: The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages).
-The combined group column reflects the percentage of participants in each accuracy level when aggregating across across all reference class conditions."
-
-# compute percentage of subjects per accuracy level per group
-observed_props_s2 <- s2_agg1 |>
-  group_by(refClass, accuracy_level) |>
-  summarise(n = n()) |>
-  group_by(refClass) |>
-  mutate(prop = n/sum(n)) |>
-  mutate(n_prop=paste0(n," (",round(prop*100,1),"%)" ), pct_grp=paste0(round(prop*100,1), "%")) |> ungroup()
-
-observed_props_s2 |> 
-  mutate(n_total=sum(n)/4) |> 
-  group_by(accuracy_level) |>
-  mutate(ns=sum(n)/4) |> 
-  mutate(Total = paste0(round(ns/n_total*100,1), "%")) |>
-  select('Reference Class'=refClass, 'Accuracy Level'=accuracy_level, '% in Group'=pct_grp, "Combined Groups %" =Total) |>
-  pivot_wider(
-    names_from = 'Reference Class',
-    values_from = c('% in Group')
-  ) |> relocate("Combined Groups %" , .after=last_col()) |> 
-  kable(escape=FALSE,booktabs=TRUE,align=c("l"))
-
-```
-
-```{r}
-#| label: fig-s2-plot
-#| tbl-cap: "Study 2: Proportion of participants in each accuracy level, colored by reference class. A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
-
-# % of entire sample 
-prop_combo_s2 <- s2_agg1 %>%
-    group_by(refClass, accuracy_level,pct_goal) %>%
-    summarise(count = n()) %>%
-    group_by(refClass) %>%
-    mutate(Probability = count / sum(count)) %>%
-    ungroup()
-
-ggplot(prop_combo_s2, aes(x = accuracy_level, y = Probability, fill = refClass)) +
-    geom_bar(stat = "identity", position = position_dodge()) +
-  scale_y_continuous(labels = scales::percent) +
-    labs(title = "S1. % of Participants within each Accuracy Bin",
-        x = "Accuracy Level",
-        y = "Percentage of Participants",
-        fill = "Reference Class") +
-    theme_minimal()
-
-```
-
-@tbl-s2-prop shows that, once again, participants in the kWh condition achieved closer alignment with the target goals (44% exact matches), followed by Percentage (27%) and USD (18%).
-These percentages are consistent with the patterns observed in Study 1, reinforcing the conclusion that providing goals in kWh supports better accuracy.
-
-```{r}
-#| label: fig-s2-plot2
-#| fig-cap: "Study 2: Proportion of participants in each accuracy level, colored by reference class, and seprated in facets based on the levels of reduction goal, and rounding.
-A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
-#| fig.width: 11
-#| fig.height: 9
-
-# % of entire sample 
-prop_combo_s2_full <- s2_agg1 %>%
-    group_by(refClass, accuracy_level, pct_goal,rounded) %>%
-    summarise(count = n()) %>%
-    group_by(refClass) %>%
-    mutate(Probability = count / sum(count)) %>%
-    ungroup()
-
-ggplot(prop_combo_s2_full, aes(x = accuracy_level, y = Probability, fill = refClass)) +
-    geom_bar(stat = "identity", position = position_dodge()) +
-  scale_y_continuous(labels = scales::percent) +
-    labs(title = "S1. % of Participants within each Accuracy Bin",
-        x = "Accuracy Level",
-        y = "Percentage of Participants",
-        fill = "Reference Class") +
-    facet_wrap(~pct_goal*rounded) +
-    theme_minimal() 
-```
-
-```{r}
 #| label: tbl-s2-ord
@@
 ordinal_model_s2_logit <- brm(
-  accuracy_level ~ refClass +rounded+pct_goal + (1|id)+ (1|state),
-  data = s2_agg1,
+  accuracy_level ~ refClass + calc+pct_goal+rounded + (1|id)+ (1|state),
+  data = s2_agg,
   family = cumulative("logit"),
   cores = 4,
-  iter = 2000,
+  iter = 3000,
   control = list(adapt_delta = 0.99), # Recommended for ordinal models
@@
             prior(normal(0, 2), class = "b")), # Priors for predictors
-  file = paste0(here::here("data/model_cache",'s2_logit_add.rds')) # Cache for efficiency
+  file = paste0(here::here("data/model_cache",'s2_acc3_add.rds')) # Cache for efficiency
 )
@@
 
-# as.data.frame(describe_posterior(ordinal_model_s2_logit, centrality = "Mean"))[, c(1,2,4,5,6)] |> 
-#   setNames(c("Parameter", "Estimate", "CI_Lower", "CI_Upper", "pd")) |> 
-#   mutate(Parameter = stringr::str_remove(Parameter, "b_")) |> 
-#   kable(escape = FALSE, booktabs = TRUE, align = c("l"), row.names = FALSE)
+t2 <- as.data.frame(describe_posterior(ordinal_model_s2_logit, centrality = "Mean"))[, c(1,2,4,5,6)] |> 
+  setNames(c("Parameter", "Estimate", "CI_Lower", "CI_Upper", "pd")) |> 
+  mutate(Parameter = stringr::str_remove(Parameter, "b_")) |> 
+  kable(escape = FALSE, booktabs = TRUE, align = c("l"), row.names = FALSE)
 
-pred_summary_s2 <- ordinal_model_s2_logit %>%
-  epred_draws(newdata = s2_agg1, re_formula = NA,ndraws=200) %>%
- # group_by("Reference Class"=refClass, rounded, "% Goal"=pct_goal, Category=.category) %>%
-  group_by("Reference Class"=refClass, Category=.category) %>%
-  summarise(
-    mean_prob = mean(.epred),
-    lower_ci = quantile(.epred, 0.025),
-    upper_ci = quantile(.epred, 0.975),
-    .groups = "drop"
-  )
+# pred_summary_s2 <- ordinal_model_s2_logit %>%
+#   epred_draws(newdata = s2_agg, re_formula = NA,ndraws=200) %>%
+#  # group_by("Reference Class"=refClass, rounded, "% Goal"=pct_goal, Category=.category) %>%
+#   group_by("Reference Class"=refClass, Category=.category) %>%
+#   summarise(
+#     mean_prob = mean(.epred),
+#     lower_ci = quantile(.epred, 0.025),
+#     upper_ci = quantile(.epred, 0.975),
+#     .groups = "drop"
+#   )
 
@@
 
-as.data.frame(fixef(ordinal_model_s2_logit)[,-2])|> as.data.frame() %>%
+or2 <- as.data.frame(fixef(ordinal_model_s2_logit)[,-2])|> as.data.frame() %>%
     rownames_to_column(var = "Parameter") %>%
@@
 |:------------------|:--------|:--------|:--------|:----|
-|Intercept[1]       |-2.13    |-3.39    |-0.86    |1.00 |
-|Intercept[2]       |-0.62    |-1.89    |0.63     |0.84 |
-|Intercept[3]       |3.15     |1.88     |4.42     |1.00 |
-|refClassPercentage |0.83     |-0.64    |2.33     |0.87 |
-|refClassUSD        |1.87     |0.31     |3.35     |0.99 |
-|roundedRounded     |-0.66    |-1.01    |-0.31    |1.00 |
-|pct_goal15%        |-0.44    |-0.79    |-0.10    |0.99 |
+|Intercept[1]       |-1.45    |-2.85    |-0.07    |0.98 |
+|Intercept[2]       |1.26     |-0.09    |2.65     |0.97 |
+|refClassPercentage |1.02     |-0.63    |2.71     |0.89 |
+|refClassUSD        |2.27     |0.53     |3.98     |0.99 |
+|calcNoCalculator   |4.10     |2.20     |6.06     |1.00 |
+|pct_goal15%        |-0.39    |-0.81    |0.04     |0.96 |
+|roundedRounded     |-0.53    |-0.96    |-0.11    |0.99 |
 
-**Experiment 2.** Ordinal Regression Model Results.
+**Experiment 2.**  Parameter estimates from the ordinal regression model.
+Positive coefficients for refClass predictors indicate increased likelihood of falling into higher error categories relative to the kWh baseline.
 
@@
 ::: {#tbl-s2-ord}
-|comparison           | odds_ratio| ci_lower| ci_upper|
-|:--------------------|----------:|--------:|--------:|
-|Percentage vs kWh    |       3.02|     0.53|    10.31|
-|USD vs kWh           |       8.80|     1.37|    28.42|
-|Rounded vs Not       |       0.53|     0.36|     0.73|
-|15% Goal vs 10% Goal |       0.66|     0.45|     0.91|
+|comparison           |odds_ratio |ci_lower |ci_upper |
+|:--------------------|:----------|:--------|:--------|
+|Percentage vs kWh    |2.78       |0.53     |15.0     |
+|USD vs kWh           |9.68       |1.69     |53.4     |
+|calcNoCalculator     |60.37      |9.02     |426.4    |
+|15% Goal vs 10% Goal |0.68       |0.44     |1.0      |
+|Rounded vs Not       |0.59       |0.38     |0.9      |
 
-**Experiment 2.** Odds ratios for group comparisons.
+**Experiment 2.** Odds ratios for group comparisons. Odds ratios greater than 1 indicate increased odds of falling into a worse accuracy category compared to the comparison condition.
 
@@
 
-We again employed Bayesian ordinal logistic regression to model the probability of participants falling into each accuracy category as a function of reference class, rounding, and goal level (@tbl-s2-ord and @tbl-s2-reg).
-Results indicated that the kWh condition served as a baseline for higher accuracy.
-Compared to kWh, the USD reference class increased the odds of falling into lower-accuracy bins (Odds Ratio = 8.80, 95% CI: 1.37 to 28.42).
-The Percentage condition showed a similar trend, though the credible intervals were more uncertain.
-Notably, the “Rounded” condition showed an advantage: rounded usage information reduced the likelihood of errors (OR = 0.53, 95% CI: 0.36 to 0.73).
-Moreover, when the goal was more challenging (15% vs. 10%), accuracy generally declined (OR = 0.66, 95% CI: 0.45 to 0.91).
-Thus, while rounding facilitated more accurate responses, the more difficult goal reduced overall accuracy.
-Crucially, the kWh condition’s advantage persisted across these additional manipulations, reinforcing the conclusion from Experiment 1 that absolute units support better accuracy in energy reduction planning.
+As in Experiment 1, accuracy was categorized into three ordinal levels: "Exact match" (0% error), "0.01-5% error," and "Over 5% error".
+The analyses for Experiment 2 employed a Bayesian ordinal regression model to examine the probability of falling into one of three accuracy categories (exact match, minor deviations, or substantial deviations) as a function of the reference class condition (kWh, Percentage, USD), while including pct_goal (10% vs.
+15%), rounded (exact vs.
+rounded usage data), and calculator usage as additional predictors.
+Random intercepts were specified for both participant and state,
+
+The ordinal regression analysis revealed that the USD reference class significantly increased the odds of higher error categories compared to the kWh reference class (OR = 9.68, 95% CI: [1.69, 53.4]).
+Participants in the USD condition were therefore substantially more likely to deviate from the target energy reduction goal compared to those in the kWh condition.
+In contrast, the Percentage condition’s odds ratio relative to kWh was more uncertain (OR = 2.78, 95% CI: 0.53, 15.0), indicating that although there may be a trend toward reduced accuracy in the Percentage condition, the evidence was not definitive.
+
+We also found that using rounded numbers modestly improved accuracy (b = -0.53, 95% CI: [-0.96, -0.11]), with participants having 0.59 times the odds of falling into a worse accuracy category when working with rounded values.
+The more challenging 15% reduction goal was associated with slightly better performance compared to the 10% goal (b = -0.39, 95% CI: [-0.81, 0.04]), though this effect was relatively small.
+Consistent with Experiment 1, the use of a calculator had a large and significant effect on accuracy.
+The coefficient for calcNoCalculator was 4.10 (95% CI: 2.20, 6.06), and the corresponding odds ratio was 60.37 (95% CI: 9.02, 426.4), indicating that participants who did not use a calculator were substantially more likely to fall into higher error categories.
+
+@fig-s2-ame shows the marginal effects of refClass on each level of accuracy_level.
+These results reveal that switching from kWh to Percentage decreased the probability of an "Exact match" by an average of 7.0 percentage points (95% CI: -19.2, 4.2) and increased the probability of "Over 5% error" by 6.9 percentage points (95% CI: -4.5, 18.6).
+Similarly, switching from kWh to USD decreased the probability of an "Exact match" by 15 percentage points (95% CI: -26.7, -3.3) and increased the probability of "Over 5% error" by 16.5 percentage points (95% CI: 3.7, 29.3).
+
+```{r}
+#| label: fig-s2-ame
+#| fig-cap: "Experiment 2.
+Average marginal effects of reference class on accuracy levels (Experiment 2).
+The points represent the average change in the probability of each accuracy level when switching from the kWh reference class to Percentage (red) or USD (green).
+Error bars indicate 95% credible intervals.
+The results show that, compared to kWh, the Percentage format decreases the probability of an \"Exact match\" by 7.0% and increases the probability of \"Over 5% error\" by 6.9%.
+The USD format has a larger negative effect on \"Exact match\" (-14.7%) and a larger positive effect on \"Over 5% error\" (+16.5%).
+The effects on the \"0.01-5% error\" category are near zero for both comparisons."
+#| fig-width: 11
+#| fig-height: 8
+
+library(ggtext)
+
+set.seed(133)
+ame2 <- avg_slopes(
+    ordinal_model_s2_logit, 
+    variables = "refClass",ndraws=850
+)
+
+# Add annotations to the data frame
+ame2_annotated <- ame2 %>%
+    mutate(label = sprintf("%.1f%%", estimate * 100))
+
+ggplot(ame2_annotated, aes(x = estimate, y = group, color = contrast, group = contrast)) +
+    geom_point(size = 3, alpha=.6,position = position_dodge(width = 0.5)) +
+    geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, alpha=.5,
+                   position = position_dodge(width = 0.5)) +
+    geom_vline(xintercept = 0, linetype = "dashed",alpha=.5) +
+    labs(
+        x = "Average Marginal Effect",
+        y = "Accuracy Level",
+        color = "Comparison",
+        title = "Average Marginal Effects of refClass on Accuracy Levels"
+    ) +
+    # Add annotations
+    geom_text(
+        aes(label = label),
+        color = "black", size = 3.5, hjust = -0.3, vjust = -0.5,
+        position = position_dodge(width = 0.5)
+    ) +
+    theme_minimal()
+
+```
 
@@
 #| label: fig-s2-ppd
+#| fig-cap: "Experiment 2.
+Posterior predictive check of the ordinal probit model, grouped by reference class (Experiment 2).
+The bars represent the observed proportions of each accuracy level within each reference class.
+The points represent the model's predicted proportions, with error bars indicating 95% credible intervals."
 
-pp_check(ordinal_model_s2_logit, type = "bars_grouped", group="refClass", fatten = 2) +
-  scale_x_continuous("Response Category", breaks = 1:4, 
-            labels = c("Exact", "0.01-2%", "2.01-15%", ">15%")) +
+pp_check(ordinal_model_s2_logit, type = "bars_grouped", group="refClass", fatten = 2,ndraws=400) +
+  scale_x_continuous("Response Category", breaks = 1:3, 
+            labels = c("Exact", "0.01-5%", ">5%")) +
   scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
@@
   theme_minimal() +
+  #scale_fill_manual(values = c("kWh" = "#66c2a5", "Percentage" = "#fc8d62", "USD" = "#8da0cb"),name="Reference Class") +
   theme(
@@
 #| label: fig-s2-els
+#| fig-cap: "Experiment 2.
+Conditional effect of energy literacy on log absolute error.
+The plot shows the relationship between energy literacy score and log absolute error, controlling for random effects of participant and state.
+Higher energy literacy scores are associated with smaller deviations from the target reduction goal, indicating more accurate planning."
 
@@
     log_abs_error ~ els + (1|id) + (1|state),
-    data = s2_agg1,
+    data = s2_agg,
     family = gaussian(),
@@
 #summary(s2_els_log_error)
+# Regression Coefficients:
+#           Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# Intercept    -1.10      0.27    -1.62    -0.56 1.00      975     1675
+# els          -3.21      0.35    -3.89    -2.52 1.00      837     1508
 
@@
 
-As in Experiment 1, we further investigated the role of individual differences in energy literacy in predicting planning accuracy.
-A Bayesian linear regression model, analogous to the one used in Experiment 1 (log_abs_error ~ els + (1|id) + (1|state)), revealed a significant negative relationship between energy literacy scores and log-transformed absolute error (Estimate = -3.21, 95% CI: -3.89 to -2.52).
-This finding indicates that participants with higher energy literacy scores tended to produce more accurate plans, exhibiting smaller deviations from the target reduction goals.
-The conditional effect plot (@fig-s2-els) visually confirms this relationship, showing a clear decreasing trend in log absolute error as energy literacy increases.
+We once again examained the effect of energy literacy on planning accuracy.
+A Bayesian linear regression model was fit with log-transformed absolute error as the outcome variable and energy literacy score as the predictor, controlling for random effects of participant and state: log_abs_error ~ els + (1|id) + (1|state).
+This revealed a significant negative relationship between energy literacy and log absolute error (Estimate = -3.21, 95% CI: -3.89 to -2.52), indicating that participants with higher energy literacy scores tended to have smaller deviations from the target reduction goal, and thus more accurate plans overall (@fig-s2-els).
 
@@
 
-Experiment 2 expanded on the findings of Experiment 1 by introducing additional manipulations of goal difficulty (10% vs.
+Experiment 2 built upon the findings of Experiment 1 by incorporating additional manipulations of goal difficulty (10% vs.
 15% reduction) and numerical presentation (rounded vs.
-exact numbers).
-The results largely corroborated and extended those of Experiment 1, providing further evidence for the advantages of presenting energy reduction goals in absolute units (kWh).
+exact numbers), while maintaining the core manipulation of reference class (kWh, Percentage, USD).
+Nevertheless, the results largely converged with those of Experiment 1, providing further converging evidence that presenting energy reduction goals in absolute units (kWh) facilitates more accurate planning compared to percentage-based or monetary formats.
 
-### Individual Differences
+Taken together, the results of Experiment 2 provide further support for the hypothesis that presenting energy reduction goals in absolute units (kWh) leads to more accurate planning compared to percentage-based or monetary formats.
 
-```{r}
-#| label: fig-s2-indv
-#| fig-cap: "Study 2: Individual performance in the energy planning task, colored by reference class.
-The dashed line represents the target reduction goal.
-Participants are shown along the y axis, those who fall above or below the dashed line have not met the target goal.
-The x-axis represents the percent change in energy usage from the prior year."
-#| fig-width: 11
-#| fig-height: 9
-
-s2_agg1 |> group_by(id,refClass,calc,pct_goal,pct_change) |> 
-    filter(plan=="plan1",rounded=="Rounded") |> 
-    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
-    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),mean_abs_error=mean(abs_error),n_accuracy=first(n_accuracy)) |>
-  ungroup() |> 
-  mutate(goal_pct = as.numeric(stringr::str_remove(pct_goal,"%"))/100) |>
-  filter(mean_abs_error <= 0.50) |>
-  mutate(id=reorder(id,pct_change)) |> 
-  ggplot(aes(y=id,x=mean_pct_change,col=refClass)) + 
-  geom_point(size=1,alpha=0.6,position = position_jitter(w=0, h=0.17)) +
-  geom_vline(aes(xintercept=goal_pct),linetype="dashed",alpha=.5) +
-  ggh4x::facet_nested_wrap(~pct_goal,axes="all",scales="free",ncol=2)  + 
-  labs(y="Participant Id", x="Percent Change", title="Individual Performance") +
-  theme(axis.text.y=element_text(face = "plain", size = rel(0.7))) + 
-    scale_x_continuous(breaks = seq(0, 0.5, by = 0.05),
-                    labels = scales::percent_format(accuracy = 1))
-
-```
-
-```{r}
-#| label: fig-s2-indv2
-#| fig-cap: "Study 2: Respones patterns for a subset of individiual participants.
-Black points are participant responses, red points are the state average, and blue points are the family average.
-The x-axis represents the appliance category, and the y-axis represents the energy usage in kWh."
-#| fig.width: 11
-#| fig.height: 10
-
-s2_long |> filter(id %in% unique(s2_long$id)[1:30]) |> 
-  filter(appliance!="TOTAL",state=="California") |> 
-  ggplot(aes(x=appliance,y=value)) + geom_point(aes(shape=plan)) +
-  geom_point(aes(y=state_avg),color="red") +
-  geom_point(aes(y=family),color="blue") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
-
-s2_long |> filter(id %in% unique(s2_long$id)[1:30]) |> 
-  filter(appliance!="TOTAL",state=="California") |> 
-  ggplot(aes(x=appliance,y=value)) + geom_point(aes(shape=plan)) +
-  geom_point(aes(y=state_avg),color="red") +
-  geom_point(aes(y=family),color="blue") +
-  facet_wrap(~id) + theme(axis.text.x = element_text(angle = 45, hjust = 1))
-```
+The effect of goal difficulty, with the more challenging 15% reduction goal associated with slightly better performance, although counterintuitive, was nonetheless relatively small and may warrant further investigation in future research.
+The large and significant effect of calculator use, consistent across both experiments, underscores the crucial role of tools that individuals are likely to employ in real-world settings.
+Finally, the consistent relationship between energy literacy and accuracy, observed across both experiments, highlights the potential value of educational interventions aimed at improving consumers' understanding of energy concepts.
 

### 2024-12-18 misc/discrep.qmd
update - marginal effects

@@
 s2_long <- readRDS(here::here("data/s2_processed.rds"))
+
+# s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
+#   filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
+#   filter(!(id %in% readRDS(here::here("data/s1_grp_outlier_ids.rds")))) |>
+#   mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))
+
+# s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
+#   filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
+#   filter(!(id %in% readRDS(here::here("data/s2_grp_outlier_ids.rds"))) ) |>
+#   mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))
 

### 2024-12-18 misc/odds_ratio.qmd
update - marginal effects

@@
                     abs_error == 0.00 ~ "Exact match",
+                    abs_error <= 0.05 ~ "0.01-5% error",
+                    TRUE ~ "Over 5% error"  # Capture all remaining cases
+                ), 
+                levels = c("Exact match", "0.01-5% error", "Over 5% error"),
+                ordered = TRUE
+            )
+        ) |> relocate(accuracy_level, .after= "pct_change")
+
+s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
+    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+    summarise(
+    mg=sum(matched_goal),
+    mgc=sum(close_match),
+    n=n(), 
+    pct=mg/n,
+    pct_close=mgc/n,
+    mean_pct_change=mean(pct_change),
+    mean_abs_error=mean(abs_error),
+    mean_log_abs_error=mean(log_abs_error),
+    n_accuracy=first(n_accuracy)) |> 
+    mutate(accuracy_level = factor(
+            case_when(
+                mean_abs_error < 0.02 ~ "Exact match",
+                mean_abs_error <= 0.05 ~ "0.01-5% error",
+                TRUE ~ "Over 5% error"  # Capture all remaining cases
+            ), 
+            levels = c("Exact match", ".01-5% error", "Over 5% error"),
+            ordered = TRUE
+        ))
+
+```
+
+```{r}
+#| label: tbl-s1-ord
+#| tbl-cap: "Study 1: Odds ratios of for group comparisons. "
+
+ordinal_model_s1 <- brm(
+    accuracy_level ~ refClass + (1|id) + (1|state),
+    data = s1_agg,
+    family = cumulative("logit"),
+    cores = 4,
+    iter = 5000,
+    save_pars = save_pars(all = TRUE),
+    control = list(adapt_delta = 0.99), 
+    prior = c(prior(normal(0, 3), class = "Intercept"), 
+                prior(normal(0, 3), class = "b")), 
+    file = paste0(here::here("data/model_cache",'s1_ordinal3.rds')) 
+)
+# 
+# ordinal_model_s1_null <- brm(
+#     accuracy_level ~  (1|id) + (1|state),
+#     data = s1_agg,
+#     family = cumulative("logit"),
+#     cores = 4,
+#     iter = 5000,
+#     save_pars = save_pars(all = TRUE),
+#     control = list(adapt_delta = 0.99), 
+#     prior = c(prior(normal(0, 3), class = "Intercept")),
+#     file = paste0(here::here("data/model_cache",'s1_null.rds')) 
+# )
+
+#colnames(mted1) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")
+
+as.data.frame(describe_posterior(ordinal_model_s1, centrality = "Mean"))[, c(1,2,4,5,6)] |> 
+  setNames(c("Parameter", "Estimate", "CI_Lower", "CI_Upper", "pd")) |> 
+  mutate(Parameter = stringr::str_remove(Parameter, "b_"))
+
+# ordinal_model_s1 |> emmeans(~refClass) |> contrast(method="pairwise") |>  
+#   gather_emmeans_draws() %>%
+#   mean_hdi(.width = .95)
+
+# describe_posterior(ordinal_model_s1, centrality = "Mode") |> 
+#     filter(stringr::str_detect(Parameter, "b_")) |> 
+#     mutate(Parameter = stringr::str_remove(Parameter, "b_")) |> 
+#     mutate(across(where(is.numeric), round, 3)) |> 
+#     kable(booktabs = TRUE)
+
+# Get predicted probabilities
+pred_summary <- ordinal_model_s1 |>
+    epred_draws(newdata = data.frame(refClass = c("kWh", "Percentage", "USD")),
+                ndraws = 1000, re_formula = NA) |>
+    group_by(refClass, Category=.category) |>
+    summarise(
+        mean_prob = mean(.epred),
+        lower_ci = quantile(.epred, 0.025),
+        upper_ci = quantile(.epred, 0.975)
+    )
+#pred_summary |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy")
+
+# Convert log-odds to odds ratios
+posterior_samples <- as.data.frame(ordinal_model_s1)
+odds_ratios <- data.frame(
+  Percentage_vs_kWh = exp(posterior_samples$b_refClassPercentage),
+  USD_vs_kWh = exp(posterior_samples$b_refClassUSD)
+)
+
+# Calculate summary statistics
+odds_ratio_summary <- data.frame(
+    comparison = c("Percentage vs kWh", "USD vs kWh"),
+    odds_ratio = c(mean(odds_ratios$Percentage_vs_kWh),
+                    mean(odds_ratios$USD_vs_kWh)),
+    ci_lower = c(quantile(odds_ratios$Percentage_vs_kWh, 0.025),
+                quantile(odds_ratios$USD_vs_kWh, 0.025)),
+    ci_upper = c(quantile(odds_ratios$Percentage_vs_kWh, 0.975),
+                quantile(odds_ratios$USD_vs_kWh, 0.975))
+)
+
+odds_ratio_summary |> kable(escape=FALSE,booktabs=TRUE,align=c("l")) 
+# |comparison        | odds_ratio| ci_lower| ci_upper|
+# |:-----------------|----------:|--------:|--------:|
+# |Percentage vs kWh |        2.3|     0.62|      6.2|
+# |USD vs kWh        |       14.1|     4.08|     38.0|
+
+#odds_ratio_summary |> pander::pandoc.table(caption="Study 1: Odds ratios of accuracy")
+
+# odds ratios of fixed effects
+as.data.frame(fixef(ordinal_model_s1)[,-2])|> as.data.frame() %>%
+    rownames_to_column(var = "Parameter") %>%
+    mutate(across(where(is.numeric), exp)) |>
+    filter(!stringr::str_detect(Parameter, "Intercept"))
+
+describe_posterior(ordinal_model_s1, centrality = "Median")[, c(1,2,4,5)] %>%
+  as_tibble() |> # remove rows with Intercept
+  mutate(across(where(is.numeric), exp)) |>
+  filter(!stringr::str_detect(Parameter, "Intercept")) |>
+  mutate(Parameter = stringr::str_remove(Parameter, "b_"))
+
+or_table <- data.frame(exp(fixef(ordinal_model_s1)))
+or_table |>
+  rownames_to_column(var = "Term") |>
+  as_flextable() |>
+  set_header_labels(values = list(
+    OR = "Odds Ratio",
+    `2.5 %` = "Lower 95% CI",
+    `97.5 %` = "Upper 95% CI"
+  )) |>
+  flextable::add_footer_lines(values = "Note: Odds ratios derived from the ordinal logistic regression model.") |> 
+  align(align = 'center', part = 'all') |>
+  fontsize(size = 10, part = "all")
+
+posterior_draws <- as_draws_df(ordinal_model_s1) |> select(starts_with("b_"))
+
+# Compute posterior summaries and exponentiate to get odds ratios
+or_summary <- posterior_draws %>%
+  summarise(across(starts_with("b_"),
+                   list(mean = ~mean(.x),
+                        sd = ~sd(.x),
+                        lower95 = ~quantile(.x, 0.025),
+                        upper95 = ~quantile(.x, 0.975)))) %>%
+  tidyr::pivot_longer(cols = everything(),
+                      names_to = c("parameter", ".value"),
+                      names_pattern = "(b_.*)_(.*)") %>%
+  mutate(odds_ratio = exp(mean),
+         lower95_or = exp(lower95),
+         upper95_or = exp(upper95))
+
+or_table <- or_summary %>%
+  select(parameter, odds_ratio, lower95_or, upper95_or) %>%
+  mutate(parameter = gsub("b_", "", parameter)) %>%
+  rename("Predictor" = parameter,
+         "OR" = odds_ratio,
+         "2.5% CI" = lower95_or,
+         "97.5% CI" = upper95_or) %>%
+  kbl(digits = 2, booktabs = TRUE, align = "lccc",
+      caption = "Odds Ratios for Ordinal Model") %>%
+  kableExtra::kable_styling(full_width = FALSE)
+
+print(or_table)
+
+```
+
+```{r}
+
+summary(ordinal_model_s1)
+
+pp_check(ordinal_model_s1, type = "bars_grouped", group="refClass", fatten = 2) +
+  scale_x_continuous("Response Category", breaks = 1:4, 
+            labels = c("Exact", "0.01-2%", "2.01-15%", ">15%")) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
+  ggtitle("Posterior Predictive Check - ordinal_model_s1") +
+  theme_minimal() +
+  theme(
+    legend.background = element_blank(),
+    legend.position = "bottom",
+    panel.grid.minor = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 1)
+  )
+
+pp_check(ordinal_model_s1, type = "dens_overlay_grouped", group="refClass", fatten = 2,ndraws=100) +   scale_x_continuous("Response Category", breaks = 1:4, 
+            labels = c("Exact", "0.01-2%", "2.01-15%", ">15%")) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
+  ggtitle("Posterior Predictive Check - ordinal_model_s1") +
+  theme_minimal() +
+  theme(
+    legend.background = element_blank(),
+    legend.position = "bottom",
+    panel.grid.minor = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 1)
+  )
+```
+
+```{r}
+
+multinomial_model_s1 <- brm(
+  accuracy_level ~ refClass + (1|id) + (1|state),
+  data = s1_agg,
+  family = categorical(), # Note: categorical() for multinomial
+  cores = 4,
+  iter = 1000,
+  save_pars = save_pars(all = TRUE),
+  control = list(adapt_delta = 0.99, max_treedepth=12),
+#   prior = c(prior(normal(0, 3), class = "Intercept"), 
+#             prior(normal(0, 3), class = "b")),
+  file = paste0(here::here("data/model_cache",'s1_multinomial.rds')) 
+)
+
+summary(multinomial_model_s1)
+pp_check(multinomial_model_s1, type = "bars_grouped", group="refClass", fatten = 2,ndraws=100) + labs(title = "Study 1: Posterior predictive check for the multinomial model") + theme_minimal()
+
+```
+
+```{r}
+#| eval: false
+
+loo_ordinal <- LOO(ordinal_model_s1, moment_match=TRUE)
+loo_multinomial <- LOO(multinomial_model_s1, moment_match=TRUE, reloo=TRUE)
+loo_compare(loo_ordinal, loo_multinomial)
+
+loo_ordinal <- LOO(ordinal_model_s1)
+loo_multinomial <- LOO(multinomial_model_s1)
+loo_compare(loo_ordinal, loo_multinomial)
+
+```
+
+## S1 exact match
+
+```{r}
+
+# Assuming s1_agg is your dataframe from Experiment 1, and it's already loaded
+
+# Fit a Bayesian logistic regression model
+exact_match_model <- brm(
+  matched_goal ~ refClass + (1|id) + (1|state),
+  data = s1_agg,
+  family = bernoulli(link = "logit"),
+  cores = 4,
+  iter = 2000,
+  control = list(adapt_delta = 0.97),
+  prior = c(
+    prior(normal(0, 2), class = "Intercept"),
+    prior(normal(0, 2), class = "b")
+  ),
+  file = paste0(here::here("data/model_cache", 's1_exact_match_model.rds'))
+)
+summary(exact_match_model)
+
+exact_match_model_c_add <- brm(
+  matched_goal ~ refClass + calc+ (1|id) + (1|state),
+  data = s1_agg,
+  family = bernoulli(link = "logit"),
+  cores = 4,
+  iter = 2000,
+  control = list(adapt_delta = 0.97),
+  prior = c(
+    prior(normal(0, 2), class = "Intercept"),
+    prior(normal(0, 2), class = "b")
+  ),
+  file = paste0(here::here("data/model_cache", 's1_exact_match_model_c_add.rds'))
+)
+summary(exact_match_model_c_add)
+
+exact_match_model_c_int <- brm(
+  matched_goal ~ refClass * calc+ (1|id) + (1|state),
+  data = s1_agg,
+  family = bernoulli(link = "logit"),
+  cores = 4,
+  iter = 2000,
+  control = list(adapt_delta = 0.97),
+  prior = c(
+    prior(normal(0, 2), class = "Intercept"),
+    prior(normal(0, 2), class = "b")
+  ),
+  file = paste0(here::here("data/model_cache", 's1_exact_match_model_c_int.rds'))
+)
+summary(exact_match_model_c_int )
+conditional_effects(exact_match_model_c_int)
+
+```
+
+## outliers
+
+```{r}
+
+identify_outliers_group <- function(x) {
+  Q1 <- quantile(x, 0.25, na.rm = TRUE)
+  Q3 <- quantile(x, 0.75, na.rm = TRUE)
+  IQR_value <- Q3 - Q1
+  lower_bound <- Q1 - 3.3 * IQR_value
+  upper_bound <- Q3 + 3.3 * IQR_value
+  return(x < lower_bound | x > upper_bound)
+}
+
+# 2. Apply the Outlier Detection to abs_error and log_abs_error within each group
+s1_agg <- s1_agg %>%
+  group_by(refClass, calc) %>%
+  mutate(
+    outlier_abs_error = identify_outliers_group(abs_error),
+    outlier_log_abs_error = identify_outliers_group(log_abs_error)
+  ) %>%
+  ungroup()
+
+#s1_agg |> relocate(outlier_abs_error,.after=abs_error) |> sort(desc(outlier_abs_error)) |> head(5)
+s1_agg |> filter(outlier_abs_error) |> select(id,refClass,calc,abs_error,log_abs_error) |> # keep unique id and refClass pairs
+  distinct(id,refClass)
+
+# 3. Summarize the Number and Proportion of Outliers by Group
+outlier_summary_by_group <- s1_agg %>%
+  group_by(refClass, calc) %>%
+  summarise(
+    total_observations = n(),
+    outliers_abs_error = sum(outlier_abs_error),
+    percent_outliers_abs_error = round((outliers_abs_error / total_observations) * 100, 2),
+    outliers_log_abs_error = sum(outlier_log_abs_error),
+    percent_outliers_log_abs_error = round((outliers_log_abs_error / total_observations) * 100, 2)
+  ) %>%
+  ungroup()
+
+# Display the outlier summary by group
+print("Outlier Summary by Reference Class and Calculator Usage:")
+print(outlier_summary_by_group)
+
+s1_agg_no_outliers <- s1_agg %>%
+  filter(!outlier_abs_error & !outlier_log_abs_error)
+
+# Re-run the exact match comparison without outliers
+
+# Create a contingency table of reference class vs. exact match without outliers
+exact_match_table_no_outliers <- s1_agg_no_outliers %>%
+  mutate(exact_match = ifelse(matched_goal, "Exact Match", "No Exact Match")) %>%
+  count(refClass, exact_match) %>%
+  pivot_wider(names_from = exact_match, values_from = n, values_fill = 0)
+
+# Perform Chi-Square Test of Independence without outliers
+chi_square_test_no_outliers <- chisq.test(exact_match_table_no_outliers[, -1])
+
+# Display results
+print("Chi-Square Test Results (Without Outliers):")
+print(chi_square_test_no_outliers)
+
+# Optionally, re-run logistic regression without outliers
+logistic_model_no_outliers <- glm(matched_goal ~ refClass, data = s1_agg_no_outliers, family = binomial)
+summary(logistic_model_no_outliers)
+
+# Calculate Odds Ratios
+odds_ratios_no_outliers <- exp(coef(logistic_model_no_outliers))
+print("Odds Ratios (Without Outliers):")
+print(odds_ratios_no_outliers)
+
+# Calculate 95% Confidence Intervals for Odds Ratios without outliers
+conf_intervals_no_outliers <- exp(confint(logistic_model_no_outliers))
+print("95% Confidence Intervals for Odds Ratios (Without Outliers):")
+print(conf_intervals_no_outliers)
+
+s1_clean <- s1_agg %>%
+  # Remove trials that are outliers in abs_error or log_abs_error
+  #filter(!(outlier_abs_error | outlier_log_abs_error))
+  # remove any id that has any outlier_abs_error
+  filter(!(id %in% (s1_agg |> filter(outlier_abs_error) |> select(id) |> distinct() |> pull(id))))
+
+length(unique(s1_agg$id)) 
+length(unique(s1_clean$id))
+
+# 2. Summary of Cleaned Data
+cleaned_summary <- s1_clean %>%
+  group_by(refClass, calc) %>%
+  summarise(
+    total_observations = n(),
+    exact_matches = sum(matched_goal),
+    percent_exact = round((exact_matches / total_observations) * 100, 2),
+    mean_abs_error = mean(abs_error, na.rm = TRUE),
+    mean_log_abs_error = mean(log_abs_error, na.rm = TRUE)
+  ) %>%
+  ungroup()
+
+# Display the cleaned data summary
+print("Cleaned Data Summary by Reference Class and Calculator Usage:")
+print(cleaned_summary)
+
+ordinal_model_s1_clean_calc <- brm(
+    accuracy_level ~ refClass +calc + (1|id) + (1|state),
+    data = s1_clean,
+    family = cumulative("logit"),
+    cores = 4,
+    iter = 2000,
+    save_pars = save_pars(all = TRUE),
+    control = list(adapt_delta = 0.97), 
+    prior = c(prior(normal(0, 3), class = "Intercept"), 
+                prior(normal(0, 3), class = "b")), 
+    file = paste0(here::here("data/model_cache",'s1_ordinal_clean_calc_3.3.rds')) 
+)
+
+summary(ordinal_model_s1_clean_calc)
+
+ordinal_model_s1_clean_calc_int <- brm(
+    accuracy_level ~ refClass * calc + (1|id) + (1|state),
+    data = s1_clean,
+    family = cumulative("logit"),
+    cores = 4,
+    iter = 2000,
+    save_pars = save_pars(all = TRUE),
+    control = list(adapt_delta = 0.97), 
+    prior = c(prior(normal(0, 3), class = "Intercept"), 
+                prior(normal(0, 3), class = "b")), 
+    file = paste0(here::here("data/model_cache",'s1_ordinal_clean_calc_int.rds')) 
+)
+
+summary(ordinal_model_s1_clean_calc_int)
+conditional_effects(ordinal_model_s1_clean_calc_int)
+
+# logistic regression on exact match with clean data (frequentist)
+logistic_model_clean <- glm(matched_goal ~ refClass, data = s1_clean, family = binomial)
+summary(logistic_model_clean)
+
+# logistic regression on exact match with clean data (frequentist) - mixed effects
+logistic_model_clean_mixed <- glmer(matched_goal ~ refClass + (1|id) + (1|state), data = s1_clean, family = binomial)
+summary(logistic_model_clean_mixed)
+
+# logistic regression on exact match with clean data (frequentist) - mixed effects
+logistic_model_clean_mixed <- glmer(matched_goal ~ refClass+calc + (1|id) , data = s1_clean, family = binomial)
+summary(logistic_model_clean_mixed)
+
+# logistic regression on exact match with clean data (bayesian)
+logistic_model_clean_bayes <- brm(
+  matched_goal ~ refClass + (1|id) + (1|state),
+  data = s1_clean,
+  family = bernoulli(link = "logit"),
+  cores = 4,
+  iter = 2000,
+  control = list(adapt_delta = 0.97),
+  prior = c(
+    prior(normal(0, 2), class = "Intercept"),
+    prior(normal(0, 2), class = "b")
+  ),
+  file = paste0(here::here("data/model_cache", 's1_exact_match_model_clean.rds'))
+)
+summary(logistic_model_clean_bayes)
+conditional_effects(logistic_model_clean_bayes)
+
+```
+
+## outlier from avg.
+
+```{r}
+
+group_stats <- s1_agg |>
+  group_by(refClass, calc) |>
+  summarise(
+    group_mean_abs_error = mean(abs_error, na.rm = TRUE),
+    group_sd_abs_error = sd(abs_error, na.rm = TRUE),
+    .groups = "drop"
+  )
+
+# Calculate participant-level average abs_error
+participant_stats <- s1_agg |>
+  group_by(id, refClass, calc) |>
+  summarise(
+    avg_abs_error = mean(abs_error, na.rm = TRUE),
+    .groups = "drop"
+  )
+
+# Join group stats with participant stats
+participant_stats <- participant_stats |>
+  left_join(group_stats, by = c("refClass", "calc"))
+
+# Identify participants to exclude based on the 2.5 SD threshold
+participants_to_exclude <- participant_stats |>
+  mutate(
+    exclusion_threshold = group_mean_abs_error + 2.5 * group_sd_abs_error,
+    exclude = avg_abs_error > exclusion_threshold
+  ) |>
+  filter(exclude)
+
+ordinal_model_s1_clean_calc <- brm(
+    accuracy_level ~ refClass +calc + (1|id) + (1|state),
+    data = s1_clean,
+    family = cumulative("logit"),
+    cores = 4,
+    iter = 2000,
+    save_pars = save_pars(all = TRUE),
+    control = list(adapt_delta = 0.97), 
+    prior = c(prior(normal(0, 3), class = "Intercept"), 
+                prior(normal(0, 3), class = "b")), 
+    file = paste0(here::here("data/model_cache",'s1_ordinal_clean_calc_3.3.rds')) 
+)
+
+s1_agg_filtered <- s1_agg |>
+  filter(!(id %in% participants_to_exclude$id))
+
+length(unique(s1_agg$id)) 
+length(unique(s1_agg_filtered$id))
+
+ordinal_model_s1_clean_calc <- brm(
+    accuracy_level ~ refClass +calc + (1|id) + (1|state),
+    data = s1_agg_filtered,
+    family = cumulative("logit"),
+    cores = 4,
+    iter = 2000,
+    control = list(adapt_delta = 0.97), 
+    prior = c(prior(normal(0, 3), class = "Intercept"), 
+                prior(normal(0, 3), class = "b")), 
+    file = paste0(here::here("data/model_cache",'s1_ordinal_grpclean_calc_2.5.rds')) 
+)
+
+summary(ordinal_model_s1_clean_calc)
+conditional_effects(ordinal_model_s1_clean_calc)
+
+mixed_model_abs_error <- lmer(log_abs_error ~ refClass + calc+ (1|id) + (1|state), data = s1_agg_filtered)
+summary(mixed_model_abs_error)
+# with p value
+library(lmerTest)
+mixed_model_abs_error <- lmer(log_abs_error ~ refClass + calc+ (1|id) + (1|state), data = s1_agg_filtered)
+summary(mixed_model_abs_error)
+
+```
+
+### clearn log error
+```{r}
+
+s1_clean |> 
+  ggplot(aes(x = abs_error)) +
+  geom_density(fill = "skyblue", alpha = 0.7) +
+  facet_wrap(~ refClass, scales = "free") +ggtitle("Error density by reference class")
+
+# plot density of log-transformed data
+s1_clean |> 
+  ggplot(aes(x = log_abs_error)) +
+  geom_density(fill = "skyblue", alpha = 0.7) +
+  facet_wrap(~ refClass, scales = "free") + ggtitle("Log-transformed error density by reference class")
+
+# mixed effects regression on abs_error (frequentist)
+mixed_model_abs_error <- lmer(abs_error ~ refClass + (1|id) + (1|state), data = s1_clean)
+summary(mixed_model_abs_error)
+
+mixed_model_abs_error <- lmer(abs_error ~ refClass*calc + (1|id) + (1|state), data = s1_clean)
+summary(mixed_model_abs_error)
+
+```
+
+# S2
+
+```{r}
+#| label: tbl-s2-agg
+#| tbl-cap: "Study 2: Summary of planning accuracy by reference class.
+The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."
+
+s2_agg <- s2_long |> 
+  filter(appliance != "TOTAL") |> 
+  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
+  summarise(
+    total_kWh = sum(value),
+    orig_kWh = sum(family),
+    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
+    state_dif = mean(state_dif),
+    .groups = "drop"
+  ) |>
+  mutate(
+    matched_goal = (pct_change == pct),
+    close_match = abs(pct_change - pct) <= 0.02,
+                error = pct_change - pct,
+                abs_error = abs(error),
+                log_abs_error=log(abs(error)+.007)) |> 
+      ungroup() |> 
+        mutate(
+            accuracy_level = factor(
+                case_when(
+                    abs_error == 0.00 ~ "Exact match",
+                    abs_error <= 0.05 ~ "0.01-5% error",
+                    TRUE ~ "Over 5% error"  # Capture all remaining cases
+                ), 
+                levels = c("Exact match","0.01-5% error", "Over 5% error"),
+                ordered = TRUE
+            )
+        )
+
+s2_agg4 <- s2_agg |> group_by(id,refClass,calc) |> 
+    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+    summarise(
+    mg=sum(matched_goal),
+    mgc=sum(close_match),
+    n=n(), 
+    pct=mg/n,
+    pct_close=mgc/n,
+    mean_pct_change=mean(pct_change),
+    mean_abs_error=mean(abs_error),
+    mean_log_abs_error=mean(log_abs_error),
+    n_accuracy=first(n_accuracy)) |> 
+    mutate(accuracy_level = factor(
+            case_when(
+                mean_abs_error < 0.02 ~ "Exact match",
+                mean_abs_error <= 0.05 ~ ".01-5% error",
+                TRUE ~ "Over 5% error"  # Capture all remaining cases
+            ), 
+            levels = c("Exact match", "01-5% error", "Over 5% error"),
+            ordered = TRUE
+        ))
+
+# overall pct of subjects who matched their goal
+s2_agg4 |> group_by(refClass) |>
+    summarise(
+    'Avg. % meeting goal' = mean(pct),
+    'Avg. Deviation From Goal' = median(mean_pct_change),
+    # sd = sd(pct),
+    # n = n(),
+    #se=sd(pct)/sqrt(n)
+) |>   mutate(across(where(is.numeric), \(x) round(x, 3))) %>% 
+  kable(escape=FALSE,booktabs=TRUE,align=c("l"))
+
+```
+
+```{r}
+
+ordinal_model_s2_logit <- brm(
+  accuracy_level ~ refClass +rounded+pct_goal + (1|id)+ (1|state),
+  data = s2_agg,
+  family = cumulative("logit"),
+  cores = 4,
+  iter = 2000,
+  control = list(adapt_delta = 0.99), # Recommended for ordinal models
+  prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
+            prior(normal(0, 2), class = "b")), # Priors for predictors
+  file = paste0(here::here("data/model_cache",'s2_logit_add.rds')) # Cache for efficiency
+)
+summary(ordinal_model_s2_logit)
+
+pred_summary_s2 <- ordinal_model_s2_logit %>%
+  epred_draws(newdata = s2_agg, re_formula = NA,ndraws=200) %>%
+  group_by("Reference Class"=refClass, rounded, "% Goal"=pct_goal, Category=.category) %>%
+  summarise(
+    mean_prob = mean(.epred),
+    lower_ci = quantile(.epred, 0.025),
+    upper_ci = quantile(.epred, 0.975),
+    .groups = "drop"
+  )
+
+# Create a new data frame with the predictor values
+new_data <- expand.grid(
+  refClass = c("kWh", "Percentage", "USD"),
+  id = unique(s2_agg$id)[1],      # Take first id
+  state = unique(s2_agg$state)[1], # Take first state,
+  rounded = unique(s2_agg$rounded)[1], # Take first rounded
+  pct_goal = unique(s2_agg$pct_goal)[1] # Take first pct_goal
+)
+
+# Generate predicted probabilities
+predicted_probabilities <- predict(ordinal_model_s2_logit, newdata = new_data, 
+                                  summary = FALSE, ndraws = 1000)
+
+# Calculate the predicted probability of "Exact match"
+predicted_probability_exact_match <- mean(predicted_probabilities <= 1)
+
+print(predicted_probability_exact_match)
+
+pred_summary_s2 |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy level by reference class",split.table=Inf)
+
+# Extract posterior samples
+posterior_samples_s2 <- as.data.frame(ordinal_model_s2_logit)
+
+odds_ratios_s2 <- posterior_samples_s2 %>%
+    transmute(
+        refClass_Percentage_vs_kWh = exp(b_refClassPercentage),
+        refClass_USD_vs_kWh = exp(b_refClassUSD),
+        rounded_Yes_vs_No = exp(b_roundedRounded),
+        pct_goal_15_vs_10 = exp(`b_pct_goal15%`)  # Note the backticks here
+    )
+
+# Calculate summary statistics
+odds_ratio_summary_s2 <- data.frame(
+    comparison = c("Percentage vs kWh", "USD vs kWh", "Rounded vs Not", "15% Goal vs 10% Goal"),
+    odds_ratio = c(
+        mean(odds_ratios_s2$refClass_Percentage_vs_kWh),
+        mean(odds_ratios_s2$refClass_USD_vs_kWh),
+        mean(odds_ratios_s2$rounded_Yes_vs_No),
+        mean(odds_ratios_s2$pct_goal_15_vs_10)
+    ),
+    ci_lower = c(
+        quantile(odds_ratios_s2$refClass_Percentage_vs_kWh, 0.025),
+        quantile(odds_ratios_s2$refClass_USD_vs_kWh, 0.025),
+        quantile(odds_ratios_s2$rounded_Yes_vs_No, 0.025),
+        quantile(odds_ratios_s2$pct_goal_15_vs_10, 0.025)
+    ),
+    ci_upper = c(
+        quantile(odds_ratios_s2$refClass_Percentage_vs_kWh, 0.975),
+        quantile(odds_ratios_s2$refClass_USD_vs_kWh, 0.975),
+        quantile(odds_ratios_s2$rounded_Yes_vs_No, 0.975),
+        quantile(odds_ratios_s2$pct_goal_15_vs_10, 0.975)
+    )
+)
+
+#             comparison odds_ratio ci_lower ci_upper
+# 1    Percentage vs kWh       1.99     0.73     4.43
+# 2           USD vs kWh       3.69     1.33     8.29
+# 3       Rounded vs Not       0.70     0.57     0.84
+# 4 15% Goal vs 10% Goal       0.78     0.63     0.94
+
+odds_ratio_summary_s2 |> kable()
+
+odds_ratio_summary_s2 |> pander::pandoc.table(caption="Study 2: Odds ratios of accuracy")
+
+# Create a new data frame with the predictor values
+# new_data <- expand.grid(
+#   refClass = c("kWh", "Percentage", "USD"),
+#   id = unique(s2_agg$id)[1],      # Take first id
+#   state = unique(s2_agg$state)[1], # Take first state,
+#   rounded = unique(s2_agg$rounded)[1], # Take first rounded
+#   pct_goal = unique(s2_agg$pct_goal)[1] # Take first pct_goal
+# )
+
+# or_table <- data.frame(exp(fixef(ordinal_model_s2_logit)))
+# or_table |>
+#   rownames_to_column(var = "Term") |>
+#   filter(!stringr::str_detect(Term, "Intercept")) |>
+#   as_flextable() |>
+#   set_header_labels(values = list(
+#     OR = "Odds Ratio",
+#     `2.5 %` = "Lower 95% CI",
+#     `97.5 %` = "Upper 95% CI"
+#   )) |>
+#   flextable::add_footer_lines(values = "Note: Odds ratios derived from the ordinal logistic regression model.") |> 
+#   align(align = 'center', part = 'all') |>
+#   fontsize(size = 10, part = "all")
+
+```
+
+# Interaction s2
+
+```{r}
+
+ordinal_model_s2_logit_int2 <- brm(
+  accuracy_level ~ refClass +(rounded*pct_goal) + (1|id)+ (1|state),
+  data = s2_agg,
+  family = cumulative("logit"),
+  cores = 4,
+  iter = 2000,
+  control = list(adapt_delta = 0.99), # Recommended for ordinal models
+  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
+            prior(normal(0, 3), class = "b")), # Priors for predictors
+  file = paste0(here::here("data/model_cache",'s2_logit_int2.rds')) # Cache for efficiency
+)
+summary(ordinal_model_s2_logit_int2)
+
+conditional_effects(ordinal_model_s2_logit_int2)
+
+ordinal_model_s2_logit_int3 <- brm(
+  accuracy_level ~ refClass * rounded*pct_goal + (1|id)+ (1|state),
+  data = s2_agg,
+  family = cumulative("logit"),
+  cores = 4,
+  iter = 3000,
+  control = list(adapt_delta = 0.99), # Recommended for ordinal models
+  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
+            prior(normal(0, 3), class = "b")), # Priors for predictors
+  file = paste0(here::here("data/model_cache",'s2_logit_int3.rds')) # Cache for efficiency
+)
+summary(ordinal_model_s2_logit_int3)
+
+conditional_effects(ordinal_model_s2_logit_int3)
+
+ordinal_model_s2_logit_c1rf_int3<- brm(
+  accuracy_level ~ refClass * rounded*pct_goal + (1|id)+ (1|state) + (1|calc),
+  data = s2_agg,
+  family = cumulative("logit"),
+  cores = 4,
+  iter = 3000,
+  control = list(adapt_delta = 0.99), # Recommended for ordinal models
+  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
+            prior(normal(0, 3), class = "b")), # Priors for predictors
+  file = paste0(here::here("data/model_cache",'s2_logit_c1rf_int3.rds')) # Cache for efficiency
+)
+summary(ordinal_model_s2_logit_c1rf_int3)
+
+conditional_effects(ordinal_model_s2_logit_c1rf_int3)
+
+ordinal_model_s2_logit_c1add_int3<- brm(
+  accuracy_level ~ (refClass * rounded*pct_goal) + calc + (1|id)+ (1|state),
+  data = s2_agg,
+  family = cumulative("logit"),
+  cores = 4,
+  iter = 3000,
+  control = list(adapt_delta = 0.97), # Recommended for ordinal models
+  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
+            prior(normal(0, 3), class = "b")), # Priors for predictors
+  file = paste0(here::here("data/model_cache",'s2_logit_c1add_int3.rds')) # Cache for efficiency
+)
+summary(ordinal_model_s2_logit_c1add_int3)
+
+conditional_effects(ordinal_model_s2_logit_c1add_int3)
+
+fixef(ordinal_model_s2_logit_c1add_int3)
+
+library(marginaleffects)
+library(tinytable)
+
+ame <- avg_slopes(
+  ordinal_model_s2_logit_c1add_int3, 
+  variables = "refClass"
+)
+print(ame)
+
+ame_by_rounded <- avg_slopes(
+  ordinal_model_s2_logit_c1add_int3,
+  variables = "refClass",
+  by = c("rounded", "accuracy_level")
+)
+print(ame_by_rounded)
+
+ame_by_rounded_pct_goal <- avg_slopes(
+  ordinal_model_s2_logit_c1add_int3,
+  variables = "refClass",
+  by = c("rounded", "pct_goal", "accuracy_level")
+)
+print(ame_by_rounded_pct_goal)
+
+conditional_slopes <- slopes(
+    ordinal_model_s2_logit_c1add_int3,
+    variables = "refClass",
+    newdata = datagrid(
+        rounded = c('Not Rounded', 'Rounded'),
+        pct_goal = c("10%", "15%")
+    )
+)
+conditional_slopes |> head(n=25)
+
+plot_slopes(
+    ordinal_model_s2_logit_c1add_int3,
+    variables = "refClass",
+    condition = c("rounded", "pct_goal")
+)
+
+avg_slopes(
+    ordinal_model_s2_logit_c1add_int3,
+    variables = "refClass",
+    hypothesis = "b1 - b2 = 0"  # Compare effects for Rounded vs Not Rounded
+)
+#  Estimate   2.5 % 97.5 %
+#    0.0569 -0.0528  0.159
+
+# Term: b1-b2=0
+# Type:  response 
+# Columns: term, estimate, conf.low, conf.high 
+avg_slopes(
+    ordinal_model_s2_logit_c1add_int3,
+    variables = "refClass",
+    hypothesis = "b1 - b3 = 0"  # Compare effects for Rounded vs Not Rounded
+)
+#  Estimate  2.5 % 97.5 %
+#   -0.0742 -0.192 0.0264
+
+# Term: b1-b3=0
+# Type:  response 
+# Columns: term, estimate, conf.low, conf.high
+
+avg_slopes(
+ordinal_model_s2_logit_c1add_int3,
+variables = "refClass",
+by = c("rounded", "pct_goal"),
+hypothesis = "(b1 - b2) - (b3 - b4) = 0"
+)
+#               Term Estimate   2.5 % 97.5 %
+#  (b1-b2)-(b3-b4)=0   0.0201 -0.0797 0.0985
+
+# Type:  response 
+# Columns: term, estimate, conf.low, conf.high
+
+# Example: Pairwise comparison of AMEs for refClass
+pairwise_comparisons <- avg_slopes(
+  ordinal_model_s2_logit_c1add_int3,
+  variables = "refClass",
+  hypothesis = "pairwise"
+)
+print(pairwise_comparisons)
+#                                                                    Term Estimate    2.5 %    97.5 %
+#  (Exact match, Percentage - kWh) - (Exact match, USD - kWh)                0.0575 -0.04893  0.170520
+#  (Exact match, Percentage - kWh) - (0.01-2% error, Percentage - kWh)      -0.0738 -0.18531  0.028426
+#  (Exact match, Percentage - kWh) - (0.01-2% error, USD - kWh)             -0.0618 -0.18001  0.051139
+#  (Exact match, Percentage - kWh) - (2.01-15% error, Percentage - kWh)     -0.1011 -0.25716  0.035666
+#  (Exact match, Percentage - kWh) - (2.01-15% error, USD - kWh)            -0.1122 -0.26659  0.023111
+#  (Exact match, Percentage - kWh) - (Over 15% error, Percentage - kWh)     -0.1485 -0.34326  0.061167
+#  (Exact match, Percentage - kWh) - (Over 15% error, USD - kWh)            -0.2098 -0.38566 -0.025617
+#  (Exact match, USD - kWh) - (0.01-2% error, Percentage - kWh)             -0.1321 -0.25511 -0.019752
+#  (Exact match, USD - kWh) - (0.01-2% error, USD - kWh)                    -0.1202 -0.23245 -0.018717
+#  (Exact match, USD - kWh) - (2.01-15% error, Percentage - kWh)            -0.1615 -0.31554 -0.024206
+#  (Exact match, USD - kWh) - (2.01-15% error, USD - kWh)                   -0.1707 -0.33209 -0.029627
+#  (Exact match, USD - kWh) - (Over 15% error, Percentage - kWh)            -0.2082 -0.38108 -0.025534
+#  (Exact match, USD - kWh) - (Over 15% error, USD - kWh)                   -0.2690 -0.47833 -0.045882
+#  (0.01-2% error, Percentage - kWh) - (0.01-2% error, USD - kWh)            0.0108 -0.00632  0.037468
+#  (0.01-2% error, Percentage - kWh) - (2.01-15% error, Percentage - kWh)   -0.0274 -0.07868  0.010135
+#  (0.01-2% error, Percentage - kWh) - (2.01-15% error, USD - kWh)          -0.0381 -0.08937 -0.000373
+#  (0.01-2% error, Percentage - kWh) - (Over 15% error, Percentage - kWh)   -0.0728 -0.16873  0.032944
+#  (0.01-2% error, Percentage - kWh) - (Over 15% error, USD - kWh)          -0.1357 -0.22795 -0.024503
+#  (0.01-2% error, USD - kWh) - (2.01-15% error, Percentage - kWh)          -0.0399 -0.09148 -0.002582
+#  (0.01-2% error, USD - kWh) - (2.01-15% error, USD - kWh)                 -0.0495 -0.10824 -0.007293
+#  (0.01-2% error, USD - kWh) - (Over 15% error, Percentage - kWh)          -0.0858 -0.17018  0.012997
+#  (0.01-2% error, USD - kWh) - (Over 15% error, USD - kWh)                 -0.1463 -0.25832 -0.026034
+#  (2.01-15% error, Percentage - kWh) - (2.01-15% error, USD - kWh)         -0.0080 -0.03343  0.004371
+#  (2.01-15% error, Percentage - kWh) - (Over 15% error, Percentage - kWh)  -0.0420 -0.11437  0.024139
+#  (2.01-15% error, Percentage - kWh) - (Over 15% error, USD - kWh)         -0.1032 -0.19343 -0.015885
+#  (2.01-15% error, USD - kWh) - (Over 15% error, Percentage - kWh)         -0.0326 -0.11179  0.042446
+#  (2.01-15% error, USD - kWh) - (Over 15% error, USD - kWh)                -0.0932 -0.17727 -0.012488
+#  (Over 15% error, Percentage - kWh) - (Over 15% error, USD - kWh)         -0.0599 -0.18057  0.053139
+
+emms_ref <- emmeans(ordinal_model_s2_logit_c1add_int3, ~ refClass, type = "link")
+contrast_ref <- contrast(emms_ref, "trt.vs.ctrl", ref = "kWh") 
+odds_ratios <- summary(contrast_ref, type = "response") # back-transform to odds
+odds_ratios
+#  contrast         odds.ratio lower.HPD upper.HPD
+#  Percentage / kWh        2.9      0.25      10.3
+#  USD / kWh               7.0      0.64      26.7
+
+# Results are averaged over the levels of: rounded, pct_goal, calc 
+# Point estimate displayed: median 
+# Results are back-transformed from the log odds ratio scale 
+# HPD interval probability: 0.95
+
+comparisons <- hypothesis(
+  ordinal_model_s2_logit_c1add_int3,
+  hypothesis = c(
+    "refClassPercentage - refClassUSD = 0",
+    "refClassPercentage = 0",
+    "refClassUSD = 0"
+  ),
+  class = "b"
+)
+
+pairs(emmeans(ordinal_model_s2_logit_c1add_int3, ~ refClass | rounded * pct_goal))
+
+```
+
+```{markdown}
+
+I need to generate and summarize publication quality group comparisons.
+Interpret the output from these initial analyses, and explain how I would report the effect of refClass in a manuscript, or if follow up analyses are necessary before we can confirm how best to report the effect.
+Also check the R code currently being used to ensure there are no mistakes.
+I mainly care about reporting the effect of refClass, but just don't want to make any mistakes in interpretation due to interactions.
+If additional analyses are necessary, provide R code for them.
+Also comment on if I should also assess group differences using odds ratios, and show R code to do this.
+
+pacman::p_load(dplyr,purrr,tidyr,stringr,here,tibble,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,ggplot2,gt,knitr,kableExtra,ggh4x,patchwork,lme4,flextable,pander)
+library(marginaleffects)
+library(tinytable)
+options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
+
+s2_agg <- s2_long |> 
+  filter(appliance != "TOTAL") |> 
+  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
+  summarise(
+    total_kWh = sum(value),
+    orig_kWh = sum(family),
+    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
+    state_dif = mean(state_dif),
+    .groups = "drop"
+  ) |>
+  mutate(
+    matched_goal = (pct_change == pct),
+    close_match = abs(pct_change - pct) <= 0.02,
+                error = pct_change - pct,
+                abs_error = abs(error),
+                log_abs_error=log(abs(error)+.007)) |> 
+      ungroup() |> 
+        mutate(
+            accuracy_level = factor(
+                case_when(
+                    abs_error == 0.00 ~ "Exact match",
                     abs_error <= 0.02 ~ "0.01-2% error",
@@
             )
-        ) |> relocate(accuracy_level, .after= "pct_change")
-
-s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
+        )
+
+s2_agg4 <- s2_agg |> group_by(id,refClass,calc) |> 
     mutate(n_accuracy = n_distinct(accuracy_level)) |> 
@@
 
-```
-
-```{r}
-#| label: tbl-s1-ord
-#| tbl-cap: "Study 1: Odds ratios of for group comparisons. "
-
-ordinal_model_s1 <- brm(
-    accuracy_level ~ refClass + (1|id) + (1|state),
-    data = s1_agg,
-    family = cumulative("logit"),
-    cores = 4,
-    iter = 5000,
-    save_pars = save_pars(all = TRUE),
-    control = list(adapt_delta = 0.99), 
-    prior = c(prior(normal(0, 3), class = "Intercept"), 
-                prior(normal(0, 3), class = "b")), 
-    file = paste0(here::here("data/model_cache",'s1_ordinal3.rds')) 
-)
-# 
-# ordinal_model_s1_null <- brm(
-#     accuracy_level ~  (1|id) + (1|state),
-#     data = s1_agg,
-#     family = cumulative("logit"),
-#     cores = 4,
-#     iter = 5000,
-#     save_pars = save_pars(all = TRUE),
-#     control = list(adapt_delta = 0.99), 
-#     prior = c(prior(normal(0, 3), class = "Intercept")),
-#     file = paste0(here::here("data/model_cache",'s1_null.rds')) 
-# )
-
-#colnames(mted1) <- c("Term", "Estimate","95% CrI Lower", "95% CrI Upper", "pd")
-
-as.data.frame(describe_posterior(ordinal_model_s1, centrality = "Mean"))[, c(1,2,4,5,6)] |> 
-  setNames(c("Parameter", "Estimate", "CI_Lower", "CI_Upper", "pd")) |> 
-  mutate(Parameter = stringr::str_remove(Parameter, "b_"))
-
-# ordinal_model_s1 |> emmeans(~refClass) |> contrast(method="pairwise") |>  
-#   gather_emmeans_draws() %>%
-#   mean_hdi(.width = .95)
-
-# describe_posterior(ordinal_model_s1, centrality = "Mode") |> 
-#     filter(stringr::str_detect(Parameter, "b_")) |> 
-#     mutate(Parameter = stringr::str_remove(Parameter, "b_")) |> 
-#     mutate(across(where(is.numeric), round, 3)) |> 
-#     kable(booktabs = TRUE)
-
-# Get predicted probabilities
-pred_summary <- ordinal_model_s1 |>
-    epred_draws(newdata = data.frame(refClass = c("kWh", "Percentage", "USD")),
-                ndraws = 1000, re_formula = NA) |>
-    group_by(refClass, Category=.category) |>
-    summarise(
-        mean_prob = mean(.epred),
-        lower_ci = quantile(.epred, 0.025),
-        upper_ci = quantile(.epred, 0.975)
-    )
-#pred_summary |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy")
-
-# Convert log-odds to odds ratios
-posterior_samples <- as.data.frame(ordinal_model_s1)
-odds_ratios <- data.frame(
-  Percentage_vs_kWh = exp(posterior_samples$b_refClassPercentage),
-  USD_vs_kWh = exp(posterior_samples$b_refClassUSD)
-)
-
-# Calculate summary statistics
-odds_ratio_summary <- data.frame(
-    comparison = c("Percentage vs kWh", "USD vs kWh"),
-    odds_ratio = c(mean(odds_ratios$Percentage_vs_kWh),
-                    mean(odds_ratios$USD_vs_kWh)),
-    ci_lower = c(quantile(odds_ratios$Percentage_vs_kWh, 0.025),
-                quantile(odds_ratios$USD_vs_kWh, 0.025)),
-    ci_upper = c(quantile(odds_ratios$Percentage_vs_kWh, 0.975),
-                quantile(odds_ratios$USD_vs_kWh, 0.975))
-)
-
-odds_ratio_summary |> kable(escape=FALSE,booktabs=TRUE,align=c("l")) 
-# |comparison        | odds_ratio| ci_lower| ci_upper|
-# |:-----------------|----------:|--------:|--------:|
-# |Percentage vs kWh |        2.3|     0.62|      6.2|
-# |USD vs kWh        |       14.1|     4.08|     38.0|
-
-#odds_ratio_summary |> pander::pandoc.table(caption="Study 1: Odds ratios of accuracy")
-
-# odds ratios of fixed effects
-as.data.frame(fixef(ordinal_model_s1)[,-2])|> as.data.frame() %>%
-    rownames_to_column(var = "Parameter") %>%
-    mutate(across(where(is.numeric), exp)) |>
-    filter(!stringr::str_detect(Parameter, "Intercept"))
-
-describe_posterior(ordinal_model_s1, centrality = "Median")[, c(1,2,4,5)] %>%
-  as_tibble() |> # remove rows with Intercept
-  mutate(across(where(is.numeric), exp)) |>
-  filter(!stringr::str_detect(Parameter, "Intercept")) |>
-  mutate(Parameter = stringr::str_remove(Parameter, "b_"))
-
-or_table <- data.frame(exp(fixef(ordinal_model_s1)))
-or_table |>
-  rownames_to_column(var = "Term") |>
-  as_flextable() |>
-  set_header_labels(values = list(
-    OR = "Odds Ratio",
-    `2.5 %` = "Lower 95% CI",
-    `97.5 %` = "Upper 95% CI"
-  )) |>
-  flextable::add_footer_lines(values = "Note: Odds ratios derived from the ordinal logistic regression model.") |> 
-  align(align = 'center', part = 'all') |>
-  fontsize(size = 10, part = "all")
-
-posterior_draws <- as_draws_df(ordinal_model_s1) |> select(starts_with("b_"))
-
-# Compute posterior summaries and exponentiate to get odds ratios
-or_summary <- posterior_draws %>%
-  summarise(across(starts_with("b_"),
-                   list(mean = ~mean(.x),
-                        sd = ~sd(.x),
-                        lower95 = ~quantile(.x, 0.025),
-                        upper95 = ~quantile(.x, 0.975)))) %>%
-  tidyr::pivot_longer(cols = everything(),
-                      names_to = c("parameter", ".value"),
-                      names_pattern = "(b_.*)_(.*)") %>%
-  mutate(odds_ratio = exp(mean),
-         lower95_or = exp(lower95),
-         upper95_or = exp(upper95))
-
-or_table <- or_summary %>%
-  select(parameter, odds_ratio, lower95_or, upper95_or) %>%
-  mutate(parameter = gsub("b_", "", parameter)) %>%
-  rename("Predictor" = parameter,
-         "OR" = odds_ratio,
-         "2.5% CI" = lower95_or,
-         "97.5% CI" = upper95_or) %>%
-  kbl(digits = 2, booktabs = TRUE, align = "lccc",
-      caption = "Odds Ratios for Ordinal Model") %>%
-  kableExtra::kable_styling(full_width = FALSE)
-
-print(or_table)
-
-```
-
-```{r}
-
-summary(ordinal_model_s1)
-
-pp_check(ordinal_model_s1, type = "bars_grouped", group="refClass", fatten = 2) +
-  scale_x_continuous("Response Category", breaks = 1:4, 
-            labels = c("Exact", "0.01-2%", "2.01-15%", ">15%")) +
-  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
-  ggtitle("Posterior Predictive Check - ordinal_model_s1") +
-  theme_minimal() +
-  theme(
-    legend.background = element_blank(),
-    legend.position = "bottom",
-    panel.grid.minor = element_blank(),
-    axis.text.x = element_text(angle = 45, hjust = 1)
-  )
-
-pp_check(ordinal_model_s1, type = "dens_overlay_grouped", group="refClass", fatten = 2,ndraws=100) +   scale_x_continuous("Response Category", breaks = 1:4, 
-            labels = c("Exact", "0.01-2%", "2.01-15%", ">15%")) +
-  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
-  ggtitle("Posterior Predictive Check - ordinal_model_s1") +
-  theme_minimal() +
-  theme(
-    legend.background = element_blank(),
-    legend.position = "bottom",
-    panel.grid.minor = element_blank(),
-    axis.text.x = element_text(angle = 45, hjust = 1)
-  )
-```
-
-```{r}
-
-multinomial_model_s1 <- brm(
-  accuracy_level ~ refClass + (1|id) + (1|state),
-  data = s1_agg,
-  family = categorical(), # Note: categorical() for multinomial
-  cores = 4,
-  iter = 1000,
-  save_pars = save_pars(all = TRUE),
-  control = list(adapt_delta = 0.99, max_treedepth=12),
-#   prior = c(prior(normal(0, 3), class = "Intercept"), 
-#             prior(normal(0, 3), class = "b")),
-  file = paste0(here::here("data/model_cache",'s1_multinomial.rds')) 
-)
-
-summary(multinomial_model_s1)
-pp_check(multinomial_model_s1, type = "bars_grouped", group="refClass", fatten = 2,ndraws=100) + labs(title = "Study 1: Posterior predictive check for the multinomial model") + theme_minimal()
-
-```
-
-```{r}
-#| eval: false
-
-loo_ordinal <- LOO(ordinal_model_s1, moment_match=TRUE)
-loo_multinomial <- LOO(multinomial_model_s1, moment_match=TRUE, reloo=TRUE)
-loo_compare(loo_ordinal, loo_multinomial)
-
-loo_ordinal <- LOO(ordinal_model_s1)
-loo_multinomial <- LOO(multinomial_model_s1)
-loo_compare(loo_ordinal, loo_multinomial)
-
-```
-
-## S1 exact match
-
-```{r}
-
-# Assuming s1_agg is your dataframe from Experiment 1, and it's already loaded
-
-# Fit a Bayesian logistic regression model
-exact_match_model <- brm(
-  matched_goal ~ refClass + (1|id) + (1|state),
-  data = s1_agg,
-  family = bernoulli(link = "logit"),
-  cores = 4,
-  iter = 2000,
-  control = list(adapt_delta = 0.97),
-  prior = c(
-    prior(normal(0, 2), class = "Intercept"),
-    prior(normal(0, 2), class = "b")
-  ),
-  file = paste0(here::here("data/model_cache", 's1_exact_match_model.rds'))
-)
-summary(exact_match_model)
-
-exact_match_model_c_add <- brm(
-  matched_goal ~ refClass + calc+ (1|id) + (1|state),
-  data = s1_agg,
-  family = bernoulli(link = "logit"),
-  cores = 4,
-  iter = 2000,
-  control = list(adapt_delta = 0.97),
-  prior = c(
-    prior(normal(0, 2), class = "Intercept"),
-    prior(normal(0, 2), class = "b")
-  ),
-  file = paste0(here::here("data/model_cache", 's1_exact_match_model_c_add.rds'))
-)
-summary(exact_match_model_c_add)
-
-exact_match_model_c_int <- brm(
-  matched_goal ~ refClass * calc+ (1|id) + (1|state),
-  data = s1_agg,
-  family = bernoulli(link = "logit"),
-  cores = 4,
-  iter = 2000,
-  control = list(adapt_delta = 0.97),
-  prior = c(
-    prior(normal(0, 2), class = "Intercept"),
-    prior(normal(0, 2), class = "b")
-  ),
-  file = paste0(here::here("data/model_cache", 's1_exact_match_model_c_int.rds'))
-)
-summary(exact_match_model_c_int )
-conditional_effects(exact_match_model_c_int)
-
-```
-
-## outliers
-
-```{r}
-
-identify_outliers_group <- function(x) {
-  Q1 <- quantile(x, 0.25, na.rm = TRUE)
-  Q3 <- quantile(x, 0.75, na.rm = TRUE)
-  IQR_value <- Q3 - Q1
-  lower_bound <- Q1 - 3.3 * IQR_value
-  upper_bound <- Q3 + 3.3 * IQR_value
-  return(x < lower_bound | x > upper_bound)
-}
-
-# 2. Apply the Outlier Detection to abs_error and log_abs_error within each group
-s1_agg <- s1_agg %>%
-  group_by(refClass, calc) %>%
-  mutate(
-    outlier_abs_error = identify_outliers_group(abs_error),
-    outlier_log_abs_error = identify_outliers_group(log_abs_error)
-  ) %>%
-  ungroup()
-
-#s1_agg |> relocate(outlier_abs_error,.after=abs_error) |> sort(desc(outlier_abs_error)) |> head(5)
-s1_agg |> filter(outlier_abs_error) |> select(id,refClass,calc,abs_error,log_abs_error) |> # keep unique id and refClass pairs
-  distinct(id,refClass)
-
-# 3. Summarize the Number and Proportion of Outliers by Group
-outlier_summary_by_group <- s1_agg %>%
-  group_by(refClass, calc) %>%
-  summarise(
-    total_observations = n(),
-    outliers_abs_error = sum(outlier_abs_error),
-    percent_outliers_abs_error = round((outliers_abs_error / total_observations) * 100, 2),
-    outliers_log_abs_error = sum(outlier_log_abs_error),
-    percent_outliers_log_abs_error = round((outliers_log_abs_error / total_observations) * 100, 2)
-  ) %>%
-  ungroup()
-
-# Display the outlier summary by group
-print("Outlier Summary by Reference Class and Calculator Usage:")
-print(outlier_summary_by_group)
-
-s1_agg_no_outliers <- s1_agg %>%
-  filter(!outlier_abs_error & !outlier_log_abs_error)
-
-# Re-run the exact match comparison without outliers
-
-# Create a contingency table of reference class vs. exact match without outliers
-exact_match_table_no_outliers <- s1_agg_no_outliers %>%
-  mutate(exact_match = ifelse(matched_goal, "Exact Match", "No Exact Match")) %>%
-  count(refClass, exact_match) %>%
-  pivot_wider(names_from = exact_match, values_from = n, values_fill = 0)
-
-# Perform Chi-Square Test of Independence without outliers
-chi_square_test_no_outliers <- chisq.test(exact_match_table_no_outliers[, -1])
-
-# Display results
-print("Chi-Square Test Results (Without Outliers):")
-print(chi_square_test_no_outliers)
-
-# Optionally, re-run logistic regression without outliers
-logistic_model_no_outliers <- glm(matched_goal ~ refClass, data = s1_agg_no_outliers, family = binomial)
-summary(logistic_model_no_outliers)
-
-# Calculate Odds Ratios
-odds_ratios_no_outliers <- exp(coef(logistic_model_no_outliers))
-print("Odds Ratios (Without Outliers):")
-print(odds_ratios_no_outliers)
-
-# Calculate 95% Confidence Intervals for Odds Ratios without outliers
-conf_intervals_no_outliers <- exp(confint(logistic_model_no_outliers))
-print("95% Confidence Intervals for Odds Ratios (Without Outliers):")
-print(conf_intervals_no_outliers)
-
-s1_clean <- s1_agg %>%
-  # Remove trials that are outliers in abs_error or log_abs_error
-  #filter(!(outlier_abs_error | outlier_log_abs_error))
-  # remove any id that has any outlier_abs_error
-  filter(!(id %in% (s1_agg |> filter(outlier_abs_error) |> select(id) |> distinct() |> pull(id))))
-
-length(unique(s1_agg$id)) 
-length(unique(s1_clean$id))
-
-# 2. Summary of Cleaned Data
-cleaned_summary <- s1_clean %>%
-  group_by(refClass, calc) %>%
-  summarise(
-    total_observations = n(),
-    exact_matches = sum(matched_goal),
-    percent_exact = round((exact_matches / total_observations) * 100, 2),
-    mean_abs_error = mean(abs_error, na.rm = TRUE),
-    mean_log_abs_error = mean(log_abs_error, na.rm = TRUE)
-  ) %>%
-  ungroup()
-
-# Display the cleaned data summary
-print("Cleaned Data Summary by Reference Class and Calculator Usage:")
-print(cleaned_summary)
-
-ordinal_model_s1_clean_calc <- brm(
-    accuracy_level ~ refClass +calc + (1|id) + (1|state),
-    data = s1_clean,
-    family = cumulative("logit"),
-    cores = 4,
-    iter = 2000,
-    save_pars = save_pars(all = TRUE),
-    control = list(adapt_delta = 0.97), 
-    prior = c(prior(normal(0, 3), class = "Intercept"), 
-                prior(normal(0, 3), class = "b")), 
-    file = paste0(here::here("data/model_cache",'s1_ordinal_clean_calc_3.3.rds')) 
-)
-
-summary(ordinal_model_s1_clean_calc)
-
-ordinal_model_s1_clean_calc_int <- brm(
-    accuracy_level ~ refClass * calc + (1|id) + (1|state),
-    data = s1_clean,
-    family = cumulative("logit"),
-    cores = 4,
-    iter = 2000,
-    save_pars = save_pars(all = TRUE),
-    control = list(adapt_delta = 0.97), 
-    prior = c(prior(normal(0, 3), class = "Intercept"), 
-                prior(normal(0, 3), class = "b")), 
-    file = paste0(here::here("data/model_cache",'s1_ordinal_clean_calc_int.rds')) 
-)
-
-summary(ordinal_model_s1_clean_calc_int)
-conditional_effects(ordinal_model_s1_clean_calc_int)
-
-# logistic regression on exact match with clean data (frequentist)
-logistic_model_clean <- glm(matched_goal ~ refClass, data = s1_clean, family = binomial)
-summary(logistic_model_clean)
-
-# logistic regression on exact match with clean data (frequentist) - mixed effects
-logistic_model_clean_mixed <- glmer(matched_goal ~ refClass + (1|id) + (1|state), data = s1_clean, family = binomial)
-summary(logistic_model_clean_mixed)
-
-# logistic regression on exact match with clean data (frequentist) - mixed effects
-logistic_model_clean_mixed <- glmer(matched_goal ~ refClass+calc + (1|id) , data = s1_clean, family = binomial)
-summary(logistic_model_clean_mixed)
-
-# logistic regression on exact match with clean data (bayesian)
-logistic_model_clean_bayes <- brm(
-  matched_goal ~ refClass + (1|id) + (1|state),
-  data = s1_clean,
-  family = bernoulli(link = "logit"),
-  cores = 4,
-  iter = 2000,
-  control = list(adapt_delta = 0.97),
-  prior = c(
-    prior(normal(0, 2), class = "Intercept"),
-    prior(normal(0, 2), class = "b")
-  ),
-  file = paste0(here::here("data/model_cache", 's1_exact_match_model_clean.rds'))
-)
-summary(logistic_model_clean_bayes)
-conditional_effects(logistic_model_clean_bayes)
-
-```
-
-## outlier from avg.
-
-```{r}
-
-group_stats <- s1_agg |>
-  group_by(refClass, calc) |>
-  summarise(
-    group_mean_abs_error = mean(abs_error, na.rm = TRUE),
-    group_sd_abs_error = sd(abs_error, na.rm = TRUE),
-    .groups = "drop"
-  )
-
-# Calculate participant-level average abs_error
-participant_stats <- s1_agg |>
-  group_by(id, refClass, calc) |>
-  summarise(
-    avg_abs_error = mean(abs_error, na.rm = TRUE),
-    .groups = "drop"
-  )
-
-# Join group stats with participant stats
-participant_stats <- participant_stats |>
-  left_join(group_stats, by = c("refClass", "calc"))
-
-# Identify participants to exclude based on the 2.5 SD threshold
-participants_to_exclude <- participant_stats |>
-  mutate(
-    exclusion_threshold = group_mean_abs_error + 2.5 * group_sd_abs_error,
-    exclude = avg_abs_error > exclusion_threshold
-  ) |>
-  filter(exclude)
-
-ordinal_model_s1_clean_calc <- brm(
-    accuracy_level ~ refClass +calc + (1|id) + (1|state),
-    data = s1_clean,
-    family = cumulative("logit"),
-    cores = 4,
-    iter = 2000,
-    save_pars = save_pars(all = TRUE),
-    control = list(adapt_delta = 0.97), 
-    prior = c(prior(normal(0, 3), class = "Intercept"), 
-                prior(normal(0, 3), class = "b")), 
-    file = paste0(here::here("data/model_cache",'s1_ordinal_clean_calc_3.3.rds')) 
-)
-
-s1_agg_filtered <- s1_agg |>
-  filter(!(id %in% participants_to_exclude$id))
-
-length(unique(s1_agg$id)) 
-length(unique(s1_agg_filtered$id))
-
-ordinal_model_s1_clean_calc <- brm(
-    accuracy_level ~ refClass +calc + (1|id) + (1|state),
-    data = s1_agg_filtered,
-    family = cumulative("logit"),
-    cores = 4,
-    iter = 2000,
-    control = list(adapt_delta = 0.97), 
-    prior = c(prior(normal(0, 3), class = "Intercept"), 
-                prior(normal(0, 3), class = "b")), 
-    file = paste0(here::here("data/model_cache",'s1_ordinal_grpclean_calc_2.5.rds')) 
-)
-
-summary(ordinal_model_s1_clean_calc)
-conditional_effects(ordinal_model_s1_clean_calc)
-
-mixed_model_abs_error <- lmer(log_abs_error ~ refClass + calc+ (1|id) + (1|state), data = s1_agg_filtered)
-summary(mixed_model_abs_error)
-# with p value
-library(lmerTest)
-mixed_model_abs_error <- lmer(log_abs_error ~ refClass + calc+ (1|id) + (1|state), data = s1_agg_filtered)
-summary(mixed_model_abs_error)
-
-```
-
-### clearn log error
-```{r}
-
-s1_clean |> 
-  ggplot(aes(x = abs_error)) +
-  geom_density(fill = "skyblue", alpha = 0.7) +
-  facet_wrap(~ refClass, scales = "free") +ggtitle("Error density by reference class")
-
-# plot density of log-transformed data
-s1_clean |> 
-  ggplot(aes(x = log_abs_error)) +
-  geom_density(fill = "skyblue", alpha = 0.7) +
-  facet_wrap(~ refClass, scales = "free") + ggtitle("Log-transformed error density by reference class")
-
-# mixed effects regression on abs_error (frequentist)
-mixed_model_abs_error <- lmer(abs_error ~ refClass + (1|id) + (1|state), data = s1_clean)
-summary(mixed_model_abs_error)
-
-mixed_model_abs_error <- lmer(abs_error ~ refClass*calc + (1|id) + (1|state), data = s1_clean)
-summary(mixed_model_abs_error)
-
-```
-
-# S2
-
-```{r}
-#| label: tbl-s2-agg
-#| tbl-cap: "Study 2: Summary of planning accuracy by reference class.
-The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."
-
-s2_agg1 <- s2_long |> 
-  filter(appliance != "TOTAL") |> 
-  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
-  summarise(
-    total_kWh = sum(value),
-    orig_kWh = sum(family),
-    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
-    state_dif = mean(state_dif),
-    .groups = "drop"
-  ) |>
-  mutate(matched_goal = (pct_change == pct),
-                error = pct_change - pct,
-                abs_error = abs(error)) |> 
-      ungroup() |> # Add ungroup here
-        mutate(
-            accuracy_level = factor(
-                case_when(
-                    abs_error == 0.00 ~ "Exact match",
-                    abs_error <= 0.02 ~ "0.01-2% error",
-                    abs_error <= 0.15 ~ "2.01-15% error",
-                    TRUE ~ "Over 15% error"  # Capture all remaining cases
-                ), 
-                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
-                ordered = TRUE
-            )
-        )
-
-s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc) |> 
-    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
-    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),mean_abs_error=mean(abs_error),n_accuracy=first(n_accuracy)) |> 
-    mutate(accuracy_level = factor(
-            case_when(
-                mean_abs_error < 0.02 ~ "Exact match",
-                mean_abs_error <= 0.02 ~ "0.01-2% error",
-                mean_abs_error <= 0.15 ~ "2.01-15% error",
-                TRUE ~ "Over 15% error"  # Capture all remaining cases
-            ), 
-            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
-            ordered = TRUE
-        ))
-
-# overall pct of subjects who matched their goal
-s2_agg4 |> group_by(refClass) |>
-    summarise(
-    'Avg. % meeting goal' = mean(pct),
-    'Avg. Deviation From Goal' = median(mean_pct_change),
-    # sd = sd(pct),
-    # n = n(),
-    #se=sd(pct)/sqrt(n)
-) |>   mutate(across(where(is.numeric), \(x) round(x, 3))) %>% 
-  kable(escape=FALSE,booktabs=TRUE,align=c("l"))
-
-```
-
-```{r}
-
-ordinal_model_s2_logit <- brm(
-  accuracy_level ~ refClass +rounded+pct_goal + (1|id)+ (1|state),
-  data = s2_agg1,
-  family = cumulative("logit"),
-  cores = 4,
-  iter = 2000,
-  control = list(adapt_delta = 0.99), # Recommended for ordinal models
-  prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
-            prior(normal(0, 2), class = "b")), # Priors for predictors
-  file = paste0(here::here("data/model_cache",'s2_logit_add.rds')) # Cache for efficiency
-)
-summary(ordinal_model_s2_logit)
-
-pred_summary_s2 <- ordinal_model_s2_logit %>%
-  epred_draws(newdata = s2_agg1, re_formula = NA,ndraws=200) %>%
-  group_by("Reference Class"=refClass, rounded, "% Goal"=pct_goal, Category=.category) %>%
-  summarise(
-    mean_prob = mean(.epred),
-    lower_ci = quantile(.epred, 0.025),
-    upper_ci = quantile(.epred, 0.975),
-    .groups = "drop"
-  )
-
-# Create a new data frame with the predictor values
-new_data <- expand.grid(
-  refClass = c("kWh", "Percentage", "USD"),
-  id = unique(s2_agg1$id)[1],      # Take first id
-  state = unique(s2_agg1$state)[1], # Take first state,
-  rounded = unique(s2_agg1$rounded)[1], # Take first rounded
-  pct_goal = unique(s2_agg1$pct_goal)[1] # Take first pct_goal
-)
-
-# Generate predicted probabilities
-predicted_probabilities <- predict(ordinal_model_s2_logit, newdata = new_data, 
-                                  summary = FALSE, ndraws = 1000)
-
-# Calculate the predicted probability of "Exact match"
-predicted_probability_exact_match <- mean(predicted_probabilities <= 1)
-
-print(predicted_probability_exact_match)
-
-pred_summary_s2 |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy level by reference class",split.table=Inf)
-
-# Extract posterior samples
-posterior_samples_s2 <- as.data.frame(ordinal_model_s2_logit)
-
-odds_ratios_s2 <- posterior_samples_s2 %>%
-    transmute(
-        refClass_Percentage_vs_kWh = exp(b_refClassPercentage),
-        refClass_USD_vs_kWh = exp(b_refClassUSD),
-        rounded_Yes_vs_No = exp(b_roundedRounded),
-        pct_goal_15_vs_10 = exp(`b_pct_goal15%`)  # Note the backticks here
-    )
-
-# Calculate summary statistics
-odds_ratio_summary_s2 <- data.frame(
-    comparison = c("Percentage vs kWh", "USD vs kWh", "Rounded vs Not", "15% Goal vs 10% Goal"),
-    odds_ratio = c(
-        mean(odds_ratios_s2$refClass_Percentage_vs_kWh),
-        mean(odds_ratios_s2$refClass_USD_vs_kWh),
-        mean(odds_ratios_s2$rounded_Yes_vs_No),
-        mean(odds_ratios_s2$pct_goal_15_vs_10)
-    ),
-    ci_lower = c(
-        quantile(odds_ratios_s2$refClass_Percentage_vs_kWh, 0.025),
-        quantile(odds_ratios_s2$refClass_USD_vs_kWh, 0.025),
-        quantile(odds_ratios_s2$rounded_Yes_vs_No, 0.025),
-        quantile(odds_ratios_s2$pct_goal_15_vs_10, 0.025)
-    ),
-    ci_upper = c(
-        quantile(odds_ratios_s2$refClass_Percentage_vs_kWh, 0.975),
-        quantile(odds_ratios_s2$refClass_USD_vs_kWh, 0.975),
-        quantile(odds_ratios_s2$rounded_Yes_vs_No, 0.975),
-        quantile(odds_ratios_s2$pct_goal_15_vs_10, 0.975)
-    )
-)
-
-#             comparison odds_ratio ci_lower ci_upper
-# 1    Percentage vs kWh       1.99     0.73     4.43
-# 2           USD vs kWh       3.69     1.33     8.29
-# 3       Rounded vs Not       0.70     0.57     0.84
-# 4 15% Goal vs 10% Goal       0.78     0.63     0.94
-
-odds_ratio_summary_s2 |> kable()
-
-odds_ratio_summary_s2 |> pander::pandoc.table(caption="Study 2: Odds ratios of accuracy")
-
-# Create a new data frame with the predictor values
-# new_data <- expand.grid(
-#   refClass = c("kWh", "Percentage", "USD"),
-#   id = unique(s2_agg1$id)[1],      # Take first id
-#   state = unique(s2_agg1$state)[1], # Take first state,
-#   rounded = unique(s2_agg1$rounded)[1], # Take first rounded
-#   pct_goal = unique(s2_agg1$pct_goal)[1] # Take first pct_goal
-# )
-
-# or_table <- data.frame(exp(fixef(ordinal_model_s2_logit)))
-# or_table |>
-#   rownames_to_column(var = "Term") |>
-#   filter(!stringr::str_detect(Term, "Intercept")) |>
-#   as_flextable() |>
-#   set_header_labels(values = list(
-#     OR = "Odds Ratio",
-#     `2.5 %` = "Lower 95% CI",
-#     `97.5 %` = "Upper 95% CI"
-#   )) |>
-#   flextable::add_footer_lines(values = "Note: Odds ratios derived from the ordinal logistic regression model.") |> 
-#   align(align = 'center', part = 'all') |>
-#   fontsize(size = 10, part = "all")
-
-```
-
-# Interaction s2
-
-```{r}
-
-ordinal_model_s2_logit_int2 <- brm(
-  accuracy_level ~ refClass +(rounded*pct_goal) + (1|id)+ (1|state),
-  data = s2_agg1,
-  family = cumulative("logit"),
-  cores = 4,
-  iter = 2000,
-  control = list(adapt_delta = 0.99), # Recommended for ordinal models
-  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
-            prior(normal(0, 3), class = "b")), # Priors for predictors
-  file = paste0(here::here("data/model_cache",'s2_logit_int2.rds')) # Cache for efficiency
-)
-summary(ordinal_model_s2_logit_int2)
-
-conditional_effects(ordinal_model_s2_logit_int2)
-
-ordinal_model_s2_logit_int3 <- brm(
-  accuracy_level ~ refClass * rounded*pct_goal + (1|id)+ (1|state),
-  data = s2_agg1,
-  family = cumulative("logit"),
-  cores = 4,
-  iter = 3000,
-  control = list(adapt_delta = 0.99), # Recommended for ordinal models
-  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
-            prior(normal(0, 3), class = "b")), # Priors for predictors
-  file = paste0(here::here("data/model_cache",'s2_logit_int3.rds')) # Cache for efficiency
-)
-summary(ordinal_model_s2_logit_int3)
-
-conditional_effects(ordinal_model_s2_logit_int3)
-
-ordinal_model_s2_logit_c1rf_int3<- brm(
-  accuracy_level ~ refClass * rounded*pct_goal + (1|id)+ (1|state) + (1|calc),
-  data = s2_agg1,
-  family = cumulative("logit"),
-  cores = 4,
-  iter = 3000,
-  control = list(adapt_delta = 0.99), # Recommended for ordinal models
-  prior = c(prior(normal(0, 3), class = "Intercept"),  # Priors for thresholds
-            prior(normal(0, 3), class = "b")), # Priors for predictors
-  file = paste0(here::here("data/model_cache",'s2_logit_c1rf_int3.rds')) # Cache for efficiency
-)
-summary(ordinal_model_s2_logit_c1rf_int3)
-
-conditional_effects(ordinal_model_s2_logit_c1rf_int3)
-
 ordinal_model_s2_logit_c1add_int3<- brm(
       accuracy_level ~ (refClass * rounded*pct_goal) + calc + (1|id)+ (1|state),
-  data = s2_agg1,
+      data = s2_agg,
       family = cumulative("logit"),
@@
     summary(ordinal_model_s2_logit_c1add_int3)
-
-conditional_effects(ordinal_model_s2_logit_c1add_int3)
+ Family: cumulative 
+  Links: mu = logit; disc = identity 
+Formula: accuracy_level ~ (refClass * rounded * pct_goal) + calc + (1 | id) + (1 | state) 
+   Data: s2_agg (Number of observations: 784) 
+  Draws: 4 chains, each with iter = 3000; warmup = 1500; thin = 1;
+         total post-warmup draws = 6000
+
+Multilevel Hyperparameters:
+~id (Number of levels: 196) 
+              Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+sd(Intercept)     4.41      0.37     3.73     5.18 1.00     1382     2871
+
+~state (Number of levels: 4) 
+              Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+sd(Intercept)     0.53      0.47     0.06     1.80 1.00     1046      864
+
+Regression Coefficients:
+                                              Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+Intercept[1]                                     -1.80      0.80    -3.38    -0.27 1.00      814     1644
+Intercept[2]                                     -0.26      0.79    -1.81     1.30 1.00      831     1824
+Intercept[3]                                      3.59      0.80     2.03     5.17 1.00      854     1713
+refClassPercentage                                1.15      0.91    -0.68     2.89 1.00      620     1179
+refClassUSD                                       1.68      0.95    -0.19     3.55 1.00      725     1126
+roundedRounded                                   -1.13      0.74    -2.60     0.30 1.00      763     1402
+pct_goal15%                                      -0.89      0.74    -2.39     0.54 1.00      764     1342
+calcNoCalculator                                  3.83      0.87     2.20     5.53 1.00      722     1685
+refClassPercentage:roundedRounded                -0.35      0.99    -2.28     1.55 1.00      874     1723
+refClassUSD:roundedRounded                        0.28      1.02    -1.74     2.24 1.00     1006     1762
+refClassPercentage:pct_goal15%                   -0.10      0.98    -2.04     1.82 1.00      874     1755
+refClassUSD:pct_goal15%                          -0.05      1.01    -2.02     1.91 1.00      994     1906
+roundedRounded:pct_goal15%                        0.54      1.37    -2.07     3.26 1.00      698     1244
+refClassPercentage:roundedRounded:pct_goal15%     0.60      1.83    -3.02     4.16 1.00      805     1526
+refClassUSD:roundedRounded:pct_goal15%            0.62      1.88    -3.04     4.32 1.00      897     1565
+
+Further Distributional Parameters:
+     Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+disc     1.00      0.00     1.00     1.00   NA       NA       NA
+
+Draws were sampled using sampling(NUTS). For each parameter, Bulk_ESS
+and Tail_ESS are effective sample size measures, and Rhat is the potential
+scale reduction factor on split chains (at convergence, Rhat = 1).
+
+ame <- avg_slopes(
+  ordinal_model_s2_logit_c1add_int3, 
+  variables = "refClass"
+)
+print(ame)
+          Group         Contrast Estimate    2.5 %   97.5 %
+ Exact match    Percentage - kWh -0.08116 -0.19571  0.03058
+ Exact match    USD - kWh        -0.14031 -0.25760 -0.02405
+ 0.01-2% error  Percentage - kWh -0.00626 -0.02044  0.00544
+ 0.01-2% error  USD - kWh        -0.01792 -0.03998 -0.00195
+ 2.01-15% error Percentage - kWh  0.02066 -0.00691  0.06562
+ 2.01-15% error USD - kWh         0.03031  0.00110  0.07913
+ Over 15% error Percentage - kWh  0.06622 -0.03016  0.15240
+ Over 15% error USD - kWh         0.12808  0.02140  0.22340
+
+ame_by_rounded <- avg_slopes(
+  ordinal_model_s2_logit_c1add_int3,
+  variables = "refClass",
+  by = c("rounded", "accuracy_level")
+)
+print(ame_by_rounded)
+
+          Group     Term         Contrast     rounded Estimate     2.5 %    97.5 %
+ Exact match    refClass Percentage - kWh Not Rounded -0.07940 -0.194010  0.031746
+ Exact match    refClass USD - kWh        Not Rounded -0.11370 -0.231058  0.000454
+ 0.01-2% error  refClass Percentage - kWh Not Rounded -0.00865 -0.026939  0.006926
+ 0.01-2% error  refClass USD - kWh        Not Rounded -0.01731 -0.043440  0.000510
+ 2.01-15% error refClass Percentage - kWh Not Rounded  0.01113 -0.008613  0.055578
+ 2.01-15% error refClass USD - kWh        Not Rounded  0.01403 -0.007740  0.060557
+ Over 15% error refClass Percentage - kWh Not Rounded  0.07524 -0.033826  0.173807
+ Over 15% error refClass USD - kWh        Not Rounded  0.11672 -0.000602  0.219258
+ Exact match    refClass Percentage - kWh Rounded     -0.08331 -0.208087  0.038742
+ Exact match    refClass USD - kWh        Rounded     -0.16720 -0.295255 -0.040737
+ 0.01-2% error  refClass Percentage - kWh Rounded     -0.00371 -0.018239  0.008463
+ 0.01-2% error  refClass USD - kWh        Rounded     -0.01848 -0.041158 -0.001891
+ 2.01-15% error refClass Percentage - kWh Rounded      0.02904 -0.010886  0.083676
+ 2.01-15% error refClass USD - kWh        Rounded      0.04645  0.005512  0.101707
+ Over 15% error refClass Percentage - kWh Rounded      0.05785 -0.033218  0.141438
+ Over 15% error refClass USD - kWh        Rounded      0.13897  0.034533  0.238851
+
+Type:  response 
+Columns: term, group, contrast, rounded, estimate, conf.low, conf.high
+
+ame_by_rounded_pct_goal <- avg_slopes(
+  ordinal_model_s2_logit_c1add_int3,
+  variables = "refClass",
+  by = c("rounded", "pct_goal", "accuracy_level")
+)
+
+ame_by_rounded_pct_goal |> head(n=25)
+
+          Group     Term         Contrast     rounded pct_goal  Estimate    2.5 %  97.5 %
+ Exact match    refClass Percentage - kWh Not Rounded      10% -0.079963 -0.21508 0.04389
+ Exact match    refClass USD - kWh        Not Rounded      10% -0.108881 -0.25134 0.01174
+ 0.01-2% error  refClass Percentage - kWh Not Rounded      10% -0.015302 -0.04340 0.01369
+ 0.01-2% error  refClass USD - kWh        Not Rounded      10% -0.026016 -0.06555 0.00249
+ 2.01-15% error refClass Percentage - kWh Not Rounded      10%  0.010750 -0.01803 0.07339
+ 2.01-15% error refClass USD - kWh        Not Rounded      10%  0.009697 -0.02100 0.07535
+ Over 15% error refClass Percentage - kWh Not Rounded      10%  0.082534 -0.05128 0.20363
+ Over 15% error refClass USD - kWh        Not Rounded      10%  0.124906 -0.01467 0.25374
+ Exact match    refClass Percentage - kWh Not Rounded      15% -0.075152 -0.20930 0.05528
+ Exact match    refClass USD - kWh        Not Rounded      15% -0.116988 -0.25518 0.02221
+ 0.01-2% error  refClass Percentage - kWh Not Rounded      15% -0.001370 -0.02112 0.00972
+ 0.01-2% error  refClass USD - kWh        Not Rounded      15% -0.007259 -0.03795 0.00572
+ 2.01-15% error refClass Percentage - kWh Not Rounded      15%  0.008287 -0.01072 0.05789
+ 2.01-15% error refClass USD - kWh        Not Rounded      15%  0.014535 -0.00597 0.07078
+ Over 15% error refClass Percentage - kWh Not Rounded      15%  0.067139 -0.05214 0.18279
+ Over 15% error refClass USD - kWh        Not Rounded      15%  0.109662 -0.02264 0.22915
+ Exact match    refClass Percentage - kWh Rounded          10% -0.059312 -0.19493 0.07424
+ Exact match    refClass USD - kWh        Rounded          10% -0.139822 -0.28116 0.00435
+ 0.01-2% error  refClass Percentage - kWh Rounded          10% -0.000549 -0.01753 0.01077
+ 0.01-2% error  refClass USD - kWh        Rounded          10% -0.010150 -0.03797 0.00555
+ 2.01-15% error refClass Percentage - kWh Rounded          10%  0.008692 -0.01282 0.05727
+ 2.01-15% error refClass USD - kWh        Rounded          10%  0.020893 -0.00579 0.07696
+ Over 15% error refClass Percentage - kWh Rounded          10%  0.049498 -0.06487 0.15795
+ Over 15% error refClass USD - kWh        Rounded          10%  0.127970 -0.00412 0.24646
+ Exact match    refClass Percentage - kWh Rounded          15% -0.109285 -0.26855 0.04866
+
+conditional_slopes <- slopes(
+    ordinal_model_s2_logit_c1add_int3,
+    variables = "refClass",
+    newdata = datagrid(
+        rounded = c('Not Rounded', 'Rounded'),
+        pct_goal = c("10%", "15%")
+    )
+)conditional_slopes |> head(n=25)
+
+          Group         Contrast     rounded pct_goal Estimate   2.5 %    97.5 %
+ Exact match    Percentage - kWh Not Rounded      10% -0.00105 -0.0183  0.001659
+ Exact match    Percentage - kWh Not Rounded      15% -0.00219 -0.0281  0.006496
+ Exact match    Percentage - kWh Rounded          10% -0.00215 -0.0317  0.014072
+ Exact match    Percentage - kWh Rounded          15% -0.00486 -0.0806  0.005323
+ Exact match    USD - kWh        Not Rounded      10% -0.00134 -0.0200  0.000222
+ Exact match    USD - kWh        Not Rounded      15% -0.00307 -0.0343  0.001715
+ Exact match    USD - kWh        Rounded          10% -0.00443 -0.0436  0.000191
+ Exact match    USD - kWh        Rounded          15% -0.00712 -0.0949 -0.000324
+ 0.01-2% error  Percentage - kWh Not Rounded      10% -0.00377 -0.0568  0.005781
+ 0.01-2% error  Percentage - kWh Not Rounded      15% -0.00778 -0.0800  0.021370
+ 0.01-2% error  Percentage - kWh Rounded          10% -0.00781 -0.0853  0.041417
+ 0.01-2% error  Percentage - kWh Rounded          15% -0.01713 -0.1707  0.016385
+ 0.01-2% error  USD - kWh        Not Rounded      10% -0.00482 -0.0621  0.000790
+ 0.01-2% error  USD - kWh        Not Rounded      15% -0.01099 -0.0970  0.005264
+ 0.01-2% error  USD - kWh        Rounded          10% -0.01575 -0.1193  0.000636
+ 0.01-2% error  USD - kWh        Rounded          15% -0.02477 -0.2094 -0.001266
+ 2.01-15% error Percentage - kWh Not Rounded      10% -0.12837 -0.4516  0.111235
+ 2.01-15% error Percentage - kWh Not Rounded      15% -0.15816 -0.4870  0.134962
+ 2.01-15% error Percentage - kWh Rounded          10% -0.12110 -0.4636  0.173061
+ 2.01-15% error Percentage - kWh Rounded          15% -0.18818 -0.5295  0.145839
+ 2.01-15% error USD - kWh        Not Rounded      10% -0.17215 -0.5106  0.023655
+ 2.01-15% error USD - kWh        Not Rounded      15% -0.23948 -0.5514  0.057446
+ 2.01-15% error USD - kWh        Rounded          10% -0.29650 -0.5931  0.028064
+ 2.01-15% error USD - kWh        Rounded          15% -0.36699 -0.6595  0.027229
+ Over 15% error Percentage - kWh Not Rounded      10%  0.13661 -0.1219  0.497865
+
+Term: refClass
+Type:  response 
+Columns: rowid, term, group, contrast, estimate, conf.low, conf.high, rounded, pct_goal, predicted_lo, predicted_hi, predicted, tmp_idx, refClass, calc, id, state, accuracy_level
+
+avg_slopes(
+        ordinal_model_s2_logit_c1add_int3,
+        variables = "refClass",
+        hypothesis = "b1 - b2 = 0"  # Compare effects for Rounded vs Not Rounded
+    )
+
+ Estimate   2.5 % 97.5 %
+   0.0575 -0.0489  0.171
+
+Term: b1-b2=0
+Type:  response 
+Columns: term, estimate, conf.low, conf.high
+
+avg_slopes(
+        ordinal_model_s2_logit_c1add_int3,
+        variables = "refClass",
+        hypothesis = "b1 - b3 = 0"  # Compare effects for Rounded vs Not Rounded
+    )
+
+ Estimate  2.5 % 97.5 %
+  -0.0738 -0.185 0.0284
+
+Term: b1-b3=0
+
+pairwise_comparisons <- avg_slopes(
+  ordinal_model_s2_logit_c1add_int3,
+  variables = "refClass",
+  hypothesis = "pairwise"
+)
+print(pairwise_comparisons)
 
 ```
+
+------
+
+"We employed a Bayesian ordinal logistic regression to evaluate the influence of reference information format (kWh, Percentage, USD) on the accuracy of planned energy savings, adjusting for the effects of goal rounding (Not Rounded, Rounded) and savings goal magnitude (10%, 15%).
+Furthermore, we incorporated random intercepts for participant (id) and state to account for potential clustering within these groups.
+Although preliminary analyses suggested that providing energy savings information in USD or percentage formats might alter the likelihood of achieving higher accuracy levels relative to kWh, these findings were not consistently supported when examining average marginal effects.
+Specifically, the average marginal effects analysis revealed that the USD format, compared to kWh, was associated with a decreased probability of an exact match (AME = -0.14, 95% CI [-0.26, -0.02]) and an increased probability of errors exceeding 15% (AME = 0.13, 95% CI [0.02, 0.22]).
+The percentage format exhibited less pronounced effects on accuracy compared to kWh.
+Moreover, the impact of reference information format appeared to be moderated by whether the savings goal was rounded and the magnitude of the goal.
+For instance, in scenarios where the goal was not rounded, the USD format was associated with an increased likelihood of errors over 15% to a greater extent than when the goal was rounded (AME = 0.12, 95% CI [0.00, 0.22] for not rounded vs.
+AME = 0.14, 95% CI [0.03, 0.24] for rounded).
+These results suggest that the effectiveness of different information formats in promoting accurate energy-saving plans may depend on the context in which the information is presented, such as the specificity and magnitude of the savings target.
+Nevertheless, it is important to interpret these findings with caution, as the 95% credible intervals indicate uncertainty regarding the magnitude and, in some cases, the direction of these effects."
+
+---------
+
+https://chatgpt.com/c/67613771-d628-8006-8af9-f57e6dedb71f
+
+```{markdown}
+
+simpler_model <- brm(
+  accuracy_level ~ refClass + rounded + pct_goal + calc + 
+    refClass:rounded + refClass:pct_goal + rounded:pct_goal + 
+    (1|id) + (1|state),
+  data = s2_agg,
+  family = cumulative("logit"),
+  cores = 4,
+  iter = 3000,
+  control = list(adapt_delta = 0.97),
+  prior = c(
+    prior(normal(0, 3), class = "Intercept"),
+    prior(normal(0, 3), class = "b")
+  )
+)
+<!-- summary(simpler_model)
+Family: cumulative 
+  Links: mu = logit; disc = identity 
+Formula: accuracy_level ~ refClass + rounded + pct_goal + calc + refClass:rounded + refClass:pct_goal + rounded:pct_goal + (1 | id) + (1 | state) 
+   Data: s2_agg (Number of observations: 784) 
+  Draws: 4 chains, each with iter = 3000; warmup = 1500; thin = 1;
+         total post-warmup draws = 6000
+
+Multilevel Hyperparameters:
+~id (Number of levels: 196) 
+              Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+sd(Intercept)     4.39      0.39     3.69     5.22 1.00     1498     2518
+
+~state (Number of levels: 4) 
+              Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+sd(Intercept)     0.52      0.42     0.05     1.66 1.00     1774     1902
+
+Regression Coefficients:
+                                  Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+Intercept[1]                         -1.78      0.78    -3.31    -0.25 1.00     1339     2219
+Intercept[2]                         -0.25      0.77    -1.72     1.29 1.00     1309     2201
+Intercept[3]                          3.60      0.79     2.10     5.18 1.00     1320     2089
+refClassPercentage                    1.01      0.86    -0.62     2.68 1.01     1010     1856
+refClassUSD                           1.59      0.88    -0.10     3.30 1.00     1207     2133
+roundedRounded                       -1.22      0.68    -2.55     0.10 1.00     1253     2330
+pct_goal15%                          -0.99      0.67    -2.33     0.35 1.00     1176     1954
+calcNoCalculator                      3.85      0.86     2.20     5.57 1.00     1210     2203
+refClassPercentage:roundedRounded    -0.07      0.45    -0.95     0.82 1.00     5545     4279
+refClassUSD:roundedRounded            0.56      0.47    -0.36     1.46 1.00     5550     4506
+refClassPercentage:pct_goal15%        0.19      0.44    -0.67     1.06 1.00     5959     4719
+refClassUSD:pct_goal15%               0.23      0.46    -0.66     1.12 1.00     5876     4387
+roundedRounded:pct_goal15%            0.78      1.18    -1.51     3.08 1.00     1019     1790 -->
+
+ame <- avg_slopes(
+  ordinal_model_s2_logit_c1add_int3, 
+  variables = "refClass"
+)
+print(ame)
+          Group         Contrast Estimate    2.5 %   97.5 %
+ Exact match    Percentage - kWh -0.07753 -0.20193  0.03311
+ Exact match    USD - kWh        -0.13708 -0.26083 -0.02975
+ 0.01-2% error  Percentage - kWh -0.00573 -0.02137  0.00450
+ 0.01-2% error  USD - kWh        -0.01786 -0.04062 -0.00192
+ 2.01-15% error Percentage - kWh  0.01551 -0.00586  0.06359
+ 2.01-15% error USD - kWh         0.02335 -0.00131  0.07491
+ Over 15% error Percentage - kWh  0.06779 -0.03259  0.16329
+ Over 15% error USD - kWh         0.13137  0.02931  0.23141
+
+ame_by_rounded <- avg_slopes(
+  ordinal_model_s2_logit_c1add_int3,
+  variables = "refClass",
+  by = c("rounded", "accuracy_level")
+)
+print(ame_by_rounded)
+
+         Group     Term         Contrast     rounded Estimate    2.5 %    97.5 %
+ Exact match    refClass Percentage - kWh Not Rounded -0.07630 -0.19556  0.030459
+ Exact match    refClass USD - kWh        Not Rounded -0.11168 -0.23341 -0.006969
+ 0.01-2% error  refClass Percentage - kWh Not Rounded -0.00793 -0.02715  0.006268
+ 0.01-2% error  refClass USD - kWh        Not Rounded -0.01665 -0.04273  0.000822
+ 2.01-15% error refClass Percentage - kWh Not Rounded  0.00835 -0.00729  0.053450
+ 2.01-15% error refClass USD - kWh        Not Rounded  0.00946 -0.00905  0.057332
+ Over 15% error refClass Percentage - kWh Not Rounded  0.07577 -0.03431  0.179538
+ Over 15% error refClass USD - kWh        Not Rounded  0.11889  0.00627  0.227794
+ Exact match    refClass Percentage - kWh Rounded     -0.07895 -0.21691  0.044418
+ Exact match    refClass USD - kWh        Rounded     -0.16185 -0.29451 -0.043900
+ 0.01-2% error  refClass Percentage - kWh Rounded     -0.00357 -0.01917  0.007500
+ 0.01-2% error  refClass USD - kWh        Rounded     -0.01860 -0.04165 -0.002423
+ 2.01-15% error refClass Percentage - kWh Rounded      0.02233 -0.00958  0.079639
+ 2.01-15% error refClass USD - kWh        Rounded      0.03701  0.00192  0.094659
+ Over 15% error refClass Percentage - kWh Rounded      0.06070 -0.03794  0.153891
+ Over 15% error refClass USD - kWh        Rounded      0.14428  0.04291  0.244697
+
+```
+
+# pre exc
+----
+
+"Reference class format significantly affected response accuracy.
+Compared to kWh format, USD format was associated with reduced probability of exact matches (AME = -13.7%, 95% CI: [-26.0%, -2.3%]) and increased probability of large errors exceeding 15% (AME = 12.6%, 95% CI: [2.3%, 22.1%]).
+Similar but non-significant trends were observed for percentage format."
+
+------
+
+"On average, across all conditions, the use of the USD format was associated with a statistically significant decrease in the probability of achieving an 'Exact match' in planning accuracy compared to the kWh format (AME = -0.137, 95% CI: [-0.260, -0.023]).
+The use of the Percentage format also decreased the likelihood of an exact match when compared to kWh (AME = -0.082, 95% CI: [-0.205, 0.030].
+However, this was not statistically significant at an alpha of .05."
+
+Reporting (Conditional AMEs by rounded):
+"The effect of refClass varied by whether plan values were rounded.
+When plan values were rounded, the use of the USD format was associated with a larger decrease in the probability of an 'Exact match' compared to kWh (AME = -0.164, 95% CI: [-0.297, -0.039]) than when plan values were not rounded (AME = -0.111, 95% CI: [-0.234, 0.002]).
+A similar, but not statistically significant, pattern was observed for the Percentage format."
+
+Reporting (Conditional AMEs by rounded and pct_goal):
+"Further analysis revealed that the effect of refClass was also moderated by the target savings goal (pct_goal).
+For example, when plans were not rounded and the savings goal was 10%, the use of the USD format had a negligible effect on the probability of an 'Exact match' compared to kWh (AME = -0.001, 95% CI: [-0.018, 0.002]).
+However, when plans were rounded and the goal was 15%, the USD format was associated with a larger, though still not statistically significant, decrease in the probability of an exact match (AME = -0.005, 95% CI: [-0.078, 0.006]).
+These findings highlight the complex interplay between information format, rounding, and savings goals in influencing planning accuracy."
+
+-------
+
+"The effect of the reference class on accuracy level was not uniform; rather, it varied depending on whether participants received rounded values and the degree of the goal (i.e., 10% vs.
+15%).
+For instance, when controlling for rounding and goal percentage, switching from kWh to USD was associated with an increased probability of being in higher error categories, whereas switching from kWh to a Percentage reference class showed a smaller and less consistent effect.
+However, these trends were contingent on the interaction terms, with some intervals overlapping zero, indicating uncertainty in the magnitude of the effects.
+Consequently, we conducted follow-up analyses of predicted probabilities and conditional slopes across experimental conditions (e.g., illustrating how the reference class altered the likelihood of exact or minimal error matching).
+These follow-up analyses, presented in Table X and Figure Y, clarify the direction and statistical strength of the observed interactions, providing a more nuanced understanding of how the reference class influences planning accuracy under varying informational contexts."
+
+------
+
+Key Observations
+
+Directional Effects: Notice that for the "Exact Match" category, the Percentage and USD formats have negative average effects, suggesting they may decrease the probability of participants exactly matching the goal compared to the kWh format.
+For the "Over 15% error" category the signs are reversed, where Percentage and USD may increase the probability of very inaccurate plans.
+
+Magnitude: The magnitude of these effects is larger for USD compared to Percentage, suggesting that this format may have a larger effect on accuracy.
+
+Category Differences: The average marginal effect of the refClass changes depending on the accuracy_level category you examine, indicating a complex relationship.
+
+Start with the AMEs (but with a warning):
+
+Begin with ame to give your reader an overall sense of the general direction of the effect.
+
+Report the table using an output that is aligned with the journal.
+
+For example, "The average marginal effect of presenting the goal as a percentage versus presenting as a kWh was -0.08 for "Exact Match", suggesting that, across all conditions, presenting as a percentage may decrease the probability of a participant exactly matching their goal relative to when the goal was presented in kWh.
+Presenting the goal as USD had a similar, but larger effect, with an average probability difference of -0.14 for "Exact Match".
+However, the direction of these effects was reversed for "Over 15% Error", suggesting that USD and Percentage formats may increase the likelihood of inaccurate plans".
+
+Crucially, acknowledge the limitations: Explicitly state that these are average effects and that the interactions mean the effect of refClass isn't constant.
+Explain that you will follow up with a more detailed conditional analysis to better describe the relationships.
+
+Move to Conditional Marginal Effects (with ame_by_rounded):
+
+Use ame_by_rounded to show how the effect of refClass varies depending on the rounding and the accuracy level, using an approach similar to the AMEs.
+
+For example: "These average marginal effects of the different refClass levels, by rounding are shown in Table X.
+Presenting the goals as percentages or USDs, relative to presenting the goal as kWh, tends to reduce the likelihood of an exact match, with both conditions reducing probability of matching the goal in non-rounded settings by about -0.08 and -0.11, respectively, and by -0.08 and -0.16 in rounded settings.
+However, when participants made errors of more than 15% the likelihood of being in this condition increased for both Percentage and USD conditions.
+This effect also appears slightly higher in the rounded conditions."
+
+You can present this using a table (if space permits) or by highlighting a few key patterns, particularly focusing on whether rounding impacts the effects of reference class.
+
+Address the Three-Way Interaction (using ame_by_rounded_pct_goal):
+
+Now, show how the refClass differs as a function of both rounded and pct_goal variables, by describing key differences and patterns.
+
+Again, report the results in a table and/or in the text.
+
+Present Conditional Slopes (and perhaps some strategic plotting):
+
+If tables begin to become unwieldy due to a large number of conditions, then consider using plots.
+For example, using the plots produced using code like plot_slopes(ordinal_model_s2_logit_c1add_int3, variables = "refClass", condition = c("rounded","pct_goal"), transform = "response") can be an excellent way to visualize complex patterns that might otherwise be difficult to grasp from a table.
+Use plots judiciously to emphasize key takeaways, and refer to specific patterns in your discussion.
+
+Reporting in Text
+
+Here's an example of how you could integrate your findings into a manuscript:
+
+"First, we examined the average marginal effects of presenting energy information using different formats (i.e.
+refClass), averaged across all other conditions.
+The results, shown in Table X, suggest that presenting a goal using percentages or USDs tend to decrease the likelihood of an exactly matching a goal compared to presenting the goal in kWh.
+However, these effects were reversed in cases where errors were above 15%, where presenting a goal as a percentage or USD increased the average likelihood of a severe error.
+Given the presence of a three way interaction, we next examined how the effect of refClass varies across different conditions of rounded and pct_goal.
+The average marginal effects of refClass, shown in Table X, demonstrated that rounding did not dramatically impact the observed patterns of likelihood of exact matches versus severe errors.
+Presenting goals as USD tended to have a more negative (or positive in the case of severe errors) effect.
+We next turned to a visualization of conditional marginal effects (Figure X).
+This revealed more nuanced patterns of interaction, where [Describe key takeaways from the plot – e.g., how effects change as a function of rounded and pct_goal].
+Therefore, while, on average, percentage and USD may reduce the accuracy of plans compared to presenting in kWh, the effect is complex, and depends on context."
+
+---------
+
+	•	USD format significantly decreases the odds of an Exact match when rounded = Rounded (Estimate = -0.164, 95% CI: -0.297, -0.039), but the effect is smaller and less clear when rounded = Not Rounded (Estimate = -0.111, CI includes 0).
+	•	USD format also significantly increases the odds of Over 15% errors when rounded = Rounded (Estimate = 0.138, 95% CI: 0.035, 0.233).
+
+  	“The effect of refClass (Percentage and USD) on accuracy level outcomes depends on rounding condition (rounded) and goal percentage (pct_goal).
+In general, the USD format significantly decreases the likelihood of achieving an Exact match and increases the likelihood of large errors (Over 15%) when rounding is applied.
+Specifically, when rounding is present and the goal percentage is high (15%), the USD format increases the odds of Over 15% errors by approximately 42% (95% CI: 5.5%, 75.5%).
+The effects of the Percentage format are smaller and often not statistically significant across conditions.”
+
+Table Summary
+
+Condition	Contrast	Accuracy Level	Estimate	95% CI	Significance
+Rounded, 15% Goal	USD vs KWh	Over 15% error	0.420	(0.055, 0.755)	Significant increase in large errors
+Rounded	USD vs KWh	Exact match	-0.164	(-0.297, -0.039)	Significant decrease in exact match
+Not Rounded	USD vs KWh	Exact match	-0.111	(-0.234, 0.002)	Non-significant decrease
+Rounded	Percentage vs KWh	All levels	Small (varies)	CIs include zero	Minimal and non-significant effects

### 2024-12-18 misc/ordinal.qmd
update - marginal effects

@@
   #echo: false
+  eval: false
   warning: false
@@
 pacman::p_load(dplyr,purrr,tidyr,here,tibble,stringr,brms,rstan,bayestestR,emmeans,tidybayes,modelsummary,
-               ggplot2,gt,knitr,kableExtra,ggh4x,lme4,flextable,pander)
+               ggplot2,gt,knitr,kableExtra,ggh4x,lme4,flextable,pander,marginaleffects,ggstance)
 options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
@@
 
-s1 <- readRDS(here::here("data/s1_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds"))))
-s2_long <- readRDS(here::here("data/s2_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds"))))
+# s1 <- readRDS(here::here("data/s1_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds"))))
+# s2_long <- readRDS(here::here("data/s2_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds"))))
+
+s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
+  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
+  filter(!(id %in% readRDS(here::here("data/s1_grp_outlier_ids.rds")))) |>
+  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))
+
+s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
+  filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
+  filter(!(id %in% readRDS(here::here("data/s2_grp_outlier_ids.rds"))) ) |>
+  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))
+
 ```
@@
 ```{r}
-
 s1_agg <- s1 |> 
     filter(appliance !="Total kWh") |> 
-    group_by(id,refClass,state,block,plan,calc,edu,pct_goal) |> 
+    group_by(id,refClass,state,block,plan,edu,pct_goal,calc) |> 
     summarise(total_kWh = sum(value),orig_kWh=sum(family), 
-                pct_change = round((orig_kWh-total_kWh)/orig_kWh,3), 
+                pct_change = abs(round((orig_kWh-total_kWh)/orig_kWh,3)), 
                 n_change = sum(value!=family),
@@
                     abs_error == 0.00 ~ "Exact match",
-                    abs_error <= 0.02 ~ "0.01-2% error",
-                    abs_error <= 0.15 ~ "2.01-15% error",
-                    TRUE ~ "Over 15% error"  # Capture all remaining cases
+                    abs_error <= 0.05 ~ "0.01-5% error",
+                    TRUE ~ "Over 5% error"  # Capture all remaining cases
                 ), 
-                levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+                levels = c("Exact match", "0.01-5% error", "Over 5% error"),
                 ordered = TRUE
             )
-        )
+        ) |> relocate(accuracy_level, .after= "pct_change")
+
+s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
+    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+    summarise(
+    mg=sum(matched_goal),
+    mgc=sum(close_match),
+    n=n(), 
+    pct=mg/n,
+    pct_close=mgc/n,
+    mean_pct_change=mean(pct_change),
+    mean_abs_error=mean(abs_error),
+    mean_log_abs_error=mean(log_abs_error),
+    n_accuracy=first(n_accuracy)) |> 
+    mutate(accuracy_level = factor(
+            case_when(
+                mean_abs_error < 0.02 ~ "Exact match",
+                mean_abs_error <= 0.05 ~ "0.01-5% error",
+                TRUE ~ "Over 5% error"  # Capture all remaining cases
+            ), 
+            levels = c("Exact match", ".01-5% error", "Over 5% error"),
+            ordered = TRUE
+        ))
 
@@
 
-s2_agg1 <- s2_long |> 
+s2_agg <- s2_long |> 
   filter(appliance != "TOTAL") |> 
-  group_by(id, state,refClass,pct,pct_goal,plan,rounded) |> 
+  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
   summarise(
@@
   ) |>
-  mutate(matched_goal = (pct_change == pct),
+  mutate(
+    matched_goal = (pct_change == pct),
+    close_match = abs(pct_change - pct) <= 0.02,
                 error = pct_change - pct,
-                abs_error = abs(error)) |> 
-      ungroup() |> # Add ungroup here
+                abs_error = abs(error),
+                log_abs_error=log(abs(error)+.007)) |> 
+      ungroup() |> 
         mutate(
@@
 
-ggplot(data = s2_agg1, aes(x = accuracy_level, fill = refClass)) +
+s2_agg4 <- s2_agg |> group_by(id,refClass,calc) |> 
+    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+    summarise(
+    mg=sum(matched_goal),
+    mgc=sum(close_match),
+    n=n(), 
+    pct=mg/n,
+    pct_close=mgc/n,
+    mean_pct_change=mean(pct_change),
+    mean_abs_error=mean(abs_error),
+    mean_log_abs_error=mean(log_abs_error),
+    n_accuracy=first(n_accuracy)) |> 
+    mutate(accuracy_level = factor(
+            case_when(
+                mean_abs_error < 0.02 ~ "Exact match",
+                mean_abs_error <= 0.05 ~ ".01-5% error",
+                TRUE ~ "Over 5% error"  # Capture all remaining cases
+            ), 
+            levels = c("Exact match", "01-5% error", "Over 5% error"),
+            ordered = TRUE
+        ))
+
+ggplot(data = s2_agg, aes(x = accuracy_level, fill = refClass)) +
   geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
@@
 
-observed_props2 <- s2_agg1 |>
+observed_props2 <- s2_agg |>
   group_by(refClass, accuracy_level) |>
@@
   # Calculate proportions
-prop_acc_s2 <- s2_agg1 %>%
+prop_acc_s2 <- s2_agg %>%
     group_by(refClass, accuracy_level) %>%
@@
 
-ggplot(data = s2_agg1, aes(x = accuracy_level, fill = refClass)) +
+ggplot(data = s2_agg, aes(x = accuracy_level, fill = refClass)) +
   geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
@@
 
-prop4_acc_s2 <- s2_agg1 %>%
+prop4_acc_s2 <- s2_agg %>%
     group_by(refClass, accuracy_level,rounded,pct_goal) %>%
@@
 
-s2_agg1 %>%
+s2_agg %>%
     group_by( accuracy_level,rounded) %>%
@@
 
-s2_agg1 <- s2_agg1 %>%
+s2_agg <- s2_agg %>%
   mutate(
@@
   accuracy_level ~ refClass +rounded+pct_goal + (1|id)+ (1|state),
-  data = s2_agg1,
+  data = s2_agg,
   family = cumulative("probit"),
@@
 pred_summary_s2 <- ordinal_model_s2_probit %>%
-  epred_draws(newdata = s2_agg1, re_formula = NA) %>%
+  epred_draws(newdata = s2_agg, re_formula = NA) %>%
   group_by(refClass, rounded, pct_goal, .category) %>%
@@
 
+marginal_effects(ordinal_model_s2_probit)
+
+```
+
+## S2 logit version
+```{r}
+
+ordinal_model_s2_logit <- brm(
+  accuracy_level ~ refClass +rounded+pct_goal + (1|id)+ (1|state),
+  data = s2_agg,
+  family = cumulative("logit"),
+  cores = 4,
+  iter = 2000,
+  control = list(adapt_delta = 0.99), # Recommended for ordinal models
+  prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
+            prior(normal(0, 2), class = "b")), # Priors for predictors
+  file = paste0(here::here("data/model_cache",'s2_logit_add.rds')) # Cache for efficiency
+)
+summary(ordinal_model_s2_logit)
+# Regression Coefficients:
+#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# Intercept[1]          -2.13      0.64    -3.39    -0.86 1.00      639     1328
+# Intercept[2]          -0.62      0.63    -1.89     0.63 1.00      658     1366
+# Intercept[3]           3.15      0.65     1.88     4.42 1.00      717     1638
+# refClassPercentage     0.83      0.75    -0.64     2.33 1.00      452     1253
+# refClassUSD            1.87      0.79     0.31     3.35 1.00      508     1260
+# roundedRounded        -0.66      0.18    -1.01    -0.31 1.00     4276     3095
+# pct_goal15%           -0.44      0.18    -0.79    -0.10 1.00     6263     3109
+
+pred_summary_s2 <- ordinal_model_s2_logit %>%
+  epred_draws(newdata = s2_agg, re_formula = NA) %>%
+  group_by(refClass, rounded, pct_goal, .category) %>%
+  summarise(
+    mean_prob = mean(.epred),
+    lower_ci = quantile(.epred, 0.025),
+    upper_ci = quantile(.epred, 0.975),
+    .groups = "drop"
+  )
+
+pred_summary_s2 |> pander::pandoc.table(caption="Study 1: Predicted probabilities of accuracy level by reference class")
+
+```
+
+### logit vs. probit
+
+```{r}
+
+loo_logit <- loo(ordinal_model_s2_logit)
+loo_probit <- loo(ordinal_model_s2_probit)
+loo_compare(loo_logit, loo_probit)
+
+waic(ordinal_model_s2_logit)
+waic(ordinal_model_s2_probit)
+
+# compare
+rstanarm::loo_compare(loo_logit, loo_probit)
+
+pp_check(ordinal_model_s2_logit, type = "bars_grouped", group="refClass", fatten = 2) +
+  scale_x_continuous("Response Category", breaks = 1:4, 
+            labels = c("Exact", "0.01-2%", "2.01-15%", ">15%")) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
+  ggtitle("Posterior Predictive Check by Reference Class - logit") +
+  theme_minimal() +
+  theme(
+    legend.background = element_blank(),
+    legend.position = "bottom",
+    panel.grid.minor = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 1))
+
+pp_check(ordinal_model_s2_probit, type = "bars_grouped", group="refClass", fatten = 2) +
+  scale_x_continuous("Response Category", breaks = 1:4, 
+            labels = c("Exact", "0.01-2%", "2.01-15%", ">15%")) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
+  ggtitle("Posterior Predictive Check by Reference Class - probit") +
+  theme_minimal() +
+  theme(
+    legend.background = element_blank(),
+    legend.position = "bottom",
+    panel.grid.minor = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 1))
+
 ```
@@
   accuracy_level ~ refClass * rounded * pct_goal + (1|id) + (1|state),
-  data = s2_agg1,
+  data = s2_agg,
   family = cumulative("logit"),
@@
 pred_summary_s2 <- ordinal_model_s2 %>%
-  epred_draws(newdata = s2_agg1, re_formula = NA) %>%
+  epred_draws(newdata = s2_agg, re_formula = NA) %>%
   group_by(refClass, rounded, pct_goal, .category) %>%
@@
 ```
+
+# 3 level
+
+```{r}
+s1_aggb <- s1 |> 
+    filter(appliance !="Total kWh") |> 
+    group_by(id,refClass,state,block,plan,edu,pct_goal,calc) |> 
+    summarise(total_kWh = sum(value),orig_kWh=sum(family), 
+                pct_change = abs(round((orig_kWh-total_kWh)/orig_kWh,3)), 
+                n_change = sum(value!=family),
+                state_p_dif=mean(state_p_dif),
+                state_f_dif=mean(state_f_dif),
+                n_less_avg = sum(less_avg),
+                duration=first(Duration__in_seconds_)) |> 
+    mutate(matched_goal = (pct_change == pct_goal), 
+                error = pct_change - pct_goal,
+                abs_error = abs(error),
+                log_abs_error=log(abs(error)+.007), 
+                close_match = abs_error <= 0.02) |>
+    ungroup() |> # Add ungroup here
+        mutate(
+            accuracy_level = factor(
+                case_when(
+                    abs_error == 0.00 ~ "Exact match",
+                    abs_error <= 0.05 ~ "0.01-5% error",
+                    TRUE ~ "Over 5% error"  # Capture all remaining cases
+                ), 
+                levels = c("Exact match", "0.01-5% error", "Over 5% error"),
+                ordered = TRUE
+            )
+        ) |> relocate(accuracy_level, .after= "pct_change")
+
+ggplot(data = s1_aggb, aes(x = refClass, fill = accuracy_level)) +
+  geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
+  scale_y_continuous(labels = scales::percent) +
+  labs(title = "Accuracy Levels by Reference Class",
+       x = "Accuracy Level",
+       y = "Percentage of Participants",
+       fill = "Reference Class") +
+  theme_minimal()
+
+library(lmerTest)
+mixed_model_abs_error <- lmer(log_abs_error ~ refClass+calc+ (1+plan|id) + (1|state), data = s1_aggb)
+summary(mixed_model_abs_error)
+
+logistic_model_clean_mixed <- glmer(matched_goal ~ refClass +calc+ (1|id) + (1|state), data = s1_aggb, family = binomial)
+summary(logistic_model_clean_mixed)
+emmeans::emmeans(logistic_model_clean_mixed, pairwise ~ refClass)
+
+mixed_model_abs_error <- lmer(log_abs_error ~  refClass+calc +(1|state)+ (1|id), data = s1_aggb)
+summary(mixed_model_abs_error)
+emmeans::emmeans(mixed_model_abs_error, pairwise ~ refClass)
+
+# categorial mixed effects regression on accuracy_level (ordered factor) with clean data (frequentist)
+ordinal_model_clean_mixed <- clmm(accuracy_level ~ refClass +calc + (1|id) + (1|state), data = s1_aggb)
+summary(ordinal_model_clean_mixed)
+emmeans::emmeans(ordinal_model_clean_mixed, pairwise ~ refClass)
+
+ordinal_model_s1 <- brm(
+    accuracy_level ~ refClass +calc + (1|id) + (1|state),
+    data = s1_aggb,
+    family = cumulative("logit"),
+    cores = 4,
+    iter = 4000,
+    control = list(adapt_delta = 0.98), 
+    prior = c(prior(normal(0, 4), class = "Intercept"), 
+                prior(normal(0, 4), class = "b")), 
+    file = paste0(here::here("data/model_cache",'s1_acc3_add.rds')) 
+)
+# summary(ordinal_model_s1)
+#  Family: cumulative 
+#   Links: mu = logit; disc = identity 
+# Formula: accuracy_level ~ refClass + calc + (1 | id) + (1 | state) 
+#    Data: s1_aggb (Number of observations: 940) 
+#   Draws: 4 chains, each with iter = 4000; warmup = 2000; thin = 1;
+#          total post-warmup draws = 8000
+# 
+# Multilevel Hyperparameters:
+# ~id (Number of levels: 235) 
+#               Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# sd(Intercept)     3.90      0.35     3.27     4.64 1.00     2072     4098
+# 
+# ~state (Number of levels: 4) 
+#               Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# sd(Intercept)     0.58      0.45     0.12     1.78 1.00     2961     2957
+# 
+# Regression Coefficients:
+#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# Intercept[1]          -4.21      0.85    -5.90    -2.58 1.00     1567     2898
+# Intercept[2]          -0.89      0.82    -2.49     0.71 1.00     1528     2564
+# refClassPercentage     1.44      0.71     0.07     2.88 1.00     1348     2624
+# refClassUSD            3.13      0.69     1.81     4.50 1.00     1373     2564
+# calcUsedCalculator    -3.30      0.73    -4.80    -1.92 1.00     1491     2765
+
+pp_check(ordinal_model_s1, type = "bars_grouped", group="refClass", fatten = 2)
+pp_check(ordinal_model_s1, type = "dens_overlay_grouped", group="refClass", fatten = 2,ndraws=100)
+
+ame1 <- avg_slopes(
+  ordinal_model_s1, 
+  variables = "refClass"
+)
+print(ame1)
+ #         Group         Contrast Estimate   2.5 %   97.5 %
+ # Exact match   Percentage - kWh  -0.1064 -0.2164 -0.00504
+ # Exact match   USD - kWh         -0.2001 -0.2988 -0.10898
+ # 0.01-5% error Percentage - kWh  -0.0211 -0.0577  0.00242
+ # 0.01-5% error USD - kWh         -0.0811 -0.1143 -0.04856
+ # Over 5% error Percentage - kWh   0.1314  0.0063  0.24907
+ # Over 5% error USD - kWh          0.2842  0.1723  0.38086
+
+as.data.frame(fixef(ordinal_model_s1)[,-2])|> as.data.frame() %>%
+    rownames_to_column(var = "Parameter") %>%
+    mutate(across(where(is.numeric), exp)) |>
+    filter(!stringr::str_detect(Parameter, "Intercept")) |> 
+    filter(!stringr::str_detect(Parameter, "calc")) |> 
+    # rename columns to |comparison           | odds_ratio| ci_lower| ci_upper|
+    rename(Comparison = Parameter, odds_ratio = Estimate, ci_lower = Q2.5, ci_upper = Q97.5) |>
+    # rename levels to Percentage vs kWh; USD vs kWh; 
+    mutate(Comparison = case_when(
+        str_detect(Comparison, "refClassPercentage") ~ "Percentage vs kWh",
+        str_detect(Comparison, "refClassUSD") ~ "USD vs kWh",
+        TRUE ~ Comparison
+    )) |> kable(escape=FALSE,booktabs=TRUE,align=c("l"))
+
+# |Comparison        |odds_ratio |ci_lower |ci_upper |
+# |:-----------------|:----------|:--------|:--------|
+# |Percentage vs kWh |4.2        |1.1      |18       |
+# |USD vs kWh        |22.9       |6.1      |90       |
+```
+
+```{r}
+
+condEffects <- function(m,xvar){
+  m |> ggplot(aes(x = {{xvar}}, y = .value, color = refClass, fill = refClass)) + 
+  stat_dist_pointinterval() + 
+  stat_halfeye(alpha=.1, height=.5) +
+  theme(legend.title=element_blank(),axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5))
+
+}
+```
+
+```{r}
+
+s2_agg <- s2_long |> 
+  filter(appliance != "TOTAL") |> 
+  group_by(id,refClass,calc, state,pct,pct_goal,plan,rounded) |> 
+  summarise(
+    total_kWh = sum(value),
+    orig_kWh = sum(family),
+    pct_change = round((orig_kWh - total_kWh) / orig_kWh, 3),
+    state_dif = mean(state_dif),
+    .groups = "drop"
+  ) |>
+  mutate(
+    matched_goal = (pct_change == pct),
+    close_match = abs(pct_change - pct) <= 0.02,
+                error = pct_change - pct,
+                abs_error = abs(error),
+                log_abs_error=log(abs(error)+.007)) |> 
+      ungroup() |> 
+        mutate(
+            accuracy_level = factor(
+                case_when(
+                    abs_error == 0.00 ~ "Exact match",
+                    abs_error <= 0.05 ~ "0.01-5% error",
+                    TRUE ~ "Over 5% error"  # Capture all remaining cases
+                ), 
+                levels = c("Exact match","0.01-5% error", "Over 5% error"),
+                ordered = TRUE
+            )
+        )
+
+ggplot(data = s2_agg, aes(x = refClass, fill = accuracy_level)) +
+  geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
+  scale_y_continuous(labels = scales::percent) +
+  labs(title = "Accuracy Levels by Reference Class",
+       x = "Accuracy Level",
+       y = "Percentage of Participants",
+       fill = "Reference Class") +
+  theme_minimal()
+
+library(lmerTest)
+mixed_model_abs_error <- lmer(log_abs_error ~ refClass + calc+ (1|id) + (1|state), data = s2_agg)
+summary(mixed_model_abs_error)
+
+logistic_model_clean_mixed <- glmer(matched_goal ~ refClass+calc+pct_goal+rounded +(1|state)+ (1|id) , data = s2_agg, family = binomial)
+summary(logistic_model_clean_mixed)
+emmeans::emmeans(logistic_model_clean_mixed, pairwise ~ refClass)
+
+mixed_model_abs_error <- lmer(log_abs_error ~  refClass+calc+pct_goal+rounded +(1|state)+ (1|id), data = s2_agg)
+summary(mixed_model_abs_error)
+emmeans::emmeans(mixed_model_abs_error, pairwise ~ refClass)
+
+library(ordinal)    
+library(parameters)
+
+# categorial mixed effects regression on accuracy_level (ordered factor) with clean data (frequentist)
+ordinal_model_clean_mixed <- clmm(accuracy_level ~ refClass + calc+pct_goal+rounded + (1|id) + (1|state), data = s2_agg)
+summary(ordinal_model_clean_mixed)
+emmeans::emmeans(ordinal_model_clean_mixed, pairwise ~ refClass)
+
+ordinal_model_clean_mixed <- clmm(accuracy_level ~ (refClass*pct_goal*rounded)+ calc + (1|id) + (1|state), data = s2_agg)
+summary(ordinal_model_clean_mixed)
+emmeans::emmeans(ordinal_model_clean_mixed, pairwise ~ refClass)
+pairs(emmeans(ordinal_model_clean_mixed, ~ refClass | rounded +pct_goal+calc))
+
+ordinal_model_clean_mixed <- clmm(accuracy_level ~ refClass + rounded + pct_goal + calc + refClass:calc+refClass:rounded + refClass:pct_goal + rounded:pct_goal + (1|id) + (1|state), data = s2_agg)
+summary(ordinal_model_clean_mixed)
+emmeans::emmeans(ordinal_model_clean_mixed, pairwise ~ refClass+calc)
+
+emms_ref <- emmeans(ordinal_model_clean_mixed, ~ refClass, type = "link")
+contrast_ref <- contrast(emms_ref, "trt.vs.ctrl", ref = "kWh") 
+odds_ratios <- summary(contrast_ref, type = "response") # back-transform to odds
+odds_ratios
+
+pairs(emmeans(ordinal_model_clean_mixed, ~ refClass | rounded +pct_goal+calc))
+pairs(emmeans(ordinal_model_clean_mixed, ~ refClass+calc))
+
+# ordinal regression
+polr_model <- MASS::polr(accuracy_level ~ refClass + calc, data = s2_agg)
+summary(polr_model)
+emmeans::emmeans(polr_model, pairwise ~ refClass+calc)
+
+m_equal <- clm(accuracy_level ~ refClass+calc+pct_goal+rounded, 
+    data = s2_agg, 
+    link="probit")
+
+parameters::model_parameters(m_equal) |> 
+  insight::print_html()
+summary(m_equal)
+
+ame_by_rounded_pct_goal <- avg_slopes(
+  m_equal,
+  variables = "refClass",
+  by = c("rounded", "pct_goal", "accuracy_level")
+)
+
+ordinal_model_s2_logit <- brm(
+  accuracy_level ~ refClass + calc+pct_goal+rounded + (1|id)+ (1|state),
+  data = s2_agg,
+  family = cumulative("logit"),
+  cores = 4,
+  iter = 3000,
+  control = list(adapt_delta = 0.99), # Recommended for ordinal models
+  prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
+            prior(normal(0, 2), class = "b")), # Priors for predictors
+  file = paste0(here::here("data/model_cache",'s2_acc3_add.rds')) # Cache for efficiency
+)
+# summary(ordinal_model_s2_logit)
+#  Family: cumulative 
+#   Links: mu = logit; disc = identity 
+# Formula: accuracy_level ~ refClass + calc + pct_goal + rounded + (1 | id) + (1 | state) 
+#    Data: s2_agg1b (Number of observations: 784) 
+#   Draws: 4 chains, each with iter = 3000; warmup = 1500; thin = 1;
+#          total post-warmup draws = 6000
+# 
+# Multilevel Hyperparameters:
+# ~id (Number of levels: 196) 
+#               Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# sd(Intercept)     4.88      0.51     3.98     5.99 1.00     1473     2687
+# 
+# ~state (Number of levels: 4) 
+#               Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# sd(Intercept)     0.38      0.39     0.01     1.42 1.00     1834     2700
+# 
+# Regression Coefficients:
+#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# Intercept[1]          -1.45      0.70    -2.85    -0.07 1.00     1200     1896
+# Intercept[2]           1.26      0.68    -0.09     2.65 1.00     1171     2064
+# refClassPercentage     1.02      0.84    -0.63     2.71 1.01      807     1901
+# refClassUSD            2.27      0.89     0.53     3.98 1.00      995     1835
+# calcNoCalculator       4.10      0.98     2.20     6.06 1.00     1876     3213
+# pct_goal15%           -0.39      0.22    -0.81     0.04 1.00     7847     4569
+# roundedRounded        -0.53      0.22    -0.96    -0.11 1.00     8353     4931
+
+conditional_effects(ordinal_model_s2_logit)
+
+pp_check(ordinal_model_s2_logit, type = "bars_grouped", group="refClass", fatten = 2) +
+  scale_x_continuous("Response Category", breaks = 1:3, 
+            labels = c("Exact", "0.01-5%", ">5%")) +
+  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
+  ggtitle("Posterior Predictive Check by Reference Class") +
+  theme_minimal() +
+  theme(
+    legend.background = element_blank(),
+    legend.position = "bottom",
+    panel.grid.minor = element_blank(),
+    axis.text.x = element_text(angle = 45, hjust = 1))
+
+ame2 <- avg_slopes(
+  ordinal_model_s2_logit, 
+  variables = "refClass",
+  ndraws=200
+)
+print(ame2)
+
+avg_slopes(
+  ordinal_model_s2_logit, 
+  variables = "refClass", type="response"
+)
+
+ame_by_rounded <- avg_slopes(
+  ordinal_model_s2_logit,
+  variables = "refClass",
+  by = c("rounded", "accuracy_level")
+)
+print(ame_by_rounded)
+
+ame_by_rounded_pct_goal <- avg_slopes(
+  ordinal_model_s2_logit,
+  variables = "refClass",
+  by = c("rounded", "pct_goal", "accuracy_level")
+)
+print(ame_by_rounded_pct_goal)
+
+ordinal_model_s2_logit |> emmeans(~refClass+pct_goal, at=list(pct_goal=c("10%","15%"),  re_formula=NA)) |> 
+  gather_emmeans_draws() |> 
+  condEffects(pct_goal)
+
+ordinal_model_s2_logit |> emmeans(~refClass+calc, at=list(calc=c('Calculator', 'No Calculator'),  re_formula=NA)) |> 
+  gather_emmeans_draws() |> 
+  condEffects(calc)
+
+emmeans::emmeans(ordinal_model_s2_logit, pairwise ~ refClass)
+# $contrasts
+#  contrast         estimate lower.HPD upper.HPD
+#  kWh - Percentage    -1.01      -2.6      0.68
+#  kWh - USD           -2.29      -4.0     -0.52
+#  Percentage - USD    -1.23      -3.1      0.60
+# 
+# Results are averaged over the levels of: calc, pct_goal, rounded 
+# Point estimate displayed: median 
+# Results are given on the log odds ratio (not the response) scale. 
+# HPD interval probability: 0.95
+
+emmeans::emmeans(ordinal_model_s2_logit, pairwise ~ refClass,type="response")
+
+emmeans::emmeans(ordinal_model_s2_logit,~ refClass,type="response") |> contrast(method="revpairwise")
+
+emmeans::emmeans(ordinal_model_s2_logit,~ refClass+calc+pct_goal+rounded) |> contrast(method="revpairwise")
+
+avg_predictions(ordinal_model_s2_logit, newdata = datagrid(refClass=unique),by="refClass")
+avg_predictions(ordinal_model_s2_logit, newdata = datagrid(id=unique,refClass=unique),by="refClass")
+avg_slopes(ordinal_model_s2_logit, variables="refClass",type="response",by="refClass")
+
+ ordinal_model_s2_logit |> emmeans(pairwise~ refClass, re_formula=NA)  |> 
+  pluck("contrasts") |>
+  gather_emmeans_draws() |> 
+  group_by(contrast,.draw) |> summarise(value=mean(.value), n=n()) |> 
+  # filter(!(contrast %in% c("ALM Constant - EXAM Constant","ALM Constant - EXAM Varied","ALM Varied - EXAM Varied ", "EXAM Constant - ALM Varied" ))) |> 
+  ggplot(aes(x=value,y=contrast,fill=contrast)) +stat_halfeye() + labs(x="Model Error Difference",y="Contrast") + theme(legend.position="none")
+
+emm_ref_all <- emmeans(ordinal_model_s2_logit, ~ refClass, type="prob", levels=levels(s2_agg1b$accuracy_level))
+emm_pairs <- pairs(emm_ref_all)
+emm_pairs
+
+avg_preds <- predictions(ordinal_model_s2_logit, 
+                                 newdata = datagrid(id=unique,refClass = unique),
+                                 by="refClass") |> as.data.frame() |> 
+  mutate(refClass = factor(refClass, levels = c("kWh", "Percentage", "USD")))
+
+predictions(ordinal_model_s2_logit, 
+                                 newdata = datagrid(id=unique,refClass = unique(s2_agg1b$refClass)),
+                                 by="refClass")
+
+predictions(ordinal_model_s2_logit, newdata = datagrid(id=unique,refClass=unique,calc=unique)) |> 
+  mutate(.value=estimate) |>
+  condEffects(refClass) + facet_wrap(~group, scales = "free")
+
+```

### 2024-12-18 nb/s2_survey.qmd
update - marginal effects

@@
 
-s2_agg4 <- s2_agg1 |> group_by(id,refClass,calc,els) |> 
+s2_agg4 <- s2_agg |> group_by(id,refClass,calc,els) |> 
     mutate(n_accuracy = n_distinct(accuracy_level)) |> 
@@
                 mean_abs_error < 0.02 ~ "Exact match",
-                mean_abs_error <= 0.02 ~ "0.01-2% error",
-                mean_abs_error <= 0.15 ~ "2.01-15% error",
-                TRUE ~ "Over 15% error"  # Capture all remaining cases
+                mean_abs_error <= 0.05 ~ ".01-5% error",
+                TRUE ~ "Over 5% error"  # Capture all remaining cases
             ), 
-            levels = c("Exact match", "0.01-2% error", "2.01-15% error", "Over 15% error"),
+            levels = c("Exact match", ".01-5% error", "Over 5% error"),
             ordered = TRUE
@@
 s2_agg4 |> group_by(accuracy_level) |> summarize(n=n())
-s2_agg1 |> group_by(accuracy_level) |> summarize(n=n())
+s2_agg |> group_by(accuracy_level) |> summarize(n=n())
 
@@
 cor.test(s2_agg4$els, s2_agg4$mean_log_abs_error)
-cor.test(s2_agg1$els, s2_agg1$log_abs_error)
+cor.test(s2_agg$els, s2_agg$log_abs_error)
 
@@
 m_equal <- clm(accuracy_level ~ els, 
-    data = s2_agg1, 
+    data = s2_agg, 
     link="probit")
@@
 library(lme4)
-mixed_model <- lmer(abs_error ~ els + (1 | id), data = s2_agg1)
+mixed_model <- lmer(abs_error ~ els + (1 | id), data = s2_agg)
 
@@
 
-polr_model_ref_interaction <- MASS::polr(accuracy_level ~ els * refClass + calc, data = s2_agg1)
+polr_model_ref_interaction <- MASS::polr(accuracy_level ~ els * refClass + calc, data = s2_agg)
 summary(polr_model_ref_interaction)
@@
 # Example for refClass = USD, calc = No Calculator
-new_data <- expand.grid(els = seq(min(s2_agg1$els), max(s2_agg1$els), length.out = 100),
+new_data <- expand.grid(els = seq(min(s2_agg$els), max(s2_agg$els), length.out = 100),
                         refClass = "USD",
@@
 predicted_probs_long <- tidyr::pivot_longer(predicted_probs_df, 
-                                            cols = `Exact match`:`Over 15% error`,
+                                            cols = `Exact match`:`Over 5% error`,
                                             names_to = "accuracy_level",
@@
 
-s2_agg <- s2_agg1 %>%
+s2_agg <- s2_agg %>%
   mutate(els_quartile = ntile(els, 4))
@@
     log_abs_error ~ els+refClass+calc+rounded+pct_goal + (1|id) + (1|state),
-    data = s2_agg1,
+    data = s2_agg,
     family = gaussian(),
@@
 
-check_model(s2_els_log_error_full_add)
+#check_model(s2_els_log_error_full_add)
 
@@
     accuracy_level ~ els+refClass+calc+rounded+pct_goal + (1|id) + (1|state),
-    data = s2_agg1,
+    data = s2_agg,
     cores = 4,
@@
     accuracy_level ~ (els*refClass)+(rounded*pct_goal)+calc + (1|id) + (1|state),
-    data = s2_agg1,
+    data = s2_agg,
     cores = 4,

### 2024-12-18 manuscript/manuscript.qmd
update intro a little

@@
 Prior research by @canfieldPerceptionsElectricityuseCommunications2017 demonstrated that presenting energy information in tabular formats enhances comprehension relative to graphs.
+Nevertheless, persistent difficulties remain in conveying energy data in formats that foster accurate consumer judgments and planning, as consumers often lack intuitive benchmarks for interpreting abstract units such as kilowatt-hours [@attariPublicPerceptionsEnergy2010].
 
@@
 Presenting energy information in absolute units (e.g., kWh) provides a clear reference class, potentially enhancing comprehension.
+Indeed, the manner in which energy information is presented has been shown to significantly influence both comprehension and subsequent behavior [@canfieldPerceptionsElectricityuseCommunications2017].
+For example, the use of tables has been found to facilitate point-reading, essential for understanding specific energy values, while graphs may be more useful in tasks requiring more complex information interpretation [@canfieldPerceptionsElectricityuseCommunications2017].
 
@@
 Specifically, there is a gap in understanding how presenting energy information in absolute units versus percentages or monetary terms affects the precision of planning appliance-specific reductions.
-Addressing this gap is crucial for developing effective interventions that promote energy conservation behaviors.
+The current study addresses these critical issues by systematically investigating the impact of varying information formats (kWh, percentage, and USD) on the accuracy of energy-planning decisions.
+By manipulating the presentation format of energy information, this research aims to elucidate how different representational formats influence planning accuracy.
 
@@
 This approach allows the estimation of threshold parameters and regression coefficients that characterize how changes in predictor variables (such as the reference class: kWh, percentage, or USD) relate to probabilities of being in each accuracy category.
+Specifically, we used a cumulative logit link function to model the ordered accuracy outcome, and we specified weakly informative priors for the regression coefficients (normal distributions with mean 0 and standard deviation of 1) and for the cutpoints (normal distributions with a mean of zero and a standard deviation of 4.0).
+The approach allows us to estimate threshold parameters and regression coefficients that characterize how changes in predictor variables (such as the reference class: kWh, percentage, or USD) relate to probabilities of being in each accuracy category.
 
@@
 
-Experiment 2 built upon the findings of Experiment 1 by incorporating additional manipulations of goal difficulty (10% vs.
-15% reduction) and numerical presentation (rounded vs.
-exact numbers), while maintaining the core manipulation of reference class (kWh, Percentage, USD).
-Nevertheless, the results largely converged with those of Experiment 1, providing further converging evidence that presenting energy reduction goals in absolute units (kWh) facilitates more accurate planning compared to percentage-based or monetary formats.
+Experiment 2 aimed to build upon the findings of Experiment 1, not only by replicating the core manipulation of reference class, but also by incorporating additional variables that might influence planning accuracy.
+These included goal difficulty and the way that the prior year's usage was presented (rounded or exact).
+The results largely converged with those of Experiment 1, providing further converging evidence that presenting energy reduction goals in absolute units (kWh) facilitates more accurate planning compared to percentage-based or monetary formats.
 
@@
 
-The effect of goal difficulty, with the more challenging 15% reduction goal associated with slightly better performance, although counterintuitive, was nonetheless relatively small and may warrant further investigation in future research.
+ The finding that the more challenging 15% reduction goal was associated with a slight improvement in accuracy is counterintuitive.
+It may be that participants put more effort into the task under this condition, or perhaps this is an artifact of the way that the task was presented.
+However, this effect was relatively small and thus should be explored in future research to better understand its underlying mechanisms.
+Furthermore, the magnitude of the effect size of this manipulation should be examined to better understand the practical implications of goal difficulty for energy conservation.
+
  The large and significant effect of calculator use, consistent across both experiments, underscores the crucial role of tools that individuals are likely to employ in real-world settings.
@@
 This advantage persisted across variations in goal difficulty and numerical presentation, suggesting a robust effect of reference class on planning accuracy.
+
+It is, however, important to note that while we find a significant benefit of presenting energy information in absolute units, some prior work suggests that consumers may prefer to receive information about costs, rather than absolute units (Karjalainen, 2011), or in some cases, that monetary information may lead to better decision-making (Blasch et al., 2019).
 

### 2025-01-04 docs/misc/ordinal.qmd
writing updates from last week I forgot to push

@@
 
-s1 <- readRDS(here::here("data/s1_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds"))))
-s2_long <- readRDS(here::here("data/s2_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds"))))
+# s1 <- readRDS(here::here("data/s1_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds"))))
+# s2_long <- readRDS(here::here("data/s2_processed.rds")) |> filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds"))))
+
+s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
+  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
+  filter(!(id %in% readRDS(here::here("data/s1_grp_outlier_ids.rds")))) |>
+  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))
+
+s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
+  filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
+  filter(!(id %in% readRDS(here::here("data/s2_grp_outlier_ids.rds"))) ) |>
+  mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))
+
+# s2 |> group_by(appliance,family,state,state_avg) |> select(appliance,family_prior_yr=family,state,state_avg) |> 
+#   slice_head() |> print(n=30) |> pander::pandoc.table(split.table=Inf)
+# 
+# s2 |> group_by(appliance,family,state,state_avg) |> select(appliance,family_prior_yr=family,state,state_avg) |>
+#   arrange(state,appliance) |>
+#   slice_head() |> print(n=30) |> kable()
+# 
+# s2 |> head() |> ungroup() |>select(-edu,-els,-block) |> kable()
 ```
@@
   accuracy_level ~ refClass +rounded+pct_goal + (1|id)+ (1|state),
-  data = s2_agg1,
+  data = s2_agg,
   family = cumulative("logit"),
@@
 pred_summary_s2 <- ordinal_model_s2_logit %>%
-  epred_draws(newdata = s2_agg1, re_formula = NA) %>%
+  epred_draws(newdata = s2_agg, re_formula = NA) %>%
   group_by(refClass, rounded, pct_goal, .category) %>%

### 2025-01-04 manuscript/appendix.qmd
writing updates from last week I forgot to push

@@
 
+```
+
+# A tibble: 4,560 × 17
+# Groups:   id, block [380]
+   id     block refClass appliance pct_goal rounded state    calc       edu                         els   pct family state_avg plan  value change state_dif
+   <chr>  <dbl> <fct>    <chr>     <chr>    <chr>   <chr>    <chr>      <fct>                     <dbl> <dbl>  <dbl>     <dbl> <fct> <dbl>  <dbl>     <dbl>
+ 1 s2_100     1 kWh      AC        10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1   1000       498 plan1   500   -500         2
+ 2 s2_100     1 kWh      AC        10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1   1000       498 plan2   300   -700      -198
+ 3 s2_100     1 kWh      HEAT      10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1  18000     16411 plan1 15000  -3000     -1411
+ 4 s2_100     1 kWh      HEAT      10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1  18000     16411 plan2 12000  -6000     -4411
+ 5 s2_100     1 kWh      WATER     10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1  12000      5832 plan1  6000  -6000       168
+ 6 s2_100     1 kWh      WATER     10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1  12000      5832 plan2  2000 -10000     -3832
+ 7 s2_100     1 kWh      REFRIG    10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1   1000      1142 plan1  2000   1000       858
+ 8 s2_100     1 kWh      REFRIG    10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1   1000      1142 plan2   500   -500      -642
+
+```{r}
+
+# widen by appliance, with separate row for value, family, and stage_avg
+
+s2 <- s2_long #|> select(-family,-state_avg,-edu,-change,-state_dif) |> pivot_wider(names_from = appliance, values_from = c(value)) 
+s2s <- s2_long |> select(-family,-state_avg,-edu,-change,-value) |> pivot_wider(names_from = appliance, values_from = c(state_dif)) |> select(id,block,refClass,pct,plan,AC:OTHER)
+# s2s
+#    id     block refClass     pct plan     AC   HEAT WATER REFRIG OTHER
+#    <chr>  <dbl> <fct>      <dbl> <fct> <dbl>  <dbl> <dbl>  <dbl> <dbl>
+#  1 s2_100     1 kWh         0.1  plan1     2  -1411   168    858   348
+#  2 s2_100     1 kWh         0.1  plan2  -198  -4411 -3832   -642 -6152
+#  3 s2_100     2 kWh         0.15 plan1  1324     19  -139    320  -424
+#  4 s2_100     2 kWh         0.15 plan2   324   -981 -1139   -318 -1424
+#  5 s2_101     1 kWh         0.1  plan1   102   -411  2168   -142   398
+#  6 s2_101     1 kWh         0.1  plan2   502  -1411   168   -142  1348
+#  7 s2_101     2 kWh         0.15 plan1   751    901  -396    682 -1883
+#  8 s2_101     2 kWh         0.15 plan2  2251    401   104    682 -1383
+#  9 s2_102     1 Percentage  0.1  plan1   958 -12954 -5354    200 -5417
+# 10 s2_102     1 Percentage  0.1  plan2   890 -15435 -4956   8544  3225
+
+# clustering planning strategies - kmeans
+
+    ‘
+s2s |>  ungroup() |> select(-id,-block,-refClass,-pct,-plan) |> scale(center = TRUE, scale = TRUE) |> kmeans(centers = 3) 
+s2s_clust <- s2s |>  ungroup() |> select(-id,-block,-refClass,-pct,-plan) |> scale(center = TRUE, scale = TRUE) |> kmeans(centers = 3) %>% .$cluster |> as.factor() 
+s2s <- cbind(s2s,c=s2s_clust)
+
+s2s |>    group_by(c) |>
+  summarise(ac=mean(AC),heat=mean(HEAT),water=mean(WATER),refrig=mean(REFRIG),other=mean(OTHER)) |>
+  pivot_longer(cols = c(ac:other), names_to = "Appliance", values_to = "Mean") |>
+  ggplot(aes(x=Appliance,y=Mean,col=c)) + 
+  geom_point() + 
+  facet_wrap(~c) +
+  theme_minimal()
+
+s2s <- s2_long |> select(-family,-state_avg,-edu,-change,-state_dif) |> pivot_wider(names_from = appliance, values_from = c(value))  |> select(id,block,refClass,pct,plan,AC:OTHER)
+
+s2s |>  ungroup() |> select(-id,-block,-refClass,-pct,-plan) |> scale(center = TRUE, scale = TRUE) |> kmeans(centers = 3)
+
+s2s_clust <- s2s |>  ungroup() |> select(-id,-block,-refClass,-pct,-plan) |> scale(center = TRUE, scale = TRUE) |> kmeans(centers = 3) %>% .$cluster |> as.factor()
+
+s2s <- cbind(s2s,c=s2s_clust) |> left_join(s2, join_by(id,refClass,block,plan,pct))
+s2s |>  group_by(c) |>
+  summarise(ac=mean(AC),heat=mean(HEAT),water=mean(WATER),refrig=mean(REFRIG),other=mean(OTHER)) |>
+  pivot_longer(cols = c(ac:other), names_to = "Appliance", values_to = "Mean") |>
+  ggplot(aes(x=Appliance,y=Mean,col=c)) + 
+  geom_point() + 
+  geom_point(aes(y=state_avg),color="black") +
+  facet_wrap(~c) +
+  theme_minimal()
+
+s2 |> group_by(appliance,family,state,state_avg) |> select(appliance,family_prior_yr=family,state,state_avg)
 ```

### 2025-01-04 manuscript/manuscript.qmd
writing updates from last week I forgot to push

@@
         wordcount-banner: true
+        echo: false
         theme: 
@@
         output-file: manuscript_hugo.md
-        suppress-bibliography: true
+        suppress-bibliography: false
     gfm:
@@
 
-- highlight the significant contribution of residential energy consumption to carbon emissions and the potential for substantial reductions.
-Make point about urgency of climate change?
-(IPCC (2014) and EIA (2012) ).
-
-- Electricity bills are a primary source of energy-use information for consumers and offer a promising avenue for enhancing communication about energy consumption [@fischerFeedbackHouseholdElectricity2008]
-
 ### Literature Review
 
-Energy poverty continues to be a pervasive issue in the United States [@memmottSociodemographicDisparitiesEnergy2021].
-This challenge partly arises from difficulties in converting information across numerical formats, impeding the development of precise energy reduction plans @reimerNumericCommunicationRisk2015.
-Prior research by @canfieldPerceptionsElectricityuseCommunications2017 demonstrated that presenting energy information in tabular formats enhances comprehension relative to graphs.
-Nevertheless, persistent difficulties remain in conveying energy data in formats that foster accurate consumer judgments and planning, as consumers often lack intuitive benchmarks for interpreting abstract units such as kilowatt-hours [@attariPublicPerceptionsEnergy2010].
+Electricity bills are a frequent touchpoint for consumers seeking to understand their energy usage, yet effectively turning that billing data into practical steps can be demanding.
+Among low-income households, these challenges contribute to disproportionate rates of energy insecurity [@memmottSociodemographicDisparitiesEnergy2021].
+At the same time, previous research shows that the clarity and format of energy-use information play a pivotal role in whether individuals adopt meaningful conservation behaviors [@fischerFeedbackHouseholdElectricity2008; @attariPublicPerceptionsEnergy2010; @canfieldPerceptionsElectricityuseCommunications2017].
+The interplay of these factors indicates that how energy data is presented can substantially shape not only how people interpret their consumption but also the degree to which they modify their habits to save energy.
 
@@
 In the context of energy conservation, using absolute units may facilitate more accurate planning and decision-making by aligning with intuitive cognitive processing.
+However, it's important to note that even intuitive formats can pose challenges.
+@weberWhyCanOnly2018 found that individuals often struggle with reasoning tasks presented in natural frequencies because they inadvertently revert to more complex probabilistic thinking.
+This suggests that merely changing the information format may not be sufficient; understanding how consumers process and utilize this information is crucial.
+
+Although prior research shows that energy-use information must be intelligible and action-oriented [@canfieldPerceptionsElectricityuseCommunications2017; @attariPublicPerceptionsEnergy2010], many consumers still have difficulty converting broad guidance (“use less energy”) into targeted appliance-level plans.
+As highlighted by @abrahamseReviewInterventionStudies2005, information alone, without concrete guidance on how to distribute reductions across appliances or times of use, may fail to encourage meaningful change.
+Moreover, a meta-analysis by @nematiImpactInformationbasedInterventions2020 revealed that interventions featuring frequent, household-level feedback, particularly when combined with social norm information and monetary incentives, are significantly more effective in driving conservation.
+Similar findings emerge in contexts beyond electricity.
+For instance, @tonkeProvidingProceduralKnowledge2024 reports that offering residents a short but actionable list of water-saving strategies—rather than a generic call to conserve—led to a measurable decrease in residential water consumption.
+These results suggest that equipping consumers with procedure-focused instructions can complement abstract energy-saving goals.
 
@@
 
-For each family scenario, the participants were shown a table containing the families utility usage from the prior year, alongside the state averages for each appliance category (both prior year usage and stage averages are always shown in kWh).
+For each family scenario, the participants were shown a table containing the families utility usage from the prior year, alongside the state averages for each appliance category (both prior year usage and state averages are always shown in kWh).
 For each scenario, participants were asked to create two possible action plans to achieve the target reduction in total household energy usage (see @fig-task).
@@
 Percentage vs.
-kWh)..
+kWh).
 We recruited 206 participants from Amazon Mechanical Turk, but data from from 10 participants were corrupted due to experimenter error, leaving a final sample of 196 participants.
@@
 
-We once again examained the effect of energy literacy on planning accuracy.
+We once again examined the effect of energy literacy on planning accuracy.
 A Bayesian linear regression model was fit with log-transformed absolute error as the outcome variable and energy literacy score as the predictor, controlling for random effects of participant and state: log_abs_error ~ els + (1|id) + (1|state).
@@
 
-# General Discusion
+# General Discussion
 

### 2025-01-04 manuscript/appendix.qmd
add demo columns to s1

@@
 ```{r}
-
+#| eval: false
 # widen by appliance, with separate row for value, family, and stage_avg
@@
 
-    ‘
 s2s |>  ungroup() |> select(-id,-block,-refClass,-pct,-plan) |> scale(center = TRUE, scale = TRUE) |> kmeans(centers = 3) 

### 2025-01-04 nb/s1_survey.qmd
add demo columns to s1

@@
 )) |> 
-  rename(calc=MATH01) |>  # replace NA's in calc with 0's
-  rename(id2=ResponseId) |>
-  mutate(id=row_number()) |> relocate(id) |> 
+  rename(calc=MATH01, 
+         id2=ResponseId,
+         gender=DEM01) |>
+  mutate(id=row_number()) |> 
+  relocate(id) |> 
+  mutate(gender = factor(gender, 
+                        levels = c(1,2,3),
+                        labels = c("Male", "Female", "Not specified"))) |> 
   mutate(calc=factor(ifelse(is.na(calc),0,calc))) |> 
-  relocate(refClass,state1,state2,calc, .after = id)
+  relocate(refClass,state1,state2,calc,gender, .after = id)
 

### 2025-01-04 nb/study1.qmd
add demo columns to s1

@@
 
-s1 <- haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav"))
+s1 <- haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav")) |> select(-IPAddress)
 
@@
 )) |> 
-  rename(calc=MATH01) |>  # replace NA's in calc with 0's
-  rename(id2=ResponseId) |>
-  mutate(id=row_number()) |> relocate(id) |> 
+  rename(calc=MATH01, 
+         id2=ResponseId,
+         gender=DEM01,
+         byear=DEM02,
+         edu=DEM04,
+         duration=Duration__in_seconds_) |>
+  mutate(id=row_number(),
+  age=as.numeric(format(RecordedDate, "%Y")) - byear) |> 
+    mutate(edu = factor(case_when(
+    edu == 1 ~ "Some schooling, but no diploma or degree",
+    edu == 2 ~ "Highschool diploma or GED",
+    edu == 3 ~ "Some college",
+    edu == 4 ~ "College degree",
+    edu == 5 ~ "Some graduate school",
+    edu == 6 ~ "Graduate degree",
+    edu == 7 ~ "Choose not to answer",
+    TRUE ~ NA_character_
+  ))) |>
+  relocate(id) |> 
+  mutate(gender = factor(gender, 
+                        levels = c(1,2,3),
+                        labels = c("Male", "Female", "Not specified"))) |> 
   mutate(calc=factor(ifelse(is.na(calc),0,calc))) |> 
-  relocate(refClass,state1,state2,calc, .after = id)
+  relocate(refClass,state1,state2,calc,gender,byear,age,RecordedDate,edu, .after = id)
 
 # Energy Literacy Score (ELS)
-
 els <- s1 |> select(id,id2,calc,refClass,state1,state2, ELS01:ELS08)
@@
     summarize(els=mean(correct=="Correct"), .groups="drop")
-
-s1 <- s1 |> rename(edu=DEM04) |>
-  mutate(edu = factor(case_when(
-    edu == 1 ~ "Some schooling, but no diploma or degree",
-    edu == 2 ~ "Highschool diploma or GED",
-    edu == 3 ~ "Some college",
-    edu == 4 ~ "College degree",
-    edu == 5 ~ "Some graduate school",
-    edu == 6 ~ "Graduate degree",
-    edu == 7 ~ "Choose not to answer",
-    TRUE ~ NA_character_
-  ))) |> relocate(edu, .after = id)
 
@@
       values_to = "value"
-    ) |> select(id,id2,state1,state2,refClass,calc,edu,modality,state,item_num,value,Duration__in_seconds_)
+    ) |> select(id,id2,state1,state2,refClass,calc,age,gender,edu,modality,state,item_num,value,duration)
 
@@
 
-#saveRDS(s1, here::here("data/s1_processed.rds"))
+saveRDS(s1, here::here("data/s1_processed.rds"))
 

### 2025-01-04 nb/study2.qmd
add s2 demo extras

@@
 )) |> 
-  rename(calc=MATH01) |>  # replace NA's in calc with 0's
-  mutate(calc=factor(ifelse(is.na(calc),0,calc))) |> 
-  mutate(calc = if_else(calc == 0, "No Calculator", "Calculator")) |> 
-  relocate(refClass,state1,state2,calc, .after = id)
-
-s2 <- s2 |> rename(edu=DEM04) |>
+  rename(calc=MATH01, 
+         gender=DEM01,
+         byear=DEM02,
+         edu=DEM04) |> 
+  mutate(duration=as.numeric(difftime(lubridate::as_datetime(EndDate), lubridate::as_datetime(StartDate), units = "secs")),
+  age=as.numeric(format(StartDate, "%Y")) - byear) |> 
   mutate(edu = factor(case_when(
@@
     TRUE ~ NA_character_
-  ))) |> relocate(edu, .after = id)
+  ))) |> 
+  mutate(gender = factor(gender, 
+                      levels = c(1,2,3),
+                      labels = c("Male", "Female", "Not specified"))) |> 
+  mutate(calc=factor(ifelse(is.na(calc),0,calc))) |>  # replace NA's in calc with 0's
+  mutate(calc = if_else(calc == 0, "No Calculator", "Calculator")) |> 
+  relocate(refClass,state1,state2,calc,gender,age,byear,edu,duration, .after = id)
 
@@
 
-s2_b1 <- s2 |> select(id,state1,refClass,TENPERCENT,ROUNDED, calc, edu,ACFAMILY1:TOTALAPTWO1) |> 
+s2_b1 <- s2 |> select(id,state1,refClass,TENPERCENT,ROUNDED, calc,age,gender,edu,duration,ACFAMILY1:TOTALAPTWO1) |> 
   mutate( 
@@
 
-s2_b2 <- s2 |> select(id,state2,refClass,TENPERCENT,ROUNDED, calc,edu,ACFAMILY2:TOTALAPTWO2) |> 
+s2_b2 <- s2 |> select(id,state2,refClass,TENPERCENT,ROUNDED,calc,age,gender,edu,duration,ACFAMILY2:TOTALAPTWO2) |> 
   mutate( 

### 2025-01-04 manuscript/appendix.qmd
fix echo issues

@@
 s1 <- readRDS(here::here("data/s1_processed.rds")) |> 
-  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |> 
-  filter(!(id %in% readRDS(here::here("data/s1_grp_outlier_ids.rds")))) |>
+  filter(!(id %in% readRDS(here::here("data/s1_discrep_ids.rds")))) |>  # n=17
+  filter(!(id %in% readRDS(here::here("data/s1_grp_outlier_ids.rds")))) |> # n=6
   mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))
@@
 s2_long <- readRDS(here::here("data/s2_processed.rds")) |> 
-  filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> 
-  filter(!(id %in% readRDS(here::here("data/s2_grp_outlier_ids.rds"))) ) |>
+  filter(!(id %in% readRDS(here::here("data/s2_discrep_ids.rds")))) |> # n=10
+  filter(!(id %in% readRDS(here::here("data/s2_grp_outlier_ids.rds"))) ) |> # n=6
   mutate(refClass = factor(refClass, levels=c("kWh","Percentage","USD")))
@@
                 n_less_avg = sum(less_avg),
-                duration=first(Duration__in_seconds_)) |> 
+                duration=first(duration)) |> 
     mutate(matched_goal = (pct_change == pct_goal), 
@@
     state_dif = mean(state_dif),
-    .groups = "drop"
-  ) |>
+    duration=first(duration),.groups = "drop") |> 
   mutate(
@@
 s2 |> group_by(appliance,family,state,state_avg) |> select(appliance,family_prior_yr=family,state,state_avg)
+```
+
+# Demographics S1
+
+```{r}
+
+n_exclusion <-readRDS(here::here("data/s1_grp_outlier_ids.rds")) |> length() # 6
+
+average_age <- s1 |> ungroup() |> 
+  summarise(average_age = mean(age, na.rm = TRUE), sd_age = sd(age, na.rm = TRUE))
+
+#   average_age sd_age
+#         <dbl>  <dbl>
+# 1        34.3   10.2
+
+gender_counts <- s1 |> 
+  group_by(id,gender) |>
+  slice(1) |>
+  ungroup() |> group_by(gender) |> 
+  summarise(count = n()) |>
+  ungroup()
+
+#   gender        count
+#   <fct>         <int>
+# 1 Male            136
+# 2 Female           92
+# 3 Not specified     1
+
+```
+
+# Demographics S2
+
+```{r}
+
+n_exclusion <-readRDS(here::here("data/s2_grp_outlier_ids.rds")) |> length() # 6
+
+average_age <- s2_long |> ungroup() |> 
+  summarise(average_age = mean(age, na.rm = TRUE), sd_age = sd(age, na.rm = TRUE))
+  # average_age sd_age
+  #       <dbl>  <dbl>
+  #        35.5   9.47
+
+gender_counts <- s2_long |> 
+  group_by(id,gender) |> 
+  slice(1) |>
+  ungroup() |> group_by(gender) |>
+  summarise(count = n()) |>
+  ungroup()
+#   gender count
+#   <fct>  <int>
+# 1 Male     102
+# 2 Female    88
+
 ```

### 2025-01-04 manuscript/manuscript.qmd
fix echo issues

@@
         output-file: rep_study.pdf
+        echo: false
         mainfont: "Linux Libertine O"

### 2025-01-09 manuscript/appendix.qmd
todays  updates - intro, discussion, few methods tweaks

@@
 
-# A tibble: 4,560 × 17
-# Groups:   id, block [380]
-   id     block refClass appliance pct_goal rounded state    calc       edu                         els   pct family state_avg plan  value change state_dif
-   <chr>  <dbl> <fct>    <chr>     <chr>    <chr>   <chr>    <chr>      <fct>                     <dbl> <dbl>  <dbl>     <dbl> <fct> <dbl>  <dbl>     <dbl>
- 1 s2_100     1 kWh      AC        10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1   1000       498 plan1   500   -500         2
- 2 s2_100     1 kWh      AC        10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1   1000       498 plan2   300   -700      -198
- 3 s2_100     1 kWh      HEAT      10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1  18000     16411 plan1 15000  -3000     -1411
- 4 s2_100     1 kWh      HEAT      10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1  18000     16411 plan2 12000  -6000     -4411
- 5 s2_100     1 kWh      WATER     10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1  12000      5832 plan1  6000  -6000       168
- 6 s2_100     1 kWh      WATER     10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1  12000      5832 plan2  2000 -10000     -3832
- 7 s2_100     1 kWh      REFRIG    10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1   1000      1142 plan1  2000   1000       858
- 8 s2_100     1 kWh      REFRIG    10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1   1000      1142 plan2   500   -500      -642
+# Subject Counts
 
 ```{r}
-#| eval: false
-# widen by appliance, with separate row for value, family, and stage_avg
 
-s2 <- s2_long #|> select(-family,-state_avg,-edu,-change,-state_dif) |> pivot_wider(names_from = appliance, values_from = c(value)) 
-s2s <- s2_long |> select(-family,-state_avg,-edu,-change,-value) |> pivot_wider(names_from = appliance, values_from = c(state_dif)) |> select(id,block,refClass,pct,plan,AC:OTHER)
-# s2s
-#    id     block refClass     pct plan     AC   HEAT WATER REFRIG OTHER
-#    <chr>  <dbl> <fct>      <dbl> <fct> <dbl>  <dbl> <dbl>  <dbl> <dbl>
-#  1 s2_100     1 kWh         0.1  plan1     2  -1411   168    858   348
-#  2 s2_100     1 kWh         0.1  plan2  -198  -4411 -3832   -642 -6152
-#  3 s2_100     2 kWh         0.15 plan1  1324     19  -139    320  -424
-#  4 s2_100     2 kWh         0.15 plan2   324   -981 -1139   -318 -1424
-#  5 s2_101     1 kWh         0.1  plan1   102   -411  2168   -142   398
-#  6 s2_101     1 kWh         0.1  plan2   502  -1411   168   -142  1348
-#  7 s2_101     2 kWh         0.15 plan1   751    901  -396    682 -1883
-#  8 s2_101     2 kWh         0.15 plan2  2251    401   104    682 -1383
-#  9 s2_102     1 Percentage  0.1  plan1   958 -12954 -5354    200 -5417
-# 10 s2_102     1 Percentage  0.1  plan2   890 -15435 -4956   8544  3225
+s1 %>%
+  group_by(id) %>%
+  summarise(total_subjects = n_distinct(id)) %>%
+  ungroup() %>%
+  summarise(total_subjects = n_distinct(id)) %>%
+  bind_cols(
+    s1 %>%
+      group_by(refClass) %>%
+      summarise(subjects_per_refClass = n_distinct(id)) %>%
+      ungroup()
+  ) %>%
+  print()
 
-# clustering planning strategies - kmeans
+s2_long %>%
+  group_by(id) %>%
+  summarise(total_subjects = n_distinct(id)) %>%
+  ungroup() %>%
+  summarise(total_subjects = n_distinct(id)) %>%
+  bind_cols(
+    s2_long %>%
+      group_by(refClass) %>%
+      summarise(subjects_per_refClass = n_distinct(id)) %>%
+      ungroup()
+  ) %>%
+  print()
 
-s2s |>  ungroup() |> select(-id,-block,-refClass,-pct,-plan) |> scale(center = TRUE, scale = TRUE) |> kmeans(centers = 3) 
-s2s_clust <- s2s |>  ungroup() |> select(-id,-block,-refClass,-pct,-plan) |> scale(center = TRUE, scale = TRUE) |> kmeans(centers = 3) %>% .$cluster |> as.factor() 
-s2s <- cbind(s2s,c=s2s_clust)
-
-s2s |>    group_by(c) |>
-  summarise(ac=mean(AC),heat=mean(HEAT),water=mean(WATER),refrig=mean(REFRIG),other=mean(OTHER)) |>
-  pivot_longer(cols = c(ac:other), names_to = "Appliance", values_to = "Mean") |>
-  ggplot(aes(x=Appliance,y=Mean,col=c)) + 
-  geom_point() + 
-  facet_wrap(~c) +
-  theme_minimal()
-
-s2s <- s2_long |> select(-family,-state_avg,-edu,-change,-state_dif) |> pivot_wider(names_from = appliance, values_from = c(value))  |> select(id,block,refClass,pct,plan,AC:OTHER)
-
-s2s |>  ungroup() |> select(-id,-block,-refClass,-pct,-plan) |> scale(center = TRUE, scale = TRUE) |> kmeans(centers = 3)
-
-s2s_clust <- s2s |>  ungroup() |> select(-id,-block,-refClass,-pct,-plan) |> scale(center = TRUE, scale = TRUE) |> kmeans(centers = 3) %>% .$cluster |> as.factor()
-
-s2s <- cbind(s2s,c=s2s_clust) |> left_join(s2, join_by(id,refClass,block,plan,pct))
-s2s |>  group_by(c) |>
-  summarise(ac=mean(AC),heat=mean(HEAT),water=mean(WATER),refrig=mean(REFRIG),other=mean(OTHER)) |>
-  pivot_longer(cols = c(ac:other), names_to = "Appliance", values_to = "Mean") |>
-  ggplot(aes(x=Appliance,y=Mean,col=c)) + 
-  geom_point() + 
-  geom_point(aes(y=state_avg),color="black") +
-  facet_wrap(~c) +
-  theme_minimal()
-
-s2 |> group_by(appliance,family,state,state_avg) |> select(appliance,family_prior_yr=family,state,state_avg)
 ```
@@
 
-n_exclusion <-readRDS(here::here("data/s1_grp_outlier_ids.rds")) |> length() # 6
+n_exclusion_s1 <-readRDS(here::here("data/s1_grp_outlier_ids.rds")) |> length() # 6
 
@@
   slice(1) |>
-  ungroup() |> group_by(gender) |> 
-  summarise(count = n()) |>
-  ungroup()
+  group_by(gender) |> 
+  summarise(count = n())
 
@@
 
+s1 %>% ungroup() |> 
+  summarise(
+    total_subjects = n_distinct(id),
+    .groups = 'drop'
+  ) %>%
+  left_join(
+    s1 %>%
+      group_by(refClass) %>%
+      summarise(
+        subjects_in_class = n_distinct(id),
+        .groups = 'drop'
+      )
+  ) %>%
+  print()
+
+ s1 %>% summarise(
+      total_subjects = n(),
+      reference_class_counts = list(
+        distinct(., id, !!sym(refClass)) %>% count(!!sym(refClass))
+      )
+    )
+
 ```
@@
 
-n_exclusion <-readRDS(here::here("data/s2_grp_outlier_ids.rds")) |> length() # 6
+n_exclusion_s2 <-readRDS(here::here("data/s2_grp_outlier_ids.rds")) |> length() # 6
 
@@
   slice(1) |>
-  ungroup() |> group_by(gender) |>
-  summarise(count = n()) |>
-  ungroup()
+  group_by(gender) |>
+  summarise(count = n()) 
 #   gender count

### 2025-01-09 manuscript/manuscript.qmd
todays  updates - intro, discussion, few methods tweaks

@@
 
-Electricity bills are a frequent touchpoint for consumers seeking to understand their energy usage, yet effectively turning that billing data into practical steps can be demanding.
-Among low-income households, these challenges contribute to disproportionate rates of energy insecurity [@memmottSociodemographicDisparitiesEnergy2021].
-At the same time, previous research shows that the clarity and format of energy-use information play a pivotal role in whether individuals adopt meaningful conservation behaviors [@fischerFeedbackHouseholdElectricity2008; @attariPublicPerceptionsEnergy2010; @canfieldPerceptionsElectricityuseCommunications2017].
-The interplay of these factors indicates that how energy data is presented can substantially shape not only how people interpret their consumption but also the degree to which they modify their habits to save energy.
+Energy costs often impose a significant burden on low-income households, leading to "energy insecurity," where basic energy needs cannot be met (Bednar & Reames, 2020).
+Frequently measured as the energy burden—the percentage of income spent on energy bills—this burden can be disproportionately high for vulnerable families, necessitating difficult trade-offs with essentials such as food or medicine [@bednarRecognitionResponseEnergy2020; @memmottSociodemographicDisparitiesEnergy2021].
+Energy insecurity has been linked to health risks and unsafe coping strategies, disproportionately impacting racial and ethnic minorities through higher rates of disconnection [@memmottSociodemographicDisparitiesEnergy2021].
+While the broader context of climate change, partly driven by residential consumption [@farghaliStrategiesEnergyContext2023], underscores the need for sustainable solutions, the financial strain on vulnerable households remains a pressing concern.
+Promoting behavior change to reduce energy consumption is crucial.
+However, the success of such interventions may hinge on how effectively energy information is communicated, with format and presentation context significantly influencing understanding and action [@canfieldPerceptionsElectricityuseCommunications2017; @fischerFeedbackHouseholdElectricity2008].
 
-The way numerical information is presented can significantly affect how individuals process and use that information [@reimerNumericCommunicationRisk2015].
-The reference class problem highlights that numbers without clear reference points can lead to misinterpretation, as the meaning of a statistic depends on the category or class it refers to [@gigerenzerSimpleToolsUnderstanding2003; @reimerNumericCommunicationRisk2015].
-Presenting energy information in absolute units (e.g., kWh) provides a clear reference class, potentially enhancing comprehension.
-Indeed, the manner in which energy information is presented has been shown to significantly influence both comprehension and subsequent behavior [@canfieldPerceptionsElectricityuseCommunications2017].
-For example, the use of tables has been found to facilitate point-reading, essential for understanding specific energy values, while graphs may be more useful in tasks requiring more complex information interpretation [@canfieldPerceptionsElectricityuseCommunications2017].
+The way numerical information is presented can significantly affect how individuals process and use that information.
+Of particular relevance are reference class effects, which occur when numerical statements are presented without a clear or intuitive basis for comparison, making it difficult to infer meaningful quantities[@gigerenzerSimpleToolsUnderstanding2003; @reimerNumericCommunicationRisk2015].
+A substantial body of evidence suggests that presenting data in terms of absolute counts or frequencies, as opposed to probabilities or percentages, can promote more accurate comprehension and facilitate.
+However, it's important to note that even intuitive formats can pose challenges.
+For instance, @weberWhyCanOnly2018 found that individuals often struggle with reasoning tasks presented in natural frequencies because they inadvertently revert to more complex probabilistic thinking.
 
-The concept of cognitive fit posits that performance improves when the information presentation format aligns with the task requirements [@vesseyCognitiveFitTheoryBased1991], and that such an alignment can reduce cognitive load and enhance accuracy in planning [@shahBarLineGraph2011].
-For instance, tables are generally more effective than graphs for conveying specific electricity usage data because they facilitate straightforward point reading [@canfieldPerceptionsElectricityuseCommunications2017].
-However, the effectiveness of the format varies with the type of information and individual differences, such as energy literacy, which significantly impacts comprehension and conservation intent.
-Moreover, the unit in which numerical information is presented influences how decision-makers evaluate and choose between options, with default units increasing value sensitivity [@herberzKiloWhatDefault2020].
-In the context of energy, presenting information in terms of multiple translations can increase preference for options aligned with activated objectives, such as pro-environmental values [@ungemachTranslatedAttributesChoice2018].
-Furthermore, mental accounting mechanisms, where individuals create mental budgets linking specific consumption acts to specific payments, significantly impact energy decisions and behaviors [@hahnelMentalAccountingMechanisms2020].
+The choice of units and the format in which information is presented have also been shown to exert a significant influence on decision-making specifically within the context of energy consumption and planning.
+For instance, tables are generally more effective than graphs for conveying specific electricity usage data because they facilitate straightforward point reading  @canfieldPerceptionsElectricityuseCommunications2017.
+Furthermore, the framing of energy costs, such as displaying monthly rather than daily or yearly expenses, can significantly affect consumers' choices [@gillConsumerMatchingCosts2022].
+Consequently, the selection of an appropriate information format is crucial for effectively supporting energy-related decisions.
+Therefore, similar to natural frequencies, kWh provide a direct measure of energy use, a characteristic that could simplify calculations and facilitate comparisons, potentially helping consumers better understand and compare the energy consumption of different appliances or activities.
+Conversely, prior research has suggested that consumers have a preference for receiving energy feedback in terms of monetary values over scientific units [@karjalainenConsumerPreferencesFeedback2011; @nematiImpactInformationbasedInterventions2020], as well as better long-term appliance selection with information presented in monetary terms [@blaschBoundedlyRationalConsumers2019].
+However, it remains uncertain how these reported benefits might generalize to the more complex domain of household energy planning, where usage patterns can be multifaceted and subject to a variety of contextual influences.
 
-Evidence from research on energy consumption feedback, normative comparisons, and eco-feedback platforms suggests that comprehensible and contextually meaningful data presentations can improve users’ ability to plan reductions, especially when these formats are integrated into daily routines [@canfieldPerceptionsElectricityuseCommunications2017; @fischerFeedbackHouseholdElectricity2008; @kimMySmartEEcofeedbackGaming2022; @schwartzWhatPeopleConsumption2015].
-Furthermore, temporal and monetary frames have been shown to alter decision quality, with monthly costs or absolute consumption levels often encouraging more energy-efficient intentions than abstract annual or percentage-based metrics [@gillConsumerMatchingCosts2022; @larrickMPGIllusion2008].
-In this context, tailoring reference classes to align with intuitive cognitive processes can help bridge the gap between aggregate reduction goals and targeted, appliance-specific conservation strategies.
+Although many individuals express a desire to conserve energy, research consistently shows that abstract goals (e.g., “reduce overall usage by 15%”) often fail to translate into effective behavior change unless accompanied by specific, actionable steps [@abrahamseReviewInterventionStudies2005; @nematiImpactInformationbasedInterventions2020].
+For instance, @abrahamseReviewInterventionStudies2005 demonstrated that merely providing general information about energy savings rarely alters consumption patterns unless consumers also receive concrete instructions or tailored feedback.
+Similarly, @tonkeProvidingProceduralKnowledge2024 reported that sending households brief but precise text messages outlining how to reduce water use (e.g., limiting irrigation times, adjusting washing machine settings) yielded meaningful decreases in consumption, underscoring the importance of procedural knowledge—namely, knowing how to operationalize a goal rather than simply why it is desirable.
+In the context of energy conservation, this implies that interventions should not only highlight potential reductions (such as a 15% target) but also guide residents in allocating those reductions across specific appliances or behaviors [@attariPublicPerceptionsEnergy2010].
+Additionally, meta-analytic findings suggest that people respond more robustly to household-level feedback that situates their usage within a personalized framework, thereby reducing the cognitive burden of figuring out next steps [@nematiImpactInformationbasedInterventions2020].
 
-Furthermore, research suggests that natural frequencies and absolute numbers are generally easier for individuals to understand compared to percentages or probabilities [@gigerenzerHowImproveBayesian1995 @hoffrageCommunicatingStatisticalInformation2000].
-In the context of energy conservation, using absolute units may facilitate more accurate planning and decision-making by aligning with intuitive cognitive processing.
-However, it's important to note that even intuitive formats can pose challenges.
-@weberWhyCanOnly2018 found that individuals often struggle with reasoning tasks presented in natural frequencies because they inadvertently revert to more complex probabilistic thinking.
-This suggests that merely changing the information format may not be sufficient; understanding how consumers process and utilize this information is crucial.
-
-Although prior research shows that energy-use information must be intelligible and action-oriented [@canfieldPerceptionsElectricityuseCommunications2017; @attariPublicPerceptionsEnergy2010], many consumers still have difficulty converting broad guidance (“use less energy”) into targeted appliance-level plans.
-As highlighted by @abrahamseReviewInterventionStudies2005, information alone, without concrete guidance on how to distribute reductions across appliances or times of use, may fail to encourage meaningful change.
-Moreover, a meta-analysis by @nematiImpactInformationbasedInterventions2020 revealed that interventions featuring frequent, household-level feedback, particularly when combined with social norm information and monetary incentives, are significantly more effective in driving conservation.
-Similar findings emerge in contexts beyond electricity.
-For instance, @tonkeProvidingProceduralKnowledge2024 reports that offering residents a short but actionable list of water-saving strategies—rather than a generic call to conserve—led to a measurable decrease in residential water consumption.
-These results suggest that equipping consumers with procedure-focused instructions can complement abstract energy-saving goals.
-
-Despite existing studies on energy-use communication and format effects, limited research has explored how different numerical representations influence consumers’ ability to create accurate energy conservation plans.
+Despite existing studies on energy-use communication and format effects, limited research has explored how different numerical representations influence consumers' ability to create accurate energy conservation plans.
 Specifically, there is a gap in understanding how presenting energy information in absolute units versus percentages or monetary terms affects the precision of planning appliance-specific reductions.
 The current study addresses these critical issues by systematically investigating the impact of varying information formats (kWh, percentage, and USD) on the accuracy of energy-planning decisions.
-By manipulating the presentation format of energy information, this research aims to elucidate how different representational formats influence planning accuracy.
-
-### Hypotheses
-
-Building on these findings and informed by prior work showing that frequencies (like absolute units in kWh) are easier to comprehend and facilitate more precise decision-making compared to percentages, our study also utilizes a tabular format, but manipulates whether participants must consider energy information presented as absolute units (kWh), percentages (%), or monetary costs (USD).
-We hypothesize that presenting information in absolute units (kWh) will lead to more accurate household energy conservation planning.
+By manipulating the presentation format of energy information, this research aims to elucidate how different representational formats influence planning accuracy.Based on the literature reviewed, we hypothesize that: 1) Presenting energy reduction goals in absolute units (kWh) will lead to greater planning accuracy compared to percentage-based or monetary formats, as absolute units provide a more direct and less ambiguous representation of energy quantities.
+2) Higher energy literacy will be associated with more accurate planning, as individuals with greater energy knowledge may be better equipped to process and utilize the provided information, regardless of format.
+We also examine the potential of several exploratory variables, such as goal difficulty and the rounding of numerical values, to further elucidate the factors that influence planning accuracy.
 
@@
 
-Participants received energy usage data for two hypothetical families and were tasked with creating action plans to meet specified reduction goals by allocating usage across five appliance categories:  heating, cooling, water heating, refrigerator, and and other appliances (e.g., TV, lighting).
+Participants were provided with energy usage data for two hypothetical families and tasked with creating action plans to meet specified energy reduction goals.
+These goals were implementing by allocating usage across five appliance categories: heating, cooling, water heating, refrigerator, and other appliances (e.g., TV, lighting).
 
-For each family scenario, the participants were shown a table containing the families utility usage from the prior year, alongside the state averages for each appliance category (both prior year usage and state averages are always shown in kWh).
+For each family scenario, participants were shown a table containing the families utility usage from the prior year, alongside the state averages for each appliance category (both prior year usage and state averages are always shown in kWh).
 For each scenario, participants were asked to create two possible action plans to achieve the target reduction in total household energy usage (see @fig-task).
@@
 
-![Example trial in the energy planning task.
-Participants are shown the prior year electricity use of a household, and are tasked with creating a plan for the next year that will meet the energy reduction goal.
-Study 1 manipulates the format of the reduction goal to be either a percentage (15% given as goal reduction), kilowatt hours (5965 kWh given), or USD (\$656)](/assets/images/Smith_10R_Wells_15E.png){#fig-task}
+kWh Condition: The target reduction is presented as an absolute number of kilowatt-hours (e.g., "Reduce total energy use by 5965 kWh").
+
+Percentage Condition: The target reduction is presented as a percentage of the family's total previous year's usage (e.g., "Reduce total energy use by 15%").
+
+USD Condition: The target reduction is presented as a monetary value, equivalent to the cost of a 15% reduction in energy use (e.g., "Reduce total energy costs by $656").
+
+![Example energy planning task trial.
+Participants saw a table with a family's previous year electricity usage (here for the Wells family in Colorado) and were asked to allocate energy usage to meet a 15% reduction goal.
+The format of the reduction goal was manipulated to be either a percentage (15% given as goal reduction), kilowatt hours (5965 kWh given), or USD (\$656)](/assets/images/Smith_10R_Wells_15E.png){#fig-task}
 
@@
                 n_less_avg = sum(less_avg),
-                duration=first(duration)) |> 
+                duration=first(duration), .groups = 'drop') |> 
     mutate(matched_goal = (pct_change == pct_goal), 
@@
                 close_match = abs_error <= 0.02) |>
-    ungroup() |> # Add ungroup here
         mutate(
@@
                     abs_error <= 0.05 ~ "0.01-5% error",
-                    TRUE ~ "Over 5% error"  # Capture all remaining cases
+                    TRUE ~ "Over 5% error" 
                 ), 
@@
     mean_abs_error=mean(abs_error),
-    mean_log_abs_error=mean(log_abs_error),
-    n_accuracy=first(n_accuracy)) |> 
+    mean_log_abs_error=mean(log_abs_error)) |> 
     mutate(accuracy_level = factor(
@@
                 mean_abs_error <= 0.05 ~ "0.01-5% error",
-                TRUE ~ "Over 5% error"  # Capture all remaining cases
+                TRUE ~ "Over 5% error" 
             ), 
@@
 #| label: fig-s1-log-dist
-#| fig-cap: "Study 1: Distribution of log absolute error by reference class."
+#| fig-cap: "Experiment 1: Distribution of the log of the absolute error between the participant's action plan and the reduction goal across different reference class conditions (kWh, Percentage, USD).
+The right side plots are further separated by calculator usage.
+A lower log absolute error suggests higher planning accuracy."
 #| fig-width: 11
@@
 
-s1_ld <- ggplot(s1_agg, aes(y = refClass, x = log_abs_error, color = refClass)) +
-  geom_density_ridges(aes(fill = refClass), alpha = 0.2, scale = 0.5,
+s1_ld <- ggplot(s1_agg, aes(y = refClass, x = log_abs_error, fill = refClass)) +
+  geom_density_ridges(aes(col = refClass), alpha = 0.2, scale = 0.5,
                       jittered_points = TRUE, point_alpha = 0.7,point_size=.4,
@@
                                                     ygap = 0.05)) +
-  geom_boxploth(width = 0.05, alpha = 0.2, outlier.shape = NA, show.legend = FALSE) +
+  geom_boxploth(width = 0.1, alpha = 0.2, outlier.shape = NA, show.legend = FALSE) +
   #scale_y_discrete(expand = expansion(mult = c(0.2, 0.4))) +
  # guides(fill = "none", color = guide_legend(reverse = TRUE)) +
-  guides(fill = "none", color = "none") +
+  guides(fill = "none") +
   labs(x = "Log Absolute Deviation", y = "Reference Class", color = "Reference Class") +
@@
 
-s1_ldc <- ggplot(s1_agg, aes(y = refClass, x = log_abs_error, color = calc)) +
-  geom_density_ridges(aes(fill = calc), alpha = 0.2, scale = 0.5,
+s1_ldc <- ggplot(s1_agg, aes(y = refClass, x = log_abs_error, fill = calc)) +
+  geom_density_ridges(aes(col = calc), alpha = 0.2, scale = 0.5,
                       jittered_points = TRUE, point_alpha = 0.7, point_size = .4,
@@
                                                     ygap = 0.05)) +
-  geom_boxploth(width = 0.05, alpha = 0.2, outlier.shape = NA, show.legend = FALSE) +
+  geom_boxploth(width = 0.1, alpha = 0.2, outlier.shape = NA, show.legend = FALSE) +
   scale_color_brewer(palette = "Set1") +
@@
 #| label: fig-s1-ppd
-#| fig-cap: "Study 1: Proportion of participants in each accuracy level, colored by reference class, and seprated in facets based on the levels of reduction goal.
-A larger % of participants in the Exact Match, or 0.01-2% error bins indicates better performance."
+#| fig-cap: "Experiment 1: Posterior predictive check for frequency of trials at each accuracy level, faceted by reference class (kWh, Percentage, USD).
+Bars show observed, dots show model predicted proportions.
+Better accuracy is indicated by higher proportions in the 'Exact Match' and '0.01-5% error' categories."
 
@@
                 log_abs_error=log(abs(error)+.007)) |> 
-      ungroup() |> 
         mutate(
@@
     'Log Abs. Deviation' = (median(mean_log_abs_error)),
-    # sd = sd(pct),
-    # n = n(),
-    #se=sd(pct)/sqrt(n)
 ) |>   mutate(across(where(is.numeric), \(x) round(x, 3))) %>% 
@@
 #| label: fig-s2-log-dist
-#| fig-cap: "Study 1: Distribution of log absolute error by reference class.
-Displays the distributions of log absolute deviation from the target reduction goal across the three reference class conditions and is also faceted by calculator usage.
-Lower values indicate greater accuracy."
+#| fig-cap: "Experiment 2: Distribution of the log of the absolute error between the participant's action plan and the reduction goal across different reference class conditions (kWh, Percentage, USD).
+The right side plots are further separated by calculator usage.
+A lower log absolute error suggests higher planning accuracy."
 #| fig-width: 11
@@
 
-s2_ld <- ggplot(s2_agg, aes(y = refClass, x = log_abs_error, color = refClass)) +
-  geom_density_ridges(aes(fill = refClass), alpha = 0.2, scale = 0.5,
+s2_ld <- ggplot(s2_agg, aes(y = refClass, x = log_abs_error, fill = refClass)) +
+  geom_density_ridges(aes(col = refClass), alpha = 0.2, scale = 0.5,
                       jittered_points = TRUE, point_alpha = 0.7,point_size=.4,
@@
                                                     ygap = 0.05)) +
-  geom_boxploth(width = 0.05, alpha = 0.2, outlier.shape = NA, show.legend = FALSE) +
+  geom_boxploth(width = 0.1, alpha = 0.3, outlier.shape = NA, show.legend = FALSE) +
   #scale_y_discrete(expand = expansion(mult = c(0.2, 0.4))) +
  # guides(fill = "none", color = guide_legend(reverse = TRUE)) +
-  guides(fill = "none", color = "none") +
+  guides(fill = "none") +
   labs(x = "Log Absolute Deviation", y = "Reference Class", color = "Reference Class") +
@@
 
-s2_ldc <- ggplot(s2_agg, aes(y = refClass, x = log_abs_error, color = calc)) +
-  geom_density_ridges(aes(fill = calc), alpha = 0.2, scale = 0.5,
+s2_ldc <- ggplot(s2_agg, aes(y = refClass, x = log_abs_error, fill = calc)) +
+  geom_density_ridges(aes(col = calc), alpha = 0.2, scale = 0.5,
                       jittered_points = TRUE, point_alpha = 0.7,point_size=.4,
@@
                                                     ygap = 0.05)) +
-  geom_boxploth(width = 0.05, alpha = 0.2, outlier.shape = NA, show.legend = FALSE) +
+  geom_boxploth(width = 0.1, alpha = 0.3, outlier.shape = NA, show.legend = FALSE) +
   scale_color_brewer(palette = "Set1") +
@@
 s2_ld + s2_ldc
+
 ```
@@
 #| fig-cap: "Experiment 2.
-Posterior predictive check of the ordinal probit model, grouped by reference class (Experiment 2).
-The bars represent the observed proportions of each accuracy level within each reference class.
+Posterior predictive check of the bayesian regression model, faceted by reference class.
+The bars represent the observed frequencies of each accuracy level within each reference class.
 The points represent the model's predicted proportions, with error bars indicating 95% credible intervals."
@@
 
-Experiment 2 aimed to build upon the findings of Experiment 1, not only by replicating the core manipulation of reference class, but also by incorporating additional variables that might influence planning accuracy.
+Experiment 2 aimed to build upon the findings of Experiment 1 by replicating the core manipulation of reference class.
+Additionally, it incorporated variables that might influence planning accuracy.
 These included goal difficulty and the way that the prior year's usage was presented (rounded or exact).
@@
 
-Across two experiments, we consistently found that presenting energy reduction goals in absolute units (kWh) led to more accurate planning compared to percentage-based or monetary representations.
-This advantage persisted across variations in goal difficulty and numerical presentation, suggesting a robust effect of reference class on planning accuracy.
+This study examined how different numerical representations of energy reduction goals influence consumers’ ability to create accurate energy conservation plans.
+Across two experiments, the findings consistently demonstrated that presenting reduction goals in absolute units (kWh) significantly enhanced planning accuracy compared to percentage-based or monetary formats.
+Furthermore, participants with higher energy literacy exhibited more precise planning across all conditions, underscoring the critical role of domain-specific knowledge in shaping decision-making outcomes.
 
-It is, however, important to note that while we find a significant benefit of presenting energy information in absolute units, some prior work suggests that consumers may prefer to receive information about costs, rather than absolute units (Karjalainen, 2011), or in some cases, that monetary information may lead to better decision-making (Blasch et al., 2019).
+These results align with the broader literature on numerical cognition, which suggests that the format in which information is presented can profoundly affect comprehension and decision-making  [@gigerenzerSimpleToolsUnderstanding2003; @reimerNumericCommunicationRisk2015].
+However, our study extends beyond simple estimations or judgments, demonstrating that the advantages of absolute units persist even in a more complex, multi-step planning task.
+Moreover, the consistent superiority of kWh over both percentages and USD provides novel insights into the specific challenges of energy-related decision-making.
+While prior research has suggested that consumers may prefer monetary feedback [@karjalainenConsumerPreferencesFeedback2011; @nematiImpactInformationbasedInterventions2020] or that monetary framing can improve long-term appliance choices [@blaschBoundedlyRationalConsumers2019], our findings indicate that when it comes to allocating specific usage reductions across appliances, absolute units are most effective.
+This may be because kWh provide a more direct and less ambiguous representation of energy quantities, facilitating the necessary calculations for accurate planning.
+Nevertheless, it is unclear whether the observed benefits of absolute units (i.e., kWh) are genuinely attributable to their absolute nature, or if other inherent characteristics of these units might be driving the effects.
+Furthermore, the percentage-based reduction targets, while potentially more salient from a goal-setting perspective, also resulted in poorer planning outcomes compared to kWh.
+Such a finding suggests that reliance on percentages can further complicate calculations by adding unnecessary transformations in the problem-solving process, in line with the "off by 100% bias" found in @fisher100BiasEffects2022, where they show that individuals often misunderstand percentage changes greater than 100%.
+The relative ease in using absolute values to arrive at an accurate plan suggests that these representations of the planning task result in the most accurate planning strategies because the number of steps required to perform the required calculations are simpler.
 
-Karjalainen 2011 - people prefer information about price [@karjalainenConsumerPreferencesFeedback2011]
+Moreover, the consistent positive relationship observed between energy literacy and planning accuracy across both experiments underscores the importance of baseline knowledge in effectively navigating energy-related information.
+Individuals with higher levels of energy literacy demonstrated a greater capacity to formulate accurate conservation plans, irrespective of the information format presented.
+This finding aligns with previous research highlighting the role of numeracy and domain-specific knowledge in improving judgments related to energy consumption [@attariPublicPerceptionsEnergy2010].
+It further suggests that interventions aimed at enhancing consumers' fundamental understanding of energy concepts could yield significant benefits in improving the effectiveness of energy conservation efforts.
 
-Mention @blaschBoundedlyRationalConsumers2019 - better long-term appliance selection with information presented in monetary terms.
+## Limitations
+
+- Each participant only completed 4 action plans
+- calculator use not controlled.
 

### 2025-01-09 misc/indiv.qmd
todays  updates - intro, discussion, few methods tweaks

@@
                 error = pct_change - pct_goal,
-            abs_error = abs(error))
+                abs_error = abs(error),
+                log_abs_error=log(abs(error)+.007), 
+                close_match = abs_error <= 0.02) |>
+    ungroup() |> # Add ungroup here
+        mutate(
+            accuracy_level = factor(
+                case_when(
+                    abs_error == 0.00 ~ "Exact match",
+                    abs_error <= 0.05 ~ "0.01-5% error",
+                    TRUE ~ "Over 5% error"  # Capture all remaining cases
+                ), 
+                levels = c("Exact match", "0.01-5% error", "Over 5% error"),
+                ordered = TRUE
+            )
+        ) |> relocate(accuracy_level, .after= "pct_change")
+
+s1_agg4 <- s1_agg |> group_by(id,refClass,calc) |> 
+    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+    summarise(
+    mg=sum(matched_goal),
+    mgc=sum(close_match),
+    n=n(), 
+    pct=mg/n,
+    pct_close=mgc/n,
+    mean_pct_change=mean(pct_change),
+    mean_abs_error=mean(abs_error),
+    mean_log_abs_error=mean(log_abs_error),
+    n_accuracy=first(n_accuracy)) |> 
+    mutate(accuracy_level = factor(
+            case_when(
+                mean_abs_error < 0.02 ~ "Exact match",
+                mean_abs_error <= 0.05 ~ "0.01-5% error",
+                TRUE ~ "Over 5% error"  # Capture all remaining cases
+            ), 
+            levels = c("Exact match", ".01-5% error", "Over 5% error"),
+            ordered = TRUE
+        ))
+```
+
+# Clustering
+
+# A tibble: 4,560 × 17
+# Groups:   id, block [380]
+   id     block refClass appliance pct_goal rounded state    calc       edu                         els   pct family state_avg plan  value change state_dif
+   <chr>  <dbl> <fct>    <chr>     <chr>    <chr>   <chr>    <chr>      <fct>                     <dbl> <dbl>  <dbl>     <dbl> <fct> <dbl>  <dbl>     <dbl>
+ 1 s2_100     1 kWh      AC        10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1   1000       498 plan1   500   -500         2
+ 2 s2_100     1 kWh      AC        10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1   1000       498 plan2   300   -700      -198
+ 3 s2_100     1 kWh      HEAT      10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1  18000     16411 plan1 15000  -3000     -1411
+ 4 s2_100     1 kWh      HEAT      10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1  18000     16411 plan2 12000  -6000     -4411
+ 5 s2_100     1 kWh      WATER     10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1  12000      5832 plan1  6000  -6000       168
+ 6 s2_100     1 kWh      WATER     10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1  12000      5832 plan2  2000 -10000     -3832
+ 7 s2_100     1 kWh      REFRIG    10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1   1000      1142 plan1  2000   1000       858
+ 8 s2_100     1 kWh      REFRIG    10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1   1000      1142 plan2   500   -500      -642
+
+```{r}
+#| eval: false
+# widen by appliance, with separate row for value, family, and stage_avg
+
+s2 <- s2_long #|> select(-family,-state_avg,-edu,-change,-state_dif) |> pivot_wider(names_from = appliance, values_from = c(value)) 
+s2s <- s2_long |> select(-family,-state_avg,-edu,-change,-value) |> pivot_wider(names_from = appliance, values_from = c(state_dif)) |> select(id,block,refClass,pct,plan,AC:OTHER)
+# s2s
+#    id     block refClass     pct plan     AC   HEAT WATER REFRIG OTHER
+#    <chr>  <dbl> <fct>      <dbl> <fct> <dbl>  <dbl> <dbl>  <dbl> <dbl>
+#  1 s2_100     1 kWh         0.1  plan1     2  -1411   168    858   348
+#  2 s2_100     1 kWh         0.1  plan2  -198  -4411 -3832   -642 -6152
+#  3 s2_100     2 kWh         0.15 plan1  1324     19  -139    320  -424
+#  4 s2_100     2 kWh         0.15 plan2   324   -981 -1139   -318 -1424
+#  5 s2_101     1 kWh         0.1  plan1   102   -411  2168   -142   398
+#  6 s2_101     1 kWh         0.1  plan2   502  -1411   168   -142  1348
+#  7 s2_101     2 kWh         0.15 plan1   751    901  -396    682 -1883
+#  8 s2_101     2 kWh         0.15 plan2  2251    401   104    682 -1383
+#  9 s2_102     1 Percentage  0.1  plan1   958 -12954 -5354    200 -5417
+# 10 s2_102     1 Percentage  0.1  plan2   890 -15435 -4956   8544  3225
+
+# clustering planning strategies - kmeans
+
+s2s |>  ungroup() |> select(-id,-block,-refClass,-pct,-plan) |> scale(center = TRUE, scale = TRUE) |> kmeans(centers = 3) 
+s2s_clust <- s2s |>  ungroup() |> select(-id,-block,-refClass,-pct,-plan) |> scale(center = TRUE, scale = TRUE) |> kmeans(centers = 3) %>% .$cluster |> as.factor() 
+s2s <- cbind(s2s,c=s2s_clust)
+
+s2s |>    group_by(c) |>
+  summarise(ac=mean(AC),heat=mean(HEAT),water=mean(WATER),refrig=mean(REFRIG),other=mean(OTHER)) |>
+  pivot_longer(cols = c(ac:other), names_to = "Appliance", values_to = "Mean") |>
+  ggplot(aes(x=Appliance,y=Mean,col=c)) + 
+  geom_point() + 
+  facet_wrap(~c) +
+  theme_minimal()
+
+s2s <- s2_long |> select(-family,-state_avg,-edu,-change,-state_dif) |> pivot_wider(names_from = appliance, values_from = c(value))  |> select(id,block,refClass,pct,plan,AC:OTHER)
+
+s2s |>  ungroup() |> select(-id,-block,-refClass,-pct,-plan) |> scale(center = TRUE, scale = TRUE) |> kmeans(centers = 3)
+
+s2s_clust <- s2s |>  ungroup() |> select(-id,-block,-refClass,-pct,-plan) |> scale(center = TRUE, scale = TRUE) |> kmeans(centers = 3) %>% .$cluster |> as.factor()
+
+s2s <- cbind(s2s,c=s2s_clust) |> left_join(s2, join_by(id,refClass,block,plan,pct))
+s2s |>  group_by(c) |>
+  summarise(ac=mean(AC),heat=mean(HEAT),water=mean(WATER),refrig=mean(REFRIG),other=mean(OTHER)) |>
+  pivot_longer(cols = c(ac:other), names_to = "Appliance", values_to = "Mean") |>
+  ggplot(aes(x=Appliance,y=Mean,col=c)) + 
+  geom_point() + 
+  geom_point(aes(y=state_avg),color="black") +
+  facet_wrap(~c) +
+  theme_minimal()
+
+s2 |> group_by(appliance,family,state,state_avg) |> select(appliance,family_prior_yr=family,state,state_avg)
+```
+
+```{r}
+
 ```

### 2025-01-12 assets/figure2.tex
updated methods and appendix.

\documentclass{article}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{graphicx}

\begin{document}

\begin{figure}[h]
\centering
\caption{Example energy planning task trials across conditions. Participants were asked to allocate energy usage to meet a 15\% reduction goal. The format of the reduction goal was manipulated to be either kilowatt hours (kWh), a percentage (\%), or U.S. dollars (USD).}

\begin{minipage}{0.48\textwidth}
\centering
\subcaption{kWh Condition}
\label{fig:kwh}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lrrr@{}}
\toprule
\textbf{Category} & \textbf{\begin{tabular}[c]{@{}c@{}}Electricity Used\\ Last Year by\\ the Wells Family\\ (kWh)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Average Electricity\\ Used Last Year by\\ Households in\\ Colorado (kWh)\end{tabular}} & \textbf{Action Plan} \\ \midrule
Cooling (Central A/C) & 697 & 498 &  \\
Heating the Home & 18,052 & 16,411 &  \\
Water Heating & 11,667 & 5,832 &  \\
Refrigerator & 1,370 & 1,142 &  \\
\begin{tabular}[c]{@{}l@{}}Other (Television,\\ Lighting, Electronics,\\ Washer/Dryer, etc.)\end{tabular} & 7,982 & 6,652 &  \\ \midrule
Total kWh & 39,768 & 30,535 &  \\ \midrule
\textbf{Target Reduction} & \multicolumn{3}{c}{\textbf{5,965 kWh}} \\ \bottomrule
\end{tabular}%
}
\end{minipage}\hfill
\begin{minipage}{0.48\textwidth}
\centering
\subcaption{Percentage Condition}
\label{fig:percent}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lrrr@{}}
\toprule
\textbf{Category} & \textbf{\begin{tabular}[c]{@{}c@{}}Electricity Used\\ Last Year by\\ the Wells Family\\ (kWh)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Average Electricity\\ Used Last Year by\\ Households in\\ Colorado (kWh)\end{tabular}} & \textbf{Action Plan} \\ \midrule
Cooling (Central A/C) & 697 & 498 &  \\
Heating the Home & 18,052 & 16,411 &  \\
Water Heating & 11,667 & 5,832 &  \\
Refrigerator & 1,370 & 1,142 &  \\
\begin{tabular}[c]{@{}l@{}}Other (Television,\\ Lighting, Electronics,\\ Washer/Dryer, etc.)\end{tabular} & 7,982 & 6,652 &  \\ \midrule
Total kWh & 39,768 & 30,535 &  \\ \midrule
\textbf{Target Reduction} & \multicolumn{3}{c}{\textbf{15\%}} \\ \bottomrule
\end{tabular}%
}
\end{minipage}

\bigskip % Add some space between the tables

\begin{minipage}{0.48\textwidth}
\centering
\subcaption{USD Condition}
\label{fig:usd}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}lrrr@{}}
\toprule
\textbf{Category} & \textbf{\begin{tabular}[c]{@{}c@{}}Electricity Used\\ Last Year by\\ the Wells Family\\ (kWh)\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Average Electricity\\ Used Last Year by\\ Households in\\ Colorado (kWh)\end{tabular}} & \textbf{Action Plan} \\ \midrule
Cooling (Central A/C) & 697 & 498 &  \\
Heating the Home & 18,052 & 16,411 &  \\
Water Heating & 11,667 & 5,832 &  \\
Refrigerator & 1,370 & 1,142 &  \\
\begin{tabular}[c]{@{}l@{}}Other (Television,\\ Lighting, Electronics,\\ Washer/Dryer, etc.)\end{tabular} & 7,982 & 6,652 &  \\ \midrule
Total kWh & 39,768 & 30,535 &  \\ \midrule
\textbf{Target Reduction} & \multicolumn{3}{c}{\textbf{\$656}} \\ \bottomrule
\end{tabular}%
}
\end{minipage}
\end{figure}

\end{document}

### 2025-01-12 assets/figure4.tex
updated methods and appendix.

\documentclass{article}
\usepackage{booktabs}
\usepackage{multirow}

\begin{document}

\begin{figure}[h]
\centering
\caption{Example energy planning task instructions and reduction goal summary. Participants were tasked with creating two action plans to help a hypothetical family reduce their energy consumption. The reduction goal was presented in one of three formats: kilowatt hours (kWh), percentage (\%), or U.S. dollars (USD).}
\label{fig:summary}
\begin{minipage}{0.8\textwidth}
\textbf{Task Instructions:}

"The \[Family Name] family wants to reduce its household electricity use next year. Please complete two possible action plans that will help the \[Family Name] family achieve this goal. Please enter how many kWh should be used next year by each appliance and the total kWh each plan would use. Enter only whole numbers. Try to provide close estimations. You may use a calculator to complete the task."
\end{minipage}

\bigskip

\begin{tabular}{@{}llrrr@{}}
\toprule
\textbf{Family} & \textbf{State} & \textbf{kWh Reduction} & \textbf{\% Reduction} & \textbf{USD Reduction} \\ \midrule
Adams & California & 3,401 & 15\% & \$510 \\
Smith & Texas & 4,357 & 15\% & \$479 \\
Wells & Colorado & 5,965 & 15\% & \$656 \\
Davis & Massachusetts & 6,944 & 15\% & \$1,042 \\ \bottomrule
\end{tabular}
\end{figure}

\end{document}

### 2025-01-12 assets/figure5.tex
updated methods and appendix.

\documentclass{article}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage[margin=1in]{geometry} % Adjust margins if needed

\begin{document}

% Common settings for all tables
\newcommand{\figwidth}{0.8\textwidth}
\newcommand{\tablefont}{\sffamily} % Sans-serif font for a cleaner look
\newcommand{\header}[1]{\textbf{#1}} % Bold column headers

% kWh Condition Figure
\begin{figure}[h]
\centering
\resizebox{\figwidth}{!}{%
\begin{tabular}{@{}lrr@{}}
\toprule
{\tablefont \header{Category}} & 
{\tablefont \begin{tabular}[c]{@{}c@{}}\header{Electricity Used Last}\\\header{Year by the Wells}\\\header{Family (kWh)}\end{tabular}} & 
{\tablefont \begin{tabular}[c]{@{}c@{}}\header{Average Electricity}\\\header{Used Last Year by}\\\header{Households in}\\\header{Colorado (kWh)}\end{tabular}} \\ 
\midrule
{\tablefont Cooling (Central A/C)} & {\tablefont 697} & {\tablefont 498} \\
{\tablefont Heating the Home} & {\tablefont 18,052} & {\tablefont 16,411} \\
{\tablefont Water Heating} & {\tablefont 11,667} & {\tablefont 5,832} \\
{\tablefont Refrigerator} & {\tablefont 1,370} & {\tablefont 1,142} \\
{\tablefont \begin{tabular}[c]{@{}l@{}}Other (Television, Lighting,\\ Electronics, Washer/Dryer,\\ etc.)\end{tabular}} & {\tablefont 7,982} & {\tablefont 6,652} \\ 
\midrule
{\tablefont Total kWh} & {\tablefont 39,768} & {\tablefont 30,535} \\
\bottomrule
\end{tabular}%
}
\bigskip

\textbf{The Wells family wants to reduce its household electricity use by 5,965 kWh next year.}

Please complete two possible action plans that will help the Wells family achieve this goal. Please enter how many kWh should be used next year by each appliance and the total kWh each plan would use. \textbf{Enter only whole numbers.} Try to provide close estimations. You may use a calculator to complete the task.

Note: The Wells family used 9,233 more kWh than the average household in Colorado last year.
\end{figure}

% Percentage Condition Figure
\begin{figure}[h]
\centering
\resizebox{\figwidth}{!}{%
\begin{tabular}{@{}lrr@{}}
\toprule
{\tablefont \header{Category}} & 
{\tablefont \begin{tabular}[c]{@{}c@{}}\header{Electricity Used Last}\\\header{Year by the Wells}\\\header{Family (kWh)}\end{tabular}} & 
{\tablefont \begin{tabular}[c]{@{}c@{}}\header{Average Electricity}\\\header{Used Last Year by}\\\header{Households in}\\\header{Colorado (kWh)}\end{tabular}} \\ 
\midrule
{\tablefont Cooling (Central A/C)} & {\tablefont 697} & {\tablefont 498} \\
{\tablefont Heating the Home} & {\tablefont 18,052} & {\tablefont 16,411} \\
{\tablefont Water Heating} & {\tablefont 11,667} & {\tablefont 5,832} \\
{\tablefont Refrigerator} & {\tablefont 1,370} & {\tablefont 1,142} \\
{\tablefont \begin{tabular}[c]{@{}l@{}}Other (Television, Lighting,\\ Electronics, Washer/Dryer,\\ etc.)\end{tabular}} & {\tablefont 7,982} & {\tablefont 6,652} \\ 
\midrule
{\tablefont Total kWh} & {\tablefont 39,768} & {\tablefont 30,535} \\
\bottomrule
\end{tabular}%
}
\bigskip

\textbf{The Wells family wants to reduce its household electricity use by 15\% next year.}

Please complete two possible action plans that will help the Wells family achieve this goal. Please enter how many kWh should be used next year by each appliance and the total kWh each plan would use. \textbf{Enter only whole numbers.} Try to provide close estimations. You may use a calculator to complete the task.

Note: The Wells family used 9,233 more kWh than the average household in Colorado last year.
\end{figure}

% USD Condition Figure
\begin{figure}[h]
\centering
\resizebox{\figwidth}{!}{%
\begin{tabular}{@{}lrr@{}}
\toprule
{\tablefont \header{Category}} & 
{\tablefont \begin{tabular}[c]{@{}c@{}}\header{Electricity Used Last}\\\header{Year by the Wells}\\\header{Family (kWh)}\end{tabular}} & 
{\tablefont \begin{tabular}[c]{@{}c@{}}\header{Average Electricity}\\\header{Used Last Year by}\\\header{Households in}\\\header{Colorado (kWh)}\end{tabular}} \\ 
\midrule
{\tablefont Cooling (Central A/C)} & {\tablefont 697} & {\tablefont 498} \\
{\tablefont Heating the Home} & {\tablefont 18,052} & {\tablefont 16,411} \\
{\tablefont Water Heating} & {\tablefont 11,667} & {\tablefont 5,832} \\
{\tablefont Refrigerator} & {\tablefont 1,370} & {\tablefont 1,142} \\
{\tablefont \begin{tabular}[c]{@{}l@{}}Other (Television, Lighting,\\ Electronics, Washer/Dryer,\\ etc.)\end{tabular}} & {\tablefont 7,982} & {\tablefont 6,652} \\ 
\midrule
{\tablefont Total kWh} & {\tablefont 39,768} & {\tablefont 30,535} \\
\bottomrule
\end{tabular}%
}
\bigskip

\textbf{The Wells family wants to reduce its household electricity bill by \$656 next year.}

Please complete two possible action plans that will help the Wells family achieve this goal. Please enter how many kWh should be used next year by each appliance and the total kWh each plan would use. \textbf{Enter only whole numbers.} Try to provide close estimations. You may use a calculator to complete the task.

Note: The Wells family used 9,233 more kWh than the average household in Colorado last year.
\end{figure}

\end{document}

### 2025-01-12 manuscript/appendix.qmd
updated methods and appendix.

@@
+---
+title: Appendix
+aliases:
+  - /appendix.html
+format: 
+  html: 
+    echo: true
+    code-fold: show
+    output-file: appendix-code.html
+    wordcount-banner: true
+    theme: 
+      - zephyr
+      - ../assets/custom.scss
+    toc: true
+    shift-heading-level-by: 1
+    code-summary: "Show the code"
+    title-block-banner: "#000D51"
+    include-in-header:
+        - text: |
+            <style type="text/css">
+            h2.hide {
+                display: none;
+            }
+            </style>
+  wordcount-html: 
+    output-file: appendix.html
+    wordcount-banner: true
+    echo: false
+    theme: 
+        - zephyr
+        - ../assets/custom.scss
+    toc: true
+    shift-heading-level-by: 1
+    title-block-banner: "#000D51"
+    include-in-header:
+        - text: |
+            <style type="text/css">
+            h2.hide {
+                display: none;
+            }
+            </style>
+  docx: default
+  #hikmah-manuscript-docx: default
+  hugo-md:
+    include: false
+    echo: false
+    html-math-method: mathjax
+    output-file: appendix_hugo.md
+    suppress-bibliography: false
+  gfm:
+    echo: true
+    output-file: appendix_gfm.md
+    suppress-bibliography: true
+  hikmah-manuscript-pdf:
+    output-file: rs_appendix.pdf
+    echo: false
+    mainfont: "Linux Libertine O"
+    mainfontoptions:
+    - "Numbers=Proportional"
+    - "Numbers=OldStyle"
+    mathfont: "Libertinus Math"
+    knitr:
+        opts_chunk:
+            dev: "cairo_pdf"
+format-links:   
+    - text: HTML
+      href: appendix.html
+      icon: file-minus
+    - text: HTML (with code)
+      href: appendix-code.html
+      icon: file-plus
+    - format: hikmah-manuscript-pdf
+      href: rs_appendix.pdf
+      text: PDF version
+      icon: file-pdf
+    - docx
+    - gfm
+
+---
+
 ## Energy Literacy Questions
 
-| **ENERGY LITERACY SURVEY** |  |  |
-| Dewaters and Powers 2011. Assesses participant<br>knowledge of energy names, use, and sources |  |  |
-| The amount of ELECTRICAL ENERGY<br>(ELECTRICITY) we use is measured in units called ...
-| ELS01 | 1 = "Kilowatt (kW)"<br>2 = "Kilowatt-hours (kWh)"<br>3 = "British Thermal Units (BTU)"<br>4 = "Volts (V)"<br>5 = "Horsepower (HP)" |
+### Energy Literacy/Knowledge Quiz
 
-| Question | Variable | Values |
-|--------------------------|--------------------------|---------------------|
-| The amount of ENERGY consumed by an electrical appliance is equal to the power rating of the appliance (watts or kilowatts) ...
-| ELS02 | 1 = "Multiplied by the cost of electricity"<br>2 = "Added to the cost of electricity"<br>3 = "Multiplied by the time it's used"<br>4 = "Divided by the time it's used"<br>5 = "Added to the time it's used" |
-| When you turn on an incandescent light bulb, which of the following energy conversion takes place?
-| ELS03 | 1 = "Electrical energy to radiant energy (light)"<br>2 = "Chemical energy to radiant energy (light)"<br>3 = "Electrical energy to radiant energy (light) and thermal energy (heat)"<br>4 = "Chemical energy to radiant energy (light) and thermal energy (heat)"<br>5 = "Electrical energy to radiant energy (light) and mechanical energy" |
-| The best reason to buy an ENERGY STAR® appliance is ...
-| ELS04 | 1 = "ENERGY STAR appliances are usually bigger"<br>2 = "ENERGY STAR appliances cost more"<br>3 = "ENERGY STAR appliances use less energy"<br>4 = "ENERGY STAR appliances are more modern looking"<br>5 = "ENERGY STAR appliances cost less" |
-| Which uses the MOST ENERGY in the average American home in one year?
-| ELS05 | 1 = "Refrigerating food and beverages"<br>2 = "Washing and drying clothing"<br>3 = "Heating and cooling rooms"<br>4 = "Heating and cooling water"<br>5 = "Lighting the home" |
-| Which of the following items uses the MOST ELECTRICITY in the average home in one year? | ELS06 | 1 = "Lights"<br>2 = "Refrigerator"<br>3 = "Telephone"<br>4 = "Television"<br>5 = "Computer" |
-| Which of the following sources provides most of the ELECTRICITY in the United States?
-| ELS07 | 1 = "Nuclear power"<br>2 = "Burning petroleum"<br>3 = "Burning coal"<br>4 = "Solar energy"<br>5 = "Water (hydro) power" |
-| Some people think that if we run out of fossil fuels we can just switch over to electric cars.
-What is wrong with this idea?
-| ELS08 | 1 = "Most electricity is currently produced from fossil fuels (coal, oil, natural gas)"<br>2 = "Switching to electric cars will make unemployment rates go up"<br>3 = "It has been proven that it is impossible to build electric cars in great quantities"<br>4 = "You can't use electricity to operate a car"<br>5 = "There is nothing wrong with this idea" |
+1.  **Electrical energy units (kWh):** The amount of ELECTRICAL ENERGY (ELECTRICITY) we use is measured in units called...
+    - [ ] Kilowatt (kW)
+    - [x] Kilowatt-hours (kWh)
+    - [ ] British Thermal Units (BTU)
+    - [ ] Volts (V)
+    - [ ] Horsepower (HP)
+
+2.  **Energy consumed and appliance power rating:** The amount of ENERGY consumed by an electrical appliance is equal to the power rating of the appliance (watts or kilowatts)...
+    - [ ] Multiplied by the cost of electricity
+    - [ ] Added to the cost of electricity
+    - [x] Multiplied by the time it's used
+    - [ ] Divided by the time it's used
+    - [ ] Added to the time it's used
+
+3.  **Incandescent lightbulb conversion:** When you turn on an incandescent light bulb, which of the following energy conversion takes place?
+    - [ ] Electrical energy to radiant energy (light)
+    - [ ] Chemical energy to radiant energy (light)
+    - [x] Electrical energy to radiant energy (light) and thermal energy (heat)
+    - [ ] Chemical energy to radiant energy (light) and thermal energy (heat)
+    - [ ] Electrical energy to radiant energy (light) and mechanical energy
+
+4.  **Reason to buy energy star appliances:** The best reason to buy an ENERGY STAR® appliance is...
+    - [ ] ENERGY STAR appliances are usually bigger
+    - [ ] ENERGY STAR appliances cost more
+    - [x] ENERGY STAR appliances use less energy
+    - [ ] ENERGY STAR appliances are more modern looking
+    - [ ] ENERGY STAR appliances cost less
+
+5.  **Which appliances uses the most energy:** Which uses the MOST ENERGY in the average American home in one year?
+    - [ ] Refrigerating food and beverages
+    - [ ] Washing and drying clothing
+    - [x] Heating and cooling rooms
+    - [ ] Heating and cooling water
+    - [ ] Lighting the home
+
+6. **Which appliance uses the most electricity:** Which of the following items uses the MOST ELECTRICITY in the average home in one year?
+    - [ ] Lights
+    - [x] Refrigerator
+    - [ ] Telephone
+    - [ ] Television
+    - [ ] Computer
+
+7.  **Which source provides most electricity in the US:** Which of the following sources provides most of the ELECTRICITY in the United States?
+    - [ ] Nuclear power
+    - [ ] Burning petroleum
+    - [x] Burning coal
+    - [ ] Solar energy
+    - [ ] Water (hydro) power
+
+8.  **Problem with electric cars:** Some people think that if we run out of fossil fuels we can just switch over to electric cars. What is wrong with this idea?
+    - [x] Most electricity is currently produced from fossil fuels (coal, oil, natural gas)
+    - [ ] Switching to electric cars will make unemployment rates go up
+    - [ ] It has been proven that it is impossible to build electric cars in great quantities
+    - [ ] You can't use electricity to operate a car
+    - [ ] There is nothing wrong with this idea
+
+DeWaters, J.
+E., & Powers, S.
+E.
+(2011).
+Energy literacy of secondary students in New York State (USA): A measure of knowledge, affect, and behavior.
+Energy Policy, 39(3), 1699–1710.
+https://doi.org/10.1016/j.enpol.2010.12.049
+
+# Experiment 1 - Energy Planning Task
+
+### State averages of consumption
+
+| Source | Texas average | Texas Family | California average | California Family | Colorado average | Colorado Family | Mass. average | Mass. Family |
+|---------|---------------|--------------|------------------|------------------|-----------------|-----------------|---------------|--------------|
+| Cooling | 4,249 | | 1,289 | | 498 | | 322 | |
+| Heating | 5,099 | | 5,597 | | 16,411 | | 19,108 | |
+| Water heating | 4,396 | | 4,601 | | 5,832 | | 5,070 | |
+| Refrigerator | 1,318 | | 1,055 | | 1,142 | | 1,025 | |
+| Other | 7,883 | | 6,916 | | 6,652 | | 6,682 | |
+| Total | 22,945 | | 19,458 | | 30,535 | | 32,207 | |
+
+### Example 1 of task given to participants - reference class is USD. - Massachusetts Family
+
+The Davis family wants to reduce its household electricity bill by $1,042 next year.
+Please complete two possible action plans that will help the Davis family achieve this goal.
+Please enter how many kWh should be used next year by each appliance and the total kWh each plan would use.
+Enter only whole numbers.
+Try to provide close estimations.
+You may use a calculator to complete the task.
+
+Note: The Davis family used 14,086 more kWh than the average household in Massachusetts last year.
+
+| Category | Electricity Used Last Year by the Davis Family (kWh) | Average Electricity Used Last Year by Households in Massachusetts (kWh) | Action Plan 1 | Action Plan 2 |
+|----------|---------------------------------------------------|---------------------------------------------------------------------|--------------|--------------|
+| Cooling (Central A/C) | 419 | 322 | | |
+| Heating the Home | 26,751 | 19,108 | | |
+| Water Heating | 10,543 | 5,070 | | |
+| Refrigerator | 1,230 | 1,025 | | |
+| Other (Television, Lighting, Electronics, Washer/Dryer, etc.) | 7,350 | 6,682 | | |
+| Total kWh | 46,293 | 32,207 | | |
+
+### Example 2 - reference class is kWh - Colorado Family
+
+The Wells family wants to reduce its household electricity use by 5,965 kWh next year.
+Please complete two possible action plans that will help the Wells family achieve this goal.
+Please enter how many kWh should be used next year by each appliance and the total kWh each plan would use.
+Enter only whole numbers.
+Try to provide close estimations.
+You may use a calculator to complete the task.
+
+Note: The Wells family used 9,233 more kWh than the average household in Colorado last year.
+
+| Category | Electricity Used Last Year by the Wells Family (kWh) | Average Electricity Used Last Year by Households in Colorado (kWh) | Action Plan 1 | Action Plan 2 |
+|----------|---------------------------------------------------|----------------------------------------------------------------|--------------|--------------|
+| Cooling (Central A/C) | 697 | 498 | | |
+| Heating the Home | 18,052 | 16,411 | | |
+| Water Heating | 11,667 | 5,832 | | |
+| Refrigerator | 1,370 | 1,142 | | |
+| Other (Television, Lighting, Electronics, Washer/Dryer, etc.) | 7,882 | 6,652 | | |
+| Total kWh | 39,768 | 30,535 | | |
+
+### Example 3 - reference class is percentage - Texas Family
+
+The Smith family wants to reduce its household electricity use by 15% next year.
+Please complete two possible action plans that will help the Smith family achieve this goal.
+Please enter how many kWh should be used next year by each appliance category and the total kWh each plan would use.
+Enter only whole numbers.
+Try to provide close estimations.
+You may use a calculator to complete the task.
+
+Note: The Smith family used 6,101 more kWh than the average household in Texas last year.
+
+| Category | Electricity Used Last Year by the Smith Family (kWh) | Average Electricity Used Last Year by Households in Texas (kWh) | Action Plan 1 | Action Plan 2 |
+|----------|---------------------------------------------------|-------------------------------------------------------------|--------------|--------------|
+| Cooling (Central A/C) | 6,573 | 4,249 | | |
+| Heating the Home | 6,118 | 5,099 | | |
+| Water Heating | 5,257 | 4,396 | | |
+| Refrigerator | 2,639 | 1,318 | | |
+| Other (Television, Lighting, Electronics, Washer/Dryer, etc.) | 8,459 | 7,883 | | |
+| Total kWh | 29,046 | 22,945 | | |
+
+### Example 4 - California Family
+
+The Adams Family
+
+| Category | Electricity Used Last Year by the Adams Family (kWh) | Average Electricity Used Last Year by Households in California (kWh) | Action Plan 1 | Action Plan 2 |
+|----------|---------------------------------------------------|------------------------------------------------------------------|--------------|--------------|
+| Cooling (Central A/C) | 2,581 | 1,289 | | |
+| Heating the Home | 6,157 | 5,597 | | |
+| Water Heating | 5,061 | 4,601 | | |
+| Refrigerator | 1,266 | 1,055 | | |
+| Other (Television, Lighting, Electronics, Washer/Dryer, etc.) | 7,608 | 6,916 | | |
+| Total kWh | 22,673 | 19,458 | | |
+
+### Breakdown of the states given to participants
+
+| Source | Texas average | Texas Family | California average | California Family | Colorado average | Colorado Family | Mass. average | Mass. Family |
+|---------|---------------|--------------|------------------|------------------|-----------------|-----------------|---------------|--------------|
+| Cooling | 4,249 | 6,573 | 1,289 | 2,581 | 498 | 697 | 322 | 419 |
+| Heating | 5,099 | 6,118 | 5,597 | 6,157 | 16,411 | 18,052 | 19,108 | 26,751 |
+| Water heating | 4,396 | 5,257 | 4,601 | 5,061 | 5,832 | 11,667 | 5,070 | 10,543 |
+| Refrigerator | 1,318 | 2,639 | 1,055 | 1,266 | 1,142 | 1,370 | 1,025 | 1,230 |
+| Other | 7,883 | 8,459 | 6,916 | 7,608 | 6,652 | 7,982 | 6,682 | 7,350 |
+| Total | 22,945 | 29,046 | 19,458 | 22,673 | 30,535 | 39,768 | 32,207 | 46,293 |
+
+# Experiment 2 - Energy Planning Task
+
+### Task amounts: 15% vs. 10% exact
+
+| State/Reference  | 15% kWh | 10% kWh | 15% USD | 10% USD | 15% Percent | 10% Percent |
+|-----------|-----------|-----------|-----------|-----------|-----------|-----------|
+| Adams / Cali     | 3,401   | 2,267   | \$510   | \$340   | 15%         | 10%         |
+| Smith / Texas    | 4,357   | 2,905   | \$479   | \$320   | 15%         | 10%         |
+| Wells / Colorado | 5,965   | 3,977   | \$656   | \$438   | 15%         | 10%         |
+| Davis / Mass     | 6,944   | 4,629   | \$1,042 | \$694   | 15%         | 10%         |
+
+### Task amounts: 15% vs. 10% rounded
+
+| State/Reference  | 15% kWh | 10% kWh | 15% USD | 10% USD | 15% Percent | 10% Percent |
+|-----------|-----------|-----------|-----------|-----------|-----------|-----------|
+| Adams / Cali     | 3,450   | 2,300   | \$518   | \$345   | 15%         | 10%         |
+| Smith / Texas    | 4,350   | 2,900   | \$479   | \$319   | 15%         | 10%         |
+| Wells / Colorado | 6,000   | 4,000   | \$660   | \$440   | 15%         | 10%         |
+| Davis / Mass     | 6,900   | 4,600   | \$1,035 | \$690   | 15%         | 10%         |
+
+### Adams Family / Cali (\$0.15 per kWh) (diff: 3,542)
+
+| Source        | State average | Exact      | Rounded    |
+|---------------|---------------|------------|------------|
+| Cooling       | 1,289         | 2,581      | 3,000      |
+| Heating       | 5,597         | 6,157      | 6,000      |
+| Water heating | 4,601         | 5,061      | 5,000      |
+| Refrigerator  | 1,055         | 1,266      | 1,000      |
+| Other         | 6,916         | 7,608      | 8,000      |
+| **Total**     | **19,458**    | **22,673** | **23,000** |
+
+### Smith Family / Texas (\$0.11 per kWh) (diff: 6,055)
+
+| Source        | State average | Exact      | Rounded    |
+|---------------|---------------|------------|------------|
+| Cooling       | 4,249         | 6,573      | 7,000      |
+| Heating       | 5,099         | 6,118      | 6,000      |
+| Water heating | 4,396         | 5,257      | 5,000      |
+| Refrigerator  | 1,318         | 2,639      | 3,000      |
+| Other         | 7,883         | 8,459      | 8,000      |
+| **Total**     | **22,945**    | **29,046** | **29,000** |
+
+### Wells Family / Colorado (\$0.11 per kWh) (diff: 9,465)
+
+| Source        | State average | Exact      | Rounded    |
+|---------------|---------------|------------|------------|
+| Cooling       | 498           | 697        | 1,000      |
+| Heating       | 16,411        | 18,052     | 18,000     |
+| Water heating | 5,832         | 11,667     | 12,000     |
+| Refrigerator  | 1,142         | 1,370      | 1,000      |
+| Other         | 6,652         | 7,982      | 8,000      |
+| **Total**     | **30,535**    | **39,768** | **40,000** |
+
+### Davis Family / Massachusetts (\$0.15 per kWh) (diff: 13,793)
+
+| Source        | State average | Exact      | Rounded    |
+|---------------|---------------|------------|------------|
+| Cooling       | 322           | 419        | 0          |
+| Heating       | 19,108        | 26,751     | 27,000     |
+| Water heating | 5,070         | 10,543     | 11,000     |
+| Refrigerator  | 1,025         | 1,230      | 1,000      |
+| Other         | 6,682         | 7,350      | 7,000      |
+| **Total**     | **32,207**    | **46,293** | **46,000** |
 
@@
 
-s1 %>% ungroup() |> 
-  summarise(
-    total_subjects = n_distinct(id),
-    .groups = 'drop'
-  ) %>%
-  left_join(
-    s1 %>%
-      group_by(refClass) %>%
-      summarise(
-        subjects_in_class = n_distinct(id),
-        .groups = 'drop'
-      )
-  ) %>%
-  print()
-
- s1 %>% summarise(
-      total_subjects = n(),
-      reference_class_counts = list(
-        distinct(., id, !!sym(refClass)) %>% count(!!sym(refClass))
-      )
-    )
+# s1 %>% ungroup() |> 
+#   summarise(
+#     total_subjects = n_distinct(id),
+#     .groups = 'drop'
+#   ) %>%
+#   left_join(
+#     s1 %>%
+#       group_by(refClass) %>%
+#       summarise(
+#         subjects_in_class = n_distinct(id),
+#         .groups = 'drop'
+#       )
+#   ) %>%
+#   print()
 
@@
 
+# diagnostics
+```{r}
+
+ordinal_model_s1 <- brm(
+    accuracy_level ~ refClass +calc + (1|id) + (1|state),
+    data = s1_agg,
+    family = cumulative("logit"),
+    cores = 4,
+    iter = 4000,
+    control = list(adapt_delta = 0.98), 
+    prior = c(prior(normal(0, 4), class = "Intercept"), 
+                prior(normal(0, 4), class = "b")), 
+    file = paste0(here::here("data/model_cache",'s1_acc3_add.rds')) 
+)
+
+plot(ordinal_model_s1, condition = "refClass")
+```
+
 ## Study2
@@
 ```{r}
-#| label: tbl-s2-prop
+#| label: tbl-s2-prop1
 #| tbl-cap: "Study 2: The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages).
@@
 ```{r}
-#| label: tbl-s2-prop
+#| label: tbl-s2-prop2
 #| tbl-cap: "Study 2: The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages).
@@
 
+library(marginaleffects)
 marginaleffects_by_pct <- avg_slopes(
@@
     )
-    print(marginaleffects_by_pct)
+
+marginaleffects_by_pct |> kable(escape=FALSE,booktabs=TRUE,align=c("l"))
 
@@
 
-avg_predictions(ordinal_model_s2_logit, newdata = datagrid(id=unique,refClass=unique),by="refClass")
+avg_predictions(ordinal_model_s2_logit, newdata = datagrid(id=unique,refClass=unique),by="refClass") |> kable(escape=FALSE,booktabs=TRUE,align=c("l"))
 
@@
 
-ggplot(s2_agg1, aes(y = pct_goal, x = log_abs_error, color = calc)) +
+ggplot(s2_agg, aes(y = pct_goal, x = log_abs_error, color = calc)) +
   geom_density_ridges(aes(fill = calc), alpha = 0.2, scale = 0.5,
@@
 
-ggplot(data = s2_agg1, aes(x = accuracy_level, fill = pct_goal)) +
+ggplot(data = s2_agg, aes(x = accuracy_level, fill = pct_goal)) +
   geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
@@
 
-ggplot(data = s2_agg1, aes(x = accuracy_level, fill = pct_goal)) +
+ggplot(data = s2_agg, aes(x = accuracy_level, fill = pct_goal)) +
 geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
@@
 
-ggplot(s2_agg1, aes(y = refClass, x = log_abs_error, color = rounded)) +
+ggplot(s2_agg, aes(y = refClass, x = log_abs_error, color = rounded)) +
   geom_density_ridges(aes(fill = rounded), alpha = 0.2, scale = 0.5,
@@
 
-ggplot(data = s2_agg1, aes(x = accuracy_level, fill = rounded)) +
+ggplot(data = s2_agg, aes(x = accuracy_level, fill = rounded)) +
 geom_bar(position = "dodge", aes(y = (..count..)/sum(..count..))) +
@@
 
-s2_agg1 |> group_by(id,refClass,calc) |> 
-    mutate(n_accuracy = n_distinct(accuracy_level)) |> 
-    summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),
-    mean_abs_error=mean(abs_error),mean_log_abs_error=mean(log_abs_error),
-    n_accuracy=first(n_accuracy))
+# s2_agg |> group_by(id,refClass,calc) |> 
+#     mutate(n_accuracy = n_distinct(accuracy_level)) |> 
+#     summarise(mg=sum(matched_goal),n=n(), pct=mg/n,mean_pct_change=mean(pct_change),
+#     mean_abs_error=mean(abs_error),mean_log_abs_error=mean(log_abs_error),
+#     n_accuracy=first(n_accuracy))
 

### 2025-01-12 manuscript/manuscript.qmd
updated methods and appendix.

@@
   - /manuscript.html
-format:
-    wordcount-html: 
-        #output-file: index.html
-        wordcount-banner: true
-        echo: false
-        theme: 
-            - zephyr
-            - ../assets/custom.scss
-        toc: true
-        shift-heading-level-by: 1
-        title-block-banner: "#000D51"
-        include-in-header:
-            - text: |
-                <style type="text/css">
-                h2.hide {
-                    display: none;
-                }
-                </style>
-    html: 
-      echo: true
-      code-fold: show
-      output-file: manuscript-code.html
-    hugo-md:
-        include: false
-        echo: false
-        html-math-method: mathjax
-        output-file: manuscript_hugo.md
-        suppress-bibliography: false
-    gfm:
-        echo: true
-        output-file: manuscript_gfm.md
-        suppress-bibliography: true
-    hikmah-manuscript-pdf:
-        output-file: rep_study.pdf
-        echo: false
-        mainfont: "Linux Libertine O"
-        mainfontoptions:
-        - "Numbers=Proportional"
-        - "Numbers=OldStyle"
-        mathfont: "Libertinus Math"
-        knitr:
-            opts_chunk:
-                dev: "cairo_pdf"
-    #hikmah-manuscript-docx: default
-    docx: default
 format-links:   
@@
     - gfm
-date: 2024-12-07
+date: 2025-01-09
 abstract: |
@@
 We implemented our task and surveys on Qualtrics, and recruited participants through Amazon Mechanical Turk.
-In Experiment 1, 252 participants were initially recruited, but data from 17 participants were corrupted due to experimenter error, leaving a final sample of 235 participants.
+In Experiment 1, 252 participants were initially recruited.
+Data from 17 participants were corrupted due to experimenter error, and six participants were excluded due to deviant performance on the task, resulting in  a final sample of 229 participants (146 males, 92 females, 1 not specified).
+The average age of participants was 34.3 years (SD = 10.2).
 Most participants (76%) reported using a calculator to complete the task.
@@
 The study employed a mixed design with reference class (kWh, percentage, USD) as a between-subjects factor and state/family scenario as a within-subjects factor.
-Each participant completed energy reduction planning tasks for two different states, with state order counterbalanced across participants.
+Each participant completed energy reduction planning tasks for two different states (i.e., family scenarios), with state order counterbalanced across participants.
 The family scenarios featured four households in different climate regions: Texas (Smith family) and California (Adams family) representing warm climates, and Colorado (Wells family) and Massachusetts (Davis family) representing cold climates.
-We obtain average utility use from each state by CITE SOURCE FOR STATE AVGS?
+Average utility use data for each state was obtained from the  CITE SOURCE FOR STATE AVGS?
+Participants also completed an 8-item questionnaire assessing participants' knowledge of energy consumption and conversion [@dewatersEnergyLiteracySecondary2011], and a question indicating whether they used a calculator for the task.
 
@@
 These goals were implementing by allocating usage across five appliance categories: heating, cooling, water heating, refrigerator, and other appliances (e.g., TV, lighting).
-
-For each family scenario, participants were shown a table containing the families utility usage from the prior year, alongside the state averages for each appliance category (both prior year usage and state averages are always shown in kWh).
-For each scenario, participants were asked to create two possible action plans to achieve the target reduction in total household energy usage (see @fig-task).
-Depending on their reference class condition, the target reduction amount presented either in kilowatt-hours (kWh), as percentages of total household usage, or in U.S.
+Participants were informed at the start of the task that they would be presented with tables of detailed energy usage data for each family, and that they would have to create 2 action plans for each of the families.
+Each action plan goal was implemented by allocating usage across five appliance categories: heating, cooling, water heating, refrigerator, and other appliances (e.g., TV, lighting, washer/dryer).For each family scenario, participants were shown a table containing the families utility usage from the prior year, alongside the state averages for each appliance category (see @fig-task).
+For each scenario, participants were asked to create two possible action plans to achieve the target reduction in total household energy usage .
+Depending on their reference class condition, the target reduction amount presented either in kilowatt-hours (kWh), as a percentage of total household usage, or in U.S.
 dollars.
@@
 
-kWh Condition: The target reduction is presented as an absolute number of kilowatt-hours (e.g., "Reduce total energy use by 5965 kWh").
-
-Percentage Condition: The target reduction is presented as a percentage of the family's total previous year's usage (e.g., "Reduce total energy use by 15%").
-
-USD Condition: The target reduction is presented as a monetary value, equivalent to the cost of a 15% reduction in energy use (e.g., "Reduce total energy costs by $656").
-
 ![Example energy planning task trial.
 Participants saw a table with a family's previous year electricity usage (here for the Wells family in Colorado) and were asked to allocate energy usage to meet a 15% reduction goal.
-The format of the reduction goal was manipulated to be either a percentage (15% given as goal reduction), kilowatt hours (5965 kWh given), or USD (\$656)](/assets/images/Smith_10R_Wells_15E.png){#fig-task}
-
-Additional data collected included:
-
--   **Energy Literacy Quiz**: An 8-item questionnaire assessing participants' knowledge of energy consumption and conversion [ @dewatersEnergyLiteracySecondary2011].
--   **Calculator Usage Tracking**: Questions determined whether participants used a calculator, paper/pen, or other methods to complete the tasks.
+The format of the reduction goal was manipulated to be either a percentage (15% given as goal reduction), kilowatt hours (5965 kWh given), or USD (\$656).
+Participants in the USD condition were provided with the conversion rate between kwH and USD.](/assets/images/Smith_10R_Wells_15E.png){#fig-task}
 
@@
 
-The experimental procedures in Experiment 2 are quite similar to those in Experiment 1.
-Experiment 2 employed a 2 (task goal: 10% vs.
-15% reduction) x 2 (last year's usage: exact vs.
-rounded) within-subjects design, with a between-subjects manipulation of the reference class (USD vs.
-Percentage vs.
-kWh).
-We recruited 206 participants from Amazon Mechanical Turk, but data from from 10 participants were corrupted due to experimenter error, leaving a final sample of 196 participants.
+In Experiment 2, We recruited 206 participants from Amazon Mechanical Turk, but data from from 10 participants were corrupted due to experimenter error, and six excluded due to deviant behavior, leaving a final sample of 190 participants (102 male; 88 female.
+Average age = 35.5, SD=9.5)
+
+Experiment 2 employed a mixed experimental design with reference class (USD, Percentage, kWh) as a between-subjects factor,  and two within-subjects factors: task goal (10% vs.
+15% reduction) and the presentation of last year's usage data (exact vs.
+rounded numbers).
+The order of presentation of the goal, rounding, and state conditions was counterbalanced across participants.
+As in Experiment 1, each participant completed the energy reduction planning task for two different family-state scenarios.
+In the “rounded” condition, both the family’s previous year usage and the state averages were rounded to the nearest whole number.
 
@@
 
+-----
+
+Storage/GoogleDrive-tegorman13@gmail.com/My%20Drive/Purdue/Representation_Study/manuscript/output/appendix.html
+
+Supplementary information and materials can be found online at [this website](https://tegorman13.github.io/Representation_Study/docs/manuscript/output/apendix.html)
+
 {{< pagebreak >}}

### 2025-01-12 nb/s1_survey.qmd
updated methods and appendix.

@@
 ```
+
+### Energy Literacy/Knowledge Quiz
+
+1.  **Electrical energy units (kWh):** The amount of ELECTRICAL ENERGY (ELECTRICITY) we use is measured in units called...
+    - [ ] Kilowatt (kW)
+    - [x] Kilowatt-hours (kWh)
+    - [ ] British Thermal Units (BTU)
+    - [ ] Volts (V)
+    - [ ] Horsepower (HP)
+
+2.  **Energy consumed and appliance power rating:** The amount of ENERGY consumed by an electrical appliance is equal to the power rating of the appliance (watts or kilowatts)...
+    - [ ] Multiplied by the cost of electricity
+    - [ ] Added to the cost of electricity
+    - [x] Multiplied by the time it's used
+    - [ ] Divided by the time it's used
+    - [ ] Added to the time it's used
+
+3.  **Incandescent lightbulb conversion:** When you turn on an incandescent light bulb, which of the following energy conversion takes place?
+    - [ ] Electrical energy to radiant energy (light)
+    - [ ] Chemical energy to radiant energy (light)
+    - [x] Electrical energy to radiant energy (light) and thermal energy (heat)
+    - [ ] Chemical energy to radiant energy (light) and thermal energy (heat)
+    - [ ] Electrical energy to radiant energy (light) and mechanical energy
+
+4.  **Reason to buy energy star appliances:** The best reason to buy an ENERGY STAR® appliance is...
+    - [ ] ENERGY STAR appliances are usually bigger
+    - [ ] ENERGY STAR appliances cost more
+    - [x] ENERGY STAR appliances use less energy
+    - [ ] ENERGY STAR appliances are more modern looking
+    - [ ] ENERGY STAR appliances cost less
+
+5.  **Which appliances uses the most energy:** Which uses the MOST ENERGY in the average American home in one year?
+    - [ ] Refrigerating food and beverages
+    - [ ] Washing and drying clothing
+    - [x] Heating and cooling rooms
+    - [ ] Heating and cooling water
+    - [ ] Lighting the home
+
+6. **Which appliance uses the most electricity:** Which of the following items uses the MOST ELECTRICITY in the average home in one year?
+    - [ ] Lights
+    - [x] Refrigerator
+    - [ ] Telephone
+    - [ ] Television
+    - [ ] Computer
+
+7.  **Which source provides most electricity in the US:** Which of the following sources provides most of the ELECTRICITY in the United States?
+    - [ ] Nuclear power
+    - [ ] Burning petroleum
+    - [x] Burning coal
+    - [ ] Solar energy
+    - [ ] Water (hydro) power
+
+8.  **Problem with electric cars:** Some people think that if we run out of fossil fuels we can just switch over to electric cars. What is wrong with this idea?
+    - [x] Most electricity is currently produced from fossil fuels (coal, oil, natural gas)
+    - [ ] Switching to electric cars will make unemployment rates go up
+    - [ ] It has been proven that it is impossible to build electric cars in great quantities
+    - [ ] You can't use electricity to operate a car
+    - [ ] There is nothing wrong with this idea

### 2025-01-12 Scripts/fun_plot.R

@@ -1,6 +1,12 @@
 pacman::p_load(ggplot2,wesanderson,glue)
 
 
+library(sysfonts)
+library(showtext)
+
+font_add_google("Manrope", "Manrope")
+showtext_auto()
+
 theme_blank <- theme(
   panel.grid = element_blank(),
   axis.text = element_blank(),

### 2025-01-12 manuscript/manuscript.qmd
update methods figure; fix some typos

@@
     - gfm
-date: 2025-01-09
+date: 2025-01-12
 abstract: |
@@
 Energy costs often impose a significant burden on low-income households, leading to "energy insecurity," where basic energy needs cannot be met (Bednar & Reames, 2020).
-Frequently measured as the energy burden—the percentage of income spent on energy bills—this burden can be disproportionately high for vulnerable families, necessitating difficult trade-offs with essentials such as food or medicine [@bednarRecognitionResponseEnergy2020; @memmottSociodemographicDisparitiesEnergy2021].
+This burden can be disproportionately high for vulnerable families, necessitating difficult tradeoffs with essentials such as food or medicine [@bednarRecognitionResponseEnergy2020; @memmottSociodemographicDisparitiesEnergy2021].
 Energy insecurity has been linked to health risks and unsafe coping strategies, disproportionately impacting racial and ethnic minorities through higher rates of disconnection [@memmottSociodemographicDisparitiesEnergy2021].
@@
 Of particular relevance are reference class effects, which occur when numerical statements are presented without a clear or intuitive basis for comparison, making it difficult to infer meaningful quantities[@gigerenzerSimpleToolsUnderstanding2003; @reimerNumericCommunicationRisk2015].
-A substantial body of evidence suggests that presenting data in terms of absolute counts or frequencies, as opposed to probabilities or percentages, can promote more accurate comprehension and facilitate.
+A substantial body of evidence suggests that presenting data in terms of absolute counts or frequencies, as opposed to probabilities or percentages, can promote more accurate comprehension and facilitate decision making.
 However, it's important to note that even intuitive formats can pose challenges.
@@
 The choice of units and the format in which information is presented have also been shown to exert a significant influence on decision-making specifically within the context of energy consumption and planning.
-For instance, tables are generally more effective than graphs for conveying specific electricity usage data because they facilitate straightforward point reading  @canfieldPerceptionsElectricityuseCommunications2017.
+For instance, tables are generally more effective than graphs for conveying specific electricity usage data because they facilitate straightforward point reading  [@canfieldPerceptionsElectricityuseCommunications2017].
 Furthermore, the framing of energy costs, such as displaying monthly rather than daily or yearly expenses, can significantly affect consumers' choices [@gillConsumerMatchingCosts2022].
 Consequently, the selection of an appropriate information format is crucial for effectively supporting energy-related decisions.
-Therefore, similar to natural frequencies, kWh provide a direct measure of energy use, a characteristic that could simplify calculations and facilitate comparisons, potentially helping consumers better understand and compare the energy consumption of different appliances or activities.
+Therefore, similar to natural frequencies, kWh provide a direct measure of energy use -  a characteristic that could simplify calculations and facilitate comparisons, potentially helping consumers better understand and compare the energy consumption of different appliances or activities.
 Conversely, prior research has suggested that consumers have a preference for receiving energy feedback in terms of monetary values over scientific units [@karjalainenConsumerPreferencesFeedback2011; @nematiImpactInformationbasedInterventions2020], as well as better long-term appliance selection with information presented in monetary terms [@blaschBoundedlyRationalConsumers2019].
-However, it remains uncertain how these reported benefits might generalize to the more complex domain of household energy planning, where usage patterns can be multifaceted and subject to a variety of contextual influences.
+However, it remains uncertain how these reported benefits might generalize to the more complex task of household energy planning, where usage patterns are multifaceted and subject to a variety of contextual influences.
 
@@
 In the context of energy conservation, this implies that interventions should not only highlight potential reductions (such as a 15% target) but also guide residents in allocating those reductions across specific appliances or behaviors [@attariPublicPerceptionsEnergy2010].
-Additionally, meta-analytic findings suggest that people respond more robustly to household-level feedback that situates their usage within a personalized framework, thereby reducing the cognitive burden of figuring out next steps [@nematiImpactInformationbasedInterventions2020].
+Additionally, meta-analytic findings suggest that people respond more robustly to household-level feedback that situates their usage within a personalized framework, thereby reducing the cognitive burden of determining next steps [@nematiImpactInformationbasedInterventions2020].
 
@@
 The format of the reduction goal was manipulated to be either a percentage (15% given as goal reduction), kilowatt hours (5965 kWh given), or USD (\$656).
-Participants in the USD condition were provided with the conversion rate between kwH and USD.](/assets/images/Smith_10R_Wells_15E.png){#fig-task}
+Participants in the USD condition were provided with the conversion rate between kwH and USD.](/assets/images/wells10e_pub.png){#fig-task}
 
@@
 
-All preprocessing and analyses were carried out in R [@rcoreteamLanguageEnvironmentStatistical2020] and the tidyverse package [@wickhamWelcomeTidyverse2019].
+All preprocessing and analyses were carried out in R [@rcoreteamLanguageEnvironmentStatistical2020] and the Tidyverse package [@wickhamWelcomeTidyverse2019].
 Mixed Bayesian regressions were fit using the brms package [@burknerBrmsPackageBayesian2017], with participants and family scenario (states) set as random effects.
@@
 
-@tbl-s1-agg that participants in the kWh condition met the target goal 38% of the time, compared to 22% for the Percentage condition and 10% for the USD condition.
+@tbl-s1-agg shows that participants in the kWh condition met the target goal 38% of the time, compared to 22% for the Percentage condition and 10% for the USD condition.
 Moreover, the kWh reference class exhibited smaller deviations from the target reduction, suggesting that participants performed more accurately when the goal was framed in kWh rather than when percentages or USD.
@@
 
-In Experiment 2, We recruited 206 participants from Amazon Mechanical Turk, but data from from 10 participants were corrupted due to experimenter error, and six excluded due to deviant behavior, leaving a final sample of 190 participants (102 male; 88 female.
+In Experiment 2, We recruited 206 participants from Amazon Mechanical Turk, but data from 10 participants were corrupted due to experimenter error, and six excluded due to deviant behavior, leaving a final sample of 190 participants (102 male; 88 female.
 Average age = 35.5, SD=9.5)
@@
 
-Storage/GoogleDrive-tegorman13@gmail.com/My%20Drive/Purdue/Representation_Study/manuscript/output/appendix.html
-
-Supplementary information and materials can be found online at [this website](https://tegorman13.github.io/Representation_Study/docs/manuscript/output/apendix.html)
+Supplementary information and materials can be found online at [this website](https://tegorman13.github.io/Representation_Study/manuscript/output/appendix.html)
 

### 2025-01-13 manuscript/manuscript.qmd
Improved discussion; fixed some typos

@@
 
-Energy costs often impose a significant burden on low-income households, leading to "energy insecurity," where basic energy needs cannot be met (Bednar & Reames, 2020).
-This burden can be disproportionately high for vulnerable families, necessitating difficult tradeoffs with essentials such as food or medicine [@bednarRecognitionResponseEnergy2020; @memmottSociodemographicDisparitiesEnergy2021].
-Energy insecurity has been linked to health risks and unsafe coping strategies, disproportionately impacting racial and ethnic minorities through higher rates of disconnection [@memmottSociodemographicDisparitiesEnergy2021].
-While the broader context of climate change, partly driven by residential consumption [@farghaliStrategiesEnergyContext2023], underscores the need for sustainable solutions, the financial strain on vulnerable households remains a pressing concern.
+Energy insecurity has emerged as a critical public health concern, especially among low-income households, who frequently face difficult choices between paying energy bills and meeting other essential needs [@bednarRecognitionResponseEnergy2020; @memmottSociodemographicDisparitiesEnergy2021].
+Such households often experience unsafe coping strategies (e.g., foregoing heating during winter months), which disproportionately affect racial and ethnic minorities and heighten risks of utility disconnection  [@memmottSociodemographicDisparitiesEnergy2021].
+Moreover, residential energy use contributes to climate change, intensifying the urgency for sustainable solutions [@farghaliStrategiesEnergyContext2023].
 Promoting behavior change to reduce energy consumption is crucial.
 However, the success of such interventions may hinge on how effectively energy information is communicated, with format and presentation context significantly influencing understanding and action [@canfieldPerceptionsElectricityuseCommunications2017; @fischerFeedbackHouseholdElectricity2008].
+Consequently, a pressing challenge lies in designing communication strategies that effectively convey energy data and motivate practical conservation decisions
 
@@
   setNames(c("Parameter", "Estimate", "CI_Lower", "CI_Upper", "pd")) |> 
-  mutate(Parameter = stringr::str_remove(Parameter, "b_")) |> kable(escape=FALSE,booktabs=TRUE,align=c("l"), row.names = FALSE)
+  mutate(Parameter = stringr::str_remove(Parameter, "b_")) |> 
+  kable(escape=FALSE,booktabs=TRUE,align=c("l"), row.names = FALSE)
 
@@
 #     )) |> kable(escape=FALSE,booktabs=TRUE,align=c("l"))
-
 ```
@@
 
-As shown in @tbl-s1-reg, the reference class coefficients are positive for both the Percentage (Estimate = 1.3, 95% CI: 0.01 to 2.66, pd = 0.98) and USD (Estimate = 2.8, 95% CI: 1.52 to 4.04, pd = 1.00) conditions, relative to the kWh baseline.
+As shown in @tbl-s1-reg, the reference class coefficients are positive for both the Percentage (Estimate = 1.44, 95% CI: 0.07 to 2.88, pd = 0.98) and USD (Estimate = 3.13, 95% CI: 1.81 to 4.50, pd = 1.00) conditions, relative to the kWh baseline.
 This indicates that, compared to the kWh condition, participants in both the Percentage and USD conditions were more likely to produce plans that fell into higher error categories.
-Moreover, the odds ratios (see Table 3) suggest that the USD condition led to a notably higher likelihood of large errors compared to the kWh baseline (OR = 15.7), while the Percentage condition also demonstrated increased odds (OR = 3.7) but was somewhat less detrimental to accuracy than USD.
+Moreover, the odds ratios (see Table 3) suggest that the USD condition led to a notably higher likelihood of large errors compared to the kWh baseline (OR = 22.9), while the Percentage condition also demonstrated increased odds (OR = 4.2) but was somewhat less detrimental to accuracy than USD.
 These results align with our descriptive findings and further clarify that framing the target reductions in absolute kWh units may facilitate significantly more accurate planning.
@@
 # conditional_effects(s1_els_log_error)
+# Regression Coefficients:
+#           Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
+# Intercept    -1.65      0.23    -2.12    -1.18 1.00      921     1465
+# els          -2.35      0.27    -2.88    -1.81 1.00      753      997
 
@@
 This study examined how different numerical representations of energy reduction goals influence consumers’ ability to create accurate energy conservation plans.
-Across two experiments, the findings consistently demonstrated that presenting reduction goals in absolute units (kWh) significantly enhanced planning accuracy compared to percentage-based or monetary formats.
-Furthermore, participants with higher energy literacy exhibited more precise planning across all conditions, underscoring the critical role of domain-specific knowledge in shaping decision-making outcomes.
+Across two experiments, participants in the kWh conditions exhibited the smallest errors in allocating energy cuts across appliances, a result that consistently held even as scenario details varied.
+These results extend prior observations on the impact of numerical format on comprehension and decision-making [@gigerenzerSimpleToolsUnderstanding2003; @reimerNumericCommunicationRisk2015].
+but they move beyond simpler estimation tasks to show how these benefits persist in a multi-step planning context.
+Notably, individuals with higher energy literacy performed better overall, a finding in line with previous work emphasizing the importance of domain knowledge for effective resource conservation [@attariPublicPerceptionsEnergy2010; @canfieldPerceptionsElectricityuseCommunications2017]
 
-These results align with the broader literature on numerical cognition, which suggests that the format in which information is presented can profoundly affect comprehension and decision-making  [@gigerenzerSimpleToolsUnderstanding2003; @reimerNumericCommunicationRisk2015].
-However, our study extends beyond simple estimations or judgments, demonstrating that the advantages of absolute units persist even in a more complex, multi-step planning task.
-Moreover, the consistent superiority of kWh over both percentages and USD provides novel insights into the specific challenges of energy-related decision-making.
-While prior research has suggested that consumers may prefer monetary feedback [@karjalainenConsumerPreferencesFeedback2011; @nematiImpactInformationbasedInterventions2020] or that monetary framing can improve long-term appliance choices [@blaschBoundedlyRationalConsumers2019], our findings indicate that when it comes to allocating specific usage reductions across appliances, absolute units are most effective.
+While prior research has suggested that consumers sometimes prefer monetary formats[@karjalainenConsumerPreferencesFeedback2011; @nematiImpactInformationbasedInterventions2020] or that monetary framing can improve appliance choices [@blaschBoundedlyRationalConsumers2019], our findings tentatively indicate that absolute energy units can facilitate the finer-grained calculations needed to plan specific usage cuts.
 This may be because kWh provide a more direct and less ambiguous representation of energy quantities, facilitating the necessary calculations for accurate planning.
+Moreover, although percentage-based targets might seem appealing from a goal-setting perspective, the additional step of converting percentages to tangible appliance reductions likely increases the chances for error [@fisher100BiasEffects2022].
+Research on consumer heuristics in energy judgments [@vandenbroekHeuristicsEnergyJudgement2019] suggests that individuals gravitate toward concrete cues, and the directness of absolute units may align well with these heuristics.
+Similarly, the concept of “default units” [@herberzKiloWhatDefault2020] further highlights how presenting energy data in a straightforward, standardized format can guide better consumer decisions.
 Nevertheless, it is unclear whether the observed benefits of absolute units (i.e., kWh) are genuinely attributable to their absolute nature, or if other inherent characteristics of these units might be driving the effects.
-Furthermore, the percentage-based reduction targets, while potentially more salient from a goal-setting perspective, also resulted in poorer planning outcomes compared to kWh.
-Such a finding suggests that reliance on percentages can further complicate calculations by adding unnecessary transformations in the problem-solving process, in line with the "off by 100% bias" found in @fisher100BiasEffects2022, where they show that individuals often misunderstand percentage changes greater than 100%.
-The relative ease in using absolute values to arrive at an accurate plan suggests that these representations of the planning task result in the most accurate planning strategies because the number of steps required to perform the required calculations are simpler.
 
-Moreover, the consistent positive relationship observed between energy literacy and planning accuracy across both experiments underscores the importance of baseline knowledge in effectively navigating energy-related information.
-Individuals with higher levels of energy literacy demonstrated a greater capacity to formulate accurate conservation plans, irrespective of the information format presented.
-This finding aligns with previous research highlighting the role of numeracy and domain-specific knowledge in improving judgments related to energy consumption [@attariPublicPerceptionsEnergy2010].
-It further suggests that interventions aimed at enhancing consumers' fundamental understanding of energy concepts could yield significant benefits in improving the effectiveness of energy conservation efforts.
-
-## Limitations
-
-- Each participant only completed 4 action plans
-- calculator use not controlled.
+A few constraints should be considered.
+First, each participant only completed a limited set of action plans (four total), which may not capture the variability of real-world decision processes.
+Second, although we recorded whether participants used calculators, we could not control how thoroughly they engaged in mathematical computations, nor could we track other external resources they might have consulted.
+Future studies could examine more extensive planning tasks, perhaps over multiple sessions, to see how stable these effects remain over time and repeated feedback cycles [@fischerFeedbackHouseholdElectricity2008].
+In addition, investigating whether certain household types or income levels respond differently to monetary versus absolute energy frames could yield further practical guidance for targeting energy-reduction interventions.
+Finally, incorporating more explicit or automated prompts for appliance-specific tips [@tonkeProvidingProceduralKnowledge2024] could clarify how best to translate high-level goals into tangible actions in diverse contexts.
+By building on these avenues, subsequent work can refine how energy information is formatted and delivered, ultimately enhancing both the precision and feasibility of conservation planning.
 

### 2025-01-15 Scripts/read_data.R
small change

@@ -35,6 +35,7 @@
 # [18] "Study2_3revised.sav"   
   
   
+  
 # Load required libraries
 library(haven)
 library(dplyr)

### 2025-01-25 manuscript/manuscript.qmd
manuscript updates; git cache reset

@@
       icon: file-pdf
-    - docx
+    - text: docx
+      href: rs_manuscript_apa.docx
     - gfm
 date: 2025-01-12
-abstract: |
-    Effective communication of energy consumption information is crucial for promoting residential energy conservation.
+abstract: "Effective communication of energy consumption information is crucial for promoting residential energy conservation.
 This study investigates how different numerical representations of energy reduction goals influence consumers' ability to create accurate conservation plans.
@@
 Furthermore, we found that higher energy literacy was associated with more accurate planning.
-These findings demonstrate that absolute units (kWh) are more effective for communicating energy-saving goals, and highlight the potential value of educational interventions to improve consumer energy literacy.
+These findings demonstrate that absolute units (kWh) are more effective for communicating energy-saving goals, and highlight the potential value of educational interventions to improve consumer energy literacy."
 ---
@@
 
-### Literature Review
-
 Energy insecurity has emerged as a critical public health concern, especially among low-income households, who frequently face difficult choices between paying energy bills and meeting other essential needs [@bednarRecognitionResponseEnergy2020; @memmottSociodemographicDisparitiesEnergy2021].
@@
 Moreover, residential energy use contributes to climate change, intensifying the urgency for sustainable solutions [@farghaliStrategiesEnergyContext2023].
-Promoting behavior change to reduce energy consumption is crucial.
+
+Given the relevance of promoting behavior change to reduce energy consumption, several strands of research have aimed to identify factors that affect energy behaviors and interventions that effectively reduce energy consumption.
+Interventions often focus on X, Y, Z.
+For example, …
+
 However, the success of such interventions may hinge on how effectively energy information is communicated, with format and presentation context significantly influencing understanding and action [@canfieldPerceptionsElectricityuseCommunications2017; @fischerFeedbackHouseholdElectricity2008].
-Consequently, a pressing challenge lies in designing communication strategies that effectively convey energy data and motivate practical conservation decisions
+Consequently, a pressing challenge lies in designing communication strategies that effectively convey energy data and motivate practical conservation decisions.
+We set out to explore if consumers are able to develop plans that would help meet certain energy consumption goals.
+
+### Representation Formats
+
+Receiving feedback about one’s energy use is an important step in identifying potential actions to reduce energy consumption.
+However, it is crucial to understand how consumers use the information that is given with their energy bill and whether consumers are able to translate a energy saving goal into an action plan.
+
+Building on the study of Canfield et al. (2017), we aimed to test xxx.
+
+Although many individuals express a desire to conserve energy, research consistently shows that abstract goals (e.g., “reduce overall usage by 15%”) often fail to translate into effective behavior change unless accompanied by specific, actionable steps [@abrahamseReviewInterventionStudies2005; @nematiImpactInformationbasedInterventions2020].
+For instance, @abrahamseReviewInterventionStudies2005 demonstrated that merely providing general information about energy savings rarely alters consumption patterns unless consumers also receive concrete instructions or tailored feedback.
+
+Similarly, @tonkeProvidingProceduralKnowledge2024 rreported that sending households brief but precise text messages outlining how to reduce water use (e.g., limiting irrigation times, adjusting washing machine settings) yielded meaningful decreases in consumption, underscoring the importance of procedural knowledge—namely, knowing how to operationalize a goal rather than simply why it is desirable.
+In the context of energy conservation, this implies that interventions should not only highlight potential reductions (such as a 15% target) but also guide residents in allocating those reductions across specific appliances or behaviors [@attariPublicPerceptionsEnergy2010].
+Additionally, meta-analytic findings suggest that people respond more robustly to household-level feedback that situates their usage within a personalized framework, thereby reducing the cognitive burden of determining next steps [@nematiImpactInformationbasedInterventions2020].
 
@@
 A substantial body of evidence suggests that presenting data in terms of absolute counts or frequencies, as opposed to probabilities or percentages, can promote more accurate comprehension and facilitate decision making.
-However, it's important to note that even intuitive formats can pose challenges.
+However, it is important to note that even intuitive formats can pose challenges.
 For instance, @weberWhyCanOnly2018 found that individuals often struggle with reasoning tasks presented in natural frequencies because they inadvertently revert to more complex probabilistic thinking.
@@
 
-Although many individuals express a desire to conserve energy, research consistently shows that abstract goals (e.g., “reduce overall usage by 15%”) often fail to translate into effective behavior change unless accompanied by specific, actionable steps [@abrahamseReviewInterventionStudies2005; @nematiImpactInformationbasedInterventions2020].
-For instance, @abrahamseReviewInterventionStudies2005 demonstrated that merely providing general information about energy savings rarely alters consumption patterns unless consumers also receive concrete instructions or tailored feedback.
-Similarly, @tonkeProvidingProceduralKnowledge2024 reported that sending households brief but precise text messages outlining how to reduce water use (e.g., limiting irrigation times, adjusting washing machine settings) yielded meaningful decreases in consumption, underscoring the importance of procedural knowledge—namely, knowing how to operationalize a goal rather than simply why it is desirable.
-In the context of energy conservation, this implies that interventions should not only highlight potential reductions (such as a 15% target) but also guide residents in allocating those reductions across specific appliances or behaviors [@attariPublicPerceptionsEnergy2010].
-Additionally, meta-analytic findings suggest that people respond more robustly to household-level feedback that situates their usage within a personalized framework, thereby reducing the cognitive burden of determining next steps [@nematiImpactInformationbasedInterventions2020].
+### Overview of Current Research
 

### 2025-01-25 nb/study2.qmd
manuscript updates; git cache reset

@@
   html: 
-
     grid:

### 2025-01-25 paper.qmd
Analysis on energy literacy
manuscript updates; git cache reset

@@
 
-  <a class="btn btn-warning btn-sm mb-1" target="_blank" href="/manuscript/output/manuscript.docx">
+  <a class="btn btn-warning btn-sm mb-1" target="_blank" href="/manuscript/output/rs_manuscript_apa.docx">
   <i class="fa-solid fa-file-word"></i>&ensp;docx <small>(download)</small>
@@
 
-<iframe src="/manuscript/output/manuscript.html" frameborder="0" style="overflow:hidden;height:75%;width:95%;border:2px solid #ccc;" height="75%" width="95%" name="manuscript"></iframe>
+<iframe src="/manuscript/output/manuscript.html" frameborder=".1" style="overflow:hidden;height:75%;width:95%;border:2px solid #ccc;" height="75%" width="98%" name="manuscript"></iframe>
 ```

### 2025-01-25 manuscript/manuscript.qmd
updates with apa formatting fixes and a few of torstens edits

@@
       icon: file-plus
-    - format: hikmah-manuscript-pdf
+    - format: apaquarto-pdf
       text: PDF version
@@
     - gfm
-date: 2025-01-12
+date: 2025-01-25
 abstract: "Effective communication of energy consumption information is crucial for promoting residential energy conservation.
@@
 
-The choice of units and the format in which information is presented have also been shown to exert a significant influence on decision-making specifically within the context of energy consumption and planning.
+The choice of units and the format in which information is presented have also been shown to exert a significant influence on decision-making within the context of energy consumption and planning.
 For instance, tables are generally more effective than graphs for conveying specific electricity usage data because they facilitate straightforward point reading  [@canfieldPerceptionsElectricityuseCommunications2017].
@@
 The current study addresses these critical issues by systematically investigating the impact of varying information formats (kWh, percentage, and USD) on the accuracy of energy-planning decisions.
-By manipulating the presentation format of energy information, this research aims to elucidate how different representational formats influence planning accuracy.Based on the literature reviewed, we hypothesize that: 1) Presenting energy reduction goals in absolute units (kWh) will lead to greater planning accuracy compared to percentage-based or monetary formats, as absolute units provide a more direct and less ambiguous representation of energy quantities.
+By manipulating the presentation format of energy information, this research aims to elucidate how different representational formats influence planning accuracy.
+Based on the literature reviewed, we hypothesize that: 1) Presenting energy reduction goals in absolute units (kWh) will lead to greater planning accuracy compared to percentage-based or monetary formats, as absolute units provide a more direct and less ambiguous representation of energy quantities.
 2) Higher energy literacy will be associated with more accurate planning, as individuals with greater energy knowledge may be better equipped to process and utilize the provided information, regardless of format.
@@
 We implemented our task and surveys on Qualtrics, and recruited participants through Amazon Mechanical Turk.
-In Experiment 1, 252 participants were initially recruited.
-Data from 17 participants were corrupted due to experimenter error, and six participants were excluded due to deviant performance on the task, resulting in  a final sample of 229 participants (146 males, 92 females, 1 not specified).
+In Experiment 1, 235 participants were initially recruited.
+Data from six participants were excluded due to deviant performance on the task, resulting in a final sample of 229 participants (146 males, 92 females, 1 not specified).
 The average age of participants was 34.3 years (SD = 10.2).
@@
 Average utility use data for each state was obtained from the  CITE SOURCE FOR STATE AVGS?
-Participants also completed an 8-item questionnaire assessing participants' knowledge of energy consumption and conversion [@dewatersEnergyLiteracySecondary2011], and a question indicating whether they used a calculator for the task.
+Participants also completed an 8-item questionnaire assessing their knowledge of energy consumption and conversion [@dewatersEnergyLiteracySecondary2011], and a question indicating whether they used a calculator for the task.
 
@@
 Participants were provided with energy usage data for two hypothetical families and tasked with creating action plans to meet specified energy reduction goals.
-These goals were implementing by allocating usage across five appliance categories: heating, cooling, water heating, refrigerator, and other appliances (e.g., TV, lighting).
-Participants were informed at the start of the task that they would be presented with tables of detailed energy usage data for each family, and that they would have to create 2 action plans for each of the families.
-Each action plan goal was implemented by allocating usage across five appliance categories: heating, cooling, water heating, refrigerator, and other appliances (e.g., TV, lighting, washer/dryer).For each family scenario, participants were shown a table containing the families utility usage from the prior year, alongside the state averages for each appliance category (see @fig-task).
+Participants were informed at the start of the study that they would be presented with tables of detailed energy usage data for each family, and that they would have to create 2 action plans for each of the families.
+Each action plan goal was implemented by allocating usage across five appliance categories: heating, cooling, water heating, refrigerator, and other appliances (e.g., TV, lighting, washer/dryer).
+For each family scenario, participants were shown a table containing the family’s utility usage from the prior year, alongside the state averages for each appliance category (see @fig-task).
 For each scenario, participants were asked to create two possible action plans to achieve the target reduction in total household energy usage .
-Depending on their reference class condition, the target reduction amount presented either in kilowatt-hours (kWh), as a percentage of total household usage, or in U.S.
+Depending on their reference class condition, the target reduction amount was presented either in kilowatt-hours (kWh), as a percentage of total household usage, or in U.S.
 dollars.
 In all conditions, the target reduction was equivalent to a 15% reduction in total household kWh.
+For example, … .
 
-![Example energy planning task trial.
-Participants saw a table with a family's previous year electricity usage (here for the Wells family in Colorado) and were asked to allocate energy usage to meet a 15% reduction goal.
+![Example energy planning task trial](/assets/images/wells10e_pub.png){#fig-task apa-note="Participants saw a table with a family's previous year electricity usage (here for the Wells family in Colorado) and were asked to allocate energy usage to meet a 15% reduction goal.
 The format of the reduction goal was manipulated to be either a percentage (15% given as goal reduction), kilowatt hours (5965 kWh given), or USD (\$656).
-Participants in the USD condition were provided with the conversion rate between kwH and USD.](/assets/images/wells10e_pub.png){#fig-task}
+Participants in the USD condition were provided with the conversion rate between kwH and USD."}
+
+## Measures
 
 ## Results
-
-### Data Analysis
 
@@
 
-::: {#tbl-s1-agg}
+::: {#tbl-s1-agg apa-note="The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal.
+Percentage meeting goal refers to … ETC"}
 
@@
 
-Study 1: Summary of planning accuracy by reference class.
-The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal
+Study 1: Summary of planning accuracy by reference class
 
@@
 
-@tbl-s1-agg shows that participants in the kWh condition met the target goal 38% of the time, compared to 22% for the Percentage condition and 10% for the USD condition.
-Moreover, the kWh reference class exhibited smaller deviations from the target reduction, suggesting that participants performed more accurately when the goal was framed in kWh rather than when percentages or USD.
+@tbl-s1-agg shows the average x,y,z, separately for the conditions of … .
+As can be seen in Table 1, participants in the kWh condition met the target goal 38% of the time, compared to 22% for the Percentage condition and 10% for the USD condition.
+Moreover, the kWh reference class exhibited smaller deviations from the target reduction, suggesting that participants performed more accurately when the goal was framed in kWh rather than when the goal was framed in percentages or USD.
 
@@
 #| label: fig-s1-log-dist
-#| fig-cap: "Experiment 1: Distribution of the log of the absolute error between the participant's action plan and the reduction goal across different reference class conditions (kWh, Percentage, USD).
+#| fig-cap: "Experiment 1: Planning Error Distributions"
+#| apa-note: "Distribution of the log of the absolute error between the participant's action plan and the reduction goal across different reference class conditions (kWh, Percentage, USD).
 The right side plots are further separated by calculator usage.
@@
 #| label: tbl-s1-prop
-#| tbl-cap: "Study 1: The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages).
+#| tbl-cap: "Study 1: Descriptive Performance"
+#| apa-note: "The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages).
 The combined group column reflects the percentage of participants in each accuracy level when aggregating across across all reference class conditions."
@@
 
-::: {#tbl-s1-reg}
+::: {#tbl-s1-reg apa-note="Ordinal regression results.
+Positive coefficients for the reference class predictors indicate that those conditions are associated with higher error categories relative to the kWh baseline."}
 
@@
 
-**Experiment 1**: Ordinal Regression results.
-Ordinal regression results.
-Positive coefficients for the reference class predictors indicate that those conditions are associated with higher error categories relative to the kWh baseline.
+**Experiment 1**: Ordinal Regression results
 
@@
 
-::: {#tbl-s1-ord}
+::: {#tbl-s1-ord apa-note="Odds ratios greater than 1 indicate increased odds of falling into a worse accuracy category compared to the kWh condition"}
 |Comparison        |odds_ratio |ci_lower |ci_upper |
@@
 
-**Experiment 1**: Odds ratios for group comparisons. Odds ratios greater than 1 indicate increased odds of falling into a worse accuracy category compared to the kWh condition.
+**Experiment 1**: Odds ratios for group comparisons
 
@@
 #| label: fig-s1-ppd
-#| fig-cap: "Experiment 1: Posterior predictive check for frequency of trials at each accuracy level, faceted by reference class (kWh, Percentage, USD).
+#| fig-cap: "Experiment 1: Posterior Predictive Check"
+#| apa-note: "Posterior predictive check for frequency of trials at each accuracy level, faceted by reference class (kWh, Percentage, USD).
 Bars show observed, dots show model predicted proportions.
@@
 #| label: fig-s1-els
-#| fig-cap: "Experiment 1.
-Conditional effect of energy literacy on log absolute error.
-The plot shows the relationship between energy literacy score and log absolute error, controlling for random effects of participant and state.
+#| fig-cap: "Experiment 1. Conditional effect of energy literacy on log absolute error"
+#| apa-note: "The plot shows the relationship between energy literacy score and log absolute error, controlling for random effects of participant and state.
 Higher energy literacy scores are associated with smaller deviations from the target reduction goal, indicating more accurate planning."
+
 s1_els_log_error <- brm(
@@
 
-::: {#tbl-s2-agg}
+::: {#tbl-s2-agg apa-note="The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."}
 
@@
 
-Experiment 2: Summary of planning accuracy by reference class.
-The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal
+Experiment 2: Summary of planning accuracy by reference class
 
@@
 #| label: fig-s2-log-dist
-#| fig-cap: "Experiment 2: Distribution of the log of the absolute error between the participant's action plan and the reduction goal across different reference class conditions (kWh, Percentage, USD).
+#| fig-cap: "Experiment 2: Planning Error Distributions"
+#| apa-note: "Distribution of the log of the absolute error between the participant's action plan and the reduction goal across different reference class conditions (kWh, Percentage, USD).
 The right side plots are further separated by calculator usage.
@@
 ```{r}
-#| label: tbl-s2-ord
+##| label: tbl-s2-ord
 ##| tbl-cap: "Study 2: Ordinal Regression Model Results."
@@
 
-::: {#tbl-s2-reg}
+::: {#tbl-s2-reg apa-note="Ordinal regression results.
+Positive coefficients for the reference class predictors indicate that those conditions are associated with higher error categories relative to the kWh baseline."}
 |Parameter          |Estimate |CI_Lower |CI_Upper |pd   |
@@
 
-**Experiment 2.**  Parameter estimates from the ordinal regression model.
-Positive coefficients for refClass predictors indicate increased likelihood of falling into higher error categories relative to the kWh baseline.
+**Experiment 2.**  Parameter estimates from the ordinal regression model
 
@@
 
-::: {#tbl-s2-ord}
+::: {#tbl-s2-ord apa-note="Odds ratios greater than 1 indicate increased odds of falling into a worse accuracy category compared to the kWh condition"} 
 |comparison           |odds_ratio |ci_lower |ci_upper |
@@
 
-**Experiment 2.** Odds ratios for group comparisons. Odds ratios greater than 1 indicate increased odds of falling into a worse accuracy category compared to the comparison condition.
+**Experiment 2.** Odds ratios for group comparisons
 
@@
 #| label: fig-s2-ame
-#| fig-cap: "Experiment 2.
-Average marginal effects of reference class on accuracy levels (Experiment 2).
-The points represent the average change in the probability of each accuracy level when switching from the kWh reference class to Percentage (red) or USD (green).
+#| fig-cap: "Experiment 2. Average marginal effects of reference class on accuracy levels"
+#| apa-note: "The points represent the average change in the probability of each accuracy level when switching from the kWh reference class to Percentage (red) or USD (green).
 Error bars indicate 95% credible intervals.
@@
 #| label: fig-s2-ppd
-#| fig-cap: "Experiment 2.
-Posterior predictive check of the bayesian regression model, faceted by reference class.
-The bars represent the observed frequencies of each accuracy level within each reference class.
-The points represent the model's predicted proportions, with error bars indicating 95% credible intervals."
+#| fig-cap: "Experiment 2. Posterior predictive check of the bayesian regression model"
+#| apa-note: "The plot shows the observed and predicted frequencies of each accuracy level, faceted by reference class (kWh, Percentage, USD).
+The model provides a reasonable fit to the observed data, with the predicted proportions closely matching the observed frequencies."
 
@@
 #| label: fig-s2-els
-#| fig-cap: "Experiment 2.
-Conditional effect of energy literacy on log absolute error.
-The plot shows the relationship between energy literacy score and log absolute error, controlling for random effects of participant and state.
+#| fig-cap: "Experiment 2. Conditional effect of energy literacy on log absolute error"
+#| apa-note: "The plot shows the relationship between energy literacy score and log absolute error, controlling for random effects of participant and state.
 Higher energy literacy scores are associated with smaller deviations from the target reduction goal, indicating more accurate planning."

### 2025-01-25 paper.qmd
updates with apa formatting fixes and a few of torstens edits

@@
 ---
-title: Representation Project
-subtitle: "Planning to Save Energy: How Information Format Affects Accuracy"
+title: Representation Study
+#subtitle: "Planning to Save Energy: How Information Format Affects Accuracy"
 toc: false
-sidebar: false
+#sidebar: false
 repo-actions: false
@@
 
-<iframe src="/manuscript/output/manuscript.html" frameborder=".1" style="overflow:hidden;height:75%;width:95%;border:2px solid #ccc;" height="75%" width="98%" name="manuscript"></iframe>
+<iframe src="/manuscript/output/manuscript.html" frameborder=".1" style="overflow:hidden;height:75%;width:90%;border:2px solid #ccc;" height="75%" width="98%" name="manuscript"></iframe>
 ```

### 2025-03-17 manuscript/intro.qmd
writing updates from last week I forgot to push
Update after fixing some apa issues; grammatical errors etc.

@@
+---
+title: "Planning to Save Energy: How Information Format Affects Accuracy"
+author:
+  - name: Thomas E. Gorman
+    url: https://tegorman13.github.io/
+    orcid: "0000-0001-5366-5442"
+    affiliations: 
+      - ref: cclab
+    attributes: 
+      corresponding: true
+affiliations:
+  - id: cclab
+    name: Communication and Cognition Lab, Purdue University
+    city: West Lafayette
+    region: IN
+    country: USA
+    postal-code: 47907
+    url: https://web.ics.purdue.edu/~treimer/
+date: 2025-03-16
+abstract: "Effective communication of energy consumption information is crucial for promoting residential energy conservation.
+This study investigates how different numerical representations of energy reduction goals influence consumers' ability to create accurate conservation plans.
+Across two experiments, we examined the impact of presenting energy information in kilowatt-hours (kWh), percentages, or U.S.
+dollars (USD) on planning accuracy.
+Participants completed a simulated household planning task in which they allocated energy usage across multiple appliances, with the goal presented in either kilowatt-hours (kWh), percentages, or monetary costs.
+Results across both experiments showed that presenting reduction goals in absolute units (kWh) led to significantly greater accuracy compared to percentage-based or monetary formats.
+Furthermore, we found that higher energy literacy was associated with more accurate planning.
+These findings demonstrate that absolute units (kWh) are more effective for communicating energy-saving goals, and highlight the potential value of educational interventions to improve consumer energy literacy."
+bibliography: ../assets/rs_bib.bib
+csl: ../assets/apa.csl
+keywords: ["Information format", "Energy literacy", "Decision-making", "Planning"]
+floatsintext: true
+---
+
 # Planning to Save Energy: How Information Format Affects Accuracy
@@
 In all conditions, the target reduction was equivalent to a 15% reduction in total household kWh.
+
+::: {#tbl-s1-ord apa-note="Odds ratios greater than 1 indicate increased odds of falling into a worse accuracy category compared to the kWh condition"}
+|Comparison        |odds_ratio |ci_lower |ci_upper |
+|:-----------------|:----------|:--------|:--------|
+|Percentage vs kWh |4.2        |1.1      |18       |
+|USD vs kWh        |22.9       |6.1      |90       |
+
+**Experiment 1**: Odds ratios for group comparisons
+
+:::
+
+For1 each family scenario, the participants were shown a table containing the families utility usage from the prior year, alongside the state averages for each appliance category (both prior year usage and stage averages are always shown in kWh).
+For each scenario, participants were asked to create two possible action plans to achieve the target reduction in total household energy usage (see fig-task).
+Depending on their reference class condition, the target reduction amount presented either in kilowatt-hours (kWh), as percentages of total household usage, or in U.S.
+dollars.
+In all conditions, the target reduction was equivalent to a 15% reduction in total household kWh.
+
+@gigerenzerSimpleToolsUnderstanding2003
+
+![Example energy planning task trial](/assets/images/wells10e_pub.png){#fig-task apa-note="Participants saw a table with a family's previous year electricity usage (here for the Wells family in Colorado) and were asked to allocate energy usage to meet a 15% reduction goal.
+The format of the reduction goal was manipulated to be either a percentage (15% given as goal reduction), kilowatt hours (5965 kWh given), or USD (\$656).
+Participants in the USD condition were provided with the conversion rate between kwH and USD."}
+
+For2 each family scenario, the participants were shown a table containing the families utility usage from the prior year, alongside the state averages for each appliance category (both prior year usage and stage averages are always shown in kWh).
+For each scenario, participants were asked to create two possible action plans to achieve the target reduction in total household energy usage (see fig-task).
+Depending on their reference class condition, the target reduction amount presented either in kilowatt-hours (kWh), as percentages of total household usage, or in U.S.
+dollars.
+In all conditions, the target reduction was equivalent to a 15% reduction in total household kWh.
+
+::: {#tbl-s1-agg apa-note="The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal.
+Percentage meeting goal refers to … ETC"}
+
+|Reference Class |Avg. % Change |% meeting goal (exact) |% meeting goal (close match) |Abs. Deviation |Log Abs. Deviation |
+|:---------------|:-------------|:----------------------|:----------------------------|:--------------|:------------------|
+|kWh             |0.22          |0.38                   |0.54                         |0.03           |-3.7               |
+|Percentage      |0.21          |0.22                   |0.40                         |0.06           |-3.1               |
+|USD             |0.23          |0.10                   |0.22                         |0.10           |-2.4               |
+
+Study 1: Summary of planning accuracy by reference class
+
+:::
+
+@tbl-s1-agg shows the average x,y,z, separately for the conditions of … .
+As can be seen in Table 1, participants in the kWh condition met the target goal 38% of the time, compared to 22% for the Percentage condition and 10% for the USD condition.
+Moreover, the kWh reference class exhibited smaller deviations from the target reduction, suggesting that participants performed more accurately when the goal was framed in kWh rather than when the goal was framed in percentages or USD.
+
+As shown in @tbl-s1-agg, participants in the kWh condition exactly met the target reduction goal 38% of the time, significantly outperforming those in the Percentage (22%) and USD (10%) conditions.
+Furthermore, the kWh reference class exhibited notably smaller mean absolute deviations (0.03) compared to Percentage (0.06) and USD (0.10), suggesting that presenting the reduction goal in absolute units facilitated more precise allocations.

### 2025-03-17 manuscript/manuscript.qmd
Update after fixing some apa issues; grammatical errors etc.

@@
     - gfm
-date: 2025-01-25
+date: 2025-03-16
 abstract: "Effective communication of energy consumption information is crucial for promoting residential energy conservation.
@@
 Given the relevance of promoting behavior change to reduce energy consumption, several strands of research have aimed to identify factors that affect energy behaviors and interventions that effectively reduce energy consumption.
-Interventions often focus on X, Y, Z.
-For example, …
+For example, @abrahamseEffectTailoredInformation2007 used a combination of tailored information about energy use and tailored feedback, in addition to setting an energy-saving goal, to promote direct and indirect energy-saving behaviors and high levels of knowledge regarding energy use.
+Direct energy-saving behaviors were classified as reducing fuel, gas, and electricity consumption, while indirect behaviors referred to producing, distributing, and disposing goods.
+After a five-month intervention, the authors found that households in the experimental group reduced their direct energy use by 5.1% and had higher levels of knowledge about energy conservation compared to the control group, which used 0.7% more energy since the beginning of the intervention.
+No difference between groups was found in indirect behaviors.
 
@@
 
-Receiving feedback about one’s energy use is an important step in identifying potential actions to reduce energy consumption.
-However, it is crucial to understand how consumers use the information that is given with their energy bill and whether consumers are able to translate a energy saving goal into an action plan.
-
-Building on the study of Canfield et al. (2017), we aimed to test xxx.
+Receiving feedback about one’s energy use is important in identifying potential actions to reduce energy consumption.
+However, it is crucial to understand how consumers use the information provided with their energy bill and whether they are able to translate an energy-saving goal into an action plan.
+@canfieldPerceptionsElectricityuseCommunications2017 specifically studied this problem by running an experiment in which participants were shown hypothetical electricity bills with information related to a household’s historical electricity use, electricity use in relation to their neighbors, and historical electricity use by appliances.
+Participants were randomly assigned to one of three formats of information representation (i.e., tables, bar graphs, and icon graphs) and were asked questions regarding their energy literacy.
+@canfieldPerceptionsElectricityuseCommunications2017 showed that tables were the easiest format to understand for consumers when evaluating every type of information related to energy use.
+Across types of information, historical electricity use elicited the highest intentions and preferences for energy savings regardless of format.
+Additionally, participants with high energy literacy had a better understanding of energy-related information across all types of information representations.
+By disentangling the effects of content, format, and individual differences in energy literacy on understanding, preferences, and intentions, @canfieldPerceptionsElectricityuseCommunications2017 demonstrated that easy-to-implement communication strategies in energy bills can lead to energy-saving behaviors.
 
@@
 The way numerical information is presented can significantly affect how individuals process and use that information.
-Of particular relevance are reference class effects, which occur when numerical statements are presented without a clear or intuitive basis for comparison, making it difficult to infer meaningful quantities[@gigerenzerSimpleToolsUnderstanding2003; @reimerNumericCommunicationRisk2015].
+Of particular relevance are reference class effects, which occur when numerical statements are presented without a clear or intuitive basis for comparison, making it difficult to infer meaningful quantities [@gigerenzerSimpleToolsUnderstanding2003; @reimerNumericCommunicationRisk2015].
 A substantial body of evidence suggests that presenting data in terms of absolute counts or frequencies, as opposed to probabilities or percentages, can promote more accurate comprehension and facilitate decision making.
@@
 
-Despite existing studies on energy-use communication and format effects, limited research has explored how different numerical representations influence consumers' ability to create accurate energy conservation plans.
+Building on the study of Canfield et al. (2017), we aimed to test xxx.
+
+Despite existing studies on energy-use communication and format effects, limited research has explored how different numerical representations influence consumers’ ability to create accurate energy conservation plans.
 Specifically, there is a gap in understanding how presenting energy information in absolute units versus percentages or monetary terms affects the precision of planning appliance-specific reductions.
@@
 #| fig-cap: "Experiment 1: Planning Error Distributions"
-#| apa-note: "Distribution of the log of the absolute error between the participant's action plan and the reduction goal across different reference class conditions (kWh, Percentage, USD).
-The right side plots are further separated by calculator usage.
+#| apa-note: "Distribution of the log of the absolute error between the participant's action plan and the reduction goal.
+Left side plot separates across different reference class conditions (kWh, Percentage, USD).
+Right side plot separates by self-reported calculator use.
 A lower log absolute error suggests higher planning accuracy."
@@
 #| label: tbl-s1-prop
-#| tbl-cap: "Study 1: Descriptive Performance"
+#| tbl-cap: "Experiment 1 1: Categorization of Participants according to Accuracy Levels"
 #| apa-note: "The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages).
@@
 Specifically, we used a cumulative logit link function to model the ordered accuracy outcome, and we specified weakly informative priors for the regression coefficients (normal distributions with mean 0 and standard deviation of 1) and for the cutpoints (normal distributions with a mean of zero and a standard deviation of 4.0).
-The approach allows us to estimate threshold parameters and regression coefficients that characterize how changes in predictor variables (such as the reference class: kWh, percentage, or USD) relate to probabilities of being in each accuracy category.
+The approach can be used to estimate threshold parameters and regression coefficients that characterize how changes in predictor variables (such as the reference class: kWh, percentage, or USD) relate to probabilities of being in each accuracy category.
 
@@
 
-**Experiment 1**: Ordinal Regression results
+**Experiment 1**: Ordinal Regression Results of a Test of Differences between conditions in Accuracy Levels
 
@@
 
-::: {#tbl-s2-agg apa-note="The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal."}
+::: {#tbl-s2-agg apa-note="The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal.
+Percentage meeting goal refers to participants who exactly met the target reduction goal (i.e., the proposed plan would reduce energy use by exactly 15%)"}
 
@@
 #| fig-cap: "Experiment 2: Planning Error Distributions"
-#| apa-note: "Distribution of the log of the absolute error between the participant's action plan and the reduction goal across different reference class conditions (kWh, Percentage, USD).
-The right side plots are further separated by calculator usage.
+#| apa-note: "Distribution of the log of the absolute error between the participant's action plan and the reduction goal.
+Left side plot separates across different reference class conditions (kWh, Percentage, USD).
+Right side plot separates by self-reported calculator use.
 A lower log absolute error suggests higher planning accuracy."

### 2025-03-17 Scripts/fun_plot.R

@@ -82,23 +82,24 @@ theme_nice_pdf <- function() {
     )
 }
 
+# Define a custom theme function for ggplot2 with specific font and styling
 theme_nice_b <- function() {
-  theme_minimal(base_family = "Manrope") +
+  theme_minimal(base_family = "Manrope") +  # Use the "Manrope" font family as the base
     theme(
-      panel.grid.minor = element_blank(),
-      plot.title = element_text(family = "Manrope Extrabold", face = "plain", size = rel(1.55)),
-      plot.subtitle = element_text(family = "Manrope Medium", face = "plain", size = rel(1.2)),
-      axis.title = element_text(family = "Manrope SemiBold", face = "plain", size = rel(1)),
-      axis.title.x = element_text(hjust = .5),
-      axis.title.y = element_text(hjust = .5),
-      axis.text = element_text(face = "plain", size = rel(0.9)),
+      panel.grid.minor = element_blank(),  # Remove minor grid lines
+      plot.title = element_text(family = "Manrope Extrabold", face = "plain", size = rel(1.55)),  # Customize plot title
+      plot.subtitle = element_text(family = "Manrope Medium", face = "plain", size = rel(1.2)),  # Customize plot subtitle
+      axis.title = element_text(family = "Manrope SemiBold", face = "plain", size = rel(1)),  # Customize axis titles
+      axis.title.x = element_text(hjust = .5),  # Center the x-axis title
+      axis.title.y = element_text(hjust = .5),  # Center the y-axis title
+      axis.text = element_text(face = "plain", size = rel(0.9)),  # Customize axis text
       strip.text = element_text(
         family = "Manrope", face = "bold",
-        size = rel(1.2), hjust = 0
+        size = rel(1.2), hjust = 0  # Customize facet strip text
       ),
-      legend.position = "top",
-      legend.text = element_text( face = "plain", size = rel(1.3)), #family = "Manrope Light",
-      strip.background = element_rect(fill = "grey90", color = NA)
+      legend.position = "top",  # Position the legend at the top
+      legend.text = element_text(face = "plain", size = rel(1.3)),  # Customize legend text
+      strip.background = element_rect(fill = "grey90", color = NA)  # Customize facet strip background
     )
 }
 

### 2025-03-31 manuscript/manuscript.qmd
small updates with edits

@@
 In all conditions, the target reduction was equivalent to a 15% reduction in total household kWh.
-For example, … .
+For example, participants assigned to the Wells family scenario (Colorado), as depicted in Figure 1, were asked to achieve a reduction of 5,965 kWh in the kWh condition, 15% in the Percentage condition, or $656 in the USD condition, all representing the same underlying 15% energy reduction target for that specific household.
 
-![Example energy planning task trial](/assets/images/wells10e_pub.png){#fig-task apa-note="Participants saw a table with a family's previous year electricity usage (here for the Wells family in Colorado) and were asked to allocate energy usage to meet a 15% reduction goal.
+![Example energy planning task trial](/assets/images/wells10e_pub.png){#fig-task width="80%" apa-note="Participants saw a table with a family's previous year electricity usage (here for the Wells family in Colorado) and were asked to allocate energy usage to meet a 15% reduction goal.
 The format of the reduction goal was manipulated to be either a percentage (15% given as goal reduction), kilowatt hours (5965 kWh given), or USD (\$656).
@@
 # overall pct of subjects who matched their goal
-s1_agg4 |> group_by('Reference Class' = refClass) |>
-    summarise(
-    'Avg. % Change' = mean(mean_pct_change),
-    '% meeting goal (exact)' = mean(pct),
-    '% meeting goal (close match)' = mean(pct_close),
-    'Abs. Deviation' = median(mean_abs_error),
-    'Log Abs. Deviation' = (median(mean_log_abs_error)),
-    # sd = sd(pct),
-    # n = n(),
-    #se=sd(pct)/sqrt(n)
-) |>   mutate(across(where(is.numeric), \(x) round(x, 3))) %>% 
-  kable(escape=FALSE,booktabs=TRUE,align=c("l"))
+# s1_agg4 |> group_by('Reference Class' = refClass) |>
+#     summarise(
+#     'Avg. % Change' = mean(mean_pct_change),
+#     '% meeting goal (exact)' = mean(pct),
+#     '% meeting goal (close match)' = mean(pct_close),
+#     'Abs. Deviation' = median(mean_abs_error),
+#     'Log Abs. Deviation' = (median(mean_log_abs_error)),
+#     # sd = sd(pct),
+#     # n = n(),
+#     #se=sd(pct)/sqrt(n)
+# ) |>   mutate(across(where(is.numeric), \(x) round(x, 3))) %>% 
+#   kable(escape=FALSE,booktabs=TRUE,align=c("l"))
 
@@
 
+summary_table <- s1_agg4 |> 
+  group_by(`Reference Class` = refClass) |> 
+  summarise(
+    N = n(),
+    `Avg. % Change` = sprintf("%.2f (%.2f)", mean(mean_pct_change), sd(mean_pct_change)),
+    `% meeting goal (exact)` = sprintf("%.2f (%.2f)", mean(pct), sd(pct)),
+    `% meeting goal (close match)` = sprintf("%.2f (%.2f)", mean(pct_close), sd(pct_close)),
+    `Abs. Deviation` = sprintf("%.2f (%.2f)", mean(mean_abs_error), sd(mean_abs_error)),
+    `Log Abs. Deviation` = sprintf("%.2f (%.2f)", mean(mean_log_abs_error), sd(mean_log_abs_error))
+  ) |> 
+  kable(escape = FALSE, booktabs = TRUE, align = c("l"))
+
+#summary_table
+
 ```
 
-::: {#tbl-s1-agg apa-note="The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal.
-Percentage meeting goal refers to … ETC"}
+::: {#tbl-s1-agg apa-note="Summary statistics by reference class.
+Values represent mean (SD) across participants for each planning accuracy metric.
+Values in parentheses represent standard deviations.
+'Close match' indicates trials where the proposed reduction was within ±2% of the target goal.
+Abs.
+Deviation and Log Abs.
+Deviation columns report medians."}
 
-|Reference Class |Avg. % Change |% meeting goal (exact) |% meeting goal (close match) |Abs. Deviation |Log Abs. Deviation |
-|:---------------|:-------------|:----------------------|:----------------------------|:--------------|:------------------|
-|kWh             |0.22          |0.38                   |0.54                         |0.03           |-3.7               |
-|Percentage      |0.21          |0.22                   |0.40                         |0.06           |-3.1               |
-|USD             |0.23          |0.10                   |0.22                         |0.10           |-2.4               |
+|Reference Class |N  |Avg. % Change |% meeting goal (exact) |% meeting goal (close match) |Abs. Deviation |Log Abs. Deviation |
+|:---------------|:--|:-------------|:----------------------|:----------------------------|:--------------|:------------------|
+|kWh             |76 |0.22 (0.14)   |0.38 (0.45)            |0.54 (0.45)                  |0.09 (0.14)    |-3.47 (1.43)       |
+|Percentage      |67 |0.21 (0.11)   |0.22 (0.37)            |0.40 (0.41)                  |0.09 (0.10)    |-3.18 (1.20)       |
+|USD             |86 |0.23 (0.13)   |0.10 (0.28)            |0.22 (0.33)                  |0.12 (0.11)    |-2.61 (1.06)       |
 
@@
 
-@tbl-s1-agg shows the average x,y,z, separately for the conditions of … .
-As can be seen in Table 1, participants in the kWh condition met the target goal 38% of the time, compared to 22% for the Percentage condition and 10% for the USD condition.
-Moreover, the kWh reference class exhibited smaller deviations from the target reduction, suggesting that participants performed more accurately when the goal was framed in kWh rather than when the goal was framed in percentages or USD.
-
-As shown in @tbl-s1-agg, participants in the kWh condition exactly met the target reduction goal 38% of the time, significantly outperforming those in the Percentage (22%) and USD (10%) conditions.
-Furthermore, the kWh reference class exhibited notably smaller mean absolute deviations (0.03) compared to Percentage (0.06) and USD (0.10), suggesting that presenting the reduction goal in absolute units facilitated more precise allocations.
+@tbl-s1-agg presents a summary of descriptive statistics regarding planning accuracy across the three experimental reference class conditions (kWh, Percentage, USD).
+Observation of the central tendencies suggests a potential advantage for the kWh condition.
+The kWh group exhibited the highest mean proportion of trials exactly matching the target goal (M = 0.38, SD = 0.45), considerably higher than that observed in the Percentage (M = 0.22, SD = 0.37) and USD (M = 0.10, SD = 0.28) conditions.
+A concordant pattern emerged for deviations, where the kWh condition displayed the lowest median absolute deviation (Mdn = 0.03) compared to the Percentage (Mdn = 0.06) and USD (Mdn = 0.10) conditions, hinting at greater precision when goals were presented in absolute energy units.
 
@@
 ```
+
+{{< pagebreak >}}
 

### 2025-03-31 manuscript/convert_docx_to_md.py

@@ -1,0 +1,152 @@
+from docx import Document
+from docx.oxml.ns import qn
+from docx.document import Document as _Document
+from docx.table import Table, _Cell
+from docx.text.paragraph import Paragraph
+import zipfile
+import xml.etree.ElementTree as ET
+import re
+
+def iter_block_items(parent):
+    """
+    Yield each paragraph and table child within parent, in document order.
+    Handles Documents, Tables, Paragraphs, and _Cells as parents.
+    """
+    if isinstance(parent, _Document):
+        parent_element = parent.element.body
+    elif isinstance(parent, Table):
+        parent_element = parent._element
+    elif isinstance(parent, Paragraph):
+        parent_element = parent._element
+    elif isinstance(parent, _Cell):
+        parent_element = parent._element
+    else:
+        raise ValueError(f"Unsupported parent type: {type(parent)}")
+
+    for child in parent_element.iterchildren():
+        if child.tag == qn('w:p'):
+            yield Paragraph(child, parent)
+        elif child.tag == qn('w:tbl'):
+            yield Table(child, parent)
+
+
+def extract_comments(doc_path):
+    """
+    Extracts comments from the DOCX file.
+    Returns a dictionary: {comment_id: (footnote_number, comment_text)}.
+    """
+    comment_map = {}
+    footnote_counter = 1
+    with zipfile.ZipFile(doc_path) as z:
+        if 'word/comments.xml' in z.namelist():
+            xml_content = z.read('word/comments.xml')
+            tree = ET.XML(xml_content)
+            ns = {'w': 'http://schemas.openxmlformats.org/wordprocessingml/2006/main'}
+            for comment in tree.findall('.//w:comment', ns):
+                comment_id = comment.attrib.get('{http://schemas.openxmlformats.org/wordprocessingml/2006/main}id')
+                texts = comment.findall('.//w:t', ns)
+                comment_text = ''.join([t.text for t in texts if t.text]).strip()
+                if comment_text:
+                    comment_map[comment_id] = (footnote_counter, comment_text)
+                    footnote_counter += 1
+    return comment_map
+
+
+def process_paragraph(para, comment_map, ns):
+    """
+    Processes a paragraph and inserts footnote markers for comments.
+    """
+    md = ""
+    for run in para.runs:
+        run_text = run.text
+        comment_refs = run._r.findall('.//w:commentReference', ns)
+        for cref in comment_refs:
+            comment_id = cref.get(qn('w:id'))
[…]

### 2025-03-31 manuscript/manuscript.qmd
major edits for natcom

@@
     - gfm
-date: 2025-03-16
+date: 2025-03-31
+keywords: 
+  - energy literacy
+  - information format
+  - planning accuracy
+  - decision making
 abstract: "Effective communication of energy consumption information is crucial for promoting residential energy conservation.
@@
 Additionally, participants with high energy literacy had a better understanding of energy-related information across all types of information representations.
+Studies such as @dewatersEnergyLiteracySecondary2011 have highlighted that while individuals may express concern for energy issues, their actual knowledge and behaviors may not align, underscoring the importance of fostering a comprehensive understanding of energy concepts to facilitate effective conservation actions.
 By disentangling the effects of content, format, and individual differences in energy literacy on understanding, preferences, and intentions, @canfieldPerceptionsElectricityuseCommunications2017 demonstrated that easy-to-implement communication strategies in energy bills can lead to energy-saving behaviors.
@@
 
-Similarly, @tonkeProvidingProceduralKnowledge2024 rreported that sending households brief but precise text messages outlining how to reduce water use (e.g., limiting irrigation times, adjusting washing machine settings) yielded meaningful decreases in consumption, underscoring the importance of procedural knowledge—namely, knowing how to operationalize a goal rather than simply why it is desirable.
+Similarly, @tonkeProvidingProceduralKnowledge2024 reported that sending households brief but precise text messages outlining how to reduce water use (e.g., limiting irrigation times, adjusting washing machine settings) yielded meaningful decreases in consumption, underscoring the importance of procedural knowledge—namely, knowing how to operationalize a goal rather than simply why it is desirable.
 In the context of energy conservation, this implies that interventions should not only highlight potential reductions (such as a 15% target) but also guide residents in allocating those reductions across specific appliances or behaviors [@attariPublicPerceptionsEnergy2010].
@@
 
-Building on the study of Canfield et al. (2017), we aimed to test xxx.
-
 Despite existing studies on energy-use communication and format effects, limited research has explored how different numerical representations influence consumers’ ability to create accurate energy conservation plans.
@@
 In Experiment 1, 235 participants were initially recruited.
-Data from six participants were excluded due to deviant performance on the task, resulting in a final sample of 229 participants (146 males, 92 females, 1 not specified).
+Data from six participants were excluded due to unusually large error magnitudes (error magnitudes exceeding 2.5 standard deviations from the group mean), resulting in a final sample of 229 participants (136 males, 92 females, 1 participant not specified).
 The average age of participants was 34.3 years (SD = 10.2).
@@
 The family scenarios featured four households in different climate regions: Texas (Smith family) and California (Adams family) representing warm climates, and Colorado (Wells family) and Massachusetts (Davis family) representing cold climates.
-Average utility use data for each state was obtained from the  CITE SOURCE FOR STATE AVGS?
 Participants also completed an 8-item questionnaire assessing their knowledge of energy consumption and conversion [@dewatersEnergyLiteracySecondary2011], and a question indicating whether they used a calculator for the task.
@@
 ![Example energy planning task trial](/assets/images/wells10e_pub.png){#fig-task width="80%" apa-note="Participants saw a table with a family's previous year electricity usage (here for the Wells family in Colorado) and were asked to allocate energy usage to meet a 15% reduction goal.
-The format of the reduction goal was manipulated to be either a percentage (15% given as goal reduction), kilowatt hours (5965 kWh given), or USD (\$656).
-Participants in the USD condition were provided with the conversion rate between kwH and USD."}
-
-## Measures
+The format of the reduction goal was manipulated to be either a percentage (15% reduction), kilowatt hours (5965 kWh), or USD (\$656).
+Participants in the USD condition were provided with the conversion rate between kWh and USD."}
 
@@
 
-All preprocessing and analyses were carried out in R [@rcoreteamLanguageEnvironmentStatistical2020] and the Tidyverse package [@wickhamWelcomeTidyverse2019].
-Mixed Bayesian regressions were fit using the brms package [@burknerBrmsPackageBayesian2017], with participants and family scenario (states) set as random effects.
+All preprocessing and analyses were carried out in `R` version 4.4.3 [@rcoreteamLanguageEnvironmentStatistical2020] and the Tidyverse package [@wickhamWelcomeTidyverse2019].
+Mixed Bayesian regressions were fit using the `brms` package [@burknerBrmsPackageBayesian2017], with participants and family scenario set as random effects.
 
@@
 ```{r}
-
 s1_agg <- s1 |> 
@@
         ))
-
 ```
@@
 
-# overall pct of subjects who matched their goal
-# s1_agg4 |> group_by('Reference Class' = refClass) |>
-#     summarise(
-#     'Avg. % Change' = mean(mean_pct_change),
-#     '% meeting goal (exact)' = mean(pct),
-#     '% meeting goal (close match)' = mean(pct_close),
-#     'Abs. Deviation' = median(mean_abs_error),
-#     'Log Abs. Deviation' = (median(mean_log_abs_error)),
-#     # sd = sd(pct),
-#     # n = n(),
-#     #se=sd(pct)/sqrt(n)
-# ) |>   mutate(across(where(is.numeric), \(x) round(x, 3))) %>% 
-#   kable(escape=FALSE,booktabs=TRUE,align=c("l"))
-
-#pander::pandoc.table(caption="Study 1: Proportion of participants who matched their goal overall")
-
 summary_table <- s1_agg4 |>
@@
         N = n(),
-    `Avg. % Change` = sprintf("%.2f (%.2f)", mean(mean_pct_change), sd(mean_pct_change)),
-    `% meeting goal (exact)` = sprintf("%.2f (%.2f)", mean(pct), sd(pct)),
-    `% meeting goal (close match)` = sprintf("%.2f (%.2f)", mean(pct_close), sd(pct_close)),
-    `Abs. Deviation` = sprintf("%.2f (%.2f)", mean(mean_abs_error), sd(mean_abs_error)),
-    `Log Abs. Deviation` = sprintf("%.2f (%.2f)", mean(mean_log_abs_error), sd(mean_log_abs_error))
+        # Keep Mean (SD) for average change proportion
+        #`Avg. % Change` = sprintf("%.2f (%.2f)", mean(mean_pct_change), sd(mean_pct_change)),
+        # Calculate Mean (SD) for percentages, multiply by 100, format as whole number
+        `Exact Match Rate(%)` = sprintf("%.0f (%.0f)", mean(pct) * 100, sd(pct) * 100),
+        `Close Match Rate(%)` = sprintf("%.0f (%.0f)", mean(pct_close) * 100, sd(pct_close) * 100),
+        # Calculate Median [IQR] for deviations, adjust decimals as needed
+        `Abs. Deviation` = sprintf("%.3f (%.3f)", median(mean_abs_error), IQR(mean_abs_error)),
+        `Log Abs. Deviation` = sprintf("%.2f (%.2f)", median(mean_log_abs_error), IQR(mean_log_abs_error))
     ) |>
-  kable(escape = FALSE, booktabs = TRUE, align = c("l"))
+    # Apply kable formatting (optional, depends on your workflow)
+    knitr::kable(escape = FALSE, booktabs = TRUE, align = c("l"))
 
@@
 
-::: {#tbl-s1-agg apa-note="Summary statistics by reference class.
-Values represent mean (SD) across participants for each planning accuracy metric.
-Values in parentheses represent standard deviations.
-'Close match' indicates trials where the proposed reduction was within ±2% of the target goal.
+::: {#tbl-s1-agg apa-note="Summary statistics by reference class condition, based on participant-level averages.
+Exact Match Rate and Close Match Rate represent the mean percentage of trials (per participant) where the proposed reduction exactly matched or closely matched (±2% tolerance) the target goal (standard deviations in parentheses), respectively.
 Abs.
 Deviation and Log Abs.
-Deviation columns report medians."}
+Deviation reflect the median (IQR) of each participant’s absolute deviation from the target goal."}
 
-|Reference Class |N  |Avg. % Change |% meeting goal (exact) |% meeting goal (close match) |Abs. Deviation |Log Abs. Deviation |
-|:---------------|:--|:-------------|:----------------------|:----------------------------|:--------------|:------------------|
-|kWh             |76 |0.22 (0.14)   |0.38 (0.45)            |0.54 (0.45)                  |0.09 (0.14)    |-3.47 (1.43)       |
-|Percentage      |67 |0.21 (0.11)   |0.22 (0.37)            |0.40 (0.41)                  |0.09 (0.10)    |-3.18 (1.20)       |
-|USD             |86 |0.23 (0.13)   |0.10 (0.28)            |0.22 (0.33)                  |0.12 (0.11)    |-2.61 (1.06)       |
+|Reference Class |N  |Exact Match Rate(%) |Close Match Rate(%) |Abs. Deviation |Log Abs. Deviation |
+|:---------------|:--|:-------------------|:-------------------|:--------------|:------------------|
+|kWh             |76 |38 (45)             |54 (45)             |0.029 (0.117)  |-3.69 (2.79)       |
+|Percentage      |67 |22 (37)             |40 (41)             |0.056 (0.104)  |-3.10 (1.87)       |
+|USD             |86 |10 (28)             |22 (33)             |0.095 (0.094)  |-2.40 (1.09)       |
 
@@
 
-@tbl-s1-agg presents a summary of descriptive statistics regarding planning accuracy across the three experimental reference class conditions (kWh, Percentage, USD).
-Observation of the central tendencies suggests a potential advantage for the kWh condition.
-The kWh group exhibited the highest mean proportion of trials exactly matching the target goal (M = 0.38, SD = 0.45), considerably higher than that observed in the Percentage (M = 0.22, SD = 0.37) and USD (M = 0.10, SD = 0.28) conditions.
-A concordant pattern emerged for deviations, where the kWh condition displayed the lowest median absolute deviation (Mdn = 0.03) compared to the Percentage (Mdn = 0.06) and USD (Mdn = 0.10) conditions, hinting at greater precision when goals were presented in absolute energy units.
+The summary statistics in @tbl-s1-agg suggest descriptive differences in planning accuracy across reference class conditions, with the kWh condition yielding the highest rates of exact matches, and lowest absolute deviations from the target reduction goal.
+However, use of these performance metrics to compare conditions is complicated by the non-normality of their distributions.
+Specifically, the substantial standard deviations associated with the match rates in @tbl-s1-agg, coupled with the visual evidence presented in @fig-s1-log-dist, underscore the markedly skewed, and potentially bi-modal distribution of the underlying planning errors.
+
+In light of these distributional characteristics, we elected to bin planning performance into ordered accuracy levels (as shown in @tbl-s1-prop).
+Consequently, we binned individual trial outcomes into three distinct, ordered categories: "Exact match" (0% absolute error), "0.01-5% error" (minor deviations), and "Over 5% error" (larger deviations).
+This binned `accuracy_level` variable subsequently served as the dependent measure in our primary inferential analyses examining the impact of reference class on planning accuracy
 
@@
 #| label: tbl-s1-prop
-#| tbl-cap: "Experiment 1 1: Categorization of Participants according to Accuracy Levels"
+#| tbl-cap: "Experiment 1: Categorization of Participants according to Accuracy Levels"
 #| apa-note: "The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages).
-The combined group column reflects the percentage of participants in each accuracy level when aggregating across across all reference class conditions."
+"
+
+# The combined group column reflects the percentage of participants in each accuracy level when aggregating across across all reference class conditions.
 
@@
   mutate(Total = paste0(round(ns/n_total*100,1), "%")) |>
-  select('Reference Class'=refClass, 'Accuracy Level'=accuracy_level, '% in Group'=pct_grp, "Combined Groups %" =Total) |>
+  select('Reference Class'=refClass, 'Accuracy Level'=accuracy_level, '% in Group'=pct_grp) |> #, "Combined Groups %" =Total) |>
   pivot_wider(
@@
     values_from = c('% in Group')
-  ) |> relocate("Combined Groups %" , .after=last_col()) |> 
+  ) |> #relocate("Combined Groups %" , .after=last_col()) |> 
   kable(escape=FALSE,booktabs=TRUE,align=c("l"))
@@
 
-We next categorized responses into three accuracy levels (exact match [0% error], minor deviations [0.01–5%], and large deviations [>5%]) for our primary statistical modeling.
-Using Bayesian ordinal regression, we modeled the ordered accuracy outcome as a function of the reference class condition, while controlling for random variation across participants and family scenarios:
+Group comparisons were conducted with Bayesian ordinal regression [@burknerOrdinalRegressionModels2019; @kruschkeDoingBayesianData2014].
+We modeled the ordered accuracy outcome as a function of the reference class condition, while controlling for random variation across participants and family scenarios (@eq-e1).
 
@@
 \text{Accuracy Level} \sim \text{Reference Class} + \text{Calculator} + (1|\text{id}) + (1|\text{Family Scenario})
-$$
+$${#eq-e1}
 
-This approach allowed us to estimate thresholds (intercepts) and regression coefficients that capture how different reference classes affect the likelihood of achieving higher accuracy categories.
-For each comparison, we provide posterior odds ratios (OR) and their 95% CIs.
 This approach allows the estimation of threshold parameters and regression coefficients that characterize how changes in predictor variables (such as the reference class: kWh, percentage, or USD) relate to probabilities of being in each accuracy category.
-Specifically, we used a cumulative logit link function to model the ordered accuracy outcome, and we specified weakly informative priors for the regression coefficients (normal distributions with mean 0 and standard deviation of 1) and for the cutpoints (normal distributions with a mean of zero and a standard deviation of 4.0).
-The approach can be used to estimate threshold parameters and regression coefficients that characterize how changes in predictor variables (such as the reference class: kWh, percentage, or USD) relate to probabilities of being in each accuracy category.
+We used a cumulative logit link function to model the ordered accuracy outcome, and we specified weakly informative priors [@kruschkeDoingBayesianData2014] for the regression coefficients (normal distributions with mean 0 and standard deviation of 1) and for the cutpoints (normal distributions with a mean of zero and a standard deviation of 4.0).
+The approach can be used to estimate threshold parameters and regression coefficients that characterize how changes in predictor variables (such as the reference class: kWh, percentage, or USD) relate to probabilities of being in each level of accuracy (0% absolute error; 0.01-5% error;  Over 5% error).
+Bayesian regression provides a probability distribution over the plausible values of the model parameters, allowing for a probabilistic interpretation of the uncertainty associated with these estimates.The width of the credible intervals (Bayesian equivalent of confidence intervals) derived from the posterior distribution directly indicates the degree of uncertainty in these effects.
 
@@
 ::: {#tbl-s1-reg apa-note="Ordinal regression results.
-Positive coefficients for the reference class predictors indicate that those conditions are associated with higher error categories relative to the kWh baseline."}
+Positive coefficient estimates for the reference class predictors indicate that those conditions are associated with higher error categories relative to the kWh baseline."}
 
@@
 |Intercept[2]       |-0.89    |-2.49    |0.71     |0.87 |
-|refClassPercentage |1.44     |0.07     |2.88     |0.98 |
-|refClassUSD        |3.13     |1.81     |4.50     |1.00 |
-|calcUsedCalculator |-3.30    |-4.80    |-1.92    |1.00 |
+|RefClass-Percentage |1.44     |0.07     |2.88     |0.98 |
+|RefClass-USD        |3.13     |1.81     |4.50     |1.00 |
+|calc-UsedCalculator |-3.30    |-4.80    |-1.92    |1.00 |
 
@@
 This indicates that, compared to the kWh condition, participants in both the Percentage and USD conditions were more likely to produce plans that fell into higher error categories.
-Moreover, the odds ratios (see Table 3) suggest that the USD condition led to a notably higher likelihood of large errors compared to the kWh baseline (OR = 22.9), while the Percentage condition also demonstrated increased odds (OR = 4.2) but was somewhat less detrimental to accuracy than USD.
+Moreover, the odds ratios (see @tbl-s1-ord) suggest that the USD condition led to a notably higher likelihood of large errors compared to the kWh baseline (OR = 22.9), while the Percentage condition also demonstrated increased odds (OR = 4.2) but was somewhat less detrimental to accuracy than USD.
 These results align with our descriptive findings and further clarify that framing the target reductions in absolute kWh units may facilitate significantly more accurate planning.
-Posterior predictive checks showed that the ordinal model provided a reasonable fit to the observed data (see @fig-s1-ppd).
+To evaluate the descriptive adequacy of our statistical model, we performed posterior predictive checks, a standard Bayesian technique wherein data simulated from the fitted ordinal model are compared against the observed data.
+As shown in @fig-s1-ppd, which compares the observed proportions of participant responses for each accuracy level (represented by bars) to the models predicted proportions (represented by points).
+The close correspondence between the observed bars and predicted points across different accuracy levels and reference class conditions suggests that the model adequately captures the patterns in the observed data.
 
@@
 #| fig-cap: "Experiment 1: Posterior Predictive Check"
-#| apa-note: "Posterior predictive check for frequency of trials at each accuracy level, faceted by reference class (kWh, Percentage, USD).
-Bars show observed, dots show model predicted proportions.
-Better accuracy is indicated by higher proportions in the 'Exact Match' and '0.01-5% error' categories."
+#| apa-note: "Posterior predictive check for the Bayesian ordinal regression model.
+Blue bars show the observed frequencies for each accuracy level (Exact match, 0.01-5% error, and Over 5% error), dots with error bars represent the model predictions.
+The plot is faceted by reference class condition (kWh, Percentage, USD)."
 
@@
 To further investigate individual factors that may influence planning accuracy, we examined the relationship between participants' energy literacy scores and their performance on the task.
-Energy literacy was assessed using an 8-item questionnaire adapted from [@dewatersEnergyLiteracySecondary2011], which covers topics such as energy units, appliance energy consumption, and sources of electricity.
-A Bayesian linear regression model was fit with log-transformed absolute error as the outcome variable and energy literacy score as the predictor, controlling for random effects of participant and state: log_abs_error ~ els + (1|id) + (1|state).
+A Bayesian linear regression model was fit with log-transformed absolute error as the outcome variable and energy literacy score as the predictor, controlling for random effects of participant and state.
 Results indicated a significant negative relationship between energy literacy and log absolute error (Estimate = -2.35, 95% CI: -2.88 to -1.81), suggesting that participants with higher energy literacy scores tended to have smaller deviations from the target reduction goal, and thus more accurate plans overall (@fig-s1-els).
@@
 Experiment 1 examined how different numerical representations of energy reduction goals affected participants’ planning accuracy.
-In line with our hypothesis that absolute units would yield better accuracy, the kWh condition supported significantly more precise energy reduction plans than did either the Percentage or USD conditions.
+In line with our hypothesis that absolute units would yield better accuracy, the kWh condition supported significantly more precise energy reduction plans than either the Percentage or USD conditions.
 Although the Percentage format was detrimental to accuracy relative to kWh, it was the USD condition that consistently produced the poorest outcomes, suggesting that monetary terms, while intuitive in everyday contexts, may not serve as effective reference classes for planning appliance-specific reductions in energy use.
@@
 
+## Introduction
+
+Building upon the preliminary evidence from Experiment 1, which indicated that representing energy reduction targets in absolute kilowatt-hour (kWh) units might yield superior planning accuracy relative to percentage or monetary formats, Experiment 2 was designed to further probe the robustness and nuances of this effect.
+A primary objective was to replicate the core finding regarding the influence of reference class on planning precision.
+Additionally, this second experiment sought to broaden the investigation by examining the potential impact of other task parameters; specifically, we manipulated the difficulty of the reduction goal (comparing a 10% versus a 15% target) and the granularity of the presented numerical usage data (using exact versus rounded figures).
+
 ## Methods
 
-In Experiment 2, We recruited 206 participants from Amazon Mechanical Turk, but data from 10 participants were corrupted due to experimenter error, and six excluded due to deviant behavior, leaving a final sample of 190 participants (102 male; 88 female.
-Average age = 35.5, SD=9.5)
+In Experiment 2, 206 participants were recruited from Amazon Mechanical Turk, but data from 10 participants were corrupted due to experimenter error, and six were excluded because their error magnitudes were statistical outliers (error magnitudes exceeding 2.5 standard deviations from the group mean), leaving a final sample of 190 participants (102 males, 88 females, with an average age = 35.5, SD=9.5)
 
@@
 rounded numbers).
-The order of presentation of the goal, rounding, and state conditions was counterbalanced across participants.
+The order of presentation of the goal, rounding, and family scenarios was counterbalanced across participants.
 As in Experiment 1, each participant completed the energy reduction planning task for two different family-state scenarios.
@@
 
+summary_table <- s2_agg4 |>
+    group_by(`Reference Class` = refClass) |>
+    summarise(
+        N = n(),
+        # Keep Mean (SD) for average change proportion
+        #`Avg. % Change` = sprintf("%.2f (%.2f)", mean(mean_pct_change), sd(mean_pct_change)),
+        # Calculate Mean (SD) for percentages, multiply by 100, format as whole number
+        `Exact Match Rate(%)` = sprintf("%.0f (%.0f)", mean(pct) * 100, sd(pct) * 100),
+        `Close Match Rate(%)` = sprintf("%.0f (%.0f)", mean(pct_close) * 100, sd(pct_close) * 100),
+        # Calculate Median [IQR] for deviations, adjust decimals as needed
+        `Abs. Deviation` = sprintf("%.3f (%.3f)", median(mean_abs_error), IQR(mean_abs_error)),
+        `Log Abs. Deviation` = sprintf("%.2f (%.2f)", median(mean_log_abs_error), IQR(mean_log_abs_error))
+    ) |>
+    # Apply kable formatting (optional, depends on your workflow)
+    knitr::kable(escape = FALSE, booktabs = TRUE, align = c("l"))
+
 ```
 
-::: {#tbl-s2-agg apa-note="The table shows performance as both the % of trials where participants matched the goal, and the mean absolute error from the target reduction goal.
-Percentage meeting goal refers to participants who exactly met the target reduction goal (i.e., the proposed plan would reduce energy use by exactly 15%)"}
+::: {#tbl-s2-agg apa-note="Summary statistics by reference class condition, based on participant-level averages.
+Exact Match Rate and Close Match Rate represent the mean percentage of trials (per participant) where the proposed reduction exactly matched or closely matched (±2% tolerance) the target goal (standard deviations in parentheses), respectively.
+Abs.
+Deviation and Log Abs.
+Deviation reflect the median (IQR) of each participant’s absolute deviation from the target goal."}
 
-|Reference Class |% meeting goal (exact) |% meeting goal (close match) |Abs. Deviation |Log Abs. Deviation |
-|:---------------|:----------------------|:----------------------------|:--------------|:------------------|
-|kWh             |0.44                   |0.52                         |0.02           |-3.9               |
-|Percentage      |0.28                   |0.42                         |0.06           |-3.2               |
-|USD             |0.20                   |0.29                         |0.10           |-2.4               |
+|Reference Class |N  |Exact Match Rate(%) |Close Match Rate(%) |Abs. Deviation |Log Abs. Deviation |
+|:---------------|:--|:-------------------|:-------------------|:--------------|:------------------|
+|kWh             |68 |44 (45)             |52 (46)             |0.022 (0.182)  |-3.87 (3.12)       |
+|Percentage      |67 |28 (36)             |42 (42)             |0.062 (0.138)  |-3.22 (2.31)       |
+|USD             |55 |20 (38)             |29 (40)             |0.102 (0.135)  |-2.42 (1.52)       |
 
@@
 |Intercept[2]       |1.26     |-0.09    |2.65     |0.97 |
-|refClassPercentage |1.02     |-0.63    |2.71     |0.89 |
-|refClassUSD        |2.27     |0.53     |3.98     |0.99 |
-|calcNoCalculator   |4.10     |2.20     |6.06     |1.00 |
+|RefClass-Percentage |1.02     |-0.63    |2.71     |0.89 |
+|RefClass-USD        |2.27     |0.53     |3.98     |0.99 |
+|calc-NoCalculator   |4.10     |2.20     |6.06     |1.00 |
 |pct_goal15%        |-0.39    |-0.81    |0.04     |0.96 |
@@
 |USD vs kWh           |9.68       |1.69     |53.4     |
-|calcNoCalculator     |60.37      |9.02     |426.4    |
+|calc-NoCalculator     |60.37      |9.02     |426.4    |
 |15% Goal vs 10% Goal |0.68       |0.44     |1.0      |
@@
 rounded usage data), and calculator usage as additional predictors.
-Random intercepts were specified for both participant and state,
+Random intercepts were specified for both participant and state (@eq-e2).
+The model was fit using the brms package in R, with weakly informative priors for the regression coefficients and cutpoints [@kruschkeDoingBayesianData2014].
+
+$$
+\text{Accuracy} \sim \text{Reference Class} + \text{Calculator} + \text{Goal} + \text{Rounded} + (1|\text{id}) + (1|\text{Family Scenario})
+$${#eq-e2}
 
@@
     geom_point(size = 3, alpha=.6,position = position_dodge(width = 0.5)) +
-    geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, alpha=.5,
+    geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2, alpha=.3,
                    position = position_dodge(width = 0.5)) +
-    geom_vline(xintercept = 0, linetype = "dashed",alpha=.5) +
+    geom_vline(xintercept = 0, linetype = "dashed",alpha=.1) +
     labs(
@@
         aes(label = label),
-        color = "black", size = 3.5, hjust = -0.3, vjust = -0.5,
+        color = "black", size = 3.6, hjust = -0.3, vjust = -0.5,
         position = position_dodge(width = 0.5)
@@
 #| fig-cap: "Experiment 2. Posterior predictive check of the bayesian regression model"
-#| apa-note: "The plot shows the observed and predicted frequencies of each accuracy level, faceted by reference class (kWh, Percentage, USD).
-The model provides a reasonable fit to the observed data, with the predicted proportions closely matching the observed frequencies."
+#| apa-note: "Posterior predictive check for the Bayesian ordinal regression model.
+Blue bars show the observed frequencies for each accuracy level (Exact match, 0.01-5% error, and Over 5% error), dots with error bars represent the model predictions.
+The plot is faceted by reference class condition (kWh, Percentage, USD)."
 
@@
 We once again examined the effect of energy literacy on planning accuracy.
-A Bayesian linear regression model was fit with log-transformed absolute error as the outcome variable and energy literacy score as the predictor, controlling for random effects of participant and state: log_abs_error ~ els + (1|id) + (1|state).
-This revealed a significant negative relationship between energy literacy and log absolute error (Estimate = -3.21, 95% CI: -3.89 to -2.52), indicating that participants with higher energy literacy scores tended to have smaller deviations from the target reduction goal, and thus more accurate plans overall (@fig-s2-els).
+A Bayesian linear regression model was fit with log-transformed absolute error as the outcome variable and energy literacy score as the predictor, controlling for random effects of participant and state.
+This revealed a significant negative relationship between energy literacy and log absolute error (Estimate = -3.21, 95% CI: [-3.89 to -2.52]), indicating that participants with higher energy literacy scores tended to have smaller deviations from the target reduction goal, and thus more accurate plans overall (@fig-s2-els).
 
@@
 These results extend prior observations on the impact of numerical format on comprehension and decision-making [@gigerenzerSimpleToolsUnderstanding2003; @reimerNumericCommunicationRisk2015].
-but they move beyond simpler estimation tasks to show how these benefits persist in a multi-step planning context.
+But they move beyond simpler estimation tasks to show how these benefits persist in a multi-step planning context.
 Notably, individuals with higher energy literacy performed better overall, a finding in line with previous work emphasizing the importance of domain knowledge for effective resource conservation [@attariPublicPerceptionsEnergy2010; @canfieldPerceptionsElectricityuseCommunications2017]
 
-While prior research has suggested that consumers sometimes prefer monetary formats[@karjalainenConsumerPreferencesFeedback2011; @nematiImpactInformationbasedInterventions2020] or that monetary framing can improve appliance choices [@blaschBoundedlyRationalConsumers2019], our findings tentatively indicate that absolute energy units can facilitate the finer-grained calculations needed to plan specific usage cuts.
+While prior research has suggested that consumers sometimes prefer monetary formats [@karjalainenConsumerPreferencesFeedback2011; @nematiImpactInformationbasedInterventions2020] or that monetary framing can improve appliance choices [@blaschBoundedlyRationalConsumers2019], our findings tentatively indicate that absolute energy units can facilitate the finer-grained calculations needed to plan specific usage cuts.
 This may be because kWh provide a more direct and less ambiguous representation of energy quantities, facilitating the necessary calculations for accurate planning.
@@
 Nevertheless, it is unclear whether the observed benefits of absolute units (i.e., kWh) are genuinely attributable to their absolute nature, or if other inherent characteristics of these units might be driving the effects.
+
+## Limitations and Future Directions
 
@@
 
------
-
-Supplementary information and materials can be found online at [this website](https://tegorman13.github.io/Representation_Study/manuscript/output/appendix.html)
-
 {{< pagebreak >}}

### 2025-03-31 manuscript/manuscript.qmd
more tweaks

@@
 Energy insecurity has emerged as a critical public health concern, especially among low-income households, who frequently face difficult choices between paying energy bills and meeting other essential needs [@bednarRecognitionResponseEnergy2020; @memmottSociodemographicDisparitiesEnergy2021].
-Such households often experience unsafe coping strategies (e.g., foregoing heating during winter months), which disproportionately affect racial and ethnic minorities and heighten risks of utility disconnection  [@memmottSociodemographicDisparitiesEnergy2021].
+Such households often experience unsafe coping strategies (e.g., foregoing heating during winter months), which disproportionately affect at-risk populations and heighten risks of utility disconnection  [@memmottSociodemographicDisparitiesEnergy2021].
 Moreover, residential energy use contributes to climate change, intensifying the urgency for sustainable solutions [@farghaliStrategiesEnergyContext2023].
@@
 Consequently, a pressing challenge lies in designing communication strategies that effectively convey energy data and motivate practical conservation decisions.
-We set out to explore if consumers are able to develop plans that would help meet certain energy consumption goals.
+We set out to explore whether consumers are able to develop plans that would help meet certain energy consumption goals.
 
@@
 By manipulating the presentation format of energy information, this research aims to elucidate how different representational formats influence planning accuracy.
-Based on the literature reviewed, we hypothesize that: 1) Presenting energy reduction goals in absolute units (kWh) will lead to greater planning accuracy compared to percentage-based or monetary formats, as absolute units provide a more direct and less ambiguous representation of energy quantities.
+Based on the literature reviewed, we hypothesized that: 1) Presenting energy reduction goals in absolute units (kWh) will lead to greater planning accuracy compared to percentage-based or monetary formats, as absolute units provide a more direct and less ambiguous representation of energy quantities.
 2) Higher energy literacy will be associated with more accurate planning, as individuals with greater energy knowledge may be better equipped to process and utilize the provided information, regardless of format.
-We also examine the potential of several exploratory variables, such as goal difficulty and the rounding of numerical values, to further elucidate the factors that influence planning accuracy.
+We also examined the potential of several exploratory variables, such as goal difficulty and the rounding of numerical values, to further elucidate the factors that influence planning accuracy.
 
@@
 The format of the reduction goal was manipulated to be either a percentage (15% reduction), kilowatt hours (5965 kWh), or USD (\$656).
-Participants in the USD condition were provided with the conversion rate between kWh and USD."}
+Participants in the USD condition were provided with the conversion rate between kWh and USD.
+All participants completed two scenarios, and for each scenario, they proposed two distinct action plans"}
 
@@
 Deviation and Log Abs.
-Deviation reflect the median (IQR) of each participant’s absolute deviation from the target goal."}
+Deviation reflect the median (IQR in parentheses) of each participant’s absolute deviation from the target goal."}
 
@@
 #| tbl-cap: "Experiment 1: Categorization of Participants according to Accuracy Levels"
-#| apa-note: "The table shows the percentage of participants who fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages).
+#| apa-note: "The table shows the percentage of action-plans which fell into each accuracy level for each reference class condition (percentages of kWh, $, and USD columns reflect within condition percentages).
 "
@@
 The approach can be used to estimate threshold parameters and regression coefficients that characterize how changes in predictor variables (such as the reference class: kWh, percentage, or USD) relate to probabilities of being in each level of accuracy (0% absolute error; 0.01-5% error;  Over 5% error).
-Bayesian regression provides a probability distribution over the plausible values of the model parameters, allowing for a probabilistic interpretation of the uncertainty associated with these estimates.The width of the credible intervals (Bayesian equivalent of confidence intervals) derived from the posterior distribution directly indicates the degree of uncertainty in these effects.
+Bayesian regression provides a probability distribution over the plausible values of the model parameters, allowing for a probabilistic interpretation of the uncertainty associated with these estimates.
+The width of the credible intervals (Bayesian equivalent of confidence intervals) derived from the posterior distribution directly indicates the degree of uncertainty in these effects.
 
@@
 ::: {#tbl-s2-reg apa-note="Ordinal regression results.
-Positive coefficients for the reference class predictors indicate that those conditions are associated with higher error categories relative to the kWh baseline."}
+Positive coefficients for the reference class predictors indicate that those conditions are associated with higher error categories relative to the kWh baseline.
+Coefficients for calc are in reference to the 'Used Calculator' group, coefficient for the pct_goal is in reference to the 10% goal, and coefficient for rounded is in reference to the 'Not Rounded' group."}
 |Parameter          |Estimate |CI_Lower |CI_Upper |pd   |
@@
 |pct_goal15%        |-0.39    |-0.81    |0.04     |0.96 |
-|roundedRounded     |-0.53    |-0.96    |-0.11    |0.99 |
+|rounded-Rounded     |-0.53    |-0.96    |-0.11    |0.99 |
 
@@
 Participants in the USD condition were therefore substantially more likely to deviate from the target energy reduction goal compared to those in the kWh condition.
-In contrast, the Percentage condition’s odds ratio relative to kWh was more uncertain (OR = 2.78, 95% CI: 0.53, 15.0), indicating that although there may be a trend toward reduced accuracy in the Percentage condition, the evidence was not definitive.
+In contrast, the Percentage condition’s odds ratio relative to kWh was more uncertain (OR = 2.78, 95% CI: [0.53, 15.0]), indicating that although there may be a trend toward reduced accuracy in the Percentage condition, the evidence was not definitive.
 
-We also found that using rounded numbers modestly improved accuracy (b = -0.53, 95% CI: [-0.96, -0.11]), with participants having 0.59 times the odds of falling into a worse accuracy category when working with rounded values.
-The more challenging 15% reduction goal was associated with slightly better performance compared to the 10% goal (b = -0.39, 95% CI: [-0.81, 0.04]), though this effect was relatively small.
+We also found that using rounded numbers modestly improved accuracy (Estimate = -0.53, 95% CI: [-0.96, -0.11]), with participants having 0.59 times the odds of falling into a worse accuracy category when working with rounded values.
+The more challenging 15% reduction goal was associated with slightly better performance compared to the 10% goal, Estimate = -0.39, 95% CI: [-0.81, 0.04], though this effect was relatively small.
 Consistent with Experiment 1, the use of a calculator had a large and significant effect on accuracy.
-The coefficient for calcNoCalculator was 4.10 (95% CI: 2.20, 6.06), and the corresponding odds ratio was 60.37 (95% CI: 9.02, 426.4), indicating that participants who did not use a calculator were substantially more likely to fall into higher error categories.
+The coefficient for calc-NoCalculator was Estimate = 4.10 95% CI: [2.20, 6.06], and the corresponding odds ratio was 60.37 (95% CI: [9.02, 426.4]), indicating that participants who did not use a calculator were substantially more likely to fall into higher error categories.
 
-@fig-s2-ame shows the marginal effects of refClass on each level of accuracy_level.
-These results reveal that switching from kWh to Percentage decreased the probability of an "Exact match" by an average of 7.0 percentage points (95% CI: -19.2, 4.2) and increased the probability of "Over 5% error" by 6.9 percentage points (95% CI: -4.5, 18.6).
-Similarly, switching from kWh to USD decreased the probability of an "Exact match" by 15 percentage points (95% CI: -26.7, -3.3) and increased the probability of "Over 5% error" by 16.5 percentage points (95% CI: 3.7, 29.3).
+@fig-s2-ame shows the marginal effects of Reference Class on each level of accuracy_level.
+These results reveal that switching from kWh to Percentage decreased the probability of an "Exact match" by an average of 7.0 percentage points, 95% CI: [-19.2, 4.2] and increased the probability of falling into the "Over 5% error" category by 6.9 percentage points, 95% CI: [-4.5, 18.6].
+Similarly, switching from kWh to USD decreased the probability of an "Exact match" by 15 percentage points, 95% CI: [-26.7, -3.3],  and increased the probability of "Over 5% error" by 16.5 percentage points, 95% CI: [3.7, 29.3].
 
@@
 It may be that participants put more effort into the task under this condition, or perhaps this is an artifact of the way that the task was presented.
-However, this effect was relatively small and thus should be explored in future research to better understand its underlying mechanisms.
+However, this effect was relatively small and thus should be explored in future research to confidently establish its existence.
 Furthermore, the magnitude of the effect size of this manipulation should be examined to better understand the practical implications of goal difficulty for energy conservation.
@@
 These results extend prior observations on the impact of numerical format on comprehension and decision-making [@gigerenzerSimpleToolsUnderstanding2003; @reimerNumericCommunicationRisk2015].
-But they move beyond simpler estimation tasks to show how these benefits persist in a multi-step planning context.
+But they move beyond simpler estimation tasks to show how these benefits persist in a goal-directed planning context.
 Notably, individuals with higher energy literacy performed better overall, a finding in line with previous work emphasizing the importance of domain knowledge for effective resource conservation [@attariPublicPerceptionsEnergy2010; @canfieldPerceptionsElectricityuseCommunications2017]
@@
 Similarly, the concept of “default units” [@herberzKiloWhatDefault2020] further highlights how presenting energy data in a straightforward, standardized format can guide better consumer decisions.
-Nevertheless, it is unclear whether the observed benefits of absolute units (i.e., kWh) are genuinely attributable to their absolute nature, or if other inherent characteristics of these units might be driving the effects.
+Nevertheless, it is unclear whether the observed benefits of absolute units (i.e., kWh) are genuinely attributable to their absolute nature, or if other characteristics of these units might be driving the effects (e.g., familiarity).
 

### 2025-04-01 manuscript/manuscript.qmd
sync

@@
 
+See @tbl-s2-agg for a breakdown of the descriptive statistics.
 The ordinal regression analysis revealed that the USD reference class significantly increased the odds of higher error categories compared to the kWh reference class (OR = 9.68, 95% CI: [1.69, 53.4]).
 Participants in the USD condition were therefore substantially more likely to deviate from the target energy reduction goal compared to those in the kWh condition.
-In contrast, the Percentage condition’s odds ratio relative to kWh was more uncertain (OR = 2.78, 95% CI: [0.53, 15.0]), indicating that although there may be a trend toward reduced accuracy in the Percentage condition, the evidence was not definitive.
+In contrast, the Percentage condition’s odds ratio relative to kWh was more uncertain (OR = 2.78, 95% CI: [0.53, 15.0]), indicating that although there may be a trend toward reduced accuracy in the Percentage condition, the evidence was not definitive (@tbl-s2-ord).
 

### 2025-04-01 manuscript/manuscript.qmd
resync

@@
 See @tbl-s2-agg for a breakdown of the descriptive statistics.
-The ordinal regression analysis revealed that the USD reference class significantly increased the odds of higher error categories compared to the kWh reference class (OR = 9.68, 95% CI: [1.69, 53.4]).
+The ordinal regression analysis (see @tbl-s2-reg) revealed that the USD reference class significantly increased the odds of higher error categories compared to the kWh reference class (OR = 9.68, 95% CI: [1.69, 53.4]).
 Participants in the USD condition were therefore substantially more likely to deviate from the target energy reduction goal compared to those in the kWh condition.

### 2025-04-09 poster/after-body.tex
code to generate poster for mpa

% This part runs after the main document content

\end{multicols} % End the 3-column layout

% Draw a simple footer line with a slightly taller background
\AddToHook{shipout/background}{% Add to background hook for each page
  \begin{tikzpicture}[remember picture, overlay]
    % Extend footer area height slightly
    \fill[purduegold] (current page.south west) rectangle ([yshift=0.53in]current page.south east);

    % Adjust Text position in the footer
    \node[anchor=south east, xshift=-0.75in, yshift=0.2in, % Shift down slightly
          text=darkgray, font=\small\sffamily] % Text color and font
          at (current page.south east) 
          {Communication and Cognition Lab | Purdue University};

    % Adjust email position in the footer
    \node[anchor=south west, xshift=0.75in, yshift=0.05in, % Shift down slightly
          text=darkgray, font=\small\sffamily] % Text color and font
          at (current page.south west) 
          {\textsuperscript{*}Corresponding author: \href{mailto:tegorman@purdue.edu}{tegorman@purdue.edu}};
  \end{tikzpicture}
}

% \end{multicols} % End the 3-column layout

% % Draw background color and text for footer area using TikZ overlay
% \AddToHook{shipout/background}{% Add to background hook for each page
%   \begin{tikzpicture}[remember picture, overlay]
%     % Fill footer area with Purdue Gold - adjust height (1in) as needed
%     \fill[purduegold] (current page.south west) rectangle ([yshift=0.8in]current page.south east);

%     % Add Text to the footer - Adjust position and content
%     \node[anchor=south east, xshift=-0.75in, yshift=0.25in, % Position from bottom right
%           text=darkgray, font=\small\sffamily] % Text color and font
%           at (current page.south east) 
%           {Communication and Cognition Lab | Purdue University};

%     % Add email to the left side of the footer
%     \node[anchor=south west, xshift=0.75in, yshift=0.25in, % Position from bottom left
%           text=darkgray, font=\small\sffamily] % Text color and font
%           at (current page.south west) 
%           {\textsuperscript{*}Corresponding author: \href{mailto:tegorman@purdue.edu}{tegorman@purdue.edu}};

%     % Optional: Add a logo - Uncomment and adjust path/size
%     % \node[anchor=south west, xshift=0.75in, yshift=0.15in] 
%     %       at (current page.south west) {
%     %   \includegraphics[height=0.5in]{path/to/purdue_logo.png} % Adjust height/path
%     % };
%   \end{tikzpicture}
% }

### 2025-04-09 poster/before-body.tex
code to generate poster for mpa

% before-body.tex

% Replace the current header background with a more subtle approach
\AddToHook{shipout/background}{% Add to background hook for each page
  \begin{tikzpicture}[remember picture, overlay]
    % Fill header area with a much thinner line instead of full background
    \fill[purduegold] (current page.north west) rectangle ([yshift=-0.2in]current page.north east);
  \end{tikzpicture}
}

% --- Logo and Title Block Setup ---
\noindent % Prevent potential paragraph indentation
\makebox[\linewidth]{% Create a box spanning the full line width
  \begin{minipage}[t]{0.15\linewidth} % Left Logo Column
    \raggedright
    \vspace*{0pt} % Align content to the top
    % First logo (Purdue logo)
    \includegraphics[width=0.80\linewidth]{assets/PU-H-Full-RGB.png}

    \vspace{0.12in} % Space between logos

    % Second logo (CLA logo)
    \includegraphics[width=0.75\linewidth]{assets/CLA_logo1.png}
  \end{minipage}%
  \hfill
  \begin{minipage}[t]{0.65\linewidth} % Center Text Column
    \centering
    \vspace*{0pt} % Align content to the top

    % --- Title ---
    {\color{darkgray}\Huge\bfseries\sffamily Planning to Save Energy: How Information Format Affects Accuracy\par}
    \vspace{0.15in} % Space after title

    % --- Authors ---
    {\color{purduegray}\Large\sffamily
    Thomas E. Gorman\textsuperscript{1,2,*},
    Torsten Reimer\textsuperscript{1,2},
    Juan Pablo Loaiza Ramirez\textsuperscript{1,2},
    Hayden Barber\textsuperscript{3}
    \par}
    \vspace{0.08in} % Space after authors

    % --- Affiliations ---
    {\color{purduegray}\large\sffamily
    \textsuperscript{1}Communication and Cognition Lab, Purdue University, West Lafayette, IN, USA \\
    \textsuperscript{2}College of Liberal Arts Research Academy, Purdue University, West Lafayette, IN, USA \\
    \textsuperscript{3}School of Communication \& Journalism, South Dakota State University, Brookings, SD, USA \\
    \vspace{0.05in} % Space after affiliations
    \par}

  \end{minipage}%
  \hfill
  \begin{minipage}[t]{0.10\linewidth} % Right Logo Column
    \raggedleft
    \vspace*{0pt} % Adjust slightly if needed e.g., \vspace*{-2pt}
    \includegraphics[scale=0.42]{assets/nsf_logo1.png}
  \end{minipage}%
} % End \makebox

% Adjust space AFTER the header block, before the multicolumn content starts.
\vspace{0.5in} % Or adjust as needed

% Start the 3-column layout
\begin{multicols}{3}
\raggedright % Justify text to the left, better for narrow columns

### 2025-04-09 poster/poster.qmd
code to generate poster for mpa

---
title: "Planning to Save Energy: How Information Format Affects Accuracy"
---

## Introduction

*   **Problem:** Energy insecurity is a critical public health concern, especially among low-income households who face difficult choices and unsafe coping strategies [@bednarRecognitionResponseEnergy2020; @memmottSociodemographicDisparitiesEnergy2021]. Furthermore, residential energy use significantly contributes to climate change, necessitating sustainable solutions [@farghaliStrategiesEnergyContext2023].
*   **Gap:** While interventions exist, success hinges on communication format [@fischerFeedbackHouseholdElectricity2008]. Specifically, how do different **numerical representations (kWh, %, USD)** of energy reduction goals influence consumers' ability to create *accurate*, specific conservation plans for appliance use?
*   **Concepts & Theory:** This study draws on principles of energy literacy [@dewatersEnergyLiteracySecondary2011], reference class effects (absolute units like kWh may simplify calculations compared to relative units like % or derived units like USD) [@gigerenzerSimpleToolsUnderstanding2003; @reimerNumericCommunicationRisk2015], and the importance of procedural knowledge (knowing *how* to act on a goal) [@tonkeProvidingProceduralKnowledge2024].

### Research Questions & Hypotheses

*   **Primary Question:** How does the format (kWh, %, USD) used to present an energy reduction goal influence the accuracy of household energy conservation planning?
*   **H1 (Reference Class):** Presenting goals in **absolute units (kWh)** will lead to higher planning accuracy compared to percentage (%) or monetary (USD) formats. 
*   **H2 (Energy Literacy):** Individuals with **higher energy literacy** will demonstrate more accurate planning, regardless of format.

## Methods

*   **Design:** Two online experiments (N=229 & N=190 via MTurk) using a simulated household energy planning task.
*   **Task:** Participants adjusted appliance energy use (kWh) for hypothetical families to meet a specified reduction goal.
*   **Manipulation:** Goal presentation format (Between-Subjects): **kWh** vs. **%** vs. **USD**.
*   **Exp 2 Factors:** Also varied Goal Difficulty (10/15%) & Numerical Rounding (exact/rounded).

### Figure 1: Procedure and Task Example

<!-- ![Participants were shown a family's previous year usage by appliance alongside state averages. They were asked to create two action plans by adjusting the 'kWh next year' for each appliance to meet a specific total reduction goal. The goal was presented differently based on condition: as a percentage (15%), absolute kWh (reduce by 5965 kWh), or monetary savings ($656, with conversion rate provided).](assets/images/wells10e_pub.png){width="70%" fig-align="center"} -->

```{=latex}
\begin{figure}[H]
\centering

\makebox[\linewidth]{%
  \begin{minipage}[t]{0.28\linewidth}
    \centering
    \subcaptionbox{Procedure overview \label{fig-mermaid}}[0.95\linewidth]{%
      \includegraphics[width=\linewidth, height=9.5in,keepaspectratio]{assets/images/mermaid_proc.png}
    }
  \end{minipage}
  \hspace{0.5em}
  \begin{minipage}[t]{0.68\linewidth}
    \centering
    \subcaptionbox{Example of planning task interface \label{fig-task}}[0.95\linewidth]{%
      \includegraphics[width=\linewidth,height=9.5in, keepaspectratio]{assets/images/tab.png}
    }
  \end{minipage}
}

\end{figure}
```

\columnbreak

## Results

*  **Key Measures:**
    *   **Planning Accuracy:** Absolute deviation between plan reduction and the goal reduction (ordinal bins: exact, minor, large error).
    *   **Energy Literacy:** 8-item knowledge scale score [@dewatersEnergyLiteracySecondary2011].
*  **Analysis:** Bayesian mixed-effects regression (Ordinal & Linear) controlling for participant and scenario random effects: $\text{Accuracy Level} \; \text{\textasciitilde} \; \text{Reference Class} + \text{Calculator} + (1|\text{id}) + (1|\text{Family Scenario})$
* **Goal Format (H1 Supported):** Planning accuracy was significantly higher when the energy reduction goal was presented in absolute units (kWh) compared to percentages (%) or U.S. dollars (USD) across both experiments. 
    * *Exp 1:* USD (*OR* = 22.9) and % (*OR* = 4.2) formats had significantly higher odds of lower accuracy than kWh.
    * *Exp 2:* Similar significant effect for USD (*OR* = 9.68); the trend persisted for % (*OR* = 2.78), though the CI included 1.0. USD/percentage formats reduced the probability of exact matches and increased large errors compared to kWh.
* **Energy Literacy (H2 Supported):** Higher energy literacy scores were significantly associated with more accurate planning (lower error) in both experiments (Exp 1: $B$ = -2.35; Exp 2: $B$  = -3.21). (See Fig 3b).
* **Exploratory Findings (Exp 2):**
    * Using rounded numbers for prior usage slightly improved accuracy (Est. = -0.53). No effect of Goal Difficulty.

### Figure 2: Reference Class Effect on Planning Error (Exp 1)

![Distribution of log absolute error by goal format condition (Exp 1). Lower values indicate higher accuracy.](assets/images/fig-s1-log-dist-1.png){width="80%"}

### Figure 3

```{=latex}
\begin{figure}[H]
\centering

\makebox[\linewidth]{%
  \begin{minipage}[t]{0.44\linewidth}
    \centering
    \subcaptionbox{Posterior predictive check for the ordinal regression model. Blue bars show the observed frequencies per accuracy level, dots represent the model predictions.\label{fig-s1-ppd}}[0.93\linewidth]{%
    \includegraphics[height=4.25in, keepaspectratio]{assets/images/fig-s1-ppd-1.png}
    }
  \end{minipage}
  \hspace{0.01em}
  \begin{minipage}[t]{0.44\linewidth}
  \centering
    \subcaptionbox{Relationship between energy literacy score and log absolute error, controlling for random effects \label{fig-s1-els}}[0.93\linewidth]{%
    \includegraphics[height=4.25in, keepaspectratio]{assets/images/fig-s1-els-1.png}
    }
  \end{minipage}
}

\end{figure}
```
\columnbreak

### Figure 4: Reference Class Effect on Planning Error (Exp 2)

![Distribution of log absolute error by goal format condition (Exp 2). Lower values indicate higher accuracy.](assets/images/fig-s2-log-dist-1.png){width="83%"}

<!-- ### Figure 5

```{=latex}
\begin{figure}[H]
\centering

\makebox[\linewidth]{%
  \begin{minipage}[t]{0.44\linewidth}
    \centering
    \subcaptionbox{Posterior predictive check for the Bayesian ordinal regression model. Blue bars show the observed frequencies for each accuracy level, dots with error bars represent the model predictions.\label{fig-s2-ppd}}[0.90\linewidth]{%
    \includegraphics[height=4.25in, keepaspectratio]{assets/images/fig-s2-ppd-1.png}
    }
  \end{minipage}
  \hspace{0.01em}
  \begin{minipage}[t]{0.44\linewidth}
  \centering
    \subcaptionbox{Relationship between energy literacy score and log absolute error, controlling for random effects of participant and family scenario \label{fig-s2-els}}[0.90\linewidth]{%
    \includegraphics[height=4.25in, keepaspectratio]{assets/images/fig-s2-els-1.png}
    }
  \end{minipage}
}

\end{figure}
``` -->

## Discussion

*   **Key Finding:** Goal format significantly impacts planning accuracy. **Presenting goals in absolute kWh is superior** to relative (%) or monetary (USD) formats, likely by simplifying calculations and reducing cognitive load [@gigerenzerSimpleToolsUnderstanding2003].
*   **Energy Literacy Matters:** Higher domain knowledge consistently predicted better planning accuracy, emphasizing the role of energy education [@canfieldPerceptionsElectricityuseCommunications2017].
*   **Limitations:** Lab simulation, short duration, self-reported calculator use.
*   **Future Directions:** Explore combining kWh format with actionable tips [@tonkeProvidingProceduralKnowledge2024]; test interactive tools/visualizations.

## References

::: {#refs}
:::

### 2025-04-09 poster/poster.tex
code to generate poster for mpa

% Options for packages loaded elsewhere
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
]{article}
\usepackage{xcolor}
\usepackage[paperwidth=42in,paperheight=25in,top=.2in,bottom=0.5in,left=0.5in,right=0.5in,columnsep=0.75in]{geometry}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{setspace}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\setlength{\emergencystretch}{3em} % prevent overfull lines

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage[T1]{fontenc}       % Use modern font encoding
\usepackage{helvet}            % Use Helvetica clone (sans-serif)
\renewcommand\familydefault{\sfdefault} % Make sans-serif the default font family
\usepackage{graphicx}          % Required for images
\usepackage{amsmath}           % For math symbols if needed
\usepackage{amssymb}           % For math symbols if needed
\usepackage{textcomp}          % For symbols like degrees
\usepackage{xcolor}            % For color definitions
\usepackage{multicol}          % For multi-column layout
\usepackage{geometry}          % To set paper size and margins (already used by Quarto, but good practice)
\usepackage{fancyhdr}          % For headers and footers
% \usepackage{titlesec}          % For customizing section titles
\usepackage{sectsty}           % Alternative/complement to titlesec for section fonts
\usepackage{enumitem}          % For customizing lists (itemize, enumerate)
\usepackage{caption}           % For customizing figure/table captions
\usepackage{subcaption} 
\usepackage{hyperref}          % For clickable links (references, URLs)
\usepackage{tikz}              % For drawing header/footer backgrounds
\usepackage{subcaption}
\usepackage{amsfonts}
\usepackage{etoolbox}
\usepackage{titletoc}        % For table of contents customization
\usepackage[explicit]{titlesec}  % For section title customization with explicit option

% --- Color Definitions ---
\definecolor{purduegold}{HTML}{CEB888} % Keep original Purdue Gold
\definecolor{purdueblack}{HTML}{000000}
\definecolor{purduegray}{HTML}{5F6062}  % Keep original Purdue Gray
\definecolor{sectionbg}{HTML}{D6C8A3}  % New tan/beige color for section headers
\definecolor{lightgold}{HTML}{FBF8F1}  % A lighter shade for backgrounds
\definecolor{darkgray}{HTML}{333333}   % Darker Gray for text contrast

% --- SET BASE FONT SIZE HERE (Overrides any default/cached \normalsize) ---
\makeatletter
\renewcommand{\normalsize}{%
    \@setfontsize\normalsize{18.5pt}{22pt}% Main font size and line spacing
    \abovedisplayskip 12pt plus 3pt minus 7pt
    \abovedisplayshortskip \z@ plus 3pt
    \belowdisplayshortskip 6.5pt plus 3.5pt minus 3pt
    \belowdisplayskip \abovedisplayskip
    \let\@listi\@listI}
\makeatother

% Adjust other font size commands to be in proportion to normalsize
\makeatletter
\renewcommand{\large}{\@setfontsize\large{18pt}{24pt}}
\renewcommand{\Large}{\@setfontsize\Large{20pt}{26pt}}
\renewcommand{\LARGE}{\@setfontsize\LARGE{24pt}{30pt}}
\renewcommand{\huge}{\@setfontsize\huge{28pt}{36pt}}
\renewcommand{\Huge}{\@setfontsize\Huge{34pt}{42pt}}
\makeatother

\normalsize  % Activate the new normal size definition

% --- Font & Section Styling ---
% Using sectsty for simplicity here
\sectionfont{\color{darkgray}\huge\bfseries\sffamily\raggedright} % Make section titles large, bold, dark gray
\subsectionfont{\color{darkgray}\Large\bfseries\sffamily\raggedright} % Make subsection titles large, bold, dark gray
\subsubsectionfont{\color{purduegray}\large\bfseries\sffamily\raggedright} % Make subsubsection titles medium, bold, gray

% --- ADJUST SPACING HERE ---
% Reduce space AFTER section titles and BEFORE subsections/subsubsections
% Format: \titlespacing*{<command>}{<left>}{<before-sep>}{<after-sep>}
\titlespacing*{\section}
  {0pt}               % Left margin
  {.3ex plus 0.2ex minus .0ex} % Space BEFORE title (reduced slightly)
  {0.3ex plus .1ex}   % Space AFTER title (SIGNIFICANTLY REDUCED from 1ex or more)

\titlespacing*{\subsection}
  {0pt}               % Left margin
  {.0ex plus 0.0ex minus .0ex} % Space BEFORE title (reduced slightly)
  {0.1ex plus .1ex}   % Space AFTER title (REDUCED)

\titlespacing*{\subsubsection}
  {0pt}               % Left margin
  {0.0ex plus 0.0ex minus .2ex} % Space BEFORE title (reduced slightly)
  {0.0ex plus .02ex}   % Space AFTER title (REDUCED)

% --- List Styling ---
\setlist[itemize]{leftmargin=1.5em, label=\textbullet, itemsep=0.4ex, topsep=0.05ex} % Adjust bullet indentation and spacing
\setlist[enumerate]{leftmargin=1.5em, itemsep=0.2ex, topsep=0.5ex}

% --- ADJUST LIST SPACING HERE ---
% Reduce space BEFORE the entire list starts
\setlist[itemize]{
  leftmargin=1.5em,     % Indentation
  label=\textbullet,    % Bullet style
  itemsep=0.2ex,        % Space between items in the list
  topsep=0.0ex          % Space BEFORE the first item (SIGNIFICANTLY REDUCED from 0.5ex or more)
}
\setlist[enumerate]{
  leftmargin=1.5em,
  itemsep=0.2ex,
  topsep=0.2ex          % Space BEFORE the first item (REDUCED)
}

% --- Section Header Styling ---
% Define the section background color first
\definecolor{sectionbg}{HTML}{D6C8A3}  % Tan/beige color for section headers

% Use a simpler approach for section styling without relying on TikZ nodes
\titleformat{\section}
  {\normalfont\Large\bfseries\sffamily}
  {}
  {0pt}
  {\colorbox{sectionbg}{\parbox{\dimexpr\columnwidth-2\fboxsep\relax}{\centering#1}}}
  []

% For unnumbered sections
\titleformat{name=\section,numberless}
  {\normalfont\Large\bfseries\sffamily}
  {}
  {0pt}
  {\colorbox{sectionbg}{\parbox{\dimexpr\columnwidth-2\fboxsep\relax}{\centering#1}}}
  []

% Adjust spacing around sections
\titlespacing*{\section}
  {0pt}
  {1.5ex plus 0.5ex minus 0.2ex}
  {1.0ex plus 0.2ex}

% --- Header and Footer Styling ---
\pagestyle{fancy} % Use fancyhdr
\fancyhf{}       % Clear default header/footer
\renewcommand{\headrulewidth}{0pt} % No header rule
\renewcommand{\footrulewidth}{0pt} % No footer rule
% Footer content is added via tikz overlay in after-body.tex

\DeclareCaptionFont{subcapstyle}{\sffamily\fontsize{16pt}{19pt}\selectfont}

% --- Figure Caption Styling ---
\captionsetup{
  font={normalsize, sf},          % Small sans-serif font for captions
  labelfont={bf, color=darkgray}, % Bold, dark gray label (e.g., "Figure 1:")
  labelsep=period,           % Use period after label
  justification=raggedright, % Left-align caption text
  singlelinecheck=false,     % Ensure justification applies even to single lines
  skip=1ex                   % Space between figure and caption
}

\captionsetup[subfigure]{font=subcapstyle}

% --- Hyperlink Styling ---
\hypersetup{
  colorlinks=true,          % Enable colored links
  linkcolor=purduegray,     % Color for internal links (e.g., refs)
  citecolor=purduegray,     % Color for citation links
  filecolor=purduegray,     % Color for file links
  urlcolor=purduegray,      % Color for URL links
  pdftitle={Planning to Save Energy: How Information Format Affects Accuracy}, % PDF metadata
  pdfauthor={Thomas E. Gorman et al.} % PDF metadata
}
\AtBeginEnvironment{CSLReferences}{\footnotesize} 
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Planning to Save Energy: How Information Format Affects Accuracy},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Planning to Save Energy: How Information Format Affects Accuracy}
\author{}
\date{}
\begin{document}
% before-body.tex

% Replace the current header background with a more subtle approach
\AddToHook{shipout/background}{% Add to background hook for each page
  \begin{tikzpicture}[remember picture, overlay]
    % Fill header area with a much thinner line instead of full background
    \fill[purduegold] (current page.north west) rectangle ([yshift=-0.2in]current page.north east);
  \end{tikzpicture}
}

% --- Logo and Title Block Setup ---
\noindent % Prevent potential paragraph indentation
\makebox[\linewidth]{% Create a box spanning the full line width
  \begin{minipage}[t]{0.15\linewidth} % Left Logo Column
    \raggedright
    \vspace*{0pt} % Align content to the top
    % First logo (Purdue logo)
    \includegraphics[width=0.80\linewidth]{assets/PU-H-Full-RGB.png}

    \vspace{0.12in} % Space between logos

    % Second logo (CLA logo)
    \includegraphics[width=0.75\linewidth]{assets/CLA_logo1.png}
  \end{minipage}%
  \hfill
  \begin{minipage}[t]{0.65\linewidth} % Center Text Column
    \centering
    \vspace*{0pt} % Align content to the top

    % --- Title ---
    {\color{darkgray}\Huge\bfseries\sffamily Planning to Save Energy: How Information Format Affects Accuracy\par}
    \vspace{0.15in} % Space after title

    % --- Authors ---
    {\color{purduegray}\Large\sffamily
    Thomas E. Gorman\textsuperscript{1,2,*},
    Torsten Reimer\textsuperscript{1,2},
    Juan Pablo Loaiza Ramirez\textsuperscript{1,2},
    Hayden Barber\textsuperscript{3}
    \par}
    \vspace{0.08in} % Space after authors

    % --- Affiliations ---
    {\color{purduegray}\large\sffamily
    \textsuperscript{1}Communication and Cognition Lab, Purdue University, West Lafayette, IN, USA \\
    \textsuperscript{2}College of Liberal Arts Research Academy, Purdue University, West Lafayette, IN, USA \\
    \textsuperscript{3}School of Communication \& Journalism, South Dakota State University, Brookings, SD, USA \\
    \vspace{0.05in} % Space after affiliations
    \par}

  \end{minipage}%
  \hfill
  \begin{minipage}[t]{0.10\linewidth} % Right Logo Column
    \raggedleft
    \vspace*{0pt} % Adjust slightly if needed e.g., \vspace*{-2pt}
    \includegraphics[scale=0.42]{assets/nsf_logo1.png}
  \end{minipage}%
} % End \makebox

% Adjust space AFTER the header block, before the multicolumn content starts.
\vspace{0.5in} % Or adjust as needed

% Start the 3-column layout
\begin{multicols}{3}
\raggedright % Justify text to the left, better for narrow columns

\setstretch{1.01}
\subsection{Introduction}\label{introduction}

\begin{itemize}
\tightlist
\item
  \textbf{Problem:} Energy insecurity is a critical public health
  concern, especially among low-income households who face difficult
  choices and unsafe coping strategies (Bednar \& Reames, 2020; Memmott
  et al., 2021). Furthermore, residential energy use significantly
  contributes to climate change, necessitating sustainable solutions
  (Farghali et al., 2023).
\item
  \textbf{Gap:} While interventions exist, success hinges on
  communication format (Fischer, 2008). Specifically, how do different
  \textbf{numerical representations (kWh, \%, USD)} of energy reduction
  goals influence consumers' ability to create \emph{accurate}, specific
  conservation plans for appliance use?
\item
  \textbf{Concepts \& Theory:} This study draws on principles of energy
  literacy (DeWaters \& Powers, 2011), reference class effects (absolute
  units like kWh may simplify calculations compared to relative units
  like \% or derived units like USD) (Gigerenzer \& Edwards, 2003;
  Reimer et al., 2015), and the importance of procedural knowledge
  (knowing \emph{how} to act on a goal) (Tonke, 2024).
\end{itemize}

\subsubsection{Research Questions \&
Hypotheses}\label{research-questions-hypotheses}

\begin{itemize}
\tightlist
\item
  \textbf{Primary Question:} How does the format (kWh, \%, USD) used to
  present an energy reduction goal influence the accuracy of household
  energy conservation planning?
\item
  \textbf{H1 (Reference Class):} Presenting goals in \textbf{absolute
  units (kWh)} will lead to higher planning accuracy compared to
  percentage (\%) or monetary (USD) formats.
\item
  \textbf{H2 (Energy Literacy):} Individuals with \textbf{higher energy
  literacy} will demonstrate more accurate planning, regardless of
  format.
\end{itemize}

\subsection{Methods}\label{methods}

\begin{itemize}
\tightlist
\item
  \textbf{Design:} Two online experiments (N=229 \& N=190 via MTurk)
  using a simulated household energy planning task.
\item
  \textbf{Task:} Participants adjusted appliance energy use (kWh) for
  hypothetical families to meet a specified reduction goal.
\item
  \textbf{Manipulation:} Goal presentation format (Between-Subjects):
  \textbf{kWh} vs.~\textbf{\%} vs.~\textbf{USD}.
\item
  \textbf{Exp 2 Factors:} Also varied Goal Difficulty (10/15\%) \&
  Numerical Rounding (exact/rounded).
\end{itemize}

\subsubsection{Figure 1: Procedure and Task
Example}\label{figure-1-procedure-and-task-example}

\begin{figure}[H]
\centering

\makebox[\linewidth]{%
  \begin{minipage}[t]{0.28\linewidth}
    \centering
    \subcaptionbox{Procedure overview \label{fig-mermaid}}[0.95\linewidth]{%
      \includegraphics[width=\linewidth, height=9.5in,keepaspectratio]{assets/images/mermaid_proc.png}
    }
  \end{minipage}
  \hspace{0.5em}
  \begin{minipage}[t]{0.68\linewidth}
    \centering
    \subcaptionbox{Example of planning task interface \label{fig-task}}[0.95\linewidth]{%
      \includegraphics[width=\linewidth,height=9.5in, keepaspectratio]{assets/images/tab.png}
    }
  \end{minipage}
}

\end{figure}

\columnbreak

\subsection{Results}\label{results}

\begin{itemize}
\tightlist
\item
  \textbf{Key Measures:}

  \begin{itemize}
  \tightlist
  \item
    \textbf{Planning Accuracy:} Absolute deviation between plan
    reduction and the goal reduction (ordinal bins: exact, minor, large
    error).
  \item
    \textbf{Energy Literacy:} 8-item knowledge scale score (DeWaters \&
    Powers, 2011).
  \end{itemize}
\item
  \textbf{Analysis:} Bayesian mixed-effects regression (Ordinal \&
  Linear) controlling for participant and scenario random effects:
  \(\text{Accuracy Level} \; \text{\textasciitilde} \; \text{Reference Class} + \text{Calculator} + (1|\text{id}) + (1|\text{Family Scenario})\)
\item
  \textbf{Goal Format (H1 Supported):} Planning accuracy was
  significantly higher when the energy reduction goal was presented in
  absolute units (kWh) compared to percentages (\%) or U.S. dollars
  (USD) across both experiments.

  \begin{itemize}
  \tightlist
  \item
    \emph{Exp 1:} USD (\emph{OR} = 22.9) and \% (\emph{OR} = 4.2)
    formats had significantly higher odds of lower accuracy than kWh.
  \item
    \emph{Exp 2:} Similar significant effect for USD (\emph{OR} = 9.68);
    the trend persisted for \% (\emph{OR} = 2.78), though the CI
    included 1.0. USD/percentage formats reduced the probability of
    exact matches and increased large errors compared to kWh.
  \end{itemize}
\item
  \textbf{Energy Literacy (H2 Supported):} Higher energy literacy scores
  were significantly associated with more accurate planning (lower
  error) in both experiments (Exp 1: \(B\) = -2.35; Exp 2: \(B\) =
  -3.21). (See Fig 3b).
\item
  \textbf{Exploratory Findings (Exp 2):}

  \begin{itemize}
  \tightlist
  \item
    Using rounded numbers for prior usage slightly improved accuracy
    (Est. = -0.53). No effect of Goal Difficulty.
  \end{itemize}
\end{itemize}

\subsubsection{Figure 2: Reference Class Effect on Planning Error (Exp
1)}\label{figure-2-reference-class-effect-on-planning-error-exp-1}

\begin{figure}[H]

{\centering \includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{assets/images/fig-s1-log-dist-1.png}

}

\caption{Distribution of log absolute error by goal format condition
(Exp 1). Lower values indicate higher accuracy.}

\end{figure}%

\subsubsection{Figure 3}\label{figure-3}

\begin{figure}[H]
\centering

\makebox[\linewidth]{%
  \begin{minipage}[t]{0.44\linewidth}
    \centering
    \subcaptionbox{Posterior predictive check for the ordinal regression model. Blue bars show the observed frequencies per accuracy level, dots represent the model predictions.\label{fig-s1-ppd}}[0.93\linewidth]{%
    \includegraphics[height=4.25in, keepaspectratio]{assets/images/fig-s1-ppd-1.png}
    }
  \end{minipage}
  \hspace{0.01em}
  \begin{minipage}[t]{0.44\linewidth}
  \centering
    \subcaptionbox{Relationship between energy literacy score and log absolute error, controlling for random effects \label{fig-s1-els}}[0.93\linewidth]{%
    \includegraphics[height=4.25in, keepaspectratio]{assets/images/fig-s1-els-1.png}
    }
  \end{minipage}
}

\end{figure}

\columnbreak

\subsubsection{Figure 4: Reference Class Effect on Planning Error (Exp
2)}\label{figure-4-reference-class-effect-on-planning-error-exp-2}

\begin{figure}[H]

{\centering \includegraphics[width=0.83\linewidth,height=\textheight,keepaspectratio]{assets/images/fig-s2-log-dist-1.png}

}

\caption{Distribution of log absolute error by goal format condition
(Exp 2). Lower values indicate higher accuracy.}

\end{figure}%

\subsection{Discussion}\label{discussion}

\begin{itemize}
\tightlist
\item
  \textbf{Key Finding:} Goal format significantly impacts planning
  accuracy. \textbf{Presenting goals in absolute kWh is superior} to
  relative (\%) or monetary (USD) formats, likely by simplifying
  calculations and reducing cognitive load (Gigerenzer \& Edwards,
  2003).
\item
  \textbf{Energy Literacy Matters:} Higher domain knowledge consistently
  predicted better planning accuracy, emphasizing the role of energy
  education (Canfield et al., 2017).
\item
  \textbf{Limitations:} Lab simulation, short duration, self-reported
  calculator use.
\item
  \textbf{Future Directions:} Explore combining kWh format with
  actionable tips (Tonke, 2024); test interactive tools/visualizations.
\end{itemize}

\subsection{References}\label{references}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-bednarRecognitionResponseEnergy2020}
Bednar, D. J., \& Reames, T. G. (2020). Recognition of and response to
energy poverty in the {United States}. \emph{Nature Energy},
\emph{5}(6), 432--439. \url{https://doi.org/10.1038/s41560-020-0582-0}

\bibitem[\citeproctext]{ref-canfieldPerceptionsElectricityuseCommunications2017}
Canfield, C., Bruine De Bruin, W., \& Wong-Parodi, G. (2017).
Perceptions of electricity-use communications: Effects of information,
format, and individual differences. \emph{Journal of Risk Research},
\emph{20}(9), 1132--1153.
\url{https://doi.org/10.1080/13669877.2015.1121909}

\bibitem[\citeproctext]{ref-dewatersEnergyLiteracySecondary2011}
DeWaters, J. E., \& Powers, S. E. (2011). Energy literacy of secondary
students in {New York State} ({USA}): {A} measure of knowledge, affect,
and behavior. \emph{Energy Policy}, \emph{39}(3), 1699--1710.
\url{https://doi.org/10.1016/j.enpol.2010.12.049}

\bibitem[\citeproctext]{ref-farghaliStrategiesEnergyContext2023}
Farghali, M., Osman, A. I., Mohamed, I. M. A., Chen, Z., Chen, L.,
Ihara, I., Yap, P.-S., \& Rooney, D. W. (2023). Strategies to save
energy in the context of the energy crisis: A review.
\emph{Environmental Chemistry Letters}, \emph{21}(4), 2003--2039.
\url{https://doi.org/10.1007/s10311-023-01591-5}

\bibitem[\citeproctext]{ref-fischerFeedbackHouseholdElectricity2008}
Fischer, C. (2008). Feedback on household electricity consumption: A
tool for saving energy? \emph{Energy Efficiency}, \emph{1}(1), 79--104.
\url{https://doi.org/10.1007/s12053-008-9009-7}

\bibitem[\citeproctext]{ref-gigerenzerSimpleToolsUnderstanding2003}
Gigerenzer, G., \& Edwards, A. (2003). Simple tools for understanding
risks: From innumeracy to insight. \emph{BMJ}, \emph{327}(7417),
741--744. \url{https://doi.org/10.1136/bmj.327.7417.741}

\bibitem[\citeproctext]{ref-memmottSociodemographicDisparitiesEnergy2021}
Memmott, T., Carley, S., Graff, M., \& Konisky, D. M. (2021).
Sociodemographic disparities in energy insecurity among low-income
households before and during the {COVID-19} pandemic. \emph{Nature
Energy}, \emph{6}(2), 186--193.
\url{https://doi.org/10.1038/s41560-020-00763-9}

\bibitem[\citeproctext]{ref-reimerNumericCommunicationRisk2015}
Reimer, T., Jones, C., \& Skubisz, C. (2015). Numeric {Communication} of
{Risk}. In \emph{The {SAGE} handbook of risk communication} (pp.
167--179).

\bibitem[\citeproctext]{ref-tonkeProvidingProceduralKnowledge2024}
Tonke, S. (2024). Providing procedural knowledge: {A} field experiment
to encourage resource conservation in {Namibia}. \emph{Journal of
Development Economics}, \emph{166}, 103202.
\url{https://doi.org/10.1016/j.jdeveco.2023.103202}

\end{CSLReferences}

% This part runs after the main document content

\end{multicols} % End the 3-column layout

% Draw a simple footer line with a slightly taller background
\AddToHook{shipout/background}{% Add to background hook for each page
  \begin{tikzpicture}[remember picture, overlay]
    % Extend footer area height slightly
    \fill[purduegold] (current page.south west) rectangle ([yshift=0.53in]current page.south east);

    % Adjust Text position in the footer
    \node[anchor=south east, xshift=-0.75in, yshift=0.2in, % Shift down slightly
          text=darkgray, font=\small\sffamily] % Text color and font
          at (current page.south east) 
          {Communication and Cognition Lab | Purdue University};

    % Adjust email position in the footer
    \node[anchor=south west, xshift=0.75in, yshift=0.05in, % Shift down slightly
          text=darkgray, font=\small\sffamily] % Text color and font
          at (current page.south west) 
          {\textsuperscript{*}Corresponding author: \href{mailto:tegorman@purdue.edu}{tegorman@purdue.edu}};
  \end{tikzpicture}
}

% \end{multicols} % End the 3-column layout

% % Draw background color and text for footer area using TikZ overlay
% \AddToHook{shipout/background}{% Add to background hook for each page
%   \begin{tikzpicture}[remember picture, overlay]
%     % Fill footer area with Purdue Gold - adjust height (1in) as needed
%     \fill[purduegold] (current page.south west) rectangle ([yshift=0.8in]current page.south east);

%     % Add Text to the footer - Adjust position and content
%     \node[anchor=south east, xshift=-0.75in, yshift=0.25in, % Position from bottom right
%           text=darkgray, font=\small\sffamily] % Text color and font
%           at (current page.south east) 
%           {Communication and Cognition Lab | Purdue University};

%     % Add email to the left side of the footer
%     \node[anchor=south west, xshift=0.75in, yshift=0.25in, % Position from bottom left
%           text=darkgray, font=\small\sffamily] % Text color and font
%           at (current page.south west) 
%           {\textsuperscript{*}Corresponding author: \href{mailto:tegorman@purdue.edu}{tegorman@purdue.edu}};

%     % Optional: Add a logo - Uncomment and adjust path/size
%     % \node[anchor=south west, xshift=0.75in, yshift=0.15in] 
%     %       at (current page.south west) {
%     %   \includegraphics[height=0.5in]{path/to/purdue_logo.png} % Adjust height/path
%     % };
%   \end{tikzpicture}
% }
\end{document}

### 2025-04-09 poster/poster_docx.qmd
code to generate poster for mpa

---
title: "Planning to Save Energy: How Information Format Affects Accuracy"
format: docx  # Changed format to docx
---

## Introduction

*   **Problem:** Energy insecurity is a critical public health concern, especially among low-income households who face difficult choices and unsafe coping strategies [@bednarRecognitionResponseEnergy2020; @memmottSociodemographicDisparitiesEnergy2021]. Furthermore, residential energy use significantly contributes to climate change, necessitating sustainable solutions [@farghaliStrategiesEnergyContext2023].
*   **Need:** Effective communication of energy consumption information is crucial for promoting residential energy conservation behaviors [@canfieldPerceptionsElectricityuseCommunications2017; @abrahamseReviewInterventionStudies2005].
*   **Gap:** While interventions exist, success hinges on communication format [@fischerFeedbackHouseholdElectricity2008]. Specifically, how do different **numerical representations (kWh, %, USD)** of energy reduction goals influence consumers' ability to create *accurate*, specific conservation plans for appliance use?
*   **Concepts & Theory:** This study draws on principles of energy literacy [@dewatersEnergyLiteracySecondary2011], reference class effects (absolute units like kWh may simplify calculations compared to relative units like % or derived units like USD) [@gigerenzerSimpleToolsUnderstanding2003; @reimerNumericCommunicationRisk2015], and the importance of procedural knowledge (knowing *how* to act on a goal) [@tonkeProvidingProceduralKnowledge2024].

### Research Questions & Hypotheses

*   **Primary Question:** How does the format (kWh, %, USD) used to present an energy reduction goal influence the accuracy of household energy conservation planning?
*   **H1 (Reference Class):** Presenting goals in **absolute units (kWh)** will lead to higher planning accuracy compared to percentage (%) or monetary (USD) formats. *Rationale: kWh provides a direct, less ambiguous measure, potentially simplifying calculations.*
*   **H2 (Energy Literacy):** Individuals with **higher energy literacy** will demonstrate more accurate planning, regardless of format. *Rationale: Better domain knowledge aids information processing.*
*   **Exploratory (Exp 2):** Investigated effects of goal difficulty (10% vs. 15% reduction) and numerical presentation (exact vs. rounded values).

### Figure 1: Procedure and Task Example

<!-- Removed commented out simple figure -->

<!-- Converted LaTeX figure block to Markdown figures -->

![Figure 1a: Procedure overview](assets/images/mermaid_proc.png){#fig-mermaid height="5.25in" fig-align="center"}

![Figure 1b: Example of planning task interface](assets/images/tab.png){#fig-task height="5.25in" fig-align="center"}

## Methods

*   **Design:** Two online experiments hosted on Qualtrics, participants recruited via Amazon Mechanical Turk (Exp 1: N=229; Exp 2: N=190).
*   **Task:** Participants completed a **simulated household energy planning task**. For two hypothetical families (in different climate zones), they adjusted projected energy usage across five appliance categories (Cooling, Heating, Water Heating, Refrigerator, Other) to meet a specific energy reduction target for the next year. Each participant created two distinct action plans per family.
*   **Key Manipulation (Between-Subjects):** The energy reduction goal was presented in one of three formats:
    *   **Kilowatt-hours (kWh):** e.g., "reduce use by 5965 kWh"
    *   **Percentage (%):** e.g., "reduce use by 15%"
    *   **U.S. Dollars (USD):** e.g., "save $656" (conversion rate provided)
    *   *(Note: All formats represented the same underlying energy reduction.)*
*   **Other Factors:**
    *   *Within-Subjects:* Family Scenario (varied baseline usage).
    *   *Exp 2 Manipulations:* Goal Difficulty (10% vs. 15% target), Numerical Rounding (prior usage data exact vs. rounded).
*   **Measures:**
    *   **Planning Accuracy:** Primary Dependent Variable. Calculated as the absolute deviation of the participant's proposed total kWh reduction from the target reduction. Analyzed as binned ordinal categories (Exact Match [0% error], Minor Deviation [0.01-5% error], Large Deviation [>5% error]) and continuous log absolute error.
    *   **Energy Literacy:** Score based on an 8-item scale assessing knowledge of energy concepts and units [@dewatersEnergyLiteracySecondary2011].
    *   **Calculator Use:** Self-reported (Yes/No).
*   **Analysis:** Mixed-effects Bayesian Ordinal and Linear Regression models (using `brms` package in R). Controlled for random effects of participant and family scenario. Priors were weakly informative. Assessed effects using posterior distributions, Credible Intervals (CI), and Odds Ratios (OR).
    *   $\text{Accuracy Level} \; \text{\textasciitilde} \; \text{Reference Class} + \text{Calculator} + (1|\text{id}) + (1|\text{Family Scenario})$
    *   $\text{Accuracy Level} \; \text{\textasciitilde} \; \text{Reference Class} + \text{Calculator} + \text{Goal} + \text{Rounded} + (1|\text{id}) + (1|\text{Family Scenario})$

<!-- Removed \columnbreak -->

## Results

* **Goal Format (H1 Supported):** Planning accuracy was significantly higher when the energy reduction goal was presented in absolute units (kWh) compared to percentages (%) or U.S. dollars (USD) across both experiments. The USD format consistently resulted in the lowest accuracy.
    * *Exp 1:* USD (*OR* = 22.9) and % (*OR* = 4.2) formats had significantly higher odds of lower accuracy than kWh.
    * *Exp 2:* Similar significant effect for USD (*OR* = 9.68); the trend persisted for % (*OR* = 2.78), though the CI included 1.0. USD/percentage formats reduced the probability of exact matches and increased large errors compared to kWh.
* **Energy Literacy (H2 Supported):** Higher energy literacy scores were significantly associated with more accurate planning (lower error) in both experiments (Exp 1: β = -2.35; Exp 2: β = -3.21)[cite: 6]. (See Fig 3b).
* **Exploratory Findings (Exp 2):**
    * Using rounded numbers for prior usage slightly improved accuracy (Est. = -0.53).
    * A more difficult goal (15% vs. 10% reduction) was associated with slightly higher accuracy, though not statistically significant (Est. = -0.39).

### Figure 2: Reference Class Effect on Planning Error (Exp 1)

![Distribution of log absolute error by goal format condition (Exp 1). Lower values indicate higher accuracy.](assets/images/fig-s1-log-dist-1.png){#fig-log-dist-1 width="75%"}

### Figure 3

<!-- Converted LaTeX figure block to Markdown figures -->

![Figure 3a: Posterior predictive check for the Bayesian ordinal regression model. Blue bars show the observed frequencies for each accuracy level, dots with error bars represent the model predictions.](assets/images/fig-s1-ppd-1.png){#fig-s1-ppd height="4.25in" fig-align="center"}

![Figure 3b: Relationship between energy literacy score and log absolute error, controlling for random effects of participant and family scenario](assets/images/fig-s1-els-1.png){#fig-s1-els height="4.25in" fig-align="center"}

<!-- Removed \columnbreak -->

### Figure 4

![Distribution of log absolute error by goal format condition (Exp 2). Lower values indicate higher accuracy.](assets/images/fig-s2-log-dist-1.png){#fig-log-dist-2 width="75%"}

### Figure 5

<!-- Converted LaTeX figure block to Markdown figures -->

![Figure 5a: Posterior predictive check for the Bayesian ordinal regression model. Blue bars show the observed frequencies for each accuracy level, dots with error bars represent the model predictions.](assets/images/fig-s2-ppd-1.png){#fig-s2-ppd height="4.25in" fig-align="center"}

![Figure 5b: Relationship between energy literacy score and log absolute error, controlling for random effects of participant and family scenario](assets/images/fig-s2-els-1.png){#fig-s2-els height="4.25in" fig-align="center"}

## Discussion

*   **Main Takeaway:** The format of energy reduction goals significantly impacts planning accuracy. **Absolute units (kWh) are superior** to relative (%) or monetary (USD) formats for helping consumers create precise energy-saving plans.
*   **Why kWh?** Likely because kWh provides a direct, concrete unit for calculation, reducing the cognitive steps (and potential for error) involved in converting percentages or monetary values back into energy units for appliance-level planning [@gigerenzerSimpleToolsUnderstanding2003]. This aligns with findings on the benefits of absolute counts/frequencies in risk communication.
*   **Energy Literacy Matters:** Consistent with prior work [@canfieldPerceptionsElectricityuseCommunications2017], individuals with greater domain knowledge perform better, highlighting the value of energy education.
*   **Practical Implication:** Communicators aiming to help households plan specific energy reductions should prioritize using kWh, even if consumers sometimes express preference for monetary information [@karjalainenConsumerPreferencesFeedback2011]. Calculators are essential tools.

### Limitations & Future Directions

*   **Limitations:** Controlled lab simulation (not real-world behavior); short task duration; self-reported calculator use (effort unknown); limited number of plans per participant.
*   **Future Directions:**
    *   Conduct field studies to assess real-world planning and behavior change over time.
    *   Investigate effects across different demographics and income levels.
    *   Explore combining optimal formats (kWh) with actionable, procedural tips [@tonkeProvidingProceduralKnowledge2024; @abrahamseEffectTailoredInformation2007].
    *   Test effectiveness of interactive planning tools or visualizations.
    *   Examine the counterintuitive goal difficulty finding further.

## References

::: {#refs}
:::

### 2025-04-09 poster/poster.qmd
update poster

@@
 
-*   **Problem:** Energy insecurity is a critical public health concern, especially among low-income households who face difficult choices and unsafe coping strategies [@bednarRecognitionResponseEnergy2020; @memmottSociodemographicDisparitiesEnergy2021].
-Furthermore, residential energy use significantly contributes to climate change, necessitating sustainable solutions [@farghaliStrategiesEnergyContext2023].
-*   **Gap:** While interventions exist, success hinges on communication format [@fischerFeedbackHouseholdElectricity2008].
-Specifically, how do different **numerical representations (kWh, %, USD)** of energy reduction goals influence consumers' ability to create *accurate*, specific conservation plans for appliance use?
-*   **Concepts & Theory:** This study draws on principles of energy literacy [@dewatersEnergyLiteracySecondary2011], reference class effects (absolute units like kWh may simplify calculations compared to relative units like % or derived units like USD) [@gigerenzerSimpleToolsUnderstanding2003; @reimerNumericCommunicationRisk2015], and the importance of procedural knowledge (knowing *how* to act on a goal) [@tonkeProvidingProceduralKnowledge2024].
+*   **Problem:** Energy insecurity is a critical public health concern [@bednarRecognitionResponseEnergy2020; @memmottSociodemographicDisparitiesEnergy2021]. 
+*   **Prior Work:** While interventions exist, success often hinges on communication format [@canfieldPerceptionsElectricityuseCommunications2017; @fischerFeedbackHouseholdElectricity2008], such as the use of tables vs.
+figures, the framing of information (e.g., savings vs.
+costs), and energy literacy.
+Many reasoning tasks involving quantities are also influenced by reference class (e.g., absolute units like kWh may simplify calculations compared to relative units like % or derived units like USD) [@gigerenzerSimpleToolsUnderstanding2003; @reimerNumericCommunicationRisk2015].
 
@@
 
-*   **Primary Question:** How does the format (kWh, %, USD) used to present an energy reduction goal influence the accuracy of household energy conservation planning?
-*   **H1 (Reference Class):** Presenting goals in **absolute units (kWh)** will lead to higher planning accuracy compared to percentage (%) or monetary (USD) formats. 
+*   **Primary Question:** How does the Reference class (kWh, %, USD) used to present a reduction goal influence the accuracy of energy planning?
+*   **H1 (Reference Class):** Goals presented in **absolute units (kWh)** will benefit planning accuracy compared to percentage (%) or monetary (USD) formats. 
 *   **H2 (Energy Literacy):** Individuals with **higher energy literacy** will demonstrate more accurate planning, regardless of format.
@@
 *   **Design:** Two online experiments (N=229 & N=190 via MTurk) using a simulated household energy planning task.
-*   **Task:** Participants adjusted appliance energy use (kWh) for hypothetical families to meet a specified reduction goal.
-*   **Manipulation:** Goal presentation format (Between-Subjects): **kWh** vs. **%** vs. **USD**.
-*   **Exp 2 Factors:** Also varied Goal Difficulty (10/15%) & Numerical Rounding (exact/rounded).
+*   **Planning Task:** Propose appliance energy use (kWh) for hypothetical families to meet a target reduction goal.
+*   **Manipulation:** Reference Class of reduction goal (Between-Subjects): **kWh** vs. **%** vs. **USD**.
+*   **Exp 2 Factors:** Goal Difficulty (10% vs. 15% reduction) and Numerical Rounding (exact vs. rounded).
 
@@
 *  **Key Measures:**
-    *   **Planning Accuracy:** Absolute deviation between plan reduction and the goal reduction (ordinal bins: exact, minor, large error).
+    *   **Planning Accuracy:**  Deviation between intended reduction goal and proposed plan (ordinal bins: exact, minor, large error).
     *   **Energy Literacy:** 8-item knowledge scale score [@dewatersEnergyLiteracySecondary2011].
-*  **Analysis:** Bayesian mixed-effects regression (Ordinal & Linear) controlling for participant and scenario random effects: $\text{Accuracy Level} \; \text{\textasciitilde} \; \text{Reference Class} + \text{Calculator} + (1|\text{id}) + (1|\text{Family Scenario})$
-* **Goal Format (H1 Supported):** Planning accuracy was significantly higher when the energy reduction goal was presented in absolute units (kWh) compared to percentages (%) or U.S.
-dollars (USD) across both experiments.
-    * *Exp 1:* USD (*OR* = 22.9) and % (*OR* = 4.2) formats had significantly higher odds of lower accuracy than kWh.
-    * *Exp 2:* Similar significant effect for USD (*OR* = 9.68); the trend persisted for % (*OR* = 2.78), though the CI included 1.0.
-USD/percentage formats reduced the probability of exact matches and increased large errors compared to kWh.
-* **Energy Literacy (H2 Supported):** Higher energy literacy scores were significantly associated with more accurate planning (lower error) in both experiments (Exp 1: $B$ = -2.35; Exp 2: $B$  = -3.21).
-(See Fig 3b).
-* **Exploratory Findings (Exp 2):**
-    * Using rounded numbers for prior usage slightly improved accuracy (Est. = -0.53). No effect of Goal Difficulty.
+*  **Analysis:** Bayesian mixed-effects regression: $\text{Accuracy Level} \; \text{\textasciitilde} \; \text{Reference Class} + \text{Calculator} + (1|\text{id}) + (1|\text{Family Scenario})$
+* **Effect of reference class (H1 Supported):** Smallest planning errors when reduction goal was presented in absolute units (kWh) compared to percentages (%) or dollars (USD) across both experiments.
+* **Energy Literacy (H2 Supported):** Higher energy literacy scores associated with more accurate planning (lower error) in both experiments.
 
@@
 
-![Distribution of log absolute error by goal format condition (Exp 1). Lower values indicate higher accuracy.](assets/images/fig-s1-log-dist-1.png){width="80%"}
+![Distribution of log absolute error by goal format condition (Exp 1). Lower values indicate higher accuracy.](assets/images/fig-s1-log-dist-1.png){width="85%"}
 
-### Figure 3
+### Figure 3 - Bayesian Regression Results (Exp 1)
 
@@
     \centering
-    \subcaptionbox{Posterior predictive check for the ordinal regression model.
-Blue bars show the observed frequencies per accuracy level, dots represent the model predictions.\label{fig-s1-ppd}}[0.93\linewidth]{%
-    \includegraphics[height=4.25in, keepaspectratio]{assets/images/fig-s1-ppd-1.png}
+    \subcaptionbox{PPC for ordinal model. Blue bars are observed frequencies per accuracy level, dots show model predictions.\label{fig-s1-ppd}}[0.93\linewidth]{%
+    \includegraphics[height=4.4in, keepaspectratio]{assets/images/fig-s1-ppd-1.png}
     }
@@
   \centering
-    \subcaptionbox{Relationship between energy literacy score and log absolute error, controlling for random effects \label{fig-s1-els}}[0.93\linewidth]{%
-    \includegraphics[height=4.25in, keepaspectratio]{assets/images/fig-s1-els-1.png}
+    \subcaptionbox{Relationship between energy literacy score and log absolute error of reduction plans \label{fig-s1-els}}[0.93\linewidth]{%
+    \includegraphics[height=4.4in, keepaspectratio]{assets/images/fig-s1-els-1.png}
     }
@@
 
-![Distribution of log absolute error by goal format condition (Exp 2). Lower values indicate higher accuracy.](assets/images/fig-s2-log-dist-1.png){width="83%"}
+![Distribution of log absolute error by goal format condition (Exp 2). Lower values indicate higher accuracy.](assets/images/fig-s2-log-dist-1.png){width="85%"}
 
-<!-- ### Figure 5
+### Figure 5 - Bayesian Regression Results (Exp 2)
 
@@
     \centering
-    \subcaptionbox{Posterior predictive check for the Bayesian ordinal regression model.
-Blue bars show the observed frequencies for each accuracy level, dots with error bars represent the model predictions.\label{fig-s2-ppd}}[0.90\linewidth]{%
-    \includegraphics[height=4.25in, keepaspectratio]{assets/images/fig-s2-ppd-1.png}
+    \subcaptionbox{PPC for ordinal model. Blue bars are observed frequencies per accuracy level, dots show model predictions.\label{fig-s2-ppd}}[0.90\linewidth]{%
+    \includegraphics[height=4.4in, keepaspectratio]{assets/images/fig-s2-ppd-1.png}
     }
@@
   \centering
-    \subcaptionbox{Relationship between energy literacy score and log absolute error, controlling for random effects of participant and family scenario \label{fig-s2-els}}[0.90\linewidth]{%
-    \includegraphics[height=4.25in, keepaspectratio]{assets/images/fig-s2-els-1.png}
+    \subcaptionbox{Relationship between energy literacy score and log absolute error of reduction plans. \label{fig-s2-els}}[0.90\linewidth]{%
+    \includegraphics[height=4.4in, keepaspectratio]{assets/images/fig-s2-els-1.png}
     }
@@
 \end{figure}
-``` -->
+```
 
@@
 
-*   **Key Finding:** Goal format significantly impacts planning accuracy.
-**Presenting goals in absolute kWh is superior** to relative (%) or monetary (USD) formats, likely by simplifying calculations and reducing cognitive load [@gigerenzerSimpleToolsUnderstanding2003].
-*   **Energy Literacy Matters:** Higher domain knowledge consistently predicted better planning accuracy, emphasizing the role of energy education [@canfieldPerceptionsElectricityuseCommunications2017].
+*   **Key Finding:** Reference class  significantly impacts planning accuracy.
+**Presenting goals in absolute kWh is superior** to relative (%) or monetary (USD) formats, likely by simplifying calculations and reducing cognitive load.
+*   **Energy Literacy Matters:** Higher domain knowledge consistently predicted better planning accuracy, emphasizing the role of energy education.
 *   **Limitations:** Lab simulation, short duration, self-reported calculator use.
-*   **Future Directions:** Explore combining kWh format with actionable tips [@tonkeProvidingProceduralKnowledge2024]; test interactive tools/visualizations.
 

### 2025-04-09 poster/poster.tex
update poster

@@
 \usepackage[explicit]{titlesec}  % For section title customization with explicit option
+\usepackage{caption}
+\captionsetup[figure]{labelformat=empty}% redefines the caption setup of the figures environment in the beamer class.
 
@@
 \renewcommand{\normalsize}{%
-    \@setfontsize\normalsize{18.5pt}{22pt}% Main font size and line spacing
+    \@setfontsize\normalsize{19.5pt}{22pt}% Main font size and line spacing
     \abovedisplayskip 12pt plus 3pt minus 7pt
@@
   {0pt}               % Left margin
-  {.0ex plus 0.0ex minus .0ex} % Space BEFORE title (reduced slightly)
+  {.1ex plus 0.1ex minus .0ex} % Space BEFORE title (reduced slightly)
   {0.1ex plus .1ex}   % Space AFTER title (REDUCED)
@@
   {0pt}               % Left margin
-  {0.0ex plus 0.0ex minus .2ex} % Space BEFORE title (reduced slightly)
+  {0.1ex plus 0.1ex minus .2ex} % Space BEFORE title (reduced slightly)
   {0.0ex plus .02ex}   % Space AFTER title (REDUCED)
@@
   \textbf{Problem:} Energy insecurity is a critical public health
-  concern, especially among low-income households who face difficult
-  choices and unsafe coping strategies (Bednar \& Reames, 2020; Memmott
-  et al., 2021). Furthermore, residential energy use significantly
-  contributes to climate change, necessitating sustainable solutions
-  (Farghali et al., 2023).
+  concern (Bednar \& Reames, 2020; Memmott et al., 2021).
 \item
-  \textbf{Gap:} While interventions exist, success hinges on
-  communication format (Fischer, 2008). Specifically, how do different
-  \textbf{numerical representations (kWh, \%, USD)} of energy reduction
-  goals influence consumers' ability to create \emph{accurate}, specific
-  conservation plans for appliance use?
-\item
-  \textbf{Concepts \& Theory:} This study draws on principles of energy
-  literacy (DeWaters \& Powers, 2011), reference class effects (absolute
-  units like kWh may simplify calculations compared to relative units
-  like \% or derived units like USD) (Gigerenzer \& Edwards, 2003;
-  Reimer et al., 2015), and the importance of procedural knowledge
-  (knowing \emph{how} to act on a goal) (Tonke, 2024).
+  \textbf{Prior Work:} While interventions exist, success often hinges
+  on communication format (Canfield et al., 2017; Fischer, 2008), such
+  as the use of tables vs.~figures, the framing of information (e.g.,
+  savings vs.~costs), and energy literacy. Many reasoning tasks
+  involving quantities are also influenced by reference class (e.g.,
+  absolute units like kWh may simplify calculations compared to relative
+  units like \% or derived units like USD) (Gigerenzer \& Edwards, 2003;
+  Reimer et al., 2015).
 \end{itemize}
@@
 \item
-  \textbf{Primary Question:} How does the format (kWh, \%, USD) used to
-  present an energy reduction goal influence the accuracy of household
-  energy conservation planning?
+  \textbf{Primary Question:} How does the Reference class (kWh, \%, USD)
+  used to present a reduction goal influence the accuracy of energy
+  planning?
 \item
-  \textbf{H1 (Reference Class):} Presenting goals in \textbf{absolute
-  units (kWh)} will lead to higher planning accuracy compared to
-  percentage (\%) or monetary (USD) formats.
+  \textbf{H1 (Reference Class):} Goals presented in \textbf{absolute
+  units (kWh)} will benefit planning accuracy compared to percentage
+  (\%) or monetary (USD) formats.
 \item
@@
 \item
-  \textbf{Task:} Participants adjusted appliance energy use (kWh) for
-  hypothetical families to meet a specified reduction goal.
+  \textbf{Planning Task:} Propose appliance energy use (kWh) for
+  hypothetical families to meet a target reduction goal.
 \item
-  \textbf{Manipulation:} Goal presentation format (Between-Subjects):
-  \textbf{kWh} vs.~\textbf{\%} vs.~\textbf{USD}.
+  \textbf{Manipulation:} Reference Class of reduction goal
+  (Between-Subjects): \textbf{kWh} vs.~\textbf{\%} vs.~\textbf{USD}.
 \item
-  \textbf{Exp 2 Factors:} Also varied Goal Difficulty (10/15\%) \&
-  Numerical Rounding (exact/rounded).
+  \textbf{Exp 2 Factors:} Goal Difficulty (10\% vs.~15\% reduction) and
+  Numerical Rounding (exact vs.~rounded).
 \end{itemize}
@@
   \item
-    \textbf{Planning Accuracy:} Absolute deviation between plan
-    reduction and the goal reduction (ordinal bins: exact, minor, large
-    error).
+    \textbf{Planning Accuracy:} Deviation between intended reduction
+    goal and proposed plan (ordinal bins: exact, minor, large error).
   \item
@@
 \item
-  \textbf{Analysis:} Bayesian mixed-effects regression (Ordinal \&
-  Linear) controlling for participant and scenario random effects:
+  \textbf{Analysis:} Bayesian mixed-effects regression:
   \(\text{Accuracy Level} \; \text{\textasciitilde} \; \text{Reference Class} + \text{Calculator} + (1|\text{id}) + (1|\text{Family Scenario})\)
 \item
-  \textbf{Goal Format (H1 Supported):} Planning accuracy was
-  significantly higher when the energy reduction goal was presented in
-  absolute units (kWh) compared to percentages (\%) or U.S. dollars
-  (USD) across both experiments.
-
-  \begin{itemize}
-  \tightlist
-  \item
-    \emph{Exp 1:} USD (\emph{OR} = 22.9) and \% (\emph{OR} = 4.2)
-    formats had significantly higher odds of lower accuracy than kWh.
-  \item
-    \emph{Exp 2:} Similar significant effect for USD (\emph{OR} = 9.68);
-    the trend persisted for \% (\emph{OR} = 2.78), though the CI
-    included 1.0. USD/percentage formats reduced the probability of
-    exact matches and increased large errors compared to kWh.
-  \end{itemize}
+  \textbf{Effect of reference class (H1 Supported):} Smallest planning
+  errors when reduction goal was presented in absolute units (kWh)
+  compared to percentages (\%) or dollars (USD) across both experiments.
 \item
   \textbf{Energy Literacy (H2 Supported):} Higher energy literacy scores
-  were significantly associated with more accurate planning (lower
-  error) in both experiments (Exp 1: \(B\) = -2.35; Exp 2: \(B\) =
-  -3.21). (See Fig 3b).
-\item
-  \textbf{Exploratory Findings (Exp 2):}
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Using rounded numbers for prior usage slightly improved accuracy
-    (Est. = -0.53). No effect of Goal Difficulty.
-  \end{itemize}
+  associated with more accurate planning (lower error) in both
+  experiments.
 \end{itemize}
@@
 
-{\centering \includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{assets/images/fig-s1-log-dist-1.png}
+{\centering \includegraphics[width=0.85\linewidth,height=\textheight,keepaspectratio]{assets/images/fig-s1-log-dist-1.png}
 
@@
 
-\subsubsection{Figure 3}\label{figure-3}
+\subsubsection{Figure 3 - Bayesian Regression Results (Exp
+1)}\label{figure-3---bayesian-regression-results-exp-1}
 
@@
 
-{\centering \includegraphics[width=0.83\linewidth,height=\textheight,keepaspectratio]{assets/images/fig-s2-log-dist-1.png}
+{\centering \includegraphics[width=0.85\linewidth,height=\textheight,keepaspectratio]{assets/images/fig-s2-log-dist-1.png}
 
@@
 
+\subsubsection{Figure 5 - Bayesian Regression Results (Exp
+2)}\label{figure-5---bayesian-regression-results-exp-2}
+
+\begin{figure}[H]
+\centering
+
+\makebox[\linewidth]{%
+  \begin{minipage}[t]{0.44\linewidth}
+    \centering
+    \subcaptionbox{PPC for ordinal model. Blue bars are observed frequencies per accuracy level, dots show model predictions.\label{fig-s2-ppd}}[0.90\linewidth]{%
+    \includegraphics[height=4.4in, keepaspectratio]{assets/images/fig-s2-ppd-1.png}
+    }
+  \end{minipage}
+  \hspace{0.01em}
+  \begin{minipage}[t]{0.44\linewidth}
+  \centering
+    \subcaptionbox{Relationship between energy literacy score and log absolute error of reduction plans. \label{fig-s2-els}}[0.90\linewidth]{%
+    \includegraphics[height=4.4in, keepaspectratio]{assets/images/fig-s2-els-1.png}
+    }
+  \end{minipage}
+}
+
+\end{figure}
+
 \subsection{Discussion}\label{discussion}
@@
 \item
-  \textbf{Key Finding:} Goal format significantly impacts planning
+  \textbf{Key Finding:} Reference class significantly impacts planning
   accuracy. \textbf{Presenting goals in absolute kWh is superior} to
   relative (\%) or monetary (USD) formats, likely by simplifying
-  calculations and reducing cognitive load (Gigerenzer \& Edwards,
-  2003).
+  calculations and reducing cognitive load.
 \item
@@
   predicted better planning accuracy, emphasizing the role of energy
-  education (Canfield et al., 2017).
+  education.
 \item
@@
   calculator use.
-\item
-  \textbf{Future Directions:} Explore combining kWh format with
-  actionable tips (Tonke, 2024); test interactive tools/visualizations.
 \end{itemize}
@@
 
-\bibitem[\citeproctext]{ref-farghaliStrategiesEnergyContext2023}
-Farghali, M., Osman, A. I., Mohamed, I. M. A., Chen, Z., Chen, L.,
-Ihara, I., Yap, P.-S., \& Rooney, D. W. (2023). Strategies to save
-energy in the context of the energy crisis: A review.
-\emph{Environmental Chemistry Letters}, \emph{21}(4), 2003--2039.
-\url{https://doi.org/10.1007/s10311-023-01591-5}
-
 \bibitem[\citeproctext]{ref-fischerFeedbackHouseholdElectricity2008}
@@
 167--179).
-
-\bibitem[\citeproctext]{ref-tonkeProvidingProceduralKnowledge2024}
-Tonke, S. (2024). Providing procedural knowledge: {A} field experiment
-to encourage resource conservation in {Namibia}. \emph{Journal of
-Development Economics}, \emph{166}, 103202.
-\url{https://doi.org/10.1016/j.jdeveco.2023.103202}
 

### 2025-04-14 manuscript/appendix.qmd
fix outputs of study_info and appendix

@@
             </style>
-  wordcount-html: 
-    output-file: appendix.html
-    wordcount-banner: true
-    echo: false
-    theme: 
-        - zephyr
-        - ../assets/custom.scss
-    toc: true
-    shift-heading-level-by: 1
-    title-block-banner: "#000D51"
-    include-in-header:
-        - text: |
-            <style type="text/css">
-            h2.hide {
-                display: none;
-            }
-            </style>
+  # wordcount-html: 
+  #   output-file: appendix.html
+  #   wordcount-banner: true
+  #   echo: false
+  #   theme: 
+  #       - zephyr
+  #       - ../assets/custom.scss
+  #   toc: true
+  #   shift-heading-level-by: 1
+  #   title-block-banner: "#000D51"
+  #   include-in-header:
+  #       - text: |
+  #           <style type="text/css">
+  #           h2.hide {
+  #               display: none;
+  #           }
+  #           </style>
   docx: default
@@
     suppress-bibliography: true
-  hikmah-manuscript-pdf:
+  # hikmah-manuscript-pdf:
+  #   output-file: rs_appendix.pdf
+  #   echo: false
+  #   mainfont: "Linux Libertine O"
+  #   mainfontoptions:
+  #   - "Numbers=Proportional"
+  #   - "Numbers=OldStyle"
+  #   mathfont: "Libertinus Math"
+  #   knitr:
+  #       opts_chunk:
+  #           dev: "cairo_pdf"
+  apaquarto-pdf: 
+    toc: false
+    echo: false
+    #output-file: rs_apa_code.pdf
     output-file: rs_appendix.pdf
-    echo: false
-    mainfont: "Linux Libertine O"
-    mainfontoptions:
-    - "Numbers=Proportional"
-    - "Numbers=OldStyle"
-    mathfont: "Libertinus Math"
-    knitr:
-        opts_chunk:
-            dev: "cairo_pdf"
+   # floatsintext: true
 format-links:   
@@
 
-@tbl-s2-prop shows that, once again, participants in the kWh condition achieved closer alignment with the target goals (44% exact matches), followed by Percentage (27%) and USD (18%).
+<!-- @tbl-s2-prop  -->
+
+shows that, once again, participants in the kWh condition achieved closer alignment with the target goals (44% exact matches), followed by Percentage (27%) and USD (18%).
 These percentages are consistent with the patterns observed in Study 1, reinforcing the conclusion that providing goals in kWh supports better accuracy.
@@
 ```{r}
+#| eval: false
 
@@
 This finding indicates that participants with higher energy literacy scores tended to produce more accurate plans, exhibiting smaller deviations from the target reduction goals.
-The conditional effect plot (@fig-s2-els) visually confirms this relationship, showing a clear decreasing trend in log absolute error as energy literacy increases.
+The conditional effect plot (fig-s2-els) visually confirms this relationship, showing a clear decreasing trend in log absolute error as energy literacy increases.
 
